### 2015-01-02 inband-STN.tex
asdf

\documentclass[11pt,pdftex,letter]{article}
%\documentclass[11pt]{llncs}
%\documentclass[11pt,pdftex]{article}
%\documentclass{sig-alternate-10pt}
%\usepackage{amsthm}
%\usepackage{algorithm}
%\usepackage[noend]{algorithmic}
\usepackage[lined,boxed,commentsnumbered]{algorithm2e}

\usepackage{amssymb}
\usepackage{comment}
\usepackage{amsmath}
%\usepackage{graphicx}
%\usepackage{color}
\usepackage{fullpage}
%\usepackage[pdftex]{graphicx}
%\DeclareGraphicsRule{*}{mps}{*}{<++>}

\ifx\pdftexversion\undefined
\usepackage[dvips]{graphicx}
\else
  \usepackage[pdftex]{graphicx}
  \DeclareGraphicsRule{*}{mps}{*}{}
\fi

\def\lf{\tiny}
\def\rrnnll{\setcounter{linenumber}{0}}
\def\nnll{\refstepcounter{linenumber}\lf\thelinenumber}
\newcounter{linenumber}

\usepackage[usenames,dvipsnames]{color}
%\usepackage[caption=true,font=footnotesize]{subfig}
\usepackage{subfigure}
%\usepackage{xspace} 	% Guesses whether a space is needed when invoked
\usepackage{listings}
\usepackage{url}
\usepackage{wrapfig}
\usepackage{cite}
%\usepackage{framed}
\usepackage{framed,color}
\definecolor{shadecolor}{rgb}{0.9,0.9,0.9}
%
\usepackage[boxed]{algorithm2e}
%\usepackage{comment}

%[[PKto change spacing
%\usepackage{titlesec}
%\titlespacing\section{0pt}{7pt}{6pt}
%]]

%\newtheorem{theorem}{Theorem}[section]
%\newtheorem{takeaway}[theorem]{Takeaway}
%\newtheorem{fact}[theorem]{Fact}

%\pdfpagewidth=8.5in
%\pdfpageheight=11in

% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

\newcommand{\CPO}{\textsc{FixTag}}
\newcommand{\DPO}{\textsc{ReuseTag}}
\newcommand{\PS}{\textsc{PS}}
\newcommand{\Bit}{\textsc{Bit}}

\newcommand{\gns}{Global Network State\xspace}
\newcommand{\GNS}{\textsc{GNS}\xspace}

%\newtheorem{theorem}{Theorem}[section]
%\newtheorem{lemma}{Lemma}[section]
%\newtheorem{claim}{Claim}[section]
%\newtheorem{scenarios}{Scenarios}[section]
%\newtheorem{observation}{Observation}
%\newtheorem{takeaway}{Takeaway}[section]
%\newtheorem{definition}{Definition}

%
%
% carry over Herald's group cool way of marking changes
\definecolor{heraldBlue}{rgb}{0.0,0.0,0.8}
\definecolor{heraldRed}{rgb}{0.8,0.0,0.0}
\definecolor{heraldGray}{rgb}{0.4,0.4,0.4}
\definecolor{heraldBlack}{rgb}{0.0,0.0,0.0} %removes comment color
\definecolor{heraldGreen}{rgb}{0.0,0.4,0.0} %removes comment color
\def\r#1{\textcolor{heraldBlue}{\em #1}}
\def\q#1{\textcolor{heraldRed}{\em #1}}
\def\d#1{\textcolor{heraldBlue}{#1}}
\def\R#1{\textcolor{heraldBlue}{#1}}
\def\D#1{\textcolor{heraldBlue}{#1}}
%
%\DeclareMathOperator{\respc}{res}

\newcommand{\len}{\text{d}}

\newcommand{\nodes}{\mathcal{N}}
\newcommand{\links}{\mathcal{E}}
\newcommand{\epoints}{\Pi}
\newcommand{\epoint}{\pi}
\newcommand{\legsw}{\mathcal{L}}
\newcommand{\sdnsw}{\mathcal{S}}
\newcommand{\sw}{\legsw\sqcup\sdnsw}
\newcommand{\fset}{FS}

\newcommand{\Cost}{\gamma}

\newcommand{\ft}{ft}

\newcommand{\eepath}{p}
\newcommand{\link}{e}

\newcommand{\dom}{\textit{dom}}
\newcommand{\pr}{\textit{pr}}
\newcommand{\CPOs}{\textit{paths}}
\newcommand{\seq}{\textit{seq}}
\newcommand{\cur}{\textit{cur}}
\newcommand{\Tag}{\textit{tag}}

\newcommand{\ports}{\Pi}
\newcommand{\seport}{\pi}
\newcommand{\inports}{\overrightarrow{\ports}}
\newcommand{\swports}{\overleftrightarrow{\ports}}
\newcommand{\seinports}{\epoints^{\bullet}}
\newcommand{\seinportsone}{\seinports_1}
\newcommand{\seinportstwo}{\seinports_2}
\newcommand{\seinportsthree}{\seinports_3}
\newcommand{\neinports}{\epoints^{\circ}}

\newcommand{\readt}{\texttt{r}}
\newcommand{\writet}{\texttt{w}}
\newcommand{\op}{\texttt{op}}

\newcommand{\cellblocks}{CB}
\newcommand{\cellblock}{c}
\newcommand{\frontier}{\mathcal{F}}
\newcommand{\MIP}{\textsc{Opt}}
\newcommand{\smartparagraph}[1]{\noindent{\bf #1}\ }
\newcommand{\eg}{{\it e.g.}}
\newcommand{\ie}{{\it i.e.}}
\newcommand{\etc}{{\it etc.}}
\newcommand{\etal}{{\it et al.}\xspace}
\newcommand{\id}{{\it id}}

\def\TR{0}
\def\NOTES{1}
\def\SAVESPACE{1}
\def\SHOWAUTHORS{1}
\def\SHOWGIT{0}
% variables may contain the above definitions to control build options at compile-time
%\input{variables}

\if \SAVESPACE 1
%\usepackage{setspace}
%\usepackage{titlesec}
%\titlespacing\section{0pt}{7pt}{6pt}
\usepackage{titlesec}
\titlespacing\section{0pt}{7pt}{6pt}

\fi

\if \NOTES 1
\newcommand{\mcnote}[1]{\textcolor{heraldBlue}{\small \bf [MC: #1]}}
\newcommand{\dlnote}[1]{\textcolor{heraldBlue}{\small \bf [DL: #1]}}
\newcommand{\ssnote}[1]{\textcolor{heraldBlue}{\small \bf [SS: #1]}}
\newcommand{\pknote}[1]{\textcolor{heraldBlue}{\small \bf [PK: #1]}}
%\newcommand{\problem}[1]{\textcolor{heraldRed}{\small \bf [PROBLEM: #1]}}
\else
\newcommand{\mcnote}[1]{}
\newcommand{\dlnote}[1]{}
\newcommand{\ssnote}[1]{}
\newcommand{\pknote}[1]{}
\newcommand{\problem}[1]{}
\fi

%Petr's definitions
\newcommand{\ack}{\textit{ack}}
\newcommand{\nack}{\textit{nack}}
\newcommand{\rmw}{\textit{rmw}}
\newcommand{\State}{\mathit{States}}
\newcommand{\ignore}[1]{}
\newcommand{\Pb}{CPC}
\newcommand{\E}{\mathcal E}
%

%[[PK environments for article style
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{observation}[theorem]{Observation}
\newenvironment{proof}[1][Proof]{\noindent\textbf{#1.} }{\hfill $\Box$\\[2mm]}
\newenvironment{proofsketch}[1][Proof sketch]{\noindent\textbf{#1.} }{\hfill $\Box$\\[2mm]}
%]]

\begin{document}
\sloppy

%\title{Distributed Network Programming}
%\title{The Distributed SDN Update Problem:\\Towards Software Transactional Networks}

%\title{On Consistent Updates in Software Defined Networks under Unreliable Control}
%\title{The Case for Reliable Software Transactional Networking}

%\title{A Distributed SDN Control Plane for Concurrent Policy Updates}

\title{Solving Consensus with OpenFlow}

\author{Liron, Stefan, Petr}

%\institute{}

\date{}

\maketitle

\thispagestyle{empty}

%\if \SAVESPACE 1
%\setlength{\floatsep}{3pt}
%\setlength{\textfloatsep}{3pt}
%\setlength{\dbltextfloatsep}{3pt}
%\setlength{\intextsep}{3pt}
%\setlength{\abovecaptionskip}{3pt}
%\fi

% A category with the (minimum) three required fields
%\category{C.2.1}{Network Architecture and Design}{Centralized Networks}
%\category{C.2.4}{Distributed Systems}{Network Operating Systems}
%\terms{Measurement, Performance}
%\keywords{}

\begin{abstract}
We solve consensus with OpenFlow, using multi-write
\end{abstract}

%\vspace{1cm}

%\begin{center}
%{\bf [Regular paper only]}
%\end{center}
%[[PK I guess not anymore?]]

%\vspace{1cm}

%\begin{center}
%{\emph{Contact Address:}\\Marco Canini, Place Sainte Barbe~2, 1348 Louvain-la-Neuve, Belgium\\Tel: $+$32 10 47 48 32,
%marco.canini@uclouvain.be}
% {\emph{Contact Address:}\\Stefan Schmid, MAR 4-4, Marchstr.~23, 10587 Berlin, Germany\\Tel: $+$49 175 930 98 75,
% stefan.schmid@tu-berlin.de}
%\end{center}

%\newpage

%\begin{center}
%{\bf Regular and student paper: Dan Levin is a full-time student.}
%\end{center}

%\section*{todo before submission}

\section{Introduction}\label{sec:intro}

\section{SDN switch configuration space as shared memory}\label{sec:todo}

We address the problem of committing consistent policies to SDN switch concurrently by multiple controllers.
We show that assuming minimalistic control protocol, SDN switch configuration space can be used to implement a consensus object thereby allowing controllers to agree on each next policy update.

Our first observation is that a match-action configuration which is accessed by multiple controllers can be considered as shared memory. Note however that standard memory map and index/offset to a word, while Flow Table entries are not indexed. To overcome this we can use the match part of an entry as the index and the action part as the word value.

In order to solve consensus we suggest to use the atomic multiple entries update capability (which is discussed in OpenFlow standard). A shared memory system that supports atomic write to (enough) multiple locations is known to allow the implementation of consensus objects. For example consider the following implementation utilizing $n^2$ memory locations, $M_{i,j}$, where $i,j\in[n]$, and the suggested values $\{v_i\}_{i\in [n]}$. All locations are initialized to zero.

In order to participate, a server i, writes its suggested value $V_i$ and atomically rewrites row i ($M_{i,*}$) with value "1" and column i ($M_{*,i}$) with value "2". After the atomic write, a server $i$ first reads the diagonal ($M_{j,j}$ for $j\in [n]$), and consider all observed servers as candidate set S.  Then server i reads all candidates intersections (locations $M_{j,k}$ where $j,k\in S$. And computes the winner server id w that was overwritten by all other candidates, i.e. such that $M_{w,k}=2$ for all $k\in S {w}$. The consensus value is $v_w$.
Note that a snapshot could replace the two reading phases.

Going back to the policy update problem, the consensus scheme allow all servers to decide on next policy by using consensus value as policy identifier, and writing the policy in advance. However this scheme doesn't actually apply the policy on incoming packets and it remains to configure the switch according to the next policy once it is decided. Making this scheme fail safe (handling failure of winning server) can be ensured by allowing any server to reconfigure the switch.

\section{The one touch policy update problem}\label{sec:todo}

In some scenarios it may be desirable to shorten the policy update time by avoiding the consensus reading and policy configuration step. We define this as the one touch policy update, namely we require each server to contact the switch only once in order to (try and) update the policy, in other words participating in reaching consensus and applying it with one atomic write. Solving this problem also has significance in understanding the strength of the SDN switch computation model.

The main idea of our solution is to adjust the previous policy update scheme in a way that performs consensus validation phase during every packet processing thereby requiring the servers only to make the first phase of the consensus - atomically writing the matrix.
In order to allow the packets to validate the consensus we make the following observation: while servers can be abstracted as memory writers, the packets processed by the switch are the readers. OpenFlow standard define an atomic write transaction as being atomic in relation to the processed packets thereby making packet processing a multiple read transaction but one read location per table is allowed. This means that if each memory location used in the consensus scheme will be stored in different table then each packet will be able to read all of them.

However, reading each memory location is not enough, as the scheme requires to make a validation that involves multiple values. We utilize the metadata field in order to store all read values, in more details, each memory value $M_{i,j}$ is stored in offset i+j*n in the metadata field. Once all values are in the metadata field, we can detect the consensus winner by trying different matches on the metadata. Although we can have one match entry per possible matrix state, this solution requires exponential number of entries. A better way which we describe next requires only $O(n^2)$ entries utilizing n tables.

Our compact validation use additional temp variable, $temp_winner$, to hold the current best consensus candidate. Each of the n validating tables checks different "column" in the matrix and update $temp_winner$ to keep track of the winner so far (after examining a prefix sub set of the columns). The entry t in table k $(0<=k<n)$  matches the case where current winner=t and location $(t,k)$ in the matrix (packet header) equals 1 and location (k,k) is non-zero. After all validating tables are processed, $temp_winner$ stores the id of the winning server which can be used to match on its policy rules (filtering out other policies) or to use the goto table command to jump to a designated policy table of the server.

\section{Dealing with multiple updates}\label{sec:todo}

TBD - adding the notion of versions and cleanup/GC. probably no good way to solve infinite versions... maybe check literature on shared memory and multiple writes.

\section{Improving the scheme with advanced OpenFlow features}\label{sec:todo}

TBD - much less resources. easily solving the infinite versions issue.

\section{Solving just concurrent policy update without consensus}\label{sec:todo}

TBD - should be simple given two previous sections

\bibliographystyle{abbrv}
\bibliography{references}  % main.bib is the name of the Bibliography in this case

\end{document}

### 2015-02-17 inband-STN.tex
stefan

@@
-\documentclass[11pt,pdftex,letter]{article}
+%\documentclass[11pt,pdftex,letter]{article}
+\documentclass[conference]{sigcomm-alternate}
+
 %\documentclass[11pt]{llncs}
@@
 
+\title{Efficient Coordination of Distributed SDN Controllers:\\The Case for Multi-Write Consensus}
+
+\title{Consistent Coordination of Distributed SDN Controllers:\\Leveraging the Consensus Power of OpenFlow}
+
 \author{Liron, Stefan, Petr}
@@
 \begin{abstract}
-We solve consensus with OpenFlow, using multi-write
+idea: reliable and scalable SDN needs distributed control; consistent network updates
+is a key challenge, especially if distributed;
+
+idea: this paper shows that Openflow can be leveraged to provide strong consensus
+abstractions! the case for the long-forgotten multi-write
+
+idea: we show how to synchronize controllers, single update but also multiple and efficient
 \end{abstract}
@@
 
-\section{SDN switch configuration space as shared memory}\label{sec:todo}
+idea: reliable and scalable SDN needs distributed control; consistent network updates
+is a key challenge, especially if distributed;
 
-We address the problem of committing consistent policies to SDN switch concurrently by multiple controllers.
-We show that assuming minimalistic control protocol, SDN switch configuration space can be used to implement a consensus object thereby allowing controllers to agree on each next policy update.
+\subsection{Our Contributions}
 
-Our first observation is that a match-action configuration which is accessed by multiple controllers can be considered as shared memory.
-Note however that standard memory map and index/offset to a word, while Flow Table entries are not indexed.
-To overcome this we can use the match part of an entry as the index and the action part as the word value.
+idea: this paper shows that Openflow can be leveraged to provide strong consensus
+abstractions! the case for the long-forgotten multi-write
+
+idea: we show how to synchronize controllers, single update but also multiple and efficient
+
+\subsection{Organization}
+
+The remainder of this paper is organized as follows.
+
+\section{Background}
+
+idea: multi-write consensus, a not well-known result!
 
@@
 Note that a snapshot could replace the two reading phases.
+
+\section{SDN switch configuration space as shared memory}\label{sec:todo}
+
+We address the problem of committing consistent policies to SDN switch concurrently by multiple controllers.
+We show that assuming minimalistic control protocol, SDN switch configuration space can be used to implement a consensus object thereby allowing controllers to agree on each next policy update.
+
+Our first observation is that a match-action configuration which is accessed by multiple controllers can be considered as shared memory.
+Note however that standard memory map and index/offset to a word, while Flow Table entries are not indexed.
+To overcome this we can use the match part of an entry as the index and the action part as the word value.
 
@@
 
+\section{Related Work}\label{sec:relwork}
+
+ideas: distributed control planes
+
+ideas: known consensus results and impossibilities
+
+ideas: the power of in-band
+
+\section{Conclusion}\label{sec:conclusion}
+
+ideas: the missing link for STN, but also others
+
 \bibliographystyle{abbrv}

### 2015-02-17 inband-STN.tex
stefan pass

--- 2015-01-02 inband-STN.tex
@@
-\documentclass[11pt,pdftex,letter]{article}
+%\documentclass[11pt,pdftex,letter]{article}
+\documentclass[conference]{sigcomm-alternate}
+
+%\documentclass{sig-alternate-10pt}
+
 %\documentclass[11pt]{llncs}
@@
 %\usepackage[noend]{algorithmic}
-\usepackage[lined,boxed,commentsnumbered]{algorithm2e}
 
@@
 %\usepackage{color}
-\usepackage{fullpage}
 %\usepackage[pdftex]{graphicx}
@@
 
-\ifx\pdftexversion\undefined
-\usepackage[dvips]{graphicx}
-\else
-  \usepackage[pdftex]{graphicx}
-  \DeclareGraphicsRule{*}{mps}{*}{}
-\fi
-
-\def\lf{\tiny}
-\def\rrnnll{\setcounter{linenumber}{0}}
-\def\nnll{\refstepcounter{linenumber}\lf\thelinenumber}
-\newcounter{linenumber}
-
-\usepackage[usenames,dvipsnames]{color}
 %\usepackage[caption=true,font=footnotesize]{subfig}
-\usepackage{subfigure}
 %\usepackage{xspace} 	% Guesses whether a space is needed when invoked
-\usepackage{listings}
-\usepackage{url}
-\usepackage{wrapfig}
 \usepackage{cite}
@@
 %
-\usepackage[boxed]{algorithm2e}
 %\usepackage{comment}
@@
 
-\newcommand{\CPO}{\textsc{FixTag}}
-\newcommand{\DPO}{\textsc{ReuseTag}}
-\newcommand{\PS}{\textsc{PS}}
-\newcommand{\Bit}{\textsc{Bit}}
-
-\newcommand{\gns}{Global Network State\xspace}
-\newcommand{\GNS}{\textsc{GNS}\xspace}
-
-%\newtheorem{theorem}{Theorem}[section]
-%\newtheorem{lemma}{Lemma}[section]
-%\newtheorem{claim}{Claim}[section]
-%\newtheorem{scenarios}{Scenarios}[section]
-%\newtheorem{observation}{Observation}
-%\newtheorem{takeaway}{Takeaway}[section]
-%\newtheorem{definition}{Definition}
-
-%
-%
-% carry over Herald's group cool way of marking changes
-\definecolor{heraldBlue}{rgb}{0.0,0.0,0.8}
-\definecolor{heraldRed}{rgb}{0.8,0.0,0.0}
-\definecolor{heraldGray}{rgb}{0.4,0.4,0.4}
-\definecolor{heraldBlack}{rgb}{0.0,0.0,0.0} %removes comment color
-\definecolor{heraldGreen}{rgb}{0.0,0.4,0.0} %removes comment color
-\def\r#1{\textcolor{heraldBlue}{\em #1}}
-\def\q#1{\textcolor{heraldRed}{\em #1}}
-\def\d#1{\textcolor{heraldBlue}{#1}}
-\def\R#1{\textcolor{heraldBlue}{#1}}
-\def\D#1{\textcolor{heraldBlue}{#1}}
-%
-%\DeclareMathOperator{\respc}{res}
-
-\newcommand{\len}{\text{d}}
-
-\newcommand{\nodes}{\mathcal{N}}
-\newcommand{\links}{\mathcal{E}}
-\newcommand{\epoints}{\Pi}
-\newcommand{\epoint}{\pi}
-\newcommand{\legsw}{\mathcal{L}}
-\newcommand{\sdnsw}{\mathcal{S}}
-\newcommand{\sw}{\legsw\sqcup\sdnsw}
-\newcommand{\fset}{FS}
-
-\newcommand{\Cost}{\gamma}
-
-\newcommand{\ft}{ft}
-
-\newcommand{\eepath}{p}
-\newcommand{\link}{e}
-
-\newcommand{\dom}{\textit{dom}}
-\newcommand{\pr}{\textit{pr}}
-\newcommand{\CPOs}{\textit{paths}}
-\newcommand{\seq}{\textit{seq}}
-\newcommand{\cur}{\textit{cur}}
-\newcommand{\Tag}{\textit{tag}}
-
-\newcommand{\ports}{\Pi}
-\newcommand{\seport}{\pi}
-\newcommand{\inports}{\overrightarrow{\ports}}
-\newcommand{\swports}{\overleftrightarrow{\ports}}
-\newcommand{\seinports}{\epoints^{\bullet}}
-\newcommand{\seinportsone}{\seinports_1}
-\newcommand{\seinportstwo}{\seinports_2}
-\newcommand{\seinportsthree}{\seinports_3}
-\newcommand{\neinports}{\epoints^{\circ}}
-
-\newcommand{\readt}{\texttt{r}}
-\newcommand{\writet}{\texttt{w}}
-\newcommand{\op}{\texttt{op}}
-
-\newcommand{\cellblocks}{CB}
-\newcommand{\cellblock}{c}
-\newcommand{\frontier}{\mathcal{F}}
-\newcommand{\MIP}{\textsc{Opt}}
-\newcommand{\smartparagraph}[1]{\noindent{\bf #1}\ }
-\newcommand{\eg}{{\it e.g.}}
-\newcommand{\ie}{{\it i.e.}}
-\newcommand{\etc}{{\it etc.}}
-\newcommand{\etal}{{\it et al.}\xspace}
-\newcommand{\id}{{\it id}}
-
-\def\TR{0}
-\def\NOTES{1}
-\def\SAVESPACE{1}
-\def\SHOWAUTHORS{1}
-\def\SHOWGIT{0}
-% variables may contain the above definitions to control build options at compile-time
-%\input{variables}
-
-\if \SAVESPACE 1
-%\usepackage{setspace}
-%\usepackage{titlesec}
-%\titlespacing\section{0pt}{7pt}{6pt}
-\usepackage{titlesec}
-\titlespacing\section{0pt}{7pt}{6pt}
-
-\fi
-
-\if \NOTES 1
-\newcommand{\mcnote}[1]{\textcolor{heraldBlue}{\small \bf [MC: #1]}}
-\newcommand{\dlnote}[1]{\textcolor{heraldBlue}{\small \bf [DL: #1]}}
-\newcommand{\ssnote}[1]{\textcolor{heraldBlue}{\small \bf [SS: #1]}}
-\newcommand{\pknote}[1]{\textcolor{heraldBlue}{\small \bf [PK: #1]}}
-%\newcommand{\problem}[1]{\textcolor{heraldRed}{\small \bf [PROBLEM: #1]}}
-\else
-\newcommand{\mcnote}[1]{}
-\newcommand{\dlnote}[1]{}
-\newcommand{\ssnote}[1]{}
-\newcommand{\pknote}[1]{}
-\newcommand{\problem}[1]{}
-\fi
-
-%Petr's definitions
-\newcommand{\ack}{\textit{ack}}
-\newcommand{\nack}{\textit{nack}}
-\newcommand{\rmw}{\textit{rmw}}
-\newcommand{\State}{\mathit{States}}
-\newcommand{\ignore}[1]{}
-\newcommand{\Pb}{CPC}
-\newcommand{\E}{\mathcal E}
-%
-
-%[[PK environments for article style
-\newtheorem{theorem}{Theorem}
-\newtheorem{conjecture}[theorem]{Conjecture}
-\newtheorem{corollary}[theorem]{Corollary}
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{observation}[theorem]{Observation}
-\newenvironment{proof}[1][Proof]{\noindent\textbf{#1.} }{\hfill $\Box$\\[2mm]}
-\newenvironment{proofsketch}[1][Proof sketch]{\noindent\textbf{#1.} }{\hfill $\Box$\\[2mm]}
-%]]
-
 \begin{document}
@@
 
-\author{Liron, Stefan, Petr}
+\title{One-Shot Coordination of Distributed SDN Controllers:\\The Case for Multi-Write Consensus}
+
+\title{One-Shot Coordination of Distributed SDN Controllers:\\On the Consensus Power of OpenFlow}
+
+\author{
+Liron Schiff\thanks{Supported by the European Research Council (ERC) Starting Grant no. 259085 and by the Israel Science Foundation Grant no.~1386/11.} $^1$, 
+Stefan Schmid$^2$, Petr Kuznetsov$^3$ \\
+\small $^1$ Tel Aviv University, Israel; $^2$ TU Berlin \& T-Labs, Germany; $^3$ ParisTech, France
+}
 
@@
 \begin{abstract}
-We solve consensus with OpenFlow, using multi-write
+The control plane of future Software-Defined Networks (SDNs)
+is likely to be \emph{distributed}: a redundant control plane
+can ensure a high availability, and a spatially distributed
+control plane allows to handle time-critical dataplane
+events close to their origin. Moreover, future SDNs may also
+allow multiple administrators or participatory users to update
+policies concurrently, e.g., by offering transactional interfaces. 
+However, not much is known today about how to implement 
+a consistent distributed control plane which supports concurrent
+policy installations.
+
+In this paper, we propose to coordinate distributed controllers
+directly on the Openflow switches: exploiting classic results
+on multi-writer consensus, we show how to use Openflow 
+to synchronize controllers and install policies in a consistent
+manner, \emph{in one shot}. Our approach is attractive as it
+does not require complicated helping mechanisms in case of controller
+failures. 
 \end{abstract}
@@
 
-\section{SDN switch configuration space as shared memory}\label{sec:todo}
+By consolidating and outsourcing the control over the dataplane switches to a logically
+centralized controller,
+Software-defined networks (SDNs) simplify the network management and facilitate faster
+innovations, as the (software) control plane is decoupled and can evolve independently.
+Moreover, Openflow, the standard SDN protocol today, introduces many generalizations,
+in terms of traffic engineering, definition of flows, as well as in-band network functionalities,
+by relying on a simple match-action paradigm which allows us to define
+forwarding rules based not only on Layer-2, but also Layer-3 and Layer-4 (and possibly beyond)
+header fields.
 
-We address the problem of committing consistent policies to SDN switch concurrently by multiple controllers.
-We show that assuming minimalistic control protocol, SDN switch configuration space can be used to implement a consensus object thereby allowing controllers to agree on each next policy update.
+While the logically centralized network perspective has the potential to simplify
+the network management---many network tasks are inherently non-local---and introduces
+many flexibilities,
+managing and operating an SDN in a more adaptive but consistent manner is non-trivial, even with SDN.
+A particularly interesting problem regards the consistent installation of new policies
+or routes.~\cite{todo,todo,todo} While implementing consistent network updates 
+is challenging already from the perspective of a single controller, there exists a wide consensus
+that the logically centralized SDN control planes will likely be \emph{actually distributed} in the future.
 
-Our first observation is that a match-action configuration which is accessed by multiple controllers can be considered as shared memory.
-Note however that standard memory map and index/offset to a word, while Flow Table entries are not indexed.
-To overcome this we can use the match part of an entry as the index and the action part as the word value.
+The reason for implementing SDN control planes in a distributed manner are manyfold.
+First, in order to provide a high availability as well as a minimal degree of
+fault-tolerance, controllers should be redundant~\cite{onix,stn,todo}: a failure
+of one controller can be masked by other controllers. Second, it has also been proposed
+to distribute controllers \emph{spatially}, in order to handle latency-sensitive and
+communication intensive control plane events close to their origin.~\cite{devoflow,kandoo,jukka,disco}
+Third, larger SDNs are likely to be operated by multiple administrators or may even offer
+participatory interfaces where different users can install and trigger policy changes
+concurrently.~\cite{participatory,stn}
+
+However, today, we do not have a good understanding of how to realize
+such distributed control planes. The problem is essentially a distributed
+computing one: Multiple controllers may simultaneously try to
+install conflicting policies, and a consensus must be reached which one of
+the two is actually installed. Moreover, controllers may also fail,
+and a mechanism must be provided to complete an update in this scenario.
+
+\subsection{Our Contributions}
+
+In this paper, we propose an approach to solve these problems.
+In particular, we promote a very direct approach to coordinate controllers,
+using the Openflow bundle feature to make an atomic consensus and update:
+using one message, a single controller will be able to install its policy,
+\emph{in one shot}. Our approach relies on a classic result in 
+distributed computing, namely multi-writer consensus. Interestingly,
+this result is not well-known even in the distributed computing community,
+and we believe that the SDN paradigm offers a very practical
+application for it. We believe that our approach nicely complements
+existing, more theoretical literature on distributed control planes; for example,
+it shows how to realize STN~\cite{stn}.
+
+\subsection{Organization}
+
+The remainder of this paper is organized as follows.
+In Section~\ref{sec:background}, we provide the theoretical
+background on our approach. Section~\ref{sec:realization}
+then discusses the basic implementation of this approach in Openflow.
+Section~\ref{sec:extensions} discusses practical issues and extensions.
+After reviewing related literature in Section~\ref{sec:relwork},
+we conclude our work in
+Section~\ref{sec:conclusion}.
+
+\section{Fundamental Idea}\label{sec:background}
+
+idea: multi-write consensus, a not well-known result!
 
@@
 
+\section{One-Touch Policy Update}\label{sec:realization}
+
+SDN switch configuration space as shared memory
+
+We address the problem of committing consistent policies to SDN switch concurrently by multiple controllers.
+We show that assuming minimalistic control protocol, SDN switch configuration space can be used to implement a consensus object thereby allowing controllers to agree on each next policy update.
+
+Our first observation is that a match-action configuration which is accessed by multiple controllers can be considered as shared memory.
+Note however that standard memory map and index/offset to a word, while Flow Table entries are not indexed.
+To overcome this we can use the match part of an entry as the index and the action part as the word value.
+
 Going back to the policy update problem, the consensus scheme allow all servers to decide on next policy by using consensus value as policy identifier, and writing the policy in advance.
@@
 Making this scheme fail safe (handling failure of winning server) can be ensured by allowing any server to reconfigure the switch.
-
-\section{The one touch policy update problem}\label{sec:todo}
 
@@
 
-\section{Dealing with multiple updates}\label{sec:todo}
+\section{Discussion and Extensions}\label{sec:extension}
+
+\subsection{Dealing with multiple updates}
 
@@
 
+\section{Related Work}\label{sec:relwork}
+
+ideas: distributed control planes
+
+ideas: known consensus results and impossibilities
+
+ideas: the power of in-band
+
+\section{Conclusion}\label{sec:conclusion}
+
+ideas: the missing link for STN, but also others
+
+{
 \bibliographystyle{abbrv}
 \bibliography{references}  % main.bib is the name of the Bibliography in this case
+}
 

### 2015-02-24 inband-STN.tex
blabla
added efficient alg.

@@
 \definecolor{shadecolor}{rgb}{0.9,0.9,0.9}
+
+%\usepackage{algorithm}
+%\usepackage[noend]{algorithmic}
+
+\usepackage{algorithmicx}
+\usepackage{algpseudocode}
+\usepackage{algorithm}
+\algrenewcommand\algorithmicreturn{\State \textbf{return}}
+\algdef{SE}[DOWHILE]{Do}{doWhile}{\algorithmicdo}[1]{\algorithmicwhile\ #1}%
+
+\algdef{SE}[TRANSACT]{StartTransaction}{EndTransaction}{\algorithmicaaa}[1]{\algorithmicwhile\ #1}%
+
+\algblock[<block>]{<start>}{<end>}
+\algblockdefx[Transaction]{startTransaction}{endTransaction} %
+[0]{ \textbf{start transaction}} %
+[1][res]{ $#1\gets$  \textbf{commit transaction}}
+
 %
@@
 
-\section{Fundamental Idea}\label{sec:background}
+\section{The Underlying Theory}\label{sec:background}
 
@@
 
-\section{One-Touch Policy Update}\label{sec:realization}
-
-SDN switch configuration space as shared memory
+\section{Basic One-Touch Mechanism}\label{sec:realization}
 
@@
 
+\subsection{SDN switch configuration space as shared memory}
+
 Our first observation is that a match-action configuration which is accessed by multiple controllers can be considered as shared memory.
-Note however that standard memory map and index/offset to a word, while Flow Table entries are not indexed.
-To overcome this we can use the match part of an entry as the index and the action part as the word value.
+Note however that standard memory map and index/offset to a word, while Flow Table entries are not necessarily indexed.
+However, we assume that each flow entry has some property that allows the operator to specifically reference and update it, for example OpenFlow entry's cookie.
+In cases where this property value can't be determined by the operator, we can use other variable entry properties to store the index, for example we can represent memory value $v$ in offset $i$ by an match-action entry whose match is $in\_port = i$ and action is "forward port $x$".
+%To overcome this we can use the match part of an entry as the index and the action part as the word value.
 
@@
 Making this scheme fail safe (handling failure of winning server) can be ensured by allowing any server to reconfigure the switch.
+
+\subsection{ One-Touch}
 
@@
 
-\section{Discussion and Extensions}\label{sec:extension}
+\section{The Practical Implementation}\label{sec:extension}
 
@@
 \section{Improving the scheme with advanced OpenFlow features}\label{sec:todo}
+Till now we considered a model that allows multiple updates to match-action entries.
+Here we extends the model with actions create-entry and delete-entry that may failed and abort the whole transaction depending on the existence on non-existence of specific entries prior to the action.
+%TBD - much less resources. easily solving the infinite versions issue.
 
-TBD - much less resources. easily solving the infinite versions issue.
+Next we show how such actions can be used to implement consensus with much less resources. We use entry id as a property used to reference specific entry.
+
+\begin{algorithm}[t]
+    \caption{Advanced Update Algorithm}
+    \label{alg:template}
+    \begin{algorithmic}[1]    
+    \Require update requests $U$
+    \Ensure installed policy is consistent with previous one
+    \If {first time}
+		%\Do 
+		\Repeat 
+			\State $my\_cur\_id\gets$ unused entry id 
+    		\State $res \gets $ create entry with id $my\_cur\_id$
+    	\Until{$res=False$}
+    	%\doWhile{$res=False$}
+    \EndIf
+ 		%\Do 
+ 		\Repeat 
+ 			\State $cur\_id\gets$ read current entry id 
+ 			\State $cur\_policy\gets$ read current policy
+ 			\State $my\_next\_id\gets$ unused entry id 
+ 			\State $new\_policy\gets$ apply update requests on $cur\_policy$
+ 			\startTransaction
+	 			\State delete entry with id $1<<(\log n) + cur\_id$
+	 			\State create entry with id $1<<(\log n) + my\_cur\_id$
+	 			\State add $new\_policy$ rules conditioned on $metadata=my\_cur\_id$
+	 			\State update policy selector to $my\_cur\_id$
+ 			\endTransaction
+     		\State $res \gets $ commit transaction 
+     	\Until{$res=False$}
+     	%\doWhile{$res=False$}
+%   \vspace{3mm}
+
+    \end{algorithmic}
+\end{algorithm}
 

### 2015-02-26 inband-STN.tex
bla

@@
 
-\title{One-Shot Coordination of Distributed SDN Controllers:\\On the Consensus Power of OpenFlow}
+%\title{One-Shot Coordination of Distributed SDN Controllers:\\On the Consensus Power of OpenFlow}
+
+\title{Coordinating Distributed SDN Controllers:\\On the Consensus Power of OpenFlow}
 
@@
 \begin{abstract}
-The control plane of future Software-Defined Networks (SDNs)
-is likely to be \emph{distributed}: a redundant control plane
+The control plane of modern Software-Defined Networks (SDNs)
+is \emph{distributed}: a redundant control plane
 can ensure a high availability, and a spatially distributed
 control plane allows to handle time-critical dataplane
-events close to their origin. Moreover, future SDNs may also
-allow multiple administrators or participatory users to update
-policies concurrently, e.g., by offering transactional interfaces. 
+events close to their origins. Moreover, an SDN may 
+be managed by multiple administrators simultaneously. 
 However, not much is known today about how to implement
-a consistent distributed control plane which supports concurrent
-policy installations.
-
-In this paper, we propose to coordinate distributed controllers
-directly on the Openflow switches: exploiting classic results
-on multi-writer consensus, we show how to use Openflow 
-to synchronize controllers and install policies in a consistent
-manner, \emph{in one shot}. Our approach is attractive as it
-does not require complicated helping mechanisms in case of controller
-failures. 
+such a distributed control plane. 
+In this paper, we show that powerful coordination objects
+can be implemented using today's Openflow switches.
+In particular, we describe the implementation of test\&set primitives 
+as well as multi-writer consensus objects which can be used
+to implement atomic transactions with all-or-nothing semantics.
+Interestingly, our approach allows controllers to detect and
+resolve very general notions of conflicts, not only, e.g., conflicts
+due to overlapping flow spaces but also conflicts based on network load.
 \end{abstract}
@@
 A particularly interesting problem regards the consistent installation of new policies
-or routes.~\cite{todo,todo,todo} While implementing consistent network updates 
+or routes.~\cite{network-update,roger-hotnets,correct} While implementing consistent network updates
 is challenging already from the perspective of a single controller, there exists a wide consensus
@@
 First, in order to provide a high availability as well as a minimal degree of
-fault-tolerance, controllers should be redundant~\cite{onix,stn,todo}: a failure
+fault-tolerance, controllers should be redundant~\cite{onix,stn,onos}: a failure
 of one controller can be masked by other controllers. Second, it has also been proposed
@@
 such distributed control planes. The problem is essentially a distributed
-computing one: Multiple controllers may simultaneously try to
-install conflicting policies, and a consensus must be reached which one of
-the two is actually installed. Moreover, controllers may also fail,
-and a mechanism must be provided to complete an update in this scenario.
+systems one: Multiple controllers may simultaneously try to
+install conflicting updates, e.g., rules (of the same priority) defined over non-disjoint
+flow space, or load-balancing policies which, if combined, actually \emph{increase} the load~\cite{log-cent}. 
+Moreover, controllers may also fail,
+before their update has been completed.
 
@@
 
-In this paper, we propose an approach to solve these problems.
-In particular, we promote a very direct approach to coordinate controllers,
-using the Openflow bundle feature to make an atomic consensus and update:
-using one message, a single controller will be able to install its policy,
-\emph{in one shot}. Our approach relies on a classic result in 
-distributed computing, namely multi-writer consensus. Interestingly,
-this result is not well-known even in the distributed computing community,
-and we believe that the SDN paradigm offers a very practical
-application for it. We believe that our approach nicely complements
+In this paper, we show that powerful synchronization primitives
+can be implemented in Openflow, allowing controllers to implement
+non-trivial atomic transactions, and detect and
+automatically resolve conflicts. In particular, we show how to implement
+the classic distributed computing primitive
+\emph{test\&set}, supporting the conditional installation of new rules,
+and we show how to use the bundle feature to implement
+multi-writer objects, an efficient means to compute consensus;
+interestingly, our implementation can also be used to compute a consensus
+and install the winner's rule \emph{atomically}, using one message.
+Moreover, our approach allows to detect and resolve very general notions of conflicts:
+conflicts may be defined, for example, for rules of the same priority over
+defined over non-independent flow spaces, but may also depend, e.g., on 
+network load (based on counters).
+We believe that our approach nicely complements
 existing, more theoretical literature on distributed control planes; for example,
-it shows how to realize STN~\cite{stn}.
+it also shows how to realize the atomic read-modify-write primitive postulated in
+STN~\cite{stn}.
 
@@
 
-ideas: distributed control planes
+There exists a wide consensus that SDN control planes need to be distributed.~\cite{onos,onix,elasticon} 
+Onix~\cite{onix} is one of the earliest proposals, and introduced the 
+the notion of Network Information Base (NIB), abstracting network state
+distribution from control logic, but 
+requiring mechanisms for the detection and resolution of conflicts. 
+Also spatially distributed control planes to improve scalability and
+latency have been studied intensively
+in the literature~\cite{kandoo,ctrl-place,hotsdn13loc}
+proposes an elastic distributed controller architecture.
 
-ideas: known consensus results and impossibilities
+Operating and updating SDNs consistently is already non-trivial
+from the perspective of a single controller. Reitblatt et al.~~\cite{network-update} 
+introduced the notion of
+per-packet consistency and proposed a 2-phase update protocol based on tagging.
+%, and described the \emph{two-phase update} technique, also used in our algorithms.
+Mahajan and Wattenhofer~\cite{roger-hotnets} considered weaker transient
+consistency guarantees, and proposed more efficient network update algorithms
+accordingly. Ludwig et al.~~\cite{hotnets14update} studied algorithms for secure
+network updates where packets are forced to traverse certain waypoints or
+middleboxes. Ghorbani et al.~~\cite{correct-virt} recently argued for the design
+of network update algorithms that provide even stronger consistency guarantees.
 
-ideas: the power of in-band
+To the best of our knowledge, the only paper explicitly addressing the consistent
+network update problem from a distributed controller perspective is STN~\cite{stn}:
+STN provides a transactional interface offering all-or-nothing semantics and serializability
+(the ``holy grail'' of safety properties), 
+allowing
+controllers to install policies in a conflict-free manner; if some controllers fail,
+other controllers can take over. STN~\cite{stn} is based on 
+atomic
+read-modify-write primitives, but while this primitive is postulated, no implementation is
+described. In this paper, we provide this missing link, and also show that other useful and
+much more powerful synchronization primitives can be implemented.
+
+Our work is also closely related to the recent work on providing more high-level
+abstractions and programming languages for SDNs, such as Frenetic~\cite{frenetic}, 
+also enabling policy composition~\cite{pyretic}.
+
+In terms of distributed computing: TODO PETR: maybe modify a bit the text below
+and also write more about multi-writer consensus etc.?
+
+\noindent\textbf{Distributed Computing.}
+There is a long tradition of defining correctness of a concurrent system via
+an equivalence to a sequential one~\cite{Pap79-serial,Lam79,HW90}.  The notion
+of sequentially composable histories is reminiscent of
+linearizability~\cite{HW90}, where a history of operations concurrently
+applied by a collection of processes is equivalent to a history in which the
+operations are in a sequential order, respecting their real-time precedence.
+In contrast, our sequentially composable histories impose requirements not
+only on high-level invocations and responses, but also on the way the traffic
+is processed. We require that the committed policies constitute a
+conflict-free sequential history, but, additionally,  we expect that each
+\emph{path} witnesses only a prefix of this history, consisting of all
+requests that were committed before the path was initiated.
+%
+The transactional interface exported by the CPC abstraction is inspired by the
+work on speculative concurrency control using software transactional memory
+(STM)~\cite{stm-st95}.
+%, thus the  term \emph{software transactional networking}.
+Our interface is however intended to model realistic network
+management operations, which makes it simpler than recent
+dynamic STM models~\cite{dstm}.
+%In particular, the sets of rules to be installed by a policy update do not depend on the state of the
+%network.
+
+Extending the interface to dynamic policies that adapt their behavior based on
+the current network state sounds like a promising research direction.  On the
+other hand, our criterion of sequential composability is more complex than
+traditional STM correctness properties in that it imposes restrictions not only
+on the high-level interface exported to the control plane, but also on the
+paths taken by the data-plane packets.
+Also, we assumed that controllers are subject to failures, which is usually not
+assumed by STM implementations.
 
@@
 
-ideas: the missing link for STN, but also others
+This paper showed that powerful synchronization primitives
+from distributed computing can be implemented in Openflow. 
+Our work complements existing research on the design of
+distributed control planes, and also provides some missing links,
+e.g., for the transactional approach taken by STN. 
+We also hope that our work can nourish the ongoing discussion of
+what additional in-band features may be useful for future
+versions of Openflow.
 

### 2015-03-05 inband-STN.tex
new intro (and story)

@@
 \usepackage{cite}
+\usepackage{url}
+
 %\usepackage{framed}
@@
 
-%\title{Distributed Network Programming}
-%\title{The Distributed SDN Update Problem:\\Towards Software Transactional Networks}
+%\title{Solving Consensus with OpenFlow}
 
-%\title{On Consistent Updates in Software Defined Networks under Unreliable Control}
-%\title{The Case for Reliable Software Transactional Networking}
-
-%\title{A Distributed SDN Control Plane for Concurrent Policy Updates}
-
-\title{Solving Consensus with OpenFlow}
-
-\title{One-Shot Coordination of Distributed SDN Controllers:\\The Case for Multi-Write Consensus}
+%\title{One-Shot Coordination of Distributed SDN Controllers:\\The Case for Multi-Write Consensus}
 
@@
 
-\title{Coordinating Distributed SDN Controllers:\\On the Consensus Power of OpenFlow}
+%\title{Coordinating Distributed SDN Controllers:\\On the Consensus Power of OpenFlow}
+
+\title{On the Synchronization Power of OpenFlow}
 
@@
 Stefan Schmid$^2$, Petr Kuznetsov$^3$ \\
-\small $^1$ Tel Aviv University, Israel; $^2$ TU Berlin \& T-Labs, Germany; $^3$ ParisTech, France
+\small $^1$ Tel Aviv University, Israel; $^2$ TU Berlin \& T-Labs,
+Germany; $^3$ T\'el\'ecom ParisTech, France
 }
@@
 \begin{abstract}
-The control plane of modern Software-Defined Networks (SDNs)
-is \emph{distributed}: a redundant control plane
-can ensure a high availability, and a spatially distributed
-control plane allows to handle time-critical dataplane
-events close to their origins. Moreover, an SDN may 
-be managed by multiple administrators simultaneously. 
-However, not much is known today about how to implement
-such a distributed control plane. 
-In this paper, we show that powerful coordination objects
-can be implemented using today's Openflow switches.
-In particular, we describe the implementation of test\&set primitives 
-as well as multi-writer consensus objects which can be used
-to implement atomic transactions with all-or-nothing semantics.
-Interestingly, our approach allows controllers to detect and
-resolve very general notions of conflicts, not only, e.g., conflicts
-due to overlapping flow spaces but also conflicts based on network load.
+Control planes of modern Software-Defined Networks (SDNs)
+are typically \emph{distributed systems}: to achieve availability,
+load balancing and responsiveness, control modules are installed on
+multiple machines.   
+%a redundant control plane
+%can ensure a high availability, and a spatially distributed
+%control plane allows to handle time-critical dataplane
+%events close to their origins. 
+%[[PK this is a contraversial point, not important for us 
+%Moreover, an SDN may be managed by multiple administrators simultaneously. 
+%]]
+Unfortunately, these goals are hard to combine with consistent network
+manipulation and certain synchronization challenges among the
+distributed controllers must be addressed. 
+%[[PK sounds contradicting to the fact that such systems already exist
+%However, not much is known today about how to implement
+%such a distributed control plane. 
+%]]
+In this paper, we show that powerful synchronization mechanisms
+can be implemented \emph{in band}, simply using today's Openflow switches.
+We show that standard OpenFlow features allow for providing simple
+but general-purpose transactions with all-or-nothing semantics. 
+We then show that these transactions can be used to reach consensus
+among software controllers and, even more surprizingly, to resolve
+\emph{on-the-fly} conflicts between concurrent policy updates in an
+effectively atomic way, without affecting the ongoing traffic.  
+%[[PK
+%In particular, we describe the implementation of test\&set primitives 
+%as well as multi-writer consensus objects which can be used
+%to implement atomic transactions with all-or-nothing semantics.
+%Interestingly, our approach allows controllers to detect and
+%resolve very general notions of conflicts, not only, e.g., conflicts
+%due to overlapping flow spaces but also conflicts based on network load.
+%]]
 \end{abstract}
@@
 By consolidating and outsourcing the control over the dataplane switches to a logically
-centralized controller,
-Software-defined networks (SDNs) simplify the network management and facilitate faster
-innovations, as the (software) control plane is decoupled and can evolve independently.
-Moreover, Openflow, the standard SDN protocol today, introduces many generalizations,
-in terms of traffic engineering, definition of flows, as well as in-band network functionalities,
-by relying on a simple match-action paradigm which allows us to define
-forwarding rules based not only on Layer-2, but also Layer-3 and Layer-4 (and possibly beyond)
-header fields.
+centralized controller, the concept Software-Defined Networking (SDN)
+simplifies network management and facilitates faster innovations, 
+as the (software) control plane is decoupled from the (hardware) data
+plane and can evolve independently.
+%[[PK do not see why it should be here
+%Moreover, Openflow, the standard SDN protocol today, introduces many generalizations,
+%in terms of traffic engineering, definition of flows, as well as in-band network functionalities,
+%by relying on a simple match-action paradigm which allows us to define
+%forwarding rules based not only on Layer-2, but also Layer-3 and Layer-4 (and possibly beyond)
+%header fields.
+%]]
+%
+The distinct network control plane has the potential to simplify
+network management
+%[[PK did not get
+%---many network tasks are inherently non-local---
+%]]
+and opens new opportunities for innovation,
+%[[PK we rather  talk about implementing it? 
+%managing and operating an SDN in 
+but implementing it 
+%]]
+%[[PK adaptive to what?
+%a more adaptive but 
+%]]
+in a scalable and  consistent manner is highly non-trivial.
+%[[PK did not get ``even''
+%, even with SDN.
+%]]
 
-While the logically centralized network perspective has the potential to simplify
-the network management---many network tasks are inherently non-local---and introduces
-many flexibilities,
-managing and operating an SDN in a more adaptive but consistent manner is non-trivial, even with SDN.
-A particularly interesting problem regards the consistent installation of new policies
-or routes.~\cite{network-update,roger-hotnets,correct} While implementing consistent network updates
-is challenging already from the perspective of a single controller, there exists a wide consensus
-that the logically centralized SDN control planes will likely be \emph{actually distributed} in the future.
-
-The reason for implementing SDN control planes in a distributed manner are manyfold.
+Indeed, there is a wide consensus that to achieve availability,
+responsiveness, and balancing of network management tasks, 
+the \emph{logically} centralized SDN
+control plane must be physically \emph{distributed}.
 First, in order to provide a high availability as well as a minimal degree of
@@
 participatory interfaces where different users can install and trigger policy changes
-concurrently.~\cite{participatory,stn}
+concurrently~\cite{participatory,stn}.
+All this, however, comes with the need for a \emph{synchronization}
+between distributed controllers, assuming that we want them to
+manipulate the network \emph{consistently}~\cite{cpc}.
+And this task is highly nontrivial~\cite{cap-theorem}.
 
-However, today, we do not have a good understanding of how to realize
-such distributed control planes. The problem is essentially a distributed
-systems one: Multiple controllers may simultaneously try to
-install conflicting updates, e.g., rules (of the same priority) defined over non-disjoint
-flow space, or load-balancing policies which, if combined, actually \emph{increase} the load~\cite{log-cent}. 
-Moreover, controllers may also fail,
-before their update has been completed.
+%[[PK not sure I see why talking about it here
+%[[
+%A particularly interesting problem regards the consistent installation of new policies
+%or routes~\cite{network-update,roger-hotnets,correct,stn}. While implementing consistent network updates
+%is challenging already from the perspective of a single controller, there exists a wide consensus
+%that the logically centralized SDN control planes will likely be \emph{actually distributed} in the future.
+%]]
 
-\subsection{Our Contributions}
+Today, unfortunately, we do not have a good understanding yet of how to realize
+such distributed control planes. The problem is essentially a
+distributed-systems 
+one: Multiple controllers may simultaneously try to
+install conflicting updates and we want to resolve these conflicts
+\emph{consistently} (no undesired behavior is observed on the data
+plane) and \emph{efficiently} (no undesired delays are imposed on the
+control application).
 
-In this paper, we show that powerful synchronization primitives
-can be implemented in Openflow, allowing controllers to implement
-non-trivial atomic transactions, and detect and
-automatically resolve conflicts. In particular, we show how to implement
-the classic distributed computing primitive
-\emph{test\&set}, supporting the conditional installation of new rules,
-and we show how to use the bundle feature to implement
-multi-writer objects, an efficient means to compute consensus;
-interestingly, our implementation can also be used to compute a consensus
-and install the winner's rule \emph{atomically}, using one message.
-Moreover, our approach allows to detect and resolve very general notions of conflicts:
-conflicts may be defined, for example, for rules of the same priority over
-defined over non-independent flow spaces, but may also depend, e.g., on 
-network load (based on counters).
-We believe that our approach nicely complements
-existing, more theoretical literature on distributed control planes; for example,
-it also shows how to realize the atomic read-modify-write primitive postulated in
-STN~\cite{stn}.
+Consider, for example, the particularly interesting problem of
+consistent installation of new forwarding policies, stipulating routes
+that packets of different header spaces should follow across the
+network~\cite{network-update,roger-hotnets,correct,stn}.
+Installing conflicting forwarding rules, e.g., rules of the same priority defined over non-disjoint
+flow spaces may lead to pathological network behavior (loops,
+bleckholes, etc.).
+Similarly, installing diverging load-balancing policies may, when combined, \emph{increase} the load~\cite{log-cent}. 
+To render things more difficult, controllers may also fail, even before their updates have been completed.
 
-\subsection{Organization}
+%\subsection{In band synchronization?}
+%
+In this paper, we look closely at the most popular specification
+language for SDN control-data plane exchange -
+OpenFlow~\cite{of-spec}. 
+At a high level, the specification represents the exchange in the form
+of \emph{control messages}, containing commands for installing or
+modifying \emph{rules} that, in turn, constitute the \emph{flow table} of
+a data-plane switch used in processing the ongoing data-plane traffic.
 
-The remainder of this paper is organized as follows.
+We observe that the standard provides the option of \emph{bundling}
+these messages in one, so that either all of the control messages are
+processed or none of them is. In particular, if, for some reason, the
+configuration request contained in one of the bundled messages cannot
+be processed, all the other requests  are rejected and an error
+message is sent to the controller that issued them.       
+Interestingly, this bundled processing can be, when properly configured, done in a
+\emph{packet-atomic} way, so that each packet is processed according
+to the switch configuration either \emph{before} or \emph{after} the
+request was processed.
+
+Intuitively, if we treat the flow table at the switch as a \emph{shared memory}
+that can be concurrently read and modified by multiple controllers, 
+the bundling features can be seen as a \emph{multi-write} or, more
+widely, \emph{multi-modify} operation that modifies
+multiple memory location in one atomic step.
+This ability is reminiscent of the hypothetical multi-write
+shared-memory model which is known to enable solutions for important
+synchronization problems, such as consensus, impossible to solve
+otherwise~\cite{multi-objects}. 
+Indeed, we describe an algorithm that solves consensus among any
+number of controllers using a single OpenFlow switch that accepts
+bundle control messages.
+
+Even more, we observe that the standard also assumes \emph{conditional} application of
+certain types of control messages. A specific $\textsf{OFPFF}\_\textsf{CHECK}\_\textsf{OVERLAP}$ flag
+set in the control message containing a new forwarding rule ensures
+that the rule is applied if and only if it does not \emph{conflict}
+with any other rule currently in the flow table.  
+This in-band conflict detection mechanism, combined with bundling,
+allows us to implement specific types of \emph{transactions} that may
+atomically install multiple rules if certain compatibility conditions
+are met.
+
+We then discuss how these synchronization mechanisms can be used to
+achieve consistent network management. 
+We first show that they can be used to replace the read-modify-write
+primitives voluntarily assumed in the recent implementation of a
+recent proposal for a consistent composition system for forwarding policies~\cite{cpc}.  
+We then use these transactions directly to implement a lightweight
+forwarding policy composition system that requires no additional
+synchronization on the control plane.     
+We then explore how more general classes of policies can be composed
+\emph{on-the-switch}.
+
+% we show that powerful synchronization primitives
+%can be implemented in Openflow, allowing controllers to implement
+%non-trivial atomic transactions, and detect and
+%automatically resolve conflicts. In particular, we show how to implement
+%the classic distributed computing primitive
+%\emph{test\&set}, supporting the conditional installation of new rules,
+%and we show how to use the bundle feature to implement
+%multi-writer objects, an efficient means to compute consensus;
+%interestingly, our implementation can also be used to compute a consensus
+%and install the winner's rule \emph{atomically}, using one message.
+%Moreover, our approach allows to detect and resolve very general notions of conflicts:
+%conflicts may be defined, for example, for rules of the same priority over
+%defined over non-independent flow spaces, but may also depend, e.g., on 
+%network load (based on counters).
+%We believe that our approach nicely complements
+%existing, more theoretical literature on distributed control planes; for example,
+%it also shows how to realize the atomic read-modify-write primitive postulated in
+%STN~\cite{stn}.
+
+%\paragraph{Organization.}
+%
+The rest of the paper is organized as follows.
 In Section~\ref{sec:background}, we provide the theoretical

### 2015-03-06 inband-STN.tex
intro update, reorganised the content a bit

@@
 Even more, we observe that the standard also assumes \emph{conditional} application of
-certain types of control messages. A specific $\textsf{OFPFF}\_\textsf{CHECK}\_\textsf{OVERLAP}$ flag
+certain types of control messages. A specific $\textsf{OFPFF\_CHECK\_OVERLAP}$ flag
 set in the control message containing a new forwarding rule ensures
@@
 \emph{on-the-switch}.
+
+Our results highlight the surprising synchronization power hidden in the
+OpenFlow standard. The synchronization mechanisms we describe in this
+paper turn out to be crucial for ensuring consistency of network
+management on the control plane while preserving consistency of the
+traffic processed on the data plane.
 
@@
 The rest of the paper is organized as follows.
-In Section~\ref{sec:background}, we provide the theoretical
-background on our approach. Section~\ref{sec:realization}
-then discusses the basic implementation of this approach in Openflow.
-Section~\ref{sec:extensions} discusses practical issues and extensions.
-After reviewing related literature in Section~\ref{sec:relwork},
-we conclude our work in
-Section~\ref{sec:conclusion}.
+In Section~\ref{sec:model}, we overview the basics of the  OpenFlow
+protocol and recall the bundle and \textsf{ofpff\_check\_overlap} features. 
+In Section~\ref{sec:sync}, we specify the implementations of our OpenFlow-based
+sycnhronization mechanisms. 
+In Section~\ref{sec:application}, we discuss a few policy composition
+algorithms that benefit from these mechanisms. 
+In Section~\ref{sec:conclusion}, we discuss open questions and related work.
+
+%we provide the theoretical
+%background on our approach. Section~\ref{sec:realization}
+%then discusses the basic implementation of this approach in Openflow.
+%Section~\ref{sec:extensions} discusses practical issues and extensions.
+%After reviewing related literature in Section~\ref{sec:relwork},
+%we conclude our work in
+%Section~\ref{sec:conclusion}.
+
+\section{SDN primer and OpenFlow features}\label{sec:model}
+
+Define the protocol (from Reclaiming the Brain"?) and discuss bundling
+and check-overlap.
+
+\section{Transactions in OpenFlow}\label{sec:sync}
+
+Multi-write and read-modify transactions.
+
+\subsection{Dealing with multiple updates}
+
+TBD - adding the notion of versions and cleanup/GC. probably no good way to solve infinite versions... maybe check literature on shared memory and multiple writes.
+
+\subsection{Improving the scheme with check-overlap}\label{sec:todo}
+Till now we considered a model that allows multiple updates to match-action entries.
+Here we extends the model with actions create-entry and delete-entry that may failed and abort the whole transaction depending on the existence on non-existence of specific entries prior to the action.
+%TBD - much less resources. easily solving the infinite versions issue.
+
+Next we show how such actions can be used to implement consensus with much less resources. We use entry id as a property used to reference specific entry.
+
+\begin{algorithm}[t]
+    \caption{Advanced Update Algorithm}
+    \label{alg:template}
+    \begin{algorithmic}[1]
+    \Require update requests $U$
+    \Ensure installed policy is consistent with previous one
+    \If {first time}
+		%\Do
+		\Repeat
+			\State $my\_cur\_id\gets$ unused entry id
+    		\State $res \gets $ create entry with id $my\_cur\_id$
+    	\Until{$res=False$}
+    	%\doWhile{$res=False$}
+    \EndIf
+ 		%\Do
+ 		\Repeat
+ 			\State $cur\_id\gets$ read current entry id
+ 			\State $cur\_policy\gets$ read current policy
+ 			\State $my\_next\_id\gets$ unused entry id
+ 			\State $new\_policy\gets$ apply update requests on $cur\_policy$
+ 			\startTransaction
+	 			\State delete entry with id $1<<(\log n) + cur\_id$
+	 			\State create entry with id $1<<(\log n) + my\_cur\_id$
+	 			\State add $new\_policy$ rules conditioned on $metadata=my\_cur\_id$
+	 			\State update policy selector to $my\_cur\_id$
+ 			\endTransaction
+     		\State $res \gets $ commit transaction
+     	\Until{$res=False$}
+     	%\doWhile{$res=False$}
+%   \vspace{3mm}
+
+    \end{algorithmic}
+\end{algorithm}
+
+\section{Applications and ramifications}\label{sec:application}
 
@@
 \section{The Practical Implementation}\label{sec:extension}
-
-\subsection{Dealing with multiple updates}
-
-TBD - adding the notion of versions and cleanup/GC. probably no good way to solve infinite versions... maybe check literature on shared memory and multiple writes.
-
-\section{Improving the scheme with advanced OpenFlow features}\label{sec:todo}
-Till now we considered a model that allows multiple updates to match-action entries.
-Here we extends the model with actions create-entry and delete-entry that may failed and abort the whole transaction depending on the existence on non-existence of specific entries prior to the action.
-%TBD - much less resources. easily solving the infinite versions issue.
-
-Next we show how such actions can be used to implement consensus with much less resources. We use entry id as a property used to reference specific entry.
-
-\begin{algorithm}[t]
-    \caption{Advanced Update Algorithm}
-    \label{alg:template}
-    \begin{algorithmic}[1]
-    \Require update requests $U$
-    \Ensure installed policy is consistent with previous one
-    \If {first time}
-		%\Do
-		\Repeat
-			\State $my\_cur\_id\gets$ unused entry id
-    		\State $res \gets $ create entry with id $my\_cur\_id$
-    	\Until{$res=False$}
-    	%\doWhile{$res=False$}
-    \EndIf
- 		%\Do
- 		\Repeat
- 			\State $cur\_id\gets$ read current entry id
- 			\State $cur\_policy\gets$ read current policy
- 			\State $my\_next\_id\gets$ unused entry id
- 			\State $new\_policy\gets$ apply update requests on $cur\_policy$
- 			\startTransaction
-	 			\State delete entry with id $1<<(\log n) + cur\_id$
-	 			\State create entry with id $1<<(\log n) + my\_cur\_id$
-	 			\State add $new\_policy$ rules conditioned on $metadata=my\_cur\_id$
-	 			\State update policy selector to $my\_cur\_id$
- 			\endTransaction
-     		\State $res \gets $ commit transaction
-     	\Until{$res=False$}
-     	%\doWhile{$res=False$}
-%   \vspace{3mm}
-
-    \end{algorithmic}
-\end{algorithm}
-
-\section{Solving just concurrent policy update without consensus}\label{sec:todo}
-
-TBD - should be simple given two previous sections
 

### 2015-03-06 inband-STN.tex
rewrite algs

--- 2015-02-26 inband-STN.tex
@@
     \begin{algorithmic}[1]
-    \Require update requests $U$
+    \Require update requirements $U$, two entries containers CUR and NEXT.
+    \Ensure installed policy is consistent with previous one
+    \Repeat
+			\State $my\_id\gets$ unused id (in both CUR and NEXT)
+    		\State $res \gets $ NEXT.Add($my\_id$, True)
+    \Until{$res=True$}
+ 		%\Do
+ 		\Repeat
+ 			\State $pid\gets$ read current policy id (only one in CUR)
+ 			\State NEXT.Add($pid$)
+ 			\If {$pid\notin CUR$}
+ 				\State {\bf continue} (restart loop)
+ 			\EndIf
+ 			\State $cur\_policy\gets$ read current policy
+ 			\State $new\_policy\gets apply\_update(cur\_policy,U)$
+ 			\startTransaction
+	 			\State CUR.Delete($b_pid$)
+	 			\State CUR.Add($my\_id$)
+				\State NEXT.Delete($my\_id$, True)
+	 			\State add $new\_policy$ rules %conditioned on $metadata=my\_a$
+	 			%\State update policy selector to $my\_a$
+ 			\endTransaction
+     		%\State $res \gets $ commit transaction
+     	\Until{$res=True$}
+     	%\doWhile{$res=False$}
+%   \vspace{3mm}
+			\Return
+
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[t]
+    \caption{Container Add abstraction}
+    \label{alg:template}
+    \begin{algorithmic}[1]
+    \Require Container id $C$, entry id $i$, unique constrain (defaults to False) $b\_unique$, caller id $caller$.
+%    \Ensure 
+    		\State $value \gets C.i$
+    		\State $mask \gets all-ones$
+    		\State $cookie \gets C.i.caller$
+    		\If {$b\_unique$}
+    			\State $flag \gets 0$
+    		\Else
+    		    \State $flag \gets CHECK\_OVERLAP$
+    		\EndIf
+    					\State $cmd\gets FlowMod(match=(value,mask), operation = ADD, cookie, flag) $
+			\Return cmd
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[t]
+    \caption{Container Delete abstraction}
+    \label{alg:template}
+    \begin{algorithmic}[1]
+    \Require Container id $C$, entry id $i$, unique constrain (defaults to False) $b\_unique$, caller id $caller$.
+%    \Ensure 
+    		\State $value \gets C.i$
+    		\State $mask \gets all-ones$
+    		\State $cookie \gets C.i.caller$
+    		\If {$b\_unique$}
+    			\State $cmd\gets FlowMod(match=(value,mask), operation=DELETE) $
+    		\Else
+    		    \State $cmd\gets FlowMod(match=(value,mask), operation=DELETE_STRICT, cookie) $
+    		\EndIf
+
+			\Return cmd
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[t]
+    \caption{Advanced Update Algorithm2}
+    \label{alg:template}
+    \begin{algorithmic}[1]
+    \Require update requirements $U$, two entries containers CUR and NEXT.
     \Ensure installed policy is consistent with previous one
@@
 		\Repeat
-			\State $my\_cur\_id\gets$ unused entry id
-    		\State $res \gets $ create entry with id $my\_cur\_id$
-    	\Until{$res=False$}
+			\State $my\_id\gets$ unused id (in both CUR and NEXT)
+    		\State $res \gets $ NEXT.Add($my\_id$, True)
+    	\Until{$res=True$}
     	%\doWhile{$res=False$}
@@
  		\Repeat
- 			\State $cur\_id\gets$ read current entry id
+ 			\State $pid\gets$ read current policy id (only one in CUR)
+ 			\State NEXT.Add($pid$)
+ 			\If {$pid\notin CUR$}
+ 				\State {\bf continue} (restart loop)
+ 			\EndIf
  			\State $cur\_policy\gets$ read current policy
- 			\State $my\_next\_id\gets$ unused entry id
- 			\State $new\_policy\gets$ apply update requests on $cur\_policy$
+ 			\State $next\_id\gets$  unused id (in both CUR and NEXT)
+ 			\State $new\_policy\gets apply\_update(cur\_policy,U)$
  			\startTransaction
-	 			\State delete entry with id $1<<(\log n) + cur\_id$
-	 			\State create entry with id $1<<(\log n) + my\_cur\_id$
-	 			\State add $new\_policy$ rules conditioned on $metadata=my\_cur\_id$
-	 			\State update policy selector to $my\_cur\_id$
+	 			\State CUR.Delete($b_pid$)
+	 			\State CUR.Add($my\_id$)
+				\State NEXT.Delete($my\_id$, True)
+				\State NEXT.Add($next\_id$, True)
+	 			\State add $new\_policy$ rules %conditioned on $metadata=my\_a$
+	 			%\State update policy selector to $my\_a$
  			\endTransaction
-     		\State $res \gets $ commit transaction
-     	\Until{$res=False$}
+     		%\State $res \gets $ commit transaction
+     	\Until{$res=True$}
+     	\State $my\_id\gets next\_id$
      	%\doWhile{$res=False$}
@@
 
+			\Return
     \end{algorithmic}

### 2015-03-07 inband-STN.tex
stefan

@@
 
-\title{On the Synchronization Power of OpenFlow}
+%\title{On the Synchronization Power of OpenFlow}
+
+\title{Coordinating Distributed SDN Controllers\\{\Large On the Synchronization Power of OpenFlow}}
 
@@
 Control planes of modern Software-Defined Networks (SDNs)
-are typically \emph{distributed systems}: to achieve availability,
-load balancing and responsiveness, control modules are installed on
-multiple machines.   
+are \emph{distributed systems}: to ensure availability and fault-tolerance,
+to improve load-balancing, and to reduce overheads, 
+control modules are physically distributed.
 %a redundant control plane
@@
 %]]
-Unfortunately, these goals are hard to combine with consistent network
-manipulation and certain synchronization challenges among the
-distributed controllers must be addressed. 
+However, a distributed control plane also introduces
+new challenges, and in order to guarantee a consistent network operation,
+the actions of different controllers may need to be synchronized and coordinated.
 %[[PK sounds contradicting to the fact that such systems already exist
@@
 can be implemented \emph{in band}, simply using today's Openflow switches.
-We show that standard OpenFlow features allow for providing simple
+We show how to leverage standard OpenFlow features for providing simple
 but general-purpose transactions with all-or-nothing semantics.
 We then show that these transactions can be used to reach consensus
-among software controllers and, even more surprizingly, to resolve
-\emph{on-the-fly} conflicts between concurrent policy updates in an
-effectively atomic way, without affecting the ongoing traffic.  
+among software controllers and, what is more, to resolve
+general conflicts between concurrent policy updates in an
+effectively atomic way and \emph{on-the-fly}, without affecting the ongoing traffic.
 %[[PK
@@
 plane and can evolve independently.
-%[[PK do not see why it should be here
-%Moreover, Openflow, the standard SDN protocol today, introduces many generalizations,
-%in terms of traffic engineering, definition of flows, as well as in-band network functionalities,
-%by relying on a simple match-action paradigm which allows us to define
-%forwarding rules based not only on Layer-2, but also Layer-3 and Layer-4 (and possibly beyond)
-%header fields.
-%]]
-%
-The distinct network control plane has the potential to simplify
-network management
-%[[PK did not get
-%---many network tasks are inherently non-local---
-%]]
-and opens new opportunities for innovation,
-%[[PK we rather  talk about implementing it? 
-%managing and operating an SDN in 
-but implementing it 
-%]]
-%[[PK adaptive to what?
-%a more adaptive but 
-%]]
-in a scalable and  consistent manner is highly non-trivial.
-%[[PK did not get ``even''
-%, even with SDN.
-%]]
+Moreover, Openflow, the standard SDN protocol today, introduces many generalizations,
+in terms of traffic engineering, definition of flows, as well as in-band network functionalities,
+by relying on a simple match-action paradigm which allows us to define
+forwarding rules based not only on Layer-2, but also Layer-3 and Layer-4 (and possibly beyond)
 
-Indeed, there is a wide consensus that to achieve availability,
-responsiveness, and balancing of network management tasks, 
-the \emph{logically} centralized SDN
-control plane must be physically \emph{distributed}.
+While the logically centralized perspective offered by SDN is attractive,
+there is a wide consensus that
+the control plane should be physically \emph{distributed}. 
 First, in order to provide a high availability as well as a minimal degree of
@@
 concurrently~\cite{participatory,stn}.
-All this, however, comes with the need for a \emph{synchronization}
-between distributed controllers, assuming that we want them to
-manipulate the network \emph{consistently}~\cite{cpc}.
-And this task is highly nontrivial~\cite{cap-theorem}.
+All this, however, comes with the need for \emph{synchronization}
+and \emph{coordination} among the distributed controllers:
+the network should be manipulated \emph{consistently}~\cite{cpc}.
+This task is nontrivial~\cite{cap-theorem}.
 
@@
 
-Today, unfortunately, we do not have a good understanding yet of how to realize
+Today, we do not have a good understanding yet of how to realize
 such distributed control planes. The problem is essentially a
-distributed-systems 
+distributed systems
 one: Multiple controllers may simultaneously try to
@@
 
-Consider, for example, the particularly interesting problem of
+Consider, for example, the practical problem of
 consistent installation of new forwarding policies, stipulating routes
@@
 flow spaces may lead to pathological network behavior (loops,
-bleckholes, etc.).
-Similarly, installing diverging load-balancing policies may, when combined, \emph{increase} the load~\cite{log-cent}. 
-To render things more difficult, controllers may also fail, even before their updates have been completed.
+blackholes, etc.).
+Similarly, installing diverging load-balancing policies may, 
+when combined, \emph{increase} the load~\cite{log-cent}.
+To render things more difficult, controllers may also fail, 
+even before their updates have been completed.
 
-%\subsection{In band synchronization?}
-%
-In this paper, we look closely at the most popular specification
-language for SDN control-data plane exchange -
-OpenFlow~\cite{of-spec}. 
-At a high level, the specification represents the exchange in the form
-of \emph{control messages}, containing commands for installing or
-modifying \emph{rules} that, in turn, constitute the \emph{flow table} of
-a data-plane switch used in processing the ongoing data-plane traffic.
-
-We observe that the standard provides the option of \emph{bundling}
-these messages in one, so that either all of the control messages are
-processed or none of them is. In particular, if, for some reason, the
-configuration request contained in one of the bundled messages cannot
-be processed, all the other requests  are rejected and an error
-message is sent to the controller that issued them.       
-Interestingly, this bundled processing can be, when properly configured, done in a
-\emph{packet-atomic} way, so that each packet is processed according
-to the switch configuration either \emph{before} or \emph{after} the
-request was processed.
-
-Intuitively, if we treat the flow table at the switch as a \emph{shared memory}
-that can be concurrently read and modified by multiple controllers, 
-the bundling features can be seen as a \emph{multi-write} or, more
-widely, \emph{multi-modify} operation that modifies
-multiple memory location in one atomic step.
+%\textbf{} 
+\paragraph{Our Contributions}
+In this paper, we show that 
+OpenFlow~\cite{of-spec}---the standard SDN protocol today---can
+be used to provide strong controller synchronization and coordination 
+primitives \emph{in-band}.
+In particular, we show how to exploit the \emph{bundling}
+feature and the $\textsf{OFPFF\_CHECK\_OVERLAP}$ flag,
+to implement a \emph{shared memory} abstraction of the switch
+and its flow tables: 
+This shared memory 
+allows us to implement
+\emph{transactions} which may
+atomically install multiple rules if certain compatibility conditions
+are met, and which also facilitate \emph{packet-atomic} rule updates. 
+These transactions can be \emph{multi-write} or even \emph{multi-modify}:
+while not supporting
+read operations, the transaction can be conditioned on the presence of
+certain rules.
 This ability is reminiscent of the hypothetical multi-write
@@
 
-Even more, we observe that the standard also assumes \emph{conditional} application of
-certain types of control messages. A specific $\textsf{OFPFF\_CHECK\_OVERLAP}$ flag
-set in the control message containing a new forwarding rule ensures
-that the rule is applied if and only if it does not \emph{conflict}
-with any other rule currently in the flow table.  
-This in-band conflict detection mechanism, combined with bundling,
-allows us to implement specific types of \emph{transactions} that may
-atomically install multiple rules if certain compatibility conditions
-are met.
-
-We then discuss how these synchronization mechanisms can be used to
-achieve consistent network management. 
-We first show that they can be used to replace the read-modify-write
-primitives voluntarily assumed in the recent implementation of a
-recent proposal for a consistent composition system for forwarding policies~\cite{cpc}.  
-We then use these transactions directly to implement a lightweight
-forwarding policy composition system that requires no additional
-synchronization on the control plane.     
-We then explore how more general classes of policies can be composed
-\emph{on-the-switch}.
-
-Our results highlight the surprising synchronization power hidden in the
+Our results highlight the synchronization power hidden in the
 OpenFlow standard. The synchronization mechanisms we describe in this
@@
 traffic processed on the data plane.
+As a case study, we also show how to replace the read-modify-write
+primitives voluntarily assumed in the recent implementation of a
+consistent composition system for forwarding policies~\cite{cpc}.
+We also show how to use these transactions directly to implement a lightweight
+forwarding policy composition system that requires no additional
+synchronization on the control plane.
+Finally, we explore how more general classes of policies can be composed
+\emph{on-the-switch}.
 
@@
 
-%\paragraph{Organization.}
-%
+\paragraph{Organization}
 The rest of the paper is organized as follows.
@@
 
-\section{SDN primer and OpenFlow features}\label{sec:model}
+\section{OpenFlow Background}\label{sec:model}
 
-Define the protocol (from Reclaiming the Brain"?) and discuss bundling
-and check-overlap.
+The synchronization and coordination mechanisms presented in this paper
+are based on Openflow~\cite{of-spec}, the standard SDN protocol today. In this section, 
+we provide the necessary background.
+
+es a small partial deployment of SDN and
+out-sources the control over a set of Openflow switches to a logically centralized software controller.
+Openflow, the standard SDN protocol today, is based on a match-action paradigm: the controller
+can install simple rules on the Openflow switches.
+
+In a nutshell, 
+OpenFlow
+is based on a match-action concept: OpenFlow switches store
+rules (installed by the controller) consisting of a match and an
+action part. A packet matched by a certain rule will be subject
+to the associated action (to be immediately applied now or later). 
+For example, an action can define a port to which the
+matched packet should be forwarded, or add or change a tag
+(a certain part in the packet header).
+Each OpenFlow switch stores one or multiple flow tables,
+each of which contains a set of rules (flow entries). Flow
+tables form a pipeline, and flow entries are ordered according
+to priorities: A packet arriving at a switch is first checked by
+the rule of the highest priority in table 0: the fields of the
+data packet are compared with the match fields of that rule,
+and if they fit, some instructions (the actions) are executed;
+subsequently, lower priority rules are checked. Depending on
+the outcome of the table 0 processing, the packet may be sent
+to additional flow tables in the pipeline; during the traversal of
+the flow tables, the packet can be equipped with meta-tags to
+transfer temporary information. Concretely, instructions can be
+used to define additional tables to be visited (goto instruction),
+to modify the set of to-be-applied actions (either by appending,
+deleting, or modifying actions), or immediately apply some
+actions to the packet. A meta-data field can be used to exchange
+information between tables; it is part of the header that can be
+inserted or removed from a packet via push and pop actions.
+In general, a packet can be matched against any of its header
+fields, and fields can be wildcarded and sometimes bitmasked.
+(For instance, the meta-data field is maskable.) If no rule
+matches, the packet is dropped.
+
+\textbf{TODO LIRON: what else is missing? describe cookies?}
+
+The Openflow standard (as of \textbf{TODO}) offers two interesting
+features which we will exploit to implement the transactional memory abstraction:
+
+\begin{enumerate}
+\item \textbf{Bundling:}
+Control messages are aggregated into one message, so that either all of the control messages are
+processed, or none of them is. In particular, if, for some reason, the
+configuration request contained in one of the bundled messages cannot
+be processed, all the other requests  are rejected and an error
+message is sent to the controller that issued them.
+Interestingly, this bundled processing can be, when properly configured, done in a
+\emph{packet-atomic} way, so that each packet is processed according
+to the switch configuration either \emph{before} or \emph{after} the
+request was processed.
+
+\item \textbf{Conditional Application:}
+The Openflow standard also supports the \emph{conditional} application of
+certain types of control messages. A specific $\textsf{OFPFF\_CHECK\_OVERLAP}$ flag
+set in the control message containing a new forwarding rule ensures
+that the rule is applied if and only if it does not \emph{conflict}
+with any other rule currently in the flow table.
+This in-band conflict detection mechanism, combined with bundling,
+allows us to implement specific types of \emph{transactions} that may
+atomically install multiple rules if certain compatibility conditions
+are met.
+\end{enumerate}
+
+\section{Vision and Problem Statement}\label{sec:vision}
+
+This paper shows how to implement a transactional memory abstraction using
+standard Openflow switches. The shared memory abstraction of the 
+flow table at the switch allows multiple controllers to atomically and 
+concurrently read and modify state,
+using the bundling feature to implement \emph{multi-write} or, more
+widely, \emph{multi-modify} transactions. .
+This ability allows us to solve consensus among any
+number of controllers, using a single OpenFlow switch that accepts
+bundle control messages.
+
+These synchronization mechanisms can be used to
+achieve consistent network management.
+We first show that they can be used to replace the read-modify-write
+primitives voluntarily assumed in the recent implementation of a
+recent proposal for a consistent composition system for forwarding policies~\cite{cpc}.
+We then use these transactions directly to implement a lightweight
+forwarding policy composition system that requires no additional
+synchronization on the control plane.
+We then explore how more general classes of policies can be composed
+\emph{on-the-switch}.
+
+TODO: write vision and contribution in full detail, as an overview
 
@@
 \subsection{Improving the scheme with check-overlap}\label{sec:todo}
+
 Till now we considered a model that allows multiple updates to match-action entries.

### 2015-03-08 inband-STN.tex
minor alg changes
rewrite algs and suggested new ones

@@
 
+\newcommand{\hide}[1]{}
 \algblock[<block>]{<start>}{<end>}
@@
     \begin{algorithmic}[1]
-    \Require update requirements $U$, two entries containers CUR and NEXT.
+    \Require update requirements $U$, multi-set (with optional add validation) NEXT, set (with full validation) CUR.
     \Ensure installed policy is consistent with previous one
-    \Repeat
-			\State $my\_id\gets$ unused id (in both CUR and NEXT)
-    		\State $res \gets $ NEXT.Add($my\_id$, True)
-    \Until{$res=True$}
  		%\Do
  		\Repeat
- 			\State $pid\gets$ read current policy id (only one in CUR)
+ 			\State read switch configuration
+ 			\State $my\_id\gets$ unused id (in both CUR and NEXT)
+ 			\State $pid\gets$ current policy id (only id in CUR)
  			\State NEXT.Add($pid$)
+ 			\State read switch configuration
  			\If {$pid\notin CUR$}
@@
  			\EndIf
- 			\State $cur\_policy\gets$ read current policy
+ 			\State $cur\_policy\gets$ current policy
  			\State $new\_policy\gets apply\_update(cur\_policy,U)$
  			\startTransaction
-	 			\State CUR.Delete($b_pid$)
-	 			\State CUR.Add($my\_id$)
+ 				\State NEXT.Add($my\_id$, True)
  				\State NEXT.Delete($my\_id$, True)
+	 			\State DeleteGroup($pid$)
+	 			\State AddGroup($my\_id$)
+				\State NEXT.Delete($pid$)
 	 			\State add $new\_policy$ rules %conditioned on $metadata=my\_a$
@@
 \begin{algorithm}[t]
-    \caption{Container Add abstraction}
+    \caption{multi-set Add abstraction}
     \label{alg:template}
     \begin{algorithmic}[1]
-    \Require Container id $C$, entry id $i$, unique constrain (defaults to False) $b\_unique$, caller id $caller$.
+    \Require id $i$, unique constrain (defaults to False) $b\_unique$, caller id $caller$ (defaults to self).
 %    \Ensure
-    		\State $value \gets C.i$
+    		\State $value \gets i$
     		\State $mask \gets all-ones$
-    		\State $cookie \gets C.i.caller$
+    		\State $cookie \gets i.caller$
+    		\State $action \gets i.caller$
     		\If {$b\_unique$}
@@
     		\EndIf
-    					\State $cmd\gets FlowMod(match=(value,mask), operation = ADD, cookie, flag) $
+    					\State $cmd\gets FlowMod(match=(value,mask), operation = ADD, cookie, flag, action) $
 			\Return cmd
@@
 \begin{algorithm}[t]
-    \caption{Container Delete abstraction}
+    \caption{multi-set Delete abstraction}
     \label{alg:template}
     \begin{algorithmic}[1]
-    \Require Container id $C$, entry id $i$, unique constrain (defaults to False) $b\_unique$, caller id $caller$.
+    \Require id $i$, unique constrain (defaults to False) $b\_unique$, caller id $caller$.
 %    \Ensure
@@
     		\Else
-    		    \State $cmd\gets FlowMod(match=(value,mask), operation=DELETE_STRICT, cookie) $
+    		    \State $cmd\gets FlowMod(match=(value,mask), operation=DELETE\_STRICT, cookie) $
     		\EndIf
@@
 \begin{algorithm}[t]
-    \caption{Advanced Update Algorithm2}
+    \caption{set Add abstraction}
     \label{alg:template}
     \begin{algorithmic}[1]
-    \Require update requirements $U$, two entries containers CUR and NEXT.
+    \Require  id $i$.
+%    \Ensure
+    		\State $command \gets OFPGC\_ADD$
+    		\State $type \gets OFPGT\_INDIRECT$
+    		\State $group_id \gets i$
+    		\State $buckets \gets [dummy\_backet]$
+    		\State $cmd\gets GroupMod(command, type, group_id, buckets) $
+			\Return cmd
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[t]
+    \caption{set Delete abstraction}
+    \label{alg:template}
+    \begin{algorithmic}[1]
+    \Require id $i$, unique constrain (defaults to False) $b\_unique$, caller id $caller$.
+%    \Ensure
+    		\State $command \gets OFPGC\_DELETE$
+    		\State $group_id \gets i$
+    		\State $cmd\gets GroupMod(command, group\_id) $
+			\Return cmd
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[t]
+    \caption{Update Algorithm with only a set}
+    \label{alg:template}
+    \begin{algorithmic}[1]
+    \Require update requirements $U$, a set CUR.
     \Ensure installed policy is consistent with previous one
-    \If {first time}
  		%\Do
  		\Repeat
-			\State $my\_id\gets$ unused id (in both CUR and NEXT)
-    		\State $res \gets $ NEXT.Add($my\_id$, True)
-    	\Until{$res=True$}
-    	%\doWhile{$res=False$}
-    \EndIf
- 		%\Do
- 		\Repeat
- 			\State $pid\gets$ read current policy id (only one in CUR)
- 			\State NEXT.Add($pid$)
- 			\If {$pid\notin CUR$}
- 				\State {\bf continue} (restart loop)
- 			\EndIf
- 			\State $cur\_policy\gets$ read current policy
- 			\State $next\_id\gets$  unused id (in both CUR and NEXT)
+ 			\State read switch configuration
+ 			\State $pid\gets$ current policy id (only id in CUR)
+ 			\State $cur\_policy\gets$ current policy
  			\State $new\_policy\gets apply\_update(cur\_policy,U)$
  			\startTransaction
-	 			\State CUR.Delete($b_pid$)
-	 			\State CUR.Add($my\_id$)
-				\State NEXT.Delete($my\_id$, True)
-				\State NEXT.Add($next\_id$, True)
+	 			\State CUR.Delete($pid$)
+	 			\State CUR.Add($pid+1$)
 	 			\State add $new\_policy$ rules %conditioned on $metadata=my\_a$
@@
      	\Until{$res=True$}
-     	\State $my\_id\gets next\_id$
      	%\doWhile{$res=False$}
 %   \vspace{3mm}
+			\Return
 
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[t]
+    \caption{Update Algorithm with pipeline-array}
+    \label{alg:template}
+    \begin{algorithmic}[1]
+    \Require update requirements $U$, array $ARR$.
+    \Ensure installed policy is consistent with previous one
 			\Return
+
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[t]
+    \caption{Update Composition Algorithm without bundle}
+    \label{alg:template}
+    \begin{algorithmic}[1]
+    \Require update requirements $U$, array $ARR$.
+    \Ensure installed policy is consistent with previous one
+			\Return
+
     \end{algorithmic}

### 2015-03-08 inband-STN.tex
stefan iteration

@@
 
-%\title{Solving Consensus with OpenFlow}
+%\title{Solving Consensus with Openflow}
 
@@
 
-%\title{One-Shot Coordination of Distributed SDN Controllers:\\On the Consensus Power of OpenFlow}
+%\title{One-Shot Coordination of Distributed SDN Controllers:\\On the Consensus Power of Openflow}
 
-%\title{Coordinating Distributed SDN Controllers:\\On the Consensus Power of OpenFlow}
+%\title{Coordinating Distributed SDN Controllers:\\On the Consensus Power of Openflow}
 
-%\title{On the Synchronization Power of OpenFlow}
+%\title{On the Synchronization Power of Openflow}
 
-\title{Coordinating Distributed SDN Controllers\\{\Large On the Synchronization Power of OpenFlow}}
+%\title{Synchronizing Distributed SDN Controllers\\{\Large The Openflow Container}}
+
+\title{Synchronizing Distributed Openflow Controllers\\{\Large Toward More Powerful In-Band Synchronization Objects}}
 
@@
 new challenges, and in order to guarantee a consistent network operation,
-the actions of different controllers may need to be synchronized and coordinated.
+the actions of different controllers may need to be synchronized.
 %[[PK sounds contradicting to the fact that such systems already exist
@@
 %]]
-In this paper, we show that powerful synchronization mechanisms
-can be implemented \emph{in band}, simply using today's Openflow switches.
-We show how to leverage standard OpenFlow features for providing simple
-but general-purpose transactions with all-or-nothing semantics.
-We then show that these transactions can be used to reach consensus
-among software controllers and, what is more, to resolve
-general conflicts between concurrent policy updates in an
-effectively atomic way and \emph{on-the-fly}, without affecting the ongoing traffic.
+In this paper, we introduce a novel synchronization object which 
+we call \emph{container}.
+The container can be implemented \emph{in-band}, using the standard Openflow 
+protocol 
+which support atomic multi-writes but only limited reads (namely 
+testing the presence of rules). We describe how to  
+leverage the container object to 
+implement powerful consensus objects, transactions with all-or-nothing
+ semantics, or atomically resolve
+general conflicts between concurrent policy updates 
+\emph{on-the-fly}, without affecting the ongoing traffic.
 %[[PK
@@
 By consolidating and outsourcing the control over the dataplane switches to a logically
-centralized controller, the concept Software-Defined Networking (SDN)
-simplifies network management and facilitates faster innovations,
-as the (software) control plane is decoupled from the (hardware) data
-plane and can evolve independently.
+centralized controller, Software-Defined Networks (SDNs)
+simplify network management and facilitate faster innovations:
+the (software) control plane can evolve independently from the 
+(hardware) data plane.
 Moreover, Openflow, the standard SDN protocol today, introduces many generalizations,
@@
 by relying on a simple match-action paradigm which allows us to define
-forwarding rules based not only on Layer-2, but also Layer-3 and Layer-4 (and possibly beyond)
+forwarding rules based not only on Layer-2, but also Layer-3 and Layer-4 (and possibly beyond).
 
@@
 concurrently~\cite{participatory,stn}.
-All this, however, comes with the need for \emph{synchronization}
-and \emph{coordination} among the distributed controllers:
+All this, however, comes with the need for \emph{synchronizing}
+the distributed controllers:
 the network should be manipulated \emph{consistently}~\cite{cpc}.
@@
 that packets of different header spaces should follow across the
-network~\cite{network-update,roger-hotnets,correct,stn}.
+network.~\cite{network-update,roger-hotnets,correct,stn}
 Installing conflicting forwarding rules, e.g., rules of the same priority defined over non-disjoint
@@
 In this paper, we show that
-OpenFlow~\cite{of-spec}---the standard SDN protocol today---can
-be used to provide strong controller synchronization and coordination 
-primitives \emph{in-band}.
-In particular, we show how to exploit the \emph{bundling}
-feature and the $\textsf{OFPFF\_CHECK\_OVERLAP}$ flag,
-to implement a \emph{shared memory} abstraction of the switch
-and its flow tables: 
-This shared memory 
-allows us to implement
-\emph{transactions} which may
-atomically install multiple rules if certain compatibility conditions
-are met, and which also facilitate \emph{packet-atomic} rule updates. 
-These transactions can be \emph{multi-write} or even \emph{multi-modify}:
-while not supporting
-read operations, the transaction can be conditioned on the presence of
-certain rules.
-This ability is reminiscent of the hypothetical multi-write
+Openflow~\cite{of-spec}---the standard SDN protocol today---can
+be used to implement powerful, high-level 
+synchronization
+objects \emph{in-band}.
+In particular, we introduce a novel synchronization object 
+called \emph{container}. A container can be seen as
+a variant of the classic \texttt{test\&set} object
+(known from distributed computing over shared memory),
+which is tailored to Openflow: while Openflow supports 
+atomic multi-write functionality (using bundling), 
+read functionality is limited to testing the presence of
+rules as well as flow space overlaps.
+
+The container object is reminiscent of the hypothetical multi-write
 shared-memory model which is known to enable solutions for important
@@
 otherwise~\cite{multi-objects}.
-Indeed, we describe an algorithm that solves consensus among any
-number of controllers using a single OpenFlow switch that accepts
-bundle control messages.
-
-Our results highlight the synchronization power hidden in the
-OpenFlow standard. The synchronization mechanisms we describe in this
-paper turn out to be crucial for ensuring consistency of network
+Indeed, we show how to exploit the container object to implement 
+consensus among controllers. 
+More generally, the container object turns out to be 
+crucial for ensuring consistency of network
 management on the control plane while preserving consistency of the
-traffic processed on the data plane.
-As a case study, we also show how to replace the read-modify-write
+traffic processed on the data plane, and we show how to implement
+\emph{packet-atomic} rule updates and more complex
+\emph{atomic transactions}.
+We also show how to replace the read-modify-write
 primitives voluntarily assumed in the recent implementation of a
 consistent composition system for forwarding policies~\cite{cpc}.
-We also show how to use these transactions directly to implement a lightweight
-forwarding policy composition system that requires no additional
-synchronization on the control plane.
-Finally, we explore how more general classes of policies can be composed
-\emph{on-the-switch}.
 
@@
 The rest of the paper is organized as follows.
-In Section~\ref{sec:model}, we overview the basics of the  OpenFlow
-protocol and recall the bundle and \textsf{ofpff\_check\_overlap} features.
-In Section~\ref{sec:sync}, we specify the implementations of our OpenFlow-based
-sycnhronization mechanisms.
-In Section~\ref{sec:application}, we discuss a few policy composition
-algorithms that benefit from these mechanisms.
-In Section~\ref{sec:conclusion}, we discuss open questions and related work.
+In Section~\ref{sec:background}, we review the basics of the  Openflow
+protocol and discuss the features which are relevant for our work.
+We introduce our vision and the container consensus object in Section~\ref{sec:vision},
+and present its implementation in Section~\ref{sec:container}.
+We then discuss applications (Section~\ref{sec:apps}) as
+well as extensions (Section~\ref{sec:discussion}).
+After reviewing related literature in Section~\ref{sec:relwork}, we conclude
+our contribution in Section~\ref{sec:conclusion}.
 
@@
 
-\section{OpenFlow Background}\label{sec:model}
+\section{Openflow Background}\label{sec:background}
 
-The synchronization and coordination mechanisms presented in this paper
+The synchronization mechanisms presented in this paper
 are based on Openflow~\cite{of-spec}, the standard SDN protocol today. In this section,
@@
 
-es a small partial deployment of SDN and
-out-sources the control over a set of Openflow switches to a logically centralized software controller.
-Openflow, the standard SDN protocol today, is based on a match-action paradigm: the controller
-can install simple rules on the Openflow switches.
-
 In a nutshell,
-OpenFlow
-is based on a match-action concept: OpenFlow switches store
+Openflow
+is based on a match-action concept: Openflow switches store
 rules (installed by the controller) consisting of a match and an
@@
 (a certain part in the packet header).
-Each OpenFlow switch stores one or multiple flow tables,
+Each Openflow switch stores one or multiple flow tables,
 each of which contains a set of rules (flow entries). Flow
@@
 
-\textbf{TODO LIRON: what else is missing? describe cookies?}
-
-The Openflow standard (as of \textbf{TODO}) offers two interesting
+The Openflow standard (as of \textbf{TODO}) offers three interesting
 features which we will exploit to implement the transactional memory abstraction:
@@
 are met.
+
+\item \textbf{Cookies:} Cookies allow to name and retrieve individual
+flow rules. Our synchronization mechanism will implement controller-specific cookies. 
 \end{enumerate}
 
-\section{Vision and Problem Statement}\label{sec:vision}
+\section{The Container Object}\label{sec:vision}
 
-This paper shows how to implement a transactional memory abstraction using
-standard Openflow switches. The shared memory abstraction of the 
-flow table at the switch allows multiple controllers to atomically and 
-concurrently read and modify state,
-using the bundling feature to implement \emph{multi-write} or, more
-widely, \emph{multi-modify} transactions. .
-This ability allows us to solve consensus among any
-number of controllers, using a single OpenFlow switch that accepts
-bundle control messages.
+The container synchronization object introduced in this paper
+is essentially a transactional memory abstraction which can be implemented
+in standard Openflow switches. 
+Before presenting the technical details in the next section,
+we give a brief overview.
 
-These synchronization mechanisms can be used to
+The postulated container object can store an arbitrary number of \emph{items}.
+As we will see, these items correspond to flow entries on the Openflow switch.
+The same item may even exist multiple times in the container (i.e., the container
+is a multi-set).
+The container
+provides two main functions:
+\begin{enumerate}
+\item \textbf{\emph{Conditional Multi-Add($S$):}} A set $S$ of items $s\in S$
+is (atomically) added to the container, if and only if no element $s\in S$
+already existed in the container. If such an element already existed,
+the operation does not take effect (i.e., the transaction is aborted).
+
+\item \textbf{\emph{Conditional Multi-Remove($S$):}} A set $S$ of items $s\in S$
+is (atomically) removed from the container, if and only if each element $s\in S$
+already existed in the container. If one or more elements did not exist,
+the operation does not take effect (i.e., the transaction is aborted).
+\end{enumerate}
+
+The container object allows controllers to atomically and
+concurrently read and modify state. This  
+allows us, for example, to solve consensus among any
+number of controllers.
+More generally, the container object
+allows us to
 achieve consistent network management.
-We first show that they can be used to replace the read-modify-write
-primitives voluntarily assumed in the recent implementation of a
-recent proposal for a consistent composition system for forwarding policies~\cite{cpc}.
-We then use these transactions directly to implement a lightweight
-forwarding policy composition system that requires no additional
-synchronization on the control plane.
-We then explore how more general classes of policies can be composed
-\emph{on-the-switch}.
 
-TODO: write vision and contribution in full detail, as an overview
+\section{Openflow Implementation}\label{sec:container}
 
-\section{Transactions in OpenFlow}\label{sec:sync}
+We now describe how to implement the container object
+in Openflow. For our implementation, we will make
+use of three main Openflow features: \emph{bundling},
+\emph{conditional application}, and \emph{cookies}.
+We will describe later how to implement some of these features
+in older Openflow versions where they are not supported. \textbf{TODO}
 
-Multi-write and read-modify transactions.
+A container contains items: they represent the flow table entries (i.e., a match-action rules, plus e.g., counters belonging to the rule).
+Items can be added and removed atomically and in bundles (e.g., to update the switch with
+a new policy). Recall that we want to add items only if they do not exist
+yet; otherwise the transaction should abort without taking effect.
+And similarly for removes: transactions should not take effect if the corresponding
+update contains rules which currently do not exist at the switch.
 
-\subsection{Dealing with multiple updates}
-
-TBD - adding the notion of versions and cleanup/GC. probably no good way to solve infinite versions... maybe check literature on shared memory and multiple writes.
-
-\subsection{Improving the scheme with check-overlap}\label{sec:todo}
-
-Till now we considered a model that allows multiple updates to match-action entries.
-Here we extends the model with actions create-entry and delete-entry that may failed and abort the whole transaction depending on the existence on non-existence of specific entries prior to the action.
-%TBD - much less resources. easily solving the infinite versions issue.
-
-Next we show how such actions can be used to implement consensus with much less resources. We use entry id as a property used to reference specific entry.
+Algorithm~\ref{alg:update} lies at the heart of our mechanism.
+It takes a set of rules $U$ (not yet a policy), 
+a multi-set $Next$ and a set (with full validation) $Cur$.
+The algorithm proceeds as follows: first, we seek to
+obtain a unique \emph{id}. FIXME: to be continued...
 
@@
     \caption{Advanced Update Algorithm}
-    \label{alg:template}
+    \label{alg:update}
     \begin{algorithmic}[1]
-    \Require update requirements $U$, multi-set (with optional add validation) NEXT, set (with full validation) CUR.
+    \Require update requirements $U$, multi-set (with optional add validation) $Next$, set (with full validation) $Cur$.
     \Ensure installed policy is consistent with previous one
@@
  		\Repeat
- 			\State read switch configuration
- 			\State $my\_id\gets$ unused id (in both CUR and NEXT)
- 			\State $pid\gets$ current policy id (only id in CUR)
- 			\State NEXT.Add($pid$)
- 			\State read switch configuration
- 			\If {$pid\notin CUR$}
- 				\State {\bf continue} (restart loop)
+ 			\State \textbf{read} switch configuration
+ 			\State $my\_id\gets$ unused \emph{id} (in both $Cur$ and $Next$)
+ 			\State $pid\gets$ current policy id (only id in $Cur$)
+ 			\State $Next$\emph{.add}($pid$)
+ 			\State \textbf{read} switch configuration
+ 			\If {$pid\notin Cur$}
+ 				\State {\bf continue} (\textbf{restart} loop)
  			\EndIf
@@
  			\startTransaction
- 				\State NEXT.Add($my\_id$, True)
- 				\State NEXT.Delete($my\_id$, True)
+ 				\State $Next$\emph{.add}($my\_id$, True)
+ 				\State $Next$\emph{.delete}($my\_id$, True)
 	 			\State DeleteGroup($pid$)
 	 			\State AddGroup($my\_id$)
-				\State NEXT.Delete($pid$)
-	 			\State add $new\_policy$ rules %conditioned on $metadata=my\_a$
+				\State $Next$\emph{.delete}($pid$)
+	 			\State \textbf{add} $new\_policy$ rules %conditioned on $metadata=my\_a$
 	 			%\State update policy selector to $my\_a$
@@
      		%\State $res \gets $ commit transaction
-     	\Until{$res=True$}
+     	\Until{$res=\texttt{true}$}
      	%\doWhile{$res=False$}
@@
 \begin{algorithm}[t]
-    \caption{multi-set Add abstraction}
-    \label{alg:template}
+    \caption{\textbf{\emph{Multi-Add()}}}
+    \label{alg:multi-add}
     \begin{algorithmic}[1]
-    \Require id $i$, unique constrain (defaults to False) $b\_unique$, caller id $caller$ (defaults to self).
+    \Require id $i$, unique constrain (defaults to \texttt{false}) $b\_unique$, caller id $caller$ (defaults to \emph{self}).
 %    \Ensure
     		\State $value \gets i$
-    		\State $mask \gets all-ones$
+    		\State $mask \gets all-1$
     		\State $cookie \gets i.caller$
@@
     		\EndIf
-    					\State $cmd\gets FlowMod(match=(value,mask), operation = ADD, cookie, flag, action) $
-			\Return cmd
+    					\State $cmd\gets \textsc{FlowMod}(match=(value,mask), operation = ADD, cookie, flag, action) $
+			\Return $cmd$
     \end{algorithmic}
@@
 \begin{algorithm}[t]
-    \caption{multi-set Delete abstraction}
+    \caption{\textbf{\emph{Multi-Remove()}}}
     \label{alg:template}
     \begin{algorithmic}[1]
-    \Require id $i$, unique constrain (defaults to False) $b\_unique$, caller id $caller$.
+    \Require id $i$, unique constrain (defaults to \texttt{false}) $b\_unique$, caller id $caller$.
 %    \Ensure
     		\State $value \gets C.i$
-    		\State $mask \gets all-ones$
+    		\State $mask \gets all-1$
     		\State $cookie \gets C.i.caller$
     		\If {$b\_unique$}
-    			\State $cmd\gets FlowMod(match=(value,mask), operation=DELETE) $
+    			\State $cmd\gets \textsc{FlowMod}(match=(value,mask), operation=DELETE) $
     		\Else
-    		    \State $cmd\gets FlowMod(match=(value,mask), operation=DELETE\_STRICT, cookie) $
+    		    \State $cmd\gets \textsc{FlowMod}(match=(value,mask), operation=DELETE\_STRICT, cookie) $
     		\EndIf
 
-			\Return cmd
+			\Return $cmd$
     \end{algorithmic}
@@
 \begin{algorithm}[t]
-    \caption{set Add abstraction}
+    \caption{\textbf{\emph{Add()}}}
     \label{alg:template}
@@
     		\State $cmd\gets GroupMod(command, type, group_id, buckets) $
-			\Return cmd
+			\Return $cmd$
     \end{algorithmic}
@@
 \begin{algorithm}[t]
-    \caption{set Delete abstraction}
+    \caption{\textbf{\emph{Remove()}}}
     \label{alg:template}
@@
     		\State $cmd\gets GroupMod(command, group\_id) $
-			\Return cmd
+			\Return $cmd$
     \end{algorithmic}
@@
     \begin{algorithmic}[1]
-    \Require update requirements $U$, a set CUR.
+    \Require update requirements $U$, a set $Cur$.
     \Ensure installed policy is consistent with previous one
@@
  		\Repeat
- 			\State read switch configuration
- 			\State $pid\gets$ current policy id (only id in CUR)
+ 			\State \textbf{read} switch configuration
+ 			\State $pid\gets$ current policy id (only id in $Cur$)
  			\State $cur\_policy\gets$ current policy
@@
  			\startTransaction
-	 			\State CUR.Delete($pid$)
-	 			\State CUR.Add($pid+1$)
+	 			\State $Cur$\emph{.delete}($pid$)
+	 			\State $Cur$\emph{.add}($pid+1$)
 	 			\State add $new\_policy$ rules %conditioned on $metadata=my\_a$
@@
 
+\section{Applications}\label{sec:apps}
+
+FIXME: STN, consensus, ...
+
+\paragraph{Consensus} FIXME: desribe how to implement consensus
+
+\paragraph{One-Shot Update} FIXME: desribe how to implement one-shot update
+
+\paragraph{STN} FIXME: desribe how to implement STN
+
+\section{Discussion and Extensions}\label{sec:discussion}
+
+FIXME: we could however even do stronger stuff: collision is only exact, not using masks: do not use the ternary nature
+
+\paragraph{Beyond Container Objects}
+FIXME: we can even do more general things!
+
+\paragraph{Older Openflow Versions} FIXME: describe matrix approach
+
+\section{In-Band Composition}\label{sec:compo}
+
+FIXME: shall we?
+
+\begin{comment}
+
+\section{Transactions in Openflow}\label{sec:sync}
+
+Multi-write and read-modify transactions.
+
+\subsection{Dealing with multiple updates}
+
+TBD - adding the notion of versions and cleanup/GC. probably no good way to solve infinite versions... maybe check literature on shared memory and multiple writes.
+
+\subsection{Improving the scheme with check-overlap}\label{sec:todo}
+
+Till now we considered a model that allows multiple updates to match-action entries.
+Here we extends the model with actions create-entry and delete-entry that may failed and abort the whole transaction depending on the existence on non-existence of specific entries prior to the action.
+%TBD - much less resources. easily solving the infinite versions issue.
+
+Next we show how such actions can be used to implement consensus with much less resources. We use entry id as a property used to reference specific entry.
+
 \section{Applications and ramifications}\label{sec:application}
@@
 
-In order to solve consensus we suggest to use the atomic multiple entries update capability (which is discussed in OpenFlow standard).
+In order to solve consensus we suggest to use the atomic multiple entries update capability (which is discussed in Openflow standard).
 A shared memory system that supports atomic write to (enough) multiple locations is known to allow the implementation of consensus objects.
@@
 Note however that standard memory map and index/offset to a word, while Flow Table entries are not necessarily indexed.
-However, we assume that each flow entry has some property that allows the operator to specifically reference and update it, for example OpenFlow entry's cookie.
+However, we assume that each flow entry has some property that allows the operator to specifically reference and update it, for example Openflow entry's cookie.
 In cases where this property value can't be determined by the operator, we can use other variable entry properties to store the index, for example we can represent memory value $v$ in offset $i$ by an match-action entry whose match is $in\_port = i$ and action is "forward port $x$".
@@
 In order to allow the packets to validate the consensus we make the following observation: while servers can be abstracted as memory writers, the packets processed by the switch are the readers.
-OpenFlow standard define an atomic write transaction as being atomic in relation to the processed packets thereby making packet processing a multiple read transaction but one read location per table is allowed.
+Openflow standard define an atomic write transaction as being atomic in relation to the processed packets thereby making packet processing a multiple read transaction but one read location per table is allowed.
 This means that if each memory location used in the consensus scheme will be stored in different table then each packet will be able to read all of them.
@@
 \section{The Practical Implementation}\label{sec:extension}
+
+\end{comment}
 
@@
 abstractions and programming languages for SDNs, such as Frenetic~\cite{frenetic},
-also enabling policy composition~\cite{pyretic}.
+also enabling policy composition~\cite{pyretic}. Our work complements these works
+on programming abstractoins
+by initiating the discussion of concurrency objects and abstractions.
 
@@
 
-This paper showed that powerful synchronization primitives
+This paper showed that powerful synchronization object
 from distributed computing can be implemented in Openflow.
@@
 e.g., for the transactional approach taken by STN.
-We also hope that our work can nourish the ongoing discussion of
-what additional in-band features may be useful for future
-versions of Openflow.
+We also hope that our work can nourish the discussion of
+which synchronization objects we can and want to implement
+in an SDN. While much progress has been made over the
+last years in the development of more intuitive programming
+languages for SDNs (e.g., Frenetic, Pyretic, Merlin), the 
+question of which synchronization abstractions to provide
+has received much less attention.
 
@@
 
+\begin{appendix}
+
+\section{Backup Stuff}
+
+using the bundling feature to implement \emph{multi-write} or, more
+widely, \emph{multi-modify} transactions. .
+, using a single Openflow switch that accepts
+bundle control messages.
+
+We first show that they can be used to replace the read-modify-write
+primitives voluntarily assumed in the recent implementation of a
+recent proposal for a consistent composition system for forwarding policies~\cite{cpc}.
+We then use these transactions directly to implement a lightweight
+forwarding policy composition system that requires no additional
+synchronization on the control plane.
+We then explore how more general classes of policies can be composed
+\emph{on-the-switch}.
+
+TODO: write vision and contribution in full detail, as an overview
+
+using the bundling feature to implement \emph{multi-write} or, more
+widely, \emph{multi-modify} transactions. .
+, using a single Openflow switch that accepts
+bundle control messages.
+
+We first show that they can be used to replace the read-modify-write
+primitives voluntarily assumed in the recent implementation of a
+recent proposal for a consistent composition system for forwarding policies~\cite{cpc}.
+We then use these transactions directly to implement a lightweight
+forwarding policy composition system that requires no additional
+synchronization on the control plane.
+We then explore how more general classes of policies can be composed
+\emph{on-the-switch}.
+
+TODO: write vision and contribution in full detail, as an overview
+
+we can do more than container:
+
+we introduce container: it's not compare+set, ... its something new!
+new object, is stronger than compare and set because bundle!
+
+\end{appendix}
+
 \end{document}

### 2015-03-08 inband-STN.tex
added short alg description.

--- 2015-03-08 inband-STN.tex
@@
 
+We use FlowMod command as abstraction to adding/deleting a number to a multi-set.
+The match part of the added flow entry represents the added number.
+In order to verify that the added number is unique we set the flag CHECK\_OVERLAP.
+Similarly, we use GroupMod command as abstraction to adding/deleting a number to a set.
+The group id of the added group represents the added number.
+
+In order to perform an update we use a set called CUR which is used to store the current policy id, and a multi-set called NEXT which is used to manage suggestions for next policy id.
+We start by reading the current switch configuration and selecting an unused id for our proposed policy.
+Next we try to prevent someone of using the current id with different policy in the future (durig our current update).
+
+We compute our new suggested policy by applying the updates request on top of current policy, supporting any kind of requests and policies.
+Then we make a transaction (using the bundle feature) to atomically check that our policy id is not blocked by any one else, to change the current policy id to ours (an action that would fail if the current policy id is no longer what we are counting of) and to actually configure our new policy.
+
+If one of the actions in the transaction fails we try again.
+There is no progress guaranty for each controller but there is one for the whole system - at least one of the controller will succeed in fulfilling its update requirements.
+
 \begin{algorithm}[t]

### 2015-03-09 inband-STN.tex
stefan merged liron

--- 2015-03-08 inband-STN.tex
@@
 
+LIRON TEXT: We use FlowMod command as abstraction to adding/deleting a number to a multi-set.
+The match part of the added flow entry represents the added number.
+In order to verify that the added number is unique we set the flag CHECK\_OVERLAP.
+Similarly, we use GroupMod command as an abstraction to adding/deleting a number to a set.
+The group id of the added group represents the added number.
+
+In order to perform an update we use a set called CUR which is used to store the current policy id, and a multi-set called NEXT which is used to manage suggestions for next policy id.
+We start by reading the current switch configuration and selecting an unused id for our proposed policy.
+Next we try to prevent someone of using the current id with different policy in the future (during our current update).
+
+We compute our new suggested policy by applying the update requests on top of current policy, supporting any kind of requests and policies.
+Then we make a transaction (using the bundle feature) to atomically check that our policy id is not blocked by anyone else, to change the current policy id to ours (an action that would fail if the current policy id is no longer what we are counting of) and to actually configure our new policy.
+
+If one of the actions in the transaction fails we try again.
+There is no progress guaranty for each controller but there is one for the whole system - at least one of the controller will succeed in fulfilling its update requirements.
+
 \section{Applications}\label{sec:apps}

### 2015-03-10 inband-STN.tex
small fixes to algs
sync interface

@@
 \newcommand{\hide}[1]{}
+\newcommand{\concat}[0]{\frown}
+\newcommand{\append}[0]{\frown}
+
 \algblock[<block>]{<start>}{<end>}
@@
 
-\section{The Container Object}\label{sec:vision}
+%\section{The Container Object}\label{sec:vision}
+\section{Transactional synchronization}\label{sec:trans}
 
-The container synchronization object introduced in this paper
-is essentially a transactional memory abstraction which can be implemented
-in standard Openflow switches.
-Before presenting the technical details in the next section,
-we give a brief overview.
+In this section, we sketch the synchronization abstraction we build on
+top of standard OpenFlow functionality. At a high level, we export a
+transactional interface which allows a controller to perform multiple
+operations on the configuration of an OpenFlow switch as an atomic
+\emph{transaction}, inspired by the concept of transactional
+memory~\cite{stm-st95,tm-book}. A transaction may \emph{commit}, in
+which case it appears as executed sequentially, or \emph{abort}, in
+which none of its operations takes effect.
 
-The postulated container object can store an arbitrary number of \emph{items}.
-As we will see, these items correspond to flow entries on the Openflow switch.
-The same item may even exist multiple times in the container (i.e., the container
-is a multi-set).
-The container
-provides two main functions:
-\begin{enumerate}
-\item \textbf{\emph{Conditional Multi-Add($S$):}} A set $S$ of items $s\in S$
-is (atomically) added to the container, if and only if no element $s\in S$
-already existed in the container. If such an element already existed,
-the operation does not take effect (i.e., the transaction is aborted).
+\subsection{Interface}\label{sec:t-if}
+We assume that transactions may comprise operations on the flow table
+of an SDN switch, such as reading flow tables or adding new rules. 
+ In addition, we also assume transactional access to  \emph{multi-set}
+abstractions that store \emph{items} and export \emph{add} and \emph{delete} operations with
+regular semantics. The multi-set abstractions are used for resolving
+conflicts between concurrent updates of the switch configuration and are themselves 
+implemented as flow-table components.
 
-\item \textbf{\emph{Conditional Multi-Remove($S$):}} A set $S$ of items $s\in S$
-is (atomically) removed from the container, if and only if each element $s\in S$
-already existed in the container. If one or more elements did not exist,
-the operation does not take effect (i.e., the transaction is aborted).
-\end{enumerate}
+A multi-set object also exports operations \emph{ms-add} and \emph{ms-delete} with the following
+semantics:
 
-The container object allows controllers to atomically and
-concurrently read and modify state. This
-allows us, for example, to solve consensus among any
-number of controllers.
-More generally, the container object
-allows us to
-achieve consistent network management.
+%The container synchronization object introduced in this paper
+%is essentially a transactional memory abstraction which can be implemented
+%in standard Openflow switches.
+%Before presenting the technical details in the next section,%
+%we give a brief overview.
 
-\section{Openflow Implementation}\label{sec:container}
+%The postulated container object can store an arbitrary number of \emph{items}.
+%As we will see, these items correspond to flow entries on the Openflow switch.
+%The same item may even exist multiple times in the container (i.e., the container
+%is a multi-set).
+%The container
+%provides two main functions:
+\begin{itemize}
+\item \textbf{$\textit{cond-add}(x)$:} if $x$ is not in the set, adds 
+  $x$ to the set; otherwise, \emph{fails} and, respectively,
+  aborts the transaction in which the operation is invoked.
+
+\item \textbf{$\textit{cond-delete}(x)$:} if $x$ is in the set, deletes an
+$x$ from the set; otherwise, \emph{fails} and
+  aborts the transaction.
+\end{itemize}
+
+Formally, we guarantee that in every execution, all committed
+transactions constitute a \emph{legal} sequential
+order~\cite{Pap79-serial}. On the progress side, we ensure that a
+transaction $T$ is only allowed to abort either because another 
+transaction updated a flow table entry read by $T$  
+or if one of $T$'s multi-set operations fails.
+
+Note that
+%, assuming such a transactional memory,  
+a $\emph{contains}(x)$ operation can be implemented as a transaction
+composed of $\textit{cond-add}(x)$ followed by $\textit{cond-delete}(x)$.  
+A \emph{set} abstraction is a special instance of a multi-set that is
+only modified with \textit{cond-add} and \textit{cond-delete}
+operations (this way the set only contains at most one copy of a given
+item).
+
+\subsection{Transactions in Openflow}\label{sec:t-impl}
 
@@
  				\State $Next$\emph{.delete}($my\_id$, True)
-	 			\State DeleteGroup($pid$)
-	 			\State AddGroup($my\_id$)
+	 			\State CUR\emph{.delete}($pid$)
+	 			\State CUR\emph{.add}($my\_id$)
 				\State $Next$\emph{.delete}($pid$)
@@
 \begin{algorithm}[t]
-    \caption{\textbf{\emph{Multi-Add()}}}
+    \caption{\textbf{\emph{MultiSet-Add()}}}
     \label{alg:multi-add}
@@
 %    \Ensure
-    		\State $value \gets i$
-    		\State $mask \gets all-1$
-    		\State $cookie \gets i.caller$
-    		\State $action \gets i.caller$
+    		\State $value \gets i\concat i$
+    		\State $mask \gets caller\concat all-1$
+    		\State $cookie \gets i\concat caller$
+    		\State $action \gets i\concat caller$
     		\If {$b\_unique$}
@@
 \begin{algorithm}[t]
-    \caption{\textbf{\emph{Multi-Remove()}}}
+    \caption{\textbf{\emph{MultiSet-Remove()}}}
     \label{alg:template}
@@
 \begin{algorithm}[t]
-    \caption{\textbf{\emph{Add()}}}
+    \caption{\textbf{\emph{Set-Add()}}}
     \label{alg:template}
@@
 \begin{algorithm}[t]
-    \caption{\textbf{\emph{Remove()}}}
+    \caption{\textbf{\emph{Set-Remove()}}}
     \label{alg:template}

### 2015-03-10 inband-STN.tex
another update on the TM and alto description

@@
 
-A multi-set object also exports operations \emph{ms-add} and \emph{ms-delete} with the following
+A multi-set object also exports operations \emph{cond-add} and \emph{cond-delete} with the following
 semantics:
@@
 
-We now describe how to implement the container object
-in Openflow. For our implementation, we will make
-use of three main Openflow features: \emph{bundling},
+We now describe how to implement our transactional synchronization
+abstraction using standard OpenFlow messages.
+We use three Openflow features: \emph{bundling},
 \emph{conditional application}, and \emph{cookies}.
-We will describe later how to implement some of these features
-in older Openflow versions where they are not supported. \textbf{TODO}
+%[[PK let us leave for the moment
+%We will describe later how to implement some of these features
+%in older Openflow versions where they are not
+%supported. \textbf{TODO}
+%]]
 
-A container contains items: they represent the flow table entries (i.e., a match-action rules, plus e.g., counters belonging to the rule).
-Items can be added and removed atomically and in bundles (e.g., to update the switch with
-a new policy). Recall that we want to add items only if they do not exist
-yet; otherwise the transaction should abort without taking effect.
-And similarly for removes: transactions should not take effect if the corresponding
-update contains rules which currently do not exist at the switch.
+Algorithms~\ref{alg:cond-add}-\ref{alg:delere} describe 
+transactional  $\textit{cond-add}(x)$, $\textit{cond-delete}(x)$,
+$\textit{add}(x)$, and $\textit{delete}(x)$. 
+Every operation within a transaction generates a number of  OpenFlow
+messages. Each of the messages is marked as a \emph{bundle} message,
+where the transaction's identifier is used as the bundle id.
+
+\textbf{TODO: explain how the lines in the pseudocode generate OF msgs.} 
+We use FlowMod command as abstraction to adding/deleting a number to a
+multi-set. The match part of the added flow entry represents the added
+number. In order to verify that the added number is unique we set the
+flag CHECK\_OVERLAP. Similarly, we use GroupMod command as an
+abstraction to adding/deleting a number to a set. The group id of the added group represents the added number.
+
+%A container contains items: they represent the flow table entries (i.e., a match-action rules, plus e.g., counters belonging to the rule).
+%Items can be added and removed atomically and in bundles (e.g., to update the switch with
+%a new policy). Recall that we want to add items only if they do not exist
+%yet; otherwise the transaction should abort without taking effect.
+%And similarly for removes: transactions should not take effect if the corresponding
+%update contains rules which currently do not exist at the switch.
+
+\begin{algorithm}[t]
+    \caption{$\textit{cond-add}(x)$}
+    \label{alg:multi-add}
+    \begin{algorithmic}[1]
+    \Require id $i$, unique constrain (defaults to \texttt{false}) $b\_unique$, caller id $caller$ (defaults to \emph{self}).
+%    \Ensure
+    		\State $value \gets i\concat i$
+    		\State $mask \gets caller\concat all-1$
+    		\State $cookie \gets i\concat caller$
+    		\State $action \gets i\concat caller$
+    		\If {$b\_unique$}
+    			\State $flag \gets 0$
+    		\Else
+    		    \State $flag \gets CHECK\_OVERLAP$
+    		\EndIf
+    					\State $cmd\gets \textsc{FlowMod}(match=(value,mask), operation = ADD, cookie, flag, action) $
+			\Return $cmd$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[t]
+    \caption{$\textit{cond-delete}(x)$}
+    \label{alg:template}
+    \begin{algorithmic}[1]
+    \Require id $i$, unique constrain (defaults to \texttt{false}) $b\_unique$, caller id $caller$.
+%    \Ensure
+    		\State $value \gets C.i$
+    		\State $mask \gets all-1$
+    		\State $cookie \gets C.i.caller$
+    		\If {$b\_unique$}
+    			\State $cmd\gets \textsc{FlowMod}(match=(value,mask), operation=DELETE) $
+    		\Else
+    		    \State $cmd\gets \textsc{FlowMod}(match=(value,mask), operation=DELETE\_STRICT, cookie) $
+    		\EndIf
+
+			\Return $cmd$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[t]
+    \caption{$\textit{add}(x)$}
+    \label{alg:add}
+    \begin{algorithmic}[1]
+    \Require  id $i$.
+%    \Ensure
+    		\State $command \gets OFPGC\_ADD$
+    		\State $type \gets OFPGT\_INDIRECT$
+    		\State $group_id \gets i$
+    		\State $buckets \gets [dummy\_backet]$
+    		\State $cmd\gets GroupMod(command, type, group_id, buckets) $
+			\Return $cmd$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[t]
+    \caption{$\textit{delete}(x)$}
+    \label{alg:delete}
+    \begin{algorithmic}[1]
+    \Require id $i$, unique constrain (defaults to False) $b\_unique$, caller id $caller$.
+%    \Ensure
+    		\State $command \gets OFPGC\_DELETE$
+    		\State $group_id \gets i$
+    		\State $cmd\gets GroupMod(command, group\_id) $
+			\Return $cmd$
+    \end{algorithmic}
+\end{algorithm}
+
+\section{Applications}\label{sec:apps}
+
+We now sketch a few examples of how our generic transactional synchronization
+abstraction  can be used by a control application.
+
+\subsection{In-band policy composition}
+
+Our first example (Algorithm~\ref{alg:simple-update}) is a simple in-band policy composition protocol that
+allows a set of  controllers to concurrently update switch  \emph{policies} , i.e., sets of
+effective rules in its flow tables, under the
+condition that these rules can be composed with the currently installed
+policy (the composition predicate is encoded in a user-specific
+$\textit{apply\_update}()$ function~\cite{cpc}.
+
+Every policy is associated with a \emph{policy identifier} 
+In the algorithm, the controller first reads the currently installed
+policy, then computes a new unique policy, checks whether the new
+policy and the current policy can be composed and, if so, replaces the
+current policy with the composition. If, meanwhile, a new
+policy (with a different identifier) has been installed, the update
+fails and takes no effect.
+
+\textbf{TODO: fill the gaps}
+
+Algorithm~\ref{alg:simple-update} allows policy identifiers to grow
+without bound. In fact, as was observed in~\cite{cpc}, the number of currently
+used policies can be made proportional to the number of concurrent controllers.
+Algorithm~\ref{alg:update} describes such a policy-update procedure.
+
+\textbf{TODO: fill the gaps}
 
@@
 obtain a unique \emph{id}. FIXME: to be continued...
+
+\textbf{TODO: fix the pseudocode, make the use of afdd/delete
+  operations uniform, make all switch operations transactional}
+
+\begin{algorithm}[t]
+    \caption{Update Algorithm with only a set}
+    \label{alg:simple-update}
+    \begin{algorithmic}[1]
+    \Require update requirements $U$, a set $Cur$.
+    \Ensure installed policy is consistent with previous one
+ 		%\Do
+ 		\Repeat
+ 			\State \textbf{read} switch configuration
+ 			\State $pid\gets$ current policy id (only id in $Cur$)
+ 			\State $cur\_policy\gets$ current policy
+ 			\State $new\_policy\gets apply\_update(cur\_policy,U)$
+ 			\startTransaction
+	 			\State $Cur$\emph{.delete}($pid$)
+	 			\State $Cur$\emph{.add}($pid+1$)
+	 			\State add $new\_policy$ rules %conditioned on $metadata=my\_a$
+	 			%\State update policy selector to $my\_a$
+ 			\endTransaction
+     		%\State $res \gets $ commit transaction
+     	\Until{$res=\texttt{true}$}
+     	%\doWhile{$res=False$}
+%   \vspace{3mm}
+			\Return
+
+    \end{algorithmic}
+\end{algorithm}
 
@@
 \begin{algorithm}[t]
-    \caption{\textbf{\emph{MultiSet-Add()}}}
-    \label{alg:multi-add}
-    \begin{algorithmic}[1]
-    \Require id $i$, unique constrain (defaults to \texttt{false}) $b\_unique$, caller id $caller$ (defaults to \emph{self}).
-%    \Ensure
-    		\State $value \gets i\concat i$
-    		\State $mask \gets caller\concat all-1$
-    		\State $cookie \gets i\concat caller$
-    		\State $action \gets i\concat caller$
-    		\If {$b\_unique$}
-    			\State $flag \gets 0$
-    		\Else
-    		    \State $flag \gets CHECK\_OVERLAP$
-    		\EndIf
-    					\State $cmd\gets \textsc{FlowMod}(match=(value,mask), operation = ADD, cookie, flag, action) $
-			\Return $cmd$
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[t]
-    \caption{\textbf{\emph{MultiSet-Remove()}}}
-    \label{alg:template}
-    \begin{algorithmic}[1]
-    \Require id $i$, unique constrain (defaults to \texttt{false}) $b\_unique$, caller id $caller$.
-%    \Ensure
-    		\State $value \gets C.i$
-    		\State $mask \gets all-1$
-    		\State $cookie \gets C.i.caller$
-    		\If {$b\_unique$}
-    			\State $cmd\gets \textsc{FlowMod}(match=(value,mask), operation=DELETE) $
-    		\Else
-    		    \State $cmd\gets \textsc{FlowMod}(match=(value,mask), operation=DELETE\_STRICT, cookie) $
-    		\EndIf
-
-			\Return $cmd$
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[t]
-    \caption{\textbf{\emph{Set-Add()}}}
-    \label{alg:template}
-    \begin{algorithmic}[1]
-    \Require  id $i$.
-%    \Ensure
-    		\State $command \gets OFPGC\_ADD$
-    		\State $type \gets OFPGT\_INDIRECT$
-    		\State $group_id \gets i$
-    		\State $buckets \gets [dummy\_backet]$
-    		\State $cmd\gets GroupMod(command, type, group_id, buckets) $
-			\Return $cmd$
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[t]
-    \caption{\textbf{\emph{Set-Remove()}}}
-    \label{alg:template}
-    \begin{algorithmic}[1]
-    \Require id $i$, unique constrain (defaults to False) $b\_unique$, caller id $caller$.
-%    \Ensure
-    		\State $command \gets OFPGC\_DELETE$
-    		\State $group_id \gets i$
-    		\State $cmd\gets GroupMod(command, group\_id) $
-			\Return $cmd$
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[t]
-    \caption{Update Algorithm with only a set}
-    \label{alg:template}
-    \begin{algorithmic}[1]
-    \Require update requirements $U$, a set $Cur$.
-    \Ensure installed policy is consistent with previous one
- 		%\Do
- 		\Repeat
- 			\State \textbf{read} switch configuration
- 			\State $pid\gets$ current policy id (only id in $Cur$)
- 			\State $cur\_policy\gets$ current policy
- 			\State $new\_policy\gets apply\_update(cur\_policy,U)$
- 			\startTransaction
-	 			\State $Cur$\emph{.delete}($pid$)
-	 			\State $Cur$\emph{.add}($pid+1$)
-	 			\State add $new\_policy$ rules %conditioned on $metadata=my\_a$
-	 			%\State update policy selector to $my\_a$
- 			\endTransaction
-     		%\State $res \gets $ commit transaction
-     	\Until{$res=\texttt{true}$}
-     	%\doWhile{$res=False$}
-%   \vspace{3mm}
-			\Return
-
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[t]
     \caption{Update Algorithm with pipeline-array}
@@
 
-LIRON TEXT: We use FlowMod command as abstraction to adding/deleting a number to a multi-set.
-The match part of the added flow entry represents the added number.
-In order to verify that the added number is unique we set the flag CHECK\_OVERLAP.
-Similarly, we use GroupMod command as an abstraction to adding/deleting a number to a set.
-The group id of the added group represents the added number.
-
+LIRON TEXT: 
 In order to perform an update we use a set called CUR which is used to store the current policy id, and a multi-set called NEXT which is used to manage suggestions for next policy id.
@@
 There is no progress guaranty for each controller but there is one for the whole system - at least one of the controller will succeed in fulfilling its update requirements.
-
-\section{Applications}\label{sec:apps}
 

### 2015-03-11 inband-STN.tex
some fixes and adding groups as a forth feature

@@
 
-The Openflow standard (as of \textbf{TODO}) offers three interesting
+The Openflow standard (as of \textbf{TODO}) offers four interesting
 features which we will exploit to implement the transactional memory abstraction:
@@
 \item \textbf{Conditional Application:}
-The Openflow standard also supports the \emph{conditional} application of
-certain types of control messages. A specific $\textsf{OFPFF\_CHECK\_OVERLAP}$ flag
-set in the control message containing a new forwarding rule ensures
-that the rule is applied if and only if it does not \emph{conflict}
-with any other rule currently in the flow table.
-This in-band conflict detection mechanism, combined with bundling,
+The Openflow standard also supports the \emph{conditional} creation of
+flow entries. A specific $\textsf{OFPFF\_CHECK\_OVERLAP}$ flag
+set in the flow creation message ensures
+that the flow entry is created if and only if it does not \emph{conflict}
+with any other flow entry currently in the flow table.
+This ?????in-band????? conflict detection mechanism, combined with bundling,
 allows us to implement specific types of \emph{transactions} that may
@@
 \item \textbf{Cookies:} Cookies allow to name and retrieve individual
-flow rules. Our synchronization mechanism will implement controller-specific cookies.
+flow entries and to use masks to delete modify/delete multiple entries. Our synchronization mechanism will implement controller-specific cookies.
+
+\item \textbf{Groups:} Groups of action buckets can be defined in a special group table.
+Each group has a unique id and any attempt to create a group with existing id or to delete un existing group fails (and can abort a bundle in process).
 \end{enumerate}

### 2015-03-11 inband-STN.tex
minor changes
update of the transactional if

@@
 \newcommand{\hide}[1]{}
-\newcommand{\concat}[0]{\frown}
+\newcommand{\concat}[0]{\oplus}
 \newcommand{\append}[0]{\frown}
+\newcommand{\Nat}{\mathbb{N}}
 
@@
 top of standard OpenFlow functionality. At a high level, we export a
-transactional interface which allows a controller to perform multiple
+transactional interface which allows each controller $i\in[N]$ ($N\geq
+2$) to perform multiple
 operations on the configuration of an OpenFlow switch as an atomic
@@
 of an SDN switch, such as reading flow tables or adding new rules. 
- In addition, we also assume transactional access to  \emph{multi-set}
-abstractions that store \emph{items} and export \emph{add} and \emph{delete} operations with
-regular semantics. The multi-set abstractions are used for resolving
-conflicts between concurrent updates of the switch configuration and are themselves 
-implemented as flow-table components.
-
-A multi-set object also exports operations \emph{cond-add} and \emph{cond-delete} with the following
+In addition we, we also assume transactional access to  two objects: a
+\emph{claim-id} object, used for claiming a policy \emph{identifier} in a
+given (bounded or unbounded) set, and a read-write \emph{register}
+object, used for storing the current used policy identifier.
+%
+More precisely, a claim-id object exports operations \emph{claim} and
+\emph{unclaim} and \emph{check} with the following sequential
 semantics:
-
 %The container synchronization object introduced in this paper
@@
 \begin{itemize}
-\item \textbf{$\textit{cond-add}(x)$:} if $x$ is not in the set, adds 
-  $x$ to the set; otherwise, \emph{fails} and, respectively,
-  aborts the transaction in which the operation is invoked.
+\item by performing $\textit{claim}(k)$, $i\in[N]$,
+  $k\in\Nat$, controller $i$ \emph{claims} 
+  identifier $k$;
 
-\item \textbf{$\textit{cond-delete}(x)$:} if $x$ is in the set, deletes an
-$x$ from the set; otherwise, \emph{fails} and
-  aborts the transaction.
+\item by performing $\textit{uncheck}(i)$, $i\in[N]$, controller $i$
+  \emph{unchecks} identifier $k$;
+
+\item $\textit{check}(k)$, $k\in\Nat$, returns $\textit{true}$ if and
+  only if $k$ is not currently claimed by any controller.
+
 \end{itemize}
@@
 transactions constitute a \emph{legal} sequential
-order~\cite{Pap79-serial}. On the progress side, we ensure that a
+order~\cite{Pap79-serial}, i.e., every transaction appears as executed
+sequentially. On the progress side, we ensure that a
 transaction $T$ is only allowed to abort either because another 
@@
 
-Note that
+%Note that
 %, assuming such a transactional memory,  
-a $\emph{contains}(x)$ operation can be implemented as a transaction
-composed of $\textit{cond-add}(x)$ followed by $\textit{cond-delete}(x)$.  
-A \emph{set} abstraction is a special instance of a multi-set that is
-only modified with \textit{cond-add} and \textit{cond-delete}
-operations (this way the set only contains at most one copy of a given
-item).
+%a $\emph{contains}(x)$ operation can be implemented as a transaction
+%composed of $\textit{cond-add}(x)$ followed by $\textit{cond-delete}(x)$.  
+%A \emph{set} abstraction is a special instance of a multi-set that is
+%only modified with \textit{cond-add} and \textit{cond-delete}
+%operations (this way the set only contains at most one copy of a given
+i%tem).
 
@@
 %    \Ensure
-    		\State $value \gets C.i$
-    		\State $mask \gets all-1$
-    		\State $cookie \gets C.i.caller$
+    		\State $value \gets i\concat i$
+    		\State $mask \gets caller\concat all-1$
+    		\State $cookie \gets i\concat caller$
     		\If {$b\_unique$}
@@
     		\State $type \gets OFPGT\_INDIRECT$
-    		\State $group_id \gets i$
+    		\State $group\_id \gets i$
     		\State $buckets \gets [dummy\_backet]$
-    		\State $cmd\gets GroupMod(command, type, group_id, buckets) $
+    		\State $cmd\gets GroupMod(command, type, group\_id, buckets) $
 			\Return $cmd$
@@
     		\State $command \gets OFPGC\_DELETE$
-    		\State $group_id \gets i$
+    		\State $group\_id \gets i$
     		\State $cmd\gets GroupMod(command, group\_id) $

### 2015-03-11 inband-STN.tex
changed abstractions

--- 2015-03-11 inband-STN.tex
@@
 \newcommand{\hide}[1]{}
-\newcommand{\concat}[0]{\frown}
+\newcommand{\concat}[0]{\oplus}
 \newcommand{\append}[0]{\frown}
@@
 \begin{algorithm}[t]
-    \caption{$\textit{cond-add}(x)$}
+    \caption{$\textit{claim}(x)$}
     \label{alg:multi-add}
     \begin{algorithmic}[1]
-    \Require id $i$, unique constrain (defaults to \texttt{false}) $b\_unique$, caller id $caller$ (defaults to \emph{self}).
+    \Require \emph{self} as the calling controller id.
 %    \Ensure
-    		\State $value \gets i\concat i$
-    		\State $mask \gets caller\concat all-1$
-    		\State $cookie \gets i\concat caller$
-    		\State $action \gets i\concat caller$
-    		\If {$b\_unique$}
+    		\State $value \gets x\concat x$
+    		\State $mask \gets self\concat all-1$
+	    	\State $match \gets (value,mask)$
+    		%\State $cookie \gets x\concat self$
+    		\State $action \gets x\concat self$
     		\State $flag \gets 0$
-    		\Else
-    		    \State $flag \gets CHECK\_OVERLAP$
-    		\EndIf
-    					\State $cmd\gets \textsc{FlowMod}(match=(value,mask), operation = ADD, cookie, flag, action) $
+    		\State $cmd\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
 			\Return $cmd$
@@
 \begin{algorithm}[t]
-    \caption{$\textit{cond-delete}(x)$}
+    \caption{$\textit{check}(x)$}
+    \label{alg:multi-add}
+    \begin{algorithmic}[1]
+    \Require \emph{self} as the calling controller id.
+%    \Ensure
+    		\State $value \gets x\concat x$
+    		\State $mask \gets self\concat all-1$
+    		\State $match \gets (value,mask)$
+    		%\State $cookie \gets x\concat self$
+    		\State $action \gets x\concat self$
+    		\State $flag \gets CHECK\_OVERLAP$
+    		\State $cmd1\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
+    		\State $cmd2\gets \textsc{FlowMod}(match, op = DELETE) $
+			\Return $cmd1$,$cmd2$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[t]
+    \caption{$\textit{unclaim}(x)$}
     \label{alg:template}
     \begin{algorithmic}[1]
-    \Require id $i$, unique constrain (defaults to \texttt{false}) $b\_unique$, caller id $caller$.
+    \Require \emph{self} as the calling controller id.
 %    \Ensure
-    		\State $value \gets C.i$
-    		\State $mask \gets all-1$
-    		\State $cookie \gets C.i.caller$
-    		\If {$b\_unique$}
-    			\State $cmd\gets \textsc{FlowMod}(match=(value,mask), operation=DELETE) $
-    		\Else
-    		    \State $cmd\gets \textsc{FlowMod}(match=(value,mask), operation=DELETE\_STRICT, cookie) $
-    		\EndIf
+    		\State $value \gets x\concat x$
+    		\State $mask \gets self\concat all-1$
+    		%\State $cookie \gets x\concat self$
+    		\State $match \gets (value,mask)$
+    		\State $cmd\gets \textsc{FlowMod}(match, op=DELETE, cookie) $
 
@@
 \begin{algorithm}[t]
-    \caption{$\textit{add}(x)$}
+    \caption{$\textit{compare\&set}(old,new)$}
     \label{alg:add}
     \begin{algorithmic}[1]
-    \Require  id $i$.
+    %\Require  id $i$.
 %    \Ensure
-    		\State $command \gets OFPGC\_ADD$
+
+    		\State $op \gets OFPGC\_DELETE$
+    		\State $group\_id \gets old$
+    		\State $cmd1\gets GroupMod(op, group\_id) $
+
+    		\State $op \gets OFPGC\_ADD$
     		\State $type \gets OFPGT\_INDIRECT$
-    		\State $group_id \gets i$
+    		\State $group\_id \gets new$
     		\State $buckets \gets [dummy\_backet]$
-    		\State $cmd\gets GroupMod(command, type, group_id, buckets) $
-			\Return $cmd$
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[t]
-    \caption{$\textit{delete}(x)$}
-    \label{alg:delete}
-    \begin{algorithmic}[1]
-    \Require id $i$, unique constrain (defaults to False) $b\_unique$, caller id $caller$.
-%    \Ensure
-    		\State $command \gets OFPGC\_DELETE$
-    		\State $group_id \gets i$
-    		\State $cmd\gets GroupMod(command, group\_id) $
-			\Return $cmd$
+    		\State $cmd2\gets GroupMod(op, type, group\_id, buckets) $
+			\Return $cmd1,cmd2$
     \end{algorithmic}
@@
 \begin{algorithm}[t]
-    \caption{Update Algorithm with only a set}
+    \caption{Update Algorithm with only a compare\&set}
     \label{alg:simple-update}
     \begin{algorithmic}[1]
-    \Require update requirements $U$, a set $Cur$.
+    \Require update requirements $U$
     \Ensure installed policy is consistent with previous one
@@
  			\State \textbf{read} switch configuration
- 			\State $pid\gets$ current policy id (only id in $Cur$)
+ 			\State $pid\gets$ current policy id 
  			\State $cur\_policy\gets$ current policy
@@
  			\startTransaction
-	 			\State $Cur$\emph{.delete}($pid$)
-	 			\State $Cur$\emph{.add}($pid+1$)
+	 			\State $compare\&set(pid,pid+1)$
 	 			\State add $new\_policy$ rules %conditioned on $metadata=my\_a$
@@
     \begin{algorithmic}[1]
-    \Require update requirements $U$, multi-set (with optional add validation) $Next$, set (with full validation) $Cur$.
+    \Require update requirements $U$
     \Ensure installed policy is consistent with previous one
@@
  			\State $pid\gets$ current policy id (only id in $Cur$)
- 			\State $Next$\emph{.add}($pid$)
+ 			\State $claim(pid)$
  			\State \textbf{read} switch configuration
  			\If {$pid\notin Cur$}
+	 			\State $unclaim(pid)$
  				\State {\bf continue} (\textbf{restart} loop)
@@
  			\startTransaction
- 				\State $Next$\emph{.add}($my\_id$, True)
- 				\State $Next$\emph{.delete}($my\_id$, True)
-	 			\State CUR\emph{.delete}($pid$)
-	 			\State CUR\emph{.add}($my\_id$)
-				\State $Next$\emph{.delete}($pid$)
+ 				\State $check(my\_id)$
+	 			\State $compare\&set(pid,my\_id)$
 	 			\State \textbf{add} $new\_policy$ rules %conditioned on $metadata=my\_a$
@@
  			\endTransaction
+	 		\State $unclaim(pid)$
      		%\State $res \gets $ commit transaction

### 2015-03-11 inband-STN.tex

--- 2015-03-11 inband-STN.tex
@@
 \begin{algorithm}[t]
-    \caption{$\textit{cond-delete}(x)$}
+    \caption{$\textit{check}(x)$}
+    \label{alg:multi-add}
+    \begin{algorithmic}[1]
+    \Require \emph{self} as the calling controller id.
+%    \Ensure
+    		\State $value \gets x\concat x$
+    		\State $mask \gets self\concat all-1$
+    		\State $match \gets (value,mask)$
+    		%\State $cookie \gets x\concat self$
+    		\State $action \gets x\concat self$
+    		\State $flag \gets CHECK\_OVERLAP$
+    		\State $cmd1\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
+    		\State $cmd2\gets \textsc{FlowMod}(match, op = DELETE) $
+			\Return $cmd1$,$cmd2$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[t]
+    \caption{$\textit{unclaim}(x)$}
     \label{alg:template}
     \begin{algorithmic}[1]
-    \Require id $i$, unique constrain (defaults to \texttt{false}) $b\_unique$, caller id $caller$.
+    \Require \emph{self} as the calling controller id.
 %    \Ensure
-    		\State $value \gets i\concat i$
-    		\State $mask \gets caller\concat all-1$
-    		\State $cookie \gets i\concat caller$
-    		\If {$b\_unique$}
-    			\State $cmd\gets \textsc{FlowMod}(match=(value,mask), operation=DELETE) $
-    		\Else
-    		    \State $cmd\gets \textsc{FlowMod}(match=(value,mask), operation=DELETE\_STRICT, cookie) $
-    		\EndIf
+    		\State $value \gets x\concat x$
+    		\State $mask \gets self\concat all-1$
+    		%\State $cookie \gets x\concat self$
+    		\State $match \gets (value,mask)$
+    		\State $cmd\gets \textsc{FlowMod}(match, op=DELETE, cookie) $
 
@@
 \begin{algorithm}[t]
-    \caption{$\textit{add}(x)$}
+    \caption{$\textit{compare\&set}(old,new)$}
     \label{alg:add}
     \begin{algorithmic}[1]
-    \Require  id $i$.
+    %\Require  id $i$.
 %    \Ensure
-    		\State $command \gets OFPGC\_ADD$
+
+    		\State $op \gets OFPGC\_DELETE$
+    		\State $group\_id \gets old$
+    		\State $cmd1\gets GroupMod(op, group\_id) $
+
+    		\State $op \gets OFPGC\_ADD$
     		\State $type \gets OFPGT\_INDIRECT$
-    		\State $group\_id \gets i$
+    		\State $group\_id \gets new$
     		\State $buckets \gets [dummy\_backet]$
-    		\State $cmd\gets GroupMod(command, type, group\_id, buckets) $
-			\Return $cmd$
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[t]
-    \caption{$\textit{delete}(x)$}
-    \label{alg:delete}
-    \begin{algorithmic}[1]
-    \Require id $i$, unique constrain (defaults to False) $b\_unique$, caller id $caller$.
-%    \Ensure
-    		\State $command \gets OFPGC\_DELETE$
-    		\State $group\_id \gets i$
-    		\State $cmd\gets GroupMod(command, group\_id) $
-			\Return $cmd$
+    		\State $cmd2\gets GroupMod(op, type, group\_id, buckets) $
+			\Return $cmd1,cmd2$
     \end{algorithmic}

### 2015-03-11 inband-STN.tex
fine tune
intro updated (again!)

@@
 
-%\title{On the Synchronization Power of Openflow}
+%[[PK sorry for stubborness :)
+\title{On the Synchronization Power of Openflow Data Plane} 
+%]]
 
@@
 
-\title{Synchronizing Distributed Openflow Controllers\\{\Large Toward More Powerful In-Band Synchronization Objects}}
+%\title{Synchronizing Distributed Openflow Controllers\\{\Large Toward More Powerful In-Band Synchronization Objects}}
 
@@
 %]]
-In this paper, we introduce a novel synchronization object which
-we call \emph{container}.
-The container can be implemented \emph{in-band}, using the standard Openflow
-protocol
-which support atomic multi-writes but only limited reads (namely
-testing the presence of rules). We describe how to
-leverage the container object to
-implement powerful consensus objects, transactions with all-or-nothing
- semantics, or atomically resolve
-general conflicts between concurrent policy updates
+In this paper, we show that the standard Openflow
+protocol is powerful enough to implement a powerful synchronization
+abstractions exporting \emph{transactional} accesses to the components
+of the data-plane configuration, with \emph{all-or-nothing} semantics. 
+The abstraction can be implemented \emph{in band}, without any
+additional synchronization between the controllers.
+We describe how to leverage the transactional abstraction to
+atomically resolve conflicts between concurrent policy updates
 \emph{on-the-fly}, without affecting the ongoing traffic.
@@
 that packets of different header spaces should follow across the
-network.~\cite{network-update,roger-hotnets,correct,stn}
-Installing conflicting forwarding rules, e.g., rules of the same priority defined over non-disjoint
+network~\cite{network-update,roger-hotnets,correct,stn}.
+Installing \emph{conflicting} forwarding rules, e.g., rules of the same priority defined over non-disjoint
 flow spaces may lead to pathological network behavior (loops,
-blackholes, etc.).
+blackholes, etc.)~\cite{cpc}.
 Similarly, installing diverging load-balancing policies may,
@@
 be used to implement powerful, high-level
-synchronization
-objects \emph{in-band}.
-In particular, we introduce a novel synchronization object
-called \emph{container}. A container can be seen as
-a variant of the classic \texttt{test\&set} object
-(known from distributed computing over shared memory),
-which is tailored to Openflow: while Openflow supports
-atomic multi-write functionality (using bundling),
-read functionality is limited to testing the presence of
-rules as well as flow space overlaps.
+synchronization objects \emph{in-band}.
+In particular, we define a transactional synchronization abstraction
+that exportes \emph{atomic} (\emph{all-or-nothing} access to multiple
+components of an SDN switch configuration, similarly to the way
+transactional memory~\cite{stm-st95,tm-book} provides atomic access to
+multiple memory locations.
+To implement this synchronization abstraction, we rely on several
+standard OpenFlow features, including \emph{message bundling},
+\emph{check-overlap} flag, and \emph{grouping}~\cite{of-spec}.
 
-The container object is reminiscent of the hypothetical multi-write
-shared-memory model which is known to enable solutions for important
-synchronization problems, such as consensus, impossible to solve
-otherwise~\cite{multi-objects}.
-Indeed, we show how to exploit the container object to implement
-consensus among controllers.
-More generally, the container object turns out to be
-crucial for ensuring consistency of network
-management on the control plane while preserving consistency of the
+%The container object is reminiscent of the hypothetical multi-write
+%shared-memory model which is known to enable solutions for important
+%synchronization problems, such as consensus, impossible to solve
+%otherwise~\cite{multi-objects}.
+%Indeed, we show how to exploit the container object to implement
+%consensus among controllers.
+We show how the transactional abstraction helps consistency of network
+management on the control plane, while preserving consistency of the
 traffic processed on the data plane, and we show how to implement
-\emph{packet-atomic} rule updates and more complex
-\emph{atomic transactions}.
-We also show how to replace the read-modify-write
-primitives voluntarily assumed in the recent implementation of a
+\emph{packet-atomic} rule updates~\cite{network-update}: even though
+an update .
+% and more complex
+%\emph{atomic transactions}.
+We also sketch how to implement the \emph{read-modify-write} primitives
+and \emph{control-plane} synchronization, assumed voluntarily to exist
+by the recent proposal of a
 consistent composition system for forwarding policies~\cite{cpc}.
@@
 In Section~\ref{sec:background}, we review the basics of the  Openflow
-protocol and discuss the features which are relevant for our work.
-We introduce our vision and the container consensus object in Section~\ref{sec:vision},
-and present its implementation in Section~\ref{sec:container}.
+protocol and discuss its features which are relevant for our work.
+In Section~\ref{sec:vision}, we introduce our vision and the container consensus object 
+and present its implementation in OpenFlow.
 We then discuss applications (Section~\ref{sec:apps}) as
 well as extensions (Section~\ref{sec:discussion}).
-After reviewing related literature in Section~\ref{sec:relwork}, we conclude
-our contribution in Section~\ref{sec:conclusion}.
+We review related literature in Section~\ref{sec:relwork}, and conclude
+in Section~\ref{sec:conclusion}.
 
@@
 %\section{The Container Object}\label{sec:vision}
-\section{Transactional synchronization}\label{sec:trans}
+\section{Transactional synchronization}\label{sec:vision}
 
@@
  			\State \textbf{read} switch configuration
- 			\State $my\_id\gets$ unused \emph{id} (in both $Cur$ and $Next$)
- 			\State $pid\gets$ current policy id (only id in $Cur$)
+ 			\State $my\_id\gets$ an unused \emph{id} 
+ 			\State $pid\gets$ current policy id 
  			\State $claim(pid)$

### 2015-03-11 inband-STN.tex
minor update of sec 3.1

@@
 \item $\textit{check}(k)$, $k\in\Nat$, returns $\textit{true}$ if and
-  only if $k$ is not currently claimed by any controller.
+  only if $k$ is not currently claimed by any controller; if
+  $\textit{false}$ is returned, we say that the check operation \emph{fails}.
 
 \end{itemize}
+
+We assume that a register can be accessed with a
+\emph{compare-and-set} (CAS) operation that takes two integer parameters
+$\textit{old}$ and $\textit{new}$, reads the register and, if the
+value of it is $\textit{old}$, replaces it with $\textit{new}$ and
+returns $\textit{true}$; otherwise, $\textit{false}$ is returned and
+we say that the CAS operation \emph{fails}.
 
@@
 transaction updated a flow table entry read by $T$  
-or if one of $T$'s multi-set operations fails.
+or if one of $T$'s operations fails.
 

### 2015-03-11 inband-STN.tex
updates and checking for new alg

@@
 \algblock[<block>]{<start>}{<end>}
-\algblockdefx[Transaction]{startTransaction}{endTransaction} %
-[0]{ \textbf{start transaction}} %
-[1][res]{ $#1\gets$  \textbf{commit transaction}}
+\algblockdefx[Bundle]{startBundle}{endBundle} %
+[0]{ \textbf{start bundle}} %
+[1][res]{ $#1\gets$  \textbf{commit bundle}}
 
@@
 We use three Openflow features: \emph{bundling},
-\emph{conditional application}, and \emph{cookies}.
+\emph{conditional application}, and \emph{groups}.
 %[[PK let us leave for the moment
@@
 
-Algorithms~\ref{alg:cond-add}-\ref{alg:delere} describe 
-transactional  $\textit{cond-add}(x)$, $\textit{cond-delete}(x)$,
-$\textit{add}(x)$, and $\textit{delete}(x)$. 
+Algorithms~\ref{alg:claim}-\ref{alg:cas} describe 
+transactional  $\textit{claim}(x)$, $\textit{unclaim}(x)$,
+$\textit{check}(x)$, and $\textit{CAS}(x)$. 
 Every operation within a transaction generates a number of  OpenFlow
@@
     \caption{$\textit{claim}(x)$}
-    \label{alg:multi-add}
+    \label{alg:claim}
     \begin{algorithmic}[1]
@@
     \caption{$\textit{check}(x)$}
-    \label{alg:multi-add}
+    \label{alg:check}
     \begin{algorithmic}[1]
@@
     		\State $action \gets x\concat self$
-    		\State $flag \gets CHECK\_OVERLAP$
+    		\State $flag \gets OFPFF\_CHECK\_OVERLAP$
     		\State $cmd1\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
@@
     \caption{$\textit{unclaim}(x)$}
-    \label{alg:template}
+    \label{alg:unclaim}
     \begin{algorithmic}[1]
@@
     \caption{$\textit{compare\&set}(old,new)$}
-    \label{alg:add}
+    \label{alg:cas}
     \begin{algorithmic}[1]
@@
  			\State $new\_policy\gets apply\_update(cur\_policy,U)$
- 			\startTransaction
+ 			\startBundle
 	 			\State $compare\&set(pid,pid+1)$
@@
 	 			%\State update policy selector to $my\_a$
- 			\endTransaction
+ 			\endBundle
      		%\State $res \gets $ commit transaction
@@
  			\State $new\_policy\gets apply\_update(cur\_policy,U)$
- 			\startTransaction
+ 			\startBundle
  				\State $check(my\_id)$
@@
 	 			%\State update policy selector to $my\_a$
- 			\endTransaction
+ 			\endBundle
 	 		\State $unclaim(pid)$
@@
 \begin{algorithm}[t]
-    \caption{Update Algorithm with pipeline-array}
+    \caption{Policy Composition Algorithm without bundle}
     \label{alg:template}
     \begin{algorithmic}[1]
-    \Require update requirements $U$, array $ARR$.
-    \Ensure installed policy is consistent with previous one
+    \Require policy $P$ (a set of FlowMod entry creation commands), three consecutive flow tables ids (per controller) $T_s,T_0,T_1$, active table index $active\in\{0,1\}$.
+    %\Ensure installed policy is consistent with previous one
+    \State $unactive \gets active + 1 \mod{2}$
+    \For {$cmd \in P$}
+	    \State $cmd.action\gets Reencode(cmd.action)$
+	    \State $cmd.table\_id\gets T_{unactive}$
+	    \State send $cmd$
+    \EndFor
+    \State modify the action of $T_s$ entry to goto $T_{unactive}$
+    \State clear table $T_{active}$
+    \State $active \gets unactive$
 	\Return
@@
 \begin{algorithm}[t]
-    \caption{Update Composition Algorithm without bundle}
+    \caption{Update Algorithm with pipeline}
     \label{alg:template}

### 2015-03-12 inband-STN.tex
passes on related work and conclusions

@@
 
-Consider, for example, the practical problem of
+Consider, for example, the problem of
 consistent installation of new forwarding policies, stipulating routes
@@
 Operating and updating SDNs consistently is already non-trivial
-from the perspective of a single controller. Reitblatt et al.~~\cite{network-update}
+from the perspective of a single controller. Reitblatt et al.~\cite{network-update}
 introduced the notion of
@@
 
+\textbf{PK: anything about your ``brain'' paper, feel related?}
+
 To the best of our knowledge, the only paper explicitly addressing the consistent
-network update problem from a distributed controller perspective is STN~\cite{stn}:
+network update problem from a distributed-controllers perspective is STN~\cite{stn}:
 STN provides a transactional interface offering all-or-nothing semantics and serializability
-(the ``holy grail'' of safety properties),
-allowing
+(the ``holy grail'' of safety properties), allowing
 controllers to install policies in a conflict-free manner; if some controllers fail,
 other controllers can take over. STN~\cite{stn} is based on
-atomic
-read-modify-write primitives, but while this primitive is postulated, no implementation is
+atomic read-modify-write primitives provided by the switches, but
+while the availability of such primitives is postulated, no implementation is
 described. In this paper, we provide this missing link, and also show that other useful and
-much more powerful synchronization primitives can be implemented.
+more powerful synchronization primitives can be implemented. 
+Also, some STN implementations in~\cite{stn} assume that the control
+plane maintains a certain amount of synchrony that allows controllers
+to solve consensus or implement a replicated state machine.     
+In this paper, we describe an OpenFlow-based consensus protocol that
+can be used for this purpose.
 
@@
 also enabling policy composition~\cite{pyretic}. Our work complements these works
-on programming abstractoins
+on programming abstractions
 by initiating the discussion of concurrency objects and abstractions.
 
+The transactional interface we describe in this paper is inspired by the
+work on speculative concurrency control using software transactional memory
+(STM)~\cite{stm-st95,tm-book}. 
+%[[PK I think this is enough]]
+
+\hide{
 In terms of distributed computing: TODO PETR: maybe modify a bit the text below
@@
 assumed by STM implementations.
+}
 
@@
 
-This paper showed that powerful synchronization object
-from distributed computing can be implemented in Openflow.
-Our work complements existing research on the design of
-distributed control planes, and also provides some missing links,
-e.g., for the transactional approach taken by STN.
-We also hope that our work can nourish the discussion of
-which synchronization objects we can and want to implement
-in an SDN. While much progress has been made over the
+While much progress has been made over the
 last years in the development of more intuitive programming
@@
 has received much less attention.
+%
+In this paper, we describe transactional abstractions that 
+can be implemented using standard Openflow features.
+%
+Our work complements existing research on the design of
+distributed control planes, and also provides some missing links,
+e.g., for the transactional approach taken by STN~\cite{stn}.
+We also hope that our work can nourish the discussion of
+which synchronization objects we can and want to implement
+in an SDN.
 

### 2015-03-12 inband-STN.tex
slight update

@@
   only if $k$ is not currently claimed by any controller; if
-  $\textit{false}$ is returned, we say that the check operation \emph{fails}.
+  $\textit{false}$ is returned, we say that the check operation
+  \emph{fails} (which causes an abort of the transaction invoking it).
 
@@
 returns $\textit{true}$; otherwise, $\textit{false}$ is returned and
-we say that the CAS operation \emph{fails}.
+we say that the CAS operation \emph{fails} (which causes an abort if
+the transaction invoking it).
 
@@
 sequentially. On the progress side, we ensure that a
-transaction $T$ is only allowed to abort either because another 
+transaction $T$ is only allowed to abort if either another 
 transaction updated a flow table entry read by $T$  
-or if one of $T$'s operations fails.
+or if one of $T$'s operations (\textit{check} or CAS) fails.
 

### 2015-03-12 inband-STN.tex
adding explanation to abstractions and algorithm

--- 2015-03-12 inband-STN.tex
@@
 \textbf{TODO: explain how the lines in the pseudocode generate OF msgs.} 
-We use FlowMod command as abstraction to adding/deleting a number to a
-multi-set. The match part of the added flow entry represents the added
-number. In order to verify that the added number is unique we set the
-flag CHECK\_OVERLAP. Similarly, we use GroupMod command as an
-abstraction to adding/deleting a number to a set. The group id of the added group represents the added number.
+We use creation and deletion of flow entries (with FlowMod commands) as abstraction to claiming/unclaiming  resource id numbers and also to checking whether a given id is claimed.
+The match part of the added flow entry represents the added
+number and the mask part of the match varies from one controller to another in order to allow multiple controllers to claim the same number without overriding the same flow entry. 
+As can be seen in \ref{alg:claim}, the match mask is a concatenation of the controller id with all ones, while the value is the id concatenated twice.
+We can see that for any two numbers $x_1$ and $x_2$ and controller ids $c_1\neq c_2$ the match patterns $m_1=(x_1\concat x_1, c_1\concat all-1)$ and $ m_1=(x_2\concat x_2, c_2\concat all-1)$ are not equal even if $x_1=x_2$ therefore they can both claim and unclaim without affecting each other.
+
+In order to check if a number is claimed we try to create the corresponding entry while setting the
+flag CHECK\_OVERLAP, thereby inflicting a failure in case an entry with same match value but different action exists.
+If the check succeeds we delete the entry.
+Considering $m_1$ and $m_2$ defined above and their respective actions $a_1=x_1\concat c1$ and $a_2=x_2\concat c2$,  they overlap with different actions iff $x_1=x_2$.
+Therefore if one is used to check the other a contradiction is detected iff $x_1=x_2$ as expected.
+Note that checking for claim must contains a deletion of the added flow entry otherwise a successive check operation might falsely fail.
+Moreover the two commands (adding and deleting of the flow entries) must be sent as part of the same bundle to support concurrent checks by different controllers.
+
+Similarly, we use a deletion of group $x$ with creation of group $y$ (with GroupMod commands) as an
+abstraction to performing a $Compare\&Set(x,y)$ operation to a special $16$bit register (we abstract only one register). 
+%The group id of the added group represents the added number. 
+Note that trying to delete a group that doesn't exist (or trying to create a group with id that already exists) fails, thereby preventing adding the new value $y$ if $x$ hasn't existed before.
+Note that the two action should be performed inside the same bundle in order to ensure the atomic nature ot the abstraction.
 
@@
 
-\subsection{In-band policy composition}
+\subsection{Policy updates composition}
 
-Our first example (Algorithm~\ref{alg:simple-update}) is a simple in-band policy composition protocol that
+Our first example (Algorithm~\ref{alg:simple-update}) is a simple policy composition protocol that
 allows a set of  controllers to concurrently update switch  \emph{policies} , i.e., sets of
@@
 In the algorithm, the controller first reads the currently installed
-policy, then computes a new unique policy, checks whether the new
+policy, then choose a new unique policy, checks whether the new
 policy and the current policy can be composed and, if so, replaces the
@@
 
-\textbf{TODO: fill the gaps}
+Algorithm~\ref{alg:simple-update} use increasing numbers as policies ids and use them in order to make sure that the new policy, with id $x$, replaces the (what considered as) current policy, with id $x-1$.
+Considering the policies ids, we send the policy installation commands as part of a bundle that also tries to perform  $Compare\&Set(x-1,x)$, thereby insuring that the commands are performed iff the current policy is still there and preventing other controllers to replace the new policy with another without observing it.
 
@@
 
-\textbf{TODO: fill the gaps}
+ Algorithm~\ref{alg:update} lies at the heart of our mechanism.
+ It takes a set of rules $U$ (not yet a policy)and proceeds as follows: first, we seek to
+ obtain a unique \emph{id}. FIXME: to be continued...
+ \textbf{LS: I am not sure if I need to tell every step of the story or maybe it best to explain the main dif from the previous one similarly to what I just wrote next}.
 
-Algorithm~\ref{alg:update} lies at the heart of our mechanism.
-It takes a set of rules $U$ (not yet a policy),
-a multi-set $Next$ and a set (with full validation) $Cur$.
-The algorithm proceeds as follows: first, we seek to
-obtain a unique \emph{id}. FIXME: to be continued...
+Algorithm~\ref{alg:update} claims the current policy id in order to prevent someone of using it with different policy in the future (during the current update) and in the bundle used to update the policy it also checks that the new policy id that it about to use (chosen from a finite set of ids) wasn't claimed by someone else.
+
+\hide{
+We compute our new suggested policy by applying the update requests on top of current policy, supporting any kind of requests and policies.
+Then we make a transaction (using the bundle feature) to atomically check that our policy id is not blocked by anyone else, to change the current policy id to ours (an action that would fail if the current policy id is no longer what we are counting of) and to actually configure our new policy.
+
+If one of the actions in the transaction fails we try again.
+There is no progress guaranty for each controller but there is one for the whole system - at least one of the controller will succeed in fulfilling its update requirements.
+}
 
@@
 \end{algorithm}
-
-LIRON TEXT: 
-In order to perform an update we use a set called CUR which is used to store the current policy id, and a multi-set called NEXT which is used to manage suggestions for next policy id.
-We start by reading the current switch configuration and selecting an unused id for our proposed policy.
-Next we try to prevent someone of using the current id with different policy in the future (during our current update).
-
-We compute our new suggested policy by applying the update requests on top of current policy, supporting any kind of requests and policies.
-Then we make a transaction (using the bundle feature) to atomically check that our policy id is not blocked by anyone else, to change the current policy id to ours (an action that would fail if the current policy id is no longer what we are counting of) and to actually configure our new policy.
-
-If one of the actions in the transaction fails we try again.
-There is no progress guaranty for each controller but there is one for the whole system - at least one of the controller will succeed in fulfilling its update requirements.
 

### 2015-03-12 inband-STN.tex
some fixes and implementation of compare\&set with FlowMod commands

@@
 number and the mask part of the match varies from one controller to another in order to allow multiple controllers to claim the same number without overriding the same flow entry. 
-As can be seen in \ref{alg:claim}, the match mask is a concatenation of the controller id with all ones, while the value is the id concatenated twice.
+As can be seen in Algorithms \ref{alg:claim}-\ref{alg:unclaim}, the match mask is a concatenation of the controller id with all ones, while the value is the id concatenated twice.
 We can see that for any two numbers $x_1$ and $x_2$ and controller ids $c_1\neq c_2$ the match patterns $m_1=(x_1\concat x_1, c_1\concat all-1)$ and $ m_1=(x_2\concat x_2, c_2\concat all-1)$ are not equal even if $x_1=x_2$ therefore they can both claim and unclaim without affecting each other.
@@
 Considering $m_1$ and $m_2$ defined above and their respective actions $a_1=x_1\concat c1$ and $a_2=x_2\concat c2$,  they overlap with different actions iff $x_1=x_2$.
-Therefore if one is used to check the other a contradiction is detected iff $x_1=x_2$ as expected.
-Note that checking for claim must contains a deletion of the added flow entry otherwise a successive check operation might falsely fail.
+Therefore if one is used in check after the other used in claims, a conflict is detected iff $x_1=x_2$ as expected.
+Note that checking for claims must contain a deletion of the added flow entry otherwise a successive check operation might mistakenly fail.
 Moreover the two commands (adding and deleting of the flow entries) must be sent as part of the same bundle to support concurrent checks by different controllers.
 
-Similarly, we use a deletion of group $x$ with creation of group $y$ (with GroupMod commands) as an
-abstraction to performing a $Compare\&Set(x,y)$ operation to a special $16$bit register (we abstract only one register). 
+Similarly, we use a deletion of group $x$ and a creation of group $y$ (with GroupMod commands) as an
+abstraction to performing a $Compare\&Set(x,y)$ operation to a special $16$bit register (we abstract only one register). The $16$ bit limit is resulted by th $16$bit group id size. 
 %The group id of the added group represents the added number. 
@@
 
+\section{Implementing $\textit{compare\&set}$ with FlowMod commands}
+\begin{algorithm}[H]
+    \caption{$\textit{compare\&set2}(old,new)$}
+    \label{alg:cas2}
+    \begin{algorithmic}[1]
+    %\Require  id $i$.
+%    \Ensure
+			\State $cmds \gets new\ list()$
+    		\State $value \gets  old$
+    		\State $mask \gets  all-1$
+    		\State $match \gets (value,mask)$
+    		%\State $cookie \gets x\concat self$
+    		\State $action \gets 2$
+    		\State $flag \gets OFPFF\_CHECK\_OVERLAP$
+    		\State $cmd1\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
+    		\State $cmd2\gets \textsc{FlowMod}(match, op = DELETE) $
+    		\State $cmd3\gets \textsc{FlowMod}(cookie = old, op = DELETE) $
+    		\State $cmds.append(cmd1)$
+    		\State $cmds.append(cmd2)$
+    		\State $cmds.append(cmd3)$
+    		\State $range\_matches \gets EncodeRange([0,new-1])\cup EncodeRange([new+1,MAX\_ID])$
+    		\For {$r\_match \in range\_matches $}
+    		    		\State $cookie \gets new$
+    		    		\State $action \gets 1$
+    		    		\State $flag \gets 0$
+    		    		\State $cmd\gets \textsc{FlowMod}(r\_match, ADD, flag, cookie, action) $
+    		    		\State $cmds.append(cmd)$
+    		\EndFor
+			\Return $cmds$
+    \end{algorithmic}
+\end{algorithm}
+
 \section{Backup Stuff}

### 2015-03-13 inband-STN.tex
a few typos corrected

@@
 In this section, we sketch the synchronization abstraction we build on
-top of standard OpenFlow functionality. At a high level, we export a
+top of the standard OpenFlow functionality. At a high level, we export a
 transactional interface which allows each controller $i\in[N]$ ($N\geq
@@
 operations on the configuration of an OpenFlow switch as an atomic
-\emph{transaction}, inspired by the concept of transactional
-memory~\cite{stm-st95,tm-book}. A transaction may \emph{commit}, in
+\emph{transaction}, inspired by the transactional-memory paradigm~\cite{stm-st95,tm-book}. A transaction may \emph{commit}, in
 which case it appears as executed sequentially, or \emph{abort}, in
@@
 of an SDN switch, such as reading flow tables or adding new rules. 
-In addition we, we also assume transactional access to  two objects: a
+In addition we, we also assume transactional access to  two kinds of objects: a
 \emph{claim-id} object, used for claiming a policy \emph{identifier} in a
@@
 %
-More precisely, a claim-id object exports operations \emph{claim} and
+A claim-id object exports operations \emph{claim},
 \emph{unclaim} and \emph{check} with the following sequential
@@
 
-\item by performing $\textit{uncheck}(i)$, $i\in[N]$, controller $i$
-  \emph{unchecks} identifier $k$;
+\item by performing $\textit{unclaim}(i)$, $i\in[N]$, controller $i$
+  \emph{unclaim} identifier $k$;
 
@@
 
-We assume that a register can be accessed with a
+A register can be accessed with a
 \emph{compare-and-set} (CAS) operation that takes two integer parameters
@@
 %operations (this way the set only contains at most one copy of a given
-i%tem).
+%item).
 

### 2015-03-13 inband-STN.tex
3.2 updated

@@
 [1][res]{ $#1\gets$  \textbf{commit bundle}}
+
+\algblockdefx[Bundle]{startTxn}{endTxn} %
+[0]{\textbf{atomic}\{} %
+[1][res]{ \} $\rightarrow #1$}
 
@@
 with any other flow entry currently in the flow table.
-This ?????in-band????? conflict detection mechanism, combined with bundling,
+This  conflict detection mechanism, combined with bundling,
 allows us to implement specific types of \emph{transactions} that may
@@
 We assume that transactions may comprise operations on the flow table
-of an SDN switch, such as reading flow tables or adding new rules. 
+of an SDN switch, such as reading flow table entries or adding new entries. 
 In addition we, we also assume transactional access to  two kinds of objects: a
@@
 
-Formally, we guarantee that in every execution, all committed
+More formally, our transactional abstraction exports the
+$\textbf{atomic}\{\;\ldots\;\}$ construct
+containing a sequence of transactional
+operations and may \emph{commit} and return responses to all its operations
+or \emph{abort} (so that none of its operation takes effect).  
+We guarantee that in every execution, all committed
 transactions constitute a \emph{legal} sequential
@@
 or if one of $T$'s operations (\textit{check} or CAS) fails.
-
+We employ the \emph{deferred-update} semantics:  
+no transaction can affect other transactions before it commits.
+In particular, if a controller crashes before committing its
+transactions, no future transaction can be affected.  
 %Note that
@@
 
-We now describe how to implement our transactional synchronization
-abstraction using standard OpenFlow messages.
-We use three Openflow features: \emph{bundling},
-\emph{conditional application}, and \emph{groups}.
+We now describe how to implement our transactional 
+abstraction using standard OpenFlow features, including
+\emph{bundling}, \emph{conditional application}, and \emph{groups}.
 %[[PK let us leave for the moment
@@
 %]]
-
-Algorithms~\ref{alg:claim}-\ref{alg:cas} describe 
-transactional  $\textit{claim}(x)$, $\textit{unclaim}(x)$,
-$\textit{check}(x)$, and $\textit{CAS}(x)$. 
+Algorithms~\ref{alg:claim}-\ref{alg:cas} describe implementations of  
+transactional operations  $\textit{claim}(x)$, $\textit{check}(x)$, $\textit{unclaim}(x)$,
+and $\textit{CAS}(x)$, respectively. 
+Algorithm~\ref{alg:commit} describes the procedure executed by
+a transaction at \emph{commit time}.
+
+We assume that every transaction has a unique identifier. 
 Every operation within a transaction generates a number of  OpenFlow
-messages. Each of the messages is marked as a \emph{bundle} message,
+messages using \textsc{FlowMod} commands.
+We therefore assume that each of these OpenFlow messages is marked as  a \emph{bundle} message,
 where the transaction's identifier is used as the bundle id.
@@
 In order to check if a number is claimed we try to create the corresponding entry while setting the
-flag CHECK\_OVERLAP, thereby inflicting a failure in case an entry with same match value but different action exists.
+flag \textsf{CHECK\_OVERLAP}, thereby inflicting a failure in case an entry with same match value but different action exists.
 If the check succeeds we delete the entry.
@@
 Note that the two action should be performed inside the same bundle in order to ensure the atomic nature ot the abstraction.
+
+To commit the transaction,  the controller sends a \textsf{OFPT\_BUNDLE\_CONTROL} message with type
+\textsf{OFPBCT\_COMMIT\_REQUEST} and waits until a matching control
+message from the switch is received. When the switch receives the
+commit request it tries to atomically process all the buffered messages
+equipped with the corresponding bundle identifier and, if the
+processing is successful for all of them, returns  a \textsf{OFPT\_BUNDLE\_CONTROL} message with type
+\textsf{OFPBCT\_CLOSE\_REPLY}; otherwise it sends a specific error
+message  pointing out the aborted transactional operation.
 
@@
     		\State $action \gets x\concat self$
-    		\State $flag \gets OFPFF\_CHECK\_OVERLAP$
+    		\State $flag \gets \textsf{OFPFF\_CHECK\_OVERLAP}$
     		\State $cmd1\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
@@
 
-    		\State $op \gets OFPGC\_DELETE$
+    		\State $op \gets \textsf{OFPGC\_DELETE}$
     		\State $group\_id \gets old$
@@
 
-    		\State $op \gets OFPGC\_ADD$
-    		\State $type \gets OFPGT\_INDIRECT$
+    		\State $op \gets \textsf{OFPGC\_ADD}$
+    		\State $type \gets \textsf{OFPGT\_INDIRECT}$
     		\State $group\_id \gets new$
@@
 			\Return $cmd1,cmd2$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[t]
+    \caption{$\textit{try-commit}()$}
+    \label{alg:commit}
+    \begin{algorithmic}[1]
+    		\State  \textbf{TODO}: Express sending the
+                  \textsf{OFPFF\_BUNDLE\_CONTROL} message with type \textsf{OFPBCT\_COMMIT\_REQUEST}
+		\Return $cmd$
     \end{algorithmic}
@@
 \begin{algorithm}[t]
-    \caption{Update Algorithm with only a compare\&set}
+    \caption{Policy update with only a compare\&set}
     \label{alg:simple-update}
@@
  			\State $new\_policy\gets apply\_update(cur\_policy,U)$
- 			\startBundle
+ 			\startTxn
 	 			\State $compare\&set(pid,pid+1)$
@@
 	 			%\State update policy selector to $my\_a$
- 			\endBundle
+ 			\endTxn
      		%\State $res \gets $ commit transaction
-     	\Until{$res=\texttt{true}$}
+     	\Until{$res\neq\texttt{abort}$}
      	%\doWhile{$res=False$}
@@
 \begin{algorithm}[t]
-    \caption{Advanced Update Algorithm}
+    \caption{Advanced policy update}
     \label{alg:update}
@@
  			\State $new\_policy\gets apply\_update(cur\_policy,U)$
- 			\startBundle
+ 			\startTxn
  				\State $check(my\_id)$
@@
 	 			%\State update policy selector to $my\_a$
- 			\endBundle
+ 			\endTxn
 	 		\State $unclaim(pid)$
      		%\State $res \gets $ commit transaction
-     	\Until{$res=\texttt{true}$}
+     	\Until{$res\neq\texttt{abort}$}
      	%\doWhile{$res=False$}
@@
 \begin{algorithm}[t]
-    \caption{Policy Composition Algorithm without bundle}
-    \label{alg:template}
+    \caption{Policy composition without bundle}
+    \label{alg:wobundle}
     \begin{algorithmic}[1]
@@
 \begin{algorithm}[t]
-    \caption{Update Algorithm with pipeline}
-    \label{alg:template}
+    \caption{Update with pipeline}
+    \label{alg:pipeline}
     \begin{algorithmic}[1]

### 2015-03-13 inband-STN.tex
pushed a short update of Sec 4, STN not yet done

@@
 Similarly, we use a deletion of group $x$ and a creation of group $y$ (with GroupMod commands) as an
-abstraction to performing a $Compare\&Set(x,y)$ operation to a special $16$bit register (we abstract only one register). The $16$ bit limit is resulted by th $16$bit group id size. 
+abstraction to performing a $CAS(x,y)$ operation to a special $16$bit register (we abstract only one register). The $16$ bit limit is resulted by th $16$bit group id size. 
 %The group id of the added group represents the added number. 
@@
 \begin{algorithm}[t]
-    \caption{$\textit{compare\&set}(old,new)$}
+    \caption{$\textit{CAS}(old,new)$}
     \label{alg:cas}
@@
 
-\subsection{Policy updates composition}
+\subsection{Policy updates}
 
@@
 
-Algorithm~\ref{alg:simple-update} use increasing numbers as policies ids and use them in order to make sure that the new policy, with id $x$, replaces the (what considered as) current policy, with id $x-1$.
-Considering the policies ids, we send the policy installation commands as part of a bundle that also tries to perform  $Compare\&Set(x-1,x)$, thereby insuring that the commands are performed iff the current policy is still there and preventing other controllers to replace the new policy with another without observing it.
+Algorithm~\ref{alg:simple-update} use increasing numbers as policies
+ids and use them in order to make sure that the new policy, with id
+$x$, replaces the (what considered as) current policy, with id
+$x-1$. Considering the policies ids, we send the policy installation
+commands as part of a bundle that also tries to perform
+$CAS(x-1,x)$, thereby insuring that the commands are
+performed iff the current policy is still there and preventing other
+controllers to replace the new policy with another without observing it.
+
+For brevity, we omit explicitly putting the $\textbf{atomic}\{...\}$ construct
+around each of the transactional operation used in the algorithm.   
+Indeed, since \textit{claim} and \textit{unclaim} operations cannot
+cause the corresponding transaction, we can assume that all
+``singleton'' (consisting of one operation only)  transactions
+containing such operations always commit.
 
@@
 without bound. In fact, as was observed in~\cite{cpc}, the number of currently
-used policies can be made proportional to the number of concurrent controllers.
-Algorithm~\ref{alg:update} describes such a policy-update procedure.
-
- Algorithm~\ref{alg:update} lies at the heart of our mechanism.
+used policies can be made proportional to the number of concurrent
+controllers.
+
  It takes a set of rules $U$ (not yet a policy)and proceeds as follows: first, we seek to
@@
 \begin{algorithm}[t]
-    \caption{Policy update with only a compare\&set}
+    \caption{Policy update with only CAS}
     \label{alg:simple-update}
@@
  			\startTxn
-	 			\State $compare\&set(pid,pid+1)$
+	 			\State $CAS(pid,pid+1)$
 	 			\State add $new\_policy$ rules %conditioned on $metadata=my\_a$
@@
  				\State $check(my\_id)$
-	 			\State $compare\&set(pid,my\_id)$
+	 			\State $CAS(pid,my\_id)$
 	 			\State \textbf{add} $new\_policy$ rules %conditioned on $metadata=my\_a$
@@
 
-FIXME: STN, consensus, ...
-
-\paragraph{Consensus} FIXME: desribe how to implement consensus
-
-\paragraph{One-Shot Update} FIXME: desribe how to implement one-shot update
-
-\paragraph{STN} FIXME: desribe how to implement STN
+\paragraph{Consensus and replicated state machines} 
+It is straightforward to implement consensus among the controllers
+using just a single register that can be accessed with an atomic CAS
+operation~\cite{Her91}. In our transactional setting, the controllers
+can use CAS to replace the default value of such a register 
+with their consensus input values (e.g., the identifiers of the
+policies that should be applied first). If the corresponding
+singleton transaction commits, the controller consideres itself 
+the \emph{winner} and decides on its input, if
+it aborts (the controller fails in the CAS) it remains to read the register to get the input of the
+winner to decide.
+
+By using multiple CAS invocations, the controllers can implement any
+generic replicated state machine~\cite{Her91,Lam98}, which will be instrumental in the
+next section.
+
+%\paragraph{One-Shot Update} FIXME: desribe how to implement one-shot update
+
+\paragraph{STN}
+
+\textbf{TODO}
 
@@
 
-\section{Implementing $\textit{compare\&set}$ with FlowMod commands}
+\section{Implementing $\textit{CAS}$ with FlowMod commands}
 \begin{algorithm}[H]
-    \caption{$\textit{compare\&set2}(old,new)$}
+    \caption{$\textit{CAS}(old,new)$}
     \label{alg:cas2}

### 2015-03-13 inband-STN.tex
a few changes

--- 2015-03-13 inband-STN.tex
@@
 Every operation within a transaction generates a number of  OpenFlow
-messages. Each of the messages is marked as a \emph{bundle} message,
-where the transaction's identifier is used as the bundle id.
+messages. Each of the messages can be wrapped 
+is marked as a 
+in a \emph{bundle} message
+%,where the transaction's identifier is used as the bundle id. 
+.
+Each bundle message carries a bundle id, where messages with same bundle id are considered to belong to same bundle and are performed atomically.
+
+%The OpenFlow messages that we use (besides )
 
@@
 \section{Implementing $\textit{compare\&set}$ with FlowMod commands}
+
+Here we describe a second implementation of the Compare\&Set abstraction that is based on entirely on flow entry creations and deletions (FlowMod commands).
+
 \begin{algorithm}[H]

### 2015-03-13 inband-STN.tex

--- 2015-03-13 inband-STN.tex
@@
 Every operation within a transaction generates a number of  OpenFlow
-messages using \textsc{FlowMod} commands.
-We therefore assume that each of these OpenFlow messages is marked as  a \emph{bundle} message,
-where the transaction's identifier is used as the bundle id.
+messages. Each of the messages can be wrapped 
+is marked as a 
+in a \emph{bundle} message
+%,where the transaction's identifier is used as the bundle id. 
+.
+Each bundle message carries a bundle id, where messages with same bundle id are considered to belong to same bundle and are performed atomically.
+
+%The OpenFlow messages that we use (besides )
 
@@
 \section{Implementing $\textit{CAS}$ with FlowMod commands}
+
+Here we describe a second implementation of the Compare\&Set abstraction that is based on entirely on flow entry creations and deletions (FlowMod commands).
+
 \begin{algorithm}[H]

### 2015-03-17 old-inband-STN.tex
stefan's changes

--- 2015-03-13 inband-STN.tex
@@
 %[[PK sorry for stubborness :)
-\title{On the Synchronization Power of Openflow Data Plane} 
+\title{On the Synchronization Power of the Openflow Data Plane}
+%\\ {\Large Towards Atomic Transactions In-Band}}
 %]]
@@
 %]]
-In this paper, we show that the standard Openflow
-protocol is powerful enough to implement a powerful synchronization
-abstractions exporting \emph{transactional} accesses to the components
-of the data-plane configuration, with \emph{all-or-nothing} semantics. 
-The abstraction can be implemented \emph{in band}, without any
-additional synchronization between the controllers.
-We describe how to leverage the transactional abstraction to
-atomically resolve conflicts between concurrent policy updates
-\emph{on-the-fly}, without affecting the ongoing traffic.
-%[[PK
-%In particular, we describe the implementation of test\&set primitives
-%as well as multi-writer consensus objects which can be used
-%to implement atomic transactions with all-or-nothing semantics.
-%Interestingly, our approach allows controllers to detect and
-%resolve very general notions of conflicts, not only, e.g., conflicts
-%due to overlapping flow spaces but also conflicts based on network load.
-%]]
+%
+A well-known concept to solve synchronization and consistency problems
+in distributed systems are \emph{transactions}: a transaction defines a sequence 
+of operations which is performed in an atomic, serializable manner, providing all-or-nothing semantics. 
+This paper shows that powerful transactions can also be implemented in software-defined networks,
+using the standard Openflow protocol. 
+%Indeed, recent Openflow versions already come with useful but primitive
+%features, in particular \emph{bundling} and \emph{rule-overlap checks}. 
+In particular, we present a
+transactional interface which supports arbitrary atomic
+transactions 
+over the switch configuration space: Our transactions 
+can include complex notions of conflicts (beyond flow overlap checks),
+as well as computations and controller interactions. We discuss different use cases
+for our interface (consistent policy installation, load-balancing, consensus),
+and also describe how to 
+perform basic forms of policy compositions in-band.
 \end{abstract}
@@
 
-%\textbf{}
+The distributed computing literature offers 
+a wide range of abstractions 
+and techniques to solve distributed synchronization and consistency problems. 
+A particularly interesting concept in distributed computing is a \emph{transactional memory}: 
+a shared memory which supports atomic transactions, describing a sequence of operations 
+to be executed with an all-or-nothing semantics. Transactional memories can be used to
+ solve many 
+classic distributed computing problems, such as consensus.
+
+This paper investigates the design of a transactional interface for software-defined
+networks. In particular, we argue that the switch configuration state can be seen
+as a shared memory which is modified concurrently by multiple controllers.
+Being able to provide more complex and atomic sequences of transactions over this configuration
+space can be appealing to resolve synchronization problems.
+
 \paragraph{Our Contributions} 
-In this paper, we show that
-Openflow~\cite{of-spec}---the standard SDN protocol today---can
-be used to implement powerful, high-level
-synchronization objects \emph{in-band}.
-In particular, we define a transactional synchronization abstraction
-that exportes \emph{atomic} (\emph{all-or-nothing} access to multiple
-components of an SDN switch configuration, similarly to the way
-transactional memory~\cite{stm-st95,tm-book} provides atomic access to
-multiple memory locations.
-To implement this synchronization abstraction, we rely on several
-standard OpenFlow features, including \emph{message bundling},
-\emph{check-overlap} flag, and \emph{grouping}~\cite{of-spec}.
-
-%The container object is reminiscent of the hypothetical multi-write
-%shared-memory model which is known to enable solutions for important
-%synchronization problems, such as consensus, impossible to solve
-%otherwise~\cite{multi-objects}.
-%Indeed, we show how to exploit the container object to implement
-%consensus among controllers.
-We show how the transactional abstraction helps consistency of network
-management on the control plane, while preserving consistency of the
-traffic processed on the data plane, and we show how to implement
-\emph{packet-atomic} rule updates~\cite{network-update}: even though
-an update .
-% and more complex
-%\emph{atomic transactions}.
-We also sketch how to implement the \emph{read-modify-write} primitives
-and \emph{control-plane} synchronization, assumed voluntarily to exist
-by the recent proposal of a
-consistent composition system for forwarding policies~\cite{cpc}.
+This paper introduces a very powerful transactional interface
+ for software-defined networks. This interface is implemented
+ in-band, and
+ using Openflow~\cite{of-spec} (the standard SDN protocol today) features only.
+Indeed, we find that Openflow's \emph{bundle} or \emph{rule overlap check}
+features can serve as synchronization primitives. 
+However, without further mechanisms, their use is limited:
+with a bundle alone it is not possible to react to
+and change switch configuration state in any non-trivial way,
+and with the rule overlap check it is not possible to react to
+any non-trivial update conflicts.
+
+In contrast, the transactional interface presented in this paper supports very general atomic
+transactions over the switch configuration space, similar to classic transactional memory~\cite{stm-st95,tm-book}. 
+For instance, a transaction allows to modify switch configurations in an interactive 
+and computational fashion, as well as to define complex notions
+of conflicts. 
+Our implementation is based on distributed identifiers and makes use of two 
+new synchronization objects for managing these identifiers:
+a \emph{claim-id object}, used for claiming an available identifier in a given
+(bounded or unbounded) set, and a \emph{read-write register object}, used for storing the 
+currently used identifier.
+Thus, our work also provides the missing link for the read-modify-write object
+postulated in~\cite{cpc}.
+Our work also contributes to the ongoing discussion of what can be implemented
+in-band in Openflow~\cite{compute,reclaim}, by taking a first look
+at the relevant problem of distributed
+synchronization.
 
@@
 protocol and discuss its features which are relevant for our work.
-In Section~\ref{sec:vision}, we introduce our vision and the container consensus object 
-and present its implementation in OpenFlow.
-We then discuss applications (Section~\ref{sec:apps}) as
-well as extensions (Section~\ref{sec:discussion}).
+%In Section~\ref{sec:vision}, we introduce our vision and the container consensus object
+%and present its implementation in OpenFlow.
+%We then discuss applications (Section~\ref{sec:apps}) as
+%well as extensions (Section~\ref{sec:discussion}).
+FIXME
 We review related literature in Section~\ref{sec:relwork}, and conclude
@@
 
-\section{Openflow Background}\label{sec:background}
+\section{Background and Motivation}\label{sec:background}
 
@@
 
-In a nutshell,
+\textbf{Background.} In a nutshell,
 Openflow
@@
 
-The Openflow standard (as of \textbf{TODO}) offers four interesting
+\textbf{Useful features for synchronization.} The Openflow standard (as of \textbf{TODO}) offers four interesting
 features which we will exploit to implement the transactional memory abstraction:
@@
 
-\item \textbf{Conditional Application:}
+\item \textbf{Check Overlap:}
 The Openflow standard also supports the \emph{conditional} creation of
@@
 
+\textbf{Limitations.}  While these features are useful primitives, 
+building complex transactions from these features
+is a non-trivial task, as we will show in the following. 
+A bundle can be understood as a transaction:
+ it defines a sequence of read and write operations 
+ on the switch state which are executed atomically.
+ However, such bundle transactions come with many 
+ limitations. A bundle transaction does not allow
+ to change the switch state in any non-trivial way,
+ as the read and write actions in the bundle do 
+ not support any conditional modifications or computations which 
+ depend on the read results. By leveraging
+ the check overlap feature, it is at least possible
+ to abort transactions depending on certain conditions
+ in the overlapping
+ flow spaces. However, in many cases, for example
+ a load-balancing application,
+ it can be interesting to define more general
+ dependencies, which depend on, from a logical perspective,
+ seemingly
+ completely independent flow spaces, 
+ which however share the underlying physical network.
+In other words, the bundle feature only allows us to 
+define ``one-shot transactions'',
+transactions which do not support any further controller interactions. 
+Given
+the limited computational options the Openflow interface of a switch offers, 
+it is difficult to implement more complex behavior.
+
 %\section{The Container Object}\label{sec:vision}
-\section{Transactional synchronization}\label{sec:vision}
+\section{Transactional Synchronization}\label{sec:vision}
+
+This section describes how to introduce more powerful transactions. We first present a new synchronization object to
 
@@
 
-\section{Applications}\label{sec:apps}
+\section{Discussion of Use-Cases}\label{sec:apps}
 
@@
 
-\subsection{Policy updates}
+FIXME: load-balancing, consensus, etc.
 
@@
 
-\section{Discussion and Extensions}\label{sec:discussion}
-
-FIXME: we could however even do stronger stuff: collision is only exact, not using masks: do not use the ternary nature
-
-\paragraph{Beyond Container Objects}
-FIXME: we can even do more general things!
-
-\paragraph{Older Openflow Versions} FIXME: describe matrix approach
-
 \section{In-Band Composition}\label{sec:compo}
+
+We conclude our technical contribution by showing that sometimes,
+powerful synchronization mechanisms can be implemented
+
+ This section show how to implement
 
@@
 of network update algorithms that provide even stronger consistency guarantees.
-
-\textbf{PK: anything about your ``brain'' paper, feel related?}
 
@@
 %[[PK I think this is enough]]
+
+Our work also contributes to the ongoing discussion of what can be implemented
+in-band in Openflow~\cite{compute,reclaim}, by assuming the relevant perspective of distributed
+synchronization.
 
@@
 \section{Implementing $\textit{CAS}$ with FlowMod commands}
+
+This section describes an implementation without group tables.
+
 \begin{algorithm}[H]

### 2015-03-19 inband-STN.tex
bundle commit code

@@
 
+\newcommand{\stefan}[1]{\textit{\textcolor{red}{[stefan]: #1}}} % Comments
+\newcommand{\liron}[1]{\textit{\textcolor{green}{[julius]: #1}}} % Comments
+\newcommand{\petr}[1]{\textit{\textcolor{blue}{[carlos]: #1}}} % Comments
+
 \algblock[<block>]{<start>}{<end>}
@@
 %]]
-In this paper, we show that the standard Openflow
-protocol is powerful enough to implement a powerful synchronization
-abstractions exporting \emph{transactional} accesses to the components
-of the data-plane configuration, with \emph{all-or-nothing} semantics. 
-The abstraction can be implemented \emph{in band}, without any
-additional synchronization between the controllers.
-We describe how to leverage the transactional abstraction to
-atomically resolve conflicts between concurrent policy updates
-\emph{on-the-fly}, without affecting the ongoing traffic.
-%[[PK
-%In particular, we describe the implementation of test\&set primitives
-%as well as multi-writer consensus objects which can be used
-%to implement atomic transactions with all-or-nothing semantics.
-%Interestingly, our approach allows controllers to detect and
-%resolve very general notions of conflicts, not only, e.g., conflicts
-%due to overlapping flow spaces but also conflicts based on network load.
-%]]
+%
+A well-known concept to solve synchronization and consistency problems
+in distributed systems are \emph{transactions}: a transaction defines a sequence
+of operations which are performed in an atomic, serializable manner, providing all-or-nothing semantics.
+This paper shows that powerful transactions can also be implemented in software-defined networks,
+using the standard Openflow protocol.
+%Indeed, recent Openflow versions already come with useful but primitive
+%features, in particular \emph{bundling} and \emph{rule-overlap checks}.
+In particular, we present a
+transactional interface which supports arbitrary atomic
+transactions
+over the switch configuration space: Our transactions
+can include complex notions of conflicts (beyond flow overlap checks),
+as well as computations and controller interactions. We discuss different use cases
+for our interface (consistent policy installation, load-balancing, consensus),
+and also describe how to
+perform basic forms of policy compositions in-band.
 \end{abstract}
@@
 
-%\textbf{}
+The distributed computing literature offers
+a wide range of abstractions
+and techniques to solve synchronization and consistency problems.
+A particularly interesting concept in distributed computing is a \emph{transactional memory}:
+a shared memory which supports atomic transactions, describing a sequence of operations
+to be executed with an all-or-nothing semantics. Transactional memories can be used to
+ solve many
+classic distributed computing problems, such as consensus.
+
+This paper investigates the design of a transactional interface for software-defined
+networks. In particular, we argue that the switch configuration state can be seen
+as a shared memory which is modified concurrently by multiple controllers.
+Being able to provide more complex and atomic sequences of transactions over this configuration
+space can be appealing to resolve synchronization problems.
+
 \paragraph{Our Contributions}
-In this paper, we show that
-Openflow~\cite{of-spec}---the standard SDN protocol today---can
-be used to implement powerful, high-level
-synchronization objects \emph{in-band}.
-In particular, we define a transactional synchronization abstraction
-that exportes \emph{atomic} (\emph{all-or-nothing} access to multiple
-components of an SDN switch configuration, similarly to the way
-transactional memory~\cite{stm-st95,tm-book} provides atomic access to
-multiple memory locations.
-To implement this synchronization abstraction, we rely on several
-standard OpenFlow features, including \emph{message bundling},
-\emph{check-overlap} flag, and \emph{grouping}~\cite{of-spec}.
-
-%The container object is reminiscent of the hypothetical multi-write
-%shared-memory model which is known to enable solutions for important
-%synchronization problems, such as consensus, impossible to solve
-%otherwise~\cite{multi-objects}.
-%Indeed, we show how to exploit the container object to implement
-%consensus among controllers.
-We show how the transactional abstraction helps consistency of network
-management on the control plane, while preserving consistency of the
-traffic processed on the data plane, and we show how to implement
-\emph{packet-atomic} rule updates~\cite{network-update}: even though
-an update .
-% and more complex
-%\emph{atomic transactions}.
-We also sketch how to implement the \emph{read-modify-write} primitives
-and \emph{control-plane} synchronization, assumed voluntarily to exist
-by the recent proposal of a
-consistent composition system for forwarding policies~\cite{cpc}.
+This paper introduces a very powerful transactional interface
+ for software-defined networks. This interface is implemented
+ in-band, and
+ using Openflow~\cite{of-spec} (the standard SDN protocol today) features only.
+Indeed, we find that Openflow's \emph{bundle} or \emph{rule overlap check}
+features can serve as synchronization primitives.
+However, without further mechanisms, their use is limited:
+with a bundle alone it is not possible to react to
+and change switch configuration state in any non-trivial way,
+and with the rule overlap check it is not possible to react to
+update conflicts defined over non-overlapping flow spaces.
+
+In contrast, the transactional interface presented in this paper supports very general atomic
+transactions over the switch configuration space, similar to classic transactional memory~\cite{stm-st95,tm-book}.
+For instance, a transaction allows to modify switch configurations in an interactive
+and computational fashion, as well as to define complex notions
+of conflicts.
+Our implementation is based on distributed identifiers and makes use of two
+new synchronization objects for managing these identifiers:
+a \emph{claim-id object}, used for claiming an available identifier in a given
+(bounded or unbounded) set, and a \emph{read-write register object}, used for storing the
+currently used identifier.
+Thus, our paper also provides the missing link for the read-modify-write object
+postulated in~\cite{cpc}.
+
+Finally, we also show that interesting synchronization mechanisms may be included
+in-band even without the bundle feature, and as a case study, show
+how to automatically \emph{compose} updates originating \emph{from different controllers}.
+
+Finally, we hope that our work can contribute to the ongoing discussion of what can be implemented
+in-band in Openflow~\cite{compute,reclaim}, by taking a first look
+at the relevant problem of distributed
+synchronization.
 
@@
 protocol and discuss its features which are relevant for our work.
-In Section~\ref{sec:vision}, we introduce our vision and the container consensus object 
-and present its implementation in OpenFlow.
-We then discuss applications (Section~\ref{sec:apps}) as
-well as extensions (Section~\ref{sec:discussion}).
+In Section~\ref{sec:motivation}, we discuss the limitations of interpreting
+the Openflow bundle as a transaction, and motivate our approach.
+We then present our approach in detail in Section~\ref{sec:main}, and discuss use cases in Section~\ref{sec:apps}.
+Section~\ref{sec:compo} describes how to perform in-band composition.
 We review related literature in Section~\ref{sec:relwork}, and conclude
-in Section~\ref{sec:conclusion}.
+in Section~\ref{sec:conclusion}. Some technical details are postponed
+to the Appendix.
 
@@
 
-\section{Openflow Background}\label{sec:background}
+\section{Background}\label{sec:background}
 
@@
 \item \textbf{Bundling:}
-Control messages are aggregated into one message, so that either all of the control messages are
+A bundle can be seen as a ``mini-transaction'':
+a bundles aggregates multiple control messages
+into one \textsc{FlowMod} message, so that either all of the control messages are
 processed, or none of them is. In particular, if, for some reason, the
@@
 message is sent to the controller that issued them.
-Interestingly, this bundled processing can be, when properly configured, done in a
+This bundled processing can be, when properly configured, done in a
 \emph{packet-atomic} way, so that each packet is processed according
@@
 
+\begin{figure*}[t]
+\centering
+\includegraphics[width=0.8\columnwidth]{loadbal.pdf}~~
+\includegraphics[width=0.8\columnwidth]{pic.pdf}\\
+\caption{\emph{Left:} Load-balancing gone wrong. \emph{Right:} Illustration of transactional abstraction which  allows
+controllers to modify switch configurations in complex ways,
+and
+depending on other flow spaces.}\label{fig:illu}
+\end{figure*}
+
+\section{Limitations and Motivation}\label{sec:motivation}
+
+Before presenting our transactional interface and implementation,
+let us first discuss the limitations of the Openflow features discussed above.
+At first sight, the bundle feature seems to provide everything
+we need: it defines a sequence of read and write operations
+ on the switch state which are executed atomically.
+ However, such bundle transactions come with many
+ limitations. A bundle transaction does not allow
+ to change the switch state in any non-trivial way,
+ as the read and write actions in the bundle do
+ not support any conditional modifications or computations which
+ depend on the read results. By leveraging
+ the check overlap feature, it is at least possible
+ to abort transactions depending on certain conditions
+ in the overlapping
+ flow spaces. However, in many cases, for example
+ a load-balancing application,
+ it can be interesting to define more general
+ dependencies, which depend on, from a logical perspective,
+ seemingly
+ completely independent flow spaces,
+ which however share the underlying physical network.
+In other words, the bundle feature only allows us to
+define ``one-shot transactions'',
+transactions which do not support any further controller interactions.
+Given
+the limited computational options the Openflow interface of a switch offers,
+it is difficult to implement more complex behavior.
+
+\textbf{Example 1.} To give an example, let us consider a controller application in charge
+of \emph{load-balancing}. The objective of the load-balancer
+is to distribute flows equally across links or servers.
+If multiple controllers concurrently and independently
+update forwarding rules, they may unvoluntarily
+unbalance the flow allocation. Note that this
+is true independently of whether the flow spaces
+affected by the respective controllers overlap or not.
+More concretely, assume that two controllers concurrently
+decide to forward traffic $1*$ and $0*$ to the same server.
+Flow space overlap checks cannot be used to detect
+such conflicts.
+FIXME: make this example better
+
+\textbf{Example 2.} To give another example,
+consider the problem of modifying previously
+installed forwarding policies.
+While a bundle allows to conditionally add new rules or remove
+old rules (and, using the flow space overlap
+check, to abort the transaction if potentially conflicting
+rules were already there), it cannot be used
+to modify rules in non-trivial ways.
+In other words, the bundle and overlap check only helps
+for the conditional creation of rules, not their modification.
+
+\stefan{It would be cool if we could make these examples more concrete,
+at least one, and get back to it later again!}
+
 %\section{The Container Object}\label{sec:vision}
-\section{Transactional synchronization}\label{sec:vision}
-
-In this section, we sketch the synchronization abstraction we build on
-top of the standard OpenFlow functionality. At a high level, we export a
+\section{Transactional Synchronization}\label{sec:main}
+
+This section describes how to introduce more powerful transactions.
+Our implementation is based on two new synchronization objects,
+allowing to coordinate controllers using identifiers.
+
+At a high level, we export a
 transactional interface which allows each controller $i\in[N]$ ($N\geq
@@
 
-\subsection{Interface}\label{sec:t-if}
-We assume that transactions may comprise operations on the flow table
-of an SDN switch, such as reading flow table entries or adding new entries. 
-In addition we, we also assume transactional access to  two kinds of objects: a
+\subsection{Synchronization Objects}\label{sec:t-if}
+
+We assume transactional access to  two kinds of objects: a
 \emph{claim-id} object, used for claiming a policy \emph{identifier} in a
@@
 
-More formally, our transactional abstraction exports the
+\subsection{Interface}\label{sec:t-if}
+
+We assume that transactions may comprise operations on the flow table
+of an SDN switch. Our transactional abstraction exports the
 $\textbf{atomic}\{\;\ldots\;\}$ construct
@@
 
-\subsection{Transactions in Openflow}\label{sec:t-impl}
+\subsection{Implementation in Openflow}\label{sec:t-impl}
 
@@
     \begin{algorithmic}[1]
-    		\State  \textbf{TODO}: Express sending the
-                  \textsf{OFPFF\_BUNDLE\_CONTROL} message with type \textsf{OFPBCT\_COMMIT\_REQUEST}
+    \Require  current bundle id $bid$.
+		    \State $cmd \gets struct ofp_bundle_ctrl_msg$
+    		\State  $cmd.bundle\_id \gets bid$
+    		\State  $cmd.type \gets $ \textsf{OFPBCT\_COMMIT\_REQUEST} 
+    		\State  $cmd.flags \gets $ \textsf{OFPBF\_ATOMIC}
+    		\State send $cmd$
 		\Return $cmd$
@@
 
-\section{Applications}\label{sec:apps}
-
-We now sketch a few examples of how our generic transactional synchronization
-abstraction  can be used by a control application.
-
-\subsection{Policy updates}
+\section{Discussion of Use-Cases}\label{sec:apps}
+
+More powerful transactions can be useful in many scenarios.
+Let us revisit our examples from earlier: with
+transactions we can easily solve the load-balancing
+and rule modification problem.
+In the following we give two more examples of our approach.
+
+\subsection{Policy Updates}
 
@@
 
-\paragraph{Consensus and replicated state machines} 
+\subsection{Consensus and Replicated State Machines}
+
 It is straightforward to implement consensus among the controllers
@@
 
-\section{Discussion and Extensions}\label{sec:discussion}
-
-FIXME: we could however even do stronger stuff: collision is only exact, not using masks: do not use the ternary nature
-
-\paragraph{Beyond Container Objects}
-FIXME: we can even do more general things!
-
-\paragraph{Older Openflow Versions} FIXME: describe matrix approach
-
 \section{In-Band Composition}\label{sec:compo}
 
-FIXME: shall we?
+We conclude our technical contribution by showing that sometimes,
+powerful synchronization mechanisms can be implemented
+in-band \emph{even without bundles}.
+Concretely, we show that sometimes it is even possible
+to automatically compose updates originating from different controllers,
+in-band and just using Openflow.
+
+To illustrate this point, we consider a simple example
+with two controllers concurrently issuing updates to a single
+given switch.
+
+FIXME: Explain, maybe even with an example?
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.8\columnwidth]{composition.pdf}\\
+\caption{Composition.}\label{fig:illu}
+\end{figure}
 
@@
 
-\section{Implementing $\textit{CAS}$ with FlowMod commands}
+\section{Implementing CAS with FlowMod commands}\label{sec:without-groups}
+
+This section describes an implementation without group tables.
 
@@
 
+\begin{comment}
 \section{Backup Stuff}
@@
 
+\end{comment}
+
 \end{appendix}

### 2015-03-19 inband-STN.tex
starting to explan new cas

@@
 \newcommand{\Nat}{\mathbb{N}}
+\newcommand{\cas}{CAS}
 
@@
 \section{Implementing CAS with FlowMod commands}\label{sec:without-groups}
-
 This section describes an implementation without group tables.
@@
 Here we describe a second implementation of the Compare\&Set abstraction that is based on entirely on flow entry creations and deletions (FlowMod commands).
+
+The CHECK\_OVERLAP flag enables us to add an entry if no contradicting entry exists and to fail if such do exists.
+However, the compare stage of \cas requires us to fail if a certain value doesn't exists, i.e.
+$\cas(x,y)$ failes iff $x \notin T$, where $T$ is the set of match entries defined in a dedicated flow table.
+...
+...
 

### 2015-03-22 inband-STN.tex
duplicated section 4.

@@
 \newcommand{\stefan}[1]{\textit{\textcolor{red}{[stefan]: #1}}} % Comments
-\newcommand{\liron}[1]{\textit{\textcolor{green}{[julius]: #1}}} % Comments
+\newcommand{\liron}[1]{\textit{\textcolor{green}{[liron]: #1}}} % Comments
 \newcommand{\petr}[1]{\textit{\textcolor{blue}{[carlos]: #1}}} % Comments
@@
 %\section{The Container Object}\label{sec:vision}
+\section{-NEW- Transactional Synchronization}\label{sec:main}
+
+This section describes how to introduce more powerful transactions.
+Our implementation is based on two new synchronization objects,
+allowing to coordinate controllers using identifiers.
+
+At a high level, we export a
+transactional interface which allows each controller $i\in[N]$ ($N\geq
+2$) to perform multiple
+operations on the configuration of an OpenFlow switch as an atomic
+\emph{transaction}, inspired by the transactional-memory paradigm~\cite{stm-st95,tm-book}. A transaction may \emph{commit}, in
+which case it appears as executed sequentially, or \emph{abort}, in
+which none of its operations takes effect.
+
+\subsection{Synchronization Objects}\label{sec:t-if}
+
+We assume transactional access to  two kinds of objects: a
+\emph{resource-claimer} object, provides claiming a resource id (e.g. policy \emph{identifier}) in a
+given (bounded or unbounded) set, and a read-write \emph{memory} abstraction
+object, which can be used for storing the current used policy identifier.
+%
+\subsubsection{A resource claimer}
+A resource-claimer object implements a weak type of locking, exports operations \emph{claim},
+\emph{unclaim} and \emph{abortIfClaimed} with the following sequential
+semantics:
+%The container synchronization object introduced in this paper
+%is essentially a transactional memory abstraction which can be implemented
+%in standard Openflow switches.
+%Before presenting the technical details in the next section,%
+%we give a brief overview.
+
+%The postulated container object can store an arbitrary number of \emph{items}.
+%As we will see, these items correspond to flow entries on the Openflow switch.
+%The same item may even exist multiple times in the container (i.e., the container
+%is a multi-set).
+%The container
+%provides two main functions:
+\begin{itemize}
+\item by performing $\textit{claim}(k)$, $i\in[N]$,
+  $k\in\Nat$, controller $i$ \emph{claims} 
+  identifier $k$;
+
+\item by performing $\textit{unclaim}(k)$, $i\in[N]$, $k\in\Nat$, controller $i$
+  \emph{unclaim} identifier $k$;
+
+\item $\textit{abortIfClaimed}(k)$, $k\in\Nat$, 
+abort current transaction if $k$ is currently claimed by any controller.
+%returns $\textit{true}$ if and
+%  only if $k$ is not currently claimed by any controller; if
+%  $\textit{false}$ is returned, we say that the check operation
+%  \emph{fails} (which causes an abort of the transaction invoking it).
+
+\end{itemize}
+
+\subsubsection{A memory abstraction}
+A memory can be accessed with the following sequential
+semantics:
+
+\begin{itemize}
+\item by performing $\textit{Write}(addr,k)$,
+  $addr,k\in\Nat$, the value at memory offset $addr$ is set to $k$;
+
+\item $\textit{abortIfUnequal}(addr,k)$, $addr,k\in\Nat$, 
+aborts current transaction if the value at memory offset $addr$ is not equal to $k$.
+%returns $\textit{true}$ if and
+%  only if $k$ is not currently claimed by any controller; if
+%  $\textit{false}$ is returned, we say that the check operation
+%  \emph{fails} (which causes an abort of the transaction invoking it).
+
+\end{itemize}
+
+Note that reading the memory can be done by just reading the switch configuration however such action doesn't affect the execution of the transaction (as $\textit{abortIfUnequal}$ does).
+
+An immediate extension of the memory interface is the \emph{compare-and-set} (CAS) operation that takes two integer parameters
+$\textit{old}$ and $\textit{new}$, reads the register and, if the
+value of it is $\textit{old}$, replaces it with $\textit{new}$ and
+returns $\textit{true}$; otherwise, $\textit{false}$ is returned and
+we say that the CAS operation \emph{fails} (which causes an abort if
+the transaction invoking it).
+
+\subsection{Interface}\label{sec:t-if}
+
+\liron{this subsection doesn't flow well for me}
+
+We assume that transactions may comprise operations on the flow table
+of an SDN switch. Our transactional abstraction exports the
+$\textbf{atomic}\{\;\ldots\;\}$ construct
+containing a sequence of transactional
+operations and may \emph{commit} and return responses to all its operations
+or \emph{abort} (so that none of its operation takes effect).  
+We guarantee that in every execution, all committed
+transactions constitute a \emph{legal} sequential
+order~\cite{Pap79-serial}, i.e., every transaction appears as executed
+sequentially. On the progress side, we ensure that a
+transaction $T$ is only allowed to abort if either another 
+transaction updated a flow table entry read by $T$  
+or if one of $T$'s operations (\textit{check} or CAS) fails.
+We employ the \emph{deferred-update} semantics:  
+no transaction can affect other transactions before it commits.
+In particular, if a controller crashes before committing its
+transactions, no future transaction can be affected.  
+%Note that
+%, assuming such a transactional memory,  
+%a $\emph{contains}(x)$ operation can be implemented as a transaction
+%composed of $\textit{cond-add}(x)$ followed by $\textit{cond-delete}(x)$.  
+%A \emph{set} abstraction is a special instance of a multi-set that is
+%only modified with \textit{cond-add} and \textit{cond-delete}
+%operations (this way the set only contains at most one copy of a given
+%item).
+
+\subsection{Implementation in Openflow}\label{sec:t-impl}
+
+We now describe how to implement our transactional 
+abstractions using standard OpenFlow features.
+
+%features , including
+%\emph{bundling}, \emph{conditional application}, and \emph{groups}.
+%[[PK let us leave for the moment
+%We will describe later how to implement some of these features
+%in older Openflow versions where they are not
+%supported. \textbf{TODO}
+%]]
+Algorithms~\ref{alg:claim}-\ref{alg:cas} describe implementations of  
+transactional operations  $\textit{claim}(x)$, $\textit{check}(x)$, $\textit{unclaim}(x)$,
+and $\textit{CAS}(x)$, respectively. 
+Algorithm~\ref{alg:commit} describes the procedure executed by
+a transaction at \emph{commit time}.
+
+We describe each operation as a message factory that returns OpenFlow commands that should be sent to the switch as part of the same transaction.
+We assume that every transaction has a unique identifier  used to start a \emph{bundle}. 
+All messages generated during a transaction invocation, are wrapped by bundle messages carrying the bundle's id and followed by bundle commit message.
+Messages with same bundle id are considered to belong to same bundle and are performed atomically.
+
+%The OpenFlow messages that we use (besides )
+
+Our implementation is based on OpenFlow flow entries modification command, the FlowMod message that explained in Section \ref{sec:background}.
+We use the CHECK\_OVERLAP optional flag to make the transaction execution dependent on switch state.
+
+We consider the match part of an entry to be a ternary string of bounded length (represented by binary value and mask) which can be applied to any set of supported packet header fields.
+In our implementation the $64$bit metedata field alone is suffice.
+Moreover we consider the action part of an entry to be an integer, which in practice can be implemented by a Set-Metadata instruction where the written value is that integer.
+
+  \liron{got till here}
+
+We use creation and deletion of flow entries as abstraction to claiming/unclaiming  resource id numbers and also to checking whether a given id is claimed.
+The match part of the added flow entry represents the added
+number and the mask part of the match varies from one controller to another in order to allow multiple controllers to claim the same number without overriding the same flow entry. 
+As can be seen in Algorithms \ref{alg:claim}-\ref{alg:unclaim}, the match mask is a concatenation of the controller id with all ones, while the value is the id concatenated twice.
+We can see that for any two numbers $x_1$ and $x_2$ and controller ids $c_1\neq c_2$ the match patterns $m_1=(x_1\concat x_1, c_1\concat all-1)$ and $ m_1=(x_2\concat x_2, c_2\concat all-1)$ are not equal even if $x_1=x_2$ therefore they can both claim and unclaim without affecting each other.
+
+In order to check if a number is claimed we try to create the corresponding entry while setting the
+flag \textsf{CHECK\_OVERLAP}, thereby inflicting a failure in case an entry with same match value but different action exists.
+If the check succeeds we delete the entry.
+Considering $m_1$ and $m_2$ defined above and their respective actions $a_1=x_1\concat c1$ and $a_2=x_2\concat c2$,  they overlap with different actions iff $x_1=x_2$.
+Therefore if one is used in check after the other used in claims, a conflict is detected iff $x_1=x_2$ as expected.
+Note that checking for claims must contain a deletion of the added flow entry otherwise a successive check operation might mistakenly fail.
+Moreover the two commands (adding and deleting of the flow entries) must be sent as part of the same bundle to support concurrent checks by different controllers.
+
+Similarly, we use a deletion of group $x$ and a creation of group $y$ (with GroupMod commands) as an
+abstraction to performing a $CAS(x,y)$ operation to a special $16$bit register (we abstract only one register). The $16$ bit limit is resulted by th $16$bit group id size. 
+%The group id of the added group represents the added number. 
+Note that trying to delete a group that doesn't exist (or trying to create a group with id that already exists) fails, thereby preventing adding the new value $y$ if $x$ hasn't existed before.
+Note that the two action should be performed inside the same bundle in order to ensure the atomic nature ot the abstraction.
+
+To commit the transaction,  the controller sends a \textsf{OFPT\_BUNDLE\_CONTROL} message with type
+\textsf{OFPBCT\_COMMIT\_REQUEST} and waits until a matching control
+message from the switch is received. When the switch receives the
+commit request it tries to atomically process all the buffered messages
+equipped with the corresponding bundle identifier and, if the
+processing is successful for all of them, returns  a \textsf{OFPT\_BUNDLE\_CONTROL} message with type
+\textsf{OFPBCT\_CLOSE\_REPLY}; otherwise it sends a specific error
+message  pointing out the aborted transactional operation.
+
+%A container contains items: they represent the flow table entries (i.e., a match-action rules, plus e.g., counters belonging to the rule).
+%Items can be added and removed atomically and in bundles (e.g., to update the switch with
+%a new policy). Recall that we want to add items only if they do not exist
+%yet; otherwise the transaction should abort without taking effect.
+%And similarly for removes: transactions should not take effect if the corresponding
+%update contains rules which currently do not exist at the switch.
+
+\begin{algorithm}[t]
+    \caption{$\textit{claim}(x)$}
+    \label{alg:claim}
+    \begin{algorithmic}[1]
+    \Require \emph{self} as the calling controller id.
+%    \Ensure
+    		\State $value \gets x\concat x$
+    		\State $mask \gets self\concat all-1$
+	    	\State $match \gets (value,mask)$
+    		%\State $cookie \gets x\concat self$
+    		\State $action \gets x\concat self$
+    		\State $flag \gets 0$
+    		\State $cmd\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
+			\Return $cmd$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[t]
+    \caption{$\textit{check}(x)$}
+    \label{alg:check}
+    \begin{algorithmic}[1]
+    \Require \emph{self} as the calling controller id.
+%    \Ensure
+    		\State $value \gets x\concat x$
+    		\State $mask \gets self\concat all-1$
+    		\State $match \gets (value,mask)$
+    		%\State $cookie \gets x\concat self$
+    		\State $action \gets x\concat self$
+    		\State $flag \gets \textsf{OFPFF\_CHECK\_OVERLAP}$
+    		\State $cmd1\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
+    		\State $cmd2\gets \textsc{FlowMod}(match, op = DELETE) $
+			\Return $cmd1$,$cmd2$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[t]
+    \caption{$\textit{unclaim}(x)$}
+    \label{alg:unclaim}
+    \begin{algorithmic}[1]
+    \Require \emph{self} as the calling controller id.
+%    \Ensure
+    		\State $value \gets x\concat x$
+    		\State $mask \gets self\concat all-1$
+    		%\State $cookie \gets x\concat self$
+    		\State $match \gets (value,mask)$
+    		\State $cmd\gets \textsc{FlowMod}(match, op=DELETE, cookie) $
+
+			\Return $cmd$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[t]
+    \caption{$\textit{CAS}(old,new)$}
+    \label{alg:cas}
+    \begin{algorithmic}[1]
+    %\Require  id $i$.
+%    \Ensure
+
+    		\State $op \gets \textsf{OFPGC\_DELETE}$
+    		\State $group\_id \gets old$
+    		\State $cmd1\gets GroupMod(op, group\_id) $
+
+    		\State $op \gets \textsf{OFPGC\_ADD}$
+    		\State $type \gets \textsf{OFPGT\_INDIRECT}$
+    		\State $group\_id \gets new$
+    		\State $buckets \gets [dummy\_backet]$
+    		\State $cmd2\gets GroupMod(op, type, group\_id, buckets) $
+			\Return $cmd1,cmd2$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[t]
+    \caption{$\textit{try-commit}()$}
+    \label{alg:commit}
+    \begin{algorithmic}[1]
+    \Require  current bundle id $bid$.
+		    \State $cmd \gets struct ofp_bundle_ctrl_msg$
+    		\State  $cmd.bundle\_id \gets bid$
+    		\State  $cmd.type \gets $ \textsf{OFPBCT\_COMMIT\_REQUEST} 
+    		\State  $cmd.flags \gets $ \textsf{OFPBF\_ATOMIC}
+    		\State send $cmd$
+		\Return $cmd$
+    \end{algorithmic}
+\end{algorithm}
+
+%\section{The Container Object}\label{sec:vision}
 \section{Transactional Synchronization}\label{sec:main}
@@
 
-\item by performing $\textit{unclaim}(i)$, $i\in[N]$, controller $i$
+\item by performing $\textit{unclaim}(k)$, $i\in[N]$, $k\in\Nat$, controller $i$
   \emph{unclaim} identifier $k$;

### 2015-03-22 inband-STN.tex
completed section 4. need to update main pseudo code.

@@
 \definecolor{shadecolor}{rgb}{0.9,0.9,0.9}
+\definecolor{mygreen}{rgb}{0,0.5,0}
 
@@
 \newcommand{\cas}{CAS}
+\newcommand{\claimcheck}{check}
+\newcommand{\compare}{compare}
 
 \newcommand{\stefan}[1]{\textit{\textcolor{red}{[stefan]: #1}}} % Comments
-\newcommand{\liron}[1]{\textit{\textcolor{green}{[liron]: #1}}} % Comments
+\newcommand{\liron}[1]{\textit{\textcolor{mygreen}{[liron]: #1}}} % Comments
 \newcommand{\petr}[1]{\textit{\textcolor{blue}{[carlos]: #1}}} % Comments
@@
 
+\liron{I think that maybe the rest of this section should be explained by the following subsections, providing in depth description of the OpenFlow features.}
+\liron{Another direction is to have a simplified explenation as now and to refference spesific sections of the standard in side our paper.}
+\subsection{The flow entry modification command (FlowMod)}
+
+\subsubsection{The CHECK\_OVERLAP flag}
+
+\subsection{Bundling}
+
 \textbf{Useful features for synchronization.} The Openflow standard (as of \textbf{TODO}) offers four interesting
@@
 A resource-claimer object implements a weak type of locking, exports operations \emph{claim},
-\emph{unclaim} and \emph{abortIfClaimed} with the following sequential
+\emph{unclaim} and \emph{\claimcheck} with the following sequential
 semantics:
@@
 
-\item $\textit{abortIfClaimed}(k)$, $k\in\Nat$, 
+\item $\textit{\claimcheck}(k)$, $k\in\Nat$, 
 abort current transaction if $k$ is currently claimed by any controller.
@@
 
-\item $\textit{abortIfUnequal}(addr,k)$, $addr,k\in\Nat$, 
+\item $\textit{\compare}(addr,k)$, $addr,k\in\Nat$, 
 aborts current transaction if the value at memory offset $addr$ is not equal to $k$.
@@
 Algorithms~\ref{alg:claim}-\ref{alg:cas} describe implementations of  
-transactional operations  $\textit{claim}(x)$, $\textit{check}(x)$, $\textit{unclaim}(x)$,
-and $\textit{CAS}(x)$, respectively. 
+transactional operations  $\textit{claim}(x)$, $\textit{\claimcheck}(x)$, $\textit{unclaim}(x)$, $\textit{write}(x)$,  $\textit{\compare}(x)$
+and $\textit{\cas}(x)$, respectively. 
 Algorithm~\ref{alg:commit} describes the procedure executed by
@@
 
-  \liron{got till here}
-
-We use creation and deletion of flow entries as abstraction to claiming/unclaiming  resource id numbers and also to checking whether a given id is claimed.
-The match part of the added flow entry represents the added
-number and the mask part of the match varies from one controller to another in order to allow multiple controllers to claim the same number without overriding the same flow entry. 
-As can be seen in Algorithms \ref{alg:claim}-\ref{alg:unclaim}, the match mask is a concatenation of the controller id with all ones, while the value is the id concatenated twice.
-We can see that for any two numbers $x_1$ and $x_2$ and controller ids $c_1\neq c_2$ the match patterns $m_1=(x_1\concat x_1, c_1\concat all-1)$ and $ m_1=(x_2\concat x_2, c_2\concat all-1)$ are not equal even if $x_1=x_2$ therefore they can both claim and unclaim without affecting each other.
+ As can be seen in Algorithms \ref{alg:claim}-\ref{alg:unclaim}, for the resource-claimer, 
+  the match part of the added flow entry represents the claimed id (id concatenated twice).
+   In addition the mask part of the match (controller id concatenated with all-ones) varies from one controller to another in order to allow multiple controllers to claim the same number without overriding the same flow entry.
+    We can see that for any two resource ids $x_1$ and $x_2$ and controller ids $c_1\neq c_2$ the match patterns $m_1=(x_1\concat x_1, c_1\concat all-1)$ and $ m_1=(x_2\concat x_2, c_2\concat all-1)$ are not equal even if $x_1=x_2$ therefore they can both claim and unclaim without affecting each other.
 
 In order to check if a number is claimed we try to create the corresponding entry while setting the
-flag \textsf{CHECK\_OVERLAP}, thereby inflicting a failure in case an entry with same match value but different action exists.
+flag \textsf{CHECK\_OVERLAP} and using different action, thereby inflicting a failure in case an entry with overlapping match value exists.
 If the check succeeds we delete the entry.
 Considering $m_1$ and $m_2$ defined above and their respective actions $a_1=x_1\concat c1$ and $a_2=x_2\concat c2$,  they overlap with different actions iff $x_1=x_2$.
-Therefore if one is used in check after the other used in claims, a conflict is detected iff $x_1=x_2$ as expected.
+Therefore if one is used in check after the other used in claims, a conflict is detected iff $x_1=x_2$, as expected.
 Note that checking for claims must contain a deletion of the added flow entry otherwise a successive check operation might mistakenly fail.
 Moreover the two commands (adding and deleting of the flow entries) must be sent as part of the same bundle to support concurrent checks by different controllers.
-
-Similarly, we use a deletion of group $x$ and a creation of group $y$ (with GroupMod commands) as an
-abstraction to performing a $CAS(x,y)$ operation to a special $16$bit register (we abstract only one register). The $16$ bit limit is resulted by th $16$bit group id size. 
-%The group id of the added group represents the added number. 
-Note that trying to delete a group that doesn't exist (or trying to create a group with id that already exists) fails, thereby preventing adding the new value $y$ if $x$ hasn't existed before.
-Note that the two action should be performed inside the same bundle in order to ensure the atomic nature ot the abstraction.
-
-To commit the transaction,  the controller sends a \textsf{OFPT\_BUNDLE\_CONTROL} message with type
-\textsf{OFPBCT\_COMMIT\_REQUEST} and waits until a matching control
-message from the switch is received. When the switch receives the
-commit request it tries to atomically process all the buffered messages
-equipped with the corresponding bundle identifier and, if the
-processing is successful for all of them, returns  a \textsf{OFPT\_BUNDLE\_CONTROL} message with type
-\textsf{OFPBCT\_CLOSE\_REPLY}; otherwise it sends a specific error
-message  pointing out the aborted transactional operation.
 
@@
 \begin{algorithm}[t]
-    \caption{$\textit{check}(x)$}
+    \caption{$\textit{\claimcheck}(x)$}
     \label{alg:check}
@@
 
+We use similar techniques for implementing memory write and abortIfUnequal operations.
+ As can be seen in Algorithms \ref{alg:write}-\ref{alg:compare}, 
+  the match part of the added flow entry is used as the memory address and the action as the written value.
+Values written to the same address will replace one another, following the command definition to replace old flow entry with a new one in casae they share the same match parts (as described in Section \ref{sec:background}).
+
+In order to check if a value is written in a givven address, we try to add it the same way as in the write operation  but we also set the \textsf{CHECK\_OVERLAP} flag.
+If the value is not there, our action differ from the action of the existsing entry thereby inflicting a failure.
+In case the value is there, our flow entry replaces the existing flow entry with identical one.
+In addition we present in Algorithm \ref{alg:cas} a simple implementation of Comapare and set (CAS) usin the two last operations.
+
 \begin{algorithm}[t]
-    \caption{$\textit{CAS}(old,new)$}
-    \label{alg:cas}
+    \caption{$\textit{write}(addr,k)$}
+    \label{alg:write}
     \begin{algorithmic}[1]
@@
 %    \Ensure
-
-    		\State $op \gets \textsf{OFPGC\_DELETE}$
-    		\State $group\_id \gets old$
-    		\State $cmd1\gets GroupMod(op, group\_id) $
-
-    		\State $op \gets \textsf{OFPGC\_ADD}$
-    		\State $type \gets \textsf{OFPGT\_INDIRECT}$
-    		\State $group\_id \gets new$
-    		\State $buckets \gets [dummy\_backet]$
-    		\State $cmd2\gets GroupMod(op, type, group\_id, buckets) $
+    		\State $value \gets addr$
+    		\State $mask \gets  all-1$
+    		\State $match \gets (value,mask)$
+    		%\State $cookie \gets x\concat self$
+    		\State $action \gets k$
+    		\State $flag \gets 0$
+    		\State $cmd\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
+			\Return $cmd$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[t]
+    \caption{$\textit{\compare}(addr,k)$}
+    \label{alg:compare}
+    \begin{algorithmic}[1]
+    %\Require  id $i$.
+%    \Ensure
+    		\State $value \gets addr$
+    		\State $mask \gets  all-1$
+    		\State $match \gets (value,mask)$
+    		%\State $cookie \gets x\concat self$
+    		\State $action \gets k$
+    		\State $flag \gets \textsf{OFPFF\_CHECK\_OVERLAP}$
+    		\State $cmd\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
+			\Return $cmd$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[t]
+    \caption{$\textit{\cas}(addr, old,new)$}
+    \label{alg:cas}
+    \begin{algorithmic}[1]
+    %\Require  id $i$.
+%    \Ensure
+
+    		\State $cmd1\gets \textsc{abortIfUnequal}((addr, old) $
+    		\State $cmd2\gets \textsc{Write}((addr, new) $
 			\Return $cmd1,cmd2$
@@
 \end{algorithm}
+
+To commit the transaction,  the controller sends a \textsf{OFPT\_BUNDLE\_CONTROL} message with type
+\textsf{OFPBCT\_COMMIT\_REQUEST} and waits until a matching control
+message from the switch is received. When the switch receives the
+commit request it tries to atomically process all the buffered messages
+equipped with the corresponding bundle identifier and, if the
+processing is successful for all of them, returns  a \textsf{OFPT\_BUNDLE\_CONTROL} message with type
+\textsf{OFPBCT\_CLOSE\_REPLY}; otherwise it sends a specific error
+message  pointing out the aborted transactional operation.
 
@@
  			\startTxn
- 				\State $check(my\_id)$
+ 				\State $abortIfClaimed(my\_id)$
 	 			\State $CAS(pid,my\_id)$
@@
 
+\section{Implementing CAS with GroupMod commands}\label{sec:with-groups}
+Similarly, we use a deletion of group $x$ and a creation of group $y$ (with GroupMod commands) as an
+abstraction to performing a $CAS(x,y)$ operation to a special $16$bit register (we abstract only one register). The $16$ bit limit is resulted by th $16$bit group id size. 
+%The group id of the added group represents the added number. 
+Note that trying to delete a group that doesn't exist (or trying to create a group with id that already exists) fails, thereby preventing adding the new value $y$ if $x$ hasn't existed before.
+Note that the two action should be performed inside the same bundle in order to ensure the atomic nature ot the abstraction.
+
+\begin{algorithm}[H]
+    \caption{$\textit{group-CAS}(old,new)$}
+    \label{alg:gcas}
+    \begin{algorithmic}[1]
+    %\Require  id $i$.
+%    \Ensure
+
+    		\State $op \gets \textsf{OFPGC\_DELETE}$
+    		\State $group\_id \gets old$
+    		\State $cmd1\gets GroupMod(op, group\_id) $
+
+    		\State $op \gets \textsf{OFPGC\_ADD}$
+    		\State $type \gets \textsf{OFPGT\_INDIRECT}$
+    		\State $group\_id \gets new$
+    		\State $buckets \gets [dummy\_backet]$
+    		\State $cmd2\gets GroupMod(op, type, group\_id, buckets) $
+			\Return $cmd1,cmd2$
+    \end{algorithmic}
+\end{algorithm}
+
 \begin{comment}

### 2015-03-22 inband-STN.tex
updating section 5

@@
 \usepackage{amsmath}
+\usepackage{xspace}
 %\usepackage{graphicx}
@@
 \newcommand{\Nat}{\mathbb{N}}
-\newcommand{\cas}{CAS}
-\newcommand{\claimcheck}{check}
-\newcommand{\compare}{compare}
+\newcommand{\cas}{CAS\xspace}
+\newcommand{\claimcheck}{check\xspace}
+\newcommand{\compare}{compare\xspace}
+\newcommand{\memwrite}{write\xspace}
 
@@
 \subsection{The flow entry modification command (FlowMod)}
+- Creating an entry with the same match part replaces the existing one.
 
 \subsubsection{The CHECK\_OVERLAP flag}
-
+- creating and entry with overlapping match (same priority) and different instructions fails.
 \subsection{Bundling}
+- failed command aborts the bundle.
 
@@
 given (bounded or unbounded) set, and a read-write \emph{memory} abstraction
-object, which can be used for storing the current used policy identifier.
+object, which we later use for storing the current used policy identifier.
 %
@@
 
-Our implementation is based on OpenFlow flow entries modification command, the FlowMod message that explained in Section \ref{sec:background}.
+Our implementation is based on OpenFlow flow entries modification command, the FlowMod message that we explain in Section \ref{sec:background}.
 We use the CHECK\_OVERLAP optional flag to make the transaction execution dependent on switch state.
@@
    In addition the mask part of the match (controller id concatenated with all-ones) varies from one controller to another in order to allow multiple controllers to claim the same number without overriding the same flow entry.
-    We can see that for any two resource ids $x_1$ and $x_2$ and controller ids $c_1\neq c_2$ the match patterns $m_1=(x_1\concat x_1, c_1\concat all-1)$ and $ m_1=(x_2\concat x_2, c_2\concat all-1)$ are not equal even if $x_1=x_2$ therefore they can both claim and unclaim without affecting each other.
+    We can see that for any two resource ids $x_1$ and $x_2$ and controller ids $c_1\neq c_2$ the match patterns $m_1=(x_1\concat x_1, c_1\concat all-1)$ and $ m_2=(x_2\concat x_2, c_2\concat all-1)$ are not equal even if $x_1=x_2$ therefore they can both be used to claim and unclaim without affecting each other.
 
@@
 Considering $m_1$ and $m_2$ defined above and their respective actions $a_1=x_1\concat c1$ and $a_2=x_2\concat c2$,  they overlap with different actions iff $x_1=x_2$.
-Therefore if one is used in check after the other used in claims, a conflict is detected iff $x_1=x_2$, as expected.
+Therefore if one is used in \claimcheck after the other used in claim, a conflict is detected iff $x_1=x_2$, as expected.
+
 Note that checking for claims must contain a deletion of the added flow entry otherwise a successive check operation might mistakenly fail.
@@
 
-We use similar techniques for implementing memory write and abortIfUnequal operations.
+We use similar techniques for implementing memory write and \compare 
+ operations.
  As can be seen in Algorithms \ref{alg:write}-\ref{alg:compare}, 
@@
 
-    		\State $cmd1\gets \textsc{abortIfUnequal}((addr, old) $
-    		\State $cmd2\gets \textsc{Write}((addr, new) $
+    		\State $cmd1\gets \textsc{\compare}((addr, old) $
+    		\State $cmd2\gets \textsc{\memwrite}((addr, new) $
 			\Return $cmd1,cmd2$
@@
 
+\hide{
 %\section{The Container Object}\label{sec:vision}
@@
 \end{algorithm}
+}
 
@@
     \begin{algorithmic}[1]
-    \Require update requirements $U$
+    \Require update requirements $U$, a predefined (memory) address for  policy id: $pid\_addr$
     \Ensure installed policy is consistent with previous one
@@
  			\startTxn
-	 			\State $CAS(pid,pid+1)$
+	 			\State $CAS(pid\_addr,pid,pid+1)$
 	 			\State add $new\_policy$ rules %conditioned on $metadata=my\_a$
@@
  			\startTxn
- 				\State $abortIfClaimed(my\_id)$
-	 			\State $CAS(pid,my\_id)$
+ 				\State $\claimcheck(my\_id)$
+	 			\State $CAS(pid\_addr, pid,my\_id)$
 	 			\State \textbf{add} $new\_policy$ rules %conditioned on $metadata=my\_a$

### 2015-03-22 inband-STN.tex
removed old section 4 (which was commented out)

@@
 
-\hide{
-%\section{The Container Object}\label{sec:vision}
-\section{Transactional Synchronization}\label{sec:main}
-
-This section describes how to introduce more powerful transactions.
-Our implementation is based on two new synchronization objects,
-allowing to coordinate controllers using identifiers.
-
-At a high level, we export a
-transactional interface which allows each controller $i\in[N]$ ($N\geq
-2$) to perform multiple
-operations on the configuration of an OpenFlow switch as an atomic
-\emph{transaction}, inspired by the transactional-memory paradigm~\cite{stm-st95,tm-book}. A transaction may \emph{commit}, in
-which case it appears as executed sequentially, or \emph{abort}, in
-which none of its operations takes effect.
-
-\subsection{Synchronization Objects}\label{sec:t-if}
-
-We assume transactional access to  two kinds of objects: a
-\emph{claim-id} object, used for claiming a policy \emph{identifier} in a
-given (bounded or unbounded) set, and a read-write \emph{register}
-object, used for storing the current used policy identifier.
-%
-A claim-id object exports operations \emph{claim},
-\emph{unclaim} and \emph{check} with the following sequential
-semantics:
-%The container synchronization object introduced in this paper
-%is essentially a transactional memory abstraction which can be implemented
-%in standard Openflow switches.
-%Before presenting the technical details in the next section,%
-%we give a brief overview.
-
-%The postulated container object can store an arbitrary number of \emph{items}.
-%As we will see, these items correspond to flow entries on the Openflow switch.
-%The same item may even exist multiple times in the container (i.e., the container
-%is a multi-set).
-%The container
-%provides two main functions:
-\begin{itemize}
-\item by performing $\textit{claim}(k)$, $i\in[N]$,
-  $k\in\Nat$, controller $i$ \emph{claims} 
-  identifier $k$;
-
-\item by performing $\textit{unclaim}(k)$, $i\in[N]$, $k\in\Nat$, controller $i$
-  \emph{unclaim} identifier $k$;
-
-\item $\textit{check}(k)$, $k\in\Nat$, returns $\textit{true}$ if and
-  only if $k$ is not currently claimed by any controller; if
-  $\textit{false}$ is returned, we say that the check operation
-  \emph{fails} (which causes an abort of the transaction invoking it).
-
-\end{itemize}
-
-A register can be accessed with a
-\emph{compare-and-set} (CAS) operation that takes two integer parameters
-$\textit{old}$ and $\textit{new}$, reads the register and, if the
-value of it is $\textit{old}$, replaces it with $\textit{new}$ and
-returns $\textit{true}$; otherwise, $\textit{false}$ is returned and
-we say that the CAS operation \emph{fails} (which causes an abort if
-the transaction invoking it).
-
-\subsection{Interface}\label{sec:t-if}
-
-We assume that transactions may comprise operations on the flow table
-of an SDN switch. Our transactional abstraction exports the
-$\textbf{atomic}\{\;\ldots\;\}$ construct
-containing a sequence of transactional
-operations and may \emph{commit} and return responses to all its operations
-or \emph{abort} (so that none of its operation takes effect).  
-We guarantee that in every execution, all committed
-transactions constitute a \emph{legal} sequential
-order~\cite{Pap79-serial}, i.e., every transaction appears as executed
-sequentially. On the progress side, we ensure that a
-transaction $T$ is only allowed to abort if either another 
-transaction updated a flow table entry read by $T$  
-or if one of $T$'s operations (\textit{check} or CAS) fails.
-We employ the \emph{deferred-update} semantics:  
-no transaction can affect other transactions before it commits.
-In particular, if a controller crashes before committing its
-transactions, no future transaction can be affected.  
-%Note that
-%, assuming such a transactional memory,  
-%a $\emph{contains}(x)$ operation can be implemented as a transaction
-%composed of $\textit{cond-add}(x)$ followed by $\textit{cond-delete}(x)$.  
-%A \emph{set} abstraction is a special instance of a multi-set that is
-%only modified with \textit{cond-add} and \textit{cond-delete}
-%operations (this way the set only contains at most one copy of a given
-%item).
-
-\subsection{Implementation in Openflow}\label{sec:t-impl}
-
-We now describe how to implement our transactional 
-abstraction using standard OpenFlow features, including
-\emph{bundling}, \emph{conditional application}, and \emph{groups}.
-%[[PK let us leave for the moment
-%We will describe later how to implement some of these features
-%in older Openflow versions where they are not
-%supported. \textbf{TODO}
-%]]
-Algorithms~\ref{alg:claim}-\ref{alg:cas} describe implementations of  
-transactional operations  $\textit{claim}(x)$, $\textit{check}(x)$, $\textit{unclaim}(x)$,
-and $\textit{CAS}(x)$, respectively. 
-Algorithm~\ref{alg:commit} describes the procedure executed by
-a transaction at \emph{commit time}.
-
-We assume that every transaction has a unique identifier. 
-Every operation within a transaction generates a number of  OpenFlow
-messages. Each of the messages can be wrapped 
-is marked as a 
-in a \emph{bundle} message
-%,where the transaction's identifier is used as the bundle id. 
-.
-Each bundle message carries a bundle id, where messages with same bundle id are considered to belong to same bundle and are performed atomically.
-
-%The OpenFlow messages that we use (besides )
-
-\textbf{TODO: explain how the lines in the pseudocode generate OF msgs.} 
-We use creation and deletion of flow entries (with FlowMod commands) as abstraction to claiming/unclaiming  resource id numbers and also to checking whether a given id is claimed.
-The match part of the added flow entry represents the added
-number and the mask part of the match varies from one controller to another in order to allow multiple controllers to claim the same number without overriding the same flow entry. 
-As can be seen in Algorithms \ref{alg:claim}-\ref{alg:unclaim}, the match mask is a concatenation of the controller id with all ones, while the value is the id concatenated twice.
-We can see that for any two numbers $x_1$ and $x_2$ and controller ids $c_1\neq c_2$ the match patterns $m_1=(x_1\concat x_1, c_1\concat all-1)$ and $ m_1=(x_2\concat x_2, c_2\concat all-1)$ are not equal even if $x_1=x_2$ therefore they can both claim and unclaim without affecting each other.
-
-In order to check if a number is claimed we try to create the corresponding entry while setting the
-flag \textsf{CHECK\_OVERLAP}, thereby inflicting a failure in case an entry with same match value but different action exists.
-If the check succeeds we delete the entry.
-Considering $m_1$ and $m_2$ defined above and their respective actions $a_1=x_1\concat c1$ and $a_2=x_2\concat c2$,  they overlap with different actions iff $x_1=x_2$.
-Therefore if one is used in check after the other used in claims, a conflict is detected iff $x_1=x_2$ as expected.
-Note that checking for claims must contain a deletion of the added flow entry otherwise a successive check operation might mistakenly fail.
-Moreover the two commands (adding and deleting of the flow entries) must be sent as part of the same bundle to support concurrent checks by different controllers.
-
-Similarly, we use a deletion of group $x$ and a creation of group $y$ (with GroupMod commands) as an
-abstraction to performing a $CAS(x,y)$ operation to a special $16$bit register (we abstract only one register). The $16$ bit limit is resulted by th $16$bit group id size. 
-%The group id of the added group represents the added number. 
-Note that trying to delete a group that doesn't exist (or trying to create a group with id that already exists) fails, thereby preventing adding the new value $y$ if $x$ hasn't existed before.
-Note that the two action should be performed inside the same bundle in order to ensure the atomic nature ot the abstraction.
-
-To commit the transaction,  the controller sends a \textsf{OFPT\_BUNDLE\_CONTROL} message with type
-\textsf{OFPBCT\_COMMIT\_REQUEST} and waits until a matching control
-message from the switch is received. When the switch receives the
-commit request it tries to atomically process all the buffered messages
-equipped with the corresponding bundle identifier and, if the
-processing is successful for all of them, returns  a \textsf{OFPT\_BUNDLE\_CONTROL} message with type
-\textsf{OFPBCT\_CLOSE\_REPLY}; otherwise it sends a specific error
-message  pointing out the aborted transactional operation.
-
-%A container contains items: they represent the flow table entries (i.e., a match-action rules, plus e.g., counters belonging to the rule).
-%Items can be added and removed atomically and in bundles (e.g., to update the switch with
-%a new policy). Recall that we want to add items only if they do not exist
-%yet; otherwise the transaction should abort without taking effect.
-%And similarly for removes: transactions should not take effect if the corresponding
-%update contains rules which currently do not exist at the switch.
-
-\begin{algorithm}[t]
-    \caption{$\textit{claim}(x)$}
-    \label{alg:claim}
-    \begin{algorithmic}[1]
-    \Require \emph{self} as the calling controller id.
-%    \Ensure
-    		\State $value \gets x\concat x$
-    		\State $mask \gets self\concat all-1$
-	    	\State $match \gets (value,mask)$
-    		%\State $cookie \gets x\concat self$
-    		\State $action \gets x\concat self$
-    		\State $flag \gets 0$
-    		\State $cmd\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
-			\Return $cmd$
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[t]
-    \caption{$\textit{check}(x)$}
-    \label{alg:check}
-    \begin{algorithmic}[1]
-    \Require \emph{self} as the calling controller id.
-%    \Ensure
-    		\State $value \gets x\concat x$
-    		\State $mask \gets self\concat all-1$
-    		\State $match \gets (value,mask)$
-    		%\State $cookie \gets x\concat self$
-    		\State $action \gets x\concat self$
-    		\State $flag \gets \textsf{OFPFF\_CHECK\_OVERLAP}$
-    		\State $cmd1\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
-    		\State $cmd2\gets \textsc{FlowMod}(match, op = DELETE) $
-			\Return $cmd1$,$cmd2$
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[t]
-    \caption{$\textit{unclaim}(x)$}
-    \label{alg:unclaim}
-    \begin{algorithmic}[1]
-    \Require \emph{self} as the calling controller id.
-%    \Ensure
-    		\State $value \gets x\concat x$
-    		\State $mask \gets self\concat all-1$
-    		%\State $cookie \gets x\concat self$
-    		\State $match \gets (value,mask)$
-    		\State $cmd\gets \textsc{FlowMod}(match, op=DELETE, cookie) $
-
-			\Return $cmd$
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[t]
-    \caption{$\textit{CAS}(old,new)$}
-    \label{alg:cas}
-    \begin{algorithmic}[1]
-    %\Require  id $i$.
-%    \Ensure
-
-    		\State $op \gets \textsf{OFPGC\_DELETE}$
-    		\State $group\_id \gets old$
-    		\State $cmd1\gets GroupMod(op, group\_id) $
-
-    		\State $op \gets \textsf{OFPGC\_ADD}$
-    		\State $type \gets \textsf{OFPGT\_INDIRECT}$
-    		\State $group\_id \gets new$
-    		\State $buckets \gets [dummy\_backet]$
-    		\State $cmd2\gets GroupMod(op, type, group\_id, buckets) $
-			\Return $cmd1,cmd2$
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[t]
-    \caption{$\textit{try-commit}()$}
-    \label{alg:commit}
-    \begin{algorithmic}[1]
-    \Require  current bundle id $bid$.
-		    \State $cmd \gets struct ofp_bundle_ctrl_msg$
-    		\State  $cmd.bundle\_id \gets bid$
-    		\State  $cmd.type \gets $ \textsf{OFPBCT\_COMMIT\_REQUEST} 
-    		\State  $cmd.flags \gets $ \textsf{OFPBF\_ATOMIC}
-    		\State send $cmd$
-		\Return $cmd$
-    \end{algorithmic}
-\end{algorithm}
-}
-
 \section{Discussion of Use-Cases}\label{sec:apps}

### 2015-03-29 inband-sync.tex
added sync file (a copy of main file)

--- 2015-03-22 inband-STN.tex
@@
 \definecolor{mygreen}{rgb}{0,0.5,0}
+\definecolor{mygrey}{rgb}{0.5,0.5,0.5}
 
@@
 %\usepackage[noend]{algorithmic}
+
+\usepackage{listings}
 
@@
 \subsection{The flow entry modification command (FlowMod)}
+\lstset{% general command to set parameter(s)
+	basicstyle=\small, % print whole listing small
+	keywordstyle=\color{black}\bfseries, % underlined bold black keywords
+	identifierstyle=, % nothing happens
+	commentstyle=\color{mygrey}, % white comments
+	stringstyle=\ttfamily, % typewriter type for strings
+	showstringspaces=false, % no special string spaces
+	language=C}
+\begin{lstlisting}
+struct ofp_flow_mod {
+	struct ofp_header header;
+	uint64_t cookie; /* Opaque controller-issued identifier. */
+	uint64_t cookie_mask; /* Mask used to restrict the cookie bits
+		that must match when the command is
+		OFPFC_MODIFY* or OFPFC_DELETE*. A value
+		of 0 indicates no restriction. */
+	/* Flow actions. */
+	uint8_t table_id; /* ID of the table to put the flow in.
+		For OFPFC_DELETE_* commands, OFPTT_ALL
+		can also be used to delete matching
+		flows from all tables. */
+	uint8_t command; /* One of OFPFC_*. */
+	uint16_t idle_timeout; /* Idle time before discarding (seconds). */
+	uint16_t hard_timeout; /* Max time before discarding (seconds). */
+	uint16_t priority; /* Priority level of flow entry. */
+	uint32_t buffer_id; /* Buffered packet to apply to, or
+		OFP_NO_BUFFER.
+		Not meaningful for OFPFC_DELETE*. */
+	uint32_t out_port; /* For OFPFC_DELETE* commands, require
+		matching entries to include this as an
+		output port. A value of OFPP_ANY
+		indicates no restriction. */
+	uint32_t out_group; /* For OFPFC_DELETE* commands, require
+		matching entries to include this as an
+		output group. A value of OFPG_ANY
+		indicates no restriction. */
+	uint16_t flags; /* Bitmap of OFPFF_* flags. */
+	uint16_t importance; /* Eviction precedence (optional). */
+	struct ofp_match match; /* Fields to match. Variable size. */
+	/* The variable size and padded match is always followed by instructions. */
+	//struct ofp_instruction_header instructions[0];
+		/* Instruction set - 0 or more. The length
+		of the instruction set is inferred from
+		the length field in the header. */
+};
+\end{lstlisting}
+
 - Creating an entry with the same match part replaces the existing one.
@@
 - failed command aborts the bundle.
+"When the commit fails, the switch must generate an error message corresponding to the
+message that could not be applied. The error message must use the xid of the offending message,
+the error data field corresponding to that message and the error code corresponding to the failure. If
+multiple messages are generating errors, the switch may return only the first error found or generate
+multiple error messages for the same bundle."
+\subsubsection{Bundle control messages}
+
+%/* Message structure for OFPT_BUNDLE_CONTROL. */
+\begin{lstlisting}
+struct ofp_bundle_ctrl_msg {
+	struct ofp_header header;
+	uint32_t bundle_id; /* Identify the bundle. */
+	uint16_t type; /* OFPBCT_*. */
+	uint16_t flags; /* Bitmap of OFPBF_* flags. */
+	/* Bundle Property list. */
+	struct ofp_bundle_prop_header properties[0]; /* Zero or more properties. */
+};
+\end{lstlisting}
+\subsubsection{Bundle add message}
+
+%/* Message structure for OFPT_BUNDLE_ADD_MESSAGE.
+\begin{lstlisting}
+* Adding a message in a bundle is done with. */
+struct ofp_bundle_add_msg {
+	struct ofp_header header;
+	uint32_t bundle_id; /* Identify the bundle. */
+	uint16_t pad; /* Align to 64 bits. */
+	uint16_t flags; /* Bitmap of OFPBF_* flags. */
+	struct ofp_header message; /* Message added to the bundle. */
+	/* If there is one property or more, 'message' is followed by:
+	* - Exactly (message.length + 7)/8*8 - (message.length) (between 0 and 7)
+	* bytes of all-zero bytes */
+	/* Bundle Property list. */
+	//struct ofp_bundle_prop_header properties[0]; /* Zero or more properties. */
+};
+\end{lstlisting}
 
@@
 
-We assume transactional access to  two kinds of objects: a
+We assume transactional access to  two kinds of objects: a read-write \emph{memory} abstraction
+object, which we later use for storing the current used policy identifier, and a
 \emph{resource-claimer} object, provides claiming a resource id (e.g. policy \emph{identifier}) in a
-given (bounded or unbounded) set, and a read-write \emph{memory} abstraction
-object, which we later use for storing the current used policy identifier.
+given (bounded or unbounded) set.
+In theory one can use the memory abstraction to implement the resource claimer, however we found it far more simple and efficient to consider the resource claimer as base object.
 %
-\subsubsection{A resource claimer}
+
+A memory can be accessed with the following sequential
+semantics:
+
+\begin{itemize}
+\item by performing $\textit{Write}(addr,k)$,
+  $addr,k\in\Nat$, the value at memory offset $addr$ is set to $k$;
+
+\item $\textit{\compare}(addr,k)$, $addr,k\in\Nat$, 
+aborts current transaction if the value at memory offset $addr$ is not equal to $k$.
+%returns $\textit{true}$ if and
+%  only if $k$ is not currently claimed by any controller; if
+%  $\textit{false}$ is returned, we say that the check operation
+%  \emph{fails} (which causes an abort of the transaction invoking it).
+
+\end{itemize}
+
+Note that reading the memory can be done by just reading the switch configuration however such action doesn't affect the execution of the transaction (as $\textit{\compare}$ does).
+
+An immediate extension of the memory interface is the \emph{compare-and-set} (CAS) operation that takes two integer parameters
+$\textit{old}$ and $\textit{new}$, reads the register and, if the
+value of it is $\textit{old}$, replaces it with $\textit{new}$ and
+returns $\textit{true}$; otherwise, $\textit{false}$ is returned and
+we say that the CAS operation \emph{fails} (which causes an abort if
+the transaction invoking it).
+
 A resource-claimer object implements a weak type of locking, exports operations \emph{claim},
@@
 \end{itemize}
-
-\subsubsection{A memory abstraction}
-A memory can be accessed with the following sequential
-semantics:
-
-\begin{itemize}
-\item by performing $\textit{Write}(addr,k)$,
-  $addr,k\in\Nat$, the value at memory offset $addr$ is set to $k$;
-
-\item $\textit{\compare}(addr,k)$, $addr,k\in\Nat$, 
-aborts current transaction if the value at memory offset $addr$ is not equal to $k$.
-%returns $\textit{true}$ if and
-%  only if $k$ is not currently claimed by any controller; if
-%  $\textit{false}$ is returned, we say that the check operation
-%  \emph{fails} (which causes an abort of the transaction invoking it).
-
-\end{itemize}
-
-Note that reading the memory can be done by just reading the switch configuration however such action doesn't affect the execution of the transaction (as $\textit{abortIfUnequal}$ does).
-
-An immediate extension of the memory interface is the \emph{compare-and-set} (CAS) operation that takes two integer parameters
-$\textit{old}$ and $\textit{new}$, reads the register and, if the
-value of it is $\textit{old}$, replaces it with $\textit{new}$ and
-returns $\textit{true}$; otherwise, $\textit{false}$ is returned and
-we say that the CAS operation \emph{fails} (which causes an abort if
-the transaction invoking it).
 
@@
 
+\subsubsection{A resource claimer}
 %features , including
@@
 In order to check if a number is claimed we try to create the corresponding entry while setting the
-flag \textsf{CHECK\_OVERLAP} and using different action, thereby inflicting a failure in case an entry with overlapping match value exists.
+flag \textsf{CHECK\_OVERLAP} and using different action ($2$ instead of $1$), thereby inflicting a failure in case an entry with overlapping match value exists.
 If the check succeeds we delete the entry.
@@
 
-Note that checking for claims must contain a deletion of the added flow entry otherwise a successive check operation might mistakenly fail.
-Moreover the two commands (adding and deleting of the flow entries) must be sent as part of the same bundle to support concurrent checks by different controllers.
+Note that for switch space reasons, checking for claims contains also a deletion of the added tester flow entry but it has no affect on claims and checks of other controllers and therefore the deletion command doesn't have to be sent as part of a bundle.
+This makes the \textsf{\claimcheck} operation atomic by nature, similar to the \textsf{claim} and \textsf{unclaim}, as being implemented by essentially one command.
 
@@
 
-\begin{algorithm}[t]
+\begin{algorithm}[H]
     \caption{$\textit{claim}(x)$}
@@
     		%\State $cookie \gets x\concat self$
-    		\State $action \gets x\concat self$
+    		\State $action \gets 1$
     		\State $flag \gets 0$
@@
 
-\begin{algorithm}[t]
+\begin{algorithm}[H]
     \caption{$\textit{\claimcheck}(x)$}
@@
     		%\State $cookie \gets x\concat self$
-    		\State $action \gets x\concat self$
+    		\State $action \gets 2$
     		\State $flag \gets \textsf{OFPFF\_CHECK\_OVERLAP}$
@@
 
-\begin{algorithm}[t]
+\begin{algorithm}[H]
     \caption{$\textit{unclaim}(x)$}
@@
     		\State $match \gets (value,mask)$
-    		\State $cmd\gets \textsc{FlowMod}(match, op=DELETE, cookie) $
+    		\State $cmd\gets \textsc{FlowMod}(match, op=DELETE) $
 
@@
 
-We use similar techniques for implementing memory write and \compare 
+\subsubsection{A memory abstraction}
+
+We use similar techniques for implementing memory \memwrite and \compare 
  operations.
@@
   the match part of the added flow entry is used as the memory address and the action as the written value.
-Values written to the same address will replace one another, following the command definition to replace old flow entry with a new one in casae they share the same match parts (as described in Section \ref{sec:background}).
-
-In order to check if a value is written in a givven address, we try to add it the same way as in the write operation  but we also set the \textsf{CHECK\_OVERLAP} flag.
-If the value is not there, our action differ from the action of the existsing entry thereby inflicting a failure.
+Values written to the same address will replace one another, following the command definition to replace old flow entry with a new one in case they share the same match parts (as described in Section \ref{sec:background}).
+
+In order to check if a value is written in a given address, we try to add it the same way as in the \memwrite operation  but we also set the \textsf{CHECK\_OVERLAP} flag.
+If the value is not there, our action differ from the action of the existing entry thereby inflicting a failure.
 In case the value is there, our flow entry replaces the existing flow entry with identical one.
-In addition we present in Algorithm \ref{alg:cas} a simple implementation of Comapare and set (CAS) usin the two last operations.
-
-\begin{algorithm}[t]
+In addition we present in Algorithm \ref{alg:cas} a simple implementation of Compare and set (CAS) using the two last operations.
+
+\begin{algorithm}[h]
     \caption{$\textit{write}(addr,k)$}
@@
 
-\begin{algorithm}[t]
+\begin{algorithm}[h]
     \caption{$\textit{\compare}(addr,k)$}
@@
 
-\begin{algorithm}[t]
+\begin{algorithm}[h]
     \caption{$\textit{\cas}(addr, old,new)$}
@@
 
-\begin{algorithm}[t]
+\begin{algorithm}[h]
     \caption{$\textit{try-commit}()$}

### 2015-03-29 inband-sync.tex
first iteration

@@
 
-%\section*{todo before submission}
+\section*{todo}
+
+stefan add old notes, bundle size complexity (how to make it minimal: store rule in one table, select it with a pointer / selector),
+practical implementation (not even noviswitch has bundles)
 
@@
 currently used identifier.
-Thus, our paper also provides the missing link for the read-modify-write object
+The claim, unclaim, and check-claim operations can be implemented individually without the Openflow
+bundle feature.
+
+Our paper therefore also provides the missing link for the read-modify-write object
 postulated in~\cite{cpc}.
 
-Finally, we also show that interesting synchronization mechanisms may be included
-in-band even without the bundle feature, and as a case study, show
-how to automatically \emph{compose} updates originating \emph{from different controllers}.
+%Finally, we also show that interesting synchronization mechanisms may be included
+%in-band even without the bundle feature, and as a case study, show
+%how to automatically \emph{compose} updates originating \emph{from different controllers}.
 
@@
 We then present our approach in detail in Section~\ref{sec:main}, and discuss use cases in Section~\ref{sec:apps}.
-Section~\ref{sec:compo} describes how to perform in-band composition.
+%Section~\ref{sec:compo} describes how to perform in-band composition.
 We review related literature in Section~\ref{sec:relwork}, and conclude
@@
 \liron{Another direction is to have a simplified explenation as now and to refference spesific sections of the standard in side our paper.}
-\subsection{The flow entry modification command (FlowMod)}
+
+\textbf{The FlowMod Command.}
+\emph{FlowMod} is the main messages used in Openflow 
+to update the switch configuration. 
+In a nutshell, \emph{FlowMod} specifies the type of flow table modification,
+a match and an action. 
+The standard behavior is that the new rule simply replaces an existing one
+with exactly the same match, and is installed
+in addition to an existing one if the match only
+partially overlaps. The latter can result
+in inconsistencies, especially if the old and the new rule
+have the same priority. \stefan{liron, is this correct?}
+\emph{FlowMod} can be used together with a the CHECK\_OVERLAP flag
+of \emph{FlowMod}, in order to detect conflicts: if the flow spaces of an old
+and the new rule overlap, 
+that is, if there is an entry with overlapping match part but different action, 
+ and if the two rules have the same priority, then 
+\emph{FlowMod} fails.
+
+As we will see, in order to implement our synchronization objects,
+the \emph{FlowMod} and CHECK\_OVERLAP features are sufficient.
+
+\textbf{Bundling.}
+The Openflow bundle can be seen as a ``mini-transaction'':
+a bundles aggregates multiple control messages (add, remove, and modify rule)
+into one, so that either all of the control messages are
+processed, or none of them is. In particular, if, for some reason, the
+configuration request contained in one of the bundled messages cannot
+be processed, all the other requests  are rejected and an error
+message is sent to the controller that issued them.
+This bundled processing can be, when properly configured, done in a
+\emph{packet-atomic} way, so that each packet is processed according
+to the switch configuration either \emph{before} or \emph{after} the
+request was processed.
+If used together with the CHECK\_OVERLAP flag, the bundle command also returns a reason 
+if the modification failed: the failed command in the bundle:
+a CHECK\_OVERLAP flag can be defined for each command in the bundle.
+
+Concretely, when the commit fails, the switch must generate an error message corresponding to the
+message that could not be applied. The error message must use the xid of the offending message,
+the error data field corresponding to that message and the error code corresponding to the failure. If
+multiple messages are generating errors, the switch may return only the first error found or generate
+multiple error messages for the same bundle.
+
+We will make use of bundling only for policy updates. 
+Bundling is \emph{not required} to implement transactions.
+
+%\textbf{Cookies:} Cookies allow to name and retrieve individual
+%flow entries and to use masks to delete modify/delete multiple entries. 
+%Our synchronization mechanism will implement controller-specific cookies.
+%
+%\item \textbf{Groups:} Groups of action buckets can be defined in a special group table.
+Each group has a unique id and any attempt to create a group with existing id or %to delete un existing group fails (and can abort a bundle in process).
+%\end{enumerate}
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=0.8\columnwidth]{loadbal.pdf}~~
+\includegraphics[width=0.8\columnwidth]{pic.pdf}\\
+\caption{\emph{Left:} Load-balancing gone wrong. \emph{Right:} Illustration of transactional abstraction which  allows
+controllers to modify switch configurations in complex ways,
+and
+depending on other flow spaces.}\label{fig:illu}
+\end{figure*}
+
+\section{Limitations and Motivation}\label{sec:motivation}
+
+what is the poewr of bundling flowmods:
+- no cookies: delete entries now by specifying exact match part
+
+- alterantive impelmentation without: group is bounded resources, can have own table etc. dont have many controlelrs anyway so not much is needed; also maybe other policies need groups
+
+Before presenting our transactional interface and implementation,
+let us first discuss the limitations of the Openflow features discussed above.
+At first sight, the bundle feature seems to provide everything
+we need: it defines a sequence of read and write operations
+ on the switch state which are executed atomically.
+ However, such bundle transactions come with many
+ limitations. A bundle transaction does not allow
+ to change the switch state in any non-trivial way,
+ as the read and write actions in the bundle do
+ not support any conditional modifications or computations which
+ depend on the read results. By leveraging
+ the check overlap feature, it is at least possible
+ to abort transactions depending on certain conditions
+ in the overlapping
+ flow spaces. However, in many cases, for example
+ a load-balancing application,
+ it can be interesting to define more general
+ dependencies, which depend on, from a logical perspective,
+ seemingly
+ completely independent flow spaces,
+ which however share the underlying physical network.
+In other words, the bundle feature only allows us to
+define ``one-shot transactions'',
+transactions which do not support any further controller interactions.
+Given
+the limited computational options the Openflow interface of a switch offers,
+it is difficult to implement more complex behavior.
+
+\textbf{Example 1.} To give an example, let us consider a controller application in charge
+of \emph{load-balancing}. The objective of the load-balancer
+is to distribute flows equally across links or servers.
+If multiple controllers concurrently and independently
+update forwarding rules, they may unvoluntarily
+unbalance the flow allocation. Note that this
+is true independently of whether the flow spaces
+affected by the respective controllers overlap or not.
+More concretely, assume that two controllers concurrently
+decide to forward traffic $1*$ and $0*$ to the same server.
+Flow space overlap checks cannot be used to detect
+such conflicts.
+FIXME: make this example better
+
+\textbf{Example 2.} To give another example,
+consider the problem of modifying previously
+installed forwarding policies.
+While a bundle allows to conditionally add new rules or remove
+old rules (and, using the flow space overlap
+check, to abort the transaction if potentially conflicting
+rules were already there), it cannot be used
+to modify rules in non-trivial ways.
+In other words, the bundle and overlap check only helps
+for the conditional creation of rules, not their modification.
+
+\stefan{It would be cool if we could make these examples more concrete,
+at least one, and get back to it later again!}
+
+%\section{The Container Object}\label{sec:vision}
+\section{-NEW- Transactional Synchronization}\label{sec:main}
+
+This section describes how to introduce more powerful transactions.
+Our implementation is based on two new synchronization objects,
+allowing to coordinate controllers using identifiers.
+
+At a high level, we export a
+transactional interface which allows each controller $i\in[N]$ ($N\geq
+2$) to perform multiple
+operations on the configuration of an OpenFlow switch as an atomic
+\emph{transaction}, inspired by the transactional-memory paradigm~\cite{stm-st95,tm-book}. A transaction may \emph{commit}, in
+which case it appears as executed sequentially, or \emph{abort}, in
+which none of its operations takes effect.
+
+\subsection{Synchronization Objects}\label{sec:t-if}
+
+We assume transactional access to  two kinds of objects: a read-write \emph{memory} abstraction
+object, which we later use for storing the current used policy identifier, and a
+\emph{resource-claimer} object, provides claiming a resource id (e.g. policy \emph{identifier}) in a
+given (bounded or unbounded) set.
+In theory one can use the memory abstraction to implement the resource claimer, however we found it far more simple and efficient to consider the resource claimer as base object.
+%
+
+A memory can be accessed with the following sequential
+semantics:
+
+\begin{itemize}
+\item by performing $\textit{Write}(addr,k)$,
+  $addr,k\in\Nat$, the value at memory offset $addr$ is set to $k$;
+
+\item $\textit{\compare}(addr,k)$, $addr,k\in\Nat$,
+aborts current transaction if the value at memory offset $addr$ is not equal to $k$.
+%returns $\textit{true}$ if and
+%  only if $k$ is not currently claimed by any controller; if
+%  $\textit{false}$ is returned, we say that the check operation
+%  \emph{fails} (which causes an abort of the transaction invoking it).
+
+\end{itemize}
+
+Note that reading the memory can be done by just reading the switch configuration however such action doesn't affect the execution of the transaction (as $\textit{\compare}$ does).
+
+An immediate extension of the memory interface is the \emph{compare-and-set} (CAS) operation that takes two integer parameters
+$\textit{old}$ and $\textit{new}$, reads the register and, if the
+value of it is $\textit{old}$, replaces it with $\textit{new}$ and
+returns $\textit{true}$; otherwise, $\textit{false}$ is returned and
+we say that the CAS operation \emph{fails} (which causes an abort if
+the transaction invoking it).
+
+A resource-claimer object implements a weak type of locking, exports operations \emph{claim},
+\emph{unclaim} and \emph{\claimcheck} with the following sequential
+semantics:
+%The container synchronization object introduced in this paper
+%is essentially a transactional memory abstraction which can be implemented
+%in standard Openflow switches.
+%Before presenting the technical details in the next section,%
+%we give a brief overview.
+
+%The postulated container object can store an arbitrary number of \emph{items}.
+%As we will see, these items correspond to flow entries on the Openflow switch.
+%The same item may even exist multiple times in the container (i.e., the container
+%is a multi-set).
+%The container
+%provides two main functions:
+\begin{itemize}
+\item by performing $\textit{claim}(k)$, $i\in[N]$,
+  $k\in\Nat$, controller $i$ \emph{claims}
+  identifier $k$;
+
+\item by performing $\textit{unclaim}(k)$, $i\in[N]$, $k\in\Nat$, controller $i$
+  \emph{unclaim} identifier $k$;
+
+\item $\textit{\claimcheck}(k)$, $k\in\Nat$,
+abort current transaction if $k$ is currently claimed by any controller.
+%returns $\textit{true}$ if and
+%  only if $k$ is not currently claimed by any controller; if
+%  $\textit{false}$ is returned, we say that the check operation
+%  \emph{fails} (which causes an abort of the transaction invoking it).
+
+\end{itemize}
+
+\subsection{Interface}\label{sec:t-if}
+
+\liron{this subsection doesn't flow well for me}
+
+We assume that transactions may comprise operations on the flow table
+of an SDN switch. Our transactional abstraction exports the
+$\textbf{atomic}\{\;\ldots\;\}$ construct
+containing a sequence of transactional
+operations and may \emph{commit} and return responses to all its operations
+or \emph{abort} (so that none of its operation takes effect).
+We guarantee that in every execution, all committed
+transactions constitute a \emph{legal} sequential
+order~\cite{Pap79-serial}, i.e., every transaction appears as executed
+sequentially. On the progress side, we ensure that a
+transaction $T$ is only allowed to abort if either another
+transaction updated a flow table entry read by $T$
+or if one of $T$'s operations (\textit{check} or CAS) fails.
+We employ the \emph{deferred-update} semantics:
+no transaction can affect other transactions before it commits.
+In particular, if a controller crashes before committing its
+transactions, no future transaction can be affected.
+%Note that
+%, assuming such a transactional memory,
+%a $\emph{contains}(x)$ operation can be implemented as a transaction
+%composed of $\textit{cond-add}(x)$ followed by $\textit{cond-delete}(x)$.
+%A \emph{set} abstraction is a special instance of a multi-set that is
+%only modified with \textit{cond-add} and \textit{cond-delete}
+%operations (this way the set only contains at most one copy of a given
+%item).
+
+\subsection{Implementation in Openflow}\label{sec:t-impl}
+
+We now describe how to implement our transactional
+abstractions using standard OpenFlow features.
+
+\subsubsection{A resource claimer}
+%features , including
+%\emph{bundling}, \emph{conditional application}, and \emph{groups}.
+%[[PK let us leave for the moment
+%We will describe later how to implement some of these features
+%in older Openflow versions where they are not
+%supported. \textbf{TODO}
+%]]
+Algorithms~\ref{alg:claim}-\ref{alg:cas} describe implementations of
+transactional operations  $\textit{claim}(x)$, $\textit{\claimcheck}(x)$, $\textit{unclaim}(x)$, $\textit{write}(x)$,  $\textit{\compare}(x)$
+and $\textit{\cas}(x)$, respectively.
+Algorithm~\ref{alg:commit} describes the procedure executed by
+a transaction at \emph{commit time}.
+
+We describe each operation as a message factory that returns OpenFlow commands that should be sent to the switch as part of the same transaction.
+We assume that every transaction has a unique identifier  used to start a \emph{bundle}.
+All messages generated during a transaction invocation, are wrapped by bundle messages carrying the bundle's id and followed by bundle commit message.
+Messages with same bundle id are considered to belong to same bundle and are performed atomically.
+
+%The OpenFlow messages that we use (besides )
+
+Our implementation is based on OpenFlow flow entries modification command, the FlowMod message that we explain in Section \ref{sec:background}.
+We use the CHECK\_OVERLAP optional flag to make the transaction execution dependent on switch state.
+
+We consider the match part of an entry to be a ternary string of bounded length (represented by binary value and mask) which can be applied to any set of supported packet header fields.
+In our implementation the $64$bit metedata field alone is suffice.
+Moreover we consider the action part of an entry to be an integer, which in practice can be implemented by a Set-Metadata instruction where the written value is that integer.
+
+ As can be seen in Algorithms \ref{alg:claim}-\ref{alg:unclaim}, for the resource-claimer,
+  the match part of the added flow entry represents the claimed id (id concatenated twice).
+   In addition the mask part of the match (controller id concatenated with all-ones) varies from one controller to another in order to allow multiple controllers to claim the same number without overriding the same flow entry.
+    We can see that for any two resource ids $x_1$ and $x_2$ and controller ids $c_1\neq c_2$ the match patterns $m_1=(x_1\concat x_1, c_1\concat all-1)$ and $ m_2=(x_2\concat x_2, c_2\concat all-1)$ are not equal even if $x_1=x_2$ therefore they can both be used to claim and unclaim without affecting each other.
+
+In order to check if a number is claimed we try to create the corresponding entry while setting the
+flag \textsf{CHECK\_OVERLAP} and using different action ($2$ instead of $1$), thereby inflicting a failure in case an entry with overlapping match value exists.
+If the check succeeds we delete the entry.
+Considering $m_1$ and $m_2$ defined above and their respective actions $a_1=x_1\concat c1$ and $a_2=x_2\concat c2$,  they overlap with different actions iff $x_1=x_2$.
+Therefore if one is used in \claimcheck after the other used in claim, a conflict is detected iff $x_1=x_2$, as expected.
+
+Note that for switch space reasons, checking for claims contains also a deletion of the added tester flow entry but it has no affect on claims and checks of other controllers and therefore the deletion command doesn't have to be sent as part of a bundle.
+This makes the \textsf{\claimcheck} operation atomic by nature, similar to the \textsf{claim} and \textsf{unclaim}, as being implemented by essentially one command.
+
+%A container contains items: they represent the flow table entries (i.e., a match-action rules, plus e.g., counters belonging to the rule).
+%Items can be added and removed atomically and in bundles (e.g., to update the switch with
+%a new policy). Recall that we want to add items only if they do not exist
+%yet; otherwise the transaction should abort without taking effect.
+%And similarly for removes: transactions should not take effect if the corresponding
+%update contains rules which currently do not exist at the switch.
+
+\begin{algorithm}[H]
+    \caption{$\textit{claim}(x)$}
+    \label{alg:claim}
+    \begin{algorithmic}[1]
+    \Require \emph{self} as the calling controller id.
+%    \Ensure
+    		\State $value \gets x\concat x$
+    		\State $mask \gets self\concat all-1$
+	    	\State $match \gets (value,mask)$
+    		%\State $cookie \gets x\concat self$
+    		\State $action \gets 1$
+    		\State $flag \gets 0$
+    		\State $cmd\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
+			\Return $cmd$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[H]
+    \caption{$\textit{\claimcheck}(x)$}
+    \label{alg:check}
+    \begin{algorithmic}[1]
+    \Require \emph{self} as the calling controller id.
+%    \Ensure
+    		\State $value \gets x\concat x$
+    		\State $mask \gets self\concat all-1$
+    		\State $match \gets (value,mask)$
+    		%\State $cookie \gets x\concat self$
+    		\State $action \gets 2$
+    		\State $flag \gets \textsf{OFPFF\_CHECK\_OVERLAP}$
+    		\State $cmd1\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
+    		\State $cmd2\gets \textsc{FlowMod}(match, op = DELETE) $
+			\Return $cmd1$,$cmd2$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[H]
+    \caption{$\textit{unclaim}(x)$}
+    \label{alg:unclaim}
+    \begin{algorithmic}[1]
+    \Require \emph{self} as the calling controller id.
+%    \Ensure
+    		\State $value \gets x\concat x$
+    		\State $mask \gets self\concat all-1$
+    		%\State $cookie \gets x\concat self$
+    		\State $match \gets (value,mask)$
+    		\State $cmd\gets \textsc{FlowMod}(match, op=DELETE) $
+
+			\Return $cmd$
+    \end{algorithmic}
+\end{algorithm}
+
+\subsubsection{A memory abstraction}
+
+We use similar techniques for implementing memory \memwrite and \compare
+ operations.
+ As can be seen in Algorithms \ref{alg:write}-\ref{alg:compare},
+  the match part of the added flow entry is used as the memory address and the action as the written value.
+Values written to the same address will replace one another, following the command definition to replace old flow entry with a new one in case they share the same match parts (as described in Section \ref{sec:background}).
+
+In order to check if a value is written in a given address, we try to add it the same way as in the \memwrite operation  but we also set the \textsf{CHECK\_OVERLAP} flag.
+If the value is not there, our action differ from the action of the existing entry thereby inflicting a failure.
+In case the value is there, our flow entry replaces the existing flow entry with identical one.
+In addition we present in Algorithm \ref{alg:cas} a simple implementation of Compare and set (CAS) using the two last operations.
+
+\begin{algorithm}[h]
+    \caption{$\textit{write}(addr,k)$}
+    \label{alg:write}
+    \begin{algorithmic}[1]
+    %\Require  id $i$.
+%    \Ensure
+    		\State $value \gets addr$
+    		\State $mask \gets  all-1$
+    		\State $match \gets (value,mask)$
+    		%\State $cookie \gets x\concat self$
+    		\State $action \gets k$
+    		\State $flag \gets 0$
+    		\State $cmd\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
+			\Return $cmd$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[h]
+    \caption{$\textit{\compare}(addr,k)$}
+    \label{alg:compare}
+    \begin{algorithmic}[1]
+    %\Require  id $i$.
+%    \Ensure
+    		\State $value \gets addr$
+    		\State $mask \gets  all-1$
+    		\State $match \gets (value,mask)$
+    		%\State $cookie \gets x\concat self$
+    		\State $action \gets k$
+    		\State $flag \gets \textsf{OFPFF\_CHECK\_OVERLAP}$
+    		\State $cmd\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
+			\Return $cmd$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[h]
+    \caption{$\textit{\cas}(addr, old,new)$}
+    \label{alg:cas}
+    \begin{algorithmic}[1]
+    %\Require  id $i$.
+%    \Ensure
+
+    		\State $cmd1\gets \textsc{\compare}((addr, old) $
+    		\State $cmd2\gets \textsc{\memwrite}((addr, new) $
+			\Return $cmd1,cmd2$
+    \end{algorithmic}
+\end{algorithm}
+
+To commit the transaction,  the controller sends a \textsf{OFPT\_BUNDLE\_CONTROL} message with type
+\textsf{OFPBCT\_COMMIT\_REQUEST} and waits until a matching control
+message from the switch is received. When the switch receives the
+commit request it tries to atomically process all the buffered messages
+equipped with the corresponding bundle identifier and, if the
+processing is successful for all of them, returns  a \textsf{OFPT\_BUNDLE\_CONTROL} message with type
+\textsf{OFPBCT\_CLOSE\_REPLY}; otherwise it sends a specific error
+message  pointing out the aborted transactional operation.
+
+\begin{algorithm}[h]
+    \caption{$\textit{try-commit}()$}
+    \label{alg:commit}
+    \begin{algorithmic}[1]
+    \Require  current bundle id $bid$.
+		    \State $cmd \gets struct ofp_bundle_ctrl_msg$
+    		\State  $cmd.bundle\_id \gets bid$
+    		\State  $cmd.type \gets $ \textsf{OFPBCT\_COMMIT\_REQUEST}
+    		\State  $cmd.flags \gets $ \textsf{OFPBF\_ATOMIC}
+    		\State send $cmd$
+		\Return $cmd$
+    \end{algorithmic}
+\end{algorithm}
+
+\section{Discussion of Use-Cases}\label{sec:apps}
+
+More powerful transactions can be useful in many scenarios.
+Let us revisit our examples from earlier: with
+transactions we can easily solve the load-balancing
+and rule modification problem.
+In the following we give two more examples of our approach.
+
+\subsection{Policy Updates}
+
+Our first example (Algorithm~\ref{alg:simple-update}) is a simple policy composition protocol that
+allows a set of  controllers to concurrently update switch  \emph{policies} , i.e., sets of
+effective rules in its flow tables, under the
+condition that these rules can be composed with the currently installed
+policy (the composition predicate is encoded in a user-specific
+$\textit{apply\_update}()$ function~\cite{cpc}.
+
+Every policy is associated with a \emph{policy identifier}
+In the algorithm, the controller first reads the currently installed
+policy, then choose a new unique policy, checks whether the new
+policy and the current policy can be composed and, if so, replaces the
+current policy with the composition. If, meanwhile, a new
+policy (with a different identifier) has been installed, the update
+fails and takes no effect.
+
+Algorithm~\ref{alg:simple-update} use increasing numbers as policies
+ids and use them in order to make sure that the new policy, with id
+$x$, replaces the (what considered as) current policy, with id
+$x-1$. Considering the policies ids, we send the policy installation
+commands as part of a bundle that also tries to perform
+$CAS(x-1,x)$, thereby insuring that the commands are
+performed iff the current policy is still there and preventing other
+controllers to replace the new policy with another without observing it.
+
+For brevity, we omit explicitly putting the $\textbf{atomic}\{...\}$ construct
+around each of the transactional operation used in the algorithm.
+Indeed, since \textit{claim} and \textit{unclaim} operations cannot
+cause the corresponding transaction, we can assume that all
+``singleton'' (consisting of one operation only)  transactions
+containing such operations always commit.
+
+Algorithm~\ref{alg:simple-update} allows policy identifiers to grow
+without bound. In fact, as was observed in~\cite{cpc}, the number of currently
+used policies can be made proportional to the number of concurrent
+controllers.
+
+ It takes a set of rules $U$ (not yet a policy)and proceeds as follows: first, we seek to
+ obtain a unique \emph{id}. FIXME: to be continued...
+ \textbf{LS: I am not sure if I need to tell every step of the story or maybe it best to explain the main dif from the previous one similarly to what I just wrote next}.
+
+Algorithm~\ref{alg:update} claims the current policy id in order to prevent someone of using it with different policy in the future (during the current update) and in the bundle used to update the policy it also checks that the new policy id that it about to use (chosen from a finite set of ids) wasn't claimed by someone else.
+
+\hide{
+We compute our new suggested policy by applying the update requests on top of current policy, supporting any kind of requests and policies.
+Then we make a transaction (using the bundle feature) to atomically check that our policy id is not blocked by anyone else, to change the current policy id to ours (an action that would fail if the current policy id is no longer what we are counting of) and to actually configure our new policy.
+
+If one of the actions in the transaction fails we try again.
+There is no progress guaranty for each controller but there is one for the whole system - at least one of the controller will succeed in fulfilling its update requirements.
+}
+
+\textbf{TODO: fix the pseudocode, make the use of afdd/delete
+  operations uniform, make all switch operations transactional}
+
+\begin{algorithm}[t]
+    \caption{Policy update with only CAS}
+    \label{alg:simple-update}
+    \begin{algorithmic}[1]
+    \Require update requirements $U$, a predefined (memory) address for  policy id: $pid\_addr$
+    \Ensure installed policy is consistent with previous one
+ 		%\Do
+ 		\Repeat
+ 			\State \textbf{read} switch configuration
+ 			\State $pid\gets$ current policy id
+ 			\State $cur\_policy\gets$ current policy
+ 			\State $new\_policy\gets apply\_update(cur\_policy,U)$
+ 			\startTxn
+	 			\State $CAS(pid\_addr,pid,pid+1)$
+	 			\State add $new\_policy$ rules %conditioned on $metadata=my\_a$
+	 			%\State update policy selector to $my\_a$
+ 			\endTxn
+     		%\State $res \gets $ commit transaction
+     	\Until{$res\neq\texttt{abort}$}
+     	%\doWhile{$res=False$}
+%   \vspace{3mm}
+			\Return
+
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[t]
+    \caption{Advanced policy update}
+    \label{alg:update}
+    \begin{algorithmic}[1]
+    \Require update requirements $U$, a predefined (memory) address for  policy id: $pid\_addr$
+    \Ensure installed policy is consistent with previous one
+ 		%\Do
+ 		\Repeat
+ 			\State \textbf{read} switch configuration
+ 			\State $my\_id\gets$ an unused \emph{id}
+ 			\State $pid\gets$ current policy id
+ 			\State $claim(pid)$
+ 			\State \textbf{read} switch configuration
+ 			\If {$pid\notin Cur$}
+	 			\State $unclaim(pid)$
+ 				\State {\bf continue} (\textbf{restart} loop)
+ 			\EndIf
+ 			\State $cur\_policy\gets$ current policy
+ 			\State $new\_policy\gets apply\_update(cur\_policy,U)$
+ 			\startTxn
+ 				\State $\claimcheck(my\_id)$
+	 			\State $CAS(pid\_addr, pid,my\_id)$
+	 			\State \textbf{add} $new\_policy$ rules %conditioned on $metadata=my\_a$
+	 			%\State update policy selector to $my\_a$
+ 			\endTxn
+	 		\State $unclaim(pid)$
+     		%\State $res \gets $ commit transaction
+     	\Until{$res\neq\texttt{abort}$}
+     	%\doWhile{$res=False$}
+%   \vspace{3mm}
+			\Return
+
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[t]
+    \caption{Policy composition without bundle}
+    \label{alg:wobundle}
+    \begin{algorithmic}[1]
+    \Require policy $P$ (a set of FlowMod entry creation commands), three consecutive flow tables ids (per controller) $T_s,T_0,T_1$, active table index $active\in\{0,1\}$.
+    %\Ensure installed policy is consistent with previous one
+    \State $unactive \gets active + 1 \mod{2}$
+    \For {$cmd \in P$}
+	    \State $cmd.action\gets Reencode(cmd.action)$
+	    \State $cmd.table\_id\gets T_{unactive}$
+	    \State send $cmd$
+    \EndFor
+    \State modify the action of $T_s$ entry to goto $T_{unactive}$
+    \State clear table $T_{active}$
+    \State $active \gets unactive$
+	\Return
+
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[t]
+    \caption{Update with pipeline}
+    \label{alg:pipeline}
+    \begin{algorithmic}[1]
+    \Require update requirements $U$, array $ARR$.
+    \Ensure installed policy is consistent with previous one
+			\Return
+
+    \end{algorithmic}
+\end{algorithm}
+
+\subsection{Consensus and Replicated State Machines}
+
+It is straightforward to implement consensus among the controllers
+using just a single register that can be accessed with an atomic CAS
+operation~\cite{Her91}. In our transactional setting, the controllers
+can use CAS to replace the default value of such a register
+with their consensus input values (e.g., the identifiers of the
+policies that should be applied first). If the corresponding
+singleton transaction commits, the controller consideres itself
+the \emph{winner} and decides on its input, if
+it aborts (the controller fails in the CAS) it remains to read the register to get the input of the
+winner to decide.
+
+By using multiple CAS invocations, the controllers can implement any
+generic replicated state machine~\cite{Her91,Lam98}, which will be instrumental in the
+next section.
+
+%\paragraph{One-Shot Update} FIXME: desribe how to implement one-shot update
+
+\paragraph{STN}
+
+\textbf{TODO}
+
+\section{In-Band Composition}\label{sec:compo}
+
+We conclude our technical contribution by showing that sometimes,
+powerful synchronization mechanisms can be implemented
+in-band \emph{even without bundles}.
+Concretely, we show that sometimes it is even possible
+to automatically compose updates originating from different controllers,
+in-band and just using Openflow.
+
+To illustrate this point, we consider a simple example
+with two controllers concurrently issuing updates to a single
+given switch.
+
+FIXME: Explain, maybe even with an example?
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.8\columnwidth]{composition.pdf}\\
+\caption{Composition.}\label{fig:illu}
+\end{figure}
+
+\begin{comment}
+
+\section{Transactions in Openflow}\label{sec:sync}
+
+Multi-write and read-modify transactions.
+
+\subsection{Dealing with multiple updates}
+
+TBD - adding the notion of versions and cleanup/GC. probably no good way to solve infinite versions... maybe check literature on shared memory and multiple writes.
+
+\subsection{Improving the scheme with check-overlap}\label{sec:todo}
+
+Till now we considered a model that allows multiple updates to match-action entries.
+Here we extends the model with actions create-entry and delete-entry that may failed and abort the whole transaction depending on the existence on non-existence of specific entries prior to the action.
+%TBD - much less resources. easily solving the infinite versions issue.
+
+Next we show how such actions can be used to implement consensus with much less resources. We use entry id as a property used to reference specific entry.
+
+\section{Applications and ramifications}\label{sec:application}
+
+\section{The Underlying Theory}\label{sec:background}
+
+idea: multi-write consensus, a not well-known result!
+
+In order to solve consensus we suggest to use the atomic multiple entries update capability (which is discussed in Openflow standard).
+A shared memory system that supports atomic write to (enough) multiple locations is known to allow the implementation of consensus objects.
+For example consider the following implementation utilizing $n^2$ memory locations, $M_{i,j}$, where $i,j\in[n]$, and the suggested values $\{v_i\}_{i\in [n]}$.
+All locations are initialized to zero.
+
+In order to participate, a server i, writes its suggested value $V_i$ and atomically rewrites row i ($M_{i,*}$) with value "1" and column i ($M_{*,i}$) with value "2".
+After the atomic write, a server $i$ first reads the diagonal ($M_{j,j}$ for $j\in [n]$), and consider all observed servers as candidate set S.
+Then server i reads all candidates intersections (locations $M_{j,k}$ where $j,k\in S$.
+And computes the winner server id w that was overwritten by all other candidates, i.e.
+such that $M_{w,k}=2$ for all $k\in S {w}$.
+The consensus value is $v_w$.
+Note that a snapshot could replace the two reading phases.
+
+\section{Basic One-Touch Mechanism}\label{sec:realization}
+
+We address the problem of committing consistent policies to SDN switch concurrently by multiple controllers.
+We show that assuming minimalistic control protocol, SDN switch configuration space can be used to implement a consensus object thereby allowing controllers to agree on each next policy update.
+
+\subsection{SDN switch configuration space as shared memory}
+
+Our first observation is that a match-action configuration which is accessed by multiple controllers can be considered as shared memory.
+Note however that standard memory map and index/offset to a word, while Flow Table entries are not necessarily indexed.
+However, we assume that each flow entry has some property that allows the operator to specifically reference and update it, for example Openflow entry's cookie.
+In cases where this property value can't be determined by the operator, we can use other variable entry properties to store the index, for example we can represent memory value $v$ in offset $i$ by an match-action entry whose match is $in\_port = i$ and action is "forward port $x$".
+%To overcome this we can use the match part of an entry as the index and the action part as the word value.
+
+Going back to the policy update problem, the consensus scheme allow all servers to decide on next policy by using consensus value as policy identifier, and writing the policy in advance.
+However this scheme doesn't actually apply the policy on incoming packets and it remains to configure the switch according to the next policy once it is decided.
+Making this scheme fail safe (handling failure of winning server) can be ensured by allowing any server to reconfigure the switch.
+
+\subsection{ One-Touch}
+
+In some scenarios it may be desirable to shorten the policy update time by avoiding the consensus reading and policy configuration step.
+We define this as the one touch policy update, namely we require each server to contact the switch only once in order to (try and) update the policy, in other words participating in reaching consensus and applying it with one atomic write.
+Solving this problem also has significance in understanding the strength of the SDN switch computation model.
+
+The main idea of our solution is to adjust the previous policy update scheme in a way that performs consensus validation phase during every packet processing thereby requiring the servers only to make the first phase of the consensus - atomically writing the matrix.
+In order to allow the packets to validate the consensus we make the following observation: while servers can be abstracted as memory writers, the packets processed by the switch are the readers.
+Openflow standard define an atomic write transaction as being atomic in relation to the processed packets thereby making packet processing a multiple read transaction but one read location per table is allowed.
+This means that if each memory location used in the consensus scheme will be stored in different table then each packet will be able to read all of them.
+
+However, reading each memory location is not enough, as the scheme requires to make a validation that involves multiple values.
+We utilize the metadata field in order to store all read values, in more details, each memory value $M_{i,j}$ is stored in offset i+j*n in the metadata field.
+Once all values are in the metadata field, we can detect the consensus winner by trying different matches on the metadata.
+Although we can have one match entry per possible matrix state, this solution requires exponential number of entries.
+A better way which we describe next requires only $O(n^2)$ entries utilizing n tables.
+
+Our compact validation use additional temp variable, $temp_winner$, to hold the current best consensus candidate.
+Each of the n validating tables checks different "column" in the matrix and update $temp_winner$ to keep track of the winner so far (after examining a prefix sub set of the columns).
+The entry t in table k $(0<=k<n)$  matches the case where current winner=t and location $(t,k)$ in the matrix (packet header) equals 1 and location (k,k) is non-zero.
+After all validating tables are processed, $temp_winner$ stores the id of the winning server which can be used to match on its policy rules (filtering out other policies) or to use the goto table command to jump to a designated policy table of the server.
+
+\section{The Practical Implementation}\label{sec:extension}
+
+\end{comment}
+
+\section{Related Work}\label{sec:relwork}
+
+There exists a wide consensus that SDN control planes need to be distributed.~\cite{onos,onix,elasticon}
+Onix~\cite{onix} is one of the earliest proposals, and introduced the
+the notion of Network Information Base (NIB), abstracting network state
+distribution from control logic, but
+requiring mechanisms for the detection and resolution of conflicts.
+Also spatially distributed control planes to improve scalability and
+latency have been studied intensively
+in the literature~\cite{kandoo,ctrl-place,hotsdn13loc}
+proposes an elastic distributed controller architecture.
+
+Operating and updating SDNs consistently is already non-trivial
+from the perspective of a single controller. Reitblatt et al.~\cite{network-update}
+introduced the notion of
+per-packet consistency and proposed a 2-phase update protocol based on tagging.
+%, and described the \emph{two-phase update} technique, also used in our algorithms.
+Mahajan and Wattenhofer~\cite{roger-hotnets} considered weaker transient
+consistency guarantees, and proposed more efficient network update algorithms
+accordingly. Ludwig et al.~~\cite{hotnets14update} studied algorithms for secure
+network updates where packets are forced to traverse certain waypoints or
+middleboxes. Ghorbani et al.~~\cite{correct-virt} recently argued for the design
+of network update algorithms that provide even stronger consistency guarantees.
+
+To the best of our knowledge, the only paper explicitly addressing the consistent
+network update problem from a distributed-controllers perspective is STN~\cite{stn}:
+STN provides a transactional interface offering all-or-nothing semantics and serializability
+(the ``holy grail'' of safety properties), allowing
+controllers to install policies in a conflict-free manner; if some controllers fail,
+other controllers can take over. STN~\cite{stn} is based on
+atomic read-modify-write primitives provided by the switches, but
+while the availability of such primitives is postulated, no implementation is
+described. In this paper, we provide this missing link, and also show that other useful and
+more powerful synchronization primitives can be implemented.
+Also, some STN implementations in~\cite{stn} assume that the control
+plane maintains a certain amount of synchrony that allows controllers
+to solve consensus or implement a replicated state machine.
+In this paper, we describe an OpenFlow-based consensus protocol that
+can be used for this purpose.
+
+Our work is also closely related to the recent work on providing more high-level
+abstractions and programming languages for SDNs, such as Frenetic~\cite{frenetic},
+also enabling policy composition~\cite{pyretic}. Our work complements these works
+on programming abstractions
+by initiating the discussion of concurrency objects and abstractions.
+
+The transactional interface we describe in this paper is inspired by the
+work on speculative concurrency control using software transactional memory
+(STM)~\cite{stm-st95,tm-book}.
+%[[PK I think this is enough]]
+
+Our work also contributes to the ongoing discussion of what can be implemented
+in-band in Openflow~\cite{compute,reclaim}, by assuming the relevant perspective of distributed
+synchronization.
+
+\hide{
+In terms of distributed computing: TODO PETR: maybe modify a bit the text below
+and also write more about multi-writer consensus etc.?
+
+\noindent\textbf{Distributed Computing.}
+There is a long tradition of defining correctness of a concurrent system via
+an equivalence to a sequential one~\cite{Pap79-serial,Lam79,HW90}.  The notion
+of sequentially composable histories is reminiscent of
+linearizability~\cite{HW90}, where a history of operations concurrently
+applied by a collection of processes is equivalent to a history in which the
+operations are in a sequential order, respecting their real-time precedence.
+In contrast, our sequentially composable histories impose requirements not
+only on high-level invocations and responses, but also on the way the traffic
+is processed. We require that the committed policies constitute a
+conflict-free sequential history, but, additionally,  we expect that each
+\emph{path} witnesses only a prefix of this history, consisting of all
+requests that were committed before the path was initiated.
+%
+The transactional interface exported by the CPC abstraction is inspired by the
+work on speculative concurrency control using software transactional memory
+(STM)~\cite{stm-st95}.
+%, thus the  term \emph{software transactional networking}.
+Our interface is however intended to model realistic network
+management operations, which makes it simpler than recent
+dynamic STM models~\cite{dstm}.
+%In particular, the sets of rules to be installed by a policy update do not depend on the state of the
+%network.
+
+Extending the interface to dynamic policies that adapt their behavior based on
+the current network state sounds like a promising research direction.  On the
+other hand, our criterion of sequential composability is more complex than
+traditional STM correctness properties in that it imposes restrictions not only
+on the high-level interface exported to the control plane, but also on the
+paths taken by the data-plane packets.
+Also, we assumed that controllers are subject to failures, which is usually not
+assumed by STM implementations.
+}
+
+\section{Conclusion}\label{sec:conclusion}
+
+While much progress has been made over the
+last years in the development of more intuitive programming
+languages for SDNs (e.g., Frenetic, Pyretic, Merlin), the
+question of which synchronization abstractions to provide
+has received much less attention.
+%
+In this paper, we describe transactional abstractions that
+can be implemented using standard Openflow features.
+%
+Our work complements existing research on the design of
+distributed control planes, and also provides some missing links,
+e.g., for the transactional approach taken by STN~\cite{stn}.
+We also hope that our work can nourish the discussion of
+which synchronization objects we can and want to implement
+in an SDN.
+
+{
+\bibliographystyle{abbrv}
+\bibliography{references}  % main.bib is the name of the Bibliography in this case
+}
+
+\cleardoublepage
+
+\onecolumn
+
+\begin{appendix}
+
+\section{The FlowMod Command}
+
 \lstset{% general command to set parameter(s)
@@
 
-\subsubsection{The CHECK\_OVERLAP flag}
-- creating and entry with overlapping match (same priority) and different instructions fails.
-\subsection{Bundling}
-- failed command aborts the bundle.
-"When the commit fails, the switch must generate an error message corresponding to the
-message that could not be applied. The error message must use the xid of the offending message,
-the error data field corresponding to that message and the error code corresponding to the failure. If
-multiple messages are generating errors, the switch may return only the first error found or generate
-multiple error messages for the same bundle."
-\subsubsection{Bundle control messages}
+\section{Bundle Control Message}
 
@@
 \end{lstlisting}
-\subsubsection{Bundle add message}
-
-%/* Message structure for OFPT_BUNDLE_ADD_MESSAGE.
+
 \begin{lstlisting}
@@
 
-\textbf{Useful features for synchronization.} The Openflow standard (as of \textbf{TODO}) offers four interesting
-features which we will exploit to implement the transactional memory abstraction:
-
-\begin{enumerate}
-\item \textbf{Bundling:}
-A bundle can be seen as a ``mini-transaction'':
-a bundles aggregates multiple control messages
-into one \textsc{FlowMod} message, so that either all of the control messages are
-processed, or none of them is. In particular, if, for some reason, the
-configuration request contained in one of the bundled messages cannot
-be processed, all the other requests  are rejected and an error
-message is sent to the controller that issued them.
-This bundled processing can be, when properly configured, done in a
-\emph{packet-atomic} way, so that each packet is processed according
-to the switch configuration either \emph{before} or \emph{after} the
-request was processed.
-
-\item \textbf{Check Overlap:}
-The Openflow standard also supports the \emph{conditional} creation of
-flow entries. A specific $\textsf{OFPFF\_CHECK\_OVERLAP}$ flag
-set in the flow creation message ensures
-that the flow entry is created if and only if it does not \emph{conflict}
-with any other flow entry currently in the flow table.
-This  conflict detection mechanism, combined with bundling,
-allows us to implement specific types of \emph{transactions} that may
-atomically install multiple rules if certain compatibility conditions
-are met.
-
-\item \textbf{Cookies:} Cookies allow to name and retrieve individual
-flow entries and to use masks to delete modify/delete multiple entries. Our synchronization mechanism will implement controller-specific cookies.
-
-\item \textbf{Groups:} Groups of action buckets can be defined in a special group table.
-Each group has a unique id and any attempt to create a group with existing id or to delete un existing group fails (and can abort a bundle in process).
-\end{enumerate}
-
-\begin{figure*}[t]
-\centering
-\includegraphics[width=0.8\columnwidth]{loadbal.pdf}~~
-\includegraphics[width=0.8\columnwidth]{pic.pdf}\\
-\caption{\emph{Left:} Load-balancing gone wrong. \emph{Right:} Illustration of transactional abstraction which  allows
-controllers to modify switch configurations in complex ways,
-and
-depending on other flow spaces.}\label{fig:illu}
-\end{figure*}
-
-\section{Limitations and Motivation}\label{sec:motivation}
-
-Before presenting our transactional interface and implementation,
-let us first discuss the limitations of the Openflow features discussed above.
-At first sight, the bundle feature seems to provide everything
-we need: it defines a sequence of read and write operations
- on the switch state which are executed atomically.
- However, such bundle transactions come with many
- limitations. A bundle transaction does not allow
- to change the switch state in any non-trivial way,
- as the read and write actions in the bundle do
- not support any conditional modifications or computations which
- depend on the read results. By leveraging
- the check overlap feature, it is at least possible
- to abort transactions depending on certain conditions
- in the overlapping
- flow spaces. However, in many cases, for example
- a load-balancing application,
- it can be interesting to define more general
- dependencies, which depend on, from a logical perspective,
- seemingly
- completely independent flow spaces,
- which however share the underlying physical network.
-In other words, the bundle feature only allows us to
-define ``one-shot transactions'',
-transactions which do not support any further controller interactions.
-Given
-the limited computational options the Openflow interface of a switch offers,
-it is difficult to implement more complex behavior.
-
-\textbf{Example 1.} To give an example, let us consider a controller application in charge
-of \emph{load-balancing}. The objective of the load-balancer
-is to distribute flows equally across links or servers.
-If multiple controllers concurrently and independently
-update forwarding rules, they may unvoluntarily
-unbalance the flow allocation. Note that this
-is true independently of whether the flow spaces
-affected by the respective controllers overlap or not.
-More concretely, assume that two controllers concurrently
-decide to forward traffic $1*$ and $0*$ to the same server.
-Flow space overlap checks cannot be used to detect
-such conflicts.
-FIXME: make this example better
-
-\textbf{Example 2.} To give another example,
-consider the problem of modifying previously
-installed forwarding policies.
-While a bundle allows to conditionally add new rules or remove
-old rules (and, using the flow space overlap
-check, to abort the transaction if potentially conflicting
-rules were already there), it cannot be used
-to modify rules in non-trivial ways.
-In other words, the bundle and overlap check only helps
-for the conditional creation of rules, not their modification.
-
-\stefan{It would be cool if we could make these examples more concrete,
-at least one, and get back to it later again!}
-
-%\section{The Container Object}\label{sec:vision}
-\section{-NEW- Transactional Synchronization}\label{sec:main}
-
-This section describes how to introduce more powerful transactions.
-Our implementation is based on two new synchronization objects,
-allowing to coordinate controllers using identifiers.
-
-At a high level, we export a
-transactional interface which allows each controller $i\in[N]$ ($N\geq
-2$) to perform multiple
-operations on the configuration of an OpenFlow switch as an atomic
-\emph{transaction}, inspired by the transactional-memory paradigm~\cite{stm-st95,tm-book}. A transaction may \emph{commit}, in
-which case it appears as executed sequentially, or \emph{abort}, in
-which none of its operations takes effect.
-
-\subsection{Synchronization Objects}\label{sec:t-if}
-
-We assume transactional access to  two kinds of objects: a read-write \emph{memory} abstraction
-object, which we later use for storing the current used policy identifier, and a
-\emph{resource-claimer} object, provides claiming a resource id (e.g. policy \emph{identifier}) in a
-given (bounded or unbounded) set.
-In theory one can use the memory abstraction to implement the resource claimer, however we found it far more simple and efficient to consider the resource claimer as base object.
-%
-
-A memory can be accessed with the following sequential
-semantics:
-
-\begin{itemize}
-\item by performing $\textit{Write}(addr,k)$,
-  $addr,k\in\Nat$, the value at memory offset $addr$ is set to $k$;
-
-\item $\textit{\compare}(addr,k)$, $addr,k\in\Nat$, 
-aborts current transaction if the value at memory offset $addr$ is not equal to $k$.
-%returns $\textit{true}$ if and
-%  only if $k$ is not currently claimed by any controller; if
-%  $\textit{false}$ is returned, we say that the check operation
-%  \emph{fails} (which causes an abort of the transaction invoking it).
-
-\end{itemize}
-
-Note that reading the memory can be done by just reading the switch configuration however such action doesn't affect the execution of the transaction (as $\textit{\compare}$ does).
-
-An immediate extension of the memory interface is the \emph{compare-and-set} (CAS) operation that takes two integer parameters
-$\textit{old}$ and $\textit{new}$, reads the register and, if the
-value of it is $\textit{old}$, replaces it with $\textit{new}$ and
-returns $\textit{true}$; otherwise, $\textit{false}$ is returned and
-we say that the CAS operation \emph{fails} (which causes an abort if
-the transaction invoking it).
-
-A resource-claimer object implements a weak type of locking, exports operations \emph{claim},
-\emph{unclaim} and \emph{\claimcheck} with the following sequential
-semantics:
-%The container synchronization object introduced in this paper
-%is essentially a transactional memory abstraction which can be implemented
-%in standard Openflow switches.
-%Before presenting the technical details in the next section,%
-%we give a brief overview.
-
-%The postulated container object can store an arbitrary number of \emph{items}.
-%As we will see, these items correspond to flow entries on the Openflow switch.
-%The same item may even exist multiple times in the container (i.e., the container
-%is a multi-set).
-%The container
-%provides two main functions:
-\begin{itemize}
-\item by performing $\textit{claim}(k)$, $i\in[N]$,
-  $k\in\Nat$, controller $i$ \emph{claims} 
-  identifier $k$;
-
-\item by performing $\textit{unclaim}(k)$, $i\in[N]$, $k\in\Nat$, controller $i$
-  \emph{unclaim} identifier $k$;
-
-\item $\textit{\claimcheck}(k)$, $k\in\Nat$, 
-abort current transaction if $k$ is currently claimed by any controller.
-%returns $\textit{true}$ if and
-%  only if $k$ is not currently claimed by any controller; if
-%  $\textit{false}$ is returned, we say that the check operation
-%  \emph{fails} (which causes an abort of the transaction invoking it).
-
-\end{itemize}
-
-\subsection{Interface}\label{sec:t-if}
-
-\liron{this subsection doesn't flow well for me}
-
-We assume that transactions may comprise operations on the flow table
-of an SDN switch. Our transactional abstraction exports the
-$\textbf{atomic}\{\;\ldots\;\}$ construct
-containing a sequence of transactional
-operations and may \emph{commit} and return responses to all its operations
-or \emph{abort} (so that none of its operation takes effect).  
-We guarantee that in every execution, all committed
-transactions constitute a \emph{legal} sequential
-order~\cite{Pap79-serial}, i.e., every transaction appears as executed
-sequentially. On the progress side, we ensure that a
-transaction $T$ is only allowed to abort if either another 
-transaction updated a flow table entry read by $T$  
-or if one of $T$'s operations (\textit{check} or CAS) fails.
-We employ the \emph{deferred-update} semantics:  
-no transaction can affect other transactions before it commits.
-In particular, if a controller crashes before committing its
-transactions, no future transaction can be affected.  
-%Note that
-%, assuming such a transactional memory,  
-%a $\emph{contains}(x)$ operation can be implemented as a transaction
-%composed of $\textit{cond-add}(x)$ followed by $\textit{cond-delete}(x)$.  
-%A \emph{set} abstraction is a special instance of a multi-set that is
-%only modified with \textit{cond-add} and \textit{cond-delete}
-%operations (this way the set only contains at most one copy of a given
-%item).
-
-\subsection{Implementation in Openflow}\label{sec:t-impl}
-
-We now describe how to implement our transactional 
-abstractions using standard OpenFlow features.
-
-\subsubsection{A resource claimer}
-%features , including
-%\emph{bundling}, \emph{conditional application}, and \emph{groups}.
-%[[PK let us leave for the moment
-%We will describe later how to implement some of these features
-%in older Openflow versions where they are not
-%supported. \textbf{TODO}
-%]]
-Algorithms~\ref{alg:claim}-\ref{alg:cas} describe implementations of  
-transactional operations  $\textit{claim}(x)$, $\textit{\claimcheck}(x)$, $\textit{unclaim}(x)$, $\textit{write}(x)$,  $\textit{\compare}(x)$
-and $\textit{\cas}(x)$, respectively. 
-Algorithm~\ref{alg:commit} describes the procedure executed by
-a transaction at \emph{commit time}.
-
-We describe each operation as a message factory that returns OpenFlow commands that should be sent to the switch as part of the same transaction.
-We assume that every transaction has a unique identifier  used to start a \emph{bundle}. 
-All messages generated during a transaction invocation, are wrapped by bundle messages carrying the bundle's id and followed by bundle commit message.
-Messages with same bundle id are considered to belong to same bundle and are performed atomically.
-
-%The OpenFlow messages that we use (besides )
-
-Our implementation is based on OpenFlow flow entries modification command, the FlowMod message that we explain in Section \ref{sec:background}.
-We use the CHECK\_OVERLAP optional flag to make the transaction execution dependent on switch state.
-
-We consider the match part of an entry to be a ternary string of bounded length (represented by binary value and mask) which can be applied to any set of supported packet header fields.
-In our implementation the $64$bit metedata field alone is suffice.
-Moreover we consider the action part of an entry to be an integer, which in practice can be implemented by a Set-Metadata instruction where the written value is that integer.
-
- As can be seen in Algorithms \ref{alg:claim}-\ref{alg:unclaim}, for the resource-claimer, 
-  the match part of the added flow entry represents the claimed id (id concatenated twice).
-   In addition the mask part of the match (controller id concatenated with all-ones) varies from one controller to another in order to allow multiple controllers to claim the same number without overriding the same flow entry.
-    We can see that for any two resource ids $x_1$ and $x_2$ and controller ids $c_1\neq c_2$ the match patterns $m_1=(x_1\concat x_1, c_1\concat all-1)$ and $ m_2=(x_2\concat x_2, c_2\concat all-1)$ are not equal even if $x_1=x_2$ therefore they can both be used to claim and unclaim without affecting each other.
-
-In order to check if a number is claimed we try to create the corresponding entry while setting the
-flag \textsf{CHECK\_OVERLAP} and using different action ($2$ instead of $1$), thereby inflicting a failure in case an entry with overlapping match value exists.
-If the check succeeds we delete the entry.
-Considering $m_1$ and $m_2$ defined above and their respective actions $a_1=x_1\concat c1$ and $a_2=x_2\concat c2$,  they overlap with different actions iff $x_1=x_2$.
-Therefore if one is used in \claimcheck after the other used in claim, a conflict is detected iff $x_1=x_2$, as expected.
-
-Note that for switch space reasons, checking for claims contains also a deletion of the added tester flow entry but it has no affect on claims and checks of other controllers and therefore the deletion command doesn't have to be sent as part of a bundle.
-This makes the \textsf{\claimcheck} operation atomic by nature, similar to the \textsf{claim} and \textsf{unclaim}, as being implemented by essentially one command.
-
-%A container contains items: they represent the flow table entries (i.e., a match-action rules, plus e.g., counters belonging to the rule).
-%Items can be added and removed atomically and in bundles (e.g., to update the switch with
-%a new policy). Recall that we want to add items only if they do not exist
-%yet; otherwise the transaction should abort without taking effect.
-%And similarly for removes: transactions should not take effect if the corresponding
-%update contains rules which currently do not exist at the switch.
-
-\begin{algorithm}[H]
-    \caption{$\textit{claim}(x)$}
-    \label{alg:claim}
-    \begin{algorithmic}[1]
-    \Require \emph{self} as the calling controller id.
-%    \Ensure
-    		\State $value \gets x\concat x$
-    		\State $mask \gets self\concat all-1$
-	    	\State $match \gets (value,mask)$
-    		%\State $cookie \gets x\concat self$
-    		\State $action \gets 1$
-    		\State $flag \gets 0$
-    		\State $cmd\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
-			\Return $cmd$
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[H]
-    \caption{$\textit{\claimcheck}(x)$}
-    \label{alg:check}
-    \begin{algorithmic}[1]
-    \Require \emph{self} as the calling controller id.
-%    \Ensure
-    		\State $value \gets x\concat x$
-    		\State $mask \gets self\concat all-1$
-    		\State $match \gets (value,mask)$
-    		%\State $cookie \gets x\concat self$
-    		\State $action \gets 2$
-    		\State $flag \gets \textsf{OFPFF\_CHECK\_OVERLAP}$
-    		\State $cmd1\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
-    		\State $cmd2\gets \textsc{FlowMod}(match, op = DELETE) $
-			\Return $cmd1$,$cmd2$
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[H]
-    \caption{$\textit{unclaim}(x)$}
-    \label{alg:unclaim}
-    \begin{algorithmic}[1]
-    \Require \emph{self} as the calling controller id.
-%    \Ensure
-    		\State $value \gets x\concat x$
-    		\State $mask \gets self\concat all-1$
-    		%\State $cookie \gets x\concat self$
-    		\State $match \gets (value,mask)$
-    		\State $cmd\gets \textsc{FlowMod}(match, op=DELETE) $
-
-			\Return $cmd$
-    \end{algorithmic}
-\end{algorithm}
-
-\subsubsection{A memory abstraction}
-
-We use similar techniques for implementing memory \memwrite and \compare 
- operations.
- As can be seen in Algorithms \ref{alg:write}-\ref{alg:compare}, 
-  the match part of the added flow entry is used as the memory address and the action as the written value.
-Values written to the same address will replace one another, following the command definition to replace old flow entry with a new one in case they share the same match parts (as described in Section \ref{sec:background}).
-
-In order to check if a value is written in a given address, we try to add it the same way as in the \memwrite operation  but we also set the \textsf{CHECK\_OVERLAP} flag.
-If the value is not there, our action differ from the action of the existing entry thereby inflicting a failure.
-In case the value is there, our flow entry replaces the existing flow entry with identical one.
-In addition we present in Algorithm \ref{alg:cas} a simple implementation of Compare and set (CAS) using the two last operations.
-
-\begin{algorithm}[h]
-    \caption{$\textit{write}(addr,k)$}
-    \label{alg:write}
-    \begin{algorithmic}[1]
-    %\Require  id $i$.
-%    \Ensure
-    		\State $value \gets addr$
-    		\State $mask \gets  all-1$
-    		\State $match \gets (value,mask)$
-    		%\State $cookie \gets x\concat self$
-    		\State $action \gets k$
-    		\State $flag \gets 0$
-    		\State $cmd\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
-			\Return $cmd$
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[h]
-    \caption{$\textit{\compare}(addr,k)$}
-    \label{alg:compare}
-    \begin{algorithmic}[1]
-    %\Require  id $i$.
-%    \Ensure
-    		\State $value \gets addr$
-    		\State $mask \gets  all-1$
-    		\State $match \gets (value,mask)$
-    		%\State $cookie \gets x\concat self$
-    		\State $action \gets k$
-    		\State $flag \gets \textsf{OFPFF\_CHECK\_OVERLAP}$
-    		\State $cmd\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
-			\Return $cmd$
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[h]
-    \caption{$\textit{\cas}(addr, old,new)$}
-    \label{alg:cas}
-    \begin{algorithmic}[1]
-    %\Require  id $i$.
-%    \Ensure
-
-    		\State $cmd1\gets \textsc{\compare}((addr, old) $
-    		\State $cmd2\gets \textsc{\memwrite}((addr, new) $
-			\Return $cmd1,cmd2$
-    \end{algorithmic}
-\end{algorithm}
-
-To commit the transaction,  the controller sends a \textsf{OFPT\_BUNDLE\_CONTROL} message with type
-\textsf{OFPBCT\_COMMIT\_REQUEST} and waits until a matching control
-message from the switch is received. When the switch receives the
-commit request it tries to atomically process all the buffered messages
-equipped with the corresponding bundle identifier and, if the
-processing is successful for all of them, returns  a \textsf{OFPT\_BUNDLE\_CONTROL} message with type
-\textsf{OFPBCT\_CLOSE\_REPLY}; otherwise it sends a specific error
-message  pointing out the aborted transactional operation.
-
-\begin{algorithm}[h]
-    \caption{$\textit{try-commit}()$}
-    \label{alg:commit}
-    \begin{algorithmic}[1]
-    \Require  current bundle id $bid$.
-		    \State $cmd \gets struct ofp_bundle_ctrl_msg$
-    		\State  $cmd.bundle\_id \gets bid$
-    		\State  $cmd.type \gets $ \textsf{OFPBCT\_COMMIT\_REQUEST} 
-    		\State  $cmd.flags \gets $ \textsf{OFPBF\_ATOMIC}
-    		\State send $cmd$
-		\Return $cmd$
-    \end{algorithmic}
-\end{algorithm}
-
-\section{Discussion of Use-Cases}\label{sec:apps}
-
-More powerful transactions can be useful in many scenarios.
-Let us revisit our examples from earlier: with
-transactions we can easily solve the load-balancing
-and rule modification problem.
-In the following we give two more examples of our approach.
-
-\subsection{Policy Updates}
-
-Our first example (Algorithm~\ref{alg:simple-update}) is a simple policy composition protocol that
-allows a set of  controllers to concurrently update switch  \emph{policies} , i.e., sets of
-effective rules in its flow tables, under the
-condition that these rules can be composed with the currently installed
-policy (the composition predicate is encoded in a user-specific
-$\textit{apply\_update}()$ function~\cite{cpc}.
-
-Every policy is associated with a \emph{policy identifier} 
-In the algorithm, the controller first reads the currently installed
-policy, then choose a new unique policy, checks whether the new
-policy and the current policy can be composed and, if so, replaces the
-current policy with the composition. If, meanwhile, a new
-policy (with a different identifier) has been installed, the update
-fails and takes no effect.
-
-Algorithm~\ref{alg:simple-update} use increasing numbers as policies
-ids and use them in order to make sure that the new policy, with id
-$x$, replaces the (what considered as) current policy, with id
-$x-1$. Considering the policies ids, we send the policy installation
-commands as part of a bundle that also tries to perform
-$CAS(x-1,x)$, thereby insuring that the commands are
-performed iff the current policy is still there and preventing other
-controllers to replace the new policy with another without observing it.
-
-For brevity, we omit explicitly putting the $\textbf{atomic}\{...\}$ construct
-around each of the transactional operation used in the algorithm.   
-Indeed, since \textit{claim} and \textit{unclaim} operations cannot
-cause the corresponding transaction, we can assume that all
-``singleton'' (consisting of one operation only)  transactions
-containing such operations always commit.
-
-Algorithm~\ref{alg:simple-update} allows policy identifiers to grow
-without bound. In fact, as was observed in~\cite{cpc}, the number of currently
-used policies can be made proportional to the number of concurrent
-controllers.
-
- It takes a set of rules $U$ (not yet a policy)and proceeds as follows: first, we seek to
- obtain a unique \emph{id}. FIXME: to be continued...
- \textbf{LS: I am not sure if I need to tell every step of the story or maybe it best to explain the main dif from the previous one similarly to what I just wrote next}.
-
-Algorithm~\ref{alg:update} claims the current policy id in order to prevent someone of using it with different policy in the future (during the current update) and in the bundle used to update the policy it also checks that the new policy id that it about to use (chosen from a finite set of ids) wasn't claimed by someone else.
-
-\hide{
-We compute our new suggested policy by applying the update requests on top of current policy, supporting any kind of requests and policies.
-Then we make a transaction (using the bundle feature) to atomically check that our policy id is not blocked by anyone else, to change the current policy id to ours (an action that would fail if the current policy id is no longer what we are counting of) and to actually configure our new policy.
-
-If one of the actions in the transaction fails we try again.
-There is no progress guaranty for each controller but there is one for the whole system - at least one of the controller will succeed in fulfilling its update requirements.
-}
-
-\textbf{TODO: fix the pseudocode, make the use of afdd/delete
-  operations uniform, make all switch operations transactional}
-
-\begin{algorithm}[t]
-    \caption{Policy update with only CAS}
-    \label{alg:simple-update}
-    \begin{algorithmic}[1]
-    \Require update requirements $U$, a predefined (memory) address for  policy id: $pid\_addr$
-    \Ensure installed policy is consistent with previous one
- 		%\Do
- 		\Repeat
- 			\State \textbf{read} switch configuration
- 			\State $pid\gets$ current policy id 
- 			\State $cur\_policy\gets$ current policy
- 			\State $new\_policy\gets apply\_update(cur\_policy,U)$
- 			\startTxn
-	 			\State $CAS(pid\_addr,pid,pid+1)$
-	 			\State add $new\_policy$ rules %conditioned on $metadata=my\_a$
-	 			%\State update policy selector to $my\_a$
- 			\endTxn
-     		%\State $res \gets $ commit transaction
-     	\Until{$res\neq\texttt{abort}$}
-     	%\doWhile{$res=False$}
-%   \vspace{3mm}
-			\Return
-
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[t]
-    \caption{Advanced policy update}
-    \label{alg:update}
-    \begin{algorithmic}[1]
-    \Require update requirements $U$, a predefined (memory) address for  policy id: $pid\_addr$
-    \Ensure installed policy is consistent with previous one
- 		%\Do
- 		\Repeat
- 			\State \textbf{read} switch configuration
- 			\State $my\_id\gets$ an unused \emph{id} 
- 			\State $pid\gets$ current policy id 
- 			\State $claim(pid)$
- 			\State \textbf{read} switch configuration
- 			\If {$pid\notin Cur$}
-	 			\State $unclaim(pid)$
- 				\State {\bf continue} (\textbf{restart} loop)
- 			\EndIf
- 			\State $cur\_policy\gets$ current policy
- 			\State $new\_policy\gets apply\_update(cur\_policy,U)$
- 			\startTxn
- 				\State $\claimcheck(my\_id)$
-	 			\State $CAS(pid\_addr, pid,my\_id)$
-	 			\State \textbf{add} $new\_policy$ rules %conditioned on $metadata=my\_a$
-	 			%\State update policy selector to $my\_a$
- 			\endTxn
-	 		\State $unclaim(pid)$
-     		%\State $res \gets $ commit transaction
-     	\Until{$res\neq\texttt{abort}$}
-     	%\doWhile{$res=False$}
-%   \vspace{3mm}
-			\Return
-
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[t]
-    \caption{Policy composition without bundle}
-    \label{alg:wobundle}
-    \begin{algorithmic}[1]
-    \Require policy $P$ (a set of FlowMod entry creation commands), three consecutive flow tables ids (per controller) $T_s,T_0,T_1$, active table index $active\in\{0,1\}$.
-    %\Ensure installed policy is consistent with previous one
-    \State $unactive \gets active + 1 \mod{2}$
-    \For {$cmd \in P$}
-	    \State $cmd.action\gets Reencode(cmd.action)$
-	    \State $cmd.table\_id\gets T_{unactive}$
-	    \State send $cmd$
-    \EndFor
-    \State modify the action of $T_s$ entry to goto $T_{unactive}$
-    \State clear table $T_{active}$
-    \State $active \gets unactive$
-	\Return
-
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[t]
-    \caption{Update with pipeline}
-    \label{alg:pipeline}
-    \begin{algorithmic}[1]
-    \Require update requirements $U$, array $ARR$.
-    \Ensure installed policy is consistent with previous one
-			\Return
-
-    \end{algorithmic}
-\end{algorithm}
-
-\subsection{Consensus and Replicated State Machines}
-
-It is straightforward to implement consensus among the controllers
-using just a single register that can be accessed with an atomic CAS
-operation~\cite{Her91}. In our transactional setting, the controllers
-can use CAS to replace the default value of such a register 
-with their consensus input values (e.g., the identifiers of the
-policies that should be applied first). If the corresponding
-singleton transaction commits, the controller consideres itself 
-the \emph{winner} and decides on its input, if
-it aborts (the controller fails in the CAS) it remains to read the register to get the input of the
-winner to decide.
-
-By using multiple CAS invocations, the controllers can implement any
-generic replicated state machine~\cite{Her91,Lam98}, which will be instrumental in the
-next section.
-
-%\paragraph{One-Shot Update} FIXME: desribe how to implement one-shot update
-
-\paragraph{STN}
-
-\textbf{TODO}
-
-\section{In-Band Composition}\label{sec:compo}
-
-We conclude our technical contribution by showing that sometimes,
-powerful synchronization mechanisms can be implemented
-in-band \emph{even without bundles}.
-Concretely, we show that sometimes it is even possible
-to automatically compose updates originating from different controllers,
-in-band and just using Openflow.
-
-To illustrate this point, we consider a simple example
-with two controllers concurrently issuing updates to a single
-given switch.
-
-FIXME: Explain, maybe even with an example?
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=0.8\columnwidth]{composition.pdf}\\
-\caption{Composition.}\label{fig:illu}
-\end{figure}
-
 \begin{comment}
-
-\section{Transactions in Openflow}\label{sec:sync}
-
-Multi-write and read-modify transactions.
-
-\subsection{Dealing with multiple updates}
-
-TBD - adding the notion of versions and cleanup/GC. probably no good way to solve infinite versions... maybe check literature on shared memory and multiple writes.
-
-\subsection{Improving the scheme with check-overlap}\label{sec:todo}
-
-Till now we considered a model that allows multiple updates to match-action entries.
-Here we extends the model with actions create-entry and delete-entry that may failed and abort the whole transaction depending on the existence on non-existence of specific entries prior to the action.
-%TBD - much less resources. easily solving the infinite versions issue.
-
-Next we show how such actions can be used to implement consensus with much less resources. We use entry id as a property used to reference specific entry.
-
-\section{Applications and ramifications}\label{sec:application}
-
-\section{The Underlying Theory}\label{sec:background}
-
-idea: multi-write consensus, a not well-known result!
-
-In order to solve consensus we suggest to use the atomic multiple entries update capability (which is discussed in Openflow standard).
-A shared memory system that supports atomic write to (enough) multiple locations is known to allow the implementation of consensus objects.
-For example consider the following implementation utilizing $n^2$ memory locations, $M_{i,j}$, where $i,j\in[n]$, and the suggested values $\{v_i\}_{i\in [n]}$.
-All locations are initialized to zero.
-
-In order to participate, a server i, writes its suggested value $V_i$ and atomically rewrites row i ($M_{i,*}$) with value "1" and column i ($M_{*,i}$) with value "2".
-After the atomic write, a server $i$ first reads the diagonal ($M_{j,j}$ for $j\in [n]$), and consider all observed servers as candidate set S.
-Then server i reads all candidates intersections (locations $M_{j,k}$ where $j,k\in S$.
-And computes the winner server id w that was overwritten by all other candidates, i.e.
-such that $M_{w,k}=2$ for all $k\in S {w}$.
-The consensus value is $v_w$.
-Note that a snapshot could replace the two reading phases.
-
-\section{Basic One-Touch Mechanism}\label{sec:realization}
-
-We address the problem of committing consistent policies to SDN switch concurrently by multiple controllers.
-We show that assuming minimalistic control protocol, SDN switch configuration space can be used to implement a consensus object thereby allowing controllers to agree on each next policy update.
-
-\subsection{SDN switch configuration space as shared memory}
-
-Our first observation is that a match-action configuration which is accessed by multiple controllers can be considered as shared memory.
-Note however that standard memory map and index/offset to a word, while Flow Table entries are not necessarily indexed.
-However, we assume that each flow entry has some property that allows the operator to specifically reference and update it, for example Openflow entry's cookie.
-In cases where this property value can't be determined by the operator, we can use other variable entry properties to store the index, for example we can represent memory value $v$ in offset $i$ by an match-action entry whose match is $in\_port = i$ and action is "forward port $x$".
-%To overcome this we can use the match part of an entry as the index and the action part as the word value.
-
-Going back to the policy update problem, the consensus scheme allow all servers to decide on next policy by using consensus value as policy identifier, and writing the policy in advance.
-However this scheme doesn't actually apply the policy on incoming packets and it remains to configure the switch according to the next policy once it is decided.
-Making this scheme fail safe (handling failure of winning server) can be ensured by allowing any server to reconfigure the switch.
-
-\subsection{ One-Touch}
-
-In some scenarios it may be desirable to shorten the policy update time by avoiding the consensus reading and policy configuration step.
-We define this as the one touch policy update, namely we require each server to contact the switch only once in order to (try and) update the policy, in other words participating in reaching consensus and applying it with one atomic write.
-Solving this problem also has significance in understanding the strength of the SDN switch computation model.
-
-The main idea of our solution is to adjust the previous policy update scheme in a way that performs consensus validation phase during every packet processing thereby requiring the servers only to make the first phase of the consensus - atomically writing the matrix.
-In order to allow the packets to validate the consensus we make the following observation: while servers can be abstracted as memory writers, the packets processed by the switch are the readers.
-Openflow standard define an atomic write transaction as being atomic in relation to the processed packets thereby making packet processing a multiple read transaction but one read location per table is allowed.
-This means that if each memory location used in the consensus scheme will be stored in different table then each packet will be able to read all of them.
-
-However, reading each memory location is not enough, as the scheme requires to make a validation that involves multiple values.
-We utilize the metadata field in order to store all read values, in more details, each memory value $M_{i,j}$ is stored in offset i+j*n in the metadata field.
-Once all values are in the metadata field, we can detect the consensus winner by trying different matches on the metadata.
-Although we can have one match entry per possible matrix state, this solution requires exponential number of entries.
-A better way which we describe next requires only $O(n^2)$ entries utilizing n tables.
-
-Our compact validation use additional temp variable, $temp_winner$, to hold the current best consensus candidate.
-Each of the n validating tables checks different "column" in the matrix and update $temp_winner$ to keep track of the winner so far (after examining a prefix sub set of the columns).
-The entry t in table k $(0<=k<n)$  matches the case where current winner=t and location $(t,k)$ in the matrix (packet header) equals 1 and location (k,k) is non-zero.
-After all validating tables are processed, $temp_winner$ stores the id of the winning server which can be used to match on its policy rules (filtering out other policies) or to use the goto table command to jump to a designated policy table of the server.
-
-\section{The Practical Implementation}\label{sec:extension}
-
-\end{comment}
-
-\section{Related Work}\label{sec:relwork}
-
-There exists a wide consensus that SDN control planes need to be distributed.~\cite{onos,onix,elasticon}
-Onix~\cite{onix} is one of the earliest proposals, and introduced the
-the notion of Network Information Base (NIB), abstracting network state
-distribution from control logic, but
-requiring mechanisms for the detection and resolution of conflicts.
-Also spatially distributed control planes to improve scalability and
-latency have been studied intensively
-in the literature~\cite{kandoo,ctrl-place,hotsdn13loc}
-proposes an elastic distributed controller architecture.
-
-Operating and updating SDNs consistently is already non-trivial
-from the perspective of a single controller. Reitblatt et al.~\cite{network-update}
-introduced the notion of
-per-packet consistency and proposed a 2-phase update protocol based on tagging.
-%, and described the \emph{two-phase update} technique, also used in our algorithms.
-Mahajan and Wattenhofer~\cite{roger-hotnets} considered weaker transient
-consistency guarantees, and proposed more efficient network update algorithms
-accordingly. Ludwig et al.~~\cite{hotnets14update} studied algorithms for secure
-network updates where packets are forced to traverse certain waypoints or
-middleboxes. Ghorbani et al.~~\cite{correct-virt} recently argued for the design
-of network update algorithms that provide even stronger consistency guarantees.
-
-To the best of our knowledge, the only paper explicitly addressing the consistent
-network update problem from a distributed-controllers perspective is STN~\cite{stn}:
-STN provides a transactional interface offering all-or-nothing semantics and serializability
-(the ``holy grail'' of safety properties), allowing
-controllers to install policies in a conflict-free manner; if some controllers fail,
-other controllers can take over. STN~\cite{stn} is based on
-atomic read-modify-write primitives provided by the switches, but
-while the availability of such primitives is postulated, no implementation is
-described. In this paper, we provide this missing link, and also show that other useful and
-more powerful synchronization primitives can be implemented. 
-Also, some STN implementations in~\cite{stn} assume that the control
-plane maintains a certain amount of synchrony that allows controllers
-to solve consensus or implement a replicated state machine.     
-In this paper, we describe an OpenFlow-based consensus protocol that
-can be used for this purpose.
-
-Our work is also closely related to the recent work on providing more high-level
-abstractions and programming languages for SDNs, such as Frenetic~\cite{frenetic},
-also enabling policy composition~\cite{pyretic}. Our work complements these works
-on programming abstractions
-by initiating the discussion of concurrency objects and abstractions.
-
-The transactional interface we describe in this paper is inspired by the
-work on speculative concurrency control using software transactional memory
-(STM)~\cite{stm-st95,tm-book}. 
-%[[PK I think this is enough]]
-
-Our work also contributes to the ongoing discussion of what can be implemented
-in-band in Openflow~\cite{compute,reclaim}, by assuming the relevant perspective of distributed
-synchronization.
-
-\hide{
-In terms of distributed computing: TODO PETR: maybe modify a bit the text below
-and also write more about multi-writer consensus etc.?
-
-\noindent\textbf{Distributed Computing.}
-There is a long tradition of defining correctness of a concurrent system via
-an equivalence to a sequential one~\cite{Pap79-serial,Lam79,HW90}.  The notion
-of sequentially composable histories is reminiscent of
-linearizability~\cite{HW90}, where a history of operations concurrently
-applied by a collection of processes is equivalent to a history in which the
-operations are in a sequential order, respecting their real-time precedence.
-In contrast, our sequentially composable histories impose requirements not
-only on high-level invocations and responses, but also on the way the traffic
-is processed. We require that the committed policies constitute a
-conflict-free sequential history, but, additionally,  we expect that each
-\emph{path} witnesses only a prefix of this history, consisting of all
-requests that were committed before the path was initiated.
-%
-The transactional interface exported by the CPC abstraction is inspired by the
-work on speculative concurrency control using software transactional memory
-(STM)~\cite{stm-st95}.
-%, thus the  term \emph{software transactional networking}.
-Our interface is however intended to model realistic network
-management operations, which makes it simpler than recent
-dynamic STM models~\cite{dstm}.
-%In particular, the sets of rules to be installed by a policy update do not depend on the state of the
-%network.
-
-Extending the interface to dynamic policies that adapt their behavior based on
-the current network state sounds like a promising research direction.  On the
-other hand, our criterion of sequential composability is more complex than
-traditional STM correctness properties in that it imposes restrictions not only
-on the high-level interface exported to the control plane, but also on the
-paths taken by the data-plane packets.
-Also, we assumed that controllers are subject to failures, which is usually not
-assumed by STM implementations.
-}
-
-\section{Conclusion}\label{sec:conclusion}
-
-While much progress has been made over the
-last years in the development of more intuitive programming
-languages for SDNs (e.g., Frenetic, Pyretic, Merlin), the
-question of which synchronization abstractions to provide
-has received much less attention.
-%
-In this paper, we describe transactional abstractions that 
-can be implemented using standard Openflow features.
-%
-Our work complements existing research on the design of
-distributed control planes, and also provides some missing links,
-e.g., for the transactional approach taken by STN~\cite{stn}.
-We also hope that our work can nourish the discussion of
-which synchronization objects we can and want to implement
-in an SDN.
-
-{
-\bibliographystyle{abbrv}
-\bibliography{references}  % main.bib is the name of the Bibliography in this case
-}
-
-\begin{appendix}
-
 \section{Implementing CAS with FlowMod commands}\label{sec:without-groups}

### 2015-03-29 inband-sync.tex
stefan sync

@@
 \algblockdefx[Bundle]{startTxn}{endTxn} %
-[0]{\textbf{atomic}\{} %
+[0]{\textbf{bundle}\{} %
 [1][res]{ \} $\rightarrow #1$}
@@
 can include complex notions of conflicts (beyond flow overlap checks),
-as well as computations and controller interactions. We discuss different use cases
+as well as computations and multiple controller interactions. We discuss different use cases
 for our interface (consistent policy installation, load-balancing, consensus),
@@
 concurrently~\cite{participatory,stn}.
-All this, however, comes with the need for \emph{synchronizing}
-the distributed controllers:
-the network should be manipulated \emph{consistently}~\cite{cpc}.
-This task is nontrivial~\cite{cap-theorem}.
 
@@
 control application).
+\emph{Synchronizing}
+the distributed controllers and manipulating the network state \emph{consistently}~\cite{cpc}.
+are non-trivial tasks~\cite{cap-theorem}.
 
@@
 
-This paper investigates the design of a transactional interface for software-defined
-networks. In particular, we argue that the switch configuration state can be seen
-as a shared memory which is modified concurrently by multiple controllers.
 Being able to provide more complex and atomic sequences of transactions over this configuration
 space can be appealing to resolve synchronization problems.
+This paper investigates the implementation of a transactional interface 
+(\`{a} la STN~\cite{stn}) modified concurrently by multiple controllers.
 
 \paragraph{Our Contributions}
-This paper introduces a very powerful transactional interface
- for software-defined networks. This interface is implemented
- in-band, and
- using Openflow~\cite{of-spec} (the standard SDN protocol today) features only.
-Indeed, we find that Openflow's \emph{bundle} or \emph{rule overlap check}
-features can serve as synchronization primitives.
-However, without further mechanisms, their use is limited:
+This paper shows how to implement a very powerful transactional interface
+ for software-defined networks. 
+ Concretely, our implementation is based on
+ standard Openflow~\cite{of-spec} primitives, which by themselves are very limited.
+For example, we find that while Openflow's \emph{bundle} or \emph{rule overlap check}
+features can serve as synchronization primitives,
+without further mechanisms, their use is very limited:
 with a bundle alone it is not possible to react to
@@
 update conflicts defined over non-overlapping flow spaces.
-
-In contrast, the transactional interface presented in this paper supports very general atomic
+Moreover, we show that many powerful synchronous operations
+can also be implemented without bundles.
+
+The transactional interface presented in this paper supports very general atomic
 transactions over the switch configuration space, similar to classic transactional memory~\cite{stm-st95,tm-book}.
@@
 The synchronization mechanisms presented in this paper
-are based on Openflow~\cite{of-spec}, the standard SDN protocol today. In this section,
-we provide the necessary background.
+are based on Openflow~\cite{of-spec}, the standard SDN protocol today.
+% In this section,
+%we provide the necessary background.
 
@@
 Openflow
-is based on a match-action concept: Openflow switches store
+follows a match-action concept: Openflow switches store
 rules (installed by the controller) consisting of a match and an
@@
 
-\textbf{The FlowMod Command.}
+\subsection{The FlowMod Command}
+
 \emph{FlowMod} is the main messages used in Openflow 
@@
 
-\textbf{Bundling.}
+\subsection{Bundling}
+
 The Openflow bundle can be seen as a ``mini-transaction'':
@@
 
+In the remainder of this paper, we will use the notation 
+$\textbf{bundle}\{\;\ldots\;\}$ to refer to represent a
+sequence of commands in a bundle.
+\stefan{liron: i think the following needs more details $:-)$ :} A bundle will always have the following structure:
+it first contains a \emph{send message} to start bundle,
+followed by a sequence of messages wrapped in bundle and bundle ID, 
+and finally end bundle control message for commit.
+A bundle always returns a binary result: \emph{error} or \emph{success}.
+
+To commit the transaction,  the controller sends a \textsf{OFPT\_BUNDLE\_CONTROL} message with type
+\textsf{OFPBCT\_COMMIT\_REQUEST} and waits until a matching control
+message from the switch is received. When the switch receives the
+commit request it tries to atomically process all the buffered messages
+equipped with the corresponding bundle identifier and, if the
+processing is successful for all of them, returns  a \textsf{OFPT\_BUNDLE\_CONTROL} message with type
+\textsf{OFPBCT\_CLOSE\_REPLY}; otherwise it sends a specific error
+message  pointing out the aborted transactional operation.
+
+Algorithm~\ref{alg:commit} describes how to commit
+the bundle. 
+\begin{algorithm}[h]
+    \caption{$\textit{try-commit}()$}
+    \label{alg:commit}
+    \begin{algorithmic}[1]
+    \Require  current bundle id $bid$.
+		    \State $cmd \gets struct ofp_bundle_ctrl_msg$
+    		\State  $cmd.bundle\_id \gets bid$
+    		\State  $cmd.type \gets $ \textsf{OFPBCT\_COMMIT\_REQUEST}
+    		\State  $cmd.flags \gets $ \textsf{OFPBF\_ATOMIC}
+    		\State send $cmd$
+		\Return $cmd$
+    \end{algorithmic}
+\end{algorithm}
+
 %\textbf{Cookies:} Cookies allow to name and retrieve individual
@@
 and
-depending on other flow spaces.}\label{fig:illu}
+depending on other flow spaces. Colors indicate different actions.}\label{fig:illu}
 \end{figure*}
 
+Finally, we remark that our solution is only based on flow tables
+and does not make use
+of group tables. 
+Moreover, we do not make use of cookies: rules
+are always identified by exact matches.
+
 \section{Limitations and Motivation}\label{sec:motivation}
 
-what is the poewr of bundling flowmods:
-- no cookies: delete entries now by specifying exact match part
-
-- alterantive impelmentation without: group is bounded resources, can have own table etc. dont have many controlelrs anyway so not much is needed; also maybe other policies need groups
+\stefan{liron, maybe we can discuss this section together: we need some more ideas}
 
@@
  on the switch state which are executed atomically.
- However, such bundle transactions come with many
- limitations. A bundle transaction does not allow
+ However, such bundle transactions come with several
+ limitations, as discussed in the following. Moreover,
+ as we will see, our mechanism also does not \emph{need}
+ the bundle feature.
+
+A bundle does not allow
  to change the switch state in any non-trivial way,
@@
 %\section{The Container Object}\label{sec:vision}
-\section{-NEW- Transactional Synchronization}\label{sec:main}
+\section{Transactional Synchronization}\label{sec:main}
 
@@
 allowing to coordinate controllers using identifiers.
-
 At a high level, we export a
@@
 
-We assume transactional access to  two kinds of objects: a read-write \emph{memory} abstraction
+We implement transactional access to  two kinds of objects: a read-write \emph{memory} abstraction
 object, which we later use for storing the current used policy identifier, and a
@@
 given (bounded or unbounded) set.
-In theory one can use the memory abstraction to implement the resource claimer, however we found it far more simple and efficient to consider the resource claimer as base object.
+In theory one can use the memory abstraction to implement the resource claimer, 
+however we found it far more simple and efficient to consider the resource claimer as base object.
 %
@@
 \begin{itemize}
-\item by performing $\textit{Write}(addr,k)$,
-  $addr,k\in\Nat$, the value at memory offset $addr$ is set to $k$;
+\item By performing $\textit{Write}(addr,k)$,
+  $addr,k\in\Nat$, the value at memory offset $addr$ is set to $k$. For
+  example, the offset $addr$ identifies the policy (identifier), 
+  but our approach is more general.
 
@@
 Note that reading the memory can be done by just reading the switch configuration however such action doesn't affect the execution of the transaction (as $\textit{\compare}$ does).
-
-An immediate extension of the memory interface is the \emph{compare-and-set} (CAS) operation that takes two integer parameters
+As we will see, this abstraction can be implemented
+without using the notion of bundles.
+
+An immediate consequence of this memory interface is 
+that we can implement a \emph{compare-and-set} (CAS) operation that takes two integer parameters
 $\textit{old}$ and $\textit{new}$, reads the register and, if the
@@
 
+Again, none of these individual operations 
+requires bundles.
+
+\begin{comment}
 \subsection{Interface}\label{sec:t-if}
@@
 of an SDN switch. Our transactional abstraction exports the
-$\textbf{atomic}\{\;\ldots\;\}$ construct
+$\textbf{bundle}\{\;\ldots\;\}$ construct
 containing a sequence of transactional
@@
 
-\subsection{Implementation in Openflow}\label{sec:t-impl}
+\end{comment}
+
+\subsection{Implementation}\label{sec:t-impl}
 
@@
 
-To commit the transaction,  the controller sends a \textsf{OFPT\_BUNDLE\_CONTROL} message with type
-\textsf{OFPBCT\_COMMIT\_REQUEST} and waits until a matching control
-message from the switch is received. When the switch receives the
-commit request it tries to atomically process all the buffered messages
-equipped with the corresponding bundle identifier and, if the
-processing is successful for all of them, returns  a \textsf{OFPT\_BUNDLE\_CONTROL} message with type
-\textsf{OFPBCT\_CLOSE\_REPLY}; otherwise it sends a specific error
-message  pointing out the aborted transactional operation.
-
-\begin{algorithm}[h]
-    \caption{$\textit{try-commit}()$}
-    \label{alg:commit}
-    \begin{algorithmic}[1]
-    \Require  current bundle id $bid$.
-		    \State $cmd \gets struct ofp_bundle_ctrl_msg$
-    		\State  $cmd.bundle\_id \gets bid$
-    		\State  $cmd.type \gets $ \textsf{OFPBCT\_COMMIT\_REQUEST}
-    		\State  $cmd.flags \gets $ \textsf{OFPBF\_ATOMIC}
-    		\State send $cmd$
-		\Return $cmd$
-    \end{algorithmic}
-\end{algorithm}
-
 \section{Discussion of Use-Cases}\label{sec:apps}
@@
 
-For brevity, we omit explicitly putting the $\textbf{atomic}\{...\}$ construct
+For brevity, we omit explicitly putting the $\textbf{bundle}\{...\}$ construct
 around each of the transactional operation used in the algorithm.
@@
 
+Given the CAS, we can also implement more complex operations,
+for example the replicated state machine postulated in STN~\cite{stn}:
+in order to choose tags in a consistent manner, CAS needs to be enhanced
+with a notion of tags.
+
 %\paragraph{One-Shot Update} FIXME: desribe how to implement one-shot update
-
-\paragraph{STN}
-
-\textbf{TODO}
-
+%
+%\paragraph{STN}
+%
+%\textbf{TODO}
+%
+
+\begin{comment}
 \section{In-Band Composition}\label{sec:compo}
@@
 \end{figure}
+
+\end{comment}
 

### 2015-03-29 inband-sync.tex
first full iteration finished

@@
 
+we add all rules of the policy also!
+but we could add rules before and only activate later => set selector later (maybe set timeout); transaction size constant => keep bundle small important: packet atomicity so maybe on hold during bundle
+
 \section{Introduction}\label{sec:intro}
@@
 with a bundle alone it is not possible to react to
-and change switch configuration state in any non-trivial way,
+and change switch configuration state in any non-trivial way (a bundle
+essentially represents a single switch interaction),
 and with the rule overlap check it is not possible to react to
@@
 A bundle always returns a binary result: \emph{error} or \emph{success}.
+Concretely, a transaction can represent between two to five switch interactions:
+if limited space, Algo 7: read, claim id, read again, transaction, unclaim;
+if if unlimited space: alg 6 is unbounded, increase policy ID by one.
+transactions = five switch interactions , claim pid (line 5+6) together in bundle => one iteraction less?
+
+In general, it is desirable to minimize the number of transactions \stefan{stefan should cite \cite{Jin2014Dionysus} here}
 
@@
 \centering
-\includegraphics[width=0.8\columnwidth]{loadbal.pdf}~~
+\includegraphics[width=0.35\columnwidth]{loadbal-before.pdf}~\includegraphics[width=0.35\columnwidth]{loadbal.pdf}~~
 \includegraphics[width=0.8\columnwidth]{pic.pdf}\\
@@
 and
-depending on other flow spaces. Colors indicate different actions.}\label{fig:illu}
+depending on other flow spaces. Colors indicate different actions. \stefan{lets make two real transactions: one bundle and one 2 or 5 interactions?}}\label{fig:illu}
 \end{figure*}
@@
 
+\stefan{liron: maybe we can make a more concrete example like we had it in medieval?}
+
 Our first example (Algorithm~\ref{alg:simple-update}) is a simple policy composition protocol that
@@
 
-Every policy is associated with a \emph{policy identifier}
+Every policy is associated with a \emph{policy identifier}.
 In the algorithm, the controller first reads the currently installed

### 2015-03-31 inband-sync.tex
few comments...

@@
 in inconsistencies, especially if the old and the new rule
-have the same priority. \stefan{liron, is this correct?}
+have the same priority. \stefan{liron, is this correct?}\liron{a bit not clear, lets talk about this}
 \emph{FlowMod} can be used together with a the CHECK\_OVERLAP flag
@@
 a CHECK\_OVERLAP flag can be defined for each command in the bundle.
+\liron{these last two sentences are not cleaar}
 
@@
 A bundle always returns a binary result: \emph{error} or \emph{success}.
+\liron{the notion of transaction presented next is confusing considering the bundle just explained}
 Concretely, a transaction can represent between two to five switch interactions:

### 2015-04-07 inband-sync.tex
updated algs

@@
 
+\liron{for alg \ref{alg:simple-update} we should say that switch configuration = pid + policy}
+
 \begin{algorithm}[t]
@@
     \begin{algorithmic}[1]
-    \Require update requirements $U$, a predefined (memory) address for  policy id: $pid\_addr$
+    \Require unsafe policy update function $U$, a predefined switch (memory) address for policy id: $pid\_addr$
     \Ensure installed policy is consistent with previous one
@@
  		\Repeat
- 			\State \textbf{read} switch configuration
- 			\State $pid\gets$ current policy id
- 			\State $cur\_policy\gets$ current policy
- 			\State $new\_policy\gets apply\_update(cur\_policy,U)$
+ 			\State $pid,policy\gets$ \textbf{read} switch configuration
+ 			\State $cmds\gets U(policy)$
  			\startTxn
 	 			\State $CAS(pid\_addr,pid,pid+1)$
-	 			\State add $new\_policy$ rules %conditioned on $metadata=my\_a$
+	 			\State \textbf{send} $cmds$ %conditioned on $metadata=my\_a$
 	 			%\State update policy selector to $my\_a$
@@
 
+\liron{for alg \ref{alg:update} we should say that switch configuration = pid + claims + policy}
+
 \begin{algorithm}[t]
@@
     \begin{algorithmic}[1]
-    \Require update requirements $U$, a predefined (memory) address for  policy id: $pid\_addr$
+        \Require unsafe policy update function $U$, a predefined switch (memory) address for policy id: $pid\_addr$, a predefined id space $C$.
     \Ensure installed policy is consistent with previous one
@@
  		\Repeat
- 			\State \textbf{read} switch configuration
- 			\State $my\_id\gets$ an unused \emph{id}
- 			\State $pid\gets$ current policy id
+		 	\State $pid,claims,policy\gets$ \textbf{read} switch configuration
  			\State $claim(pid)$
- 			\State \textbf{read} switch configuration
- 			\If {$pid\notin Cur$}
+ 			\State $pid2,claims,policy\gets$ \textbf{read} switch configuration
+ 			\If {$pid\neq pid2$}
 	 			\State $unclaim(pid)$
@@
  			\EndIf
- 			\State $cur\_policy\gets$ current policy
- 			\State $new\_policy\gets apply\_update(cur\_policy,U)$
+ 			\State $my\_id\gets$ choose a number from $C\setminus claims$
+ 			\State $cmds\gets U(policy)$
  			\startTxn
@@
 	 			\State $CAS(pid\_addr, pid,my\_id)$
-	 			\State \textbf{add} $new\_policy$ rules %conditioned on $metadata=my\_a$
+	 			\State \textbf{send} $cmds$ %conditioned on $metadata=my\_a$
 	 			%\State update policy selector to $my\_a$

### 2015-04-12 inband-sync.tex
small changes to back and adding draft figs

@@
 to update the switch configuration. 
-In a nutshell, \emph{FlowMod} specifies the type of flow table modification,
-a match and an action. 
-The standard behavior is that the new rule simply replaces an existing one
+In a nutshell, \emph{FlowMod} specifies a new flow entry or the modifications to existing flow entries.
+The main properties of each flow entry are: 1) A match defined by a ternary patterns (a value and a mask) over packet header fields, 2) Instructions defined by  a list of modifications to packet action set 3) a cookie value used to reference a single or multiple packets.
+
+%,a match and an action. 
+The standard behavior is that the new flow entry simply replaces an existing one
 with exactly the same match, and is installed
@@
 have the same priority. \stefan{liron, is this correct?}\liron{a bit not clear, lets talk about this}
+
 \emph{FlowMod} can be used together with a the CHECK\_OVERLAP flag
@@
 FIXME: make this example better
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1\columnwidth]{no-sync.png}\\
+\caption{}\label{fig:no-sync}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1\columnwidth]{sync-with-cas.png}\\
+\caption{}\label{fig:cas-sync}
+\end{figure}
 

### 2015-04-12 inband-sync.tex
aaa

@@
 to update the switch configuration. 
-In a nutshell, \emph{FlowMod} specifies a new flow entry or the modifications to existing flow entries.
-The main properties of each flow entry are: 1) A match defined by a ternary patterns (a value and a mask) over packet header fields, 2) Instructions defined by  a list of modifications to packet action set 3) a cookie value used to reference a single or multiple packets.
+In a nutshell, \emph{FlowMod} specifies a new flow entry or modifications to existing flow entries.
+A flow entry is mainly composed from a match defined by a ternary patterns (a value and a mask) over packet header fields, and from actions.
 
@@
 The Openflow bundle can be seen as a ``mini-transaction'':
-a bundles aggregates multiple control messages (add, remove, and modify rule)
-into one, so that either all of the control messages are
+a bundles aggregates multiple control commands (e.g., FlowMod messages)
+into one execution, so that either all of the control messages are
 processed, or none of them is. In particular, if, for some reason, the

### 2015-04-12 inband-sync.tex
fixing FlowMod. still working on Bundle

@@
 in inconsistencies, especially if the old and the new rule
-have the same priority. \stefan{liron, is this correct?}\liron{a bit not clear, lets talk about this}
+have the same priority.
 
@@
 the \emph{FlowMod} and CHECK\_OVERLAP features are sufficient.
+
+In this paper, we denote the creation of a \emph{FlowMod} message by calling a function named \emph{FlowMod} with the following interface $FlowMod(match, action, flags)$, where flags are used to activate the CHECK\_OVERLAP feature.
+
+Note that the \emph{FlowMod} message (and flow entries) includes more parameters which we don't utilize in this paper and one should use default values for them in order to implement the described schemes.
 
@@
 request was processed.
-If used together with the CHECK\_OVERLAP flag, the bundle command also returns a reason 
-if the modification failed: the failed command in the bundle:
-a CHECK\_OVERLAP flag can be defined for each command in the bundle.
-\liron{these last two sentences are not cleaar}
+If multiple FlowMod messages with the CHECK\_OVERLAP flag are sent as a single bundle, the bundle command also returns a reason 
+if the checks failed: the failed command in the bundle.
+%:a CHECK\_OVERLAP flag can be defined for each command in the bundle.\liron{these last two sentences are not cleaar}
 
@@
 \stefan{liron: i think the following needs more details $:-)$ :} A bundle will always have the following structure:
-it first contains a \emph{send message} to start bundle,
-followed by a sequence of messages wrapped in bundle and bundle ID, 
-and finally end bundle control message for commit.
+it first contains a \emph{create bundle message} to start bundle,
+followed by a sequence of messages (e.g., FlowMod) each wrapped in a \emph{add to bundle message} with the bundle ID, 
+and finally \emph{commit bundle} control message for commit.
 A bundle always returns a binary result: \emph{error} or \emph{success}.
+
 \liron{the notion of transaction presented next is confusing considering the bundle just explained}

### 2015-04-16 inband-sync.tex
moved two algs to appendix "to be deleted"
some updates to bundle subsection

@@
 into one execution, so that either all of the control messages are
-processed, or none of them is. In particular, if, for some reason, the
+performed, or none of them is. In particular, if, for some reason, the
 configuration request contained in one of the bundled messages cannot
-be processed, all the other requests  are rejected and an error
+be performed, all the other requests  are rejected and an error
 message is sent to the controller that issued them.
@@
 Bundling is \emph{not required} to implement transactions.
+\liron{what does is means?}
 
 In the remainder of this paper, we will use the notation 
+%$\textbf{bundle}\{\;\ldots\;\}\rightarrow res$ to refer to represent a
 $\textbf{bundle}\{\;\ldots\;\}$ to refer to represent a
-sequence of commands in a bundle.
-\stefan{liron: i think the following needs more details $:-)$ :} A bundle will always have the following structure:
-it first contains a \emph{create bundle message} to start bundle,
-followed by a sequence of messages (e.g., FlowMod) each wrapped in a \emph{add to bundle message} with the bundle ID, 
-and finally \emph{commit bundle} control message for commit.
+sequence of commands sent in a bundle, returning $True$ if and only if all are performed successfully.
+\stefan{liron: i think the following needs more details $:-)$ :}\liron{is it ok now?} A bundle will always have the following structure:
+it first contains a \emph{create bundle message} ( \textsf{OFPT\_BUNDLE\_CONTROL} message with type
+\textsf{OFPBCT\_OPEN\_REQUEST}) to start bundle,
+followed by a sequence of commands (e.g., FlowMod) each wrapped in a \emph{add to bundle message} ( \textsf{OFPT\_BUNDLE\_ADD\_MESSAGE}) with the bundle ID, 
+and finally \emph{commit bundle} control message (\textsf{OFPT\_BUNDLE\_CONTROL} message with type \textsf{OFPBCT\_COMMIT\_REQUEST}) for commit.
 A bundle always returns a binary result: \emph{error} or \emph{success}.
@@
 
+\hide{
 To commit the transaction,  the controller sends a \textsf{OFPT\_BUNDLE\_CONTROL} message with type
@@
 \end{algorithm}
-
+}
 %\textbf{Cookies:} Cookies allow to name and retrieve individual
@@
 %   \vspace{3mm}
-			\Return
-
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[t]
-    \caption{Policy composition without bundle}
-    \label{alg:wobundle}
-    \begin{algorithmic}[1]
-    \Require policy $P$ (a set of FlowMod entry creation commands), three consecutive flow tables ids (per controller) $T_s,T_0,T_1$, active table index $active\in\{0,1\}$.
-    %\Ensure installed policy is consistent with previous one
-    \State $unactive \gets active + 1 \mod{2}$
-    \For {$cmd \in P$}
-	    \State $cmd.action\gets Reencode(cmd.action)$
-	    \State $cmd.table\_id\gets T_{unactive}$
-	    \State send $cmd$
-    \EndFor
-    \State modify the action of $T_s$ entry to goto $T_{unactive}$
-    \State clear table $T_{active}$
-    \State $active \gets unactive$
-	\Return
-
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[t]
-    \caption{Update with pipeline}
-    \label{alg:pipeline}
-    \begin{algorithmic}[1]
-    \Require update requirements $U$, array $ARR$.
-    \Ensure installed policy is consistent with previous one
 			\Return
@@
 
+\section{To Be Deleted...}
+
+\begin{algorithm}[h]
+    \caption{Policy composition without bundle}
+    \label{alg:wobundle}
+    \begin{algorithmic}[1]
+    \Require policy $P$ (a set of FlowMod entry creation commands), three consecutive flow tables ids (per controller) $T_s,T_0,T_1$, active table index $active\in\{0,1\}$.
+    %\Ensure installed policy is consistent with previous one
+    \State $unactive \gets active + 1 \mod{2}$
+    \For {$cmd \in P$}
+	    \State $cmd.action\gets Reencode(cmd.action)$
+	    \State $cmd.table\_id\gets T_{unactive}$
+	    \State send $cmd$
+    \EndFor
+    \State modify the action of $T_s$ entry to goto $T_{unactive}$
+    \State clear table $T_{active}$
+    \State $active \gets unactive$
+	\Return
+
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[h]
+    \caption{Update with pipeline}
+    \label{alg:pipeline}
+    \begin{algorithmic}[1]
+    \Require update requirements $U$, array $ARR$.
+    \Ensure installed policy is consistent with previous one
+			\Return
+
+    \end{algorithmic}
+\end{algorithm}
+
 \begin{comment}

### 2015-04-19 inband-sync.tex
old iteration of stefan. now the new one starts :)

--- 2015-04-12 inband-sync.tex
@@
 \newcommand{\memwrite}{write\xspace}
+
+\newcommand{\FlowMod}{\emph{FlowMod}\xspace}
+\newcommand{\match}{\emph{match}\xspace}
+\newcommand{\action}{\emph{action}\xspace}
+\newcommand{\flags}{\emph{flags}\xspace}
+\newcommand{\checko}{\texttt{chkov}\xspace}
 
@@
 can include complex notions of conflicts (beyond flow overlap checks),
-as well as computations and multiple controller interactions. We discuss different use cases
-for our interface (consistent policy installation, load-balancing, consensus),
-and also describe how to
-perform basic forms of policy compositions in-band.
+as well as computations and multiple controller interactions. 
+We show how our interface can be used for applications such as 
+in-band controller consensus, consistent policy installation, or load-balancing.
 \end{abstract}
@@
 
-\section*{todo}
-
-stefan add old notes, bundle size complexity (how to make it minimal: store rule in one table, select it with a pointer / selector),
-practical implementation (not even noviswitch has bundles)
-
-we add all rules of the policy also!
-but we could add rules before and only activate later => set selector later (maybe set timeout); transaction size constant => keep bundle small important: packet atomicity so maybe on hold during bundle
+%
+%
+%\section*{todo}
+%
+%stefan add old notes, bundle size complexity (how to make it minimal: store rule in one table, select it with a pointer / selector),
+%practical implementation (not even noviswitch has bundles)
+%
+%we add all rules of the policy also!
+but we could add rules before and only activate later => set selector later (maybe set timeout); transaction size constant => keep %bundle small important: packet atomicity so maybe on hold during bundle
+%
+%
 
@@
 
-Being able to provide more complex and atomic sequences of transactions over this configuration
-space can be appealing to resolve synchronization problems.
+By providing more complex and atomic transactions over the
+switch configuration
+space, several synchronization problems can be resolved.
 This paper investigates the implementation of a transactional interface
@@
 \paragraph{Our Contributions}
-This paper shows how to implement a very powerful transactional interface
+This paper shows how to implement a powerful transactional interface
 for software-defined networks.
- Concretely, our implementation is based on
- standard Openflow~\cite{of-spec} primitives, which by themselves are very limited.
-For example, we find that while Openflow's \emph{bundle} or \emph{rule overlap check}
-features can serve as synchronization primitives,
-without further mechanisms, their use is very limited:
-with a bundle alone it is not possible to react to
-and change switch configuration state in any non-trivial way (a bundle
-essentially represents a single switch interaction),
-and with the rule overlap check it is not possible to react to
-update conflicts defined over non-overlapping flow spaces.
-Moreover, we show that many powerful synchronous operations
-can also be implemented without bundles.
-
-The transactional interface presented in this paper supports very general atomic
-transactions over the switch configuration space, similar to classic transactional memory~\cite{stm-st95,tm-book}.
-For instance, a transaction allows to modify switch configurations in an interactive
-and computational fashion, as well as to define complex notions
-of conflicts.
-Our implementation is based on distributed identifiers and makes use of two
-new synchronization objects for managing these identifiers:
+This transactional interface is reminiscent of
+classic transactional memory abstractions~\cite{stm-st95,tm-book}
+and
+can be used to resolve
+synchronization problems of distributed control planes efficiently. 
+We show that this transactional interface can be used to implement
+consensus among distributed controllers or to consistently install a policy (also providing 
+the missing link for the read-modify-write object
+postulated in~\cite{cpc}).
+The atomic transactions supported by our interface
+may also define complex notions
+of conditions when transactions stand in conflict and should be serialized:
+For instance, conflicts may be defined for transactions which seemingly
+operate over \emph{independent flow spaces}, but where dependencies are introduced
+in the mapping of the flows to the physical infrastructure. This is attractive,
+e.g., for a load-balancing application.
+
+Interestingly, our transactional interface can be implemented using standard Openflow~\cite{of-spec} primitives, 
+which by themselves are of limited use.
+Concretely, we introduce, and implement in standard Openflow,
+two new synchronization objects for managing \emph{distributed identifiers}:
 a \emph{claim-id object}, used for claiming an available identifier in a given
@@
 currently used identifier.
-The claim, unclaim, and check-claim operations can be implemented individually without the Openflow
-bundle feature.
-
-Our paper therefore also provides the missing link for the read-modify-write object
-postulated in~\cite{cpc}.
+
+We hope that our work can also contribute to the ongoing discussion of what can be implemented
+in-band in Openflow~\cite{compute,reclaim}, by taking a first look
+at the relevant problem of distributed
+synchronization.
 
@@
 %how to automatically \emph{compose} updates originating \emph{from different controllers}.
-
-Finally, we hope that our work can contribute to the ongoing discussion of what can be implemented
-in-band in Openflow~\cite{compute,reclaim}, by taking a first look
-at the relevant problem of distributed
-synchronization.
 
@@
 protocol and discuss its features which are relevant for our work.
-In Section~\ref{sec:motivation}, we discuss the limitations of interpreting
-the Openflow bundle as a transaction, and motivate our approach.
+In Section~\ref{sec:motivation}, we discuss the limitations of the Openflow
+primitives and motivate the need for transactions.
 We then present our approach in detail in Section~\ref{sec:main}, and discuss use cases in Section~\ref{sec:apps}.
@@
 
-The synchronization mechanisms presented in this paper
-are based on Openflow~\cite{of-spec}, the standard SDN protocol today.
+The transactional interface presented in this paper can be implemented in 
+Openflow~\cite{of-spec}, the standard SDN protocol today.
 % In this section,
 %we provide the necessary background.
-
 In a nutshell,
@@
 Each Openflow switch stores one or multiple flow tables,
-each of which contains a set of rules (flow entries). Flow
+each of which contains a set of rules: the \emph{flow entries}.
+
+Flow
 tables form a pipeline, and flow entries are ordered according
@@
 the outcome of the table 0 processing, the packet may be sent
-to additional flow tables in the pipeline; during the traversal of
+to additional flow tables in the pipeline. 
+Instructions can be
+used to define additional tables to be visited (a ``goto instruction''),
+to modify the set of actions to be applied later (either by appending,
+deleting, or modifying actions), or to immediately apply some
+actions to the packet.
+
+During the traversal of
 the flow tables, the packet can be equipped with meta-tags to
-transfer temporary information. Concretely, instructions can be
-used to define additional tables to be visited (goto instruction),
-to modify the set of to-be-applied actions (either by appending,
-deleting, or modifying actions), or immediately apply some
-actions to the packet. A meta-data field can be used to exchange
-information between tables; it is part of the header that can be
+transfer temporary information between tables.
+The meta-data field is part of the header that can be
 inserted or removed from a packet via push and pop actions.
@@
 
-\liron{I think that maybe the rest of this section should be explained by the following subsections, providing in depth description of the OpenFlow features.}
-\liron{Another direction is to have a simplified explenation as now and to refference spesific sections of the standard in side our paper.}
-
-\subsection{The FlowMod Command}
-
-\emph{FlowMod} is the main messages used in Openflow 
-to update the switch configuration. 
-In a nutshell, \emph{FlowMod} specifies a new flow entry or modifications to existing flow entries.
-A flow entry is mainly composed from a match defined by a ternary patterns (a value and a mask) over packet header fields, and from actions.
-
+%\liron{I think that maybe the rest of this section should be explained by the following subsections, providing in depth description of the OpenFlow features.}
+%\liron{Another direction is to have a simplified explenation as now and to refference spesific sections of the standard in side our paper.}
+
+\paragraph{The FlowMod Command}
+The main messages used in Openflow
+to update the switch configuration is called \emph{FlowMod}.
+In a nutshell, \emph{FlowMod} either specifies a new flow entry or 
+modifications to existing flow entries.
+A flow entry consists of: (1) a match defined by a ternary pattern over packet header fields
+(a value and a mask), and (2) a set of actions.
 %,a match and an action.
-The standard behavior is that the new flow entry simply replaces an existing one
-with exactly the same match, and is installed
-in addition to an existing one if the match only
-partially overlaps. The latter can result
+The standard behavior is that the new flow entry will simply \emph{replace} 
+an existing one if the existing one has exactly the \emph{same match}, 
+but will be installed
+\emph{in addition} to existing entries if they only
+partially overlap (or define flowspace subsets or supersets). 
+The latter can result
 in inconsistencies, especially if the old and the new rule
-have the same priority. \stefan{liron, is this correct?}\liron{a bit not clear, lets talk about this}
-
-\emph{FlowMod} can be used together with a the CHECK\_OVERLAP flag
-of \emph{FlowMod}, in order to detect conflicts: if the flow spaces of an old
+have the same priority.
+
+In order to avoid conflicts, \emph{FlowMod} can be used together with a the $\checko$ flag ($CHECK\_OVERLAP$):
+if the flow spaces of an old
 and the new rule overlap,
-that is, if there is an entry with overlapping match part but different action, 
+that is, if there is an entry with an overlapping match part but a
+different action,
  and if the two rules have the same priority, then
-\emph{FlowMod} fails.
-
-As we will see, in order to implement our synchronization objects,
-the \emph{FlowMod} and CHECK\_OVERLAP features are sufficient.
-
-\subsection{Bundling}
-
-The Openflow bundle can be seen as a ``mini-transaction'':
-a bundles aggregates multiple control commands (e.g., FlowMod messages)
-into one execution, so that either all of the control messages are
-processed, or none of them is. In particular, if, for some reason, the
-configuration request contained in one of the bundled messages cannot
-be processed, all the other requests  are rejected and an error
+\emph{FlowMod} will fail.
+
+We will exploit the \emph{FlowMod} command in combination with
+$\checko$ to implement synchronization objects which form
+the basis of our transactional interface.
+We will use the following notation to create a \emph{FlowMod} message:
+$\FlowMod(\match, \action, \flags)$, $\match$ denotes the match,
+$\action$ the action, and $\flags$ the flags to activate the $\checko$ feature.
+The \emph{FlowMod} message (and also the flow entries) include more parameters,
+which however are irrelevant for our implementation: we suggest to
+use the default values.
+
+\paragraph{Bundling}
+Another interesting feature in Openflow 
+is the \emph{bundle}. A bundle 
+aggregates multiple control commands (e.g., \emph{FlowMod} messages)
+into one message, with \emph{all-or-nothing semantics}: 
+either all of the commands are
+performed, or none of them is. In particular, if, for some reason, the
+configuration request contained in one of the bundled commands cannot
+be performed, all the other requests  are rejected and an error
 message is sent to the controller that issued them.
-This bundled processing can be, when properly configured, done in a
-\emph{packet-atomic} way, so that each packet is processed according
+When properly configured,
+the bundled processing can be done in a
+\emph{packet-atomic} way, in the sense that 
+each packet is processed according
 to the switch configuration either \emph{before} or \emph{after} the
-request was processed.
-If used together with the CHECK\_OVERLAP flag, the bundle command also returns a reason 
-if the modification failed: the failed command in the bundle:
-a CHECK\_OVERLAP flag can be defined for each command in the bundle.
-\liron{these last two sentences are not cleaar}
-
+bundle (i.e., all its commands) took effect.
+If multiple \emph{FlowMod} messages with the $\checko$ flag are sent as a single bundle, 
+the bundle command also returns a reason
+why the overlap checks failed: it returns the failed command in the bundle.
 Concretely, when the commit fails, the switch must generate an error message corresponding to the
-message that could not be applied. The error message must use the xid of the offending message,
+message that could not be applied. The error message must use the \stefan{what is this:} \emph{xid} of the offending message,
 the error data field corresponding to that message and the error code corresponding to the failure. If
@@
 
-We will make use of bundling only for policy updates. 
-Bundling is \emph{not required} to implement transactions.
-
-In the remainder of this paper, we will use the notation 
-$\textbf{bundle}\{\;\ldots\;\}$ to refer to represent a
-sequence of commands in a bundle.
-\stefan{liron: i think the following needs more details $:-)$ :} A bundle will always have the following structure:
-it first contains a \emph{send message} to start bundle,
-followed by a sequence of messages wrapped in bundle and bundle ID, 
-and finally end bundle control message for commit.
-A bundle always returns a binary result: \emph{error} or \emph{success}.
-\liron{the notion of transaction presented next is confusing considering the bundle just explained}
-Concretely, a transaction can represent between two to five switch interactions:
-if limited space, Algo 7: read, claim id, read again, transaction, unclaim;
-if if unlimited space: alg 6 is unbounded, increase policy ID by one.
-transactions = five switch interactions , claim pid (line 5+6) together in bundle => one iteraction less?
-
-In general, it is desirable to minimize the number of transactions \stefan{stefan should cite \cite{Jin2014Dionysus} here}
-
+%:a $\checko$ flag can be defined for each command in the bundle.\liron{these last two sentences are not cleaar}
+
+In this paper, we will use the notation
+%$\textbf{bundle}\{\;\ldots\;\}\rightarrow res$ to refer to represent a
+$\textbf{bundle}\{\;\ldots\;\}$ to refer to a
+sequence of commands sent in a bundle. The bundle returns 
+$\top$ if and only if 
+all commands are performed successfully, and otherwise $\bot$.
+The bundles used in this paper will always have the following structure:
+\begin{enumerate}
+\item First, it contains a \emph{create bundle} message: a \textsf{OFPT\_BUNDLE\_CONTROL} message of type
+\textsf{OFPBCT\_OPEN\_REQUEST}.
+\item The follows a sequence of commands (e.g., \emph{FlowMod}), each wrapped in an 
+\emph{add to bundle message} (\textsf{OFPT\_BUNDLE\_ADD\_MESSAGE}) with the \emph{bundle identifier}.
+\item Finally, a \emph{commit bundle} control message (\textsf{OFPT\_BUNDLE\_CONTROL}) of 
+type \textsf{OFPBCT\_COMMIT\_REQUEST}.
+\end{enumerate}
+Again, the bundle always returns a binary result: \emph{error} ($\bot$) or \emph{success} ($\top$).
+
+%\liron{the notion of transaction presented next is confusing considering the bundle just explained}
+%Concretely, a transaction can represent between two to five switch interactions:
+%if limited space, Algo 7: read, claim id, read again, transaction, unclaim;
+%if if unlimited space: alg 6 is unbounded, increase policy ID by one.
+%transactions = five switch interactions , claim pid (line 5+6) together in bundle => one iteraction less?
+
+%In general, it is desirable to minimize the number of transactions \stefan{stefan should cite \cite{Jin2014Dionysus} here}
+
+\hide{
 To commit the transaction,  the controller sends a \textsf{OFPT\_BUNDLE\_CONTROL} message with type
@@
 
-\begin{figure*}[t]
-\centering
-\includegraphics[width=0.35\columnwidth]{loadbal-before.pdf}~\includegraphics[width=0.35\columnwidth]{loadbal.pdf}~~
-\includegraphics[width=0.8\columnwidth]{pic.pdf}\\
-\caption{\emph{Left:} Load-balancing gone wrong. \emph{Right:} Illustration of transactional abstraction which  allows
-controllers to modify switch configurations in complex ways,
-and
-depending on other flow spaces. Colors indicate different actions. \stefan{lets make two real transactions: one bundle and one 2 or 5 interactions?}}\label{fig:illu}
-\end{figure*}
+%\begin{figure*}[t]
+%\centering
+%\includegraphics[width=0.35\columnwidth]{loadbal-before.pdf}~\includegraphics[width=0.35\columnwidth]{loadbal.pdf}~~
+%\includegraphics[width=0.8\columnwidth]{pic.pdf}\\
+%\caption{\emph{Left:} Load-balancing gone wrong. \emph{Right:} Illustration of transactional abstraction which  allows
+%controllers to modify switch configurations in complex ways,
+%and
+%depending on other flow spaces. Colors indicate different actions. \stefan{lets make two real transactions: one bundle and one 2 or 5 interactions?}}\label{fig:illu}
+%\end{figure*}
 
@@
 and does not make use
-of group tables. 
-Moreover, we do not make use of cookies: rules
+of \emph{Openflow group tables}, nor do we make use of \emph{cookies}: rules
 are always identified by exact matches.
@@
 
-\stefan{liron, maybe we can discuss this section together: we need some more ideas}
-
-Before presenting our transactional interface and implementation,
-let us first discuss the limitations of the Openflow features discussed above.
-At first sight, the bundle feature seems to provide everything
-we need: it defines a sequence of read and write operations
+At first sight, the Openflow bundle feature seems to provide everything
+we need to implement our transactional interface: the bundle
+ defines a sequence of read and write operations
  on the switch state which are executed atomically.
- However, such bundle transactions come with several
- limitations, as discussed in the following. Moreover,
- as we will see, our mechanism also does not \emph{need}
- the bundle feature.
-
-A bundle does not allow
- to change the switch state in any non-trivial way,
- as the read and write actions in the bundle do
+ However, while the bundle can indeed be seen 
+ as a ``mini-transaction'', it comes with several
+ limitations, which we discuss in the following.
+
+First, if employed without any additional features, a bundle does not allow
+ to change the switch state in any non-trivial way:
+the read and write actions in the bundle do
  not support any conditional modifications or computations which
- depend on the read results. By leveraging
- the check overlap feature, it is at least possible
- to abort transactions depending on certain conditions
- in the overlapping
- flow spaces. However, in many cases, for example
- a load-balancing application,
- it can be interesting to define more general
- dependencies, which depend on, from a logical perspective,
- seemingly
- completely independent flow spaces,
- which however share the underlying physical network.
-In other words, the bundle feature only allows us to
-define ``one-shot transactions'',
-transactions which do not support any further controller interactions.
-Given
-the limited computational options the Openflow interface of a switch offers,
-it is difficult to implement more complex behavior.
-
-\textbf{Example 1.} To give an example, let us consider a controller application in charge
-of \emph{load-balancing}. The objective of the load-balancer
+ depend on the read results. 
+But also using the bundle together with the check overlap feature ($\checko$)
+comes with limitations: while it is possible
+ to \emph{abort} transactions depending on certain conditions
+ in the \emph{overlapping
+ flow spaces}, oftentimes it is desirable to be able
+ to define more general conditions and notions of conflicts
+ under which transactions should not take effect. Moreover,
+ sometimes it is desirable to react to conflicts in smarter ways than
+ to simply abort, e.g.,
+ by supporting \emph{modifications}.
+Let us consider two examples.
+
+\begin{enumerate}
+\item \textbf{Example 1: More complex dependencies.} Let us consider a controller application in charge
+of \emph{load-balancing}. 
+In a load-balancing application, seemingly independent actions, defined over
+completely independent logical flow spaces,
+may in fact be dependent due to collocation in the underlying physical network.
+To be more concrete, say that the objective of the load-balancer
 is to distribute flows equally across links or servers.
 If multiple controllers concurrently and independently
-update forwarding rules, they may unvoluntarily
-unbalance the flow allocation. Note that this
-is true independently of whether the flow spaces
-affected by the respective controllers overlap or not.
-More concretely, assume that two controllers concurrently
-decide to forward traffic $1*$ and $0*$ to the same server.
+update forwarding rules (say, two ``independent'' flows with source IP address ranges $1*$ and $0*$), 
+they may unvoluntarily
+unbalance the flow allocation, as the managed flows are embedded
+on the same physical links.  
 Flow space overlap checks cannot be used to detect
 such conflicts.
-FIXME: make this example better
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1\columnwidth]{no-sync.png}\\
-\caption{}\label{fig:no-sync}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1\columnwidth]{sync-with-cas.png}\\
-\caption{}\label{fig:cas-sync}
-\end{figure}
-
-\textbf{Example 2.} To give another example,
-consider the problem of modifying previously
+
+\stefan{todo: explain figure}
+
+\item \textbf{Example 2: Supporting modifications.} Consider the problem of modifying previously
 installed forwarding policies.
@@
 old rules (and, using the flow space overlap
-check, to abort the transaction if potentially conflicting
+check feature $\checko$, to abort the transaction if potentially conflicting
 rules were already there), it cannot be used
@@
 for the conditional creation of rules, not their modification.
-
-\stefan{It would be cool if we could make these examples more concrete,
-at least one, and get back to it later again!}
+\end{enumerate}
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=1\columnwidth]{no-sync.png}~\includegraphics[width=1\columnwidth]{sync-with-cas.png}\\
+\caption{\emph{Left:} No synchronization. \emph{Right:} CAS.}
+\end{figure*}
+
+As we will see, we do not even need the bundle feature
+to implement the powerful synchronization objects of our transactional interface.
+However, the bundle feature can still be useful in the context of
+consistent policy updates.
+
+%At first sight, a bundle seems to be a powerful concept
+%for the implementation of a transactional interface. However,
+%while the  bundle can indeed be seen as a ``mini-transaction'',
+%in itself it is not very useful. In fact, as we will discuss
+%in more detail in the next section, we will not make
+%use of the bundle in the implementation of our synchronization
+%objects, but will employ it for our case study
+%of consistent policy updates.%
+%
+%Before presenting our transactional interface and implementation,
+%let us first discuss the limitations of the Openflow features discussed above.
 
 %\section{The Container Object}\label{sec:vision}
-\section{Transactional Synchronization}\label{sec:main}
-
-This section describes how to introduce more powerful transactions.
-Our implementation is based on two new synchronization objects,
-allowing to coordinate controllers using identifiers.
-At a high level, we export a
+\section{Transaction $\&$ Synchronization}\label{sec:main}
+
+Given this motivation, we now present a more powerful transactional interface.
+At a high level, we aim to export a
 transactional interface which allows each controller $i\in[N]$ ($N\geq
@@
 
+Our approach is based on the distributed management of \emph{identifiers (IDs)}. In particular, 
+we introduce two new synchronization objects (and describe their implementation in standard Openflow),
+which allows controllers to coordinate 
+themselves based on these identifiers.
+
 \subsection{Synchronization Objects}\label{sec:t-if}
 
-We implement transactional access to  two kinds of objects: a read-write \emph{memory} abstraction
-object, which we later use for storing the current used policy identifier, and a
-\emph{resource-claimer} object, provides claiming a resource id (e.g. policy \emph{identifier}) in a
+We implement transactional access to  two kinds of objects: 
+a read-write \emph{memory} abstraction
+object, which we will use for storing the current used policy identifier, and a
+\emph{resource-claimer} object, which allows controllers to claim a resource ID, 
+e.g., a policy \emph{identifier} in a
 given (bounded or unbounded) set.
-In theory one can use the memory abstraction to implement the resource claimer, 
-however we found it far more simple and efficient to consider the resource claimer as base object.
-%
-
-A memory can be accessed with the following sequential
-semantics:
-
+%While In theory one can use the memory abstraction to implement the resource claimer,
+%however we found it far more simple and efficient to consider the resource claimer as base object.
+%
+
+With these base objects, we provide the following shared memory abstraction, with 
+write-compare access semantics:
 \begin{itemize}
@@
   $addr,k\in\Nat$, the value at memory offset $addr$ is set to $k$. For
-  example, the offset $addr$ identifies the policy (identifier), 
+  example, the offset $addr$ may be used to identify the policy (identifier),
   but our approach is more general.
 
-\item $\textit{\compare}(addr,k)$, $addr,k\in\Nat$,
-aborts current transaction if the value at memory offset $addr$ is not equal to $k$.
+\item The operation $\textit{\compare}(addr,k)$, $addr,k\in\Nat$,
+aborts current transaction, if the value at memory offset $addr$ is not equal to $k$.
 %returns $\textit{true}$ if and
@@
 %  \emph{fails} (which causes an abort of the transaction invoking it).
-
 \end{itemize}
 
-Note that reading the memory can be done by just reading the switch configuration however such action doesn't affect the execution of the transaction (as $\textit{\compare}$ does).
-As we will see, this abstraction can be implemented
-without using the notion of bundles.
-
-An immediate consequence of this memory interface is 
-that we can implement a \emph{compare-and-set} (CAS) operation that takes two integer parameters
+As we will see, the bundle feature is not required to implement this abstraction.
+Also note that memory read operations correspond to reading the switch 
+configuration, and are simple as they do not affect the execution of the transaction 
+(unlike the $\textit{\compare}$ operation).
+
+With this memory interface we can immediately implement 
+a powerful \emph{compare-and-set} (CAS) operation which takes two integer parameters
 $\textit{old}$ and $\textit{new}$, reads the register and, if the
 value of it is $\textit{old}$, replaces it with $\textit{new}$ and
-returns $\textit{true}$; otherwise, $\textit{false}$ is returned and
+returns $\top$; otherwise, $\bot$ is returned and
 we say that the CAS operation \emph{fails} (which causes an abort if
@@
 
-A resource-claimer object implements a weak type of locking, exports operations \emph{claim},
+Our resource-claimer object implements a weak type of locking, and exports operations \emph{claim},
 \emph{unclaim} and \emph{\claimcheck} with the following sequential
@@
 \begin{itemize}
-\item by performing $\textit{claim}(k)$, $i\in[N]$,
-  $k\in\Nat$, controller $i$ \emph{claims}
-  identifier $k$;
-
-\item by performing $\textit{unclaim}(k)$, $i\in[N]$, $k\in\Nat$, controller $i$
-  \emph{unclaim} identifier $k$;
-
-\item $\textit{\claimcheck}(k)$, $k\in\Nat$,
-abort current transaction if $k$ is currently claimed by any controller.
+\item With $\textit{claim}(k)$, $i\in[N]$,
+  $k\in\Nat$, a controller $i$ \emph{claims}
+  identifier $k$.
+
+\item With $\textit{unclaim}(k)$, $i\in[N]$, $k\in\Nat$, controller $i$
+  \emph{unclaims} identifier $k$;
+
+\item If $k$ is currently claimed by any controller, 
+$\textit{\claimcheck}(k)$, $k\in\Nat$,
+aborts the current transaction.
 %returns $\textit{true}$ if and
@@
 
-\subsubsection{A resource claimer}
+\paragraph{The resource claimer}
 %features , including
@@
 Algorithms~\ref{alg:claim}-\ref{alg:cas} describe implementations of
-transactional operations  $\textit{claim}(x)$, $\textit{\claimcheck}(x)$, $\textit{unclaim}(x)$, $\textit{write}(x)$,  $\textit{\compare}(x)$
+transactional operations  $\textit{claim}(x)$, $\textit{\claimcheck}(x)$, 
+$\textit{unclaim}(x)$, $\textit{write}(x)$,  $\textit{\compare}(x)$
 and $\textit{\cas}(x)$, respectively.
@@
 
-We describe each operation as a message factory that returns OpenFlow commands that should be sent to the switch as part of the same transaction.
-We assume that every transaction has a unique identifier  used to start a \emph{bundle}.
-All messages generated during a transaction invocation, are wrapped by bundle messages carrying the bundle's id and followed by bundle commit message.
-Messages with same bundle id are considered to belong to same bundle and are performed atomically.
+We describe each operation as a message factory that returns OpenFlow commands which 
+should be sent to the switch as part of the same transaction.
+We assume that every transaction has a unique identifier and is used to start a \emph{bundle}.
+All messages generated during a transaction invocation, are wrapped by bundle messages carrying 
+the bundle's ID and followed by bundle commit message.
+Messages with same bundle ID are considered to belong to same bundle and are performed atomically.
 
@@
 
-Our implementation is based on OpenFlow flow entries modification command, the FlowMod message that we explain in Section \ref{sec:background}.
-We use the CHECK\_OVERLAP optional flag to make the transaction execution dependent on switch state.
-
-We consider the match part of an entry to be a ternary string of bounded length (represented by binary value and mask) which can be applied to any set of supported packet header fields.
-In our implementation the $64$bit metedata field alone is suffice.
-Moreover we consider the action part of an entry to be an integer, which in practice can be implemented by a Set-Metadata instruction where the written value is that integer.
+Our implementation is based on the \emph{FlowMod} command. 
+We use the $\checko$ optional flag to make the transaction execution dependent on switch state.
+We consider the match part of an entry to be a ternary string of bounded length (represented by binary value and mask) 
+which can be applied to any set of supported packet header fields. In our implementation, the $64$bit meta-data field alone is 
+already sufficient. Moreover we consider the action part of an entry to be an integer, which in practice can be implemented by a 
+\emph{set meta-data} instruction where the written value is that integer.
 
 As can be seen in Algorithms \ref{alg:claim}-\ref{alg:unclaim}, for the resource-claimer,
-  the match part of the added flow entry represents the claimed id (id concatenated twice).
-   In addition the mask part of the match (controller id concatenated with all-ones) varies from one controller to another in order to allow multiple controllers to claim the same number without overriding the same flow entry.
-    We can see that for any two resource ids $x_1$ and $x_2$ and controller ids $c_1\neq c_2$ the match patterns $m_1=(x_1\concat x_1, c_1\concat all-1)$ and $ m_2=(x_2\concat x_2, c_2\concat all-1)$ are not equal even if $x_1=x_2$ therefore they can both be used to claim and unclaim without affecting each other.
-
-In order to check if a number is claimed we try to create the corresponding entry while setting the
-flag \textsf{CHECK\_OVERLAP} and using different action ($2$ instead of $1$), thereby inflicting a failure in case an entry with overlapping match value exists.
-If the check succeeds we delete the entry.
-Considering $m_1$ and $m_2$ defined above and their respective actions $a_1=x_1\concat c1$ and $a_2=x_2\concat c2$,  they overlap with different actions iff $x_1=x_2$.
-Therefore if one is used in \claimcheck after the other used in claim, a conflict is detected iff $x_1=x_2$, as expected.
-
-Note that for switch space reasons, checking for claims contains also a deletion of the added tester flow entry but it has no affect on claims and checks of other controllers and therefore the deletion command doesn't have to be sent as part of a bundle.
-This makes the \textsf{\claimcheck} operation atomic by nature, similar to the \textsf{claim} and \textsf{unclaim}, as being implemented by essentially one command.
+the match part of the added flow entry represents the claimed ID (the ID is concatenated twice).
+In addition the mask part of the match (controller ID concatenated with all-ones) varies from one controller 
+to another in order to allow multiple controllers to claim the same number without overriding the same flow entry.
+We can see that for any two resource IDs $x_1$ and $x_2$ and controller IDs $c_1\neq c_2$, 
+the match patterns $m_1=(x_1\concat x_1, c_1\concat all-1)$ and $ m_2=(x_2\concat x_2, c_2\concat all-1)$ are not equal, 
+even if $x_1=x_2$; therefore, they can both be used to claim and unclaim identifiers without affecting each other.
+
+In order to check if a number is claimed, we try to create the 
+corresponding entry while setting the
+flag \textsf{\checko}, and using different actions ($2$ instead of $1$), thereby inflicting 
+a failure in case an entry with overlapping match value exists. 
+If the check succeeds, we delete the entry. Considering $m_1$ and $m_2$ as defined above, 
+and their respective actions $a_1=x_1\concat c1$ and $a_2=x_2\concat c2$,  they overlap 
+with different actions iff $x_1=x_2$. Therefore, if one is used in \claimcheck after the other used in the claim operation, 
+a conflict is detected iff $x_1=x_2$, as desired.
+
+Note that for switch space reasons, when checking for claims, also the deletion of the added tester flow entry is included,
+which however has no effect on claims and checks of other controllers; hence, the deletion command does not 
+have to be sent as part of a bundle. Therefore, the \textsf{\claimcheck} operation is atomic by nature, 
+similar to the \textsf{claim} and \textsf{unclaim} operations, and is implemented as essentially one command.
 
@@
     \begin{algorithmic}[1]
-    \Require \emph{self} as the calling controller id.
+    \Require \emph{self} as the calling controller ID.
 %    \Ensure
@@
     		\State $action \gets 2$
-    		\State $flag \gets \textsf{OFPFF\_CHECK\_OVERLAP}$
+    		\State $flag \gets \textsf{OFPFF\_\checko}$
     		\State $cmd1\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
@@
 
-\subsubsection{A memory abstraction}
+\paragraph{A memory abstraction}
 
@@
 
-In order to check if a value is written in a given address, we try to add it the same way as in the \memwrite operation  but we also set the \textsf{CHECK\_OVERLAP} flag.
+In order to check if a value is written in a given address, we try to add it the same way as in the \memwrite operation  but we also set the \textsf{\checko} flag.
 If the value is not there, our action differ from the action of the existing entry thereby inflicting a failure.
@@
     		\State $action \gets k$
-    		\State $flag \gets \textsf{OFPFF\_CHECK\_OVERLAP}$
+    		\State $flag \gets \textsf{OFPFF\_\checko}$
     		\State $cmd\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
@@
 
-The CHECK\_OVERLAP flag enables us to add an entry if no contradicting entry exists and to fail if such do exists.
+The $\checko$ flag enables us to add an entry if no contradicting entry exists and to fail if such do exists.
 However, the compare stage of \cas requires us to fail if a certain value doesn't exists, i.e.
@@
     		\State $action \gets 2$
-    		\State $flag \gets OFPFF\_CHECK\_OVERLAP$
+    		\State $flag \gets OFPFF\_\checko$
     		\State $cmd1\gets \textsc{FlowMod}(match, op = ADD, flag, action) $

### 2015-04-19 inband-sync.tex
stefan next pass

@@
 \newcommand{\memwrite}{write\xspace}
+
+\newcommand{\addr}{\emph{addr}\xspace}
+\newcommand{\pid}{\emph{pid}\xspace}
+\newcommand{\add}{\textbf{add}\xspace}
+\newcommand{\dele}{\textbf{delete}\xspace}
 
@@
 \item The follows a sequence of commands (e.g., \emph{FlowMod}), each wrapped in an 
-\emph{add to bundle message} (\textsf{OFPT\_BUNDLE\_ADD\_MESSAGE}) with the \emph{bundle identifier}.
+\emph{add to bundle message} (\textsf{OFPT\_BUNDLE\_\add\_MESSAGE}) with the \emph{bundle identifier}.
 \item Finally, a \emph{commit bundle} control message (\textsf{OFPT\_BUNDLE\_CONTROL}) of 
@@
 %if if unlimited space: alg 6 is unbounded, increase policy ID by one.
-%transactions = five switch interactions , claim pid (line 5+6) together in bundle => one iteraction less?
+%transactions = five switch interactions , claim \pid (line 5+6) together in bundle => one iteraction less?
 
@@
 \begin{itemize}
-\item By performing $\textit{Write}(addr,k)$,
-  $addr,k\in\Nat$, the value at memory offset $addr$ is set to $k$. For
-  example, the offset $addr$ may be used to identify the policy (identifier),
+\item By performing $\textit{write}(\addr,k)$,
+  $\addr,k\in\Nat$, the value at memory offset $\addr$ is set to $k$. For
+  example, the offset $\addr$ may be used to identify the policy (identifier),
   but our approach is more general.
 
-\item The operation $\textit{\compare}(addr,k)$, $addr,k\in\Nat$,
-aborts current transaction, if the value at memory offset $addr$ is not equal to $k$.
+\item The operation $\textit{\compare}(\addr,k)$, $\addr,k\in\Nat$,
+aborts current transaction, if the value at memory offset $\addr$ is not equal to $k$.
 %returns $\textit{true}$ if and
@@
     		\State $flag \gets 0$
-    		\State $cmd\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
+    		\State $cmd\gets \emph{FlowMod}(match, op = \add, flag, action) $
 			\Return $cmd$
@@
     		\State $flag \gets \textsf{OFPFF\_\checko}$
-    		\State $cmd1\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
-    		\State $cmd2\gets \textsc{FlowMod}(match, op = DELETE) $
+    		\State $cmd1\gets \emph{FlowMod}(match, op = \add, flag, action) $
+    		\State $cmd2\gets \emph{FlowMod}(match, op = \dele) $
 			\Return $cmd1$,$cmd2$
@@
     		\State $match \gets (value,mask)$
-    		\State $cmd\gets \textsc{FlowMod}(match, op=DELETE) $
+    		\State $cmd\gets \emph{FlowMod}(match, op=\dele) $
 
@@
 \paragraph{A memory abstraction}
-
-We use similar techniques for implementing memory \memwrite and \compare
- operations.
+The memory \memwrite and \compare
+ operations can be implemented using similar techniques.
  As can be seen in Algorithms \ref{alg:write}-\ref{alg:compare},
-  the match part of the added flow entry is used as the memory address and the action as the written value.
-Values written to the same address will replace one another, following the command definition to replace old flow entry with a new one in case they share the same match parts (as described in Section \ref{sec:background}).
-
-In order to check if a value is written in a given address, we try to add it the same way as in the \memwrite operation  but we also set the \textsf{\checko} flag.
-If the value is not there, our action differ from the action of the existing entry thereby inflicting a failure.
-In case the value is there, our flow entry replaces the existing flow entry with identical one.
-In addition we present in Algorithm \ref{alg:cas} a simple implementation of Compare and set (CAS) using the two last operations.
+  the match part of the added flow entry is used as the memory address and the action as 
+  the written value. Values written to the same address will replace one another,
+following the command definition to replace old flow entry with a new one in case they 
+share the same match parts.
+
+In order to check if a value is written in a given address, 
+we try to add it the same way as in the \memwrite operation,  
+but we also set the \textsf{\checko} flag. If the value is not there, 
+our action differs from the action of the existing entry, 
+thereby inflicting a failure.
+In case the value is there, our flow entry replaces the existing flow entry 
+with an identical one.
+In addition, we present in Algorithm \ref{alg:cas} a simple implementation 
+of \emph{compare-and-set (CAS)}, using the two last operations.
 
 \begin{algorithm}[h]
-    \caption{$\textit{write}(addr,k)$}
+    \caption{$\textit{write}(\addr,k)$}
     \label{alg:write}
@@
 %    \Ensure
-    		\State $value \gets addr$
+    		\State $value \gets \addr$
     		\State $mask \gets  all-1$
@@
 \begin{algorithm}[h]
-    \caption{$\textit{\compare}(addr,k)$}
+    \caption{$\textit{\compare}(\addr,k)$}
     \label{alg:compare}
@@
     		\State $flag \gets \textsf{OFPFF\_\checko}$
-    		\State $cmd\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
+    		\State $cmd\gets \emph{FlowMod}(match, op = \add, flag, action) $
 			\Return $cmd$
@@
 \begin{algorithm}[h]
-    \caption{$\textit{\cas}(addr, old,new)$}
+    \caption{$\textit{\cas}(\addr, old,new)$}
     \label{alg:cas}
@@
 
-    		\State $cmd1\gets \textsc{\compare}((addr, old) $
-    		\State $cmd2\gets \textsc{\memwrite}((addr, new) $
+    		\State $cmd1\gets \textsc{\compare}((\addr, old) $
+    		\State $cmd2\gets \textsc{\memwrite}((\addr, new) $
 			\Return $cmd1,cmd2$
@@
 
-\section{Discussion of Use-Cases}\label{sec:apps}
-
-More powerful transactions can be useful in many scenarios.
-Let us revisit our examples from earlier: with
-transactions we can easily solve the load-balancing
-and rule modification problem.
-In the following we give two more examples of our approach.
+\section{Discussion of Use Cases}\label{sec:apps}
+
+Let us discuss some use cases of our transactional interface.
+
+\subsection{Consensus and Replicated State Machines}
+
+Using our approach, it is straightforward to implement consensus among the controllers
+using just a single register that can be accessed with an atomic CAS
+operation~\cite{Her91}. In our transactional setting, the controllers
+can use CAS to replace the default value of such a register
+with their consensus input values (e.g., the identifiers of the
+policies that should be applied first). If the corresponding
+singleton transaction commits, the controller considers itself
+the \emph{winner} and decides on its input, if
+it aborts (the controller fails in the CAS) it remains to read the register to get the input of the
+winner to decide.
+
+By using multiple CAS invocations, the controllers can implement any
+generic replicated state machine~\cite{Her91,Lam98}, which will be instrumental in the
+next section.
+
+Given the CAS, we can also implement more complex operations,
+for example the replicated state machine postulated in STN~\cite{stn}:
+in order to choose tags in a consistent manner, CAS needs to be enhanced
+with a notion of tags.
 
@@
 
-\stefan{liron: maybe we can make a more concrete example like we had it in medieval?}
-
-Our first example (Algorithm~\ref{alg:simple-update}) is a simple policy composition protocol that
+As a second example, let us consider a simple policy composition protocol that
 allows a set of  controllers to concurrently update switch  \emph{policies} , i.e., sets of
-effective rules in its flow tables, under the
+effective flow tables rules, under the
 condition that these rules can be composed with the currently installed
-policy (the composition predicate is encoded in a user-specific
+policy. 
+Let us assume that the composition predicate is encoded in a user-specific
 $\textit{apply\_update}()$ function~\cite{cpc}.
 
-Every policy is associated with a \emph{policy identifier}.
+Algorithm~\ref{alg:simple-update} is based on the bundle feature and associates each policy with a \emph{policy identifier (pid)}.
 In the algorithm, the controller first reads the currently installed
-policy, then choose a new unique policy, checks whether the new
+policy, then chooses a new unique policy identifier \emph{pid}, checks whether the new
 policy and the current policy can be composed and, if so, replaces the
@@
 
-Algorithm~\ref{alg:simple-update} use increasing numbers as policies
+We use increasing numbers as policies
 ids and use them in order to make sure that the new policy, with id
@@
 
-\liron{for alg \ref{alg:simple-update} we should say that switch configuration = pid + policy}
+\liron{for alg \ref{alg:simple-update} we should say that switch configuration = \pid + policy}
 
@@
     \begin{algorithmic}[1]
-    \Require unsafe policy update function $U$, a predefined switch (memory) address for policy id: $pid\_addr$
+    \Require unsafe policy update function $U$, a predefined switch (memory) address for policy id: $\pid_{\addr}$
     \Ensure installed policy is consistent with previous one
@@
  		\Repeat
- 			\State $pid,policy\gets$ \textbf{read} switch configuration
+ 			\State $\pid,policy\gets$ \textbf{read} switch configuration
  			\State $cmds\gets U(policy)$
  			\startTxn
-	 			\State $CAS(pid\_addr,pid,pid+1)$
+	 			\State $CAS(\pid_{\addr},\pid,\pid+1)$
 	 			\State \textbf{send} $cmds$ %conditioned on $metadata=my\_a$
@@
 
-\liron{for alg \ref{alg:update} we should say that switch configuration = pid + claims + policy}
+\liron{for alg \ref{alg:update} we should say that switch configuration = \pid + claims + policy}
 
@@
     \begin{algorithmic}[1]
-        \Require unsafe policy update function $U$, a predefined switch (memory) address for policy id: $pid\_addr$, a predefined id space $C$.
+        \Require unsafe policy update function $U$, a predefined switch (memory) address for policy id: $\pid_{\addr}$, a predefined id space $C$.
     \Ensure installed policy is consistent with previous one
@@
  		\Repeat
-		 	\State $pid,claims,policy\gets$ \textbf{read} switch configuration
- 			\State $claim(pid)$
- 			\State $pid2,claims,policy\gets$ \textbf{read} switch configuration
- 			\If {$pid\neq pid2$}
-	 			\State $unclaim(pid)$
+		 	\State $\pid,claims,policy\gets$ \textbf{read} switch configuration
+ 			\State $claim(\pid)$
+ 			\State $\pid 2,claims,policy\gets$ \textbf{read} switch configuration
+ 			\If {$\pid\neq \pid 2$}
+	 			\State $unclaim(\pid)$
  				\State {\bf continue} (\textbf{restart} loop)
@@
  				\State $\claimcheck(my\_id)$
-	 			\State $CAS(pid\_addr, pid,my\_id)$
+	 			\State $CAS(\pid_{\addr}, \pid,my\_id)$
 	 			\State \textbf{send} $cmds$ %conditioned on $metadata=my\_a$
@@
  			\endTxn
-	 		\State $unclaim(pid)$
+	 		\State $unclaim(\pid)$
      		%\State $res \gets $ commit transaction
@@
 \end{algorithm}
-
-\subsection{Consensus and Replicated State Machines}
-
-It is straightforward to implement consensus among the controllers
-using just a single register that can be accessed with an atomic CAS
-operation~\cite{Her91}. In our transactional setting, the controllers
-can use CAS to replace the default value of such a register
-with their consensus input values (e.g., the identifiers of the
-policies that should be applied first). If the corresponding
-singleton transaction commits, the controller consideres itself
-the \emph{winner} and decides on its input, if
-it aborts (the controller fails in the CAS) it remains to read the register to get the input of the
-winner to decide.
-
-By using multiple CAS invocations, the controllers can implement any
-generic replicated state machine~\cite{Her91,Lam98}, which will be instrumental in the
-next section.
-
-Given the CAS, we can also implement more complex operations,
-for example the replicated state machine postulated in STN~\cite{stn}:
-in order to choose tags in a consistent manner, CAS needs to be enhanced
-with a notion of tags.
 
@@
 
-\cleardoublepage
+\clearpage
 
@@
 		that must match when the command is
-		OFPFC_MODIFY* or OFPFC_DELETE*. A value
+		OFPFC_MODIFY* or OFPFC_\dele*. A value
 		of 0 indicates no restriction. */
@@
 	uint8_t table_id; /* ID of the table to put the flow in.
-		For OFPFC_DELETE_* commands, OFPTT_ALL
+		For OFPFC_\dele_* commands, OFPTT_ALL
 		can also be used to delete matching
@@
 		OFP_NO_BUFFER.
-		Not meaningful for OFPFC_DELETE*. */
-	uint32_t out_port; /* For OFPFC_DELETE* commands, require
+		Not meaningful for OFPFC_\dele*. */
+	uint32_t out_port; /* For OFPFC_\dele* commands, require
 		matching entries to include this as an
@@
 		indicates no restriction. */
-	uint32_t out_group; /* For OFPFC_DELETE* commands, require
+	uint32_t out_group; /* For OFPFC_\dele* commands, require
 		matching entries to include this as an
@@
     		\State $flag \gets OFPFF\_\checko$
-    		\State $cmd1\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
-    		\State $cmd2\gets \textsc{FlowMod}(match, op = DELETE) $
-    		\State $cmd3\gets \textsc{FlowMod}(cookie = old, op = DELETE) $
+    		\State $cmd1\gets \emph{FlowMod}(match, op = \add, flag, action) $
+    		\State $cmd2\gets \emph{FlowMod}(match, op = \dele) $
+    		\State $cmd3\gets \emph{FlowMod}(cookie = old, op = \dele) $
     		\State $cmds.append(cmd1)$
@@
     		    		\State $flag \gets 0$
-    		    		\State $cmd\gets \textsc{FlowMod}(r\_match, ADD, flag, cookie, action) $
+    		    		\State $cmd\gets \emph{FlowMod}(r\_match, \add, flag, cookie, action) $
     		    		\State $cmds.append(cmd)$
@@
 
-    		\State $op \gets \textsf{OFPGC\_DELETE}$
+    		\State $op \gets \textsf{OFPGC\_\dele}$
     		\State $group\_id \gets old$
@@
 
-    		\State $op \gets \textsf{OFPGC\_ADD}$
+    		\State $op \gets \textsf{OFPGC\_\add}$
     		\State $type \gets \textsf{OFPGT\_INDIRECT}$

### 2015-04-19 inband-sync.tex
final for now

@@
 \newcommand{\flags}{\emph{flags}\xspace}
-\newcommand{\checko}{\texttt{chkov}\xspace}
+\newcommand{\checko}{\texttt{check\_overlap}\xspace}
 
@@
 as well as computations and multiple controller interactions. 
-We show how our interface can be used for applications such as 
-in-band controller consensus, consistent policy installation, or load-balancing.
+We discuss applications of our transactional memory interface to  
+in-band controller consensus, consistent policy installation, and load-balancing.
 \end{abstract}
@@
 
-In order to avoid conflicts, \emph{FlowMod} can be used together with a the $\checko$ flag ($CHECK\_OVERLAP$):
+In order to avoid conflicts, \emph{FlowMod} can be used together with a the $\checko$ flag:
 if the flow spaces of an old
@@
 \begin{enumerate}
-\item First, it contains a \emph{create bundle} message: a \textsf{OFPT\_BUNDLE\_CONTROL} message of type
-\textsf{OFPBCT\_OPEN\_REQUEST}.
+\item First, it contains a \emph{create bundle} message: a \texttt{ofpt\_bundle\_control} message of type
+\texttt{OFPBCT\_OPEN\_REQUEST}.
 \item The follows a sequence of commands (e.g., \emph{FlowMod}), each wrapped in an 
-\emph{add to bundle message} (\textsf{OFPT\_BUNDLE\_\add\_MESSAGE}) with the \emph{bundle identifier}.
-\item Finally, a \emph{commit bundle} control message (\textsf{OFPT\_BUNDLE\_CONTROL}) of 
-type \textsf{OFPBCT\_COMMIT\_REQUEST}.
+\emph{add to bundle message} (\texttt{ofpt\_bundle\_add\_message}) with the \emph{bundle identifier}.
+\item Finally, a \emph{commit bundle} control message (\texttt{ofpt\_bundle\_control}) of
+type \texttt{ofpbct\_commit\_request}.
 \end{enumerate}
@@
 \hide{
-To commit the transaction,  the controller sends a \textsf{OFPT\_BUNDLE\_CONTROL} message with type
-\textsf{OFPBCT\_COMMIT\_REQUEST} and waits until a matching control
+To commit the transaction,  the controller sends a \texttt{OFPT\_BUNDLE\_CONTROL} message with type
+\texttt{OFPBCT\_COMMIT\_REQUEST} and waits until a matching control
 message from the switch is received. When the switch receives the
@@
 equipped with the corresponding bundle identifier and, if the
-processing is successful for all of them, returns  a \textsf{OFPT\_BUNDLE\_CONTROL} message with type
-\textsf{OFPBCT\_CLOSE\_REPLY}; otherwise it sends a specific error
+processing is successful for all of them, returns  a \texttt{OFPT\_BUNDLE\_CONTROL} message with type
+\texttt{OFPBCT\_CLOSE\_REPLY}; otherwise it sends a specific error
 message  pointing out the aborted transactional operation.
@@
     		\State  $cmd.bundle\_id \gets bid$
-    		\State  $cmd.type \gets $ \textsf{OFPBCT\_COMMIT\_REQUEST}
-    		\State  $cmd.flags \gets $ \textsf{OFPBF\_ATOMIC}
-    		\State send $cmd$
+    		\State  $cmd.type \gets $ \texttt{ofpbct\_commit\_request}
+    		\State  $cmd.flags \gets $ \texttt{ofpbf\_atomic}
+    		\State \textbf{send} $cmd$
 		\Return $cmd$
@@
  defines a sequence of read and write operations
- on the switch state which are executed atomically.
+ on the switch state which is executed atomically.
  However, while the bundle can indeed be seen 
@@
 the read and write actions in the bundle do
- not support any conditional modifications or computations which
+ not support any conditional \emph{modifications} or computations which
  depend on the read results. 
@@
 \begin{enumerate}
-\item \textbf{Example 1: More complex dependencies.} Let us consider a controller application in charge
+\item \textbf{Example 1: More complex dependencies.} Let us consider two controllers 
+in charge
 of \emph{load-balancing}. 
 In a load-balancing application, seemingly independent actions, defined over
-completely independent logical flow spaces,
+completely independent logical flow spaces (say, two flows with source IP address ranges $1*$ and $0*$),
 may in fact be dependent due to collocation in the underlying physical network.
-To be more concrete, say that the objective of the load-balancer
-is to distribute flows equally across links or servers.
-If multiple controllers concurrently and independently
-update forwarding rules (say, two ``independent'' flows with source IP address ranges $1*$ and $0*$), 
+If the two controllers concurrently and independently
+update forwarding rules,  
 they may unvoluntarily
@@
 to modify rules in non-trivial ways.
-In other words, the bundle and overlap check only helps
-for the conditional creation of rules, not their modification.
+In other words, the bundle and overlap check only help
+for the \emph{conditional creation} of rules, not their \emph{modification}.
 \end{enumerate}
@@
 we introduce two new synchronization objects (and describe their implementation in standard Openflow),
-which allows controllers to coordinate 
+which allow controllers to coordinate 
 themselves based on these identifiers.
@@
   $\addr,k\in\Nat$, the value at memory offset $\addr$ is set to $k$. For
-  example, the offset $\addr$ may be used to identify the policy (identifier),
+  example, the offset $\addr$ may correspond to the policy identifier (\emph{pid}),
   but our approach is more general.
 
-\item The operation $\textit{\compare}(\addr,k)$, $\addr,k\in\Nat$,
-aborts current transaction, if the value at memory offset $\addr$ is not equal to $k$.
+\item By performing an operation $\textit{\compare}(\addr,k)$, $\addr,k\in\Nat$,
+a transaction can be aborted if the value at memory offset $\addr$ is not equal to $k$.
 %returns $\textit{true}$ if and
@@
 
-As we will see, the bundle feature is not required to implement this abstraction.
+As we will see, the bundle feature is not required to implement these semantics.
 Also note that memory read operations correspond to reading the switch 
@@
 
-With this memory interface we can immediately implement 
-a powerful \emph{compare-and-set} (CAS) operation which takes two integer parameters
+Given this memory interface, we can immediately implement 
+a powerful \emph{Compare-And-Set} (CAS) operation which takes two integer parameters
 $\textit{old}$ and $\textit{new}$, reads the register and, if the
@@
 returns $\top$; otherwise, $\bot$ is returned and
-we say that the CAS operation \emph{fails} (which causes an abort if
-the transaction invoking it).
+we say that the CAS operation \emph{fails} (which causes an abort of
+the invoking transaction).
 
@@
 \begin{itemize}
-\item With $\textit{claim}(k)$, $i\in[N]$,
-  $k\in\Nat$, a controller $i$ \emph{claims}
+\item With $\textit{claim}(k)$, 
+  $k\in\Nat$, a controller $i\in[N]$ \emph{claims}
   identifier $k$.
 
-\item With $\textit{unclaim}(k)$, $i\in[N]$, $k\in\Nat$, controller $i$
-  \emph{unclaims} identifier $k$;
+\item With $\textit{unclaim}(k)$, $k\in\Nat$, controller $i\in[N]$
+  \emph{unclaims} identifier $k$.
 
@@
 corresponding entry while setting the
-flag \textsf{\checko}, and using different actions ($2$ instead of $1$), thereby inflicting 
+flag \texttt{\checko}, and using different actions ($2$ instead of $1$), thereby inflicting
 a failure in case an entry with overlapping match value exists. 
@@
 which however has no effect on claims and checks of other controllers; hence, the deletion command does not 
-have to be sent as part of a bundle. Therefore, the \textsf{\claimcheck} operation is atomic by nature, 
-similar to the \textsf{claim} and \textsf{unclaim} operations, and is implemented as essentially one command.
+have to be sent as part of a bundle. Therefore, the \texttt{\claimcheck} operation is atomic by nature,
+similar to the \texttt{claim} and \texttt{unclaim} operations, and is implemented as essentially one command.
 
@@
     		\State $action \gets 2$
-    		\State $flag \gets \textsf{OFPFF\_\checko}$
+    		\State $flag \gets \texttt{OFPFF\_\checko}$
     		\State $cmd1\gets \emph{FlowMod}(match, op = \add, flag, action) $
@@
 we try to add it the same way as in the \memwrite operation,  
-but we also set the \textsf{\checko} flag. If the value is not there, 
+but we also set the \texttt{\checko} flag. If the value is not there,
 our action differs from the action of the existing entry, 
@@
 In addition, we present in Algorithm \ref{alg:cas} a simple implementation 
-of \emph{compare-and-set (CAS)}, using the two last operations.
+of \emph{Compare-And-Set (CAS)}, using the two last operations.
 
@@
     		\State $action \gets k$
-    		\State $flag \gets \textsf{OFPFF\_\checko}$
+    		\State $flag \gets \texttt{OFPFF\_\checko}$
     		\State $cmd\gets \emph{FlowMod}(match, op = \add, flag, action) $
@@
 As a second example, let us consider a simple policy composition protocol that
-allows a set of  controllers to concurrently update switch  \emph{policies} , i.e., sets of
+allows a set of  controllers to concurrently update switch  \emph{policies}, i.e., sets of
 effective flow tables rules, under the
@@
 We use increasing numbers as policies
-ids and use them in order to make sure that the new policy, with id
+IDs and use them in order to make sure that the new policy, with id
 $x$, replaces the (what considered as) current policy, with id
-$x-1$. Considering the policies ids, we send the policy installation
+$x-1$. Considering the policies IDs, we send the policy installation
 commands as part of a bundle that also tries to perform
@@
 
-Algorithm~\ref{alg:update} claims the current policy id in order to prevent someone of using it with different policy in the future (during the current update) and in the bundle used to update the policy it also checks that the new policy id that it about to use (chosen from a finite set of ids) wasn't claimed by someone else.
+Algorithm~\ref{alg:update} claims the current policy id in order to prevent someone of using it with different policy in the future (during the current update) and in the bundle used to update the policy it also checks that the new policy id that it about to use (chosen from a finite set of IDs) wasn't claimed by someone else.
 
@@
      		%\State $res \gets $ commit transaction
-     	\Until{$res\neq\texttt{abort}$}
+     	\Until{$res\neq \bot$}
      	%\doWhile{$res=False$}
@@
      		%\State $res \gets $ commit transaction
-     	\Until{$res\neq\texttt{abort}$}
+     	\Until{$res\neq\bot$}
      	%\doWhile{$res=False$}
@@
 		that must match when the command is
-		OFPFC_MODIFY* or OFPFC_\dele*. A value
+		\texttt{ofpfc_{modify}*} or \texttt{ofpfc_{delete}*. A value
 		of 0 indicates no restriction. */
@@
 	uint8_t table_id; /* ID of the table to put the flow in.
-		For OFPFC_\dele_* commands, OFPTT_ALL
+		For OFPFC_DELETE_* commands, OFPTT_ALL
 		can also be used to delete matching
@@
 
-%/* Message structure for OFPT_BUNDLE_CONTROL. */
+%/* Message structure for OFPT_BUNDLE_CTRL. */
 \begin{lstlisting}
@@
     \begin{algorithmic}[1]
-    \Require policy $P$ (a set of FlowMod entry creation commands), three consecutive flow tables ids (per controller) $T_s,T_0,T_1$, active table index $active\in\{0,1\}$.
+    \Require policy $P$ (a set of FlowMod entry creation commands), three consecutive flow tables IDs (per controller) $T_s,T_0,T_1$, active table index $active\in\{0,1\}$.
     %\Ensure installed policy is consistent with previous one
@@
 
-    		\State $op \gets \textsf{OFPGC\_\dele}$
+    		\State $op \gets \texttt{OFPGC\_\dele}$
     		\State $group\_id \gets old$
@@
 
-    		\State $op \gets \textsf{OFPGC\_\add}$
-    		\State $type \gets \textsf{OFPGT\_INDIRECT}$
+    		\State $op \gets \texttt{OFPGC\_\add}$
+    		\State $type \gets \texttt{OFPGT\_INDIRECT}$
     		\State $group\_id \gets new$

### 2015-04-19 inband-sync.tex
small changes and comments

@@
 we say that the CAS operation \emph{fails} (which causes an abort of
-the invoking transaction).
+the invoking transaction). \liron{CAS requieres to run in a bundle while the other operations disscussed here are not. Maybe we should remove CAS to the consensus subsection as it may confuses here}
 
@@
 In addition, we present in Algorithm \ref{alg:cas} a simple implementation 
-of \emph{Compare-And-Set (CAS)}, using the two last operations.
+of \emph{Compare-And-Set (CAS)}, using the two last operations. \liron{meybe move CAS to consensus subsection}
 
@@
 $\textit{apply\_update}()$ function~\cite{cpc}.
+\liron{TODO: fix the description of the composition predicate - receives a policy and return the commands. we also need to say how a normal controller code can be instrumented to use our paradigm}
 
@@
 In the algorithm, the controller first reads the currently installed
-policy, then chooses a new unique policy identifier \emph{pid}, checks whether the new
+policy and its id \emph{pid}, checks whether the new
 policy and the current policy can be composed and, if so, replaces the
@@
 policy (with a different identifier) has been installed, the update
-fails and takes no effect.
+fails and takes no effect. \liron{the term composition doesn't make sense now with the new user update function... }
 
@@
 IDs and use them in order to make sure that the new policy, with id
-$x$, replaces the (what considered as) current policy, with id
-$x-1$. Considering the policies IDs, we send the policy installation
+$pid+1$, replaces the (what considered as) current policy, with id
+$pid$. Considering the policies IDs, we send the policy installation
 commands as part of a bundle that also tries to perform
-$CAS(x-1,x)$, thereby insuring that the commands are
+$CAS(pid,pid+1)$, thereby insuring that the commands are
 performed iff the current policy is still there and preventing other
@@
 ``singleton'' (consisting of one operation only)  transactions
-containing such operations always commit.
+containing such operations always commit. \liron{I think we can ommit this paragraph entirely - no need for bundle when it is not written}
 
@@
 
-It takes a set of rules $U$ (not yet a policy)and proceeds as follows: first, we seek to
- obtain a unique \emph{id}. FIXME: to be continued...
- \textbf{LS: I am not sure if I need to tell every step of the story or maybe it best to explain the main dif from the previous one similarly to what I just wrote next}.
-
-Algorithm~\ref{alg:update} claims the current policy id in order to prevent someone of using it with different policy in the future (during the current update) and in the bundle used to update the policy it also checks that the new policy id that it about to use (chosen from a finite set of IDs) wasn't claimed by someone else.
-
-\hide{
-We compute our new suggested policy by applying the update requests on top of current policy, supporting any kind of requests and policies.
-Then we make a transaction (using the bundle feature) to atomically check that our policy id is not blocked by anyone else, to change the current policy id to ours (an action that would fail if the current policy id is no longer what we are counting of) and to actually configure our new policy.
-
-If one of the actions in the transaction fails we try again.
-There is no progress guaranty for each controller but there is one for the whole system - at least one of the controller will succeed in fulfilling its update requirements.
-}
-
-\textbf{TODO: fix the pseudocode, make the use of afdd/delete
-  operations uniform, make all switch operations transactional}
-
 \liron{for alg \ref{alg:simple-update} we should say that switch configuration = \pid + policy}
 
-\begin{algorithm}[t]
+\begin{algorithm}[h]
     \caption{Policy update with only CAS}
@@
 
+\hide{
+It takes a set of rules $U$ (not yet a policy)and proceeds as follows: first, we seek to
+ obtain a unique \emph{id}. FIXME: to be continued...
+ \textbf{LS: I am not sure if I need to tell every step of the story or maybe it best to explain the main dif from the previous one similarly to what I just wrote next}.
+}
+
+Algorithm~\ref{alg:update} claims the current policy id in order to prevent someone of using it with different policy in the future (during the current update) and in the bundle used to update the policy it also checks that the new policy id that it about to use (chosen from a finite set of IDs) wasn't claimed by someone else.
+
+\hide{
+We compute our new suggested policy by applying the update requests on top of current policy, supporting any kind of requests and policies.
+Then we make a transaction (using the bundle feature) to atomically check that our policy id is not blocked by anyone else, to change the current policy id to ours (an action that would fail if the current policy id is no longer what we are counting of) and to actually configure our new policy.
+
+If one of the actions in the transaction fails we try again.
+There is no progress guaranty for each controller but there is one for the whole system - at least one of the controller will succeed in fulfilling its update requirements.
+}
+
+\textbf{TODO: fix the pseudocode, make the use of add/delete
+  operations uniform, make all switch operations transactional}
+  \liron{what does this TODO mean??}
+
 \liron{for alg \ref{alg:update} we should say that switch configuration = \pid + claims + policy}
 
-\begin{algorithm}[t]
+\begin{algorithm}[h]
     \caption{Advanced policy update}

### 2015-04-21 inband-sync.tex
iteration by stefan according to our skype meeting today

@@
 Concretely, when the commit fails, the switch must generate an error message corresponding to the
-message that could not be applied. The error message must use the \stefan{what is this:} \emph{xid} of the offending message,
+message that could not be applied. The error message must use the identifier (\emph{xid}) of the offending message,
 the error data field corresponding to that message and the error code corresponding to the failure. If
@@
 \item First, it contains a \emph{create bundle} message: a \texttt{ofpt\_bundle\_control} message of type
-\texttt{OFPBCT\_OPEN\_REQUEST}.
+\texttt{ofpbct\_open\_request}.
 \item The follows a sequence of commands (e.g., \emph{FlowMod}), each wrapped in an
@@
 
-Finally, we remark that our solution is only based on flow tables
+\paragraph{Remarks}
+As we will see, in our implementation,
+the switch \emph{configuration} consists of two parts:
+the ``normal'' configuration which is defined by the network policy and
+directly effects the traffic (e.g., forwarding rules), and some additional
+configuration which serves as synchronization information 
+for our mechanism. 
+By installing Openflow \emph{flow monitors}, 
+a controller can read the entire configuration 
+(and subscribe for updates). 
+%
+%can i match multiple flows in multiple tables? yes, but only for FlowMod, not GroupMod: but that is good enough!%
+
+We also remark that our solution is only based on flow tables
 and does not make use
@@
 (unlike the $\textit{\compare}$ operation).
-
-Given this memory interface, we can immediately implement 
-a powerful \emph{Compare-And-Set} (CAS) operation which takes two integer parameters
-$\textit{old}$ and $\textit{new}$, reads the register and, if the
-value of it is $\textit{old}$, replaces it with $\textit{new}$ and
-returns $\top$; otherwise, $\bot$ is returned and
-we say that the CAS operation \emph{fails} (which causes an abort of
-the invoking transaction). \liron{CAS requieres to run in a bundle while the other operations disscussed here are not. Maybe we should remove CAS to the consensus subsection as it may confuses here}
 
@@
     		\State $action \gets 2$
-    		\State $flag \gets \texttt{OFPFF\_\checko}$
+    		\State $flag \gets \texttt{ofpff\_\checko}$
     		\State $cmd1\gets \emph{FlowMod}(match, op = \add, flag, action) $
@@
 with an identical one.
-In addition, we present in Algorithm \ref{alg:cas} a simple implementation 
-of \emph{Compare-And-Set (CAS)}, using the two last operations. \liron{meybe move CAS to consensus subsection}
 
@@
 
-\subsection{Consensus and Replicated State Machines}
-
-Using our approach, it is straightforward to implement consensus among the controllers
+\subsection{CAS and Consensus}
+
+Given our memory interface and using the bundle concept, we can implement
+a powerful \emph{Compare-And-Set} (CAS) operation which takes two integer parameters
+$\textit{old}$ and $\textit{new}$, reads the register and, if the
+value of it is $\textit{old}$, replaces it with $\textit{new}$ and
+returns $\top$; otherwise, $\bot$ is returned and
+we say that the CAS operation \emph{fails} (which causes an abort of
+the invoking transaction). See Algorithm \ref{alg:cas}.
+
+It is then straightforward to implement consensus among the controllers
 using just a single register that can be accessed with an atomic CAS
@@
 
-Given the CAS, we can also implement more complex operations,
+Moreover, given the CAS, we can also implement more complex operations,
 for example the replicated state machine postulated in STN~\cite{stn}:
@@
 
-As a second example, let us consider a simple policy composition protocol that
+As a second example, let us consider a simple policy update protocol that
 allows a set of  controllers to concurrently update switch  \emph{policies}, i.e., sets of
@@
 policy (with a different identifier) has been installed, the update
-fails and takes no effect. \liron{the term composition doesn't make sense now with the new user update function... }
+fails and takes no effect.
+% \liron{the term composition doesn't make sense now with the new user update function... }
 
@@
 
-For brevity, we omit explicitly putting the $\textbf{bundle}\{...\}$ construct
-around each of the transactional operation used in the algorithm.
-Indeed, since \textit{claim} and \textit{unclaim} operations cannot
-cause the corresponding transaction, we can assume that all
-``singleton'' (consisting of one operation only)  transactions
-containing such operations always commit. \liron{I think we can ommit this paragraph entirely - no need for bundle when it is not written}
+%For brevity, we omit explicitly putting the $\textbf{bundle}\{...\}$ construct
+%around each of the transactional operation used in the algorithm.
+%Indeed, since \textit{claim} and \textit{unclaim} operations cannot
+%cause the corresponding transaction, we can assume that all
+%``singleton'' (consisting of one operation only)  transactions
+%containing such operations always commit. \liron{I think we can ommit this paragraph entirely - no need for bundle when it is not written}
 
 Algorithm~\ref{alg:simple-update} allows policy identifiers to grow
-without bound. In fact, as was observed in~\cite{cpc}, the number of currently
+arbitrarily. In fact, as was observed in~\cite{cpc}, the number of currently
 used policies can be made proportional to the number of concurrent

### 2015-04-21 inband-sync.tex
small fix

@@
 themselves based on these identifiers.
+
+As we will see, in the Openflow implementation of our mechanism, 
+in addition to the ``normal'' configuration stored at the switch
+(e.g., forwarding rules according to network policies), 
+controllers will maintain additional information 
+in the switch configuration: this information is used by the
+controllers to synchronize.
 

### 2015-04-22 inband-sync.tex
small changes in sec 4

@@
 
+\section{Discussion of Use Cases}\label{sec:apps}
+
+Let us discuss some use cases of our transactional interface.
+
+\subsection{CAS and Consensus}
+
+Given our memory interface and using the bundle concept, we can implement
+a powerful \emph{Compare-And-Set} (CAS) operation which takes two integer parameters
+$\textit{old}$ and $\textit{new}$, reads the register and, if the
+value of it is $\textit{old}$, replaces it with $\textit{new}$ and
+returns $\top$; otherwise, $\bot$ is returned and
+we say that the CAS operation \emph{fails} (which causes an abort of
+the invoking transaction). As can be seen Algorithm \ref{alg:cas}, CAS is implemented by two commands. These commands should be sent as part of the same bundle
+
 \begin{algorithm}[h]
@@
 \end{algorithm}
-
-\section{Discussion of Use Cases}\label{sec:apps}
-
-Let us discuss some use cases of our transactional interface.
-
-\subsection{CAS and Consensus}
-
-Given our memory interface and using the bundle concept, we can implement
-a powerful \emph{Compare-And-Set} (CAS) operation which takes two integer parameters
-$\textit{old}$ and $\textit{new}$, reads the register and, if the
-value of it is $\textit{old}$, replaces it with $\textit{new}$ and
-returns $\top$; otherwise, $\bot$ is returned and
-we say that the CAS operation \emph{fails} (which causes an abort of
-the invoking transaction). See Algorithm \ref{alg:cas}.
 
@@
 $\textit{apply\_update}()$ function~\cite{cpc}.
-\liron{TODO: fix the description of the composition predicate - receives a policy and return the commands. we also need to say how a normal controller code can be instrumented to use our paradigm}
-
-Algorithm~\ref{alg:simple-update} is based on the bundle feature and associates each policy with a \emph{policy identifier (pid)}.
+Algorithm~\ref{alg:simple-update} uses the bundle feature to update the switch configuration, where the configuration contains both the policy and an abstracted memory address that holds the \emph{policy identifier (pid)} .
 In the algorithm, the controller first reads the currently installed
-policy and its id \emph{pid}, checks whether the new
-policy and the current policy can be composed and, if so, replaces the
-current policy with the composition. If, meanwhile, a new
+policy and its id \emph{pid}, executes the local update function ($\textit{apply\_update}()$) which results with update commands and then tries to apply them atomically with updating the $pid$.
+If, meanwhile, a new
 policy (with a different identifier) has been installed, the update
@@
 
+\liron{we can explain how a normal controller code can be instrumented to use our paradigm}
+
 We use increasing numbers as policies
@@
 $pid+1$, replaces the (what considered as) current policy, with id
-$pid$. Considering the policies IDs, we send the policy installation
+$pid$. Considering the policies IDs, we send the policy update
 commands as part of a bundle that also tries to perform
@@
 %containing such operations always commit. \liron{I think we can ommit this paragraph entirely - no need for bundle when it is not written}
-
-Algorithm~\ref{alg:simple-update} allows policy identifiers to grow
-arbitrarily. In fact, as was observed in~\cite{cpc}, the number of currently
-used policies can be made proportional to the number of concurrent
-controllers.
-
-\liron{for alg \ref{alg:simple-update} we should say that switch configuration = \pid + policy}
 
@@
 
+Algorithm~\ref{alg:simple-update} allows policy identifiers to grow
+arbitrarily. In fact, as was observed in~\cite{cpc}, the number of currently
+used policies can be made proportional to the number of concurrent
+controllers.
+
 \hide{
@@
 
-Algorithm~\ref{alg:update} claims the current policy id in order to prevent someone of using it with different policy in the future (during the current update) and in the bundle used to update the policy it also checks that the new policy id that it about to use (chosen from a finite set of IDs) wasn't claimed by someone else.
+Algorithm~\ref{alg:update} uses bounded number of policy ids by claiming the current policy id preventing someone of using it with different policy in the future (during the current update) and in the bundle used to update the policy it also checks that the new policy id that it about to use (chosen from a finite set of IDs) wasn't claimed by someone else.
 
@@
 }
-
-\textbf{TODO: fix the pseudocode, make the use of add/delete
-  operations uniform, make all switch operations transactional}
-  \liron{what does this TODO mean??}
-
-\liron{for alg \ref{alg:update} we should say that switch configuration = \pid + claims + policy}
 
@@
 \end{algorithm}
+
+Note that in the case of Algorithm \ref{alg:update}, switch configuration is used to store $pid$ and resource claims (both implemented by flow entries) in addition to the policy.
 

### 2015-05-14 inband-sync.tex
all petr's comments are either fixed or added to the tex

@@
 \newcommand{\liron}[1]{\textit{\textcolor{mygreen}{[liron]: #1}}} % Comments
-\newcommand{\petr}[1]{\textit{\textcolor{blue}{[carlos]: #1}}} % Comments
+\newcommand{\petr}[1]{\textit{\textcolor{blue}{[petr]: #1}}} % Comments
 
@@
 \section{Introduction}\label{sec:intro}
-
 By consolidating and outsourcing the control over the dataplane switches to a logically
@@
 (\`{a} la STN~\cite{stn}) modified concurrently by multiple controllers.
+\petr{"interface modified concurrently" does not sound right -> "accessed concurrently"?}
 
@@
 \section{Background}\label{sec:background}
-
+\petr{ the 1st sentence does not belong here}\liron{I think it serves as a motivation for the section, but it fines by me to remove it}
 The transactional interface presented in this paper can be implemented in
@@
 action part. A packet matched by a certain rule will be subject
-to the associated action (to be immediately applied now or later).
+to the associated action (to be immediately applied now or later).\petr{"to be immediately applied now or later" - sounds funny}\liron{I agree it sounds a bit strange, maybe we should explain this later after we disscuss the pipeline}
 For example, an action can define a port to which the
@@
 data packet are compared with the match fields of that rule,
-and if they fit, some instructions (the actions) are executed;
+and if they fit, some instructions (the actions)  are executed; \petr{ what is the difference between instruction or action?} \liron{we should explain that: every packet is associated with an action set.
+This set is empty at the beginning and is changed according to the flow table matches.
+Each match is asociated with instructions that can add or modify actions in the action set and also to apply (now) all actions in the action set.
+By default action set is applied at the end of the pipeline.}
 subsequently, lower priority rules are checked. Depending on
@@
 A flow entry consists of: (1) a match defined by a ternary pattern over packet header fields
-(a value and a mask), and (2) a set of actions.
+(a value and a mask), (2) a set of actions, and (3) a priority.
 %,a match and an action.
@@
 \paragraph{Bundling}
+
 Another interesting feature in Openflow
-is the \emph{bundle}. A bundle
-aggregates multiple control commands (e.g., \emph{FlowMod} messages)
-into one message, with \emph{all-or-nothing semantics}:
+is the \emph{bundle}.
+A bundle allows to send multiple commands (e.g., \emph{FlowMod} messages) with the same bundle id and then execute them  by sending a bundle commit message.
+Theses (bundled) commands are performed with \emph{all-or-nothing semantics}:
 either all of the commands are
@@
 to the switch configuration either \emph{before} or \emph{after} the
-bundle (i.e., all its commands) took effect.
+bundle (i.e., all its commands) took effect. \liron{we need to disscuss controller atomicity and order preservation inside a bundle}
 If multiple \emph{FlowMod} messages with the $\checko$ flag are sent as a single bundle,
@@
 all commands are performed successfully, and otherwise $\bot$.
+\petr{What about replacing $\top-\bot$ with ack-nack? Can we also get a reason why a bundle did not commit?}
 The bundles used in this paper will always have the following structure:
@@
 for our mechanism. 
-By installing Openflow \emph{flow monitors}, 
-a controller can read the entire configuration 
-(and subscribe for updates). 
+By installing Openflow \emph{flow monitors} with the \textsf{OFPFMF\_INITIAL} optional flag,  
+a controller can read the current configuration. 
+% %(and subscribe for updates). 
 %
 %can i match multiple flows in multiple tables? yes, but only for FlowMod, not GroupMod: but that is good enough!%
+\liron{I changed a bit the description of flow monitors following Petr's comment, I hope it's more clear now.}
 
@@
 
+ generic way for multiple distributed controllers to safely update switch configuration with no need of any control plane
+
 At first sight, the Openflow bundle feature seems to provide everything
-we need to implement our transactional interface: the bundle
+we need to implement our transactional interface \petr{did we say what we need already?}: the bundle
  defines a sequence of read and write operations
@@
  limitations, which we discuss in the following.
-
+ \petr{unclear what "reads and writes" are here, not discussed before. How openflow can be used to implement reads and writes?}
+
+\liron{following Petr's comments, I suggest the following paragraph}
+At first sight, the Openflow bundle feature seems to provide a simple and generic way for a controller to atomically update switch configuration, similarly to a transaction in shared memory system.
+However, while the bundle can indeed be seen
+ as a ``mini-transaction'', it comes with several
+ limitations, which we discuss in the following.
+
+\liron{ 
 First, if employed without any additional features, a bundle does not allow
  to change the switch state in any non-trivial way:
-the read and write actions in the bundle do
- not support any conditional \emph{modifications} or computations which
- depend on the read results.
-But also using the bundle together with the check overlap feature ($\checko$)
+the commands in the bundle essentially do
+ not support any conditional \emph{modifications} or computations which depend on any current switch configuration (except to limited types of constraints, e.g., free space).
+ }
+
+But also bundling FlowMod commands with the check overlap feature ($\checko$)
 comes with limitations: while it is possible
- to \emph{abort} transactions depending on certain conditions
+ to \emph{abort} a bundle depending on certain conditions
  in the \emph{overlapping
@@
  to define more general conditions and notions of conflicts
- under which transactions should not take effect. Moreover,
+ under which an atomic update should not take effect.
+ \liron{replaced transaction in the last sentence with " atomic update"}
+
+  Moreover,
  sometimes it is desirable to react to conflicts in smarter ways than
@@
 \item \textbf{Example 2: Supporting modifications.} Consider the problem of modifying previously
-installed forwarding policies.
+installed forwarding rules.
 While a bundle allows to conditionally add new rules or remove
@@
 check feature $\checko$, to abort the transaction if potentially conflicting
-rules were already there), it cannot be used
-to modify rules in non-trivial ways.
+rules were already there), it cannot be easily used
+to conditionally modify existing rules .
 In other words, the bundle and overlap check only help
@@
 
-Our approach is based on the distributed management of \emph{identifiers (IDs)}. In particular,
+Our approach is based on the distributed management of \emph{identifiers (IDs)} \petr{unclear what "identifiers" mean here}. In particular,
 we introduce two new synchronization objects (and describe their implementation in standard Openflow),
@@
 controllers to synchronize. 
-
+\petr{"As we will see,..." - looks funny compared with "As we will see ..." at the end of Sec 2}
 \subsection{Synchronization Objects}\label{sec:t-if}
@@
 
+\petr{it is unclear why exactly we need these write and compare operations}
+
 As we will see, the bundle feature is not required to implement these semantics.
@@
 (unlike the $\textit{\compare}$ operation).
+\petr{read operations "do not affect the execution of the transaction" appears strange, given that a transaction, if committed, "appears to execute sequentially".
+Something seems wrong here...}
 
@@
 requires bundles.
+
+\petr{Would be great actually to motivate why we use reads, writes and claim ids.
+Maybe skecth a solution to some important problems using these primitives in the transactions.} \liron{we mention that its kind of locking.
+but I agree we can motivate it better}
 
@@
 
+\petr{what is "offset" here (offset addr may correspond")?}\liron{I don't understand the comment}
+
 \subsection{Implementation}\label{sec:t-impl}
@@
 
-\paragraph{The resource claimer}
 %features , including
@@
 %]]
-Algorithms~\ref{alg:claim}-\ref{alg:cas} describe implementations of
+Algorithms~\ref{alg:claim}-\ref{alg:compare} describe implementations of
 transactional operations  $\textit{claim}(x)$, $\textit{\claimcheck}(x)$,
-$\textit{unclaim}(x)$, $\textit{write}(x)$,  $\textit{\compare}(x)$
-and $\textit{\cas}(x)$, respectively.
-Algorithm~\ref{alg:commit} describes the procedure executed by
-a transaction at \emph{commit time}.
-
-We describe each operation as a message factory that returns OpenFlow commands which
+$\textit{unclaim}(x)$, $\textit{write}(x)$ and $\textit{\compare}(x)$, respectively.
+%Algorithm~\ref{alg:commit} describes the procedure executed by
+%a transaction at \emph{commit time}.
+
+We describe each operation as a message factory \petr{did not get what is "message factory"} that returns OpenFlow commands which
 should be sent to the switch as part of the same transaction.
@@
 Messages with same bundle ID are considered to belong to same bundle and are performed atomically.
+\liron{I don't think we need all of this bundle staff here. we need to move it to next sextion - policy updates}
 
@@
 Our implementation is based on the \emph{FlowMod} command.
-We use the $\checko$ optional flag to make the transaction execution dependent on switch state.
+We use the $\checko$ optional flag to make the transaction execution dependent on switch state \liron{ last sentence might a bit contradict what we said in the limitation section}.
 We consider the match part of an entry to be a ternary string of bounded length (represented by binary value and mask)
@@
 \emph{set meta-data} instruction where the written value is that integer.
+
+\paragraph{The resource claimer}
+
+\petr{
+the "resource claimed" part a hard read, the trick of using masks and values is smart, deserves to be explained better
+}
+
+\petr{Alg 2 returns "cmd1, cmd2", what does it mean? Isn't it supposed to return a boolean?}
+\petr{what is this "$\concat$" operation? simple concatenation?}
+\petr{sometimes FlowMod commands have flag and action, sometimes not, sometimes flag=0, semantics of all this is unclear.}
 
@@
 \subsection{CAS and Consensus}
-
+\petr{ some motivation of why CAS is important/useful is missing}
 Given our memory interface and using the bundle concept, we can implement
@@
 with a notion of tags.
+
+\petr{when talking about replicated state machines, we should make it clear that we that we talk about replication among controllers.}
+
+\petr{"more complex operations": more complex than what?}
+
+\petr{the discussion of CAS in STN is confusing, given that we never described STN.}
 
@@
 \end{comment}
+
+\petr{this is the first time we talk concretely about "policies" ("forwarding policies"?), but they were mentioned a few times before. Shouldn't we have a short "SDN primer" in Sec 3?}
+
+\petr{did not get what a "composition predicate" is}
+
+\petr{why apply-update is "local" ("local update function")?}
+
+\petr{the 1st paragraph is hard to digest}
+
+\petr{Sec 5.2 relies too heavily on [3]. Do we expect the reader unfamiliar with [3] to get it?}
 

### 2015-05-17 inband-sync.tex
more fixes

@@
 \newcommand{\checko}{\texttt{check\_overlap}\xspace}
-
+\newcommand{\ufunc}{U} %apply\_update
 \newcommand{\stefan}[1]{\textit{\textcolor{red}{[stefan]: #1}}} % Comments
@@
 
-\liron{ 
+\liron{ Also updated the following paragraph:}
 First, if employed without any additional features, a bundle does not allow
@@
  not support any conditional \emph{modifications} or computations which depend on any current switch configuration (except to limited types of constraints, e.g., free space).
- }
 
@@
 %a transaction at \emph{commit time}.
-
+Each operation produces OpenFlow command messages that should be sent to the switch, possibly as part of a bundle. 
+\hide{ 
 We describe each operation as a message factory \petr{did not get what is "message factory"} that returns OpenFlow commands which
@@
 \liron{I don't think we need all of this bundle staff here. we need to move it to next sextion - policy updates}
-
+}
 %The OpenFlow messages that we use (besides )
 
-Our implementation is based on the \emph{FlowMod} command.
-We use the $\checko$ optional flag to make the transaction execution dependent on switch state \liron{ last sentence might a bit contradict what we said in the limitation section}.
-We consider the match part of an entry to be a ternary string of bounded length (represented by binary value and mask)
-which can be applied to any set of supported packet header fields. In our implementation, the $64$bit meta-data field alone is
-already sufficient. Moreover we consider the action part of an entry to be an integer, which in practice can be implemented by a
-\emph{set meta-data} instruction where the written value is that integer.
+%Our implementation is based on the \emph{FlowMod} command. 
+To simplify the algorithms we set the FlowMod command field values with integers, according to the following guidelines:
+\begin{itemize}
+
+\item {\bf  Flag:} we often use the $\checko$ optional flag value to make the command execution dependent on switch state. In other cases, when the flag value remains zero, we omit the assignment.
+
+\item {\bf  Match:} we consider the match part of an entry to be a ternary string of bounded length, represented by binary strings (integers) named value and mask.
+In general, the match can be applied to any set of supported packet header fields.
+In our implementation, the $64$bit meta-data field alone is sufficient.
+We sometimes use the concatenation of two $32$bit integers, denoted by $a\concat b$, to form one match value or mask, where $a\concat b := (a<<32)+b$.
+
+\item {\bf Action:} we consider the action part of an entry to be an integer, which in practice can be implemented by a \emph{set meta-data} instruction where the written value is that integer.
+
+\end{itemize}
+
+In addition we use "op" as shortened form to a field that indicates whether the command should $\add$ or $\dele$ the specified flow entry.
+This corresponds to the FlowMod field named "command" which can receive values OFPFC\_ADD and OFPFC\_DELETE.
 
 \paragraph{The resource claimer}
-
 \petr{
@@
 \petr{sometimes FlowMod commands have flag and action, sometimes not, sometimes flag=0, semantics of all this is unclear.}
+\liron{There is now some text above that should help to understand this part (the resource claimer) and the next one. I hope it does.}
 
@@
 
-    		\State $cmd1\gets \textsc{\compare}((\addr, old) $
-    		\State $cmd2\gets \textsc{\memwrite}((\addr, new) $
+    		\State $cmd1\gets \textsc{\compare}(\addr, old) $
+    		\State $cmd2\gets \textsc{\memwrite}(\addr, new) $
 			\Return $cmd1,cmd2$
@@
 Let us assume that the composition predicate is encoded in a user-specific
-$\textit{apply\_update}()$ function~\cite{cpc}.
+$\textit{\ufunc}()$ function~\cite{cpc}.
 Algorithm~\ref{alg:simple-update} uses the bundle feature to update the switch configuration, where the configuration contains both the policy and an abstracted memory address that holds the \emph{policy identifier (pid)} .
 In the algorithm, the controller first reads the currently installed
-policy and its id \emph{pid}, executes the local update function ($\textit{apply\_update}()$) which results with update commands and then tries to apply them atomically with updating the $pid$.
+policy and its id \emph{pid}, executes the local update function ($\textit{\ufunc}$) which results with update commands and then tries to apply them atomically with updating the $pid$.
 If, meanwhile, a new
@@
     \begin{algorithmic}[1]
-    \Require unsafe policy update function $U$, a predefined switch (memory) address for policy id: $\pid_{\addr}$
+    \Require unsafe policy update function $\ufunc$, a predefined switch (memory) address for policy id: $\pid_{\addr}$
     \Ensure installed policy is consistent with previous one
@@
  			\State $\pid,policy\gets$ \textbf{read} switch configuration
- 			\State $cmds\gets U(policy)$
+ 			\State $update\_cmds\gets \ufunc(policy)$
  			\startTxn
-	 			\State $CAS(\pid_{\addr},\pid,\pid+1)$
-	 			\State \textbf{send} $cmds$ %conditioned on $metadata=my\_a$
+	 			\State $CAS(\pid_{\addr},\pid,\pid+1)$,
+	 			\State $update\_cmds$ %conditioned on $metadata=my\_a$
 	 			%\State update policy selector to $my\_a$
@@
     \begin{algorithmic}[1]
-        \Require unsafe policy update function $U$, a predefined switch (memory) address for policy id: $\pid_{\addr}$, a predefined id space $C$.
+        \Require unsafe policy update function $\ufunc$, a predefined switch (memory) address for policy id: $\pid_{\addr}$, a predefined id space $C$.
     \Ensure installed policy is consistent with previous one
@@
  			\State $my\_id\gets$ choose a number from $C\setminus claims$
- 			\State $cmds\gets U(policy)$
+ 			\State $update\_cmds\gets \ufunc(policy)$
  			\startTxn
- 				\State $\claimcheck(my\_id)$
-	 			\State $CAS(\pid_{\addr}, \pid,my\_id)$
-	 			\State \textbf{send} $cmds$ %conditioned on $metadata=my\_a$
+ 				\State $\claimcheck(my\_id)$,
+	 			\State $CAS(\pid_{\addr}, \pid,my\_id)$,
+	 			\State $update\_cmds$ %conditioned on $metadata=my\_a$
 	 			%\State update policy selector to $my\_a$

### 2015-05-20 inband-sync-20-may.tex
Partial commit, still working on Sec 4

--- 2015-05-17 inband-sync.tex
@@
 
-\liron{following Petr's comments, I suggest the following paragraph}
+\liron{following Petr's comments (regarding last paragraph), I suggest the following paragraph (instead of the old version which is still in the text before)}
 At first sight, the Openflow bundle feature seems to provide a simple and generic way for a controller to atomically update switch configuration, similarly to a transaction in shared memory system.

### 2015-05-21 inband-sync.tex
one more partial update

@@
 
+\newcommand{\error}{\texttt{error}}
+\newcommand{\xid}{\texttt{xid}}
+\newcommand{\ecode}{\texttt{error-code}}
+\newcommand{\bundle}{\textbf{atomic}}
+\newcommand{\ack}{\textit{ack}}
+
 \algblock[<block>]{<start>}{<end>}
@@
 to improve load-balancing, and to reduce overheads,
-control modules are physically distributed.
+control modules operating on the data plane are physically distributed.
 %a redundant control plane
@@
 %]]
-However, a distributed control plane also introduces
-new challenges, and in order to guarantee a consistent network operation,
-the actions of different controllers may need to be synchronized.
+However, %a distributed control plane also introduces
+%new challenges, 
+in order to guarantee \emph{consistency} of network operation, actions
+performed on the data plane by different controllers may need to be \emph{synchronized},  which is a
+highly nontrivial task.
 %[[PK sounds contradicting to the fact that such systems already exist
@@
 %
-A well-known concept to solve synchronization and consistency problems
-in distributed systems are \emph{transactions}: a transaction defines a sequence
-of operations which are performed in an atomic, serializable manner, providing all-or-nothing semantics.
-This paper shows that powerful transactions can also be implemented in software-defined networks,
-using the standard Openflow protocol.
+In this paper, we propose a synchronization
+framework for control plane  based on the well-known concept
+\emph{transactions}.    
+%
+A transaction aggregates a sequence  
+of control messages and specific synchronization primitives, 
+providing all-or-nothing semantics.
+We show that a powerful transactional interface can be implemented \emph{in
+  band}, using the standard Openflow commands on the data-plane configuration.
 %Indeed, recent Openflow versions already come with useful but primitive
 %features, in particular \emph{bundling} and \emph{rule-overlap checks}.
-In particular, we present a
-transactional interface which supports arbitrary atomic
-transactions
-over the switch configuration space: Our transactions
-can include complex notions of conflicts (beyond flow overlap checks),
-as well as computations and multiple controller interactions.
-We discuss applications of our transactional memory interface to
-in-band controller consensus, consistent policy installation, and load-balancing.
+%In particular, we present a
+%transactional interface which supports arbitrary atomic
+%transactions
+%over the switch configuration space: 
+%Our transactions
+%can include complex notions of conflicts (beyond flow overlap checks),
+%as well as computations and multiple controller interactions.
+We discuss how to use our synchronization framework to implement
+control-plane replicated service, consistent policy composition, and load-balancing.
 \end{abstract}
@@
 
-While the logically centralized perspective offered by SDN is attractive,
+While the perspective of \emph{logically centralized} control plane
+offered by SDN is intuitive and attractive,
 there is a wide consensus that
-the control plane should be physically \emph{distributed}.
-First, in order to provide a high availability as well as a minimal degree of
-fault-tolerance, controllers should be redundant~\cite{onix,stn,onos}: a failure
-of one controller can be masked by other controllers. Second, it has also been proposed
-to distribute controllers \emph{spatially}, in order to handle latency-sensitive and
-communication intensive control plane events close to their origin.~\cite{devoflow,kandoo,jukka,disco}
+the control plane should be  \emph{physically distributed}.
+First, in order to provide high availability, 
+%as well as a minimal degree of
+%fault-tolerance, 
+%[[PK fault-tolerance is necessary to provide hight availability?]]
+controllers should be redundant~\cite{onix,stn,onos}: a failure
+of one controller can be masked by other controllers. 
+Second, it has also been proposed to distribute controllers \emph{spatially}, in order to handle latency-sensitive and
+communication intensive 
+%control-plane 
+data-plane
+events close to their origin~\cite{devoflow,kandoo,jukka,disco}
 Third, larger SDNs are likely to be operated by multiple administrators or may even offer
@@
 
-\section{Background}\label{sec:background}
-\petr{ the 1st sentence does not belong here}\liron{I think it serves as a motivation for the section, but it fines by me to remove it}
-The transactional interface presented in this paper can be implemented in
-Openflow~\cite{of-spec}, the standard SDN protocol today.
-% In this section,
-%we provide the necessary background.
-In a nutshell,
-Openflow
-follows a match-action concept: Openflow switches store
-rules (installed by the controller) consisting of a match and an
-action part. A packet matched by a certain rule will be subject
-to the associated action (to be immediately applied now or later).\petr{"to be immediately applied now or later" - sounds funny}\liron{I agree it sounds a bit strange, maybe we should explain this later after we disscuss the pipeline}
-For example, an action can define a port to which the
-matched packet should be forwarded, or add or change a tag
-(a certain part in the packet header).
-Each Openflow switch stores one or multiple flow tables,
-each of which contains a set of rules: the \emph{flow entries}.
-
-Flow
-tables form a pipeline, and flow entries are ordered according
-to priorities: A packet arriving at a switch is first checked by
-the rule of the highest priority in table 0: the fields of the
-data packet are compared with the match fields of that rule,
-and if they fit, some instructions (the actions)  are executed; \petr{ what is the difference between instruction or action?} \liron{we should explain that: every packet is associated with an action set.
-This set is empty at the beginning and is changed according to the flow table matches.
-Each match is asociated with instructions that can add or modify actions in the action set and also to apply (now) all actions in the action set.
-By default action set is applied at the end of the pipeline.}
-subsequently, lower priority rules are checked. Depending on
-the outcome of the table 0 processing, the packet may be sent
-to additional flow tables in the pipeline.
-Instructions can be
-used to define additional tables to be visited (a ``goto instruction''),
-to modify the set of actions to be applied later (either by appending,
-deleting, or modifying actions), or to immediately apply some
-actions to the packet.
-
-During the traversal of
-the flow tables, the packet can be equipped with meta-tags to
+\section{Background: SDN and Openflow}\label{sec:background}
+%\petr{ the 1st sentence does not belong here}\liron{I think it serves as a motivation for the section, but it fines by me to remove it}
+
+%In this section, we recall the basic SDN principles, with a focus on
+%Openflow~\cite{of-spec}, the standard SDN protocol today.
+
+\paragraph{Control and data planes}
+Software-defined networking decouples \emph{control plane}, software
+controllers that configure the network, from the \emph{data plane}, the
+devices that implement the network.  
+% 
+In Openflow~\cite{of-spec}, the \emph{de facto} standard SDN protocol
+today, the data plane consists of \emph{Openflow switches}. 
+In a nutshell, the configuration of an Openflow switch is a set of
+\emph{rules}. 
+A rule is essentially a \emph{match-action} pair:
+the match part of a rule identifies the space of packet headers that are
+subject to the rule (e.g., all packets to a specific destination) and
+the action part identifies how the switch should process the matched
+packets (e.g., forward them to a specific port).
+
+More precisely, the match field of a rule is 
+a  a ternary pattern over packet header fields
+(a value and a mask).
+\petr{example?}
+In general, a packet can be matched against any of its header
+fields, and fields can be wildcarded and sometimes bitmasked.
+\petr{unclear what ``sometimes'' means here}
+%(For instance, the meta-data field is maskable.)
+
+A set of rules, associated with \emph{priority} levels, form \emph{flow
+  entries} in a \emph{flow table}, and the 
+entries are ordered according to their priorities.
+Furthermore, an Openflow switch may maintain multiple flow tables
+forming a \emph{pipeline}.
+Data-plane packets are checked against flow entries in the pipelined
+tables to determine the set of actions the switch should perform on the packet.      
+A data-plane packet arriving at a switch is first checked against
+the rule with the highest priority and in table $0$. 
+If the header of the packet fits the match fields of that rule, 
+a default \emph{instruction} associates the packet the corresponding action.
+Otherwise, the packet is checked against flow  entries with lower
+priorities.  
+Specific instructions can be used to define
+additional tables to be visited (\textit{goto} instructions), to
+modify the set of actions associated with the packet (by adding,
+deleting or modifying actions), or immediately apply some actions in the action set. 
+During the traversal of the flow tables, the packet can also be equipped with \emph{meta-tags} to
 transfer temporary information between tables.
 The meta-data field is part of the header that can be
-inserted or removed from a packet via push and pop actions.
-In general, a packet can be matched against any of its header
-fields, and fields can be wildcarded and sometimes bitmasked.
-(For instance, the meta-data field is maskable.) If no rule
-matches, the packet is dropped.
+inserted or removed from a packet via \emph{push} and \emph{pop}
+actions. 
+\petr{do we need the discussion of multiple tables/pipelining here? I still find it
+  very confusing}
+
+The actions associated with the packet are performed at the end of the pipeline.
+If no rule matches the packet, the packet is dropped.
+
+\liron{we should explain that: every packet is associated with an
+  action set. This set is empty at the beginning and is changed
+  according to the flow table matches. Each match is asociated with
+  instructions that can add or modify actions in the action set and
+  also to apply (now) all actions in the action set. By default action
+  set is applied at the end of the pipeline.}
 
@@
 
-\paragraph{The FlowMod Command}
-The main messages used in Openflow
-to update the switch configuration is called \emph{FlowMod}.
-In a nutshell, \emph{FlowMod} either specifies a new flow entry or
-modifications to existing flow entries.
-A flow entry consists of: (1) a match defined by a ternary pattern over packet header fields
-(a value and a mask), (2) a set of actions, and (3) a priority.
+\paragraph{Interaction between control and data planes: FlowMod commands}
+The flow tables installed on the switches across the data plane
+determine the network \emph{policy}.
+Controllers can change the policy via sending  
+control messages containing \emph{FlowMod commands}.
+In a nutshell, a \emph{FlowMod} command either specifies a new flow entry or
+a modifications to an existing flow entry. \petr{The description below
+  does not explain how a flow entry can be modified.}  
+A flow entry consists of: (1)~a match 
+field
+%defined by a ternary pattern over packet header fields
+(a value and a mask), (2)~a set of actions, and (3)~a priority.
 %,a match and an action.
-The standard behavior is that the new flow entry will simply \emph{replace}
-an existing one if the existing one has exactly the \emph{same match},
-but will be installed
-\emph{in addition} to existing entries if they only
-partially overlap (or define flowspace subsets or supersets).
-The latter can result
-in inconsistencies, especially if the old and the new rule
-have the same priority.
-
-In order to avoid conflicts, \emph{FlowMod} can be used together with a the $\checko$ flag:
-if the flow spaces of an old
-and the new rule overlap,
-that is, if there is an entry with an overlapping match part but a
-different action,
+The standard processing of a FlowMod command received by a switch is
+as follows. 
+If the switch already maintains a flow entry with \emph{exactly} the
+same match, then the new flow entry will simply \emph{replace} it.
+\petr{What if the two rules have different priorities?}
+Otherwise, the flow entry will be installed in addition to existing ones. 
+%but will be installed
+%\emph{in addition} to existing entries if they only
+%partially overlap (or define flowspace subsets or supersets).
+
+If rules with overlapping but not identical match fields have the same priority, 
+inconsistencies can arise, since there is no clearly specified action
+that should be applied to a packet matching the two entries.
+In order to avoid such conflicts, a \emph{FlowMod} command can be equipped with a the $\checko$ flag:
+if the switch maintains is a flow entry with an overlapping but not
+identical match part with the same priority but a different action,
  and if the two rules have the same priority, then
 \emph{FlowMod} will fail.
-
-We will exploit the \emph{FlowMod} command in combination with
-$\checko$ to implement synchronization objects which form
-the basis of our transactional interface.
+\petr{Is ``different action'' important here?}
+The $\checko$ flag in a \emph{FlowMod} command helps resolving simple
+conflicts between controllers, and will be instrumental
+in implementing our synchronization abstractions.
+
+%[[PK not sure this is the right place 
+%implement synchronization objects which form
+%the basis of our transactional interface.
+%]]
 We will use the following notation to create a \emph{FlowMod} message:
 $\FlowMod(\match, \action, \flags)$, $\match$ denotes the match,
-$\action$ the action, and $\flags$ the flags to activate the $\checko$ feature.
+$\action$ the action, and $\flags$ the flags to activate the $\checko$
+feature.
+\petr{Sometimes we also write something like op=add in flowmod, should
+  we specify it here?}  
 The \emph{FlowMod} message (and also the flow entries) include more parameters,
@@
 
+By installing Openflow \emph{flow monitors} with the \textsf{OFPFMF\_INITIAL} optional flag,  
+a controller can read the current configuration. 
+% %(and subscribe for updates). 
+%
+%can i match multiple flows in multiple tables? yes, but only for FlowMod, not GroupMod: but that is good enough!%
+\liron{I changed a bit the description of flow monitors following Petr's comment, I hope it's more clear now.}
+\petr{Not very clear what reading the configuration means: is it done
+  in an atomic way?}
+
+%[[PK should go to the discussion?
+%We also remark that our solution is only based on flow tables
+%and does not make use
+%of \emph{Openflow group tables}, nor do we make use of \emph{cookies}: rules
+%are always identified by exact matches.
+%]]
+
 \paragraph{Bundling}
 
-Another interesting feature in Openflow
-is the \emph{bundle}.
-A bundle allows to send multiple commands (e.g., \emph{FlowMod} messages) with the same bundle id and then execute them  by sending a bundle commit message.
-Theses (bundled) commands are performed with \emph{all-or-nothing semantics}:
-either all of the commands are
-performed, or none of them is. In particular, if, for some reason, the
-configuration request contained in one of the bundled commands cannot
-be performed, all the other requests  are rejected and an error
+Another important feature in Openflow is the \emph{bundle}. 
+A controller can send multiple $\FlowMod$ commands  equipped with 
+the same \emph{bundle identifier}.
+These commands are buffered temporarily by
+the switch until a \emph{bundle commit} message is received from the
+controller. 
+The bundled commands are then performed with \emph{all-or-nothing semantics}:
+either all of the commands are performed, or none of them is. 
+In particular, if a configuration request contained in one of the bundled commands cannot
+be applied (e.g.,  all the command is rejected because of the set
+$\checko$ flag and a conflicting flow entry), all  other commands in
+the bundle are rejected and an error
 message is sent to the controller that issued them.
+%If multiple \emph{FlowMod} messages with the $\checko$ flag are sent as a single bundle,
+The error message contains $\xid$, the identifier  first failed command in the
+bundle, and $ecode$, the \emph{error code} corresponding to the failure.   
+%[[PK too detailed and not very clear
+%Concretely, when the commit fails, the switch must generate an error message corresponding to the
+%message that could not be applied. The error message must use the identifier (\emph{xid}) of the offending message,
+%the error data field corresponding to that message and the error code corresponding to the failure. If
+%multiple messages are generating errors, the switch may return only the first error found or generate
+%multiple error messages for the same bundle.
+%]]
+
 When properly configured,
-the bundled processing can be done in a
-\emph{packet-atomic} way, in the sense that
+a bundle can be processed in a \emph{packet-atomic} way, in the sense that
 each packet is processed according
 to the switch configuration either \emph{before} or \emph{after} the
-bundle (i.e., all its commands) took effect. \liron{we need to disscuss controller atomicity and order preservation inside a bundle}
-If multiple \emph{FlowMod} messages with the $\checko$ flag are sent as a single bundle,
-the bundle command also returns a reason
-why the overlap checks failed: it returns the failed command in the bundle.
-Concretely, when the commit fails, the switch must generate an error message corresponding to the
-message that could not be applied. The error message must use the identifier (\emph{xid}) of the offending message,
-the error data field corresponding to that message and the error code corresponding to the failure. If
-multiple messages are generating errors, the switch may return only the first error found or generate
-multiple error messages for the same bundle.
+bundle (i.e., all its commands) are executed. 
+\liron{we need to disscuss controller atomicity and order preservation inside a bundle}
+\petr{Right! :)}
 
 %:a $\checko$ flag can be defined for each command in the bundle.\liron{these last two sentences are not cleaar}
-
-In this paper, we will use the notation
-%$\textbf{bundle}\{\;\ldots\;\}\rightarrow res$ to refer to represent a
-$\textbf{bundle}\{\;\ldots\;\}$ to refer to a
-sequence of commands sent in a bundle. The bundle returns
-$\top$ if and only if
-all commands are performed successfully, and otherwise $\bot$.
-\petr{What about replacing $\top-\bot$ with ack-nack? Can we also get a reason why a bundle did not commit?}
-The bundles used in this paper will always have the following structure:
-\begin{enumerate}
-\item First, it contains a \emph{create bundle} message: a \texttt{ofpt\_bundle\_control} message of type
-\texttt{ofpbct\_open\_request}.
-\item The follows a sequence of commands (e.g., \emph{FlowMod}), each wrapped in an
-\emph{add to bundle message} (\texttt{ofpt\_bundle\_add\_message}) with the \emph{bundle identifier}.
-\item Finally, a \emph{commit bundle} control message (\texttt{ofpt\_bundle\_control}) of
-type \texttt{ofpbct\_commit\_request}.
-\end{enumerate}
-Again, the bundle always returns a binary result: \emph{error} ($\bot$) or \emph{success} ($\top$).
 
@@
 
-\paragraph{Remarks}
-As we will see, in our implementation,
-the switch \emph{configuration} consists of two parts:
-the ``normal'' configuration which is defined by the network policy and
-directly effects the traffic (e.g., forwarding rules), and some additional
-configuration which serves as synchronization information 
-for our mechanism. 
-By installing Openflow \emph{flow monitors} with the \textsf{OFPFMF\_INITIAL} optional flag,  
-a controller can read the current configuration. 
-% %(and subscribe for updates). 
-%
-%can i match multiple flows in multiple tables? yes, but only for FlowMod, not GroupMod: but that is good enough!%
-\liron{I changed a bit the description of flow monitors following Petr's comment, I hope it's more clear now.}
-
-We also remark that our solution is only based on flow tables
-and does not make use
-of \emph{Openflow group tables}, nor do we make use of \emph{cookies}: rules
-are always identified by exact matches.
-
-\section{Limitations and Motivation}\label{sec:motivation}
-
- generic way for multiple distributed controllers to safely update switch configuration with no need of any control plane
-
-At first sight, the Openflow bundle feature seems to provide everything
-we need to implement our transactional interface \petr{did we say what we need already?}: the bundle
- defines a sequence of read and write operations
- on the switch state which is executed atomically.
- However, while the bundle can indeed be seen
- as a ``mini-transaction'', it comes with several
- limitations, which we discuss in the following.
- \petr{unclear what "reads and writes" are here, not discussed before. How openflow can be used to implement reads and writes?}
-
-\liron{following Petr's comments, I suggest the following paragraph}
-At first sight, the Openflow bundle feature seems to provide a simple and generic way for a controller to atomically update switch configuration, similarly to a transaction in shared memory system.
-However, while the bundle can indeed be seen
- as a ``mini-transaction'', it comes with several
- limitations, which we discuss in the following.
-
-\liron{ Also updated the following paragraph:}
-First, if employed without any additional features, a bundle does not allow
- to change the switch state in any non-trivial way:
-the commands in the bundle essentially do
- not support any conditional \emph{modifications} or computations which depend on any current switch configuration (except to limited types of constraints, e.g., free space).
-
-But also bundling FlowMod commands with the check overlap feature ($\checko$)
-comes with limitations: while it is possible
- to \emph{abort} a bundle depending on certain conditions
- in the \emph{overlapping
- flow spaces}, oftentimes it is desirable to be able
- to define more general conditions and notions of conflicts
- under which an atomic update should not take effect.
- \liron{replaced transaction in the last sentence with " atomic update"}
-
-  Moreover,
- sometimes it is desirable to react to conflicts in smarter ways than
- to simply abort, e.g.,
- by supporting \emph{modifications}.
+\section{Limitations and Motivations}\label{sec:motivation}
+
+% generic way for multiple distributed controllers to safely update switch configuration with no need of any control plane
+
+At first sight, bundling \emph{FlowMod} commands with  the $\checko$
+flag already provides a very powerful synchronization mechanism.
+Indeed, its ``min-transaction'' nature allows a controller 
+to install multiple flow entries in an atomic way, making
+%[[PK we did not say what reads and writes are... 
+%At first sight, the Openflow bundle feature seems to provide everything
+%we need to implement our transactional interface \petr{did we say what we need already?}: the bundle
+% defines a sequence of read and write operations
+% on the switch state which is executed atomically.
+%]]
+However, as is 
+%while the bundle can indeed be seen as a ``mini-transaction'', 
+bundling has important limitations.
+%\petr{unclear what "reads and writes" are here, not discussed before. How openflow can be used to implement reads and writes?}
+
+First, %if employed without any additional features, 
+a bundle does not provide any means to modify the switch configuration 
+based on a generic application-specific condition, depending on the
+current configuration, except to limited types of constraints, e.g.,
+or \emph{overlapping flow spaces} or insufficient free space.
+Indeed, sometimes it is desirable to  define more generic conditions
+under which configuration updates should or should not take place.
+
+%But also bundling FlowMod commands with the check overlap feature ($\checko$)
+%comes with limitations: while it is possible
+% to \emph{abort} a bundle depending on certain conditions
+% in the \emph{overlapping
+% flow spaces}, oftentimes it is desirable to be able
+% to define more general conditions and notions of conflicts
+% under which an atomic update should not take effect.
+% \liron{replaced transaction in the last sentence with " atomic update"}
+
+Second, sometimes it is desirable to react to conflicts in smarter ways than
+ to simply abort, e.g., by supporting \emph{conditional modifications}.
 Let us consider two examples.
@@
 \begin{enumerate}
-\item \textbf{Example 1: More complex dependencies.} Let us consider two controllers
+\item \textbf{Example 1: Handling complex dependencies.} Let us consider two controllers
 in charge
@@
 update forwarding rules,
-they may unvoluntarily
+they may involuntarily
 unbalance the flow allocation, as the managed flows are embedded
@@
 
-\stefan{todo: explain figure}
-
-\item \textbf{Example 2: Supporting modifications.} Consider the problem of modifying previously
-installed forwarding rules.
+Consider the example in the left part of Figure~\ref{fig:CAS-example}. Two controllers,
+$1$ and $2$ read the configuration, try add two more flows to a switch
+with two outgoing links, balancing the load.  
+Imagine that each of the controllers reads the switch configuration and
+learns that currently one flow is forwarded to the left link and two
+flows are forwarded to the right link.
+Without synchronization, the two controllers naturally choose to
+install their flows on the left link, which results in an undesirable
+unbalanced state. Note that, assuming that all these flows are
+completely independent, bundling and $\checko$ features do not really
+help here.
+
+\petr{postpone the explanation of the right part of the figure until we come to CAS?}
+
+\item \textbf{Example 2: Supporting modifications.} 
+Consider the problem of modifying previously installed forwarding rules.
 While a bundle allows to conditionally add new rules or remove
@@
 check feature $\checko$, to abort the transaction if potentially conflicting
-rules were already there), it cannot be easily used
-to conditionally modify existing rules .
+rules were already there), it is not straightforward how to use it to  
+modify existing rules, especially provided a specific condition.
 In other words, the bundle and overlap check only help
@@
 
+Intuitively, one way to address these issues is to use a part of the
+configuration space of the data-plane to store 
+``meta-configuration'', i.e., information on contention and conflicts between
+concurrently installed policy updates. Controllers can read and modify
+this meta-data by applying synchronization primitives. 
+In the following section, we
+describe how this space can be organized and consistently maintained
+by multiple controllers despite of concurrency and conflicts.
+
 \begin{figure*}[t]
@@
 \includegraphics[width=1\columnwidth]{no-sync.png}~\includegraphics[width=1\columnwidth]{sync-with-cas.png}\\
-\caption{\emph{Left:} No synchronization. \emph{Right:} CAS.}
+\caption{\emph{Left:} No synchronization. \emph{Right:}
+  CAS.}\label{fig:CAS-example}
 \end{figure*}
 
-As we will see, we do not even need the bundle feature
-to implement the powerful synchronization objects of our transactional interface.
-However, the bundle feature can still be useful in the context of
-consistent policy updates.
+%[[PK not really discussed?
+%As we will see, we do not even need the bundle feature
+%to implement the powerful synchronization objects of our transactional interface.
+%However, the bundle feature can still be useful in the context of
+%consistent policy updates.
 
@@
 %\section{The Container Object}\label{sec:vision}
-\section{Transaction $\&$ Synchronization}\label{sec:main}
-
+
+\section{Synchronization abstractions}\label{sec:main}
+
+We now describe how our  meta-configuration space at
+an Openflow switch can be implemented and manipulated using our
+synchronization framework.
+
+\subsection{Configuration space}
+
+We distinguish two parts of the configuration space of a switch:
+the ``normal'' configuration which determines the network policy and
+is used to process the data-plane traffic (e.g., forwarding rules), 
+and the ``meta'' configuration that contains information used by the
+controllers to synchronize their actions.
+
+\petr{How to make sure that traffic is not affected by the ``meta'' part?}
+
+\subsection{Transactional interface}
+
+Our framework provides the controllers with a transactional interface to the
+switch configuration.     
+A transaction consists of a sequence of \emph{operations}, which
+should either be executed in its entirety or completely rejected. 
+An operation can be a regular FlowMod command or a synchronization
+primitive, which, as we
+show below translate into specific FlowMod commands operating on the
+switch meta-configuration.
+
+We will use the notation
+%$\textbf{bundle}\{\;\ldots\;\}\rightarrow res$ to refer to represent a
+$\bundle\{op_1,\;\ldots\;op_k\}$ to refer to a
+sequence of commands $op_1,\ldots,op_k$ sent in a bundle.
+%\petr{What about replacing $\top-\bot$ with ack-nack? Can we also get a reason why a bundle did not commit?}
+The bundle construct is implemented as a sequence of the following
+control messages:
+\begin{enumerate}
+\item A \emph{create bundle} message: a \texttt{ofpt\_bundle\_control} message of type
+\texttt{ofpbct\_open\_request} equipped with \textit{bid}, the bundle identifier.
+\item A sequence of commands, each wrapped in an
+\emph{add to bundle message} (\texttt{ofpt\_bundle\_add\_message})
+equipped with \emph{bid}.
+\item Finally, a \emph{commit bundle} control message (\texttt{ofpt\_bundle\_control}) of
+type \texttt{ofpbct\_commit\_request} with \textit{bid}.
+\end{enumerate}
+The construct returns $\ack$ if the bundle \emph{commits} and all commands are performed successfully, and
+$\error(\xid,\ecode)$ received from the switch otherwise.
+%Again, the bundle always returns a binary result: \emph{error} ($\bot$) or \emph{success} ($\top$).
+\petr{Should also provide the reason why the bundle has failed?}
+
+\hide{
 Given this motivation, we now present a more powerful transactional interface.
@@
 which none of its operations takes effect.
+}
+
+\subsection{Synchronization primitives}
 
@@
 \petr{
-the "resource claimed" part a hard read, the trick of using masks and values is smart, deserves to be explained better
+the "resource claimed" part a hard read, the trick of using s and values is smart, deserves to be explained better
 }

### 2015-05-21 inband-sync.tex
i added controller atomicity and order preservation, see that it make sense...

@@
 
-By installing Openflow \emph{flow monitors} with the \textsf{OFPFMF\_INITIAL} optional flag,  
-a controller can read the current configuration. 
+In order to read the current configuration of a switch (the existing flow entries), the controller should send
+%By installing 
+the OpenFlow \emph{add flow monitor} command with the \textsf{OFPFMF\_INITIAL} optional flag.
+%In reply the switch will send all the flow entries defined in the switch.
+In this paper, we indicate sending this message and gathering the responds (the reply messages containing the flow entries) by \emph{read configuration} statement which also returns the configuration.
+%, a controller can read the current configuration. 
 % %(and subscribe for updates). 
@@
 
-When properly configured,
-a bundle can be processed in a \emph{packet-atomic} way, in the sense that
-each packet is processed according
-to the switch configuration either \emph{before} or \emph{after} the
+%When properly configured,
+Bundle execution is considered controller atomic in the sense that all controllers can observe switch states (configurations) that either \emph{before} or \emph{after} the
 bundle (i.e., all its commands) are executed. 
-\liron{we need to disscuss controller atomicity and order preservation inside a bundle}
-\petr{Right! :)}
+By specifying the flag OFPBF\_ATOMIC,  
+a bundle can also be  \emph{packet-atomic}, in a similar sense, which can prevent packets to being processed by intermediate (possibly invalid) configuration during a bundle.
+In addition, by specifying the flag OFPBF\_ORDERED,  bundle commands are executed according to the order they were added to the bundle, thereby allowing to perform dependent actions inside a bundle.
+\liron{i added controller atomicity and order preservation, see that it make sense...}
 

### 2015-05-21 inband-sync.tex
one more

@@
 \newcommand{\ecode}{\texttt{error-code}}
-\newcommand{\bundle}{\textbf{atomic}}
+\newcommand{\exec}{\textbf{execute}}
+\newcommand{\execatomic}{\textbf{execute-atomic}}
 \newcommand{\ack}{\textit{ack}}
@@
 More precisely, the match field of a rule is 
-a  a ternary pattern over packet header fields
+a ternary pattern over packet header fields
 (a value and a mask).
@@
 In a nutshell, a \emph{FlowMod} command either specifies a new flow entry or
-a modifications to an existing flow entry. \petr{The description below
+a modification to an existing flow entry. \petr{The description below
   does not explain how a flow entry can be modified.}  
@@
 %,a match and an action.
-The standard processing of a FlowMod command received by a switch is
+
+A FlowMod command can either add a flow entry or delete 
+The standard processing of a FlowMod $\add$ command received by a switch is
 as follows. 
 If the switch already maintains a flow entry with \emph{exactly} the
-same match, then the new flow entry will simply \emph{replace} it.
+same match and priority level, then the new flow entry will simply \emph{replace} it.
 \petr{What if the two rules have different priorities?}
-Otherwise, the flow entry will be installed in addition to existing ones. 
+Otherwise, the flow entry will be installed in addition to existing
+ones.
+A $\dele$ command simply remove an existing flow entry with the same
+match and priority  or does nothing if such an entry is not present.
+\petr{Is this correct?}
+
 %but will be installed
@@
 We will use the following notation to create a \emph{FlowMod} message:
-$\FlowMod(\match, \action, \flags)$, $\match$ denotes the match,
+$\FlowMod(\match, \add/\dele, \action, \flags)$, where the first field
+$\match$ denotes the match,
+the second introduces 
 $\action$ the action, and $\flags$ the flags to activate the $\checko$
 feature.
+
 \petr{Sometimes we also write something like op=add in flowmod, should
@@
 %By installing 
-the OpenFlow \emph{add flow monitor} command with the \textsf{OFPFMF\_INITIAL} optional flag.
+the OpenFlow \emph{add flow monitor} command with the \textsf{ofpt\_initial} optional flag.
 %In reply the switch will send all the flow entries defined in the switch.
@@
 The error message contains $\xid$, the identifier  first failed command in the
-bundle, and $ecode$, the \emph{error code} corresponding to the failure.   
+bundle, and $\ecode$, the \emph{error code} corresponding to the failure.   
 %[[PK too detailed and not very clear
@@
 
+The bundles used in this paper will always have the following structure:
+\begin{enumerate}
+\item First, a bundle contains a \emph{create bundle} message: a \texttt{ofpt\_bundle\_control} message of type
+\texttt{ofpbct\_open\_request}.
+\item The follows a sequence of commands (e.g., \emph{FlowMod}), each wrapped in an
+\emph{add to bundle message} (\texttt{ofpt\_bundle\_add\_message}) with the \emph{bundle identifier}.
+\item Finally, a \emph{commit bundle} control message (\texttt{ofpt\_bundle\_control}) of
+type \texttt{ofpbct\_commit\_request}.
+\end{enumerate}
+Again, the bundle always returns a binary result: \emph{error} ($\bot$) or \emph{success} ($\top$).
+
 %When properly configured,
@@
 bundle (i.e., all its commands) are executed. 
-By specifying the flag OFPBF\_ATOMIC,  
+By specifying the flag \texttt{OFPBF\_ATOMIC},  
 a bundle can also be  \emph{packet-atomic}, in a similar sense, which can prevent packets to being processed by intermediate (possibly invalid) configuration during a bundle.
-In addition, by specifying the flag OFPBF\_ORDERED,  bundle commands are executed according to the order they were added to the bundle, thereby allowing to perform dependent actions inside a bundle.
+In addition, by specifying the flag \texttt{OFPBF\_ORDERED},  bundle commands are executed according to the order they were added to the bundle, thereby allowing to perform dependent actions inside a bundle.
 \liron{i added controller atomicity and order preservation, see that it make sense...}
@@
 is used to process the data-plane traffic (e.g., forwarding rules), 
-and the ``meta'' configuration that contains information used by the
+and the \emph{meta-configuration} that contains information used by the
 controllers to synchronize their actions.
 
-\petr{How to make sure that traffic is not affected by the ``meta'' part?}
-
-\subsection{Transactional interface}
-
-Our framework provides the controllers with a transactional interface to the
-switch configuration.     
-A transaction consists of a sequence of \emph{operations}, which
+The meta-configuration is implemented as a set of flow entries
+designed in a way that it does not affect the processing of data-plane
+traffic. One way to do it is to set a low priority to
+the meta flow entries and ensure that the regular flow entries
+maintain higher priority levels.
+Alternatively, we could leverage the standard ability of maintaining multiple
+flow tables at an OpenFlow switch, and use a separate flow table for
+the meta-configuration,  making sure that the
+\emph{pipelining} mechanism~\cite{of-spec} will never involve this
+flow table.
+
+%\petr{How to make sure that traffic is not affected by the ``meta'' part?}
+
+\subsection{Synchronization}
+
+Our framework provides the controllers with the ability to execute
+sequences of \emph{operations} on the 
+switch configuration in an \emph{atomic} (transactional) or \emph{non-atomic} mode.     
+In the atomic mode, all operations the sequence (which we also refer
+to as a \emph{transaction}) 
 should either be executed in its entirety or completely rejected. 
-An operation can be a regular FlowMod command or a synchronization
+In the non-atomic mode, no atomicity guarantees are provided. 
+An operation here can be a regular FlowMod command or a synchronization
 primitive, which, as we
-show below translate into specific FlowMod commands operating on the
+show below, translates into a sequence of FlowMod commands operating on the
 switch meta-configuration.
 
-We will use the notation
+We propose the following constructs
 %$\textbf{bundle}\{\;\ldots\;\}\rightarrow res$ to refer to represent a
-$\bundle\{op_1,\;\ldots\;op_k\}$ to refer to a
-sequence of commands $op_1,\ldots,op_k$ sent in a bundle.
-%\petr{What about replacing $\top-\bot$ with ack-nack? Can we also get a reason why a bundle did not commit?}
-The bundle construct is implemented as a sequence of the following
-control messages:
-\begin{enumerate}
-\item A \emph{create bundle} message: a \texttt{ofpt\_bundle\_control} message of type
-\texttt{ofpbct\_open\_request} equipped with \textit{bid}, the bundle identifier.
-\item A sequence of commands, each wrapped in an
-\emph{add to bundle message} (\texttt{ofpt\_bundle\_add\_message})
-equipped with \emph{bid}.
-\item Finally, a \emph{commit bundle} control message (\texttt{ofpt\_bundle\_control}) of
-type \texttt{ofpbct\_commit\_request} with \textit{bid}.
-\end{enumerate}
-The construct returns $\ack$ if the bundle \emph{commits} and all commands are performed successfully, and
-$\error(\xid,\ecode)$ received from the switch otherwise.
+$\exec\{op_1,\;\ldots\;op_k\}$ and $\execatomic\{op_1,\;\ldots\;op_k\}$ 
+which refer to executing a sequence of commands $op_1,\ldots,op_k$ in
+a regular and atomic way, respectively.
+The construct returns $\ack$ if the sequence of operations is performed and all commands are performed successfully, and
+$\error(\xid,\ecode)$ received from the switch otherwise, providing
+the identifier of the first command in the sequence that failed
+together with the corresponding error code. 
 %Again, the bundle always returns a binary result: \emph{error} ($\bot$) or \emph{success} ($\top$).
-\petr{Should also provide the reason why the bundle has failed?}
+%\petr{Should also provide the reason why the bundle has failed?}
 

### 2015-05-21 inband-sync.tex
intro, first version

@@
 
-%\title{Solving Consensus with Openflow}
+%\title{Solving Consensus with OPenFlow}
 
@@
 
-%\title{One-Shot Coordination of Distributed SDN Controllers:\\On the Consensus Power of Openflow}
-
-%\title{Coordinating Distributed SDN Controllers:\\On the Consensus Power of Openflow}
+%\title{One-Shot Coordination of Distributed SDN Controllers:\\On the Consensus Power of OPenFlow}
+
+%\title{Coordinating Distributed SDN Controllers:\\On the Consensus Power of OPenFlow}
 
 %[[PK sorry for stubborness :)
-\title{On the Synchronization Power of the Openflow Data Plane}
+\title{On the Synchronization Power of the OPenFlow Data Plane}
 %\\ {\Large Towards Atomic Transactions In-Band}}
@@
 
-%\title{Synchronizing Distributed SDN Controllers\\{\Large The Openflow Container}}
-
-%\title{Synchronizing Distributed Openflow Controllers\\{\Large Toward More Powerful In-Band Synchronization Objects}}
+%\title{Synchronizing Distributed SDN Controllers\\{\Large The OPenFlow Container}}
+
+%\title{Synchronizing Distributed OPenFlow Controllers\\{\Large Toward More Powerful In-Band Synchronization Objects}}
 
@@
 \begin{abstract}
-Control planes of modern Software-Defined Networks (SDNs)
-are \emph{distributed systems}: to ensure availability and fault-tolerance,
+Control planes of forthcoming Software-Defined Networks (SDNs)
+will be \emph{distributed}: to ensure availability and fault-tolerance,
 to improve load-balancing, and to reduce overheads,
-control modules operating on the data plane are physically distributed.
+control modules operating on the data plane should be physically distributed.
 %a redundant control plane
@@
 We show that a powerful transactional interface can be implemented \emph{in
-  band}, using the standard Openflow commands on the data-plane configuration.
-%Indeed, recent Openflow versions already come with useful but primitive
+  band}, using the standard OpenFlow commands on the data-plane configuration.
+%Indeed, recent OpenFlow versions already come with useful but primitive
 %features, in particular \emph{bundling} and \emph{rule-overlap checks}.
@@
 (hardware) data plane.
-Moreover, Openflow, the standard SDN protocol today, introduces many generalizations,
+Moreover, OpenFlow, the standard SDN protocol today, introduces many generalizations,
 in terms of traffic engineering, definition of flows, as well as in-band network functionalities,
@@
 Second, it has also been proposed to distribute controllers \emph{spatially}, in order to handle latency-sensitive and
-communication intensive 
+communication-intensive 
 %control-plane 
 data-plane
-events close to their origin~\cite{devoflow,kandoo,jukka,disco}
-Third, larger SDNs are likely to be operated by multiple administrators or may even offer
+events close to their origin~\cite{devoflow,kandoo,jukka,disco}.
+Third, larger SDNs are likely to be operated by multiple administrators~\cite{stn} or may even offer
 participatory interfaces where different users can install and trigger policy changes
-concurrently~\cite{participatory,stn}.
+concurrently~\cite{participatory}.
 
@@
 
-The distributed computing literature offers
-a wide range of abstractions
-and techniques to solve synchronization and consistency problems.
-A particularly interesting concept in distributed computing is a \emph{transactional memory}:
-a shared memory which supports atomic transactions, describing a sequence of operations
-to be executed with an all-or-nothing semantics. Transactional memories can be used to
- solve many
-classic distributed computing problems, such as consensus.
-
-By providing more complex and atomic transactions over the
-switch configuration
-space, several synchronization problems can be resolved.
-This paper investigates the implementation of a transactional interface
-(\`{a} la STN~\cite{stn}) modified concurrently by multiple controllers.
-\petr{"interface modified concurrently" does not sound right -> "accessed concurrently"?}
-
-\paragraph{Our Contributions}
-This paper shows how to implement a powerful transactional interface
-for software-defined networks.
-This transactional interface is reminiscent of
-classic transactional memory abstractions~\cite{stm-st95,tm-book}
-and
-can be used to resolve
-synchronization problems of distributed control planes efficiently.
-We show that this transactional interface can be used to implement
-consensus among distributed controllers or to consistently install a policy (also providing
+The concurrent computing literature offers
+a wide range of synchronization abstractions~\cite{HS08-book}. 
+E.g., the popular \emph{transactional
+  memory} abstraction~\cite{stm-st95,tm-book} provides a collection of
+concurrent processes with the ability of aggregating sequences of
+shared-memory operations in atomic \emph{transactions} with
+all-or-nothing semantics.
+
+%[[PK 
+%and techniques to solve synchronization and consistency problems.
+%A particularly interesting concept in distributed computing is a \emph{transactional memory}:
+%a shared memory which supports atomic transactions, describing a sequence of operations
+%to be executed with an all-or-nothing semantics. Transactional memories can be used to
+% solve many
+%classic distributed computing problems, such as consensus.
+%]]
+
+\paragraph{Contributions}
+This paper applies the principle of atomicity in concurrent computing
+to distributed SDN control planes. 
+%
+Extending the recent work on Software Transactional
+Networking~\cite{stn}, we propose synchronization constructs that
+allow a controller to represent multiple configuration commands an
+the data plane as an atomic transaction. 
+%
+If none of the transaction's commands \emph{conflicts} (where conflicts can be
+defined in a general application-specific way) with the current
+configuration, the transaction appears to be executed atomically.
+Otherwise, the transaction is aborted in its entirety and does not
+affect neither traffic nor other controllers.
+
+%By providing more complex and atomic transactions over the
+%switch configuration
+%space, several synchronization problems can be resolved.
+%This paper investigates the implementation of a transactional interface
+%(\`{a} la STN~\cite{stn}) modified concurrently by multiple controllers.
+%\petr{"interface modified concurrently" does not sound right -> "accessed concurrently"?}
+
+We show that powerful synchronization constructs can be implemented
+directly \emph{in OpenFlow}, i.e., using the \emoh{de facto} standard SDN
+protocol. 
+In our implementation, a part of the data-plane configuration space is
+used as a shared memory storing information about  contention and
+conflicts between distributed controllers.
+A transaction can then contain standard OpenFlow control operation
+as well as  \emph{synchronization primitives} operating on this shared
+memory. 
+Our synchronization primitives allow the controllers to define complex notions
+of conflicts that can be used for coordinating configuration updates that
+operate over \emph{independent flow spaces}, but where dependencies are introduced
+in the mapping of the flows to the physical infrastructure. This is
+important, e.g., for load-balancing applications.
+
+%[[PK 
+%This paper shows how to implement a powerful transactional interface
+%for software-defined networks.
+%This transactional interface is reminiscent of
+%classic transactional memory abstractions~\cite{stm-st95,tm-book}
+%and
+%can be used to resolve
+%synchronization problems of distributed control planes efficiently.
+%]]
+
+We then show that, using our abstraction, the controllers can 
+solve several important synchronization problems. 
+We describe an implementation of the fundamental \emph{compare-and-swap} primitive, 
+which in turn can be used to solve consensus and
+implement a generic replicated service.
+We then implement a simple CAS-based concurrent policy-update mechanism. 
+Finally, we use our synchronization abstraction to provide
 the missing link for the read-modify-write object
 postulated in~\cite{cpc}).
-The atomic transactions supported by our interface
-may also define complex notions
-of conditions when transactions stand in conflict and should be serialized:
-For instance, conflicts may be defined for transactions which seemingly
-operate over \emph{independent flow spaces}, but where dependencies are introduced
-in the mapping of the flows to the physical infrastructure. This is attractive,
-e.g., for a load-balancing application.
-
-Interestingly, our transactional interface can be implemented using standard Openflow~\cite{of-spec} primitives,
-which by themselves are of limited use.
-Concretely, we introduce, and implement in standard Openflow,
-two new synchronization objects for managing \emph{distributed identifiers}:
-a \emph{claim-id object}, used for claiming an available identifier in a given
-(bounded or unbounded) set, and a \emph{read-write register object}, used for storing the
-currently used identifier.
 
 We hope that our work can also contribute to the ongoing discussion of what can be implemented
-in-band in Openflow~\cite{compute,reclaim}, by taking a first look
-at the relevant problem of distributed
-synchronization.
+in-band in OpenFlow~\cite{compute,reclaim}, by taking a first look
+at the problem of synchronization.
 
@@
 % we show that powerful synchronization primitives
-%can be implemented in Openflow, allowing controllers to implement
+%can be implemented in OpenFlow, allowing controllers to implement
 %non-trivial atomic transactions, and detect and
@@
 The rest of the paper is organized as follows.
-In Section~\ref{sec:background}, we review the basics of the  Openflow
+In Section~\ref{sec:background}, we review the basics of the  OpenFlow
 protocol and discuss its features which are relevant for our work.
-In Section~\ref{sec:motivation}, we discuss the limitations of the Openflow
+In Section~\ref{sec:motivation}, we discuss the limitations of the OpenFlow
 primitives and motivate the need for transactions.
@@
 %background on our approach. Section~\ref{sec:realization}
-%then discusses the basic implementation of this approach in Openflow.
+%then discusses the basic implementation of this approach in OpenFlow.
 %Section~\ref{sec:extensions} discusses practical issues and extensions.
@@
 
-\section{Background: SDN and Openflow}\label{sec:background}
+\section{Background: SDN and OpenFlow}\label{sec:background}
 %\petr{ the 1st sentence does not belong here}\liron{I think it serves as a motivation for the section, but it fines by me to remove it}
@@
 %In this section, we recall the basic SDN principles, with a focus on
-%Openflow~\cite{of-spec}, the standard SDN protocol today.
+%OpenFlow~\cite{of-spec}, the standard SDN protocol today.
 
@@
 % 
-In Openflow~\cite{of-spec}, the \emph{de facto} standard SDN protocol
-today, the data plane consists of \emph{Openflow switches}. 
-In a nutshell, the configuration of an Openflow switch is a set of
+In OpenFlow~\cite{of-spec}, the \emph{de facto} standard SDN protocol
+today, the data plane consists of \emph{OpenFlow switches}. 
+In a nutshell, the configuration of an OpenFlow switch is a set of
 \emph{rules}. 
@@
 entries are ordered according to their priorities.
-Furthermore, an Openflow switch may maintain multiple flow tables
+Furthermore, an OpenFlow switch may maintain multiple flow tables
 forming a \emph{pipeline}.
@@
 %and does not make use
-%of \emph{Openflow group tables}, nor do we make use of \emph{cookies}: rules
+%of \emph{OpenFlow group tables}, nor do we make use of \emph{cookies}: rules
 %are always identified by exact matches.
@@
 
-Another important feature in Openflow is the \emph{bundle}. 
+Another important feature in OpenFlow is the \emph{bundle}. 
 A controller can send multiple $\FlowMod$ commands  equipped with 
@@
 %[[PK we did not say what reads and writes are... 
-%At first sight, the Openflow bundle feature seems to provide everything
+%At first sight, the OpenFlow bundle feature seems to provide everything
 %we need to implement our transactional interface \petr{did we say what we need already?}: the bundle
@@
 %Before presenting our transactional interface and implementation,
-%let us first discuss the limitations of the Openflow features discussed above.
+%let us first discuss the limitations of the OpenFlow features discussed above.
 
@@
 We now describe how our  meta-configuration space at
-an Openflow switch can be implemented and manipulated using our
+an OpenFlow switch can be implemented and manipulated using our
 synchronization framework.
@@
 Our approach is based on the distributed management of \emph{identifiers (IDs)} \petr{unclear what "identifiers" mean here}. In particular,
-we introduce two new synchronization objects (and describe their implementation in standard Openflow),
+we introduce two new synchronization objects (and describe their implementation in standard OpenFlow),
 which allow controllers to coordinate
@@
 
-As we will see, in the Openflow implementation of our mechanism, 
+As we will see, in the OpenFlow implementation of our mechanism, 
 in addition to the ``normal'' configuration stored at the switch
@@
 %is essentially a transactional memory abstraction which can be implemented
-%in standard Openflow switches.
+%in standard OpenFlow switches.
 %Before presenting the technical details in the next section,%
@@
 %The postulated container object can store an arbitrary number of \emph{items}.
-%As we will see, these items correspond to flow entries on the Openflow switch.
+%As we will see, these items correspond to flow entries on the OpenFlow switch.
 %The same item may even exist multiple times in the container (i.e., the container
@@
 %We will describe later how to implement some of these features
-%in older Openflow versions where they are not
+%in older OpenFlow versions where they are not
 %supported. \textbf{TODO}
@@
 to automatically compose updates originating from different controllers,
-in-band and just using Openflow.
+in-band and just using OpenFlow.
 
@@
 
-\section{Transactions in Openflow}\label{sec:sync}
+\section{Transactions in OpenFlow}\label{sec:sync}
 
@@
 
-In order to solve consensus we suggest to use the atomic multiple entries update capability (which is discussed in Openflow standard).
+In order to solve consensus we suggest to use the atomic multiple entries update capability (which is discussed in OpenFlow standard).
 A shared memory system that supports atomic write to (enough) multiple locations is known to allow the implementation of consensus objects.
@@
 Note however that standard memory map and index/offset to a word, while Flow Table entries are not necessarily indexed.
-However, we assume that each flow entry has some property that allows the operator to specifically reference and update it, for example Openflow entry's cookie.
+However, we assume that each flow entry has some property that allows the operator to specifically reference and update it, for example OpenFlow entry's cookie.
 In cases where this property value can't be determined by the operator, we can use other variable entry properties to store the index, for example we can represent memory value $v$ in offset $i$ by an match-action entry whose match is $in\_port = i$ and action is "forward port $x$".
@@
 In order to allow the packets to validate the consensus we make the following observation: while servers can be abstracted as memory writers, the packets processed by the switch are the readers.
-Openflow standard define an atomic write transaction as being atomic in relation to the processed packets thereby making packet processing a multiple read transaction but one read location per table is allowed.
+OpenFlow standard define an atomic write transaction as being atomic in relation to the processed packets thereby making packet processing a multiple read transaction but one read location per table is allowed.
 This means that if each memory location used in the consensus scheme will be stored in different table then each packet will be able to read all of them.
@@
 Our work also contributes to the ongoing discussion of what can be implemented
-in-band in Openflow~\cite{compute,reclaim}, by assuming the relevant perspective of distributed
+in-band in OpenFlow~\cite{compute,reclaim}, by assuming the relevant perspective of distributed
 synchronization.
@@
 In this paper, we describe transactional abstractions that
-can be implemented using standard Openflow features.
+can be implemented using standard OpenFlow features.
 %
@@
 widely, \emph{multi-modify} transactions. .
-, using a single Openflow switch that accepts
+, using a single OpenFlow switch that accepts
 bundle control messages.

### 2015-05-22 inband-sync.tex
Sec 2-4

@@
 \newcommand{\hide}[1]{}
-\newcommand{\concat}[0]{\oplus}
+\newcommand{\concat}[0]{\cdot}
 \newcommand{\append}[0]{\frown}
@@
 
-%\title{Solving Consensus with OPenFlow}
+%\title{Solving Consensus with OpenFlow}
 
@@
 %[[PK sorry for stubborness :)
-\title{On the Synchronization Power of the OPenFlow Data Plane}
+\title{On the Synchronization Power of the OpenFlow Data Plane}
 %\\ {\Large Towards Atomic Transactions In-Band}}
@@
 We show that powerful synchronization constructs can be implemented
-directly \emph{in OpenFlow}, i.e., using the \emoh{de facto} standard SDN
+directly \emph{in OpenFlow}, i.e., using the \emph{de facto} standard SDN
 protocol. 
@@
 
-More precisely, the match field of a rule is 
-a ternary pattern over packet header fields
-(a value and a mask).
-\petr{example?}
-In general, a packet can be matched against any of its header
-fields, and fields can be wildcarded and sometimes bitmasked.
-\petr{unclear what ``sometimes'' means here}
-%(For instance, the meta-data field is maskable.)
-
-A set of rules, associated with \emph{priority} levels, form \emph{flow
-  entries} in a \emph{flow table}, and the 
-entries are ordered according to their priorities.
+More precisely, the match part of a rule is 
+a ternary pattern over packet \emph{header} fields.
+The pattern is represented as a \emph{value} and a \emph{mask}. 
+In the mask, certain header fields, e.g., port numbers or destination and source
+addresses, can be \emph{wildcarded}, stipulating that their content does
+not affect the match. 
+For examples, \texttt{src=10.1.14.0, ff.ff.ff.f0} matches all packets
+with source addresses in the range \texttt{10.1.14.0}--\texttt{10.1.14.127}.
+Certain field, such as \emph{metadata}, can be arbitrarily
+bitmasked. 
+In our implementations, we are going to use bitmasking of the metadata
+filed for performing conditional operations on  its content.   
+\petr{Example?}
+
+A \emph{configuration} of an OpenFlow switch is represented as a
+collection \emph{flow table}.
+A flow table is a set of \emph{flow entries}, each containing a rule,
+with a match, an action, and a \emph{priority} level, and flow entries
+are ordered according to their priorities.  
+In the simplest setting, a switch maintains one flow table (table 0).  
+A data-plane packet arriving at a switch is first checked against
+the rule with the highest priority and in table $0$. 
+If the header of the packet fits the match fields of that rule, 
+a default \emph{instruction} associates the packet with the corresponding action.
+Otherwise, the packet is checked against flow  entries with lower
+priorities, and if no matching rule is found, the packet is dropped.
+
+%[[PK move pipelining to the discussion?
+\hide{
 Furthermore, an OpenFlow switch may maintain multiple flow tables
@@
 actions.
+
+The actions associated with the packet are performed at the end of the pipeline.
+If no rule matches the packet, the packet is dropped.
 \petr{do we need the discussion of multiple tables/pipelining here? I still find it
   very confusing}
-
-The actions associated with the packet are performed at the end of the pipeline.
-If no rule matches the packet, the packet is dropped.
 
@@
   set is applied at the end of the pipeline.}
+}
 
@@
 
-\paragraph{Interaction between control and data planes: FlowMod commands}
+\paragraph{Interaction between the planes: FlowMod commands}
 The flow tables installed on the switches across the data plane
@@
 In a nutshell, a \emph{FlowMod} command either specifies a new flow entry or
-a modification to an existing flow entry. \petr{The description below
-  does not explain how a flow entry can be modified.}  
-A flow entry consists of: (1)~a match 
-field
-%defined by a ternary pattern over packet header fields
-(a value and a mask), (2)~a set of actions, and (3)~a priority.
-%,a match and an action.
-
-A FlowMod command can either add a flow entry or delete 
+a modification to an existing flow entry.  
+%[[PK already said
+%A flow entry consists of: (1)~a match 
+%field
+%(a value and a mask), (2)~a set of actions, and (3)~a priority.
+%]]
+
+A FlowMod command can either add a flow entry or delete a flow entry.
 The standard processing of a FlowMod $\add$ command received by a switch is
@@
 same match and priority level, then the new flow entry will simply \emph{replace} it.
-\petr{What if the two rules have different priorities?}
 Otherwise, the flow entry will be installed in addition to existing
@@
 \emph{FlowMod} will fail.
-\petr{Is ``different action'' important here?}
 The $\checko$ flag in a \emph{FlowMod} command helps resolving simple
@@
 We will use the following notation to create a \emph{FlowMod} message:
-$\FlowMod(\match, \add/\dele, \action, \flags)$, where the first field
-$\match$ denotes the match,
-the second introduces 
-$\action$ the action, and $\flags$ the flags to activate the $\checko$
+$\FlowMod(\match, op, \action, \flags)$, where 
+%$\match$ denotes the match,
+$op$ is $\add$ or $\dele$, and $\flags$ are used, e.g.,  to activate the $\checko$
 feature.
-
-\petr{Sometimes we also write something like op=add in flowmod, should
-  we specify it here?}  
-The \emph{FlowMod} message (and also the flow entries) include more parameters,
-which however are irrelevant for our implementation: we suggest to
+A standard \emph{FlowMod} message (and also the flow entries) may specify
+more parameters, which are however irrelevant for our implementation, we suggest to
 use the default values.
@@
 In order to read the current configuration of a switch (the existing flow entries), the controller should send
-%By installing 
 the OpenFlow \emph{add flow monitor} command with the \textsf{ofpt\_initial} optional flag.
-%In reply the switch will send all the flow entries defined in the switch.
-In this paper, we indicate sending this message and gathering the responds (the reply messages containing the flow entries) by \emph{read configuration} statement which also returns the configuration.
-%, a controller can read the current configuration. 
-% %(and subscribe for updates). 
-%
-%can i match multiple flows in multiple tables? yes, but only for FlowMod, not GroupMod: but that is good enough!%
-\liron{I changed a bit the description of flow monitors following Petr's comment, I hope it's more clear now.}
-\petr{Not very clear what reading the configuration means: is it done
-  in an atomic way?}
-
+In this paper, we indicate sending this message and gathering the
+responds (the reply messages containing the flow entries) 
+by \textit{read-config} statement which returns the configuration.
+
+\hide{
 %[[PK should go to the discussion?
-%We also remark that our solution is only based on flow tables
-%and does not make use
-%of \emph{OpenFlow group tables}, nor do we make use of \emph{cookies}: rules
-%are always identified by exact matches.
+We also remark that our solution is only based on flow tables
+and does not make use
+of \emph{OpenFlow group tables}, nor do we make use of \emph{cookies}: rules
+are always identified by exact matches.
 %]]
+}
 
@@
 
-Another important feature in OpenFlow is the \emph{bundle}. 
+Another important feature in OpenFlow is \emph{bundling}. 
 A controller can send multiple $\FlowMod$ commands  equipped with 
@@
 The bundled commands are then performed with \emph{all-or-nothing semantics}:
-either all of the commands are performed, or none of them is. 
+either all of them are performed, or none of them is. 
 In particular, if a configuration request contained in one of the bundled commands cannot
-be applied (e.g.,  all the command is rejected because of the set
+be applied (e.g., the command is rejected because of the set
 $\checko$ flag and a conflicting flow entry), all  other commands in
@@
 %If multiple \emph{FlowMod} messages with the $\checko$ flag are sent as a single bundle,
-The error message contains $\xid$, the identifier  first failed command in the
+The error message contains $\xid$, the identifier of the first failed command in the
 bundle, and $\ecode$, the \emph{error code} corresponding to the failure.
+
+A bundle begins with a  \texttt{ofpt\_bundle\_control} message of type
+\texttt{ofpbct\_open\_request} (creating the bundle), wraps
+each of its \emph{FlowMod} command in a message  
+\texttt{ofpt\_bundle\_add\_message} equipped with the \emph{bundle
+  identifier}, and ends with a \texttt{ofpt\_bundle\_control} message of
+type \texttt{ofpbct\_commit\_request} (committing the bundle).
+
 %[[PK too detailed and not very clear
@@
 
-The bundles used in this paper will always have the following structure:
-\begin{enumerate}
-\item First, a bundle contains a \emph{create bundle} message: a \texttt{ofpt\_bundle\_control} message of type
-\texttt{ofpbct\_open\_request}.
-\item The follows a sequence of commands (e.g., \emph{FlowMod}), each wrapped in an
-\emph{add to bundle message} (\texttt{ofpt\_bundle\_add\_message}) with the \emph{bundle identifier}.
-\item Finally, a \emph{commit bundle} control message (\texttt{ofpt\_bundle\_control}) of
-type \texttt{ofpbct\_commit\_request}.
-\end{enumerate}
-Again, the bundle always returns a binary result: \emph{error} ($\bot$) or \emph{success} ($\top$).
+%Again, the bundle always returns a binary result: \emph{error} ($\bot$) or \emph{success} ($\top$).
 
 %When properly configured,
-Bundle execution is considered controller atomic in the sense that all controllers can observe switch states (configurations) that either \emph{before} or \emph{after} the
+Bundle execution is considered \emph{controller-atomic} 
+in the sense that all controllers can observe switch states (configurations) that either \emph{before} or \emph{after} the
 bundle (i.e., all its commands) are executed. 
-By specifying the flag \texttt{OFPBF\_ATOMIC},  
+By specifying the flag \texttt{ofpbf\_atomic},  
 a bundle can also be  \emph{packet-atomic}, in a similar sense, which can prevent packets to being processed by intermediate (possibly invalid) configuration during a bundle.
-In addition, by specifying the flag \texttt{OFPBF\_ORDERED},  bundle commands are executed according to the order they were added to the bundle, thereby allowing to perform dependent actions inside a bundle.
-\liron{i added controller atomicity and order preservation, see that it make sense...}
+In addition, by specifying the flag \texttt{ofpbf\_ordered},  bundle commands are executed according to the order they were added to the bundle, thereby allowing to perform dependent actions inside a bundle.
+%\liron{i added controller atomicity and order preservation, see that it make sense...}
 
@@
 based on a generic application-specific condition, depending on the
-current configuration, except to limited types of constraints, e.g.,
-or \emph{overlapping flow spaces} or insufficient free space.
-Indeed, sometimes it is desirable to  define more generic conditions
-under which configuration updates should or should not take place.
+current configuration, except to limited types of constraints, such as
+overlapping flow spaces or insufficient free space.
+But we might need more generic conditions under which configuration
+updates 
+should or should not take place.
 
@@
 Second, sometimes it is desirable to react to conflicts in smarter ways than
- to simply abort, e.g., by supporting \emph{conditional modifications}.
+by simply aborting an operation, e.g., by supporting \emph{conditional modifications}.
 Let us consider two examples.
@@
 
-\petr{postpone the explanation of the right part of the figure until we come to CAS?}
+%\petr{postpone the explanation of the right part of the figure until we come to CAS?}
 
@@
 
-\subsection{Configuration space}
-
+\paragraph{Configuration space}
+%
 We distinguish two parts of the configuration space of a switch:
@@
 controllers to synchronize their actions.
+Flow entries in the meta-configuration can be referred to as shared
+memory locations, provided a total addressing scheme.
 
@@
 
-\subsection{Synchronization}
-
+\paragraph{Synchronization}
+ %
 Our framework provides the controllers with the ability to execute
@@
 
-\subsection{Synchronization primitives}
+\paragraph{Synchronization primitives}
+%
+\hide{
 
@@
 \subsection{Synchronization Objects}\label{sec:t-if}
-
-We implement transactional access to  two kinds of objects:
-a read-write \emph{memory} abstraction
-object, which we will use for storing the current used policy identifier, and a
-\emph{resource-claimer} object, which allows controllers to claim a resource ID,
-e.g., a policy \emph{identifier} in a
-given (bounded or unbounded) set.
+}
+
+Following the work on consistent policy
+updates~\cite{network-update,stn}, we focus on the consistent use of  \emph{policy
+identifiers}.
+Intuitively, a controller should 
+equip each policy update it wants to install with a distinct
+identifier
+%[[PK do not want to say ``unique'': we also need to motivate claimer-id]] 
+and the installation should succeed only if the identifier of the
+currently installed policy has not changed since the last time the
+controller read it.
+
+Therefore, we provide the following transactional operations.
+A $\textit{write}(\addr,k)$ operation sets the content at memory location
+$\addr$ to $k$. 
+%
+%For  example, the offset $\addr$ may correspond to the policy identifier (\emph{pid}),
+%  but our approach is more general.
+%
+A $\textit{\compare}(\addr,k)$ operation \emph{aborts} the transaction
+if and only if the content at
+$\addr$ is \emph{not} $k$.
+Intuitively, by combining compare and write with a set of regular
+FlowMod commands in a transaction we can make sure that the
+configuration is changed and the current policy identifier is updated
+only if the check operation \emph{succeeds} (does not abort).
+
+To make sure that policy identifiers do not grow without bound~\cite{stn}, we
+also provide a \emph{id-claimer} object.
+The object implements a weak type
+of locking by allowing a controller to claim a \emph{resource identifier} in a
+given set, release an identifier,  and check if a given identifier
+is currently claimed by any controller.
 %While In theory one can use the memory abstraction to implement the resource claimer,
@@
 %
-
-With these base objects, we provide the following shared memory abstraction, with
-write-compare access semantics:
-\begin{itemize}
-\item By performing $\textit{write}(\addr,k)$,
-  $\addr,k\in\Nat$, the value at memory offset $\addr$ is set to $k$. For
-  example, the offset $\addr$ may correspond to the policy identifier (\emph{pid}),
-  but our approach is more general.
-
-\item By performing an operation $\textit{\compare}(\addr,k)$, $\addr,k\in\Nat$,
-a transaction can be aborted if the value at memory offset $\addr$ is not equal to $k$.
-%returns $\textit{true}$ if and
-%  only if $k$ is not currently claimed by any controller; if
-%  $\textit{false}$ is returned, we say that the check operation
-%  \emph{fails} (which causes an abort of the transaction invoking it).
-\end{itemize}
-
-\petr{it is unclear why exactly we need these write and compare operations}
-
-As we will see, the bundle feature is not required to implement these semantics.
-Also note that memory read operations correspond to reading the switch
-configuration, and are simple as they do not affect the execution of the transaction
-(unlike the $\textit{\compare}$ operation).
-\petr{read operations "do not affect the execution of the transaction" appears strange, given that a transaction, if committed, "appears to execute sequentially".
-Something seems wrong here...}
-
-Our resource-claimer object implements a weak type of locking, and exports operations \emph{claim},
+More precisely, the id-claimer object exports three operations \emph{claim},
 \emph{unclaim} and \emph{\claimcheck} with the following sequential
@@
 \item With $\textit{claim}(k)$,
-  $k\in\Nat$, a controller $i\in[N]$ \emph{claims}
-  identifier $k$.
-
-\item With $\textit{unclaim}(k)$, $k\in\Nat$, controller $i\in[N]$
+%  $k\in\Nat$, 
+a controller %$i\in[N]$ 
+\emph{claims}  identifier $k$.
+
+\item With $\textit{unclaim}(k)$, %$k\in\Nat$, 
+the controller %$i\in[N]$
   \emph{unclaims} identifier $k$.
+An identifier $k$ is called \emph{claimed} at a given point of an
+execution if there is a controller that has performed $\textit{claim}(k)$
+but has not yet performed  $\textit{unclaim}(k)$.
 
 \item If $k$ is currently claimed by any controller,
-$\textit{\claimcheck}(k)$, $k\in\Nat$,
+$\textit{\claimcheck}(k)$ %, $k\in\Nat$,
 aborts the current transaction.
@@
 
-Again, none of these individual operations
-requires bundles.
-
-\petr{Would be great actually to motivate why we use reads, writes and claim ids.
-Maybe skecth a solution to some important problems using these primitives in the transactions.} \liron{we mention that its kind of locking.
-but I agree we can motivate it better}
+%As we will see, the bundle feature is not required to implement these semantics.
+%Also note that memory read operations correspond to reading the switch
+%configuration, and are simple as they do not affect the execution of the transaction
+%(unlike the $\textit{\compare}$ operation).
+\petr{read operations "do not affect the execution of the transaction" appears strange, given that a transaction, if committed, "appears to execute sequentially".
+Something seems wrong here...}
 
@@
 
-\petr{what is "offset" here (offset addr may correspond")?}\liron{I don't understand the comment}
-
-\subsection{Implementation}\label{sec:t-impl}
-
+\paragraph{Implementation}\label{sec:t-impl}
+%
 We now describe how to implement our transactional
 abstractions using standard OpenFlow features.
-
-%features , including
-%\emph{bundling}, \emph{conditional application}, and \emph{groups}.
-%[[PK let us leave for the moment
-%We will describe later how to implement some of these features
-%in older OpenFlow versions where they are not
-%supported. \textbf{TODO}
-%]]
-Algorithms~\ref{alg:claim}-\ref{alg:compare} describe implementations of
-transactional operations  $\textit{claim}(x)$, $\textit{\claimcheck}(x)$,
-$\textit{unclaim}(x)$, $\textit{write}(x)$ and $\textit{\compare}(x)$, respectively.
-%Algorithm~\ref{alg:commit} describes the procedure executed by
-%a transaction at \emph{commit time}.
-Each operation produces OpenFlow command messages that should be sent to the switch, possibly as part of a bundle. 
+Essentially, we translate each of the transactional operations into a
+sequence of FlowMod commands.
+
+The implementation of $\exec\{op_1,\ldots,op_k\}$ and
+$\execatomic\{op_1,\ldots,op_k\}$ constructs
+invoked by a control application is simple.  
+First we translate  $op_1,\ldots,op_k$ into a sequence of FlowMod
+commands. If $op_i$ is already a FlowMid command, the translation is 
+trivial. If $op_i$ is a synchronization primitive in
+$\{\textit{claim},  \textit{checkclaim}, \textit{unclaim},
+\textit{write}, \textit{compare}\}$, we use
+Algorithms~\ref{alg:claim}, \ref{alg:check}, \ref{alg:unclaim},
+\ref{alg:write}, and \ref{alg:compare}, respectively, to get the
+corresponding command(s).
+
+To run  $\exec\{op_1,\ldots,op_k\}$, we simply send the commands in
+the resulting sequence, one by one, to the switch. 
+%
+To run  $\execatomic\{op_1,\ldots,op_k\}$, we 
+additionally create a bundle, wrap each of the resulting commands in a
+bundle message with the corresponding bundle identifier, complete it
+with a bundle commit message.
+The we wait until the
+corresponding responses are received. 
+If an error message is received
+it is returned to the control application,
+otherwise $\ack$ is returned.
+
+We now describe Algorithms~\ref{alg:claim}-\ref{alg:write} in more
+detail. 
+Each of the operations generates one or two FlowMod commands. 
+The \textit{match} part in these commands specifies only the $64$bit meta-data
+field in the header space, leaving the remaining fields in default
+values.  
+In defining the value and the mask of the math part, we 
+use notation $a\concat b$ to represent a concatenation of two
+$32$bit integers $a$ and $b$: $a\concat b := (a<<32)+b$.
+We also write \texttt{*} for  a $32$bit 
+string of $1$'s used in a mask.  
+%
+The action part in these commands is defined as an integer, which in
+practice can be implemented by a \emph{set meta-data} instruction where the written value is that integer.
+
 \hide{
-We describe each operation as a message factory \petr{did not get what is "message factory"} that returns OpenFlow commands which
-should be sent to the switch as part of the same transaction.
-We assume that every transaction has a unique identifier and is used to start a \emph{bundle}.
-All messages generated during a transaction invocation, are wrapped by bundle messages carrying
-the bundle's ID and followed by bundle commit message.
-Messages with same bundle ID are considered to belong to same bundle and are performed atomically.
-\liron{I don't think we need all of this bundle staff here. we need to move it to next sextion - policy updates}
-}
-%The OpenFlow messages that we use (besides )
-
-%Our implementation is based on the \emph{FlowMod} command. 
-To simplify the algorithms we set the FlowMod command field values with integers, according to the following guidelines:
+%[[PK too lengthy, partially said before
+To simplify the presentation we set the FlowMod command field values with integers, according to the following guidelines:
 \begin{itemize}
@@
 In addition we use "op" as shortened form to a field that indicates whether the command should $\add$ or $\dele$ the specified flow entry.
-This corresponds to the FlowMod field named "command" which can receive values OFPFC\_ADD and OFPFC\_DELETE.
-
-\paragraph{The resource claimer}
-\petr{
-the "resource claimed" part a hard read, the trick of using s and values is smart, deserves to be explained better
+This corresponds to the FlowMod field named "command" which can receive values \textt{ofpfc\_add} and \texttt{ofpfc\_delete}.
+%]]
 }
 
-\petr{Alg 2 returns "cmd1, cmd2", what does it mean? Isn't it supposed to return a boolean?}
-\petr{what is this "$\concat$" operation? simple concatenation?}
-\petr{sometimes FlowMod commands have flag and action, sometimes not, sometimes flag=0, semantics of all this is unclear.}
-\liron{There is now some text above that should help to understand this part (the resource claimer) and the next one. I hope it does.}
-
-As can be seen in Algorithms \ref{alg:claim}-\ref{alg:unclaim}, for the resource-claimer,
-the match part of the added flow entry represents the claimed ID (the ID is concatenated twice). 
-In addition the mask part of the match (controller ID concatenated with all-ones) varies from one controller
-to another in order to allow multiple controllers to claim the same number without overriding the same flow entry. 
-We can see that for any two resource IDs $x_1$ and $x_2$ and controller IDs $c_1\neq c_2$,
-the match patterns $m_1=(x_1\concat x_1, c_1\concat all-1)$ and $ m_2=(x_2\concat x_2, c_2\concat all-1)$ are not equal,
-even if $x_1=x_2$; therefore, they can both be used to claim and unclaim identifiers without affecting each other.
-
-In order to check if a number is claimed, we try to create the
-corresponding entry while setting the
-flag \texttt{\checko}, and using different actions ($2$ instead of $1$), thereby inflicting
-a failure in case an entry with overlapping match value exists.
-If the check succeeds, we delete the entry. Considering $m_1$ and $m_2$ as defined above,
+\paragraph{The id-claimer}
+
+In Algorithms \ref{alg:claim}-\ref{alg:unclaim}, 
+the match part of the constructed commands represents a concatenation
+of two claimed identifiers. 
+In addition, every controller uses a unique mask part of the match
+defined as a distinct $32$bit controller identifier concatenated with
+\texttt{*}.
+This allows multiple controllers to claim the same number without
+overriding the same flow entry.
+%
+For example, for any two resource identifiers $x_1$ and $x_2$ and controller identifiers $c_1\neq c_2$,
+the match patterns $m_1=(x_1\concat x_1, c_1\concat \texttt{*})$ and $ m_2=(x_2\concat x_2, c_2\concat \texttt{*})$ are distinct,
+even if $x_1=x_2$. 
+Therefore, they can both be used to claim and unclaim identifiers without affecting each other.
+
+In order to check if an identifier is claimed, a controller tries to add the
+a new entry setting the flag \texttt{\checko}, and using a different
+action ($2$ instead of $1$), 
+thereby inflicting a failure in case an entry with overlapping match value exists.
+If the check succeeds, we delete the entry. 
+In the example above above,
 and their respective actions $a_1=x_1\concat c1$ and $a_2=x_2\concat c2$,  they overlap
@@
     		\State $value \gets x\concat x$
-    		\State $mask \gets self\concat all-1$
+    		\State $mask \gets self\concat \texttt{*}$
 	    	\State $match \gets (value,mask)$
@@
     \begin{algorithmic}[1]
-    \Require \emph{self} as the calling controller ID.
+    \Require \emph{self} as the calling controller identifier.
 %    \Ensure
     		\State $value \gets x\concat x$
-    		\State $mask \gets self\concat all-1$
+    		\State $mask \gets self\concat \texttt{*}$
     		\State $match \gets (value,mask)$
@@
     		\State $value \gets x\concat x$
-    		\State $mask \gets self\concat all-1$
+    		\State $mask \gets self\concat \texttt{*}$
     		%\State $cookie \gets x\concat self$
@@
     		\State $value \gets \addr$
-    		\State $mask \gets  all-1$
+    		\State $mask \gets  \texttt{*}. \texttt{*}.$
     		\State $match \gets (value,mask)$
@@
     		\State $value \gets \addr$
-    		\State $mask \gets  all-1$
+    		\State $mask \gets  \texttt{*}.\texttt{*}$
     		\State $match \gets (value,mask)$
@@
 We use increasing numbers as policies
-IDs and use them in order to make sure that the new policy, with id
+identifiers and use them in order to make sure that the new policy, with id
 $pid+1$, replaces the (what considered as) current policy, with id
-$pid$. Considering the policies IDs, we send the policy update
+$pid$. Considering the policies identifiers, we send the policy update
 commands as part of a bundle that also tries to perform
@@
  		\Repeat
- 			\State $\pid,policy\gets$ \textbf{read} switch configuration
+ 			\State $\pid,policy\gets$ \textbf{read-config} %switch configuration
  			\State $update\_cmds\gets \ufunc(policy)$
@@
 
-Algorithm~\ref{alg:update} uses bounded number of policy ids by claiming the current policy id preventing someone of using it with different policy in the future (during the current update) and in the bundle used to update the policy it also checks that the new policy id that it about to use (chosen from a finite set of IDs) wasn't claimed by someone else.
+Algorithm~\ref{alg:update} uses bounded number of policy ids by claiming the current policy id preventing someone of using it with different policy in the future (during the current update) and in the bundle used to update the policy it also checks that the new policy id that it about to use (chosen from a finite set of identifiers) wasn't claimed by someone else.
 
@@
  		\Repeat
-		 	\State $\pid,claims,policy\gets$ \textbf{read} switch configuration
+		 	\State $\pid,claims,policy\gets$ \textbf{read-config} %switch configuration
  			\State $claim(\pid)$
- 			\State $\pid 2,claims,policy\gets$ \textbf{read} switch configuration
+ 			\State $\pid 2,claims,policy\gets$ \textbf{read-config} %switch configuration
  			\If {$\pid\neq \pid 2$}
@@
 	/* Flow actions. */
-	uint8_t table_id; /* ID of the table to put the flow in.
+	uint8_t table_id; /* identifier of the table to put the flow in.
 		For OFPFC_DELETE_* commands, OFPTT_ALL
@@
     \begin{algorithmic}[1]
-    \Require policy $P$ (a set of FlowMod entry creation commands), three consecutive flow tables IDs (per controller) $T_s,T_0,T_1$, active table index $active\in\{0,1\}$.
+    \Require policy $P$ (a set of FlowMod entry creation commands), three consecutive flow tables identifiers (per controller) $T_s,T_0,T_1$, active table index $active\in\{0,1\}$.
     %\Ensure installed policy is consistent with previous one
@@
     		\State $value \gets  old$
-    		\State $mask \gets  all-1$
+    		\State $mask \gets  \texttt{*}$
     		\State $match \gets (value,mask)$

### 2015-05-22 inband-sync.tex
notiation

@@
 $32$bit integers $a$ and $b$: $a\concat b := (a<<32)+b$.
-We also write \texttt{*} for  a $32$bit 
+We also write \texttt{1$\ldots$1} for  a $32$bit 
 string of $1$'s used in a mask.  
@@
 defined as a distinct $32$bit controller identifier concatenated with
-\texttt{*}.
-This allows multiple controllers to claim the same number without
-overriding the same flow entry.
+\texttt{1$\ldots$1}.
+This allows multiple controllers to claim the same identifier without
+overriding existing flow entries.
 %
 For example, for any two resource identifiers $x_1$ and $x_2$ and controller identifiers $c_1\neq c_2$,
-the match patterns $m_1=(x_1\concat x_1, c_1\concat \texttt{*})$ and $ m_2=(x_2\concat x_2, c_2\concat \texttt{*})$ are distinct,
+the match patterns $m_1=(x_1\concat x_1, c_1\concat \texttt{1$\ldots$1})$ and $ m_2=(x_2\concat x_2, c_2\concat \texttt{1$\ldots$1})$ are distinct,
 even if $x_1=x_2$. 
@@
     		\State $value \gets x\concat x$
-    		\State $mask \gets self\concat \texttt{*}$
+    		\State $mask \gets self\concat \texttt{1$\ldots$1}$
 	    	\State $match \gets (value,mask)$
@@
     		\State $value \gets x\concat x$
-    		\State $mask \gets self\concat \texttt{*}$
+    		\State $mask \gets self\concat \texttt{1$\ldots$1}$
     		\State $match \gets (value,mask)$
@@
     		\State $value \gets x\concat x$
-    		\State $mask \gets self\concat \texttt{*}$
+    		\State $mask \gets self\concat \texttt{1$\ldots$1}$
     		%\State $cookie \gets x\concat self$
@@
 
-\paragraph{A memory abstraction}
+\paragraph{Write and compare}
 The memory \memwrite and \compare
@@
     		\State $value \gets \addr$
-    		\State $mask \gets  \texttt{*}. \texttt{*}.$
+    		\State $mask \gets  \texttt{1$\ldots$1}. \texttt{1$\ldots$1}.$
     		\State $match \gets (value,mask)$
@@
     		\State $value \gets \addr$
-    		\State $mask \gets  \texttt{*}.\texttt{*}$
+    		\State $mask \gets  \texttt{1$\ldots$1}.\texttt{1$\ldots$1}$
     		\State $match \gets (value,mask)$
@@
     		\State $value \gets  old$
-    		\State $mask \gets  \texttt{*}$
+    		\State $mask \gets  \texttt{1$\ldots$1}$
     		\State $match \gets (value,mask)$

### 2015-05-22 inband-sync.tex
bundle flags

@@
 not affect the match. 
-For examples, \texttt{src=10.1.14.0, ff.ff.ff.f0} matches all packets
-with source addresses in the range \texttt{10.1.14.0}--\texttt{10.1.14.127}.
-Certain field, such as \emph{metadata}, can be arbitrarily
+Certain field, such as \emph{src}, \emph{dst}, \emph{metadata}, can be arbitrarily
 bitmasked. 
 In our implementations, we are going to use bitmasking of the metadata
-filed for performing conditional operations on  its content.   
-\petr{Example?}
+field for performing conditional operations on  its content.   
+For examples, \texttt{port=10.1.14.0, ff.ff.ff.f0} matches all packets
+with source addresses in the range \texttt{10.1.14.0}--\texttt{10.1.14.127}.
 
@@
 Bundle execution is considered \emph{controller-atomic} 
-in the sense that all controllers can observe switch states (configurations) that either \emph{before} or \emph{after} the
-bundle (i.e., all its commands) are executed. 
-By specifying the flag \texttt{ofpbf\_atomic},  
-a bundle can also be  \emph{packet-atomic}, in a similar sense, which can prevent packets to being processed by intermediate (possibly invalid) configuration during a bundle.
-In addition, by specifying the flag \texttt{ofpbf\_ordered},  bundle commands are executed according to the order they were added to the bundle, thereby allowing to perform dependent actions inside a bundle.
+in the sense that all controllers can only observe switch 
+configurations \emph{before} or \emph{after} the
+bundle commands are executed. 
+We set the flag \texttt{ofpbf\_atomic},  
+making bundles \emph{packet-atomic}, 
+in the sense that every data-plane packet is processed by a configuration
+before or after the bundle, and not by a configuration resulting from
+an incomplete bundle execution.
+We also set the flag \texttt{ofpbf\_ordered},  to make sure that the 
+bundle commands are executed in the order they were added to the
+bundle, thereby respecting dependencies between bundle commands.
 %\liron{i added controller atomicity and order preservation, see that it make sense...}
@@
 
-\begin{enumerate}
-\item \textbf{Example 1: Handling complex dependencies.} Let us consider two controllers
+%\begin{enumerate}
+\vspace{2mm}
+\noindent \textbf{Example 1: Handling complex dependencies.} Let us consider two controllers
 in charge
@@
 
-\item \textbf{Example 2: Supporting modifications.} 
+\vspace{2mm}
+\noindent\textbf{Example 2: Supporting modifications.} 
 Consider the problem of modifying previously installed forwarding rules.
@@
 for the \emph{conditional creation} of rules, not their \emph{modification}.
-\end{enumerate}
 

### 2015-05-22 inband-sync.tex
flow monitor

@@
 
-In order to read the current configuration of a switch (the existing flow entries), the controller should send
+To read the current configuration of a switch (the existing flow entries), the controller should send
 the OpenFlow \emph{add flow monitor} command with the \textsf{ofpt\_initial} optional flag.
-In this paper, we indicate sending this message and gathering the
-responds (the reply messages containing the flow entries) 
-by \textit{read-config} statement which returns the configuration.
+In this paper, we write \textit{read-config} for sending 
+this message, receiving a response, and returning the received configuration.
 

### 2015-05-22 inband-sync.tex
committed up to Sec 4

@@
 
-\section{Discussion of Use Cases}\label{sec:apps}
-
-Let us discuss some use cases of our transactional interface.
-
-\subsection{CAS and Consensus}
-\petr{ some motivation of why CAS is important/useful is missing}
-Given our memory interface and using the bundle concept, we can implement
-a powerful \emph{Compare-And-Set} (CAS) operation which takes two integer parameters
-$\textit{old}$ and $\textit{new}$, reads the register and, if the
-value of it is $\textit{old}$, replaces it with $\textit{new}$ and
-returns $\top$; otherwise, $\bot$ is returned and
-we say that the CAS operation \emph{fails} (which causes an abort of
-the invoking transaction). As can be seen Algorithm \ref{alg:cas}, CAS is implemented by two commands. These commands should be sent as part of the same bundle
+\section{Use Cases}\label{sec:apps}
+
+We now discuss several applications of our transactional interface.
+
+\subsection{Compare-and-Set}
+The pathological scenario described in the example in
+Figure~\ref{fig:CAS-example} can be easily resolved if we put
+policy update in a transactional $\execatomic}$ construct together
+ with a \emph{compare-and-set} (CAS) operation on the policy identifier.
+A $\cas(\addr,\textit{old},\textit{new})$ operation on a memory location $\addr$ 
+checks if the content of $\addr$ is $\textit{old}$ and if so, 
+replaces it with $\textit{new}$ and returns $\true$ (the CAS
+\emh{succeeds}), otherwise it returns $\true$ (the CAS
+\emh{fails} (which causes an abort of
+the invoking transaction). 
+It is straightforward to implement the CAS operation by executing
+$\textit{\compare}(\addr, old)$ followed by $\textit{\memwrite}(\addr,
+new)$ withing the same bundle.
+
+%[[PK said it in the text
+\hide{
+As can be seen Algorithm \ref{alg:cas}, CAS is implemented by two commands. These commands should be sent as part of the same bundle
 
@@
 \end{algorithm}
-
-It is then straightforward to implement consensus among the controllers
-using just a single register that can be accessed with an atomic CAS
-operation~\cite{Her91}. In our transactional setting, the controllers
+%]]
+
+CAS can be used to implement a wide range powerful abstractions
+shared by the controllers, such as queues, sets, hash tables,
+etc.~\cite{Her91} in the \emph{wait-free} manner, i.e., tolerating
+asynchrony and failure of any number of controllers.      
+In our transactional setting, the controllers
 can use CAS to replace the default value of such a register

### 2015-05-22 inband-sync.tex
partial Sec 5

@@
 \newcommand{\ack}{\textit{ack}}
+\newcommand{\true}{\textit{true}}
+\newcommand{\false}{\textit{false}}
 
@@
 
-\section{Use Cases}\label{sec:apps}
+\section{Applications}\label{sec:apps}
 
@@
 
-\subsection{Compare-and-Set}
+\paragraph{Compare-and-Set}
 The pathological scenario described in the example in
 Figure~\ref{fig:CAS-example} can be easily resolved if we put
-policy update in a transactional $\execatomic}$ construct together
+policy update in our transactional $\execatomic$ construct together
  with a \emph{compare-and-set} (CAS) operation on the policy identifier.
@@
 replaces it with $\textit{new}$ and returns $\true$ (the CAS
-\emh{succeeds}), otherwise it returns $\true$ (the CAS
-\emh{fails} (which causes an abort of
+\emph{succeeds}), otherwise it returns $\true$ (the CAS
+\emph{fails} (which causes an abort of
 the invoking transaction). 
@@
 \end{algorithm}
+}
 %]]
 
-CAS can be used to implement a wide range powerful abstractions
-shared by the controllers, such as queues, sets, hash tables,
-etc.~\cite{Her91} in the \emph{wait-free} manner, i.e., tolerating
+A single CAS invocation can be used to solve the fundamental
+problem of fault-tolerant \emph{consensus}~\cite{FLP85} in which the controllers
+need to agree on one of their private inputs, despite failures of an
+arbitrary number of controllers.    
+Indeed, each controller simply performs, within a transaction, a CAS operation to replace the default value
+in a memory location at the meta-configuration of a data-plane switch.
+If the transaction commits, the controller considers itself
+the \emph{winner} and decides on its input, 
+if it aborts (the controller fails in the CAS), it remains to read the
+memory to get the input of the winner and decide on it.
+
+Using multiple CAS invocations, the controllers can implement any
+generic replicated state machine~\cite{Her91,Lam98} on the control plane, 
+in the \emph{wait-free} manner, i.e., tolerating
 asynchrony and failure of any number of controllers.
-In our transactional setting, the controllers
-can use CAS to replace the default value of such a register
-with their consensus input values (e.g., the identifiers of the
-policies that should be applied first). If the corresponding
-singleton transaction commits, the controller considers itself
-the \emph{winner} and decides on its input, if
-it aborts (the controller fails in the CAS) it remains to read the register to get the input of the
-winner to decide.
-
-By using multiple CAS invocations, the controllers can implement any
-generic replicated state machine~\cite{Her91,Lam98}, which will be instrumental in the
-next section.
-
-Moreover, given the CAS, we can also implement more complex operations,
-for example the replicated state machine postulated in STN~\cite{stn}:
-in order to choose tags in a consistent manner, CAS needs to be enhanced
-with a notion of tags.
-
-\petr{when talking about replicated state machines, we should make it clear that we that we talk about replication among controllers.}
-
-\petr{"more complex operations": more complex than what?}
-
-\petr{the discussion of CAS in STN is confusing, given that we never described STN.}
-
-\subsection{Policy Updates}
+%which will be instrumental in the
+%next section.
+
+%Moreover, given the CAS, we can also implement more complex operations,
+%for example the replicated state machine postulated in STN~\cite{stn}:
+%in order to choose tags in a consistent manner, CAS needs to be enhanced%
+%with a notion of tags.
+
+%\petr{when talking about replicated state machines, we should make it clear that we that we talk about replication among controllers.}
+
+%\petr{"more complex operations": more complex than what?}
+
+%\petr{the discussion of CAS in STN is confusing, given that we never described STN.}
+
+\paragraph{Policy Updates}
 

### 2015-05-22 inband-sync.tex
completed first version

@@
 
-\newcommand{\addr}{\emph{addr}\xspace}
-\newcommand{\pid}{\emph{pid}\xspace}
+\newcommand{\addr}{\textit{addr}\xspace}
+\newcommand{\paddr}{\textit{paddr}\xspace}
+\newcommand{\pid}{\textit{pid}\xspace}
 \newcommand{\add}{\textbf{add}\xspace}
@@
 \newcommand{\checko}{\texttt{check\_overlap}\xspace}
-\newcommand{\ufunc}{U} %apply\_update
+\newcommand{\ufunc}{update} %apply\_update
 \newcommand{\stefan}[1]{\textit{\textcolor{red}{[stefan]: #1}}} % Comments
@@
 \algblockdefx[Bundle]{startTxn}{endTxn} %
-[0]{\textbf{bundle}\{} %
+[0]{\textbf{\execatomic}\{} %
 [1][res]{ \} $\rightarrow #1$}
@@
 solve several important synchronization problems. 
-We describe an implementation of the fundamental \emph{compare-and-swap} primitive, 
+We describe an implementation of the fundamental \emph{compare-and-set} primitive, 
 which in turn can be used to solve consensus and
@@
     		\State $action \gets k$
-    		\State $flag \gets \texttt{OFPFF\_\checko}$
+    		\State $flag \gets \texttt{ofpff\_\checko}$
     		\State $cmd\gets \emph{FlowMod}(match, op = \add, flag, action) $
@@
 As a second example, let us consider a simple policy update protocol that
-allows a set of  controllers to concurrently update switch  \emph{policies}, i.e., sets of
-effective flow tables rules, under the
-condition that these rules can be composed with the currently installed
+allows a set of  controllers to concurrently update switch policies, i.e., sets of
+%[[PK policies have been defined in Background]]
+effective flow entries, under the
+condition that these rules can be \emph{composed} with the currently installed
 policy.
+
 Let us assume that the composition predicate is encoded in a user-specific
-$\textit{\ufunc}()$ function~\cite{cpc}.
-Algorithm~\ref{alg:simple-update} uses the bundle feature to update the switch configuration, where the configuration contains both the policy and an abstracted memory address that holds the \emph{policy identifier (pid)} .
+$\textsc{\ufunc}()$ function~\cite{cpc}.
+Algorithm~\ref{alg:simple-update} uses the bundle feature to update
+the switch configuration, where the configuration contains both the
+policy and a meta-configuration memory address $\paddr$ that holds the \emph{policy identifier (pid)} .
 In the algorithm, the controller first reads the currently installed
-policy and its id \emph{pid}, executes the local update function ($\textit{\ufunc}$) which results with update commands and then tries to apply them atomically with updating the $pid$.
-If, meanwhile, a new
-policy (with a different identifier) has been installed, the update
+policy together with its {\pid},  applies the
+$\textsc{\ufunc}()$ function to the current policy which results in a
+set of update FlowMod commands and then tries to apply them 
+atomically together with a CAS operation on $\paddr$ replacing $\pid$
+with $\pid+1$. 
+The latter ensures that if, concurrently, a new policy (with a different identifier) has been installed, the update
 fails and takes no effect.
@@
 \liron{we can explain how a normal controller code can be instrumented to use our paradigm}
-
-We use increasing numbers as policies
-identifiers and use them in order to make sure that the new policy, with id
-$pid+1$, replaces the (what considered as) current policy, with id
-$pid$. Considering the policies identifiers, we send the policy update
-commands as part of a bundle that also tries to perform
-$CAS(pid,pid+1)$, thereby insuring that the commands are
-performed iff the current policy is still there and preventing other
-controllers to replace the new policy with another without observing it.
+\petr{did not get the comment}
+
+%[[PK clear from the algorithm?]]
+% and use them in order to make sure that the new policy, with id
+%$pid+1$, replaces the (what considered as) current policy, with id
+%$pid$. 
+%Considering the policies identifiers, we send the policy update
+%commands as part of a bundle that also tries to perform
+%$CAS(pid,pid+1)$, thereby insuring that the commands are
+%performed iff the current policy is still there and preventing other
+%controllers to replace the new policy with another without observing it.
+
+\petr{what does it mean for an update function to be ``unsafe''?}
 
@@
     \begin{algorithmic}[1]
-    \Require unsafe policy update function $\ufunc$, a predefined switch (memory) address for policy id: $\pid_{\addr}$
+    \Require unsafe policy update function $\ufunc$, policy id $\addr$
     \Ensure installed policy is consistent with previous one
@@
      		%\State $res \gets $ commit transaction
-     	\Until{$res\neq \bot$}
+     	\Until{$res=\ack$}
      	%\doWhile{$res=False$}
@@
 
-Algorithm~\ref{alg:simple-update} allows policy identifiers to grow
-arbitrarily. In fact, as was observed in~\cite{cpc}, the number of currently
-used policies can be made proportional to the number of concurrent
+Note that in Algorithm~\ref{alg:simple-update}, policies identifiers
+grow without bound.
+However, as was observed in~\cite{cpc}, the number of currently
+used identifiers can be made proportional to the number of concurrent
 controllers.
-
+%
 \hide{
@@
 }
-
-Algorithm~\ref{alg:update} uses bounded number of policy ids by claiming the current policy id preventing someone of using it with different policy in the future (during the current update) and in the bundle used to update the policy it also checks that the new policy id that it about to use (chosen from a finite set of identifiers) wasn't claimed by someone else.
+%
+
+Algorithm~\ref{alg:update} uses our id-claimer abstraction 
+to bound the number of policy identifiers.
+Claiming a policy id prevents another controller from using it with a
+different policy.
+A controller claims the current policy identifier, chooses a new id
+from a finite set of identifiers,   
+and then checks that the new policy id is not claimed by another controller.
+Note that in Algorithm \ref{alg:update}, the meta-configuration is
+used to store $pid$ and id-claimer.
+\petr{The description is not very clear.}
 
 \hide{
-We compute our new suggested policy by applying the update requests on top of current policy, supporting any kind of requests and policies.
-Then we make a transaction (using the bundle feature) to atomically check that our policy id is not blocked by anyone else, to change the current policy id to ours (an action that would fail if the current policy id is no longer what we are counting of) and to actually configure our new policy.
+We compute our new suggested policy by applying the update requests on
+top of current policy, supporting any kind of requests and
+policies. Then we make a transaction (using the bundle feature) to
+atomically check that our policy id is not blocked by another controller, to change the current policy id to ours (an action that would fail if the current policy id is no longer what we are counting of) and to actually configure our new policy.
 
@@
 		 	\State $\pid,claims,policy\gets$ \textbf{read-config} %switch configuration
- 			\State $claim(\pid)$
+ 			\State $\textbf{execute}\{claim(\pid)\}$
  			\State $\pid 2,claims,policy\gets$ \textbf{read-config} %switch configuration
  			\If {$\pid\neq \pid 2$}
-	 			\State $unclaim(\pid)$
+	 			\State $\textbf{execute} \{unclaim(\pid)\}$
  				\State {\bf continue} (\textbf{restart} loop)
@@
  			\endTxn
-	 		\State $unclaim(\pid)$
+	 		\State $\textbf{execute} \{unclaim(\pid)\}$
      		%\State $res \gets $ commit transaction
-     	\Until{$res\neq\bot$}
+     	\Until{$res=\ack$}
      	%\doWhile{$res=False$}
@@
 \end{algorithm}
-
-Note that in the case of Algorithm \ref{alg:update}, switch configuration is used to store $pid$ and resource claims (both implemented by flow entries) in addition to the policy.
 
@@
 \end{comment}
-
-\petr{this is the first time we talk concretely about "policies" ("forwarding policies"?), but they were mentioned a few times before. Shouldn't we have a short "SDN primer" in Sec 3?}
-
-\petr{did not get what a "composition predicate" is}
-
-\petr{why apply-update is "local" ("local update function")?}
-
-\petr{the 1st paragraph is hard to digest}
-
-\petr{Sec 5.2 relies too heavily on [3]. Do we expect the reader unfamiliar with [3] to get it?}
 
@@
 Our work also contributes to the ongoing discussion of what can be implemented
-in-band in OpenFlow~\cite{compute,reclaim}, by assuming the relevant perspective of distributed
-synchronization.
+in-band in OpenFlow~\cite{compute,reclaim}, by focusing on the
+relevant problem of synchronization.
 
@@
 
-\section{Conclusion}\label{sec:conclusion}
+\section{Concluding Remarks}\label{sec:conclusion}
 
@@
 in an SDN.
+
+Even though we believe that we propose efficient and consistent
+transactional framework for in-band synchronization, we do not claim
+that we offer an exhaustive set of synchronization primitives. Indeed,
+one may explore standard OpenFlow features to enable more powerful
+primitives. For instance, \emph{cookies}~\cite{openflow} can be used for modifying
+flow tables, based on more sophisticated conditions than the default
+``exact match''. 
+Also, the standard pipelining feature allows the switches to maintain multiple flow
+tables which can be traversed in arbitrary patterns, defining the sets
+of corresponding actions. 
+Exploring the power of such features is left to future work.
 

### 2015-05-25 inband-sync.tex
made a fix in implementation section.

@@
 a ternary pattern over packet \emph{header} fields.
-The pattern is represented as a \emph{value} and a \emph{mask}. 
+The pattern is represented as a \emph{value} and a \emph{mask}. \liron{we need to discuss about this, I am not sure this description is accurate.}
 In the mask, certain header fields, e.g., port numbers or destination and source
@@
 
-We now describe Algorithms~\ref{alg:claim}-\ref{alg:write} in more
+We now describe Algorithms~\ref{alg:claim}-\ref{alg:compare} in more
 detail. 
@@
 If the check succeeds, we delete the entry. 
-In the example above above,
-and their respective actions $a_1=x_1\concat c1$ and $a_2=x_2\concat c2$,  they overlap
-with different actions iff $x_1=x_2$. Therefore, if one is used in \claimcheck after the other used in the claim operation,
+Con sidering, $m_1$ and $m_2$ from the example above,
+with respective actions $a_1=1$ and $a_2=2$,  they overlap
+with different actions iff $x_1=x_2$. Therefore, if $m_1$ is used in \claimcheck after $m_2$ used in the claim operation,
 a conflict is detected iff $x_1=x_2$, as desired.

### 2015-05-26 inband-sync.tex
couple of typos corrected

@@
  			\startTxn
-	 			\State $CAS(\pid_{\addr},\pid,\pid+1)$,
+	 			\State $CAS(\paddr,\pid,\pid+1)$,
 	 			\State $update\_cmds$ %conditioned on $metadata=my\_a$
@@
     \begin{algorithmic}[1]
-        \Require unsafe policy update function $\ufunc$, a predefined switch (memory) address for policy id: $\pid_{\addr}$, a predefined id space $C$.
+        \Require unsafe policy update function $\ufunc$, a predefined switch (memory) address for policy id: $\paddr$, a predefined id space $C$.
     \Ensure installed policy is consistent with previous one
@@
  				\State $\claimcheck(my\_id)$,
-	 			\State $CAS(\pid_{\addr}, \pid,my\_id)$,
+	 			\State $CAS(\paddr, \pid,my\_id)$,
 	 			\State $update\_cmds$ %conditioned on $metadata=my\_a$
@@
 
-There exists a wide consensus that SDN control planes need to be distributed.~\cite{onos,onix,elasticon}
+There exists a wide consensus that SDN control planes need to be
+distributed~\cite{onos,onix,elasticon}
 Onix~\cite{onix} is one of the earliest proposals, and introduced the

### 2015-05-26 inband-sync.tex
typos again
typos again, sorry

@@
 The pattern is represented as a \emph{value} and a \emph{mask}. \liron{we need to discuss about this, I am not sure this description is accurate.}
-In the mask, certain header fields, e.g., port numbers or destination and source
+In the mask, certain header fields, e.g., TCP or UDP port numbers or destination and source
 addresses, can be \emph{wildcarded}, stipulating that their content does
 not affect the match. 
-Certain field, such as \emph{src}, \emph{dst}, \emph{metadata}, can be arbitrarily
+Certain fields, such as \emph{ipv4\_src }, \emph{ipv4\_dst}, \emph{metadata}, can be arbitrarily
 bitmasked. 
@@
 field for performing conditional operations on  its content.   
-For examples, \texttt{port=10.1.14.0, ff.ff.ff.f0} matches all packets
-with source addresses in the range \texttt{10.1.14.0}--\texttt{10.1.14.127}.
+For example, \texttt{ipv4\_src=10.1.14.0, ff.ff.ff.f0} matches all packets
+with source IP addresses in the range \texttt{10.1.14.0}--\texttt{10.1.14.127}.
 
@@
     		\State $value \gets \addr$
-    		\State $mask \gets  \texttt{1$\ldots$1}. \texttt{1$\ldots$1}.$
+    		\State $mask \gets  \texttt{1$\ldots$1}. \texttt{1$\ldots$1}$
     		\State $match \gets (value,mask)$

### 2015-05-26 inband-sync.tex
typos

@@
 \emph{Synchronizing}
-the distributed controllers and manipulating the network state \emph{consistently}~\cite{cpc}.
+the distributed controllers and manipulating the network state \emph{consistently}~\cite{cpc}
 are non-trivial tasks~\cite{cap-theorem}.

### 2015-05-26 inband-sync.tex
a pass on intro and CAS

@@
 directly \emph{in OpenFlow}, i.e., using the \emph{de facto} standard SDN
-protocol. 
+protocol. This allows us to efficiently alleviate the problems related to
+coordinating controllers via a control-plane
+out-of-band network, in the presence of asynchrony and failures. 
+Indeed, the inherent costs  of such protocols are often considered too high, both in
+terms of the necessary computability assumptions about the underlying
+system~\cite{FLP85,CHT96,cap-theorem}, and the amount of communication
+complexity~\cite{Lam98,Lam06}.
+In contrast, our \emph{in-band} solution allows the
+controllers to solve fundamental agreement tasks in just one message
+exchange with the data plane, tolerating asynchrony and failures of
+any number of controllers.
+
 In our implementation, a part of the data-plane configuration space is
@@
 solve several important synchronization problems. 
-We describe an implementation of the fundamental \emph{compare-and-set} primitive, 
+We describe an implementation of the fundamental
+\emph{compare-and-set} (CAS) primitive, 
 which in turn can be used to solve consensus and
-implement a generic replicated service.
+implement a generic replicated service~\cite{Her91,Lam98} in a consistent and
+fault-tolerant way, which became a mainstream building block in
+modern system design~\cite{bigtable,spanner,zookeeper}.  
 We then implement a simple CAS-based concurrent policy-update mechanism. 
@@
 the missing link for the read-modify-write object
-postulated in~\cite{cpc}).
+postulated in~\cite{cpc}). \petr{We do not seem to discuss the latter?}
 
@@
 ones.
-A $\dele$ command simply remove an existing flow entry with the same
+A $\dele$ command simply removes an existing flow entry with the same
 match and priority  or does nothing if such an entry is not present.
@@
 
-A single CAS invocation can be used to solve the fundamental
-problem of fault-tolerant \emph{consensus}~\cite{FLP85} in which the controllers
+CAS can be used to solve the fundamental
+problem of fault-tolerant \emph{consensus}~\cite{FLP85}
+\hide{%[[PK, feels like we do not really need this level of detail here
+ in which the controllers
 need to agree on one of their private inputs, despite failures of an
@@
 memory to get the input of the winner and decide on it.
-
-Using multiple CAS invocations, the controllers can implement any
-generic replicated state machine~\cite{Her91,Lam98} on the control plane, 
-in the \emph{wait-free} manner, i.e., tolerating
+}
+and, more generally, 
+to implement any generic replicated state machine~\cite{Her91,Lam98} 
+in the control plane,  in the \emph{wait-free} manner, i.e., tolerating
 asynchrony and failure of any number of controllers.
@@
     \begin{algorithmic}[1]
-    \Require unsafe policy update function $\ufunc$, policy id $\addr$
+    \Require unsafe policy update function \textsc{\ufunc}, policy id $\addr$
     \Ensure installed policy is consistent with previous one
@@
  			\State $\pid,policy\gets$ \textbf{read-config} %switch configuration
- 			\State $update\_cmds\gets \ufunc(policy)$
+ 			\State $update\_cmds\gets \textsc{\ufunc}(policy)$
  			\startTxn
@@
     \begin{algorithmic}[1]
-        \Require unsafe policy update function $\ufunc$, a predefined switch (memory) address for policy id: $\paddr$, a predefined id space $C$.
+        \Require unsafe policy update function \textsc{\ufunc}, a predefined switch (memory) address for policy id: $\paddr$, a predefined id space $C$.
     \Ensure installed policy is consistent with previous one
@@
  			\State $my\_id\gets$ choose a number from $C\setminus claims$
- 			\State $update\_cmds\gets \ufunc(policy)$
+ 			\State $update\_cmds\gets \textsc{\ufunc}(policy)$
  			\startTxn

### 2015-05-29 inband-sync.tex
added load balancing example pseudo code
only intermediate push

@@
 performed on the data plane by different controllers may need to be \emph{synchronized},  which is a
-highly nontrivial task.
+ nontrivial task.
 %[[PK sounds contradicting to the fact that such systems already exist
@@
 In this paper, we propose a synchronization
-framework for control plane  based on the well-known concept
+framework for control planes  based on the well-known concept of
 \emph{transactions}.
@@
 providing all-or-nothing semantics.
-We show that a powerful transactional interface can be implemented \emph{in
-  band}, using the standard OpenFlow commands on the data-plane configuration.
+We show that a powerful transactional interface can be implemented \emph{inband}, using 
+standard OpenFlow commands on the data-plane configuration.
 %Indeed, recent OpenFlow versions already come with useful but primitive
@@
 We discuss how to use our synchronization framework to implement
-control-plane replicated service, consistent policy composition, and load-balancing.
+a replicated control-plane service, consistent policy composition, and load-balancing.
 \end{abstract}
+
+\category{C.2.2}{Computer-Communication Networks}{Network Protocols}
+\category{C.2.3}{Network Operations}{Network Management}
+%\category{G.1.6}{Optimization}{Linear Programming}
+
+%\terms{Algorithms}
+
+\keywords{Software-Defined Networking, Distributed Control Plane, In-band Mechanisms}
 
@@
 \emph{Synchronizing}
-the distributed controllers and manipulating the network state \emph{consistently}~\cite{cpc}
-are non-trivial tasks~\cite{cap-theorem}.
+the distributed controllers and manipulating the network state \emph{consistently}
+are non-trivial tasks.~\cite{cpc,sharon,cap-theorem}
 
@@
 Networking~\cite{stn}, we propose synchronization constructs that
-allow a controller to represent multiple configuration commands an
+allow a controller to represent multiple configuration commands on
 the data plane as an atomic transaction.
 %
-If none of the transaction's commands \emph{conflicts} (where conflicts can be
-defined in a general application-specific way) with the current
-configuration, the transaction appears to be executed atomically.
+If none of the transaction commands \emph{conflicts}  with the current
+configuration, where conflicts can be
+defined in a general, application-specific manner, the transaction appears to be executed atomically.
 Otherwise, the transaction is aborted in its entirety and does not
@@
 Our synchronization primitives allow the controllers to define complex notions
-of conflicts that can be used for coordinating configuration updates that
-operate over \emph{independent flow spaces}, but where dependencies are introduced
+of conflicts that can be used for coordinating configuration updates operating 
+over \emph{independent flow spaces}, but where dependencies are introduced
 in the mapping of the flows to the physical infrastructure. This is
@@
 which in turn can be used to solve consensus and
-implement a generic replicated service~\cite{Her91,Lam98} in a consistent and
+implement a generic replicated control-plane service~\cite{Her91,Lam98} in a consistent and
 fault-tolerant way, which became a mainstream building block in
 modern system design~\cite{bigtable,spanner,zookeeper}.
-We then implement a simple CAS-based concurrent policy-update mechanism. 
+We then discuss a simple CAS-based concurrent policy-update mechanism.
 Finally, we use our synchronization abstraction to provide
 the missing link for the read-modify-write object
-postulated in~\cite{cpc}). \petr{We do not seem to discuss the latter?}
+postulated in~\cite{cpc}. \petr{We do not seem to discuss the latter?}
 
@@
 In Section~\ref{sec:background}, we review the basics of the  OpenFlow
-protocol and discuss its features which are relevant for our work.
+protocol and discuss the features which are relevant for our work.
 In Section~\ref{sec:motivation}, we discuss the limitations of the OpenFlow
@@
 \paragraph{Control and data planes}
-Software-defined networking decouples \emph{control plane}, software
+Software-defined networking decouples the \emph{control plane}, software
 controllers that configure the network, from the \emph{data plane}, the
@@
 determine the network \emph{policy}.
-Controllers can change the policy via sending  
+Controllers can change the policy by sending
 control messages containing \emph{FlowMod commands}.
@@
 In order to avoid such conflicts, a \emph{FlowMod} command can be equipped with a the $\checko$ flag:
-if the switch maintains is a flow entry with an overlapping but not
+if the switch maintains a flow entry with an overlapping but not
 identical match part with the same priority but a different action,
@@
 A standard \emph{FlowMod} message (and also the flow entries) may specify
-more parameters, which are however irrelevant for our implementation, we suggest to
+more parameters, which are however irrelevant for our implementation, and we suggest to
 use the default values.
@@
 flag already provides a very powerful synchronization mechanism.
-Indeed, its ``min-transaction'' nature allows a controller 
-to install multiple flow entries in an atomic way, making
+Indeed, its ``mini-transaction'' nature allows a controller
+to install multiple flow entries in an atomic way. 
 %[[PK we did not say what reads and writes are...
@@
 First, %if employed without any additional features,
+except for some limited constraint types (such as
+overlapping flow spaces or insufficient free space),
 a bundle does not provide any means to modify the switch configuration
-based on a generic application-specific condition, depending on the
-current configuration, except to limited types of constraints, such as
-overlapping flow spaces or insufficient free space.
-But we might need more generic conditions under which configuration
+based on application-specific conditions, which depend on the
+current configuration, 
+However, as we will argue, a support for more generic conditions under which configuration
 updates
-should or should not take place.
+should or should not take place, is often desirable.
 
@@
 Consider the example in the left part of Figure~\ref{fig:CAS-example}. Two controllers,
-$1$ and $2$ read the configuration, try add two more flows to a switch
+$1$ and $2$ read the configuration, try to add two more flows to a switch
 with two outgoing links, balancing the load.
@@
 
+\begin{algorithm}[h]
+    \caption{$\textit{Naive Load Balancing}$}
+    \label{alg:naive-lb}
+\begin{algorithmic}[1]
+
+    \Require \emph{new\_flows} as flow entries of the new flows, \emph{ports} as a list of switch output ports.
+    \State $policy \gets \textbf{read\_config}$
+    \State $update\_cmds \gets Calc\_LB\_update(policy)$ %(new\_flows, ports, policy)$
+    \State $\textbf{execute-atomic}\{update\_cmds\}$
+    \Statex
+
+    \Statex \emph{\bf Function} $\textit{Calc\_LB\_update(policy)}$ % (new\_flows, ports, policy)}$
+
+	\State $flows\_per\_port \gets [0, 0, \ldots, 0]$
+	\For {$entry \in policy$}
+	    \State $port \gets entry.output\_port$
+	    \State $flows\_per\_port[port]++$
+	\EndFor
+	\State $new\_flow\_mod\_cmds \gets \{\}$
+	\For {$entry \in new\_flows$}
+	    \State $port \gets Argmin_{p\in ports} flows\_per\_port[p]$
+	    \State $entry.output\_port \gets port$
+	    \State $flows\_per\_port[port]++$
+	    \State $cmd \gets FlowMod(add,f)$
+	    \State $new\_flow\_mod\_cmds.add(cmd)$
+	\EndFor
+	\Return $new\_flow\_mod\_cmds$
+    \State \emph{\bf EndFunction}
+\end{algorithmic}
+\end{algorithm}
+
 %\petr{postpone the explanation of the right part of the figure until we come to CAS?}
@@
 Consider the problem of modifying previously installed forwarding rules.
-While a bundle allows to conditionally add new rules or remove
-old rules (and, using the flow space overlap
+In principle, a bundle allows to conditionally add new rules or remove
+old rules, and, using the flow space overlap
 check feature $\checko$, to abort the transaction if potentially conflicting
-rules were already there), it is not straightforward how to use it to  
-modify existing rules, especially provided a specific condition.
+rules were already there. However, it is not obvious how  
+to use the bundle to
+modify \emph{existing} rules, especially if the modification depends on 
+specific conditions.
 In other words, the bundle and overlap check only help
@@
 
-\begin{figure*}[t]
+\begin{figure}[t]
 \centering
-\includegraphics[width=1\columnwidth]{no-sync.png}~\includegraphics[width=1\columnwidth]{sync-with-cas.png}\\
-\caption{\emph{Left:} No synchronization. \emph{Right:}
-  CAS.}\label{fig:CAS-example}
-\end{figure*}
+\includegraphics[width=.56\columnwidth]{loadbal-1.pdf}~\hspace{-.7cm}~\includegraphics[width=.56\columnwidth]{loadbal-2.pdf}\\
+\caption{\emph{Left:} Without synchronization, the two controllers
+may end up unbalancing the flow allocations. \emph{Right:}
+ With synchronization, i.e., by bundling the flow installation with a Compare-and-Set (CAS) primitive
+ depending on the policy id (\emph{pid}), this problem is avoided.}\label{fig:CAS-example}
+\end{figure}
 
@@
 designed in a way that it does not affect the processing of data-plane
-traffic. One way to do it is to set a low priority to
+traffic. One way to achieve this is to set a low priority to
 the meta flow entries and ensure that the regular flow entries
@@
 switch configuration in an \emph{atomic} (transactional) or \emph{non-atomic} mode.
-In the atomic mode, all operations the sequence (which we also refer
-to as a \emph{transaction}) 
-should either be executed in its entirety or completely rejected. 
+In the atomic mode, either \emph{all} operations in the sequence, henceforth simply
+called \emph{transaction},
+should be executed, or none (the transaction is rejected).
 In the non-atomic mode, no atomicity guarantees are provided.
@@
 Intuitively, a controller should
-equip each policy update it wants to install with a distinct
-identifier
+equip each to-be-installed policy with a distinct
+identifier,
 %[[PK do not want to say ``unique'': we also need to motivate claimer-id]]
@@
 To make sure that policy identifiers do not grow without bound~\cite{stn}, we
-also provide a \emph{id-claimer} object.
+also provide an \emph{id-claimer} object.
 The object implements a weak type
@@
 First we translate  $op_1,\ldots,op_k$ into a sequence of FlowMod
-commands. If $op_i$ is already a FlowMid command, the translation is 
+commands. If $op_i$ is already a FlowMod command, the translation is
 trivial. If $op_i$ is a synchronization primitive in
@@
 
-To run  $\exec\{op_1,\ldots,op_k\}$, we simply send the commands in
+To implement  $\exec\{op_1,\ldots,op_k\}$, we simply send the commands in
 the resulting sequence, one by one, to the switch.
 %
-To run  $\execatomic\{op_1,\ldots,op_k\}$, we 
+To implement  $\execatomic\{op_1,\ldots,op_k\}$, we
 additionally create a bundle, wrap each of the resulting commands in a
@@
 with a bundle commit message.
-The we wait until the
+Then we wait until the
 corresponding responses are received.
-If an error message is received
+If an error message is received,
 it is returned to the control application,
@@
 values.
-In defining the value and the mask of the math part, we 
-use notation $a\concat b$ to represent a concatenation of two
+In defining the value and the mask, 
+% of the math part, 
+we
+use the notation $a\concat b$ to represent a concatenation of two
 $32$bit integers $a$ and $b$: $a\concat b := (a<<32)+b$.
@@
 
-In order to check if an identifier is claimed, a controller tries to add the
+In order to check if an identifier is claimed, a controller tries to add 
 a new entry setting the flag \texttt{\checko}, and using a different
@@
 
-Note that for switch space reasons, when checking for claims, also the deletion of the added tester flow entry is included,
+Note that for space reasons, when checking for claims, also the deletion of the added tester flow entry is included,
 which however has no effect on claims and checks of other controllers; hence, the deletion command does not
@@
   the written value. Values written to the same address will replace one another,
-following the command definition to replace old flow entry with a new one in case they
+following the command definition to replace the old flow entry with a new one in case they
 share the same match parts.
@@
 with an identical one.
+
+%HERE: we can read state by reading flow monitor => read configuraiton abstraction => state of claimer ID objects list of all claimed identifiers
 
@@
 
+\textbf{Stefan: Give intro
+}%very powerful primitive => any sequential program in repliated way can be imlemented, see also STN paper where we assume we can do this.
+
 \paragraph{Compare-and-Set}
 The pathological scenario described in the example in
-Figure~\ref{fig:CAS-example} can be easily resolved if we put
-policy update in our transactional $\execatomic$ construct together
+Figure~\ref{fig:CAS-example} can be resolved easily if we put
+the policy update in our transactional $\execatomic$ construct together
  with a \emph{compare-and-set} (CAS) operation on the policy identifier.
@@
 \emph{succeeds}), otherwise it returns $\true$ (the CAS
-\emph{fails} (which causes an abort of
+\emph{fails} which causes an abort of
 the invoking transaction).
@@
 $\textit{\compare}(\addr, old)$ followed by $\textit{\memwrite}(\addr,
-new)$ withing the same bundle.
+new)$ within the same bundle.
 
@@
 
+\begin{comment}
 \section{Related Work}\label{sec:relwork}
@@
 There exists a wide consensus that SDN control planes need to be
-distributed~\cite{onos,onix,elasticon}
+distributed~\cite{onos,onix,elasticon}.
 Onix~\cite{onix} is one of the earliest proposals, and introduced the
@@
 latency have been studied intensively
-in the literature~\cite{kandoo,ctrl-place,hotsdn13loc}
+in the literature~\cite{sharon,kandoo,ctrl-place,hotsdn13loc}
 proposes an elastic distributed controller architecture.
 
-Operating and updating SDNs consistently is already non-trivial
-from the perspective of a single controller. Reitblatt et al.~\cite{network-update}
+Operating and updating SDNs efficiently and consistently is already non-trivial
+in a setting consisting of a single controller and a single switch.~\cite{sharon} 
+Reitblatt et al.~\cite{network-update}
 introduced the notion of
-per-packet consistency and proposed a 2-phase update protocol based on tagging.
+per-packet consistency and proposed a tagging-based, 2-phase update protocol 
+which allows a single controller to update network routes.
 %, and described the \emph{two-phase update} technique, also used in our algorithms.
-Mahajan and Wattenhofer~\cite{roger-hotnets} considered weaker transient
+Mahajan and Wattenhofer~\cite{roger-hotnets} and Ludwig et al.~\cite{podc15}
+considered weaker transient
 consistency guarantees, and proposed more efficient network update algorithms
@@
 while the availability of such primitives is postulated, no implementation is
-described. In this paper, we provide this missing link, and also show that other useful and
-more powerful synchronization primitives can be implemented.
-Also, some STN implementations in~\cite{stn} assume that the control
-plane maintains a certain amount of synchrony that allows controllers
+described. Also NetPaxos~\cite{netpaxos} observes the usefulness of
+providing synchronization primitives inband, and proposes simple OpenFlow
+extensions to implement those.
+
+In this paper, we provide this missing link, and also show that other useful and
+more powerful synchronization primitives can be implemented in today's OpenFlow.
+Also, the STN protocol~\cite{stn} assumes that the control
+plane maintains a certain amount of synchrony which allows controllers
 to solve consensus or implement a replicated state machine.
@@
 
+\end{comment}
+
 \section{Concluding Remarks}\label{sec:conclusion}
@@
 }
+
+\begin{comment}
 
@@
 
+\end{appendix}
+
 \end{comment}
 
-\end{appendix}
-
 \end{document}

### 2015-05-29 inband-sync.tex
small updates

@@
 providing all-or-nothing semantics.
-We show that a powerful transactional interface can be implemented \emph{inband}, using 
+We argue that a powerful transactional interface should and can be implemented \emph{inband}, using 
 standard OpenFlow commands on the data-plane configuration.
@@
 We discuss how to use our synchronization framework to implement
-a replicated control-plane service, consistent policy composition, and load-balancing.
+ consistent policy compositions and replicated control-plane services.
 \end{abstract}
@@
 
-\keywords{Software-Defined Networking, Distributed Control Plane, In-band Mechanisms}
+\keywords{Software-Defined Networking, Distributed Control Planes, In-band Mechanisms}
 
@@
 \section{Introduction}\label{sec:intro}
+
 By consolidating and outsourcing the control over the dataplane switches to a logically
@@
 %[[PK fault-tolerance is necessary to provide hight availability?]]
-controllers should be redundant~\cite{onix,stn,onos}: a failure
+controllers should be redundant~\cite{onix,stn,onos,elasticon}: a failure 
 of one controller can be masked by other controllers.
@@
 data-plane
-events close to their origin~\cite{devoflow,kandoo,jukka,disco}.
+events close to their origin~\cite{sharon,devoflow,kandoo,jukka,disco}.
 Third, larger SDNs are likely to be operated by multiple administrators~\cite{stn} or may even offer
@@
 that packets of different header spaces should follow across the
-network~\cite{network-update,roger-hotnets,correct,stn}.
+network~\cite{network-update,podc15,roger-hotnets,correct,stn}.
 Installing \emph{conflicting} forwarding rules, e.g., rules of the same priority defined over non-disjoint
 flow spaces may lead to pathological network behavior (loops,
-blackholes, etc.)~\cite{cpc}.
+blackholes, routes bypassing a firewall, etc.).~\cite{hotnets14update,roger-hotnets}
 Similarly, installing diverging load-balancing policies may,
@@
 %
-Extending the recent work on Software Transactional
-Networking~\cite{stn}, we propose synchronization constructs that
+In particular, we propose synchronization constructs which
 allow a controller to represent multiple configuration commands on
@@
 
-%By providing more complex and atomic transactions over the
-%switch configuration
-%space, several synchronization problems can be resolved.
-%This paper investigates the implementation of a transactional interface
-%(\`{a} la STN~\cite{stn}) modified concurrently by multiple controllers.
-%\petr{"interface modified concurrently" does not sound right -> "accessed concurrently"?}
-
-We show that powerful synchronization constructs can be implemented
-directly \emph{in OpenFlow}, i.e., using the \emph{de facto} standard SDN
-protocol. This allows us to efficiently alleviate the problems related to
+We promote to implement our synchronization constructs \emph{inband},
+on the OpenFlow switch.
+An inband implementation allows us to efficiently alleviate the problems related to
 coordinating controllers via a control-plane
@@
 
-In our implementation, a part of the data-plane configuration space is
+Interestingly, our mechanisms can be implemented using the \emph{standard OpenFlow} 
+protocol, and do not require any extensions as assumed and postulated in recent literature (e.g.,
+STN~\cite{stn} or NetPaxos~\cite{netpaxos}). 
+Our work may hence also contribute to the ongoing discussion of what can be implemented
+in-band in today's OpenFlow protocol~\cite{compute,reclaim},
+as well as to possible high-level concurrency objects and 
+language abstractions.~\cite{pyretic}
+
+%By providing more complex and atomic transactions over the
+%switch configuration
+%space, several synchronization problems can be resolved.
+%This paper investigates the implementation of a transactional interface
+%(\`{a} la STN~\cite{stn}) modified concurrently by multiple controllers.
+%\petr{"interface modified concurrently" does not sound right -> "accessed concurrently"?}
+
+Concretely, in our implementation, a part of the data plane configuration space is
 used as a shared memory storing information about  contention and
 conflicts between distributed controllers.
-A transaction can then contain standard OpenFlow control operation
+A transaction can then contain standard OpenFlow control operations
 as well as  \emph{synchronization primitives} operating on this shared
@@
 postulated in~\cite{cpc}. \petr{We do not seem to discuss the latter?}
-
-We hope that our work can also contribute to the ongoing discussion of what can be implemented
-in-band in OpenFlow~\cite{compute,reclaim}, by taking a first look
-at the problem of synchronization.
 
@@
 %Section~\ref{sec:compo} describes how to perform in-band composition.
-We review related literature in Section~\ref{sec:relwork}, and conclude
-in Section~\ref{sec:conclusion}. Some technical details are postponed
-to the Appendix.
+We 
+%review related literature in Section~\ref{sec:relwork}, and 
+conclude our paper
+in Section~\ref{sec:conclusion}.
+% Some technical details are postponed
+%to the Appendix.
 
@@
 the bundle.
-\begin{algorithm}[h]
+\begin{algorithm}[t]
     \caption{$\textit{try-commit}()$}
@@
 
-\begin{algorithm}[h]
+\begin{algorithm}[t]
     \caption{$\textit{Naive Load Balancing}$}
@@
 
-\begin{algorithm}[H]
+\begin{algorithm}[t]
     \caption{$\textit{claim}(x)$}
@@
 
-\begin{algorithm}[H]
+\begin{algorithm}[t]
     \caption{$\textit{\claimcheck}(x)$}
@@
 
-\begin{algorithm}[H]
+\begin{algorithm}[t]
     \caption{$\textit{unclaim}(x)$}
@@
 
-\begin{algorithm}[h]
+\begin{algorithm}[t]
     \caption{$\textit{write}(\addr,k)$}
@@
 
-\begin{algorithm}[h]
+\begin{algorithm}[t]
     \caption{$\textit{\compare}(\addr,k)$}
@@
 
-\begin{algorithm}[h]
+\begin{algorithm}[t]
     \caption{$\textit{\cas}(\addr, old,new)$}
@@
 
-\begin{algorithm}[h]
+\begin{algorithm}[t]
     \caption{Policy update with only CAS}
@@
 
-\begin{algorithm}[h]
+\begin{algorithm}[t]
     \caption{Advanced policy update}

### 2015-05-29 inband-sync-many-cites.tex
started shortening

--- 2015-05-29 inband-sync.tex
@@
 We argue that a powerful transactional interface should and can be implemented \emph{inband}, using 
-standard OpenFlow commands on the data-plane configuration.
+standard OpenFlow commands on the data plane configuration.
 %Indeed, recent OpenFlow versions already come with useful but primitive
@@
 We discuss how to use our synchronization framework to implement
- consistent policy compositions and replicated control-plane services.
+ consistent policy compositions and replicated control plane services.
 \end{abstract}
 
+\begin{comment}
 \category{C.2.2}{Computer-Communication Networks}{Network Protocols}
@@
 %\end{center}
+
+\end{comment}
 
@@
 communication-intensive
-%control-plane
-data-plane
+%control plane
+data plane
 events close to their origin~\cite{sharon,devoflow,kandoo,jukka,disco}.
@@
 participatory interfaces where different users can install and trigger policy changes
-concurrently~\cite{participatory}.
+concurrently~\cite{participatory,opensdwn}.
 
@@
 
-\paragraph{Contributions}
+\noindent\textbf{Contributions}
 This paper applies the principle of atomicity in concurrent computing
@@
 An inband implementation allows us to efficiently alleviate the problems related to
-coordinating controllers via a control-plane
+coordinating controllers via a control plane
 out-of-band network, in the presence of asynchrony and failures.
@@
 
-Interestingly, our mechanisms can be implemented using the \emph{standard OpenFlow} 
-protocol, and do not require any extensions as assumed and postulated in recent literature (e.g.,
-STN~\cite{stn} or NetPaxos~\cite{netpaxos}). 
-Our work may hence also contribute to the ongoing discussion of what can be implemented
-in-band in today's OpenFlow protocol~\cite{compute,reclaim},
-as well as to possible high-level concurrency objects and 
-language abstractions.~\cite{pyretic}
-
+Interestingly, our mechanisms are simple and can be implemented using the \emph{standard OpenFlow}
+protocol. 
+%
 %By providing more complex and atomic transactions over the
@@
 %\petr{"interface modified concurrently" does not sound right -> "accessed concurrently"?}
-
-Concretely, in our implementation, a part of the data plane configuration space is
-used as a shared memory storing information about  contention and
+%
+Concretely, in our implementation, 
+as part of the data plane configuration space, 
+we propose to use a shared memory storing information about  contention and
 conflicts between distributed controllers.
@@
 which in turn can be used to solve consensus and
-implement a generic replicated control-plane service~\cite{Her91,Lam98} in a consistent and
-fault-tolerant way, which became a mainstream building block in
+implement a generic replicated control plane service~\cite{Her91,Lam98} in a consistent and
+fault-tolerant way: a mainstream building block in
 modern system design~\cite{bigtable,spanner,zookeeper}.
-We then discuss a simple CAS-based concurrent policy-update mechanism.
-Finally, we use our synchronization abstraction to provide
+We also discuss a simple CAS-based concurrent policy-update mechanism,
+and also use our synchronization abstraction to provide
 the missing link for the read-modify-write object
@@
 
-\paragraph{Organization}
+\noindent\textbf{Organization}
 The rest of the paper is organized as follows.
@@
 protocol and discuss the features which are relevant for our work.
-In Section~\ref{sec:motivation}, we discuss the limitations of the OpenFlow
+In Section~\ref{sec:motivation}, we highlight the limitations of the OpenFlow
 primitives and motivate the need for transactions.
@@
 
-\paragraph{Control and data planes}
+\noindent\textbf{Control and data planes}
 Software-defined networking decouples the \emph{control plane}, software
@@
 In the simplest setting, a switch maintains one flow table (table 0).
-A data-plane packet arriving at a switch is first checked against
+A data plane packet arriving at a switch is first checked against
 the rule with the highest priority and in table $0$.
@@
 forming a \emph{pipeline}.
-Data-plane packets are checked against flow entries in the pipelined
+Data plane packets are checked against flow entries in the pipelined
 tables to determine the set of actions the switch should perform on the packet.
-A data-plane packet arriving at a switch is first checked against
+A data plane packet arriving at a switch is first checked against
 the rule with the highest priority and in table $0$.
@@
 
-\paragraph{Interaction between the planes: FlowMod commands}
+\noindent\textbf{Interaction between the planes: FlowMod commands}
 The flow tables installed on the switches across the data plane
@@
 a modification to an existing flow entry.
+%
+%
 %[[PK already said
@@
 %]]
-
+%
+%
 A FlowMod command can either add a flow entry or delete a flow entry.
@@
 A $\dele$ command simply removes an existing flow entry with the same
-match and priority  or does nothing if such an entry is not present.
+match and priority,  or does nothing if such an entry is not present.
 \petr{Is this correct?}
@@
 
-\paragraph{Bundling}
+\noindent\textbf{Bundling}
 
@@
 %When properly configured,
-Bundle execution is considered \emph{controller-atomic}
+The bundle execution is considered \emph{controller-atomic}
 in the sense that all controllers can only observe switch
@@
 making bundles \emph{packet-atomic},
-in the sense that every data-plane packet is processed by a configuration
+in the sense that every data plane packet is processed by a configuration
 before or after the bundle, and not by a configuration resulting from
@@
 
-\section{Limitations and Motivations}\label{sec:motivation}
+\section{Limitations and Motivation}\label{sec:motivation}
 
@@
 At first sight, bundling \emph{FlowMod} commands with  the $\checko$
-flag already provides a very powerful synchronization mechanism.
+flag already provides a useful synchronization mechanism.
 Indeed, its ``mini-transaction'' nature allows a controller
@@
 %]]
-However, as is
+However, as is,
 %while the bundle can indeed be seen as a ``mini-transaction'',
-bundling has important limitations.
+bundling has important limitations, as we will argue in the following.
 %\petr{unclear what "reads and writes" are here, not discussed before. How openflow can be used to implement reads and writes?}
@@
 in charge
-of \emph{load-balancing}.
-In a load-balancing application, seemingly independent actions, defined over
-completely independent logical flow spaces (say, two flows with source IP address ranges $1*$ and $0*$),
-may in fact be dependent due to collocation in the underlying physical network.
-If the two controllers concurrently and independently
-update forwarding rules,
+of \emph{load-balancing}, see Figure~\ref{fig:CAS-example} 
+for a simplistic example with two links. 
+In such a scenarios, seemingly independent actions, defined over
+completely independent logical flow spaces (say, two flows with source IP address ranges $1*$ and $0*$,
+respectively),
+may actually be dependent: the flows share the underlying physical network.
+Accordingly, if multiple controllers concurrently and independently
+update forwarding rules according to a naive load-balancing algorithm \`{a} la Algorithm~\ref{alg:naive-lb},
 they may involuntarily
-unbalance the flow allocation, as the managed flows are embedded
-on the same physical links.
-Flow space overlap checks cannot be used to detect
+unbalance the flow allocation.
+Simple flow space overlap checks cannot be used to detect
 such conflicts.
-
-Consider the example in the left part of Figure~\ref{fig:CAS-example}. Two controllers,
-$1$ and $2$ read the configuration, try to add two more flows to a switch
-with two outgoing links, balancing the load.
-Imagine that each of the controllers reads the switch configuration and
-learns that currently one flow is forwarded to the left link and two
-flows are forwarded to the right link.
-Without synchronization, the two controllers naturally choose to
-install their flows on the left link, which results in an undesirable
-unbalanced state. Note that, assuming that all these flows are
+Note that, assuming that all these flows are
 completely independent, bundling and $\checko$ features do not really
@@
 
+%Consider the example in the left part of Figure~\ref{fig:CAS-example}. Two controllers,
+%$1$ and $2$ read the configuration, try to add two more flows to a switch
+%with two outgoing links, balancing the load.
+%Imagine that each of the controllers reads the switch configuration and
+%learns that currently one flow is forwarded to the left link and two
+%flows are forwarded to the right link.
+%Without synchronization, the two controllers naturally choose to
+%install their flows on the left link, which results in an undesirable
+%unbalanced state.
+
 \begin{algorithm}[t]
-    \caption{$\textit{Naive Load Balancing}$}
+    \caption{$\textit{Naive Flow Balancing}$}
     \label{alg:naive-lb}
@@
 
-    \Require \emph{new\_flows} as flow entries of the new flows, \emph{ports} as a list of switch output ports.
+    \Require \emph{new\_flows}: flow entries of the new flows, \emph{ports}: a list of switch output ports.
     \State $policy \gets \textbf{read\_config}$
     \State $update\_cmds \gets Calc\_LB\_update(policy)$ %(new\_flows, ports, policy)$    
-    \State $\textbf{execute-atomic}\{update\_cmds\}$
+    \State $\textbf{send}\{update\_cmds\}$
+%    \State $\textbf{execute-atomic}\{update\_cmds\}$
     \Statex
@@
 	\For {$entry \in new\_flows$}
-	    \State $port \gets Argmin_{p\in ports} flows\_per\_port[p]$
+	    \State $port \gets \arg\min_{p\in ports} flows\_per\_port[p]$
 	    \State $entry.output\_port \gets port$
@@
 \noindent\textbf{Example 2: Supporting modifications.}
-Consider the problem of modifying previously installed forwarding rules.
+Consider the arguably very basic problem of \emph{modifying} previously installed forwarding rules
+in a consistent manner.
 In principle, a bundle allows to conditionally add new rules or remove
@@
 
-Intuitively, one way to address these issues is to use a part of the
-configuration space of the data-plane to store
-``meta-configuration'', i.e., information on contention and conflicts between
-concurrently installed policy updates. Controllers can read and modify
+\vspace{2mm}
+\noindent\textbf{Our Approach.}
+This paper proposes mechanisms to solve these two examples,
+by using a minimal 
+``meta-configuration'' space, i.e., controllers store information 
+on contention and conflicts between
+concurrently installed policy updates on the switch. Controllers can read and modify
 this meta-data by applying synchronization primitives.
@@
 
+We will later come back to these examples, and show that Algorithm~\ref{alg:naive-lb} 
+can actually be used without any modifications within our framework. 
+Consistent load-balancing can also be achieved using our atomic Compare-and-Set instruction based on policy ids (pids),
+see Figure~\ref{fig:CAS-example} (\emph{right}).
+
 \begin{figure}[t]
@@
 \includegraphics[width=.56\columnwidth]{loadbal-1.pdf}~\hspace{-.7cm}~\includegraphics[width=.56\columnwidth]{loadbal-2.pdf}\\
-\caption{\emph{Left:} Without synchronization, the two controllers
-may end up unbalancing the flow allocations. \emph{Right:}
+\caption{\emph{Left:} Without synchronization, the two controllers naturally choose to
+install their flows on the left link, which results in an undesirable
+unbalanced state.
+ \emph{Right:}
  With synchronization, i.e., by bundling the flow installation with a Compare-and-Set (CAS) primitive
@@
 
-\section{Synchronization abstractions}\label{sec:main}
+\section{Synchronization Abstractions}\label{sec:main}
 
@@
 
-\paragraph{Configuration space}
+\noindent\textbf{Configuration space}
 %
@@
 the ``normal'' configuration which determines the network policy and
-is used to process the data-plane traffic (e.g., forwarding rules),
+is used to process the data plane traffic (e.g., forwarding rules),
 and the \emph{meta-configuration} that contains information used by the
@@
 The meta-configuration is implemented as a set of flow entries
-designed in a way that it does not affect the processing of data-plane
+designed in a way that it does not affect the processing of data plane
 traffic. One way to achieve this is to set a low priority to
@@
 
-\paragraph{Synchronization}
+\noindent\textbf{Synchronization}
  %
@@
 $\exec\{op_1,\;\ldots\;op_k\}$ and $\execatomic\{op_1,\;\ldots\;op_k\}$
-which refer to executing a sequence of commands $op_1,\ldots,op_k$ in
+which execute a sequence of commands $op_1,\ldots,op_k$ in
 a regular and atomic way, respectively.
@@
 
-\paragraph{Synchronization primitives}
+\noindent\textbf{Synchronization primitives}
 %
@@
 
-\paragraph{Implementation}\label{sec:t-impl}
+\noindent\textbf{Implementation}\label{sec:t-impl}
 %
@@
 commands. If $op_i$ is already a FlowMod command, the translation is
-trivial. If $op_i$ is a synchronization primitive in
-$\{\textit{claim},  \textit{checkclaim}, \textit{unclaim},
-\textit{write}, \textit{compare}\}$, we use
-Algorithms~\ref{alg:claim}, \ref{alg:check}, \ref{alg:unclaim},
-\ref{alg:write}, and \ref{alg:compare}, respectively, to get the
-corresponding command(s).
+trivial. If $op_i$ is a synchronization primitive, we employ 
+Algorithm~\ref{alg:claim} for $\{\textit{claim}$, 
+Algorithm~\ref{alg:checkclaim} for $\textit{checkclaim}$, 
+Algorithm~\ref{alg:unclaim}
+for $\textit{unclaim}$,
+Algorithm~\ref{alg:write} for $\textit{write}$ and
+Algorithm~\ref{alg:compare} for $\textit{compare}\}$.
 
@@
 additionally create a bundle, wrap each of the resulting commands in a
-bundle message with the corresponding bundle identifier, complete it
+bundle message with the corresponding bundle identifier, and complete it
 with a bundle commit message.
@@
 corresponding responses are received.
-If an error message is received,
-it is returned to the control application,
-otherwise $\ack$ is returned.
+If an error message is return,
+it is forwarded to the control application,
+otherwise it is $\ack$ed.
 
@@
 The \textit{match} part in these commands specifies only the $64$bit meta-data
-field in the header space, leaving the remaining fields in default
+field in the header space, leaving the remaining fields at default
 values.
@@
 
-\paragraph{The id-claimer}
+\noindent\textbf{The id-claimer}
 
@@
 If the check succeeds, we delete the entry.
-Con sidering, $m_1$ and $m_2$ from the example above,
+Considering $m_1$ and $m_2$ from the example above,
 with respective actions $a_1=1$ and $a_2=2$,  they overlap
-with different actions iff $x_1=x_2$. Therefore, if $m_1$ is used in \claimcheck after $m_2$ used in the claim operation,
+with different actions iff $x_1=x_2$. Therefore, if $m_1$ is used in the \claimcheck after $m_2$ was used in the claim operation,
 a conflict is detected iff $x_1=x_2$, as desired.
 
-Note that for space reasons, when checking for claims, also the deletion of the added tester flow entry is included,
+Note that in order to save space, when checking for claims, also the deletion of the added tester flow entry is included,
 which however has no effect on claims and checks of other controllers; hence, the deletion command does not
@@
 
-\paragraph{Write and compare}
+\noindent\textbf{Write and compare}
 The memory \memwrite and \compare
@@
 
-In order to check if a value is written in a given address,
+In order to check if a value is written at a given address,
 we try to add it the same way as in the \memwrite operation,
@@
 
-We now discuss several applications of our transactional interface.
-
-\textbf{Stefan: Give intro
-}%very powerful primitive => any sequential program in repliated way can be imlemented, see also STN paper where we assume we can do this.
-
-\paragraph{Compare-and-Set}
-The pathological scenario described in the example in
+We believe that our synchronization mechanisms 
+can be useful in many situations. 
+For example, the pathological scenario described in 
 Figure~\ref{fig:CAS-example} can be resolved easily if we put
 the policy update in our transactional $\execatomic$ construct together
- with a \emph{compare-and-set} (CAS) operation on the policy identifier.
+ with a \emph{Compare-and-Set} (CAS) operation on the policy identifier.
 A $\cas(\addr,\textit{old},\textit{new})$ operation on a memory location $\addr$
@@
 replaces it with $\textit{new}$ and returns $\true$ (the CAS
-\emph{succeeds}), otherwise it returns $\true$ (the CAS
+\emph{succeeds}), otherwise it returns $\false$ (the CAS
 \emph{fails} which causes an abort of
@@
 
-CAS can be used to solve the fundamental
+More generally, CAS can be used to solve the fundamental
 problem of fault-tolerant \emph{consensus}~\cite{FLP85}
@@
 Indeed, each controller simply performs, within a transaction, a CAS operation to replace the default value
-in a memory location at the meta-configuration of a data-plane switch.
+in a memory location at the meta-configuration of a data plane switch.
 If the transaction commits, the controller considers itself
@@
 }
-and, more generally,
+and allows us
 to implement any generic replicated state machine~\cite{Her91,Lam98}
@@
 asynchrony and failure of any number of controllers.
+As such, CAS can also be used to implement STN~\cite{stn},
+by providing the missing synchronization object.
+
 %which will be instrumental in the
@@
 
-\paragraph{Policy Updates}
-
-As a second example, let us consider a simple policy update protocol that
+Our mechanism can also be used as a generic template to
+implement concurrent policy compositions of a user-specific
+$\textsc{\ufunc}()$ function~\cite{cpc}.
+That is, let us consider a simple policy update protocol that
 allows a set of  controllers to concurrently update switch policies, i.e., sets of
@@
 policy.
-
-Let us assume that the composition predicate is encoded in a user-specific
-$\textsc{\ufunc}()$ function~\cite{cpc}.
 Algorithm~\ref{alg:simple-update} uses the bundle feature to update
@@
 
-\liron{we can explain how a normal controller code can be instrumented to use our paradigm}
-\petr{did not get the comment}
+%\liron{we can explain how a normal controller code can be instrumented to use our paradigm}
+%\petr{did not get the comment}
 
@@
 
-\petr{what does it mean for an update function to be ``unsafe''?}
+%\petr{what does it mean for an update function to be ``unsafe''?}
 
@@
 
-Note that in Algorithm~\ref{alg:simple-update}, policies identifiers
-grow without bound.
-However, as was observed in~\cite{cpc}, the number of currently
-used identifiers can be made proportional to the number of concurrent
-controllers.
+%Note that in Algorithm~\ref{alg:simple-update}, policies identifiers
+%grow without bound.
+%However, as was observed in~\cite{cpc}, the number of currently
+%used identifiers can be made proportional to the number of concurrent
+%controllers.
 %
@@
 \petr{The description is not very clear.}
+
+In other words, Algorithm~\ref{alg:simple-update} can also be used
+to render our naive load-balancing algorithm (Algorithm~\ref{alg:naive-lb}) from Section~\ref{sec:motivation}
+consistent, without any modifications.
 
@@
 
-%\paragraph{One-Shot Update} FIXME: desribe how to implement one-shot update
-%
-%\paragraph{STN}
+%\noindent\textbf{One-Shot Update} FIXME: desribe how to implement one-shot update
+%
+%\noindent\textbf{STN}
 %
@@
 on the high-level interface exported to the control plane, but also on the
-paths taken by the data-plane packets.
+paths taken by the data plane packets.
 Also, we assumed that controllers are subject to failures, which is usually not
@@
 
-\section{Concluding Remarks}\label{sec:conclusion}
-
+\section{Conclusion}\label{sec:conclusion}
+
+This paper initiated the study of mechanisms which allow distributed controllers 
+to synchronize their possibly concurrent actions. We believe that our in-band 
+approach is natural and attractive as it can avoid the 
+complexities and limitations of distributed out-of-band agreement protocols.
+Moreover, our mechanisms are simple and can be implemented using the \emph{standard OpenFlow}
+protocol: they do not require any protocol/hardware extensions as postulated in recent literature~\cite{stn,netpaxos}.
+Our work may hence also contribute to the ongoing discussion of what can be implemented
+in-band in today's OpenFlow protocol~\cite{reclaim},
+as well as to useful high-level concurrency objects and
+language abstractions.~\cite{pyretic}
+
+%,
+%nor on
+%nor do we make any assumptions on packet orderings.
+%and do not rely on any
+%
+%.
+
+\begin{comment}
 While much progress has been made over the
@@
 Exploring the power of such features is left to future work.
+\end{comment}
 
 {
+\begin{footnotesize}
 \bibliographystyle{abbrv}
 \bibliography{references}  % main.bib is the name of the Bibliography in this case
+\end{footnotesize}
 }

### 2015-05-29 inband-sync.tex
first 6-page version

@@
 %[[PK fault-tolerance is necessary to provide hight availability?]]
-controllers should be redundant~\cite{onix,stn,onos,elasticon}: a failure 
+controllers should be redundant~\cite{onix,onos,elasticon}: a failure 
 of one controller can be masked by other controllers.
@@
 communication-intensive
-%control-plane
-data-plane
-events close to their origin~\cite{sharon,devoflow,kandoo,jukka,disco}.
+%control plane
+data plane
+events close to their origin~\cite{devoflow,kandoo}.
 Third, larger SDNs are likely to be operated by multiple administrators~\cite{stn} or may even offer
@@
 the distributed controllers and manipulating the network state \emph{consistently}
-are non-trivial tasks.~\cite{cpc,sharon,cap-theorem}
+are non-trivial tasks.~\cite{sharon}
 
@@
 that packets of different header spaces should follow across the
-network~\cite{network-update,podc15,roger-hotnets,correct,stn}.
+network~\cite{network-update,roger-hotnets,podc15}.
 Installing \emph{conflicting} forwarding rules, e.g., rules of the same priority defined over non-disjoint
@@
 Similarly, installing diverging load-balancing policies may,
-when combined, \emph{increase} the load~\cite{log-cent}.
+when combined, \emph{increase} the load.
 To render things more difficult, controllers may also fail,
@@
 The concurrent computing literature offers
-a wide range of synchronization abstractions~\cite{HS08-book}.
+a wide range of synchronization abstractions.
 E.g., the popular \emph{transactional
-  memory} abstraction~\cite{stm-st95,tm-book} provides a collection of
+  memory} abstraction~\cite{tm-book} provides a collection of
 concurrent processes with the ability of aggregating sequences of
@@
 terms of the necessary computability assumptions about the underlying
-system~\cite{FLP85,CHT96,cap-theorem}, and the amount of communication
-complexity~\cite{Lam98,Lam06}.
+system~\cite{FLP85}, and the amount of communication
+complexity~\cite{Lam06}.
 In contrast, our \emph{in-band} solution allows the
@@
 which in turn can be used to solve consensus and
-implement a generic replicated control-plane service~\cite{Her91,Lam98} in a consistent and
-fault-tolerant way, which became a mainstream building block in
-modern system design~\cite{bigtable,spanner,zookeeper}.
-We then discuss a simple CAS-based concurrent policy-update mechanism.
-Finally, we use our synchronization abstraction to provide
+implement a generic replicated control plane service~\cite{Her91} in a consistent and
+fault-tolerant way: a mainstream building block in
+modern system design.
+We also discuss a simple CAS-based concurrent policy-update mechanism,
+and also use our synchronization abstraction to provide
 the missing link for the read-modify-write object
-postulated in~\cite{cpc}. \petr{We do not seem to discuss the latter?}
+postulated in~\cite{stn}. \petr{We do not seem to discuss the latter?}
 
@@
 %
-In OpenFlow~\cite{of-spec}, the \emph{de facto} standard SDN protocol
+In OpenFlow, the \emph{de facto} standard SDN protocol
 today, the data plane consists of \emph{OpenFlow switches}.
@@
 the meta-configuration,  making sure that the
-\emph{pipelining} mechanism~\cite{of-spec} will never involve this
+\emph{pipelining} mechanism will never involve this
 flow table.
@@
 }
-and, more generally,
-to implement any generic replicated state machine~\cite{Her91,Lam98}
+and allows us
+to implement any generic replicated state machine~\cite{Her91}
 in the control plane,  in the \emph{wait-free} manner, i.e., tolerating
 asynchrony and failure of any number of controllers.
+As such, CAS can also be used to implement STN~\cite{stn},
+by providing the missing synchronization object.
+
 %which will be instrumental in the
@@
 
-\paragraph{Policy Updates}
-
-As a second example, let us consider a simple policy update protocol that
+Our mechanism can also be used as a generic template to
+implement concurrent policy compositions of a user-specific
+$\textsc{\ufunc}()$ function~\cite{stn}.
+That is, let us consider a simple policy update protocol that
 allows a set of  controllers to concurrently update switch policies, i.e., sets of
@@
 
-Note that in Algorithm~\ref{alg:simple-update}, policies identifiers
-grow without bound.
-However, as was observed in~\cite{cpc}, the number of currently
-used identifiers can be made proportional to the number of concurrent
-controllers.
+%Note that in Algorithm~\ref{alg:simple-update}, policies identifiers
+%grow without bound.
+%However, as was observed in~\cite{stn}, the number of currently
+%used identifiers can be made proportional to the number of concurrent
+%controllers.
 %
@@
 %
-The transactional interface exported by the CPC abstraction is inspired by the
+The transactional interface exported by the STN abstraction is inspired by the
 work on speculative concurrency control using software transactional memory
@@
 Exploring the power of such features is left to future work.
+\end{comment}
 
 {
+%\begin{footnotesize}
 \bibliographystyle{abbrv}
 \bibliography{references}  % main.bib is the name of the Bibliography in this case
+%\end{footnotesize}
 }
@@
 primitives voluntarily assumed in the recent implementation of a
-recent proposal for a consistent composition system for forwarding policies~\cite{cpc}.
+recent proposal for a consistent composition system for forwarding policies~\cite{stn}.
 We then use these transactions directly to implement a lightweight

### 2015-05-29 inband-sync.tex
bla

@@
 Interestingly, our mechanisms are simple and can be implemented using the \emph{standard OpenFlow}
-protocol. 
+protocol (version 1.4). 
 %
@@
 
-Another important feature in OpenFlow is \emph{bundling}.
+Another important feature in OpenFlow is \emph{bundling} (as of OpenFlow version 1.4). \stefan{liron, is that correct?}
 A controller can send multiple $\FlowMod$ commands  equipped with

### 2015-05-30 inband-sync.tex
small

@@
 %[[PK sorry for stubborness :)
-\title{On the Synchronization Power of the OpenFlow Data Plane}
+%\title{On the Synchronization Power of the OpenFlow Data Plane}
+\title{Network Concurrency Control Done In Band}
 %\\ {\Large Towards Atomic Transactions In-Band}}
@@
 \author{
-Liron Schiff\thanks{Supported by the European Research Council (ERC) Starting Grant no. 259085 and by the Israel Science Foundation Grant no.~1386/11.} $^1$,
+Liron Schiff$^1$, % \thanks{Supported by the European Research Council (ERC) Starting Grant no. 259085 and by the Israel Science Foundation Grant no.~1386/11.} $^1$,
 Stefan Schmid$^2$, Petr Kuznetsov$^3$ \\
@@
 We discuss how to use our synchronization framework to implement
- consistent policy compositions and replicated control plane services.
+ consistent policy composition and fault-tolerant control-plane applications.
 \end{abstract}

### 2015-05-30 inband-sync.tex
intro

@@
 providing all-or-nothing semantics.
-We argue that a powerful transactional interface should and can be implemented \emph{inband}, using 
+We argue that a powerful transactional interface should and can be
+implemented \emph{in-band}, using 
 standard OpenFlow commands on the data plane configuration.
@@
 (hardware) data plane.
-Moreover, OpenFlow, the standard SDN protocol today, introduces many generalizations,
-in terms of traffic engineering, definition of flows, as well as in-band network functionalities,
-by relying on a simple match-action paradigm which allows us to define
-forwarding rules based not only on Layer-2, but also Layer-3 and Layer-4 (and possibly beyond).
-
+%Moreover, OpenFlow, the standard SDN protocol today, introduces many generalizations,
+%in terms of traffic engineering, definition of flows, as well as in-band network functionalities,
+%by relying on a simple match-action paradigm which allows us to define
+%forwarding rules based not only on Layer-2, but also Layer-3 and Layer-4 (and possibly beyond).
+%
 While the perspective of \emph{logically centralized} control plane
@@
 such distributed control planes. The problem is essentially a
-distributed systems
+distributed-systems
 one: Multiple controllers may simultaneously try to
@@
 concurrent processes with the ability of aggregating sequences of
-shared-memory operations in atomic \emph{transactions} with
+shared-memory operations in \emph{atomic  transactions} with
 all-or-nothing semantics.
@@
 
-\noindent\textbf{Contributions}
+\hspace{1mm}
+\noindent\textbf{Contributions.}
 This paper applies the principle of atomicity in concurrent computing
@@
 %
-If none of the transaction commands \emph{conflicts}  with the current
+If none of the transaction's commands \emph{conflicts}  with the current
 configuration, where conflicts can be
 defined in a general, application-specific manner, the transaction appears to be executed atomically.
-Otherwise, the transaction is aborted in its entirety and does not
+Otherwise, the transaction is \emph{aborted} in its entirety and does not
 affect neither traffic nor other controllers.
 
-We promote to implement our synchronization constructs \emph{inband},
-on the OpenFlow switch.
-An inband implementation allows us to efficiently alleviate the problems related to
+We promote to implement our synchronization constructs \emph{in-band},
+on the switch.
+An in-band implementation allows us to efficiently alleviate the problems related to
 coordinating controllers via a control plane
@@
 Concretely, in our implementation, 
-as part of the data plane configuration space, 
-we propose to use a shared memory storing information about  contention and
-conflicts between distributed controllers.
+we use a part of the data-plane configuration space as a \emph{shared
+  memory}  that stores information about  contention and conflicts
+between the controllers.
 A transaction can then contain standard OpenFlow control operations
@@
 memory.
-Our synchronization primitives allow the controllers to define complex notions
-of conflicts that can be used for coordinating configuration updates operating 
-over \emph{independent flow spaces}, but where dependencies are introduced
-in the mapping of the flows to the physical infrastructure. This is
-important, e.g., for load-balancing applications.
+Our synchronization primitives allow the controllers to define
+general notions of conflicts between configuration updates, 
+not restricted to simple conflicts on the overlapping flow spaces.
+In particular, we can define dependencies between mappings of
+independent flows to the physical infrastructure, which is important,
+e.g., for load-balancing applications.     
+%which can be used even for coordinating 
+%configuration updates operating 
+%over \emph{independent flow spaces}, but where dependencies are introduced
+%in the mapping of the flows to the physical infrastructure. This is
+%important, e.g., for load-balancing applications.
 
@@
 the missing link for the read-modify-write object
-postulated in~\cite{stn}. \petr{We do not seem to discuss the latter?}
+postulated in~\cite{stn}. %\petr{We do not seem to discuss the latter?}
 
@@
 As such, CAS can also be used to implement STN~\cite{stn},
-by providing the missing synchronization object.
+by providing the postulated \emph{read-modify-write} object.
 

### 2015-05-30 inband-sync.tex
background

@@
 %OpenFlow~\cite{of-spec}, the standard SDN protocol today.
-
-\noindent\textbf{Control and data planes}
-Software-defined networking decouples the \emph{control plane}, software
+\hspace{1mm}
+\noindent\textbf{Control and data planes.} Software-defined networking
+decouples the \emph{control plane}, software
 controllers that configure the network, from the \emph{data plane}, the
@@
 a ternary pattern over packet \emph{header} fields.
-The pattern is represented as a \emph{value} and a \emph{mask}. \liron{we need to discuss about this, I am not sure this description is accurate.}
+The pattern is represented as a \emph{value} and a \emph{mask}. 
+%\liron{we need to discuss about this, I am not sure this description is accurate.}
 In the mask, certain header fields, e.g., TCP or UDP port numbers or destination and source
@@
 not affect the match.
-Certain fields, such as \emph{ipv4\_src }, \emph{ipv4\_dst}, \emph{metadata}, can be arbitrarily
+Certain fields, such as \emph{ipv4\_src}, \emph{ipv4\_dst}, \emph{metadata}, can be arbitrarily
 bitmasked.
-In our implementations, we are going to use bitmasking of the metadata
+In our implementations, we are going to use bitmasking of the \emph{metadata}
 field for performing conditional operations on  its content.
-For example, \texttt{ipv4\_src=10.1.14.0, ff.ff.ff.f0} matches all packets
-with source IP addresses in the range \texttt{10.1.14.0}--\texttt{10.1.14.127}.
+%For example, \texttt{ipv4\_src=10.1.14.0, ff.ff.ff.f0} matches all packets
+%with source IP addresses in the range \texttt{10.1.14.0}--\texttt{10.1.14.127}.
 
 A \emph{configuration} of an OpenFlow switch is represented as a
-collection \emph{flow table}.
+a collection of \emph{flow tables}.
 A flow table is a set of \emph{flow entries}, each containing a rule,
@@
 %\liron{Another direction is to have a simplified explenation as now and to refference spesific sections of the standard in side our paper.}
-
+\hspace{1mm}
 \noindent\textbf{Interaction between the planes: FlowMod commands}
@@
 match and priority,  or does nothing if such an entry is not present.
-\petr{Is this correct?}
+%\petr{Is this correct?}
 
@@
 
-If rules with overlapping but not identical match fields have the same priority,
-inconsistencies can arise, since there is no clearly specified action
-that should be applied to a packet matching the two entries.
-In order to avoid such conflicts, a \emph{FlowMod} command can be equipped with a the $\checko$ flag:
+%If rules with overlapping but not identical match fields have the same priority,
+%inconsistencies can arise, since there is no clearly specified action
+%that should be applied to a packet matching the two entries.
+In order to avoid inconsistencies caused by different rules with
+overlapping match fields, a \emph{FlowMod} command can be equipped with a the $\checko$ flag:
 if the switch maintains a flow entry with an overlapping but not
@@
 feature.
-A standard \emph{FlowMod} message (and also the flow entries) may specify
-more parameters, which are however irrelevant for our implementation, and we suggest to
-use the default values.
+For simplicity, we omit other standard parameters, assuming them to carry
+default values.
+%A standard \emph{FlowMod} message (and also the flow entries) may specify
+%more parameters, which are however irrelevant for our implementation, and we suggest to
+%use the default values.
 
@@
 
-\noindent\textbf{Bundling}
-
-Another important feature in OpenFlow is \emph{bundling} (as of OpenFlow version 1.4). \stefan{liron, is that correct?}
+\hspace
+\noindent\textbf{Bundling.}
+%
+Another important feature in OpenFlow is \emph{bundling}. %(as of OpenFlow version 1.4). \stefan{liron, is that correct?}
 A controller can send multiple $\FlowMod$ commands  equipped with

### 2015-05-30 inband-sync.tex
typo
mostly cosmetics

@@
 
-\hspace{1mm}
+\vspace{1mm}
 \noindent\textbf{Contributions.}
@@
 
-Interestingly, our mechanisms are simple and can be implemented using the \emph{standard OpenFlow}
-protocol (version 1.4). 
+Interestingly, our mechanisms are simple and can be implemented using the standard \emph{OpenFlow}
+protocol (version 1.4~\cite{openflow}). 
 %
@@
 %OpenFlow~\cite{of-spec}, the standard SDN protocol today.
-\hspace{1mm}
+\vspace{1mm}
 \noindent\textbf{Control and data planes.} Software-defined networking
@@
 A \emph{configuration} of an OpenFlow switch is represented as a
-a collection of \emph{flow tables}.
+collection of \emph{flow tables}.
 A flow table is a set of \emph{flow entries}, each containing a rule,
@@
 %\liron{Another direction is to have a simplified explenation as now and to refference spesific sections of the standard in side our paper.}
-\hspace{1mm}
+\vspace{1mm}
 \noindent\textbf{Interaction between the planes: FlowMod commands}
@@
 
-\hspace
+\vspace{1mm}
 \noindent\textbf{Bundling.}
@@
 %\begin{enumerate}
-\vspace{2mm}
+\vspace{1mm}
 \noindent \textbf{Example 1: Handling complex dependencies.} Let us consider two controllers
@@
 Accordingly, if multiple controllers concurrently and independently
-update forwarding rules according to a naive load-balancing algorithm \`{a} la Algorithm~\ref{alg:naive-lb},
+update forwarding rules according to a na\"ive load-balancing algorithm \`{a} la Algorithm~\ref{alg:naive-lb},
 they may involuntarily
@@
 
+{\small
 \begin{algorithm}[t]
-    \caption{$\textit{Naive Flow Balancing}$}
+    \caption{$\textit{Na\"ive Flow Balancing}$}
     \label{alg:naive-lb}
@@
 
-    \Require \emph{new\_flows}: flow entries of the new flows, \emph{ports}: a list of switch output ports.
-    \State $policy \gets \textbf{read\_config}$
-    \State $update\_cmds \gets Calc\_LB\_update(policy)$ %(new\_flows, ports, policy)$    
-    \State $\textbf{send}\{update\_cmds\}$
+    \Require \emph{new\_flows}: flow entries of the new flows,
+    \emph{ports}: a list of switch output ports, 
+    \State $\textit{policy} \gets \textbf{read\_config}$
+    \State $\textit{update\_cmds} \gets \textit{lb\_update}(policy)$ %(new\_flows, ports, policy)$    
+    \State $\textbf{send}\{\textit{update\_cmds}\}$
 %    \State $\textbf{execute-atomic}\{update\_cmds\}$
@@
 
-    \Statex \emph{\bf Function} $\textit{Calc\_LB\_update(policy)}$ % (new\_flows, ports, policy)}$
-
-	\State $flows\_per\_port \gets [0, 0, \ldots, 0]$
-	\For {$entry \in policy$}
-	    \State $port \gets entry.output\_port$
-	    \State $flows\_per\_port[port]++$
+    \Statex \emph{\bf Function} $\textit{lb\_update}(\textit{policy})$ % (new\_flows, ports, policy)}$
+
+	\State $\textit{flows\_per\_port} \gets [0, 0, \ldots, 0]$
+	\ForAll {$\textit{entry} \in \textit{policy}$}
+	    \State $\textit{port} \gets \textit{entry.output\_port}$
+	    \State $\textit{flows\_per\_port}[\textit{port}]++$
 	\EndFor
-	\State $new\_flow\_mod\_cmds \gets \{\}$
-	\For {$entry \in new\_flows$}
-	    \State $port \gets \arg\min_{p\in ports} flows\_per\_port[p]$
-	    \State $entry.output\_port \gets port$
-	    \State $flows\_per\_port[port]++$
-	    \State $cmd \gets FlowMod(add,f)$
-	    \State $new\_flow\_mod\_cmds.add(cmd)$
+	\State $\textit{new\_flow\_mod\_cmds} \gets \{\}$
+	\For {$\textit{entry} \in \textit{new\_flows}$}
+	    \State $\textit{port} \gets \arg\min_{p\in \textit{ports}} \textit{flows\_per\_port}[p]$
+	    \State $\textit{entry.output\_port} \gets \textit{port}$
+	    \State $\textit{flows\_per\_port}[\textit{port}]++$
+	    \State $\textit{cmd} \gets \textit{FlowMod}(add,f)$
+	    \State $\textit{new\_flow\_mod\_cmds}.\textit{add}(\textit{cmd})$
 	\EndFor
-	\Return $new\_flow\_mod\_cmds$
-    \State \emph{\bf EndFunction}
+	\Return $\textit{new\_flow\_mod\_cmds}$
+  %  \State \emph{\bf EndFunction}
 \end{algorithmic}
 \end{algorithm}
+}
 
@@
 
-\vspace{2mm}
+\vspace{1mm}
 \noindent\textbf{Example 2: Supporting modifications.}
@@
 
-\noindent\textbf{Configuration space}
+\vspace{1mm}
+\noindent\textbf{Configuration space.}
 %
@@
 %\petr{How to make sure that traffic is not affected by the ``meta'' part?}
-
-\noindent\textbf{Synchronization}
+\vspace{1mm}
+\noindent\textbf{Synchronization.}
  %
@@
 
-\noindent\textbf{Synchronization primitives}
+\vspace{1mm}
+\noindent\textbf{Synchronization primitives.}
 %
@@
 }
-
-Following the work on consistent policy
-updates~\cite{network-update,stn}, we focus on the consistent use of  \emph{policy
-identifiers}.
+%
+We focus first on the consistent use of  \emph{policy identifiers},
+an important notion in consistent policy updates~\cite{network-update,stn}. 
 Intuitively, a controller should
@@
 
-\noindent\textbf{Implementation}\label{sec:t-impl}
+\vspace{1mm}
+\noindent\textbf{Implementation.} %\label{sec:t-impl}
 %
@@
 Algorithm~\ref{alg:claim} for $\{\textit{claim}$, 
-Algorithm~\ref{alg:checkclaim} for $\textit{checkclaim}$, 
+Algorithm~\ref{alg:check} for $\textit{checkclaim}$, 
 Algorithm~\ref{alg:unclaim}
@@
 
-\noindent\textbf{The id-claimer}
-
+\vspace{1mm}
+\noindent\textbf{The id-claimer.}
+%
 In Algorithms \ref{alg:claim}-\ref{alg:unclaim},
@@
 
+{\small
+
 \begin{algorithm}[t]
@@
 %    \Ensure
-    		\State $value \gets x\concat x$
-    		\State $mask \gets self\concat \texttt{1$\ldots$1}$
-	    	\State $match \gets (value,mask)$
+    		\State $\textit{value} \gets x\concat x$
+    		\State $\textit{mask} \gets \textit{self}\concat \texttt{1$\ldots$1}$
+	    	\State $\textit{match} \gets (\textit{value},\textit{mask})$
     		%\State $cookie \gets x\concat self$
-    		\State $action \gets 1$
-    		\State $flag \gets 0$
-    		\State $cmd\gets \emph{FlowMod}(match, op = \add, flag, action) $
-			\Return $cmd$
+    		\State $\textit{action} \gets 1$
+    		\State $\textit{flag} \gets 0$
+    		\State $\textit{cmd}\gets \textit{FlowMod}(match, op = \add, \textit{flag}, \textit{action}) $
+			\Return $\textit{cmd}$
     \end{algorithmic}
@@
 %    \Ensure
-    		\State $value \gets x\concat x$
-    		\State $mask \gets self\concat \texttt{1$\ldots$1}$
-    		\State $match \gets (value,mask)$
+    		\State $\textit{value} \gets x\concat x$
+    		\State $\textit{mask} \gets self\concat \texttt{1$\ldots$1}$
+    		\State $\textit{match} \gets (value,mask)$
     		%\State $cookie \gets x\concat self$
-    		\State $action \gets 2$
-    		\State $flag \gets \texttt{ofpff\_\checko}$
-    		\State $cmd1\gets \emph{FlowMod}(match, op = \add, flag, action) $
-    		\State $cmd2\gets \emph{FlowMod}(match, op = \dele) $
-			\Return $cmd1$,$cmd2$
+    		\State $\textit{action} \gets 2$
+    		\State $\textit{flag} \gets \texttt{ofpff\_\checko}$
+    		\State $\textit{cmd1}\gets \textit{FlowMod}(\textit{match}, op = \add, \textit{flag}, \textit{action}) $
+    		\State $\textit{cmd2}\gets \textit{FlowMod}(\textit{match}, op = \dele) $
+			\Return $\textit{cmd1}$,$\textit{cmd2}$
     \end{algorithmic}
@@
 %    \Ensure
-    		\State $value \gets x\concat x$
-    		\State $mask \gets self\concat \texttt{1$\ldots$1}$
+    		\State $\textit{value} \gets x\concat x$
+    		\State $\textit{mask} \gets self\concat \texttt{1$\ldots$1}$
     		%\State $cookie \gets x\concat self$
-    		\State $match \gets (value,mask)$
-    		\State $cmd\gets \emph{FlowMod}(match, op=\dele) $
+    		\State $\textit{match} \gets (\textit{value},\textit{mask})$
+    		\State $\textit{cmd}\gets \textit{FlowMod}(\textit{match}, op=\dele) $
 
@@
 \end{algorithm}
-
-\noindent\textbf{Write and compare}
+}
+
+\vspace{1mm}
+\noindent\textbf{Write and compare.}
 The memory \memwrite and \compare
@@
 $\textit{\compare}(\addr, old)$ followed by $\textit{\memwrite}(\addr,
-new)$ within the same bundle.
+new)$ within the {\execatomic} construct.
 
@@
 
+{\small
 \begin{algorithm}[t]
@@
     \begin{algorithmic}[1]
-    \Require unsafe policy update function \textsc{\ufunc}, policy id $\addr$
+    \Require policy update function \textsc{\ufunc}, policy id address $\paddr$
     \Ensure installed policy is consistent with previous one
@@
  		\Repeat
- 			\State $\pid,policy\gets$ \textbf{read-config} %switch configuration
- 			\State $update\_cmds\gets \textsc{\ufunc}(policy)$
+ 			\State $\pid,\textit{policy}\gets$ \textbf{read-config} %switch configuration}
+ 			\State $\textit{update\_cmds}\gets \textsc{\ufunc}(policy)$
  			\startTxn
 	 			\State $CAS(\paddr,\pid,\pid+1)$,
-	 			\State $update\_cmds$ %conditioned on $metadata=my\_a$
+	 			\State $\textit{update\_cmds}$ %conditioned on $metadata=my\_a$
 	 			%\State update policy selector to $my\_a$
@@
      		%\State $res \gets $ commit transaction
-     	\Until{$res=\ack$}
+     	\Until{$\textit{res}=\ack$}
      	%\doWhile{$res=False$}
 %   \vspace{3mm}
-			\Return
+			\Return $\textit{res}$
 
@@
 Note that in Algorithm \ref{alg:update}, the meta-configuration is
-used to store $pid$ and id-claimer.
+used to store $\textit{pid}$ and id-claimer.
 \petr{The description is not very clear.}
@@
 In other words, Algorithm~\ref{alg:simple-update} can also be used
-to render our naive load-balancing algorithm (Algorithm~\ref{alg:naive-lb}) from Section~\ref{sec:motivation}
+to render our na\"ive load-balancing scheme (Algorithm~\ref{alg:naive-lb} in Section~\ref{sec:motivation})
 consistent, without any modifications.
@@
     \begin{algorithmic}[1]
-        \Require unsafe policy update function \textsc{\ufunc}, a predefined switch (memory) address for policy id: $\paddr$, a predefined id space $C$.
+        \Require policy update function \textsc{\ufunc}, policy id address $\paddr$, id space $C$.
     \Ensure installed policy is consistent with previous one
@@
  			\State $\textbf{execute}\{claim(\pid)\}$
- 			\State $\pid 2,claims,policy\gets$ \textbf{read-config} %switch configuration
+ 			\State $\pid 2,\textit{claims},\textit{policy}\gets$ \textbf{read-config} %switch configuration
  			\If {$\pid\neq \pid 2$}
-	 			\State $\textbf{execute} \{unclaim(\pid)\}$
+	 			\State $\textbf{execute} \{\textit{unclaim}(\pid)\}$
  				\State {\bf continue} (\textbf{restart} loop)
  			\EndIf
- 			\State $my\_id\gets$ choose a number from $C\setminus claims$
- 			\State $update\_cmds\gets \textsc{\ufunc}(policy)$
+ 			\State $\textit{my\_id}\gets$ choose a number from $C\setminus claims$
+ 			\State $\textit{update\_cmds}\gets \textsc{\ufunc}(policy)$
  			\startTxn
@@
 	 			\State $CAS(\paddr, \pid,my\_id)$,
-	 			\State $update\_cmds$ %conditioned on $metadata=my\_a$
+	 			\State $\textit{update\_cmds}$ %conditioned on $metadata=my\_a$
 	 			%\State update policy selector to $my\_a$
  			\endTxn
-	 		\State $\textbf{execute} \{unclaim(\pid)\}$
+	 		\State $\textbf{execute} \{\textit{unclaim}(\pid)\}$
      		%\State $res \gets $ commit transaction
@@
 %   \vspace{3mm}
-			\Return
+			\Return $\textit{res}$
 

### 2015-06-01 inband-sync.tex
expanding alg 8 description

@@
 
+%Algorithm~\ref{alg:update} uses our id-claimer abstraction
+%to bound the number of policy identifiers.
+%Claiming a policy id prevents another controller from using it with a
+%different policy.
+%A controller claims the current policy identifier, chooses a new id
+%from a finite set of identifiers,
+%and then checks that the new policy id is not claimed by another controller.
+%Note that in Algorithm \ref{alg:update}, the meta-configuration is
+%used to store $\textit{pid}$ and id-claimer.
+%\petr{The description is not very clear.}
+
+{\color{red}
 Algorithm~\ref{alg:update} uses our id-claimer abstraction
 to bound the number of policy identifiers.
-Claiming a policy id prevents another controller from using it with a
+The controller begins by reading the current policy id and claiming it in order to prevent another controller from using it with a
 different policy.
-A controller claims the current policy identifier, chooses a new id
-from a finite set of identifiers,
-and then checks that the new policy id is not claimed by another controller.
+Then it reads the configuration again in case the policy was changed just before claiming its id.
+Then considering other claims the controller chooses an unused id for the update and computes the the update commands using the UPDATE function.
+Finally a transaction is made to verify that the chosen id is not claimed and that the policy wasn't changed by someone else, and to update the policy.
 Note that in Algorithm \ref{alg:update}, the meta-configuration is
-used to store $\textit{pid}$ and id-claimer.
-\petr{The description is not very clear.}
+used to store $\textit{pid}$ and id-claimer (the claims).
+}
 

### 2015-06-01 inband-sync.tex
a few minor changes

@@
 the OpenFlow \emph{add flow monitor} command with the \textsf{ofpt\_initial} optional flag.
-In this paper, we write \textit{read-config} for sending
+In this paper, we write \textbf{read-config} for sending
 this message, receiving a response, and returning the received configuration.
@@
 \vspace{1mm}
-\noindent \textbf{Example 1: Handling complex dependencies.} Let us consider two controllers
+\noindent \textbf{Example~1: Handling complex dependencies.} Let us consider two controllers
 in charge
@@
 In such a scenarios, seemingly independent actions, defined over
-completely independent logical flow spaces (say, two flows with source IP address ranges $1*$ and $0*$,
-respectively),
+completely independent logical flow spaces (say, flows with
+originated in disjoint ranges of IP addresses),
 may actually be dependent: the flows share the underlying physical network.
@@
     \State $\textit{policy} \gets \textbf{read\_config}$
-    \State $\textit{update\_cmds} \gets \textit{lb\_update}(policy)$ %(new\_flows, ports, policy)$    
+    \State $\textit{update\_cmds} \gets \textsc{lb\_update}(policy)$ %(new\_flows, ports, policy)$    
     \State $\textbf{send}\{\textit{update\_cmds}\}$
@@
 
-    \Statex \emph{\bf Function} $\textit{lb\_update}(\textit{policy})$ % (new\_flows, ports, policy)}$
+    \Statex \emph{\bf Function} $\textsc{lb\_update}(\textit{policy})$ % (new\_flows, ports, policy)}$
 
@@
 \vspace{1mm}
-\noindent\textbf{Example 2: Supporting modifications.}
+\noindent\textbf{Example~2: Supporting modifications.}
 Consider the arguably very basic problem of \emph{modifying} previously installed forwarding rules
@@
 have to be sent as part of a bundle. Therefore, the \texttt{\claimcheck} operation is atomic by nature,
-similar to the \texttt{claim} and \texttt{unclaim} operations, and is implemented as essentially one command.
+similar to the \texttt{claim} and \texttt{unclaim} operations, and is
+implemented as essentially one command.
+
+By employing the \textbf{read-config} command, we may easily implement
+an additional operation that returns all currently claimed ids: it is
+sufficient to go over the rules matching every controller identifier.
 
@@
 implement concurrent policy compositions of a user-specific
-$\textsc{\ufunc}()$ function~\cite{stn}.
+$\textsc{\ufunc}()$ function~\cite{stn}, e.g., our load-balancing
+function $\textsc{lb\_update}$ (Example~1 in Section~\ref{sec:motivation}).   
 That is, let us consider a simple policy update protocol that
@@
 Algorithm~\ref{alg:update} uses our id-claimer abstraction
-to bound the number of policy identifiers.
-The controller begins by reading the current policy id and claiming it in order to prevent another controller from using it with a
-different policy.
-Then it reads the configuration again in case the policy was changed just before claiming its id.
-Then considering other claims the controller chooses an unused id for the update and computes the the update commands using the UPDATE function.
-Finally a transaction is made to verify that the chosen id is not claimed and that the policy wasn't changed by someone else, and to update the policy.
-Note that in Algorithm \ref{alg:update}, the meta-configuration is
-used to store $\textit{pid}$ and id-claimer (the claims).
+to bound the number of used policy identifiers.
+Here the controller first reads the current policy id and claims it in order to prevent another controller from using it with a
+different policy (lines~\ref{update:read-config1}-\ref{update:read-config2}).
+%Then it reads the configuration again in case the policy was changed just before claiming its id.
+%Then considering other claims the controller chooses 
+Then the controller claims the first \emph{unused} id and computes the update commands to be executed (line~\ref{update:compute}).
+Finally, within an atomic transaction (\textbf{execatomic}), it verifies that the chosen policy id is not claimed and that 
+the policy has not been concurrently modified, and updates the policy.
+%[[PK added a note to the id-claimer description
+%Note that in Algorithm \ref{alg:update}, the meta-configuration is
+%used to store $\textit{pid}$ and id-claimer (the claims).
+
+Algorithm~\ref{alg:simple-update} can therefore also be used
+to render our na\"ive load-balancing scheme (Algorithm~\ref{alg:naive-lb} in Section~\ref{sec:motivation})
+consistent, without any modifications in the update function.
 }
-
-In other words, Algorithm~\ref{alg:simple-update} can also be used
-to render our na\"ive load-balancing scheme (Algorithm~\ref{alg:naive-lb} in Section~\ref{sec:motivation})
-consistent, without any modifications.
 
@@
  		\Repeat
-		 	\State $\pid,claims,policy\gets$ \textbf{read-config} %switch configuration
+		 	\State $\pid,claims,policy\gets$ \textbf{read-config} \label{update:read-config1} %switch configuration 
  			\State $\textbf{execute}\{claim(\pid)\}$
- 			\State $\pid 2,\textit{claims},\textit{policy}\gets$ \textbf{read-config} %switch configuration
+ 			\State $\pid 2,\textit{claims},\textit{policy}\gets$ \textbf{read-config} \label{update:read-config2} %switch configuration 
  			\If {$\pid\neq \pid 2$}
@@
  			\State $\textit{my\_id}\gets$ choose a number from $C\setminus claims$
- 			\State $\textit{update\_cmds}\gets \textsc{\ufunc}(policy)$
+ 			\State $\textit{update\_cmds}\gets \textsc{\ufunc}(policy)$ \label{update:compute}
  			\startTxn
@@
 protocol: they do not require any protocol/hardware extensions as postulated in recent literature~\cite{stn,netpaxos}.
+{\color{red}
+We also conjecture that they can be extended to implement atomic modifications of multiple switches.   
+}
 Our work may hence also contribute to the ongoing discussion of what can be implemented

### 2015-06-01 inband-sync.tex
first complete draft

@@
 %\title{On the Synchronization Power of the OpenFlow Data Plane}
-\title{Network Concurrency Control Done In Band}
+\title{In-Band Synchronization for\\Distributed SDN Control Planes}
 %\\ {\Large Towards Atomic Transactions In-Band}}
@@
 We also discuss a simple CAS-based concurrent policy-update mechanism,
-and also use our synchronization abstraction to provide
+and use our synchronization abstraction to provide
 the missing link for the read-modify-write object
@@
 In our implementations, we are going to use bitmasking of the \emph{metadata}
-field for performing conditional operations on  its content.
+field for performing conditional operations based on  its content.
 %For example, \texttt{ipv4\_src=10.1.14.0, ff.ff.ff.f0} matches all packets
@@
 \vspace{1mm}
-\noindent\textbf{Interaction between the planes: FlowMod commands}
+\noindent\textbf{Interaction between the planes: FlowMod commands.}
 The flow tables installed on the switches across the data plane
@@
 match and priority,  or does nothing if such an entry is not present.
+%Stefan: alternatively could also delete an entry by using the cookie field
+
 %\petr{Is this correct?}
@@
 should or should not take place, is often desirable.
-
+%
 %But also bundling FlowMod commands with the check overlap feature ($\checko$)
@@
 % \liron{replaced transaction in the last sentence with " atomic update"}
-
+%
 Second, sometimes it is desirable to react to conflicts in smarter ways than
@@
 for a simplistic example with two links.
-In such a scenarios, seemingly independent actions, defined over
-completely independent logical flow spaces (say, flows with
-originated in disjoint ranges of IP addresses),
+In such a scenario, seemingly independent actions, defined over
+completely independent logical flow spaces (say, two different
+TCP micro-flows),
+%originated in disjoint ranges of IP addresses),
 may actually be dependent: the flows share the underlying physical network.
@@
 by using a minimal
-``meta-configuration'' space, i.e., controllers store information 
+``meta-configuration'' space. That is, controllers store information
 on contention and conflicts between
@@
 maintain higher priority levels.
-Alternatively, we could leverage the standard ability of maintaining multiple
+Alternatively, we could leverage the possibility to maintain multiple
 flow tables at an OpenFlow switch, and use a separate flow table for
@@
 trivial. If $op_i$ is a synchronization primitive, we employ
-Algorithm~\ref{alg:claim} for $\{\textit{claim}$, 
+Algorithm~\ref{alg:claim} for $\textit{claim}$,
 Algorithm~\ref{alg:check} for $\textit{checkclaim}$,
@@
 Algorithm~\ref{alg:write} for $\textit{write}$ and
-Algorithm~\ref{alg:compare} for $\textit{compare}\}$.
+Algorithm~\ref{alg:compare} for $\textit{compare}$.
 
@@
 fails and takes no effect.
+Algorithm~\ref{alg:simple-update} can therefore also be used
+to render our na\"ive load-balancing Algorithm~\ref{alg:naive-lb}
+consistent, without any modifications in the update function.
+
 % \liron{the term composition doesn't make sense now with the new user update function... }
@@
 
-{\color{red}
-Algorithm~\ref{alg:update} uses our id-claimer abstraction
+Algorithm~\ref{alg:update} is a more efficient alternative 
+to Algorithm~\ref{alg:simple-update}. It uses our id-claimer abstraction
 to bound the number of used policy identifiers.
-Here the controller first reads the current policy id and claims it in order to prevent another controller from using it with a
-different policy (lines~\ref{update:read-config1}-\ref{update:read-config2}).
+Here, the controller first reads the current policy id and claims to prevent another controller from using it 
+for a
+different policy (Lines~\ref{update:read-config1}-\ref{update:read-config2}).
 %Then it reads the configuration again in case the policy was changed just before claiming its id.
 %Then considering other claims the controller chooses
-Then the controller claims the first \emph{unused} id and computes the update commands to be executed (line~\ref{update:compute}).
-Finally, within an atomic transaction (\textbf{execatomic}), it verifies that the chosen policy id is not claimed and that 
-the policy has not been concurrently modified, and updates the policy.
+Then the controller claims the first \emph{unused} id and computes the update commands to be executed (Line~\ref{update:compute}).
+Finally, within an atomic transaction (\textbf{execatomic}), it verifies whether the chosen policy id has not
+been claimed and that
+the policy has not been concurrently modified, and then updates the policy.
 %[[PK added a note to the id-claimer description
@@
 %used to store $\textit{pid}$ and id-claimer (the claims).
-
-Algorithm~\ref{alg:simple-update} can therefore also be used
-to render our na\"ive load-balancing scheme (Algorithm~\ref{alg:naive-lb} in Section~\ref{sec:motivation})
-consistent, without any modifications in the update function.
-}
 
@@
 protocol: they do not require any protocol/hardware extensions as postulated in recent literature~\cite{stn,netpaxos}.
-{\color{red}
-We also conjecture that they can be extended to implement atomic modifications of multiple switches.   
-}
 Our work may hence also contribute to the ongoing discussion of what can be implemented
@@
 language abstractions.~\cite{pyretic}
+We also plan to release a small library with our synchronization primitives. 
+Moreover, we are investigating possible generalizations of our transactional guarantees,
+across multiple switches~\cite{stn}, as well as opportunities for a more fine-grained
+locking, improving concurrency.
 

### 2015-06-01 inband-sync.tex
phrasing

@@
 to improve load-balancing, and to reduce overheads,
-control modules operating on the data plane should be physically distributed.
+modules of the control plane %operating on the data plane 
+should be physically distributed.
 %a redundant control plane
@@
 
-We believe that our synchronization mechanisms
-can be useful in many situations.
-For example, the pathological scenario described in
-Figure~\ref{fig:CAS-example} can be resolved easily if we put
-the policy update in our transactional $\execatomic$ construct together
- with a \emph{Compare-and-Set} (CAS) operation on the policy identifier.
-A $\cas(\addr,\textit{old},\textit{new})$ operation on a memory location $\addr$
+Our synchronization mechanisms can be useful in many situations.
+For example, it can be used to implement the fundamental
+\emph{Compare-and-Set} (CAS) operation. % on the policy identifier.
+When executed within a
+transaction, a $\cas(\addr,\textit{old},\textit{new})$ operation 
+on a memory location $\addr$
 checks if the content of $\addr$ is $\textit{old}$ and if so,
-replaces it with $\textit{new}$ and returns $\true$ (the CAS
-\emph{succeeds}), otherwise it returns $\false$ (the CAS
-\emph{fails} which causes an abort of
-the invoking transaction).
+replaces it with $\textit{new}$ %and returns $\true$ 
+(the CAS \emph{succeeds}), otherwise it 
+%returns $\false$ 
+causes an abort of the invoking transaction
+(the CAS \emph{fails}).
 It is straightforward to implement the CAS operation by executing
@@
 
-More generally, CAS can be used to solve the fundamental
+CAS can be used to solve the fundamental
 problem of fault-tolerant \emph{consensus}~\cite{FLP85}
@@
 implement concurrent policy compositions of a user-specific
-$\textsc{\ufunc}()$ function~\cite{stn}, e.g., our load-balancing
-function $\textsc{lb\_update}$ (Example~1 in Section~\ref{sec:motivation}).
-That is, let us consider a simple policy update protocol that
+$\textsc{\ufunc}()$ function~\cite{stn}.
+This way, for example, by choosing our load-balancing
+function $\textsc{lb\_update}$ (Example~1 in
+Section~\ref{sec:motivation}), 
+the pathological scenario described in 
+Figure~\ref{fig:CAS-example} can be resolved.
+% if we put
+%the policy update in our transactional $\execatomic$ construct together
+% with a \emph{Compare-and-Set} (CAS) operation on the policy identifier.
+
+Let us first  consider a simple policy update protocol that
 allows a set of  controllers to concurrently update switch policies, i.e., sets of
@@
 Then the controller claims the first \emph{unused} id and computes the update commands to be executed (Line~\ref{update:compute}).
-Finally, within an atomic transaction (\textbf{execatomic}), it verifies whether the chosen policy id has not
+Finally, within an atomic transaction (\textbf{\execatomic}), it verifies whether the chosen policy id has not
 been claimed and that
@@
 Moreover, we are investigating possible generalizations of our transactional guarantees,
-across multiple switches~\cite{stn}, as well as opportunities for a more fine-grained
-locking, improving concurrency.
+across multiple switches~\cite{stn}, as well as opportunities for more fine-grained
+\emph{partial} updates that concern subsets of flow entries, thus improving concurrency.
 

### 2015-08-11 ccr-rev1/inband-sync.tex
CCR rev 1 added

--- 2015-06-01 inband-sync.tex
@@
 %
-While the perspective of \emph{logically centralized} control plane
+While the perspective of a \emph{logically centralized} control plane
 offered by SDN is intuitive and attractive,
@@
 distributed-systems
-one: Multiple controllers may simultaneously try to
+one: multiple controllers may simultaneously try to
 install conflicting updates and we want to resolve these conflicts
@@
 If none of the transaction's commands \emph{conflicts}  with the current
-configuration, where conflicts can be
+configuration, where a conflict can be
 defined in a general, application-specific manner, the transaction appears to be executed atomically.
-Otherwise, the transaction is \emph{aborted} in its entirety and does not
-affect neither traffic nor other controllers.
-
-We promote to implement our synchronization constructs \emph{in-band},
+Otherwise, the transaction is \emph{aborted} in its entirety and 
+affects neither traffic nor other controllers.
+
+We propose to implement our synchronization constructs \emph{in-band},
 on the switch.
@@
 primitives and motivate the need for transactions.
-We then present our approach in detail in Section~\ref{sec:main}, and discuss use cases in Section~\ref{sec:apps}.
+We present our approach in detail in Section~\ref{sec:main}, and
+discuss use cases in Section~\ref{sec:apps}.
 %Section~\ref{sec:compo} describes how to perform in-band composition.
@@
 field for performing conditional operations based on  its content.
-%For example, \texttt{ipv4\_src=10.1.14.0, ff.ff.ff.f0} matches all packets
-%with source IP addresses in the range \texttt{10.1.14.0}--\texttt{10.1.14.127}.
+%[[PK put the example back in response of review 2
+For example, \texttt{ipv4\_src=10.1.14.0, ff.ff.ff.f0} matches all packets
+with source IP addresses in the range \texttt{10.1.14.0}--\texttt{10.1.14.127}.
+%]]
 
@@
 current configuration,
-However, as we will argue, a support for more generic conditions under which configuration
+However, as we will argue, support for more generic conditions under which configuration
 updates
@@
 in a consistent manner.
-In principle, a bundle allows to conditionally add new rules or remove
+In principle, a bundle allows us to conditionally add new rules or remove
 old rules, and, using the flow space overlap
@@
 Flow entries in the meta-configuration can be referred to as shared
-memory locations, provided a total addressing scheme.
+memory locations, each provided with a distinct \emph{address}.
 
@@
 
-We propose the following constructs
+We introduce two constructs
 %$\textbf{bundle}\{\;\ldots\;\}\rightarrow res$ to refer to represent a
@@
 a regular and atomic way, respectively.
-The construct returns $\ack$ if the sequence of operations is performed and all commands are performed successfully, and
-$\error(\xid,\ecode)$ received from the switch otherwise, providing
+The construct $\execatomic\{op_1,\;\ldots\;op_k\}$  
+returns $\ack$ if and only if all commands in the sequence are performed successfully; otherwise
+$\error(\xid,\ecode)$ is returned, providing
 the identifier of the first command in the sequence that failed
@@
 
+We now describe sample scenarios of how of our synchronization mechanisms
+can be used for control applications. We aim at two applications: a
+simple load-balancing scheme and a more sophisticated forwarding policy
+composition, but we begin with implementing a fundamental higher-level
+\emph{Compare-and-Set} (CAS) primitive. We then sketch show to use CAS, along with
+     {\exec} and {\execatomic} constructs, in our two control applications.
+
 Our synchronization mechanisms can be useful in many situations.
-For example, it can be used to implement the fundamental
-\emph{Compare-and-Set} (CAS) operation. % on the policy identifier.
+
+\vspace{1mm}
+\noindent\textbf{Compare-and-Set.}
+%For example, we can implement the fundamental
+%\emph{Compare-and-Set} (CAS) primitive. % on the policy identifier.
 When executed within a
@@
 
+\vspace{1mm}
+\noindent\textbf{Load balancing.}
 CAS can be used to solve the fundamental
-problem of fault-tolerant \emph{consensus}~\cite{FLP85}
+problem of fault-tolerant \emph{consensus}~\cite{FLP85},
 \hide{%[[PK, feels like we do not really need this level of detail here
@@
 }
-and allows us
+which allows us
 to implement any generic replicated state machine~\cite{Her91}
@@
 
+\vspace{1mm}
+\noindent\textbf{Generic policy composition.}
 Our mechanism can also be used as a generic template to

### 2015-08-13 reviews/response-rev1.txt
added reviews and response

Dear Dina, dear reviewers:

thank you very much for your time and good feedback!

We revised our paper according to your suggestions.

In the following, we discuss the most important changes.

We hope that the new version meets your expectations, and are happy to take any further feedback.

With kind regards,

Liron, Petr, Stefan

### 2015-08-25 reviews/response-rev1.txt
rev plan

@@
 Liron, Petr, Stefan
+
+Reviewer 1
+==========
+
+> Overall, the work seems useful and correct.
+> My main concern is scope.
+> The paper is essentially a note on OpenFlow hacking, and thus might be
+> of limited interest to many CCR readers. However, OpenFlow and similar
+> SDN approaches have seen sufficient adoption and research interest in
+> recent years that there is utility in this paper appearing in a venue with
+> broad reach.
+
+FIXME: discuss more broadly the interface and meta data approach and examples?
+
+> A lesser concern is novelty. With the addition of atomic bundles to
+> OpenFlow, it's not completely surprising that the atomic CAS operation
+> can be implemented, but there is definite value in having it spelled out
+> cleanly.
+
+FIXME: it is non trivial and inband perspective novel
+
+> I found the brief argument for inband
+> control vs outofband control unconvincing. There are many highperformance
+> systems that use consensus, and a full architectural
+> comparison of the two approaches to
+> SDN is beyond the scope of this paper. I certainly believe that inband
+> control is a useful and interesting design choice, but not just because of
+> FLP. It also seems almost certain that a full system would require outofband
+> coordination between controllers: a full transactional system needs
+> application specific ways of resolving conflicts, which will require
+> coordination.
+
+FIXME: right, that is what we are currently exploring also in related work
+
+> I also found 2nd example of the limitations of bundles confusing. A
+> simple concrete example would be greatly helpful, and strength the
+> argument.
+
+FIXME
+
+> Typos/corrections:
+> p.1 col. 2 para. (Contributions): 2nd "conflicts" => "a conflict"
+
+FIXME
+
+> "does not affect" => "affects neither"
+
+FIXME
+
+> " " para. 3: "promote" => "propose" (?)
+
+FIXME
+
+> p. 3 col. 2 para (Configuration space): "provided a total addressing scheme" sentence is confusing.
+
+FIXME
+
+Reviewer 2
+==========
+
+> The idea of the paper is really nice and I believe it can be used in
+> practice. However, the presentation of the paper needs major work it
+> feels to be switching between highlevel
+> overview versus lowlevel details is
+> background section necessary? Cannot you introduce details like
+> bundles, overlap_checking and actual OpenFlow flags progressively and
+> on examples?
+> maybe start with another example with controllers updating
+> overlapping flows. Then explain ability to do overlap_check. Then
+> introduce loadbalancing example and the need for external
+> synchronization. Transactions/bundles are kindof
+> orthogonal and could
+> be explained after you state that you want executeatomic{ops} and
+> that there is a native support for it in OpenFlow if ops are OF commands.
+> Figure 1 time axis is too inconspicous
+> when you described execute{op1,...} and executeatomic{
+> op1, ...} I was totally confused what is the definition good for. Only later on I
+> realized that operations there are actually not OpenFlow operations but
+> your custom synchronization primitives. Please, clarify that in the text.
+> what does "mask=1...1" mean? Is it a) exact match or b) everything
+> wildcarded? From analysing algorithm 5 I believe it is a)
+
+> Do the algorithms 24 really need to use value=x.x?
+> Cannot it be "0..0 . x" ? Or "1...1 . x"? Or for an arbitrary fixed
+> constant "c.x" ? Maybe elaborate in the text on that point because it
+> looks like magic to me.
+> In any case, you should explain more why "x masked by self1"
+> overlaps with "x masked by self2" exactly iff self1 != self2
+> On my first reading I missed that claim() operation can actually allow
+> more controllers to claim the same value. Please clarify in the text.
+> What is the value of such operation? Actually, it looks like a kindof
+> "read lock". following the thought, additionally to claim() and unclaim(), could you
+> have algorithms for exclusive lock() and unlock()? Maybe even explain it
+> sooner than claim() so that people can think about wellknown
+> semantic first.
+> There looks to be an inconsistency of what happens with check_overlap
+> flag for the exactly same rule:
+> You say "a FlowMod command can be equipped with a the check
+> overlap flag: if the switch maintains a flow entry with an overlapping but
+> not identical match part with the same priority but a different action,
+> and if the two rules have the same priority, then FlowMod will fail."
+> OF 1.4 spec says "For add requests (OFPFC_ADD) with the
+> OFPFF_CHECK_OVERLAP flag set, the switch must first check for any
+> overlapping flow entries in the requested table. Two flow entries overlap
+> if a single packet may match
+> both, and both entries have the same priority. If an overlap conflict
+> exists between an existing flow entry and the add request, the switch
+> must refuse the addition and respond with an ofp_error_msg with
+> OFPET_FLOW_MOD_FAILED type and OFPFMFC_OVERLAP code."
+> the consequence is that I believe your explanation allows for no error
+> if we add exactly the same rule but the spec requires an error. How do
+> real switches implement this? Does this affect the validity of your
+> algorithms?
+> in Algorithm 6 you seems to be assuming that same match with same
+> actions won't cause an error but same match different actions will. This
+> is neither consistent with your explanation of overlap check nor the spec
+> Algorithm 3 I believe commands cmd1 and cmd2 must be either
+> bundled together with guaranteed ordering or separated by a barrier,
+> otherwise the switch might reorder them and render the cleanup
+
+> Why is Algorithm 8 "more efficient alternative to Algorithm 7" ? It
+> looks to be doing much more work
+
+Reviewer 3
+==========
+
+> This is a intriguing little article that shows that a recent version of
+> OpenFlow (which version?!) has the raw ingredients needed to implement a very general
+> transactional interface for updating flow tables. It is obvious that this is
+> an accident. The OpenFlow primitives that the article uses actually have a
+> very narrow interface, but the paper uses them in a clever way to build very
+> general transactions: a transaction is an atomic sequence of flowtable
+> writes and assertions. i.e., the writes succeed or fail atomically only if all the
+> assertions are true.
+> This article is primarily a theoretical exercise and doesn't address
+> performance concerns:
+> Do real switches have the space to use all these extra rules for
+> coordination?
+> Do switches actually implement the features of OpenFlow used by this article
+> (i.e., check_overlap, bundles, arbitrary bitmasking of the METADATA field,
+> etc.) and do they implement them fast?
+> Being optimistic, I think that the real value of this article may be to
+> show that OpenFlow might as well provide a more general transactional interface.
+> Without any empirical evidence, I doubt that the scheme proposed in the article
+> will actually work in practice.
+
+> A deeper issue that this article doesn't address is whether this kind of
+> abstraction is even desirable. A simpler abstraction may be to have one
+> controller for a small set of switches and have the controllers coordinate
+> with each other.
+> Transactions are a powerful primitive, but apart from the simple load
+> balancer in Algorithm 1, the paper doesn't delve into how applications should be
+> built to take advantage of the transactional primitive. I think this is what
+> the article sorely lacks. Without it, the article just shows off cool hack.
+
+> MINOR
+> pg 1 col 1: "perspective of [a] logically centralized"
+
+> pg 1 col 1: "a distributedsystems one: [multiple]"
+
+> pg 1 col 2: perhaps paper article
+
+> pg 1 col 2: "We [propose] to implement our"
+
+> pg 3 col 1: "as we will argue, asupport for more generic" (delete a)
+
+> pg 3 col 2: "a bundle allows to conditionally"
+
+> In Example 1, why not have the two controllers communicate
+
+> Example 2 says bundling and checkoverlap don't help modify. But, Section 2
+> says that the semantics of addflow is to modify if a flow with exactly
+> the same priority and match exist
+
+EDITOR
+======
+
+> Dear authors,
+> Thank you for your submission. The reviews for this paper raise several
+> important questions and provide good feedback. We hope they will help
+> in improving the paper. Beyond addressing the technical questions and
+> presentationrelated
+> issues mentioned in the reviews, a revised version
+> of the paper should include:
+> Examples
+> showing how transactions can be used by control
+> applications. This will make the paper storyline richer.
+> Experiments
+> with actual switches to give readers some confidence that
+> the proposed transactions can be efficiently implemented.
+> Regards. Area Editor.

### 2015-08-31 reviews/response-rev1.txt
some updates

@@
 
-FIXME: discuss more broadly the interface and meta data approach and examples?
+Thank you very much for your feedback. Indeed, we hope and believe that
+our work is of relevance beyond specific Openflow versions. In our revision,
+we aim to put the fundamental aspects and the benefit of inband synchronization
+into a broader context. Having that said, in our revision we also report
+in more detail on our proof-of-concept implementation based on POX.
 
@@
 
-FIXME: it is non trivial and inband perspective novel
+We agree, and believe that the implementations with and without bundles
+is a non-trivial step forward.
 
@@
 
-FIXME: right, that is what we are currently exploring also in related work
+We agree. With this paper, we aim to show feasibility of inband synchronization, 
+and we offer a simple interface of synchronization primitives. While inband mechanisms
+may reduce the round-trip times and reduce the number of elements a protocol relies upon,
+further research is needed to fully answer the out-of-band vs inband
+debate.
 
@@
 
-FIXME
+We reorganized the paper, and hope that the contributions are now
+put into perspective better.
 
@@
 
-FIXME
+Done, thanks.
 
@@
 
-FIXME
+Thanks.
 
@@
 
-FIXME
+We changed this part, thank you.
 
@@
 
+This is a very good suggestion, and in our revision, we restructured the paper. 
+We hope that the new structure puts the contribution into perspective better.
+
 > Do the algorithms 24 really need to use value=x.x?
@@
 > overlaps with "x masked by self2" exactly iff self1 != self2
+
+Thanks, we tried to clarify this.
+
 > On my first reading I missed that claim() operation can actually allow
@@
 > semantic first.
+
+Thank you, we revised this part and tried to clarify.
+
 > There looks to be an inconsistency of what happens with check_overlap
@@
 > algorithms?
+
+We elaborated more on these technical parts, and also now include a short
+discussion of a proof-of-concept implementation.
+
 > in Algorithm 6 you seems to be assuming that same match with same
@@
 > is neither consistent with your explanation of overlap check nor the spec
+
+We agree that this part was confusing, and changed the description now.
+
 > Algorithm 3 I believe commands cmd1 and cmd2 must be either
@@
 
+FIXME
+
 > Why is Algorithm 8 "more efficient alternative to Algorithm 7" ? It
 > looks to be doing much more work
+
+Indeed, the algorithm seems to be longer, and we now clarified what we mean by more efficient.
 
@@
 
+We indeed understand our article as a first step, and a proof of concept.
+However, much more work is needed to evaluate its performance and scalability,
+and we aim to do this in our future research. Nevertheless, our mechanisms
+can readily be implemented in today's Openflow protocol. In our revision,
+we also report on such an implementation.
+
 > A deeper issue that this article doesn't address is whether this kind of
@@
 
+This is a very good point. While it is difficult to prove that our abstraction
+is the best or even the right one, we believe that our approach represents a meaningful
+first step. Moreover, we believe that our contribution goes beyond an "openflow hack".
+By restructuring the paper, we seek to put our contribution into a more general
+and broader perspective.
+
 > MINOR
@@
 
+Thanks, done.
+
 > pg 1 col 1: "a distributedsystems one: [multiple]"
+
+Fixed.
 
@@
 
+Fixed.
+
 > pg 1 col 2: "We [propose] to implement our"
+
+Fixed.
 
@@
 
+Fixed.
+
 > pg 3 col 2: "a bundle allows to conditionally"
 
+Fixed.
+
 > In Example 1, why not have the two controllers communicate
-
 > Example 2 says bundling and checkoverlap don't help modify. But, Section 2
@@
 > the same priority and match exist
+
+In our revision, we restructured these parts and hope we found
+a omre intuitive way to put this into perspective.
 
@@
 > Regards. Area Editor.
+
+Thank you for your good comments. In our revision, we restructured
+the paper and also elaborate more on the proof of concept implementation,
+which shows the feasibility of our approach. Of course, for an extensive
+study of the performance and overheads, more work is needed. However,
+we hope that our article provides a worthwhile first step and
+contribution for the community.

### 2015-09-01 ccr-rev1/inband-sync.tex
minor update
interim

@@
 in order to guarantee \emph{consistency} of network operation, actions
-performed on the data plane by different controllers may need to be \emph{synchronized},  which is a
+performed on the data plane by different controllers may need to 
+be \emph{synchronized},  which is a
  nontrivial task.
@@
 %
-A transaction aggregates a sequence
-of control messages and specific synchronization primitives,
-providing all-or-nothing semantics.
-We argue that a powerful transactional interface should and can be
-implemented \emph{in-band}, using
-standard OpenFlow commands on the data plane configuration.
+%A transaction aggregates a sequence
+%of control messages and specific synchronization primitives,
+%providing all-or-nothing semantics.
+In particular, we present a powerful interface 
+to implement transactions \emph{in-band}, on the 
+switch itself
+and using
+the data plane configuration as a shared memory.
+We argue that this in-band approach is attractive as 
+it keeps the failure scope local and 
+does not require additional 
+out-of-band coordination mechanisms.
+This allows us 
+to realize fundamental consensus primitives in the presence
+of controller failures, and we discuss applications
+for 
+ consistent policy composition and fault-tolerant control-planes.
+%
 %Indeed, recent OpenFlow versions already come with useful but primitive
@@
 %as well as computations and multiple controller interactions.
-We discuss how to use our synchronization framework to implement
- consistent policy composition and fault-tolerant control-plane applications.
+%
+Interestingly, 
+by leveraging the match-action
+paradigm to realize simple key-value stores, 
+we can implement our synchronization framework in
+today's standard Openflow protocol, 
+and we report on
+a proof-of-concept accordingly.
 \end{abstract}
@@
 
-Our synchronization mechanisms can be useful in many situations.
+%Our synchronization mechanisms can be useful in many situations.
 

### 2015-09-01 ccr-rev1/inband-sync.tex
df

@@
 flow spaces may lead to pathological network behavior (loops,
-blackholes, routes bypassing a firewall, etc.).~\cite{hotnets14update,roger-hotnets}
+blackholes, routes bypassing a firewall, etc.)~\cite{hotnets14update,roger-hotnets}.
 Similarly, installing diverging load-balancing policies may,
@@
 terms of the necessary computability assumptions about the underlying
-system~\cite{FLP85}, and the amount of communication
+system~\cite{FLP85}, and the high communication
 complexity~\cite{Lam06}.
@@
 
-Interestingly, our mechanisms are simple and can be implemented using the standard \emph{OpenFlow}
+The key idea of our approach is to use the switch configuration space
+as a \emph{transactional shared memory}. 
+The shared memory of the data-plane configuration space 
+can store information about  contention and conflicts
+between the controllers.
+A transaction can then contain standard OpenFlow control operations
+as well as  \emph{synchronization primitives} operating on this shared
+memory.
+We investigate useful synchronization abstractions and propose 
+a simple interface which revolves around policy identifiers (IDs)
+which can be claimed and released atomically and hence consistently. 
+More generally, our synchronization primitives allow the controllers to define
+general notions of conflicts between configuration updates,
+not restricted to simple conflicts on the overlapping flow spaces.
+In particular, we can define dependencies between mappings of
+independent flows to the physical infrastructure, which is important,
+e.g., for load-balancing applications.
+
+We complement our conceptual contributions by
+demonstrating that our approach can readily be implemented
+in today's standard \emph{OpenFlow}
  protocol (version 1.4~\cite{openflow}).
@@
 %
-Concretely, in our implementation,
-we use a part of the data-plane configuration space as a \emph{shared
-  memory}  that stores information about  contention and conflicts
-between the controllers.
-A transaction can then contain standard OpenFlow control operations
-as well as  \emph{synchronization primitives} operating on this shared
-memory.
-Our synchronization primitives allow the controllers to define
-general notions of conflicts between configuration updates,
-not restricted to simple conflicts on the overlapping flow spaces.
-In particular, we can define dependencies between mappings of
-independent flows to the physical infrastructure, which is important,
-e.g., for load-balancing applications.
+In particular, we show that OpenFlow's match-action paradigm can be instrumented
+to realize simple key-value stores in the configuration space.
+
 %which can be used even for coordinating
@@
 
-We then show that, using our abstraction, the controllers can
+Using our abstraction, the controllers can
 solve several important synchronization problems.
-We describe an implementation of the fundamental
+We describe an OpenFlow implementation of the fundamental
 \emph{compare-and-set} (CAS) primitive,

### 2015-09-01 ccr-rev1/inband-sync.tex
bla

@@
 The concurrent computing literature offers
-a wide range of synchronization abstractions.
-E.g., the popular \emph{transactional
+a wide range of synchronization abstractions
+to realize consistent distributed systems.
+In particular, the popular \emph{transactional
   memory} abstraction~\cite{tm-book} provides a collection of
@@
 between the controllers.
-A transaction can then contain standard OpenFlow control operations
+A transaction can then contain standard control and update operations
 as well as  \emph{synchronization primitives} operating on this shared
@@
 
-We complement our conceptual contributions by
-demonstrating that our approach can readily be implemented
-in today's standard \emph{OpenFlow}
- protocol (version 1.4~\cite{openflow}).
 %
@@
 %
-In particular, we show that OpenFlow's match-action paradigm can be instrumented
-to realize simple key-value stores in the configuration space.
 
@@
 
+As a case study, and to demostrate the feasibility
+of our conceptual approach, we 
+consider the standard \emph{OpenFlow}
+ protocol (version 1.4~\cite{openflow}).
+ In particular, we show that OpenFlow's match-action paradigm can be instrumented
+to realize simple key-value stores in the configuration space.
+
 Using our abstraction, the controllers can
 solve several important synchronization problems.
-We describe an OpenFlow implementation of the fundamental
-\emph{compare-and-set} (CAS) primitive,
-which in turn can be used to solve consensus and
-implement a generic replicated control plane service~\cite{Her91} in a consistent and
+For example, we describe an OpenFlow implementation of the fundamental
+\emph{compare-and-set} (CAS) primitive.
+The atomic CAS operation is known to provide an infinite consensus number,
+and can for instance be used to 
+implement generic replicated control plane services~\cite{Her91} in a consistent and
 fault-tolerant way: a mainstream building block in
-modern system design.
-We also discuss a simple CAS-based concurrent policy-update mechanism,
-and use our synchronization abstraction to provide
+modern distributed system design (Chubby, Google Spanner, Amazon Webservices, etc.).
+We discuss a simple CAS-based concurrent policy-update mechanism,
+and present a generic template to implement concurrent policy
+compositions of a user-specic update functions.
+Our synchronization abstraction also provides
 the missing link for the read-modify-write object
@@
 The rest of the paper is organized as follows.
-In Section~\ref{sec:background}, we review the basics of the  OpenFlow
-protocol and discuss the features which are relevant for our work.
-In Section~\ref{sec:motivation}, we highlight the limitations of the OpenFlow
-primitives and motivate the need for transactions.
-We present our approach in detail in Section~\ref{sec:main}, and
+In Section~\ref{sec:motivation}, we motivate the 
+transactional approach.
+We present our abstractions in more detail in Section~\ref{sec:main}.
+In Section~\ref{sec:background}, we discuss
+an OpenFlow implementation, 
+and we 
 discuss use cases in Section~\ref{sec:apps}.
@@
 
-\section{Background: SDN and OpenFlow}\label{sec:background}
-%\petr{ the 1st sentence does not belong here}\liron{I think it serves as a motivation for the section, but it fines by me to remove it}
-
-%In this section, we recall the basic SDN principles, with a focus on
-%OpenFlow~\cite{of-spec}, the standard SDN protocol today.
-\vspace{1mm}
-\noindent\textbf{Control and data planes.} Software-defined networking
-decouples the \emph{control plane}, software
-controllers that configure the network, from the \emph{data plane}, the
-devices that implement the network.
-%
-In OpenFlow, the \emph{de facto} standard SDN protocol
-today, the data plane consists of \emph{OpenFlow switches}.
-In a nutshell, the configuration of an OpenFlow switch is a set of
-\emph{rules}.
-A rule is essentially a \emph{match-action} pair:
-the match part of a rule identifies the space of packet headers that are
-subject to the rule (e.g., all packets to a specific destination) and
-the action part identifies how the switch should process the matched
-packets (e.g., forward them to a specific port).
-
-More precisely, the match part of a rule is
-a ternary pattern over packet \emph{header} fields.
-The pattern is represented as a \emph{value} and a \emph{mask}.
-%\liron{we need to discuss about this, I am not sure this description is accurate.}
-In the mask, certain header fields, e.g., TCP or UDP port numbers or destination and source
-addresses, can be \emph{wildcarded}, stipulating that their content does
-not affect the match.
-Certain fields, such as \emph{ipv4\_src}, \emph{ipv4\_dst}, \emph{metadata}, can be arbitrarily
-bitmasked.
-In our implementations, we are going to use bitmasking of the \emph{metadata}
-field for performing conditional operations based on  its content.
-%[[PK put the example back in response of review 2
-For example, \texttt{ipv4\_src=10.1.14.0, ff.ff.ff.f0} matches all packets
-with source IP addresses in the range \texttt{10.1.14.0}--\texttt{10.1.14.127}.
-%]]
-
-A \emph{configuration} of an OpenFlow switch is represented as a
-collection of \emph{flow tables}.
-A flow table is a set of \emph{flow entries}, each containing a rule,
-with a match, an action, and a \emph{priority} level, and flow entries
-are ordered according to their priorities.
-In the simplest setting, a switch maintains one flow table (table 0).
-A data plane packet arriving at a switch is first checked against
-the rule with the highest priority and in table $0$.
-If the header of the packet fits the match fields of that rule,
-a default \emph{instruction} associates the packet with the corresponding action.
-Otherwise, the packet is checked against flow  entries with lower
-priorities, and if no matching rule is found, the packet is dropped.
-
-%[[PK move pipelining to the discussion?
-\hide{
-Furthermore, an OpenFlow switch may maintain multiple flow tables
-forming a \emph{pipeline}.
-Data plane packets are checked against flow entries in the pipelined
-tables to determine the set of actions the switch should perform on the packet.
-A data plane packet arriving at a switch is first checked against
-the rule with the highest priority and in table $0$.
-If the header of the packet fits the match fields of that rule,
-a default \emph{instruction} associates the packet the corresponding action.
-Otherwise, the packet is checked against flow  entries with lower
-priorities.
-Specific instructions can be used to define
-additional tables to be visited (\textit{goto} instructions), to
-modify the set of actions associated with the packet (by adding,
-deleting or modifying actions), or immediately apply some actions in the action set.
-During the traversal of the flow tables, the packet can also be equipped with \emph{meta-tags} to
-transfer temporary information between tables.
-The meta-data field is part of the header that can be
-inserted or removed from a packet via \emph{push} and \emph{pop}
-actions.
-
-The actions associated with the packet are performed at the end of the pipeline.
-If no rule matches the packet, the packet is dropped.
-\petr{do we need the discussion of multiple tables/pipelining here? I still find it
-  very confusing}
-
-\liron{we should explain that: every packet is associated with an
-  action set. This set is empty at the beginning and is changed
-  according to the flow table matches. Each match is asociated with
-  instructions that can add or modify actions in the action set and
-  also to apply (now) all actions in the action set. By default action
-  set is applied at the end of the pipeline.}
-}
-
-%\liron{I think that maybe the rest of this section should be explained by the following subsections, providing in depth description of the OpenFlow features.}
-%\liron{Another direction is to have a simplified explenation as now and to refference spesific sections of the standard in side our paper.}
-\vspace{1mm}
-\noindent\textbf{Interaction between the planes: FlowMod commands.}
-The flow tables installed on the switches across the data plane
-determine the network \emph{policy}.
-Controllers can change the policy by sending
-control messages containing \emph{FlowMod commands}.
-In a nutshell, a \emph{FlowMod} command either specifies a new flow entry or
-a modification to an existing flow entry.
-%
-%
-%[[PK already said
-%A flow entry consists of: (1)~a match
-%field
-%(a value and a mask), (2)~a set of actions, and (3)~a priority.
-%]]
-%
-%
-A FlowMod command can either add a flow entry or delete a flow entry.
-The standard processing of a FlowMod $\add$ command received by a switch is
-as follows.
-If the switch already maintains a flow entry with \emph{exactly} the
-same match and priority level, then the new flow entry will simply \emph{replace} it.
-Otherwise, the flow entry will be installed in addition to existing
-ones.
-A $\dele$ command simply removes an existing flow entry with the same
-match and priority,  or does nothing if such an entry is not present.
-%Stefan: alternatively could also delete an entry by using the cookie field
-
-%\petr{Is this correct?}
-
-%but will be installed
-%\emph{in addition} to existing entries if they only
-%partially overlap (or define flowspace subsets or supersets).
-
-%If rules with overlapping but not identical match fields have the same priority,
-%inconsistencies can arise, since there is no clearly specified action
-%that should be applied to a packet matching the two entries.
-In order to avoid inconsistencies caused by different rules with
-overlapping match fields, a \emph{FlowMod} command can be equipped with a the $\checko$ flag:
-if the switch maintains a flow entry with an overlapping but not
-identical match part with the same priority but a different action,
- and if the two rules have the same priority, then
-\emph{FlowMod} will fail.
-The $\checko$ flag in a \emph{FlowMod} command helps resolving simple
-conflicts between controllers, and will be instrumental
-in implementing our synchronization abstractions.
-
-%[[PK not sure this is the right place
-%implement synchronization objects which form
-%the basis of our transactional interface.
-%]]
-We will use the following notation to create a \emph{FlowMod} message:
-$\FlowMod(\match, op, \action, \flags)$, where
-%$\match$ denotes the match,
-$op$ is $\add$ or $\dele$, and $\flags$ are used, e.g.,  to activate the $\checko$
-feature.
-For simplicity, we omit other standard parameters, assuming them to carry
-default values.
-%A standard \emph{FlowMod} message (and also the flow entries) may specify
-%more parameters, which are however irrelevant for our implementation, and we suggest to
-%use the default values.
-
-To read the current configuration of a switch (the existing flow entries), the controller should send
-the OpenFlow \emph{add flow monitor} command with the \textsf{ofpt\_initial} optional flag.
-In this paper, we write \textbf{read-config} for sending
-this message, receiving a response, and returning the received configuration.
-
-\hide{
-%[[PK should go to the discussion?
-We also remark that our solution is only based on flow tables
-and does not make use
-of \emph{OpenFlow group tables}, nor do we make use of \emph{cookies}: rules
-are always identified by exact matches.
-%]]
-}
-
-\vspace{1mm}
-\noindent\textbf{Bundling.}
-%
-Another important feature in OpenFlow is \emph{bundling}. %(as of OpenFlow version 1.4). \stefan{liron, is that correct?}
-A controller can send multiple $\FlowMod$ commands  equipped with
-the same \emph{bundle identifier}.
-These commands are buffered temporarily by
-the switch until a \emph{bundle commit} message is received from the
-controller.
-The bundled commands are then performed with \emph{all-or-nothing semantics}:
-either all of them are performed, or none of them is.
-In particular, if a configuration request contained in one of the bundled commands cannot
-be applied (e.g., the command is rejected because of the set
-$\checko$ flag and a conflicting flow entry), all  other commands in
-the bundle are rejected and an error
-message is sent to the controller that issued them.
-%If multiple \emph{FlowMod} messages with the $\checko$ flag are sent as a single bundle,
-The error message contains $\xid$, the identifier of the first failed command in the
-bundle, and $\ecode$, the \emph{error code} corresponding to the failure.
-
-A bundle begins with a  \texttt{ofpt\_bundle\_control} message of type
-\texttt{ofpbct\_open\_request} (creating the bundle), wraps
-each of its \emph{FlowMod} command in a message
-\texttt{ofpt\_bundle\_add\_message} equipped with the \emph{bundle
-  identifier}, and ends with a \texttt{ofpt\_bundle\_control} message of
-type \texttt{ofpbct\_commit\_request} (committing the bundle).
-
-%[[PK too detailed and not very clear
-%Concretely, when the commit fails, the switch must generate an error message corresponding to the
-%message that could not be applied. The error message must use the identifier (\emph{xid}) of the offending message,
-%the error data field corresponding to that message and the error code corresponding to the failure. If
-%multiple messages are generating errors, the switch may return only the first error found or generate
-%multiple error messages for the same bundle.
-%]]
-
-%Again, the bundle always returns a binary result: \emph{error} ($\bot$) or \emph{success} ($\top$).
-
-%When properly configured,
-The bundle execution is considered \emph{controller-atomic}
-in the sense that all controllers can only observe switch
-configurations \emph{before} or \emph{after} the
-bundle commands are executed.
-We set the flag \texttt{ofpbf\_atomic},
-making bundles \emph{packet-atomic},
-in the sense that every data plane packet is processed by a configuration
-before or after the bundle, and not by a configuration resulting from
-an incomplete bundle execution.
-We also set the flag \texttt{ofpbf\_ordered},  to make sure that the
-bundle commands are executed in the order they were added to the
-bundle, thereby respecting dependencies between bundle commands.
-%\liron{i added controller atomicity and order preservation, see that it make sense...}
-
-%:a $\checko$ flag can be defined for each command in the bundle.\liron{these last two sentences are not cleaar}
-
-%\liron{the notion of transaction presented next is confusing considering the bundle just explained}
-%Concretely, a transaction can represent between two to five switch interactions:
-%if limited space, Algo 7: read, claim id, read again, transaction, unclaim;
-%if if unlimited space: alg 6 is unbounded, increase policy ID by one.
-%transactions = five switch interactions , claim \pid (line 5+6) together in bundle => one iteraction less?
-
-%In general, it is desirable to minimize the number of transactions \stefan{stefan should cite \cite{Jin2014Dionysus} here}
-
-\hide{
-To commit the transaction,  the controller sends a \texttt{OFPT\_BUNDLE\_CONTROL} message with type
-\texttt{OFPBCT\_COMMIT\_REQUEST} and waits until a matching control
-message from the switch is received. When the switch receives the
-commit request it tries to atomically process all the buffered messages
-equipped with the corresponding bundle identifier and, if the
-processing is successful for all of them, returns  a \texttt{OFPT\_BUNDLE\_CONTROL} message with type
-\texttt{OFPBCT\_CLOSE\_REPLY}; otherwise it sends a specific error
-message  pointing out the aborted transactional operation.
-
-Algorithm~\ref{alg:commit} describes how to commit
-the bundle.
-\begin{algorithm}[t]
-    \caption{$\textit{try-commit}()$}
-    \label{alg:commit}
-    \begin{algorithmic}[1]
-    \Require  current bundle id $bid$.
-		    \State $cmd \gets struct ofp_bundle_ctrl_msg$
-    		\State  $cmd.bundle\_id \gets bid$
-    		\State  $cmd.type \gets $ \texttt{ofpbct\_commit\_request}
-    		\State  $cmd.flags \gets $ \texttt{ofpbf\_atomic}
-    		\State \textbf{send} $cmd$
-		\Return $cmd$
-    \end{algorithmic}
-\end{algorithm}
-}
-
-%\textbf{Cookies:} Cookies allow to name and retrieve individual
-%flow entries and to use masks to delete modify/delete multiple entries.
-%Our synchronization mechanism will implement controller-specific cookies.
-%
-%\item \textbf{Groups:} Groups of action buckets can be defined in a special group table.
-Each group has a unique id and any attempt to create a group with existing id or %to delete un existing group fails (and can abort a bundle in process).
-%\end{enumerate}
-
-%\begin{figure*}[t]
-%\centering
-%\includegraphics[width=0.35\columnwidth]{loadbal-before.pdf}~\includegraphics[width=0.35\columnwidth]{loadbal.pdf}~~
-%\includegraphics[width=0.8\columnwidth]{pic.pdf}\\
-%\caption{\emph{Left:} Load-balancing gone wrong. \emph{Right:} Illustration of transactional abstraction which  allows
-%controllers to modify switch configurations in complex ways,
-%and
-%depending on other flow spaces. Colors indicate different actions. \stefan{lets make two real transactions: one bundle and one 2 or 5 interactions?}}\label{fig:illu}
-%\end{figure*}
-
-\section{Limitations and Motivation}\label{sec:motivation}
+\section{Motivation and Example}\label{sec:motivation}
 
@@
 
-At first sight, bundling \emph{FlowMod} commands with  the $\checko$
-flag already provides a useful synchronization mechanism.
-Indeed, its ``mini-transaction'' nature allows a controller
-to install multiple flow entries in an atomic way.
-%[[PK we did not say what reads and writes are...
-%At first sight, the OpenFlow bundle feature seems to provide everything
-%we need to implement our transactional interface \petr{did we say what we need already?}: the bundle
-% defines a sequence of read and write operations
-% on the switch state which is executed atomically.
-%]]
-However, as is,
-%while the bundle can indeed be seen as a ``mini-transaction'',
-bundling has important limitations, as we will argue in the following.
-%\petr{unclear what "reads and writes" are here, not discussed before. How openflow can be used to implement reads and writes?}
-
-First, %if employed without any additional features,
-except for some limited constraint types (such as
-overlapping flow spaces or insufficient free space),
-a bundle does not provide any means to modify the switch configuration
-based on application-specific conditions, which depend on the
-current configuration,
-However, as we will argue, support for more generic conditions under which configuration
-updates
-should or should not take place, is often desirable.
+A transactional distributed system offers all-or-nothing semantics, 
+and aborts transactions in case of conflicts. 
+For instance, in the context of SDNs, a transaction trying to add a forwarding rule whose
+domain overlaps with an existing rule of the same priority
+can constitute a conflict. 
+However, as we will argue, there are often conflicts which are less obvious,
+and go beyond simple range overlaps. 
+Moreover, sometimes it is desirable to react to conflicts in smarter ways than
+by simply aborting an operation, e.g., by supporting \emph{conditional modifications}.
+Let us consider two examples.
+
 %
@@
 %
-Second, sometimes it is desirable to react to conflicts in smarter ways than
-by simply aborting an operation, e.g., by supporting \emph{conditional modifications}.
-Let us consider two examples.
+
+%Let us consider an example.
 
@@
 \vspace{1mm}
-\noindent \textbf{Example~1: Handling complex dependencies.} Let us consider two controllers
+\noindent \textbf{Example: Handling complex dependencies.} We consider two controllers
 in charge
@@
 Accordingly, if multiple controllers concurrently and independently
-update forwarding rules according to a na\"ive load-balancing algorithm \`{a} la Algorithm~\ref{alg:naive-lb},
+update forwarding rules according to a na\"ive load-balancing algorithm,
 they may involuntarily
@@
 such conflicts.
-Note that, assuming that all these flows are
-completely independent, bundling and $\checko$ features do not really
-help here.
 
@@
 
-{\small
-\begin{algorithm}[t]
-    \caption{$\textit{Na\"ive Flow Balancing}$}
-    \label{alg:naive-lb}
-\begin{algorithmic}[1]
-
-    \Require \emph{new\_flows}: flow entries of the new flows,
-    \emph{ports}: a list of switch output ports,
-    \State $\textit{policy} \gets \textbf{read\_config}$
-    \State $\textit{update\_cmds} \gets \textsc{lb\_update}(policy)$ %(new\_flows, ports, policy)$
-    \State $\textbf{send}\{\textit{update\_cmds}\}$
-%    \State $\textbf{execute-atomic}\{update\_cmds\}$
-    \Statex
-
-    \Statex \emph{\bf Function} $\textsc{lb\_update}(\textit{policy})$ % (new\_flows, ports, policy)}$
-
-	\State $\textit{flows\_per\_port} \gets [0, 0, \ldots, 0]$
-	\ForAll {$\textit{entry} \in \textit{policy}$}
-	    \State $\textit{port} \gets \textit{entry.output\_port}$
-	    \State $\textit{flows\_per\_port}[\textit{port}]++$
-	\EndFor
-	\State $\textit{new\_flow\_mod\_cmds} \gets \{\}$
-	\For {$\textit{entry} \in \textit{new\_flows}$}
-	    \State $\textit{port} \gets \arg\min_{p\in \textit{ports}} \textit{flows\_per\_port}[p]$
-	    \State $\textit{entry.output\_port} \gets \textit{port}$
-	    \State $\textit{flows\_per\_port}[\textit{port}]++$
-	    \State $\textit{cmd} \gets \textit{FlowMod}(add,f)$
-	    \State $\textit{new\_flow\_mod\_cmds}.\textit{add}(\textit{cmd})$
-	\EndFor
-	\Return $\textit{new\_flow\_mod\_cmds}$
-  %  \State \emph{\bf EndFunction}
-\end{algorithmic}
-\end{algorithm}
-}
-
 %\petr{postpone the explanation of the right part of the figure until we come to CAS?}
 
-\vspace{1mm}
-\noindent\textbf{Example~2: Supporting modifications.}
-Consider the arguably very basic problem of \emph{modifying} previously installed forwarding rules
-in a consistent manner.
-In principle, a bundle allows us to conditionally add new rules or remove
-old rules, and, using the flow space overlap
-check feature $\checko$, to abort the transaction if potentially conflicting
-rules were already there. However, it is not obvious how
-to use the bundle to
-modify \emph{existing} rules, especially if the modification depends on
-specific conditions.
-In other words, the bundle and overlap check only help
-for the \emph{conditional creation} of rules, not their \emph{modification}.
-
-\vspace{2mm}
-\noindent\textbf{Our Approach.}
-This paper proposes mechanisms to solve these two examples,
+This paper proposes mechanisms to solve these synchronization problems 
 by using a minimal
@@
 by multiple controllers despite of concurrency and conflicts.
-
-We will later come back to these examples, and show that Algorithm~\ref{alg:naive-lb}
-can actually be used without any modifications within our framework.
-Consistent load-balancing can also be achieved using our atomic Compare-and-Set instruction based on policy ids (pids),
-see Figure~\ref{fig:CAS-example} (\emph{right}).
 
@@
 
-\section{Synchronization Abstractions}\label{sec:main}
-
-We now describe how our  meta-configuration space at
-an OpenFlow switch can be implemented and manipulated using our
+\section{In-Band Synchronization}\label{sec:main}
+
+In order to synchronize controllers, we propose
+to use a meta-configuration space at
+the switches which can be manipulated using our
 synchronization framework.
@@
 designed in a way that it does not affect the processing of data plane
-traffic. One way to achieve this is to set a low priority to
+traffic. One way to achieve this, for example, is to set a low priority to
 the meta flow entries and ensure that the regular flow entries
-maintain higher priority levels.
-Alternatively, we could leverage the possibility to maintain multiple
-flow tables at an OpenFlow switch, and use a separate flow table for
-the meta-configuration,  making sure that the
-\emph{pipelining} mechanism will never involve this
-flow table.
+maintain higher priority levels. We will discuss other options later
+in the context of OpenFlow.
 
@@
  %
-Our framework provides the controllers with the ability to execute
+We provide the controllers with the ability to execute
 sequences of \emph{operations} on the
@@
 In the non-atomic mode, no atomicity guarantees are provided.
-An operation here can be a regular FlowMod command or a synchronization
-primitive, which, as we
-show below, translates into a sequence of FlowMod commands operating on the
-switch meta-configuration.
 
@@
 \end{comment}
+
+\section{Case Study: OpenFlow}\label{sec:background}
+%\petr{ the 1st sentence does not belong here}\liron{I think it serves as a motivation for the section, but it fines by me to remove it}
+
+%In this section, we recall the basic SDN principles, with a focus on
+%OpenFlow~\cite{of-spec}, the standard SDN protocol today.
+\vspace{1mm}
+\noindent\textbf{Control and data planes.} Software-defined networking
+decouples the \emph{control plane}, software
+controllers that configure the network, from the \emph{data plane}, the
+devices that implement the network.
+%
+In OpenFlow, the \emph{de facto} standard SDN protocol
+today, the data plane consists of \emph{OpenFlow switches}.
+In a nutshell, the configuration of an OpenFlow switch is a set of
+\emph{rules}.
+A rule is essentially a \emph{match-action} pair:
+the match part of a rule identifies the space of packet headers that are
+subject to the rule (e.g., all packets to a specific destination) and
+the action part identifies how the switch should process the matched
+packets (e.g., forward them to a specific port).
+
+More precisely, the match part of a rule is
+a ternary pattern over packet \emph{header} fields.
+The pattern is represented as a \emph{value} and a \emph{mask}.
+%\liron{we need to discuss about this, I am not sure this description is accurate.}
+In the mask, certain header fields, e.g., TCP or UDP port numbers or destination and source
+addresses, can be \emph{wildcarded}, stipulating that their content does
+not affect the match.
+Certain fields, such as \emph{ipv4\_src}, \emph{ipv4\_dst}, \emph{metadata}, can be arbitrarily
+bitmasked.
+In our implementations, we are going to use bitmasking of the \emph{metadata}
+field for performing conditional operations based on  its content.
+%[[PK put the example back in response of review 2
+For example, \texttt{ipv4\_src=10.1.14.0, ff.ff.ff.f0} matches all packets
+with source IP addresses in the range \texttt{10.1.14.0}--\texttt{10.1.14.127}.
+%]]
+
+A \emph{configuration} of an OpenFlow switch is represented as a
+collection of \emph{flow tables}.
+A flow table is a set of \emph{flow entries}, each containing a rule,
+with a match, an action, and a \emph{priority} level, and flow entries
+are ordered according to their priorities.
+In the simplest setting, a switch maintains one flow table (table 0).
+A data plane packet arriving at a switch is first checked against
+the rule with the highest priority and in table $0$.
+If the header of the packet fits the match fields of that rule,
+a default \emph{instruction} associates the packet with the corresponding action.
+Otherwise, the packet is checked against flow  entries with lower
+priorities, and if no matching rule is found, the packet is dropped.
+
+%[[PK move pipelining to the discussion?
+\hide{
+Furthermore, an OpenFlow switch may maintain multiple flow tables
+forming a \emph{pipeline}.
+Data plane packets are checked against flow entries in the pipelined
+tables to determine the set of actions the switch should perform on the packet.
+A data plane packet arriving at a switch is first checked against
+the rule with the highest priority and in table $0$.
+If the header of the packet fits the match fields of that rule,
+a default \emph{instruction} associates the packet the corresponding action.
+Otherwise, the packet is checked against flow  entries with lower
+priorities.
+Specific instructions can be used to define
+additional tables to be visited (\textit{goto} instructions), to
+modify the set of actions associated with the packet (by adding,
+deleting or modifying actions), or immediately apply some actions in the action set.
+During the traversal of the flow tables, the packet can also be equipped with \emph{meta-tags} to
+transfer temporary information between tables.
+The meta-data field is part of the header that can be
+inserted or removed from a packet via \emph{push} and \emph{pop}
+actions.
+
+The actions associated with the packet are performed at the end of the pipeline.
+If no rule matches the packet, the packet is dropped.
+\petr{do we need the discussion of multiple tables/pipelining here? I still find it
+  very confusing}
+
+\liron{we should explain that: every packet is associated with an
+  action set. This set is empty at the beginning and is changed
+  according to the flow table matches. Each match is asociated with
+  instructions that can add or modify actions in the action set and
+  also to apply (now) all actions in the action set. By default action
+  set is applied at the end of the pipeline.}
+}
+
+%\liron{I think that maybe the rest of this section should be explained by the following subsections, providing in depth description of the OpenFlow features.}
+%\liron{Another direction is to have a simplified explenation as now and to refference spesific sections of the standard in side our paper.}
+\vspace{1mm}
+\noindent\textbf{Interaction between the planes: FlowMod commands.}
+The flow tables installed on the switches across the data plane
+determine the network \emph{policy}.
+Controllers can change the policy by sending
+control messages containing \emph{FlowMod commands}.
+In a nutshell, a \emph{FlowMod} command either specifies a new flow entry or
+a modification to an existing flow entry.
+%
+%
+%[[PK already said
+%A flow entry consists of: (1)~a match
+%field
+%(a value and a mask), (2)~a set of actions, and (3)~a priority.
+%]]
+%
+%
+A FlowMod command can either add a flow entry or delete a flow entry.
+The standard processing of a FlowMod $\add$ command received by a switch is
+as follows.
+If the switch already maintains a flow entry with \emph{exactly} the
+same match and priority level, then the new flow entry will simply \emph{replace} it.
+Otherwise, the flow entry will be installed in addition to existing
+ones.
+A $\dele$ command simply removes an existing flow entry with the same
+match and priority,  or does nothing if such an entry is not present.
+%Stefan: alternatively could also delete an entry by using the cookie field
+
+%\petr{Is this correct?}
+
+%but will be installed
+%\emph{in addition} to existing entries if they only
+%partially overlap (or define flowspace subsets or supersets).
+
+%If rules with overlapping but not identical match fields have the same priority,
+%inconsistencies can arise, since there is no clearly specified action
+%that should be applied to a packet matching the two entries.
+In order to avoid inconsistencies caused by different rules with
+overlapping match fields, a \emph{FlowMod} command can be equipped with a the $\checko$ flag:
+if the switch maintains a flow entry with an overlapping but not
+identical match part with the same priority but a different action,
+ and if the two rules have the same priority, then
+\emph{FlowMod} will fail.
+The $\checko$ flag in a \emph{FlowMod} command helps resolving simple
+conflicts between controllers, and will be instrumental
+in implementing our synchronization abstractions.
+
+%[[PK not sure this is the right place
+%implement synchronization objects which form
+%the basis of our transactional interface.
+%]]
+We will use the following notation to create a \emph{FlowMod} message:
+$\FlowMod(\match, op, \action, \flags)$, where
+%$\match$ denotes the match,
+$op$ is $\add$ or $\dele$, and $\flags$ are used, e.g.,  to activate the $\checko$
+feature.
+For simplicity, we omit other standard parameters, assuming them to carry
+default values.
+%A standard \emph{FlowMod} message (and also the flow entries) may specify
+%more parameters, which are however irrelevant for our implementation, and we suggest to
+%use the default values.
+
+To read the current configuration of a switch (the existing flow entries), the controller should send
+the OpenFlow \emph{add flow monitor} command with the \textsf{ofpt\_initial} optional flag.
+In this paper, we write \textbf{read-config} for sending
+this message, receiving a response, and returning the received configuration.
+
+\hide{
+%[[PK should go to the discussion?
+We also remark that our solution is only based on flow tables
+and does not make use
+of \emph{OpenFlow group tables}, nor do we make use of \emph{cookies}: rules
+are always identified by exact matches.
+%]]
+}
+
+\vspace{1mm}
+\noindent\textbf{Bundling.}
+%
+Another important feature in OpenFlow is \emph{bundling}. %(as of OpenFlow version 1.4). \stefan{liron, is that correct?}
+A controller can send multiple $\FlowMod$ commands  equipped with
+the same \emph{bundle identifier}.
+These commands are buffered temporarily by
+the switch until a \emph{bundle commit} message is received from the
+controller.
+The bundled commands are then performed with \emph{all-or-nothing semantics}:
+either all of them are performed, or none of them is.
+In particular, if a configuration request contained in one of the bundled commands cannot
+be applied (e.g., the command is rejected because of the set
+$\checko$ flag and a conflicting flow entry), all  other commands in
+the bundle are rejected and an error
+message is sent to the controller that issued them.
+%If multiple \emph{FlowMod} messages with the $\checko$ flag are sent as a single bundle,
+The error message contains $\xid$, the identifier of the first failed command in the
+bundle, and $\ecode$, the \emph{error code} corresponding to the failure.
+
+A bundle begins with a  \texttt{ofpt\_bundle\_control} message of type
+\texttt{ofpbct\_open\_request} (creating the bundle), wraps
+each of its \emph{FlowMod} command in a message
+\texttt{ofpt\_bundle\_add\_message} equipped with the \emph{bundle
+  identifier}, and ends with a \texttt{ofpt\_bundle\_control} message of
+type \texttt{ofpbct\_commit\_request} (committing the bundle).
+
+%[[PK too detailed and not very clear
+%Concretely, when the commit fails, the switch must generate an error message corresponding to the
+%message that could not be applied. The error message must use the identifier (\emph{xid}) of the offending message,
+%the error data field corresponding to that message and the error code corresponding to the failure. If
+%multiple messages are generating errors, the switch may return only the first error found or generate
+%multiple error messages for the same bundle.
+%]]
+
+%Again, the bundle always returns a binary result: \emph{error} ($\bot$) or \emph{success} ($\top$).
+
+%When properly configured,
+The bundle execution is considered \emph{controller-atomic}
+in the sense that all controllers can only observe switch
+configurations \emph{before} or \emph{after} the
+bundle commands are executed.
+We set the flag \texttt{ofpbf\_atomic},
+making bundles \emph{packet-atomic},
+in the sense that every data plane packet is processed by a configuration
+before or after the bundle, and not by a configuration resulting from
+an incomplete bundle execution.
+We also set the flag \texttt{ofpbf\_ordered},  to make sure that the
+bundle commands are executed in the order they were added to the
+bundle, thereby respecting dependencies between bundle commands.
+%\liron{i added controller atomicity and order preservation, see that it make sense...}
+
+%:a $\checko$ flag can be defined for each command in the bundle.\liron{these last two sentences are not cleaar}
+
+%\liron{the notion of transaction presented next is confusing considering the bundle just explained}
+%Concretely, a transaction can represent between two to five switch interactions:
+%if limited space, Algo 7: read, claim id, read again, transaction, unclaim;
+%if if unlimited space: alg 6 is unbounded, increase policy ID by one.
+%transactions = five switch interactions , claim \pid (line 5+6) together in bundle => one iteraction less?
+
+%In general, it is desirable to minimize the number of transactions \stefan{stefan should cite \cite{Jin2014Dionysus} here}
+
+\hide{
+To commit the transaction,  the controller sends a \texttt{OFPT\_BUNDLE\_CONTROL} message with type
+\texttt{OFPBCT\_COMMIT\_REQUEST} and waits until a matching control
+message from the switch is received. When the switch receives the
+commit request it tries to atomically process all the buffered messages
+equipped with the corresponding bundle identifier and, if the
+processing is successful for all of them, returns  a \texttt{OFPT\_BUNDLE\_CONTROL} message with type
+\texttt{OFPBCT\_CLOSE\_REPLY}; otherwise it sends a specific error
+message  pointing out the aborted transactional operation.
+
+Algorithm~\ref{alg:commit} describes how to commit
+the bundle.
+\begin{algorithm}[t]
+    \caption{$\textit{try-commit}()$}
+    \label{alg:commit}
+    \begin{algorithmic}[1]
+    \Require  current bundle id $bid$.
+		    \State $cmd \gets struct ofp_bundle_ctrl_msg$
+    		\State  $cmd.bundle\_id \gets bid$
+    		\State  $cmd.type \gets $ \texttt{ofpbct\_commit\_request}
+    		\State  $cmd.flags \gets $ \texttt{ofpbf\_atomic}
+    		\State \textbf{send} $cmd$
+		\Return $cmd$
+    \end{algorithmic}
+\end{algorithm}
+}
+
+%\textbf{Cookies:} Cookies allow to name and retrieve individual
+%flow entries and to use masks to delete modify/delete multiple entries.
+%Our synchronization mechanism will implement controller-specific cookies.
+%
+%\item \textbf{Groups:} Groups of action buckets can be defined in a special group table.
+Each group has a unique id and any attempt to create a group with existing id or %to delete un existing group fails (and can abort a bundle in process).
+%\end{enumerate}
+
+%\begin{figure*}[t]
+%\centering
+%\includegraphics[width=0.35\columnwidth]{loadbal-before.pdf}~\includegraphics[width=0.35\columnwidth]{loadbal.pdf}~~
+%\includegraphics[width=0.8\columnwidth]{pic.pdf}\\
+%\caption{\emph{Left:} Load-balancing gone wrong. \emph{Right:} Illustration of transactional abstraction which  allows
+%controllers to modify switch configurations in complex ways,
+%and
+%depending on other flow spaces. Colors indicate different actions. \stefan{lets make two real transactions: one bundle and one 2 or 5 interactions?}}\label{fig:illu}
+%\end{figure*}
 
@@
 This paper initiated the study of mechanisms which allow distributed controllers
-to synchronize their possibly concurrent actions. We believe that our in-band
-approach is natural and attractive as it can avoid the
-complexities and limitations of distributed out-of-band agreement protocols.
+to synchronize their possibly concurrent actions. We believe that our centralized in-band
+approach is natural and attractive, as it avoids the
+complexities and limitations of distributed out-of-band agreement protocols, which may also
+introduce additional points of potential failures.
 Moreover, our mechanisms are simple and can be implemented using the \emph{standard OpenFlow}
@@
 language abstractions.~\cite{pyretic}
-We also plan to release a small library with our synchronization primitives. 
-Moreover, we are investigating possible generalizations of our transactional guarantees,
-across multiple switches~\cite{stn}, as well as opportunities for more fine-grained
-\emph{partial} updates that concern subsets of flow entries, thus improving concurrency.
+In general, we understand our paper as a first step, and more
+research is required to investigate and compare useful
+inband synchronization abstractions, also in terms of performance.
+%In the long run, we seek to develop in-band mechanisms
+%which allow us to seemlessly transition between single controller and
+%multiple controller mode.
+
+%We also plan to release a small library with our synchronization primitives. 
+%Moreover, we are investigating possible generalizations of our transactional guarantees,
+%across multiple switches~\cite{stn}, as well as opportunities for more fine-grained
+%\emph{partial} updates that concern subsets of flow entries, thus improving concurrency.
 

### 2015-09-01 ccr-rev1/inband-sync.tex
revised version

@@
 
-%\if \SAVESPACE 1
-%\setlength{\floatsep}{3pt}
-%\setlength{\textfloatsep}{3pt}
-%\setlength{\dbltextfloatsep}{3pt}
-%\setlength{\intextsep}{3pt}
-%\setlength{\abovecaptionskip}{3pt}
-%\fi
-
-% A category with the (minimum) three required fields
-%\category{C.2.1}{Network Architecture and Design}{Centralized Networks}
-%\category{C.2.4}{Distributed Systems}{Network Operating Systems}
-%\terms{Measurement, Performance}
-%\keywords{}
-
 \begin{abstract}
@@
 should be physically distributed.
-%a redundant control plane
-%can ensure a high availability, and a spatially distributed
-%control plane allows to handle time-critical dataplane
-%events close to their origins.
-%[[PK this is a contraversial point, not important for us
-%Moreover, an SDN may be managed by multiple administrators simultaneously.
-%]]
 However, %a distributed control plane also introduces
@@
  nontrivial task.
-%[[PK sounds contradicting to the fact that such systems already exist
-%However, not much is known today about how to implement
-%such a distributed control plane.
-%]]
-%
 In this paper, we propose a synchronization
@@
 \emph{transactions}.
-%
-%A transaction aggregates a sequence
-%of control messages and specific synchronization primitives,
-%providing all-or-nothing semantics.
 In particular, we present a powerful interface 
@@
  consistent policy composition and fault-tolerant control-planes.
-%
-%Indeed, recent OpenFlow versions already come with useful but primitive
-%features, in particular \emph{bundling} and \emph{rule-overlap checks}.
-%In particular, we present a
-%transactional interface which supports arbitrary atomic
-%transactions
-%over the switch configuration space:
-%Our transactions
-%can include complex notions of conflicts (beyond flow overlap checks),
-%as well as computations and multiple controller interactions.
-%
 Interestingly, 
@@
 
-%\vspace{1cm}
-
-%\begin{center}
-%{\bf [Regular paper only]}
-%\end{center}
-%[[PK I guess not anymore?]]
-
-%\vspace{1cm}
-
-%\begin{center}
-%{\emph{Contact Address:}\\Marco Canini, Place Sainte Barbe~2, 1348 Louvain-la-Neuve, Belgium\\Tel: $+$32 10 47 48 32,
-%marco.canini@uclouvain.be}
-% {\emph{Contact Address:}\\Stefan Schmid, MAR 4-4, Marchstr.~23, 10587 Berlin, Germany\\Tel: $+$49 175 930 98 75,
-% stefan.schmid@tu-berlin.de}
-%\end{center}
-
 \end{comment}
-
-%\newpage
-
-%\begin{center}
-%{\bf Regular and student paper: Dan Levin is a full-time student.}
-%\end{center}
-
-%
-%
-%\section*{todo}
-%
-%stefan add old notes, bundle size complexity (how to make it minimal: store rule in one table, select it with a pointer / selector),
-%practical implementation (not even noviswitch has bundles)
-%
-%we add all rules of the policy also!
-but we could add rules before and only activate later => set selector later (maybe set timeout); transaction size constant => keep %bundle small important: packet atomicity so maybe on hold during bundle
-%
-%
 
@@
 (hardware) data plane.
-%Moreover, OpenFlow, the standard SDN protocol today, introduces many generalizations,
-%in terms of traffic engineering, definition of flows, as well as in-band network functionalities,
-%by relying on a simple match-action paradigm which allows us to define
-%forwarding rules based not only on Layer-2, but also Layer-3 and Layer-4 (and possibly beyond).
-%
 While the perspective of a \emph{logically centralized} control plane
@@
 First, in order to provide high availability,
-%as well as a minimal degree of
-%fault-tolerance,
-%[[PK fault-tolerance is necessary to provide hight availability?]]
 controllers should be redundant~\cite{onix,onos,elasticon}: a failure
@@
 concurrently~\cite{participatory}.
-
-%[[PK not sure I see why talking about it here
-%[[
-%A particularly interesting problem regards the consistent installation of new policies
-%or routes~\cite{network-update,roger-hotnets,correct,stn}. While implementing consistent network updates
-%is challenging already from the perspective of a single controller, there exists a wide consensus
-%that the logically centralized SDN control planes will likely be \emph{actually distributed} in the future.
-%]]
 
@@
 all-or-nothing semantics.
-
-%[[PK
-%and techniques to solve synchronization and consistency problems.
-%A particularly interesting concept in distributed computing is a \emph{transactional memory}:
-%a shared memory which supports atomic transactions, describing a sequence of operations
-%to be executed with an all-or-nothing semantics. Transactional memories can be used to
-% solve many
-%classic distributed computing problems, such as consensus.
-%]]
 
@@
 
-%
-%By providing more complex and atomic transactions over the
-%switch configuration
-%space, several synchronization problems can be resolved.
-%This paper investigates the implementation of a transactional interface
-%(\`{a} la STN~\cite{stn}) modified concurrently by multiple controllers.
-%\petr{"interface modified concurrently" does not sound right -> "accessed concurrently"?}
-%
-
-%which can be used even for coordinating
-%configuration updates operating
-%over \emph{independent flow spaces}, but where dependencies are introduced
-%in the mapping of the flows to the physical infrastructure. This is
-%important, e.g., for load-balancing applications.
-
-%[[PK
-%This paper shows how to implement a powerful transactional interface
-%for software-defined networks.
-%This transactional interface is reminiscent of
-%classic transactional memory abstractions~\cite{stm-st95,tm-book}
-%and
-%can be used to resolve
-%synchronization problems of distributed control planes efficiently.
-%]]
-
 As a case study, and to demostrate the feasibility
@@
 
-%Finally, we also show that interesting synchronization mechanisms may be included
-%in-band even without the bundle feature, and as a case study, show
-%how to automatically \emph{compose} updates originating \emph{from different controllers}.
-
-% we show that powerful synchronization primitives
-%can be implemented in OpenFlow, allowing controllers to implement
-%non-trivial atomic transactions, and detect and
-%automatically resolve conflicts. In particular, we show how to implement
-%the classic distributed computing primitive
-%\emph{test\&set}, supporting the conditional installation of new rules,
-%and we show how to use the bundle feature to implement
-%multi-writer objects, an efficient means to compute consensus;
-%interestingly, our implementation can also be used to compute a consensus
-%and install the winner's rule \emph{atomically}, using one message.
-%Moreover, our approach allows to detect and resolve very general notions of conflicts:
-%conflicts may be defined, for example, for rules of the same priority over
-%defined over non-independent flow spaces, but may also depend, e.g., on
-%network load (based on counters).
-%We believe that our approach nicely complements
-%existing, more theoretical literature on distributed control planes; for example,
-%it also shows how to realize the atomic read-modify-write primitive postulated in
-%STN~\cite{stn}.
-
 \noindent\textbf{Organization}
@@
 discuss use cases in Section~\ref{sec:apps}.
-%Section~\ref{sec:compo} describes how to perform in-band composition.
 We
-%review related literature in Section~\ref{sec:relwork}, and
 conclude our paper
 in Section~\ref{sec:conclusion}.
-% Some technical details are postponed
-%to the Appendix.
-
-%we provide the theoretical
-%background on our approach. Section~\ref{sec:realization}
-%then discusses the basic implementation of this approach in OpenFlow.
-%Section~\ref{sec:extensions} discusses practical issues and extensions.
-%After reviewing related literature in Section~\ref{sec:relwork},
-%we conclude our work in
-%Section~\ref{sec:conclusion}.
 
 \section{Motivation and Example}\label{sec:motivation}
-
-% generic way for multiple distributed controllers to safely update switch configuration with no need of any control plane
 
@@
 
-%
-%But also bundling FlowMod commands with the check overlap feature ($\checko$)
-%comes with limitations: while it is possible
-% to \emph{abort} a bundle depending on certain conditions
-% in the \emph{overlapping
-% flow spaces}, oftentimes it is desirable to be able
-% to define more general conditions and notions of conflicts
-% under which an atomic update should not take effect.
-% \liron{replaced transaction in the last sentence with " atomic update"}
-%
-
-%Let us consider an example.
-
-%\begin{enumerate}
 \vspace{1mm}
@@
 TCP micro-flows),
-%originated in disjoint ranges of IP addresses),
 may actually be dependent: the flows share the underlying physical network.
@@
 such conflicts.
-
-%Consider the example in the left part of Figure~\ref{fig:CAS-example}. Two controllers,
-%$1$ and $2$ read the configuration, try to add two more flows to a switch
-%with two outgoing links, balancing the load.
-%Imagine that each of the controllers reads the switch configuration and
-%learns that currently one flow is forwarded to the left link and two
-%flows are forwarded to the right link.
-%Without synchronization, the two controllers naturally choose to
-%install their flows on the left link, which results in an undesirable
-%unbalanced state.
-
-%\petr{postpone the explanation of the right part of the figure until we come to CAS?}
 
@@
 
-%[[PK not really discussed?
-%As we will see, we do not even need the bundle feature
-%to implement the powerful synchronization objects of our transactional interface.
-%However, the bundle feature can still be useful in the context of
-%consistent policy updates.
-
-%At first sight, a bundle seems to be a powerful concept
-%for the implementation of a transactional interface. However,
-%while the  bundle can indeed be seen as a ``mini-transaction'',
-%in itself it is not very useful. In fact, as we will discuss
-%in more detail in the next section, we will not make
-%use of the bundle in the implementation of our synchronization
-%objects, but will employ it for our case study
-%of consistent policy updates.%
-%
-%Before presenting our transactional interface and implementation,
-%let us first discuss the limitations of the OpenFlow features discussed above.
-
-%\section{The Container Object}\label{sec:vision}
-
 \section{In-Band Synchronization}\label{sec:main}
@@
 together with the corresponding error code.
-%Again, the bundle always returns a binary result: \emph{error} ($\bot$) or \emph{success} ($\top$).
-%\petr{Should also provide the reason why the bundle has failed?}
-
-\hide{
-Given this motivation, we now present a more powerful transactional interface.
-At a high level, we aim to export a
-transactional interface which allows each controller $i\in[N]$ ($N\geq
-2$) to perform multiple
-operations on the configuration of an OpenFlow switch as an atomic
-\emph{transaction}, inspired by the transactional-memory paradigm~\cite{stm-st95,tm-book}. A transaction may \emph{commit}, in
-which case it appears as executed sequentially, or \emph{abort}, in
-which none of its operations takes effect.
-}
 
@@
 %
-\hide{
-
-Our approach is based on the distributed management of \emph{identifiers (IDs)} \petr{unclear what "identifiers" mean here}. In particular,
-we introduce two new synchronization objects (and describe their implementation in standard OpenFlow),
-which allow controllers to coordinate
-themselves based on these identifiers.
-
-As we will see, in the OpenFlow implementation of our mechanism,
-in addition to the ``normal'' configuration stored at the switch
-(e.g., forwarding rules according to network policies),
-controllers will maintain additional information
-in the switch configuration: this information is used by the
-controllers to synchronize.
-\petr{"As we will see,..." - looks funny compared with "As we will see ..." at the end of Sec 2}
-\subsection{Synchronization Objects}\label{sec:t-if}
-}
-%
+
 We focus first on the consistent use of  \emph{policy identifiers},
@@
 $\addr$ to $k$.
-%
-%For  example, the offset $\addr$ may correspond to the policy identifier (\emph{pid}),
-%  but our approach is more general.
-%
 A $\textit{\compare}(\addr,k)$ operation \emph{aborts} the transaction
@@
 is currently claimed by any controller.
-%While In theory one can use the memory abstraction to implement the resource claimer,
-%however we found it far more simple and efficient to consider the resource claimer as base object.
-%
 More precisely, the id-claimer object exports three operations \emph{claim},
@@
 semantics:
-%The container synchronization object introduced in this paper
-%is essentially a transactional memory abstraction which can be implemented
-%in standard OpenFlow switches.
-%Before presenting the technical details in the next section,%
-%we give a brief overview.
-
-%The postulated container object can store an arbitrary number of \emph{items}.
-%As we will see, these items correspond to flow entries on the OpenFlow switch.
-%The same item may even exist multiple times in the container (i.e., the container
-%is a multi-set).
-%The container
-%provides two main functions:
+
 \begin{itemize}
@@
 aborts the current transaction.
-%returns $\textit{true}$ if and
-%  only if $k$ is not currently claimed by any controller; if
-%  $\textit{false}$ is returned, we say that the check operation
-%  \emph{fails} (which causes an abort of the transaction invoking it).
 
 \end{itemize}
-
-%As we will see, the bundle feature is not required to implement these semantics.
-%Also note that memory read operations correspond to reading the switch
-%configuration, and are simple as they do not affect the execution of the transaction
-%(unlike the $\textit{\compare}$ operation).
-\petr{read operations "do not affect the execution of the transaction" appears strange, given that a transaction, if committed, "appears to execute sequentially".
-Something seems wrong here...}
-
-\begin{comment}
-\subsection{Interface}\label{sec:t-if}
-
-\liron{this subsection doesn't flow well for me}
-
-We assume that transactions may comprise operations on the flow table
-of an SDN switch. Our transactional abstraction exports the
-$\textbf{bundle}\{\;\ldots\;\}$ construct
-containing a sequence of transactional
-operations and may \emph{commit} and return responses to all its operations
-or \emph{abort} (so that none of its operation takes effect).
-We guarantee that in every execution, all committed
-transactions constitute a \emph{legal} sequential
-order~\cite{Pap79-serial}, i.e., every transaction appears as executed
-sequentially. On the progress side, we ensure that a
-transaction $T$ is only allowed to abort if either another
-transaction updated a flow table entry read by $T$
-or if one of $T$'s operations (\textit{check} or CAS) fails.
-We employ the \emph{deferred-update} semantics:
-no transaction can affect other transactions before it commits.
-In particular, if a controller crashes before committing its
-transactions, no future transaction can be affected.
-%Note that
-%, assuming such a transactional memory,
-%a $\emph{contains}(x)$ operation can be implemented as a transaction
-%composed of $\textit{cond-add}(x)$ followed by $\textit{cond-delete}(x)$.
-%A \emph{set} abstraction is a special instance of a multi-set that is
-%only modified with \textit{cond-add} and \textit{cond-delete}
-%operations (this way the set only contains at most one copy of a given
-%item).
-
-\end{comment}
 
@@
 
-%In this section, we recall the basic SDN principles, with a focus on
-%OpenFlow~\cite{of-spec}, the standard SDN protocol today.
-\vspace{1mm}
-\noindent\textbf{Control and data planes.} Software-defined networking
-decouples the \emph{control plane}, software
-controllers that configure the network, from the \emph{data plane}, the
-devices that implement the network.
-%
-In OpenFlow, the \emph{de facto} standard SDN protocol
-today, the data plane consists of \emph{OpenFlow switches}.
-In a nutshell, the configuration of an OpenFlow switch is a set of
+As a case study, in this section, we show how to implement
+our transactional approach using today's standard SDN protocol,
+ OpenFlow.
+
+\subsection{Background on OpenFlow}
+
+The configuration of an OpenFlow switch is a set of
 \emph{rules}.
@@
 packets (e.g., forward them to a specific port).
-
 More precisely, the match part of a rule is
@@
 
-\hide{
-%[[PK should go to the discussion?
-We also remark that our solution is only based on flow tables
-and does not make use
-of \emph{OpenFlow group tables}, nor do we make use of \emph{cookies}: rules
-are always identified by exact matches.
-%]]
-}
-
 \vspace{1mm}
@@
 bundle, thereby respecting dependencies between bundle commands.
-%\liron{i added controller atomicity and order preservation, see that it make sense...}
-
-%:a $\checko$ flag can be defined for each command in the bundle.\liron{these last two sentences are not cleaar}
-
-%\liron{the notion of transaction presented next is confusing considering the bundle just explained}
-%Concretely, a transaction can represent between two to five switch interactions:
-%if limited space, Algo 7: read, claim id, read again, transaction, unclaim;
-%if if unlimited space: alg 6 is unbounded, increase policy ID by one.
-%transactions = five switch interactions , claim \pid (line 5+6) together in bundle => one iteraction less?
-
-%In general, it is desirable to minimize the number of transactions \stefan{stefan should cite \cite{Jin2014Dionysus} here}
-
-\hide{
-To commit the transaction,  the controller sends a \texttt{OFPT\_BUNDLE\_CONTROL} message with type
-\texttt{OFPBCT\_COMMIT\_REQUEST} and waits until a matching control
-message from the switch is received. When the switch receives the
-commit request it tries to atomically process all the buffered messages
-equipped with the corresponding bundle identifier and, if the
-processing is successful for all of them, returns  a \texttt{OFPT\_BUNDLE\_CONTROL} message with type
-\texttt{OFPBCT\_CLOSE\_REPLY}; otherwise it sends a specific error
-message  pointing out the aborted transactional operation.
-
-Algorithm~\ref{alg:commit} describes how to commit
-the bundle.
-\begin{algorithm}[t]
-    \caption{$\textit{try-commit}()$}
-    \label{alg:commit}
-    \begin{algorithmic}[1]
-    \Require  current bundle id $bid$.
-		    \State $cmd \gets struct ofp_bundle_ctrl_msg$
-    		\State  $cmd.bundle\_id \gets bid$
-    		\State  $cmd.type \gets $ \texttt{ofpbct\_commit\_request}
-    		\State  $cmd.flags \gets $ \texttt{ofpbf\_atomic}
-    		\State \textbf{send} $cmd$
-		\Return $cmd$
-    \end{algorithmic}
-\end{algorithm}
-}
-
-%\textbf{Cookies:} Cookies allow to name and retrieve individual
-%flow entries and to use masks to delete modify/delete multiple entries.
-%Our synchronization mechanism will implement controller-specific cookies.
-%
-%\item \textbf{Groups:} Groups of action buckets can be defined in a special group table.
-Each group has a unique id and any attempt to create a group with existing id or %to delete un existing group fails (and can abort a bundle in process).
-%\end{enumerate}
-
-%\begin{figure*}[t]
-%\centering
-%\includegraphics[width=0.35\columnwidth]{loadbal-before.pdf}~\includegraphics[width=0.35\columnwidth]{loadbal.pdf}~~
-%\includegraphics[width=0.8\columnwidth]{pic.pdf}\\
-%\caption{\emph{Left:} Load-balancing gone wrong. \emph{Right:} Illustration of transactional abstraction which  allows
-%controllers to modify switch configurations in complex ways,
-%and
-%depending on other flow spaces. Colors indicate different actions. \stefan{lets make two real transactions: one bundle and one 2 or 5 interactions?}}\label{fig:illu}
-%\end{figure*}
-
-\vspace{1mm}
-\noindent\textbf{Implementation.} %\label{sec:t-impl}
-%
-We now describe how to implement our transactional
-abstractions using standard OpenFlow features.
-Essentially, we translate each of the transactional operations into a
-sequence of FlowMod commands.
+
+\subsection{Implementation}
+
+At first sight, bundling \emph{FlowMod} commands with  the $\checko$
+flag already provides a useful synchronization mechanism.
+Indeed, its ``mini-transaction'' nature allows a controller
+to install multiple flow entries in an atomic way.
+However, as is,
+bundling has important limitations, as we will argue in the following.
+
+First, %if employed without any additional features,
+except for some limited constraint types (such as
+overlapping flow spaces or insufficient free space),
+a bundle does not provide any means to modify the switch configuration
+based on application-specific conditions, which depend on the
+current configuration,
+However, as we have for example seen in our load-balancing
+example, a support for more generic conditions under which configuration
+updates
+should or should not take place, is often desirable.
+Moreover, it is sometimes desirable to react to conflicts in smarter ways than
+by simply aborting an operation, e.g., by supporting \emph{conditional modifications}.
+
+Nevertheless, we now show that OpenFlow can be used 
+to realize our transactional
+abstractions. Concretely, the OpenFlow match-action paradigm can be used
+to implement a key-value store.
+The OpenFlow
+configuration space can be realized using low-priority rules.
+Alternatively, we can leverage the possibility to maintain multiple
+flow tables at an OpenFlow switch, and use a separate flow table for
+the meta-configuration,  making sure that the
+\emph{pipelining} mechanism will never involve this
+flow table.
+
+Recall that a transaction consists of a sequence of operations,
+each of which can either be a regular FlowMod command or a synchronization
+primitive, which, as we
+show below, translates into a sequence of FlowMod commands operating on the
+switch meta-configuration.
 
@@
 sufficient to go over the rules matching every controller identifier.
-
-%A container contains items: they represent the flow table entries (i.e., a match-action rules, plus e.g., counters belonging to the rule).
-%Items can be added and removed atomically and in bundles (e.g., to update the switch with
-%a new policy). Recall that we want to add items only if they do not exist
-%yet; otherwise the transaction should abort without taking effect.
-%And similarly for removes: transactions should not take effect if the corresponding
-%update contains rules which currently do not exist at the switch.
 
@@
 
+\subsection{Proof-of-Concept}
+
+As a proof of concept, we implemented the algorithms as a python library that wraps ovs-ofctl - a well known command line tool for monitoring and administering OpenFlow switches. 
+We evaluated the framework using a series of tests that used our low-level abstractions as well as the transactional policy update algorithm.
+
+We used three processes to make 3000 concurrent policy updates, each update consists of replacing a single rule with another one and we verified that all transactions completed correctly.
+The tests were conducted using a software switch, Open vSwitch (version 2.4.0) and virtual network simulator, Mininet.
+This framework can be used as synchronization tool by itself and we also work on to implementing it as as integral part of existing controllers such as Ryu.
+
 \section{Applications}\label{sec:apps}
@@
 We now describe sample scenarios of how of our synchronization mechanisms
-can be used for control applications. We aim at two applications: a
-simple load-balancing scheme and a more sophisticated forwarding policy
+can be used for control applications. We aim at two applications: the
+simple load-balancing scheme we discussed as an example in the beginning,
+and a more sophisticated forwarding policy
 composition, but we begin with implementing a fundamental higher-level
@@
 
+{\small
+\begin{algorithm}[t]
+\caption{$\textit{Na\"ive Flow Balancing}$}
+\label{alg:naive-lb}
+\begin{algorithmic}[1]
+
+\Require \emph{new\_flows}: flow entries of the new flows,
+\emph{ports}: a list of switch output ports,
+\State $\textit{policy} \gets \textbf{read\_config}$
+\State $\textit{update\_cmds} \gets \textsc{lb\_update}(policy)$ %(new\_flows, ports, policy)$
+\State $\textbf{send}\{\textit{update\_cmds}\}$
+% \State $\textbf{execute-atomic}\{update\_cmds\}$
+\Statex
+
+\Statex \emph{\bf Function} $\textsc{lb\_update}(\textit{policy})$ % (new\_flows, ports, policy)}$
+
+\State $\textit{flows\_per\_port} \gets [0, 0, \ldots, 0]$
+\ForAll {$\textit{entry} \in \textit{policy}$}
+\State $\textit{port} \gets \textit{entry.output\_port}$
+\State $\textit{flows\_per\_port}[\textit{port}]++$
+\EndFor
+\State $\textit{new\_flow\_mod\_cmds} \gets \{\}$
+\For {$\textit{entry} \in \textit{new\_flows}$}
+\State $\textit{port} \gets \arg\min_{p\in \textit{ports}} \textit{flows\_per\_port}[p]$
+\State $\textit{entry.output\_port} \gets \textit{port}$
+\State $\textit{flows\_per\_port}[\textit{port}]++$
+\State $\textit{cmd} \gets \textit{FlowMod}(add,f)$
+\State $\textit{new\_flow\_mod\_cmds}.\textit{add}(\textit{cmd})$
+\EndFor
+\Return $\textit{new\_flow\_mod\_cmds}$
+% \State \emph{\bf EndFunction}
+\end{algorithmic}
+\end{algorithm}
+}
+
 \vspace{1mm}
 \noindent\textbf{Load balancing.}
-CAS can be used to solve the fundamental
-problem of fault-tolerant \emph{consensus}~\cite{FLP85},
-\hide{%[[PK, feels like we do not really need this level of detail here
- in which the controllers
-need to agree on one of their private inputs, despite failures of an
-arbitrary number of controllers.
-Indeed, each controller simply performs, within a transaction, a CAS operation to replace the default value
-in a memory location at the meta-configuration of a data plane switch.
-If the transaction commits, the controller considers itself
-the \emph{winner} and decides on its input,
-if it aborts (the controller fails in the CAS), it remains to read the
-memory to get the input of the winner and decide on it.
-}
-which allows us
-to implement any generic replicated state machine~\cite{Her91}
-in the control plane,  in the \emph{wait-free} manner, i.e., tolerating
-asynchrony and failure of any number of controllers.
-As such, CAS can also be used to implement STN~\cite{stn},
-by providing the postulated \emph{read-modify-write} object.
-
-%which will be instrumental in the
-%next section.
-
-%Moreover, given the CAS, we can also implement more complex operations,
-%for example the replicated state machine postulated in STN~\cite{stn}:
-%in order to choose tags in a consistent manner, CAS needs to be enhanced%
-%with a notion of tags.
-
-%\petr{when talking about replicated state machines, we should make it clear that we that we talk about replication among controllers.}
-
-%\petr{"more complex operations": more complex than what?}
-
-%\petr{the discussion of CAS in STN is confusing, given that we never described STN.}
-
-\vspace{1mm}
-\noindent\textbf{Generic policy composition.}
+A simple strategy to balance load in our 2-link example
+is given in Algorithm~\ref{alg:naive-lb}.
+However, this algorithm fails in the presence of
+multiple controllers, even if 
+bundling and $\checko$ features are used:
+the flows are defined over ``independent'' 
+flow spaces.
+
 Our mechanism can also be used as a generic template to
@@
 This way, for example, by choosing our load-balancing
-function $\textsc{lb\_update}$ (Example~1 in
-Section~\ref{sec:motivation}), 
+function $\textsc{lb\_update}$, 
 the pathological scenario described in 
@@
 %used to store $\textit{pid}$ and id-claimer (the claims).
-
-\hide{
-We compute our new suggested policy by applying the update requests on
-top of current policy, supporting any kind of requests and
-policies. Then we make a transaction (using the bundle feature) to
-atomically check that our policy id is not blocked by another controller, to change the current policy id to ours (an action that would fail if the current policy id is no longer what we are counting of) and to actually configure our new policy.
-
-If one of the actions in the transaction fails we try again.
-There is no progress guaranty for each controller but there is one for the whole system - at least one of the controller will succeed in fulfilling its update requirements.
-}
 
@@
 
-%\noindent\textbf{One-Shot Update} FIXME: desribe how to implement one-shot update
-%
-%\noindent\textbf{STN}
-%
-%\textbf{TODO}
-%
-
-\begin{comment}
-\section{In-Band Composition}\label{sec:compo}
-
-We conclude our technical contribution by showing that sometimes,
-powerful synchronization mechanisms can be implemented
-in-band \emph{even without bundles}.
-Concretely, we show that sometimes it is even possible
-to automatically compose updates originating from different controllers,
-in-band and just using OpenFlow.
-
-To illustrate this point, we consider a simple example
-with two controllers concurrently issuing updates to a single
-given switch.
-
-FIXME: Explain, maybe even with an example?
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=0.8\columnwidth]{composition.pdf}\\
-\caption{Composition.}\label{fig:illu}
-\end{figure}
-
-\end{comment}
-
-\begin{comment}
-
-\section{Transactions in OpenFlow}\label{sec:sync}
-
-Multi-write and read-modify transactions.
-
-\subsection{Dealing with multiple updates}
-
-TBD - adding the notion of versions and cleanup/GC. probably no good way to solve infinite versions... maybe check literature on shared memory and multiple writes.
-
-\subsection{Improving the scheme with check-overlap}\label{sec:todo}
-
-Till now we considered a model that allows multiple updates to match-action entries.
-Here we extends the model with actions create-entry and delete-entry that may failed and abort the whole transaction depending on the existence on non-existence of specific entries prior to the action.
-%TBD - much less resources. easily solving the infinite versions issue.
-
-Next we show how such actions can be used to implement consensus with much less resources. We use entry id as a property used to reference specific entry.
-
-\section{Applications and ramifications}\label{sec:application}
-
-\section{The Underlying Theory}\label{sec:background}
-
-idea: multi-write consensus, a not well-known result!
-
-In order to solve consensus we suggest to use the atomic multiple entries update capability (which is discussed in OpenFlow standard).
-A shared memory system that supports atomic write to (enough) multiple locations is known to allow the implementation of consensus objects.
-For example consider the following implementation utilizing $n^2$ memory locations, $M_{i,j}$, where $i,j\in[n]$, and the suggested values $\{v_i\}_{i\in [n]}$.
-All locations are initialized to zero.
-
-In order to participate, a server i, writes its suggested value $V_i$ and atomically rewrites row i ($M_{i,*}$) with value "1" and column i ($M_{*,i}$) with value "2".
-After the atomic write, a server $i$ first reads the diagonal ($M_{j,j}$ for $j\in [n]$), and consider all observed servers as candidate set S.
-Then server i reads all candidates intersections (locations $M_{j,k}$ where $j,k\in S$.
-And computes the winner server id w that was overwritten by all other candidates, i.e.
-such that $M_{w,k}=2$ for all $k\in S {w}$.
-The consensus value is $v_w$.
-Note that a snapshot could replace the two reading phases.
-
-\section{Basic One-Touch Mechanism}\label{sec:realization}
-
-We address the problem of committing consistent policies to SDN switch concurrently by multiple controllers.
-We show that assuming minimalistic control protocol, SDN switch configuration space can be used to implement a consensus object thereby allowing controllers to agree on each next policy update.
-
-\subsection{SDN switch configuration space as shared memory}
-
-Our first observation is that a match-action configuration which is accessed by multiple controllers can be considered as shared memory.
-Note however that standard memory map and index/offset to a word, while Flow Table entries are not necessarily indexed.
-However, we assume that each flow entry has some property that allows the operator to specifically reference and update it, for example OpenFlow entry's cookie.
-In cases where this property value can't be determined by the operator, we can use other variable entry properties to store the index, for example we can represent memory value $v$ in offset $i$ by an match-action entry whose match is $in\_port = i$ and action is "forward port $x$".
-%To overcome this we can use the match part of an entry as the index and the action part as the word value.
-
-Going back to the policy update problem, the consensus scheme allow all servers to decide on next policy by using consensus value as policy identifier, and writing the policy in advance.
-However this scheme doesn't actually apply the policy on incoming packets and it remains to configure the switch according to the next policy once it is decided.
-Making this scheme fail safe (handling failure of winning server) can be ensured by allowing any server to reconfigure the switch.
-
-\subsection{ One-Touch}
-
-In some scenarios it may be desirable to shorten the policy update time by avoiding the consensus reading and policy configuration step.
-We define this as the one touch policy update, namely we require each server to contact the switch only once in order to (try and) update the policy, in other words participating in reaching consensus and applying it with one atomic write.
-Solving this problem also has significance in understanding the strength of the SDN switch computation model.
-
-The main idea of our solution is to adjust the previous policy update scheme in a way that performs consensus validation phase during every packet processing thereby requiring the servers only to make the first phase of the consensus - atomically writing the matrix.
-In order to allow the packets to validate the consensus we make the following observation: while servers can be abstracted as memory writers, the packets processed by the switch are the readers.
-OpenFlow standard define an atomic write transaction as being atomic in relation to the processed packets thereby making packet processing a multiple read transaction but one read location per table is allowed.
-This means that if each memory location used in the consensus scheme will be stored in different table then each packet will be able to read all of them.
-
-However, reading each memory location is not enough, as the scheme requires to make a validation that involves multiple values.
-We utilize the metadata field in order to store all read values, in more details, each memory value $M_{i,j}$ is stored in offset i+j*n in the metadata field.
-Once all values are in the metadata field, we can detect the consensus winner by trying different matches on the metadata.
-Although we can have one match entry per possible matrix state, this solution requires exponential number of entries.
-A better way which we describe next requires only $O(n^2)$ entries utilizing n tables.
-
-Our compact validation use additional temp variable, $temp_winner$, to hold the current best consensus candidate.
-Each of the n validating tables checks different "column" in the matrix and update $temp_winner$ to keep track of the winner so far (after examining a prefix sub set of the columns).
-The entry t in table k $(0<=k<n)$  matches the case where current winner=t and location $(t,k)$ in the matrix (packet header) equals 1 and location (k,k) is non-zero.
-After all validating tables are processed, $temp_winner$ stores the id of the winning server which can be used to match on its policy rules (filtering out other policies) or to use the goto table command to jump to a designated policy table of the server.
-
-\section{The Practical Implementation}\label{sec:extension}
-
-\end{comment}
-
-\begin{comment}
-\section{Related Work}\label{sec:relwork}
-
-There exists a wide consensus that SDN control planes need to be
-distributed~\cite{onos,onix,elasticon}.
-Onix~\cite{onix} is one of the earliest proposals, and introduced the
-the notion of Network Information Base (NIB), abstracting network state
-distribution from control logic, but
-requiring mechanisms for the detection and resolution of conflicts.
-Also spatially distributed control planes to improve scalability and
-latency have been studied intensively
-in the literature~\cite{sharon,kandoo,ctrl-place,hotsdn13loc}
-proposes an elastic distributed controller architecture.
-
-Operating and updating SDNs efficiently and consistently is already non-trivial
-in a setting consisting of a single controller and a single switch.~\cite{sharon}
-Reitblatt et al.~\cite{network-update}
-introduced the notion of
-per-packet consistency and proposed a tagging-based, 2-phase update protocol
-which allows a single controller to update network routes.
-%, and described the \emph{two-phase update} technique, also used in our algorithms.
-Mahajan and Wattenhofer~\cite{roger-hotnets} and Ludwig et al.~\cite{podc15}
-considered weaker transient
-consistency guarantees, and proposed more efficient network update algorithms
-accordingly. Ludwig et al.~~\cite{hotnets14update} studied algorithms for secure
-network updates where packets are forced to traverse certain waypoints or
-middleboxes. Ghorbani et al.~~\cite{correct-virt} recently argued for the design
-of network update algorithms that provide even stronger consistency guarantees.
-
-To the best of our knowledge, the only paper explicitly addressing the consistent
-network update problem from a distributed-controllers perspective is STN~\cite{stn}:
-STN provides a transactional interface offering all-or-nothing semantics and serializability
-(the ``holy grail'' of safety properties), allowing
-controllers to install policies in a conflict-free manner; if some controllers fail,
-other controllers can take over. STN~\cite{stn} is based on
-atomic read-modify-write primitives provided by the switches, but
-while the availability of such primitives is postulated, no implementation is
-described. Also NetPaxos~\cite{netpaxos} observes the usefulness of
-providing synchronization primitives inband, and proposes simple OpenFlow
-extensions to implement those.
-
-In this paper, we provide this missing link, and also show that other useful and
-more powerful synchronization primitives can be implemented in today's OpenFlow.
-Also, the STN protocol~\cite{stn} assumes that the control
-plane maintains a certain amount of synchrony which allows controllers
-to solve consensus or implement a replicated state machine.
-In this paper, we describe an OpenFlow-based consensus protocol that
-can be used for this purpose.
-
-Our work is also closely related to the recent work on providing more high-level
-abstractions and programming languages for SDNs, such as Frenetic~\cite{frenetic},
-also enabling policy composition~\cite{pyretic}. Our work complements these works
-on programming abstractions
-by initiating the discussion of concurrency objects and abstractions.
-
-The transactional interface we describe in this paper is inspired by the
-work on speculative concurrency control using software transactional memory
-(STM)~\cite{stm-st95,tm-book}.
-%[[PK I think this is enough]]
-
-Our work also contributes to the ongoing discussion of what can be implemented
-in-band in OpenFlow~\cite{compute,reclaim}, by focusing on the
-relevant problem of synchronization.
-
-\hide{
-In terms of distributed computing: TODO PETR: maybe modify a bit the text below
-and also write more about multi-writer consensus etc.?
-
-\noindent\textbf{Distributed Computing.}
-There is a long tradition of defining correctness of a concurrent system via
-an equivalence to a sequential one~\cite{Pap79-serial,Lam79,HW90}.  The notion
-of sequentially composable histories is reminiscent of
-linearizability~\cite{HW90}, where a history of operations concurrently
-applied by a collection of processes is equivalent to a history in which the
-operations are in a sequential order, respecting their real-time precedence.
-In contrast, our sequentially composable histories impose requirements not
-only on high-level invocations and responses, but also on the way the traffic
-is processed. We require that the committed policies constitute a
-conflict-free sequential history, but, additionally,  we expect that each
-\emph{path} witnesses only a prefix of this history, consisting of all
-requests that were committed before the path was initiated.
-%
-The transactional interface exported by the STN abstraction is inspired by the
-work on speculative concurrency control using software transactional memory
-(STM)~\cite{stm-st95}.
-%, thus the  term \emph{software transactional networking}.
-Our interface is however intended to model realistic network
-management operations, which makes it simpler than recent
-dynamic STM models~\cite{dstm}.
-%In particular, the sets of rules to be installed by a policy update do not depend on the state of the
-%network.
-
-Extending the interface to dynamic policies that adapt their behavior based on
-the current network state sounds like a promising research direction.  On the
-other hand, our criterion of sequential composability is more complex than
-traditional STM correctness properties in that it imposes restrictions not only
-on the high-level interface exported to the control plane, but also on the
-paths taken by the data plane packets.
-Also, we assumed that controllers are subject to failures, which is usually not
-assumed by STM implementations.
-}
-
-\end{comment}
-
 \section{Conclusion}\label{sec:conclusion}
@@
 inband synchronization abstractions, also in terms of performance.
-%In the long run, we seek to develop in-band mechanisms
-%which allow us to seemlessly transition between single controller and
-%multiple controller mode.
-
-%We also plan to release a small library with our synchronization primitives. 
-%Moreover, we are investigating possible generalizations of our transactional guarantees,
-%across multiple switches~\cite{stn}, as well as opportunities for more fine-grained
-%\emph{partial} updates that concern subsets of flow entries, thus improving concurrency.
-
-%,
-%nor on
-%nor do we make any assumptions on packet orderings.
-%and do not rely on any
-%
-%.
-
-\begin{comment}
-While much progress has been made over the
-last years in the development of more intuitive programming
-languages for SDNs (e.g., Frenetic, Pyretic, Merlin), the
-question of which synchronization abstractions to provide
-has received much less attention.
-%
-In this paper, we describe transactional abstractions that
-can be implemented using standard OpenFlow features.
-%
-Our work complements existing research on the design of
-distributed control planes, and also provides some missing links,
-e.g., for the transactional approach taken by STN~\cite{stn}.
-We also hope that our work can nourish the discussion of
-which synchronization objects we can and want to implement
-in an SDN.
-
-Even though we believe that we propose efficient and consistent
-transactional framework for in-band synchronization, we do not claim
-that we offer an exhaustive set of synchronization primitives. Indeed,
-one may explore standard OpenFlow features to enable more powerful
-primitives. For instance, \emph{cookies}~\cite{openflow} can be used for modifying
-flow tables, based on more sophisticated conditions than the default
-``exact match''.
-Also, the standard pipelining feature allows the switches to maintain multiple flow
-tables which can be traversed in arbitrary patterns, defining the sets
-of corresponding actions.
-Exploring the power of such features is left to future work.
-\end{comment}
-
 {
@@
 
-\begin{comment}
-
-\clearpage
-
-\onecolumn
-
-\begin{appendix}
-
-\section{The FlowMod Command}
-
-\lstset{% general command to set parameter(s)
-	basicstyle=\small, % print whole listing small
-	keywordstyle=\color{black}\bfseries, % underlined bold black keywords
-	identifierstyle=, % nothing happens
-	commentstyle=\color{mygrey}, % white comments
-	stringstyle=\ttfamily, % typewriter type for strings
-	showstringspaces=false, % no special string spaces
-	language=C}
-\begin{lstlisting}
-struct ofp_flow_mod {
-	struct ofp_header header;
-	uint64_t cookie; /* Opaque controller-issued identifier. */
-	uint64_t cookie_mask; /* Mask used to restrict the cookie bits
-		that must match when the command is
-		\texttt{ofpfc_{modify}*} or \texttt{ofpfc_{delete}*. A value
-		of 0 indicates no restriction. */
-	/* Flow actions. */
-	uint8_t table_id; /* identifier of the table to put the flow in.
-		For OFPFC_DELETE_* commands, OFPTT_ALL
-		can also be used to delete matching
-		flows from all tables. */
-	uint8_t command; /* One of OFPFC_*. */
-	uint16_t idle_timeout; /* Idle time before discarding (seconds). */
-	uint16_t hard_timeout; /* Max time before discarding (seconds). */
-	uint16_t priority; /* Priority level of flow entry. */
-	uint32_t buffer_id; /* Buffered packet to apply to, or
-		OFP_NO_BUFFER.
-		Not meaningful for OFPFC_\dele*. */
-	uint32_t out_port; /* For OFPFC_\dele* commands, require
-		matching entries to include this as an
-		output port. A value of OFPP_ANY
-		indicates no restriction. */
-	uint32_t out_group; /* For OFPFC_\dele* commands, require
-		matching entries to include this as an
-		output group. A value of OFPG_ANY
-		indicates no restriction. */
-	uint16_t flags; /* Bitmap of OFPFF_* flags. */
-	uint16_t importance; /* Eviction precedence (optional). */
-	struct ofp_match match; /* Fields to match. Variable size. */
-	/* The variable size and padded match is always followed by instructions. */
-	//struct ofp_instruction_header instructions[0];
-		/* Instruction set - 0 or more. The length
-		of the instruction set is inferred from
-		the length field in the header. */
-};
-\end{lstlisting}
-
-- Creating an entry with the same match part replaces the existing one.
-
-\section{Bundle Control Message}
-
-%/* Message structure for OFPT_BUNDLE_CTRL. */
-\begin{lstlisting}
-struct ofp_bundle_ctrl_msg {
-	struct ofp_header header;
-	uint32_t bundle_id; /* Identify the bundle. */
-	uint16_t type; /* OFPBCT_*. */
-	uint16_t flags; /* Bitmap of OFPBF_* flags. */
-	/* Bundle Property list. */
-	struct ofp_bundle_prop_header properties[0]; /* Zero or more properties. */
-};
-\end{lstlisting}
-
-\begin{lstlisting}
-* Adding a message in a bundle is done with. */
-struct ofp_bundle_add_msg {
-	struct ofp_header header;
-	uint32_t bundle_id; /* Identify the bundle. */
-	uint16_t pad; /* Align to 64 bits. */
-	uint16_t flags; /* Bitmap of OFPBF_* flags. */
-	struct ofp_header message; /* Message added to the bundle. */
-	/* If there is one property or more, 'message' is followed by:
-	* - Exactly (message.length + 7)/8*8 - (message.length) (between 0 and 7)
-	* bytes of all-zero bytes */
-	/* Bundle Property list. */
-	//struct ofp_bundle_prop_header properties[0]; /* Zero or more properties. */
-};
-\end{lstlisting}
-
-\section{To Be Deleted...}
-
-\begin{algorithm}[h]
-    \caption{Policy composition without bundle}
-    \label{alg:wobundle}
-    \begin{algorithmic}[1]
-    \Require policy $P$ (a set of FlowMod entry creation commands), three consecutive flow tables identifiers (per controller) $T_s,T_0,T_1$, active table index $active\in\{0,1\}$.
-    %\Ensure installed policy is consistent with previous one
-    \State $unactive \gets active + 1 \mod{2}$
-    \For {$cmd \in P$}
-	    \State $cmd.action\gets Reencode(cmd.action)$
-	    \State $cmd.table\_id\gets T_{unactive}$
-	    \State send $cmd$
-    \EndFor
-    \State modify the action of $T_s$ entry to goto $T_{unactive}$
-    \State clear table $T_{active}$
-    \State $active \gets unactive$
-	\Return
-
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[h]
-    \caption{Update with pipeline}
-    \label{alg:pipeline}
-    \begin{algorithmic}[1]
-    \Require update requirements $U$, array $ARR$.
-    \Ensure installed policy is consistent with previous one
-			\Return
-
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{comment}
-\section{Implementing CAS with FlowMod commands}\label{sec:without-groups}
-This section describes an implementation without group tables.
-
-Here we describe a second implementation of the Compare\&Set abstraction that is based on entirely on flow entry creations and deletions (FlowMod commands).
-
-The $\checko$ flag enables us to add an entry if no contradicting entry exists and to fail if such do exists.
-However, the compare stage of \cas requires us to fail if a certain value doesn't exists, i.e.
-$\cas(x,y)$ failes iff $x \notin T$, where $T$ is the set of match entries defined in a dedicated flow table.
-...
-...
-
-\begin{algorithm}[H]
-    \caption{$\textit{CAS}(old,new)$}
-    \label{alg:cas2}
-    \begin{algorithmic}[1]
-    %\Require  id $i$.
-%    \Ensure
-			\State $cmds \gets new\ list()$
-    		\State $value \gets  old$
-    		\State $mask \gets  \texttt{1$\ldots$1}$
-    		\State $match \gets (value,mask)$
-    		%\State $cookie \gets x\concat self$
-    		\State $action \gets 2$
-    		\State $flag \gets OFPFF\_\checko$
-    		\State $cmd1\gets \emph{FlowMod}(match, op = \add, flag, action) $
-    		\State $cmd2\gets \emph{FlowMod}(match, op = \dele) $
-    		\State $cmd3\gets \emph{FlowMod}(cookie = old, op = \dele) $
-    		\State $cmds.append(cmd1)$
-    		\State $cmds.append(cmd2)$
-    		\State $cmds.append(cmd3)$
-    		\State $range\_matches \gets EncodeRange([0,new-1])\cup EncodeRange([new+1,MAX\_ID])$
-    		\For {$r\_match \in range\_matches $}
-    		    		\State $cookie \gets new$
-    		    		\State $action \gets 1$
-    		    		\State $flag \gets 0$
-    		    		\State $cmd\gets \emph{FlowMod}(r\_match, \add, flag, cookie, action) $
-    		    		\State $cmds.append(cmd)$
-    		\EndFor
-			\Return $cmds$
-    \end{algorithmic}
-\end{algorithm}
-
-\section{Implementing CAS with GroupMod commands}\label{sec:with-groups}
-Similarly, we use a deletion of group $x$ and a creation of group $y$ (with GroupMod commands) as an
-abstraction to performing a $CAS(x,y)$ operation to a special $16$bit register (we abstract only one register). The $16$ bit limit is resulted by th $16$bit group id size.
-%The group id of the added group represents the added number.
-Note that trying to delete a group that doesn't exist (or trying to create a group with id that already exists) fails, thereby preventing adding the new value $y$ if $x$ hasn't existed before.
-Note that the two action should be performed inside the same bundle in order to ensure the atomic nature ot the abstraction.
-
-\begin{algorithm}[H]
-    \caption{$\textit{group-CAS}(old,new)$}
-    \label{alg:gcas}
-    \begin{algorithmic}[1]
-    %\Require  id $i$.
-%    \Ensure
-
-    		\State $op \gets \texttt{OFPGC\_\dele}$
-    		\State $group\_id \gets old$
-    		\State $cmd1\gets GroupMod(op, group\_id) $
-
-    		\State $op \gets \texttt{OFPGC\_\add}$
-    		\State $type \gets \texttt{OFPGT\_INDIRECT}$
-    		\State $group\_id \gets new$
-    		\State $buckets \gets [dummy\_backet]$
-    		\State $cmd2\gets GroupMod(op, type, group\_id, buckets) $
-			\Return $cmd1,cmd2$
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{comment}
-\section{Backup Stuff}
-
-using the bundling feature to implement \emph{multi-write} or, more
-widely, \emph{multi-modify} transactions. .
-, using a single OpenFlow switch that accepts
-bundle control messages.
-
-We first show that they can be used to replace the read-modify-write
-primitives voluntarily assumed in the recent implementation of a
-recent proposal for a consistent composition system for forwarding policies~\cite{stn}.
-We then use these transactions directly to implement a lightweight
-forwarding policy composition system that requires no additional
-synchronization on the control plane.
-We then explore how more general classes of policies can be composed
-\emph{on-the-switch}.
-
-TODO: write vision and contribution in full detail, as an overview
-
-using the bundling feature to implement \emph{multi-write} or, more
-widely, \emph{multi-modify} transactions. .
-, using a single OpenFlow switch that accepts
-bundle control messages.
-
-We first show that they can be used to replace the read-modify-write
-primitives voluntarily assumed in the recent implementation of a
-recent proposal for a consistent composition system for forwarding policies~\cite{stn}.
-We then use these transactions directly to implement a lightweight
-forwarding policy composition system that requires no additional
-synchronization on the control plane.
-We then explore how more general classes of policies can be composed
-\emph{on-the-switch}.
-
-TODO: write vision and contribution in full detail, as an overview
-
-we can do more than container:
-
-we introduce container: it's not compare+set, ... its something new!
-new object, is stronger than compare and set because bundle!
-
-\end{appendix}
-
-\end{comment}
-
 \end{document}

### 2015-09-01 ccr-rev1/inband-sync.tex
a version

@@
 \author{
-Liron Schiff$^1$, % \thanks{Supported by the European Research Council (ERC) Starting Grant no. 259085 and by the Israel Science Foundation Grant no.~1386/11.} $^1$,
-Stefan Schmid$^2$, Petr Kuznetsov$^3$ \\
-\small $^1$ Tel Aviv University, Israel; $^2$ TU Berlin \& T-Labs,
-Germany; $^3$ T\'el\'ecom ParisTech, France
+Liron Schiff$^1$  \quad % \thanks{Supported by the European Research Council (ERC) Starting Grant no. 259085 and by the Israel Science Foundation Grant no.~1386/11.} $^1$,
+Stefan Schmid$^2$ \quad Petr Kuznetsov$^3$ \\
+\small $^1$ Tel Aviv University, Israel \quad $^2$ TU Berlin \& T-Labs,
+Germany \quad $^3$ T\'el\'ecom ParisTech, France
 }
@@
 
-\vspace{1mm}
-\noindent\textbf{Contributions.}
+%\vspace{1mm}
+%\noindent\textbf{Contributions.}
 This paper applies the principle of atomicity in concurrent computing
@@
 and present a generic template to implement concurrent policy
-compositions of a user-specic update functions.
+compositions of a user-specific update functions.
 Our synchronization abstraction also provides
@@
 
-\noindent\textbf{Organization}
-The rest of the paper is organized as follows.
+%\noindent\textbf{Organization}
+The remainder of the paper is organized as follows.
 In Section~\ref{sec:motivation}, we motivate the 
@@
 can constitute a conflict. 
-However, as we will argue, there are often conflicts which are less obvious,
-and go beyond simple range overlaps. 
+However, as we will argue, 
+there exist many applications with less obvious conflicts, which go 
+beyond simple range overlaps. 
 Moreover, sometimes it is desirable to react to conflicts in smarter ways than
 by simply aborting an operation, e.g., by supporting \emph{conditional modifications}.
-Let us consider two examples.
-
-\vspace{1mm}
-\noindent \textbf{Example: Handling complex dependencies.} We consider two controllers
+
+Let us consider an example.
+Imagine two controllers
 in charge
-of \emph{load-balancing}, see Figure~\ref{fig:CAS-example}
-for a simplistic example with two links.
+of \emph{load-balancing}, see Figure~\ref{fig:CAS-example}.
 In such a scenario, seemingly independent actions, defined over
@@
 
-This paper proposes mechanisms to solve these synchronization problems 
-by using a minimal
-``meta-configuration'' space. That is, controllers store information
-on contention and conflicts between
-concurrently installed policy updates on the switch. Controllers can read and modify
-this meta-data by applying synchronization primitives.
-In the following section, we
-describe how this space can be organized and consistently maintained
-by multiple controllers despite of concurrency and conflicts.
+This paper proposes mechanisms to solve such more general 
+synchronization problems.
 
@@
 In order to synchronize controllers, we propose
-to use a meta-configuration space at
-the switches which can be manipulated using our
-synchronization framework.
+to use a minimal
+``meta-configuration'' space on the switch,
+to store information
+on contention and conflicts between
+concurrently installed policy updates. Controllers can read and modify
+this meta-data by applying synchronization primitives.
+In the following section, we
+describe how this space can be organized and consistently maintained
+by multiple controllers despite of concurrency and conflicts.
 
@@
 designed in a way that it does not affect the processing of data plane
-traffic. One way to achieve this, for example, is to set a low priority to
-the meta flow entries and ensure that the regular flow entries
-maintain higher priority levels. We will discuss other options later
-in the context of OpenFlow.
+traffic. 
+We will later discuss later how this can be achieved, for example in 
+today's Openflow.
+
+%One way to achieve this, for example, is to set a low priority to
+%the meta flow entries and ensure that the regular flow entries
+%maintain higher priority levels. We will discuss other options later
+%in the context of OpenFlow.
 
@@
 %
-
-We focus first on the consistent use of  \emph{policy identifiers},
+We aim to enforce a consistent use of  \emph{policy identifiers},
 an important notion in consistent policy updates~\cite{network-update,stn}.
@@
 Intuitively, by combining compare and write with a set of regular
-FlowMod commands in a transaction we can make sure that the
+switch update commands 
+in a transaction (such as FlowMod in OpenFlow)  we can make sure that the
 configuration is changed and the current policy identifier is updated
@@
 
-\subsection{Background on OpenFlow}
+\subsection{Background}
 
@@
 
-At first sight, bundling \emph{FlowMod} commands with  the $\checko$
+At first sight, it may seem that bundling \emph{FlowMod} commands with  the $\checko$
 flag already provides a useful synchronization mechanism.
@@
 updates
-should or should not take place, is often desirable.
+should or should not take place, is essential.
 Moreover, it is sometimes desirable to react to conflicts in smarter ways than
@@
 corresponding responses are received.
-If an error message is return,
+If an error message is returned,
 it is forwarded to the control application,
@@
 
-As a proof of concept, we implemented the algorithms as a python library that wraps ovs-ofctl - a well known command line tool for monitoring and administering OpenFlow switches. 
-We evaluated the framework using a series of tests that used our low-level abstractions as well as the transactional policy update algorithm.
-
-We used three processes to make 3000 concurrent policy updates, each update consists of replacing a single rule with another one and we verified that all transactions completed correctly.
+To investigate the feasibility of our approach,
+we implemented a proof-of-concept of our algorithms
+as a python library which wraps ovs-ofctl, 
+a well-known command line tool for monitoring and administering OpenFlow switches. 
+The code is made available together with this paper.
+
+We experimented with the low-level abstractions as well as the 
+transactional policy update algorithm of our proof-of-concept, using
+different workloads. 
 The tests were conducted using a software switch, Open vSwitch (version 2.4.0) and virtual network simulator, Mininet.
-This framework can be used as synchronization tool by itself and we also work on to implementing it as as integral part of existing controllers such as Ryu.
+
+For instance, we experimented with a setting where three processes
+(controllers) issue thousands of concurrent policy updates
+to the switch, where 
+each update consists in replacing a single rule with another one.
+When verifying the transactions, we could confirm that 
+despite the concurrency, all transactions have been successfully completed.
+Of course, due to the concurrency, multiple transactions also needed to abort,
+due to conflicts.
+
+While our framework can be used as synchronization tool itself, we 
+currently also work on implementing it as an integral part of existing controllers 
+such as Ryu.
 
@@
 
-We now describe sample scenarios of how of our synchronization mechanisms
-can be used for control applications. We aim at two applications: the
-simple load-balancing scheme we discussed as an example in the beginning,
-and a more sophisticated forwarding policy
-composition, but we begin with implementing a fundamental higher-level
-\emph{Compare-and-Set} (CAS) primitive. We then sketch show to use CAS, along with
-     {\exec} and {\execatomic} constructs, in our two control applications.
-
-%Our synchronization mechanisms can be useful in many situations.
-
-\vspace{1mm}
-\noindent\textbf{Compare-and-Set.}
-%For example, we can implement the fundamental
-%\emph{Compare-and-Set} (CAS) primitive. % on the policy identifier.
+Given our mechanisms, we can easily implement
+powerful synchronization primitives such at compare-and-set:
 When executed within a
@@
 
-%[[PK said it in the text
-\hide{
-As can be seen Algorithm \ref{alg:cas}, CAS is implemented by two commands. These commands should be sent as part of the same bundle
-
-\begin{algorithm}[t]
-    \caption{$\textit{\cas}(\addr, old,new)$}
-    \label{alg:cas}
-    \begin{algorithmic}[1]
-    %\Require  id $i$.
-%    \Ensure
-
-    		\State $cmd1\gets \textsc{\compare}(\addr, old) $
-    		\State $cmd2\gets \textsc{\memwrite}(\addr, new) $
-			\Return $cmd1,cmd2$
-    \end{algorithmic}
-\end{algorithm}
-}
-%]]
-
-{\small
-\begin{algorithm}[t]
-\caption{$\textit{Na\"ive Flow Balancing}$}
-\label{alg:naive-lb}
-\begin{algorithmic}[1]
-
-\Require \emph{new\_flows}: flow entries of the new flows,
-\emph{ports}: a list of switch output ports,
-\State $\textit{policy} \gets \textbf{read\_config}$
-\State $\textit{update\_cmds} \gets \textsc{lb\_update}(policy)$ %(new\_flows, ports, policy)$
-\State $\textbf{send}\{\textit{update\_cmds}\}$
-% \State $\textbf{execute-atomic}\{update\_cmds\}$
-\Statex
-
-\Statex \emph{\bf Function} $\textsc{lb\_update}(\textit{policy})$ % (new\_flows, ports, policy)}$
-
-\State $\textit{flows\_per\_port} \gets [0, 0, \ldots, 0]$
-\ForAll {$\textit{entry} \in \textit{policy}$}
-\State $\textit{port} \gets \textit{entry.output\_port}$
-\State $\textit{flows\_per\_port}[\textit{port}]++$
-\EndFor
-\State $\textit{new\_flow\_mod\_cmds} \gets \{\}$
-\For {$\textit{entry} \in \textit{new\_flows}$}
-\State $\textit{port} \gets \arg\min_{p\in \textit{ports}} \textit{flows\_per\_port}[p]$
-\State $\textit{entry.output\_port} \gets \textit{port}$
-\State $\textit{flows\_per\_port}[\textit{port}]++$
-\State $\textit{cmd} \gets \textit{FlowMod}(add,f)$
-\State $\textit{new\_flow\_mod\_cmds}.\textit{add}(\textit{cmd})$
-\EndFor
-\Return $\textit{new\_flow\_mod\_cmds}$
-% \State \emph{\bf EndFunction}
-\end{algorithmic}
-\end{algorithm}
-}
-
-\vspace{1mm}
-\noindent\textbf{Load balancing.}
+Let us now revisit our load-balancing application
+from the beginning of the paper.
 A simple strategy to balance load in our 2-link example
@@
 
-Our mechanism can also be used as a generic template to
+As we show in the following, our mechanisms can be used 
+to render such algorithms concurrent.
+In particular, we can implement a
+generic template to
 implement concurrent policy compositions of a user-specific
@@
 
-Let us first  consider a simple policy update protocol that
+Let us consider a simple policy update protocol that
 allows a set of  controllers to concurrently update switch policies, i.e., sets of
@@
 
-%Note that in Algorithm~\ref{alg:simple-update}, policies identifiers
-%grow without bound.
-%However, as was observed in~\cite{stn}, the number of currently
-%used identifiers can be made proportional to the number of concurrent
-%controllers.
-%
-\hide{
-It takes a set of rules $U$ (not yet a policy)and proceeds as follows: first, we seek to
- obtain a unique \emph{id}. FIXME: to be continued...
- \textbf{LS: I am not sure if I need to tell every step of the story or maybe it best to explain the main dif from the previous one similarly to what I just wrote next}.
-}
-%
-
-%Algorithm~\ref{alg:update} uses our id-claimer abstraction
-%to bound the number of policy identifiers.
-%Claiming a policy id prevents another controller from using it with a
-%different policy.
-%A controller claims the current policy identifier, chooses a new id
-%from a finite set of identifiers,
-%and then checks that the new policy id is not claimed by another controller.
-%Note that in Algorithm \ref{alg:update}, the meta-configuration is
-%used to store $\textit{pid}$ and id-claimer.
-%\petr{The description is not very clear.}
-
 Algorithm~\ref{alg:update} is a more efficient alternative 
@@
 different policy (Lines~\ref{update:read-config1}-\ref{update:read-config2}).
-%Then it reads the configuration again in case the policy was changed just before claiming its id.
-%Then considering other claims the controller chooses
 Then the controller claims the first \emph{unused} id and computes the update commands to be executed (Line~\ref{update:compute}).
@@
 the policy has not been concurrently modified, and then updates the policy.
-%[[PK added a note to the id-claimer description
-%Note that in Algorithm \ref{alg:update}, the meta-configuration is
-%used to store $\textit{pid}$ and id-claimer (the claims).
 

### 2015-09-01 ccr-rev1/inband-sync.tex
blabla

@@
 %\noindent\textbf{Organization}
-The remainder of the paper is organized as follows.
-In Section~\ref{sec:motivation}, we motivate the 
-transactional approach.
-We present our abstractions in more detail in Section~\ref{sec:main}.
-In Section~\ref{sec:background}, we discuss
-an OpenFlow implementation, 
-and we 
-discuss use cases in Section~\ref{sec:apps}.
-We
-conclude our paper
-in Section~\ref{sec:conclusion}.
+%The remainder of the paper is organized as follows.
+%In Section~\ref{sec:motivation}, we motivate the 
+%transactional approach.
+%We present our abstractions in more detail in Section~\ref{sec:main}.
+%In Section~\ref{sec:background}, we discuss
+%an OpenFlow implementation, 
+%and we 
+%discuss use cases in Section~\ref{sec:apps}.
+%We
+%conclude our paper
+%in Section~\ref{sec:conclusion}.
 
@@
 %[[PK put the example back in response of review 2
-For example, \texttt{ipv4\_src=10.1.14.0, ff.ff.ff.f0} matches all packets
-with source IP addresses in the range \texttt{10.1.14.0}--\texttt{10.1.14.127}.
+For example, \texttt{ipv4\_src=10.1.14.0/ff.ff.ff.f0} matches all packets
+with source IP addresses in the range \texttt{10.1.14.0}--\texttt{10.1.14.15}.
 %]]
@@
 In order to avoid inconsistencies caused by different rules with
-overlapping match fields, a \emph{FlowMod} command can be equipped with a the $\checko$ flag:
+overlapping match fields, a \emph{FlowMod} command can be equipped with 
+the $\checko$ flag:
 if the switch maintains a flow entry with an overlapping but not
-identical match part with the same priority but a different action,
- and if the two rules have the same priority, then
+identical match part with the same priority, then
 \emph{FlowMod} will fail.
@@
 \end{algorithm}
-
-\subsection{Proof-of-Concept}
-
-To investigate the feasibility of our approach,
-we implemented a proof-of-concept of our algorithms
-as a python library which wraps ovs-ofctl, 
-a well-known command line tool for monitoring and administering OpenFlow switches. 
-The code is made available together with this paper.
-
-We experimented with the low-level abstractions as well as the 
-transactional policy update algorithm of our proof-of-concept, using
-different workloads. 
-The tests were conducted using a software switch, Open vSwitch (version 2.4.0) and virtual network simulator, Mininet.
-
-For instance, we experimented with a setting where three processes
-(controllers) issue thousands of concurrent policy updates
-to the switch, where 
-each update consists in replacing a single rule with another one.
-When verifying the transactions, we could confirm that 
-despite the concurrency, all transactions have been successfully completed.
-Of course, due to the concurrency, multiple transactions also needed to abort,
-due to conflicts.
-
-While our framework can be used as synchronization tool itself, we 
-currently also work on implementing it as an integral part of existing controllers 
-such as Ryu.
 
@@
 
+\section{Proof-of-Concept}
+
+To investigate the feasibility of our approach,
+we implemented a proof-of-concept of our algorithms
+as a python library which wraps ovs-ofctl, 
+a well-known command line tool for monitoring and administering OpenFlow switches. 
+The code is made available together with this paper.
+
+We experimented with the low-level abstractions as well as the 
+transactional policy update algorithm of our proof-of-concept, using
+different workloads. 
+The tests were conducted using a software switch, Open vSwitch (version 2.4.0) and virtual network simulator, Mininet.
+
+For instance, we experimented with a setting where three processes
+(controllers) issue thousands of concurrent policy updates
+to the switch, where 
+each update consists in replacing a single rule with another one.
+When verifying the transactions, we could confirm that 
+despite the concurrency, all transactions have been successfully completed.
+Of course, due to the concurrency, multiple transactions also needed to abort,
+due to conflicts.
+
+While our framework can be used as synchronization tool itself, we 
+currently also work on implementing it as an integral part of existing controllers 
+such as Ryu.
+
 \section{Conclusion}\label{sec:conclusion}

### 2015-09-01 reviews/response-rev1-stefan.txt
asdf

--- 2015-08-31 reviews/response-rev1.txt
@@
 
-We revised our paper according to your suggestions.
+We revised our paper according to your suggestions. In particular, we re-structured the paper significantly to 
+better differentiate between the conceptual and the OpenFlow specific parts, and we emphasize more
+and earlier the aplications. We also performed some experiments of our transactional interface, and briefly report
+on the results. Essentially, the experiments confirmed the feasibility and correctness of our algorithms for the
+OpenFlow case: they showed that despite the high degree of concurrency, 100% of the transactions eventually complete
+in a consistent manner.
 
-In the following, we discuss the most important changes.
+In the following, we discuss the changes in more detail.
 

### 2015-09-01 ccr-rev1/inband-sync.tex
updated pseudo codes

@@
 \newcommand{\false}{\textit{false}}
+
+\newcommand{\ones}[1]{1^{#1}}
+\newcommand{\zeros}[1]{0^{#1}}
+\newcommand{\claimmagic}{\textit{CLAIM}}
+\newcommand{\memmagic}{\textit{WRITE}}
 
@@
 use the notation $a\concat b$ to represent a concatenation of two
-$32$bit integers $a$ and $b$: $a\concat b := (a<<32)+b$.
-We also write \texttt{1$\ldots$1} for  a $32$bit
-string of $1$'s used in a mask.
+ integers $a$ and $b$: $a\concat b := (a<<s)+b$, where $s$ is the size of $b$ in bits.
+We also write \texttt{$\ones{s}$} (\texttt{$\zeros{s}$}) for  a $s$ bit 
+string of $1$'s ($0$'s).
 %
@@
 In our implementation, the $64$bit meta-data field alone is sufficient.
-We sometimes use the concatenation of two $32$bit integers, denoted by $a\concat b$, to form one match value or mask, where $a\concat b := (a<<32)+b$.
+We sometimes use the concatenation of two integers, denoted by $a\concat b$, to form one match value or mask, where $a\concat b := (a<<s)+b$ and $s$ is the size of $b$ in bits.
 
@@
 the match part of the constructed commands represents a concatenation
-of two claimed identifiers.
-In addition, every controller uses a unique mask part of the match
-defined as a distinct $32$bit controller identifier concatenated with
-\texttt{1$\ldots$1}.
+of claimed identifier with a distinct $16$bit controller identifier.
+%In addition, every controller uses a unique mask part of the match
+%defined as a distinct $32$bit controller identifier concatenated with
+%\texttt{1$\ldots$1}.
 This allows multiple controllers to claim the same identifier without
@@
 For example, for any two resource identifiers $x_1$ and $x_2$ and controller identifiers $c_1\neq c_2$,
-the match patterns $m_1=(x_1\concat x_1, c_1\concat \texttt{1$\ldots$1})$ and $ m_2=(x_2\concat x_2, c_2\concat \texttt{1$\ldots$1})$ are distinct,
+the match patterns $m_1=(c_1\concat x_1, \ones{48})$ and $ m_2=(c_2\concat x_2, \ones{48})$ are distinct,
 even if $x_1=x_2$.
@@
 a new entry setting the flag \texttt{\checko}, and using a different
-action ($2$ instead of $1$),
-thereby inflicting a failure in case an entry with overlapping match value exists.
+match where the controller id is shifted $16$ bits to the left,
+thereby inflicting a failure in case an entry with the same claimed value exists.
 If the check succeeds, we delete the entry.
-Considering $m_1$ and $m_2$ from the example above,
-with respective actions $a_1=1$ and $a_2=2$,  they overlap
-with different actions iff $x_1=x_2$. Therefore, if $m_1$ is used in the \claimcheck after $m_2$ was used in the claim operation,
-a conflict is detected iff $x_1=x_2$, as desired.
+In any case since the check entry contains the controller id, it doesn't conflicts with other check entries and there is no negative impact of delaying the deletion.
+%Considering $m_1$ and $m_2$ from the example above,
+%with respective actions $a_1=1$ and $a_2=2$,  they overlap
+%with different actions iff $x_1=x_2$. Therefore, if $m_1$ is used in the \claimcheck after $m_2$ was used in the claim operation,
+%a conflict is detected iff $x_1=x_2$, as desired.
 
@@
 sufficient to go over the rules matching every controller identifier.
+In order to distinguish between claim entries and other policy entries we set the action part to be some magic value, $\claimmagic$.
 
@@
 %    \Ensure
-    		\State $\textit{value} \gets x\concat x$
-    		\State $\textit{mask} \gets \textit{self}\concat \texttt{1$\ldots$1}$
+    		\State $\textit{value} \gets self\concat x$
+    		\State $\textit{mask} \gets \ones{48}$
 	    	\State $\textit{match} \gets (\textit{value},\textit{mask})$
     		%\State $cookie \gets x\concat self$
-    		\State $\textit{action} \gets 1$
+    		\State $\textit{action} \gets \claimmagic$
     		\State $\textit{flag} \gets 0$
@@
 %    \Ensure
-    		\State $\textit{value} \gets x\concat x$
-    		\State $\textit{mask} \gets self\concat \texttt{1$\ldots$1}$
+    		\State $\textit{value} \gets self\concat \ones{16}\concat x$
+    		\State $\textit{mask} \gets \ones{16}\concat \zeros{16}\concat \ones{32}$
     		\State $\textit{match} \gets (value,mask)$
     		%\State $cookie \gets x\concat self$
-    		\State $\textit{action} \gets 2$
+    		\State $\textit{action} \gets \claimmagic$
     		\State $\textit{flag} \gets \texttt{ofpff\_\checko}$
@@
 %    \Ensure
-    		\State $\textit{value} \gets x\concat x$
-    		\State $\textit{mask} \gets self\concat \texttt{1$\ldots$1}$
+    		\State $\textit{value} \gets self\concat x$
+    		\State $\textit{mask} \gets \ones{48}$
     		%\State $cookie \gets x\concat self$
@@
  As can be seen in Algorithms \ref{alg:write}-\ref{alg:compare},
-  the match part of the added flow entry is used as the memory address and the action as
-  the written value. Values written to the same address will replace one another,
-following the command definition to replace the old flow entry with a new one in case they
-share the same match parts.
+  the value part of the added flow entry match contains the memory address mask part contains
+  the written value. In order to rewrite a memory location, the old entry needs to be deleted and a new one need to be added atomically.
+  %Values written to the same address will replace one another,
+%following the command definition to replace the old flow entry with a new one in case they
+%share the same match parts.
 
@@
 but we also set the \texttt{\checko} flag. If the value is not there,
-our action differs from the action of the existing entry,
+our action differs from the existing entry, but still overlaps with it since only their masks differ,
 thereby inflicting a failure.
@@
 %    \Ensure
-    		\State $value \gets \addr$
-    		\State $mask \gets  \texttt{1$\ldots$1}. \texttt{1$\ldots$1}$
+    		\State $value \gets \ones{32}\concat\addr$
+    		\State $mask \gets  \ones{32}$
+    		\State $match \gets (value,mask)$
+    		\State $\textit{cmd1}\gets \textit{FlowMod}(\textit{match}, op = \dele) $
+
+  		   \State $value \gets \ones{32}\concat\addr$
+  		   \State $mask \gets  k\concat\ones{32}$
   		   \State $match \gets (value,mask)$
  		    		%\State $cookie \gets x\concat self$
-    		\State $action \gets k$
+ 		   \State $action \gets \memmagic$
  		   \State $flag \gets 0$
-    		\State $cmd\gets \emph{FlowMod}(match, op = \add, flag, action) $
-			\Return $cmd$
+		   \State $\textit{cmd2}\gets \textit{FlowMod}(\textit{match}, op = \add, \textit{flag}, \textit{action}) $
+
+			\Return $\textit{cmd1}$,$\textit{cmd2}$
     \end{algorithmic}
@@
 %    \Ensure
-    		\State $value \gets \addr$
-    		\State $mask \gets  \texttt{1$\ldots$1}.\texttt{1$\ldots$1}$
+    		\State $value \gets \ones{32}\concat\addr$
+    		\State $mask \gets  k\concat\ones{32}$
     		\State $match \gets (value,mask)$
     		%\State $cookie \gets x\concat self$
-    		\State $action \gets k$
+    		\State $action \gets \memmagic$
     		\State $flag \gets \texttt{ofpff\_\checko}$

### 2015-09-01 ccr-rev1/inband-sync.tex
small fix

@@
 \newcommand{\zeros}[1]{0^{#1}}
-\newcommand{\claimmagic}{\textit{CLAIM}}
-\newcommand{\memmagic}{\textit{WRITE}}
+\newcommand{\claimmagic}{\textit{CLAIM\_MAGIC}}
+\newcommand{\memmagic}{\textit{WRITE\_MAGIC}}
 
@@
  As can be seen in Algorithms \ref{alg:write}-\ref{alg:compare},
-  the value part of the added flow entry match contains the memory address mask part contains
+  the value part of the added flow entry match contains the memory address and the mask part contains
   the written value. In order to rewrite a memory location, the old entry needs to be deleted and a new one need to be added atomically.
@@
 %share the same match parts.
+Similarly to claim entries, we set the action to a magic value, $\memmagic$, in order to distinguish them from policy entries.
 

### 2015-09-01 reviews/response-rev1-stefan.txt
dsaf

@@
 into a broader context. Having that said, in our revision we also report
-in more detail on our proof-of-concept implementation based on POX.
+in more detail on our proof-of-concept implementation.
 
@@
 
-FIXME
+Clarified.
 
@@
 In our revision, we restructured these parts and hope we found
-a omre intuitive way to put this into perspective.
+a more intuitive way to put this into perspective.
 

### 2015-09-01 ccr-rev1/inband-sync.tex
dsaf

@@
 switch itself
-and using
-the data plane configuration as a shared memory.
+and using part of
+the data plane configuration space as a shared memory.
 We argue that this in-band approach is attractive as 
@@
 This allows us 
-to realize fundamental consensus primitives in the presence
-of controller failures, and we discuss applications
+to implement fundamental consensus primitives in the presence
+of controller failures, and we discuss their applications
 for 
@@
 out-of-band network, in the presence of asynchrony and failures.
-Indeed, the inherent costs  of such protocols are often considered too high, both in
+Indeed, the inherent costs  of such conventional \emph{out-of-band} 
+protocols are often considered too high, both in
 terms of the necessary computability assumptions about the underlying
@@
 
-This paper proposes mechanisms to solve such more general 
+This paper proposes mechanisms to solve these and more general 
 synchronization problems.
@@
 based on application-specific conditions, which depend on the
-current configuration,
-However, as we have for example seen in our load-balancing
+current configuration.
+However, as we have seen in our load-balancing
 example, a support for more generic conditions under which configuration
@@
 If the check succeeds, we delete the entry.
-In any case since the check entry contains the controller id, it doesn't conflicts with other check entries and there is no negative impact of delaying the deletion.
+In any case since the check entry contains the controller id, it does not conflict with other check entries and there is no negative impact of delaying the deletion.
 %Considering $m_1$ and $m_2$ from the example above,
@@
 A simple strategy to balance load in our 2-link example
-is given in Algorithm~\ref{alg:naive-lb}.
-However, this algorithm fails in the presence of
+ fails in the presence of
 multiple controllers, even if 
@@
 Algorithm~\ref{alg:simple-update} can therefore also be used
-to render our na\"ive load-balancing Algorithm~\ref{alg:naive-lb}
+to render a na\"ive, single-controller load-balancing algorithm
 consistent, without any modifications in the update function.
@@
 different workloads. 
-The tests were conducted using a software switch, Open vSwitch (version 2.4.0) and virtual network simulator, Mininet.
+The tests were conducted using a software switch, Open vSwitch (version 2.4.0) and
+a virtual network simulator, Mininet.
 
@@
 When verifying the transactions, we could confirm that 
-despite the concurrency, all transactions have been successfully completed.
+despite the concurrency, 100 \% of the transactions have been successfully completed.
 Of course, due to the concurrency, multiple transactions also needed to abort,

### 2015-11-02 reviews/response-rev1-stefan.txt
ccr reviews

--- 2015-09-01 reviews/response-rev1-stefan.txt
@@
 we aim to put the fundamental aspects and the benefit of inband synchronization
-into a broader context. Having that said, in our revision we also report
-in more detail on our proof-of-concept implementation based on POX.
+into a broader context, and restructured the article accordingly. Having that said, in our revision we also report
+in more detail on our proof-of-concept implementation and some preliminary experiments.
 
@@
 further research is needed to fully answer the out-of-band vs inband
-debate.
+debate. In the revision, we tried to better define the scope and discuss the limitations of our work.
 

### 2015-11-07 artefacts/CCR-camera-ready.tex
cr prepared

--- 2015-09-01 ccr-rev1/inband-sync.tex
@@
 \newcommand{\paddr}{\textit{paddr}\xspace}
-\newcommand{\pid}{\textit{pid}\xspace}
+\newcommand{\pid}{\textit{pol-id}\xspace}
 \newcommand{\add}{\textbf{add}\xspace}
@@
 \newcommand{\exec}{\textbf{execute}}
-\newcommand{\execatomic}{\textbf{execute-atomic}}
+\newcommand{\execatomic}{\textbf{execute-transaction}}
 \newcommand{\ack}{\textit{ack}}
@@
 %\title{On the Synchronization Power of the OpenFlow Data Plane}
-\title{In-Band Synchronization for\\Distributed SDN Control Planes}
+\title{In-Band Synchronization for\\Distributed SDN Control Planes\thanks{Supported by the
+European Institute of Innovation \& Technology (EIT) project 13153 (Software Defined Networking) 
+and the FP7 EU project UNIFY.}}
 %\\ {\Large Towards Atomic Transactions In-Band}}
@@
 Stefan Schmid$^2$ \quad Petr Kuznetsov$^3$ \\
-\small $^1$ Tel Aviv University, Israel \quad $^2$ TU Berlin \& T-Labs,
-Germany \quad $^3$ T\'el\'ecom ParisTech, France
+\small $^1$ Tel Aviv University, Israel \quad $^2$ Aalborg University,
+Denmark \quad $^3$ T\'el\'ecom ParisTech, France
 }
@@
 \emph{transactions}.
-In particular, we present a powerful interface 
-to implement transactions \emph{in-band}, on the 
-switch itself
-and using part of
-the data plane configuration space as a shared memory.
+In particular, we make the case for 
+implementing transactions \emph{in-band}, on the 
+switch itself.
 We argue that this in-band approach is attractive as 
@@
 out-of-band coordination mechanisms.
-This allows us 
-to implement fundamental consensus primitives in the presence
+It allows us to realize fundamental consensus primitives in the presence
 of controller failures, and we discuss their applications
@@
 Interestingly, 
-by leveraging the match-action
+using part of
+the data plane configuration space as a shared memory
+and by leveraging 
+ the match-action
 paradigm to realize simple key-value stores, 
 we can implement our synchronization framework in
-today's standard Openflow protocol, 
+today's standard OpenFlow protocol, 
 and we report on
@@
 
-\begin{comment}
 \category{C.2.2}{Computer-Communication Networks}{Network Protocols}
@@
 
-%\terms{Algorithms}
+\terms{Networks}
 
 \keywords{Software-Defined Networking, Distributed Control Planes, In-band Mechanisms}
-
-\end{comment}
 
@@
 the distributed controllers and manipulating the network state \emph{consistently}
-are non-trivial tasks.~\cite{sharon}
+are non-trivial tasks~\cite{sharon}.
 
@@
 consider the standard \emph{OpenFlow}
- protocol (version 1.4~\cite{openflow}).
+ protocol (version 1.4~\cite{openflow}). \stefan{liron can we say that
+ for example novikit supports all these features?}
  In particular, we show that OpenFlow's match-action paradigm can be instrumented
@@
 the missing link for the read-modify-write object
-postulated in~\cite{stn}. %\petr{We do not seem to discuss the latter?}
+postulated in CPC and STN~\cite{stn,infocom15}. %\petr{We do not seem to discuss the latter?}
 
@@
 We will later discuss later how this can be achieved, for example in 
-today's Openflow.
+today's OpenFlow.
 
@@
 configuration is changed and the current policy identifier is updated
-only if the check operation \emph{succeeds} (does not abort).
+only if the check operation \emph{succeeds} (i.e., does not abort).
 
@@
 the OpenFlow \emph{add flow monitor} command with the \textsf{ofpt\_initial} optional flag.
-In this paper, we write \textbf{read-config} for sending
+In this paper, we write \textbf{read-state} for sending
 this message, receiving a response, and returning the received configuration.
@@
 
-By employing the \textbf{read-config} command, we may easily implement
+By employing the \textbf{read-state} command, we may easily implement
 an additional operation that returns all currently claimed ids: it is
@@
 \end{algorithm}
+
+\vspace{1mm}
+\noindent\textbf{Note on efficiency.}
+
+\stefan{We conclude by observing that our implementation is switch space efficient,
+in the sense that it requires only a constant number of flow entries per memory address
+and resources scale linearly in the number of controllers. Also, in case of controller failures,
+at most a constant amount of space is lost.}
+
+\stefan{We also note that while we described our scheme in a simplified
+way where numbers continuously increase, it is straight-forward to use a modulo
+scheme. While in our case 32-bit numbers are used, 
+the flow table width (number of bits which can be matched in parallel)
+of modern switches is 40-bits and more.}
 
@@
  		\Repeat
- 			\State $\pid,\textit{policy}\gets$ \textbf{read-config} %switch configuration}
+ 			\State $\pid,\textit{policy}\gets$ \textbf{read-state} %switch configuration}
  			\State $\textit{update\_cmds}\gets \textsc{\ufunc}(policy)$
@@
 for a
-different policy (Lines~\ref{update:read-config1}-\ref{update:read-config2}).
+different policy (Lines~\ref{update:read-state1}-\ref{update:read-state2}).
 Then the controller claims the first \emph{unused} id and computes the update commands to be executed (Line~\ref{update:compute}).
@@
  		\Repeat
-		 	\State $\pid,claims,policy\gets$ \textbf{read-config} \label{update:read-config1} %switch configuration
+		 	\State $\pid,claims,policy\gets$ \textbf{read-state} \label{update:read-state1} %switch configuration
  			\State $\textbf{execute}\{claim(\pid)\}$
- 			\State $\pid 2,\textit{claims},\textit{policy}\gets$ \textbf{read-config} \label{update:read-config2} %switch configuration
+ 			\State $\pid 2,\textit{claims},\textit{policy}\gets$ \textbf{read-state} \label{update:read-state2} %switch configuration
  			\If {$\pid\neq \pid 2$}
@@
 a well-known command line tool for monitoring and administering OpenFlow switches. 
-The code is made available together with this paper.
+\stefan{The code is made publicly available at TODO LIRON.}
 
@@
 
-This paper initiated the study of mechanisms which allow distributed controllers
-to synchronize their possibly concurrent actions. We believe that our centralized in-band
-approach is natural and attractive, as it avoids the
-complexities and limitations of distributed out-of-band agreement protocols, which may also
+%This paper initiated the study of mechanisms which allow distributed controllers
+%to synchronize their possibly concurrent actions.
+
+We believe that our centralized in-band
+synchronization approach is natural and attractive, as it avoids the
+complexities and limitations of distributed out-of-band agreement protocols, which may 
 introduce additional points of potential failures.
-Moreover, our mechanisms are simple and can be implemented using the \emph{standard OpenFlow}
+Moreover, the switch is the element implementing the
+controller policies anyway.
+Interestingly, the mechanisms presented in this paper are simple and can be implemented using the \emph{standard OpenFlow}
 protocol: they do not require any protocol/hardware extensions as postulated in recent literature~\cite{stn,netpaxos}.
@@
 as well as to useful high-level concurrency objects and
-language abstractions.~\cite{pyretic}
+language abstractions~\cite{pyretic}.
 In general, we understand our paper as a first step, and more

### 2015-11-10 ccr-camera/inband-sync.tex
starting camera-ready
addressed comments and added my own

--- 2015-09-01 ccr-rev1/inband-sync.tex
@@
 \definecolor{mygrey}{rgb}{0.5,0.5,0.5}
+\definecolor{mypurple}{rgb}{1,0,1}
 
@@
 \newcommand{\stefan}[1]{\textit{\textcolor{red}{[stefan]: #1}}} % Comments
-\newcommand{\liron}[1]{\textit{\textcolor{mygreen}{[liron]: #1}}} % Comments
+\newcommand{\liron}[1]{\textit{\textcolor{mypurple}{[liron]: #1}}} % Comments
 \newcommand{\petr}[1]{\textit{\textcolor{blue}{[petr]: #1}}} % Comments
@@
 controllers to solve fundamental agreement tasks in just one message
-exchange with the data plane, tolerating asynchrony and failures of
+exchange \liron{--------why one? ---------}  with the data plane, tolerating asynchrony and failures of
 any number of controllers.
@@
 independent flows to the physical infrastructure, which is important,
-e.g., for load-balancing applications.
+e.g., for load-balancing applications. \liron{I think this is a bit hard to understand}
 
@@
 consider the standard \emph{OpenFlow}
- protocol (version 1.4~\cite{openflow}).
+ protocol (version 1.4~\cite{openflow}). \stefan{liron can we say that
+ for example novikit supports all these features?}\liron{No. Novikit support OF 1.3 and therefore lacks support for bundles.}
  In particular, we show that OpenFlow's match-action paradigm can be instrumented
@@
  With synchronization, i.e., by bundling the flow installation with a Compare-and-Set (CAS) primitive
- depending on the policy id (\emph{pid}), this problem is avoided.}\label{fig:CAS-example}
+ depending on the policy id (\emph{\pid}), this problem is avoided.}\label{fig:CAS-example}
 \end{figure}
@@
 For example, for any two resource identifiers $x_1$ and $x_2$ and controller identifiers $c_1\neq c_2$,
-the match patterns $m_1=(c_1\concat x_1, \ones{48})$ and $ m_2=(c_2\concat x_2, \ones{48})$ are distinct,
+the match patterns
+%$m_1=(c_1\concat x_1, \ones{48})$ and $ m_2=(c_2\concat x_2, \ones{48})$ 
+$m_1=(*^{16}\concat c_1\concat x_1)$ and $ m_2=(*^{16}\concat c_2\concat x_2)$ 
+are distinct,
 even if $x_1=x_2$.
@@
 a new entry setting the flag \texttt{\checko}, and using a different
-match where the controller id is shifted $16$ bits to the left,
+match, $m$, where the controller id, $c$, is shifted $16$ bits to the left, i.e., $m=( c_1\concat *^{16}\concat x_1)$,
 thereby inflicting a failure in case an entry with the same claimed value exists.
@@
 
-By employing the \textbf{read-config} command, we may easily implement
+By employing the \textbf{read-state} command, we may easily implement
 an additional operation that returns all currently claimed ids: it is
 sufficient to go over the rules matching every controller identifier.
-In order to distinguish between claim entries and other policy entries we set the action part to be some magic value, $\claimmagic$.
+In order to distinguish between claim entries and other policy entries, returned by  \textbf{read-state}, we set the action part to be some magic value, $\claimmagic$.
 
@@
 %    \Ensure
-    		\State $\textit{value} \gets self\concat x$
-    		\State $\textit{mask} \gets \ones{48}$
+    		\State $\textit{value} \gets \zeros{16}\concat self\concat x$
+    		\State $\textit{mask} \gets \zeros{16}\concat\ones{16}\concat\ones{32}$
 	    	\State $\textit{match} \gets (\textit{value},\textit{mask})$
@@
     		\State $\textit{action} \gets \claimmagic$
-    		\State $\textit{flag} \gets 0$
-    		\State $\textit{cmd}\gets \textit{FlowMod}(match, op = \add, \textit{flag}, \textit{action}) $
+    		%\State $\textit{flag} \gets 0$
+    		\State $\textit{cmd}\gets \textit{FlowMod}(match, op = \add, \textit{action}) $
 			\Return $\textit{cmd}$
@@
 %    \Ensure
-    		\State $\textit{value} \gets self\concat x$
-    		\State $\textit{mask} \gets \ones{48}$
+    		\State $\textit{value} \gets \zeros{16}\concat self\concat x$
+    		\State $\textit{mask} \gets \zeros{16}\concat\ones{16}\concat\ones{32}$
     		%\State $cookie \gets x\concat self$
@@
  operations can be implemented using similar techniques.
+ \liron{
  As can be seen in Algorithms \ref{alg:write}-\ref{alg:compare},
-  the value part of the added flow entry match contains the memory address and the mask part contains
-  the written value. In order to rewrite a memory location, the old entry needs to be deleted and a new one need to be added atomically.
+  the value part of the added flow entry match contains the memory address, $addr$, and the mask part contains
+  the written value $k$. More specifically, we use a ternary match string of the form $k'\concat \addr$, where $k'$ is result os replacing every zero bit in $k$ with a wildcard ($*$).}
+  In order to rewrite a memory location, the old entry needs to be deleted and a new one need to be added atomically.
   %Values written to the same address will replace one another,
@@
 %share the same match parts.
-Similarly to claim entries, we set the action to a magic value, $\memmagic$, in order to distinguish them from policy entries.
+Similarly to claim entries, we set the action to a magic value, $\memmagic$, in order to distinguish them from policy entries returned by \textbf{read-state}.
 
@@
 but we also set the \texttt{\checko} flag. If the value is not there,
-our action differs from the existing entry, but still overlaps with it since only their masks differ,
+our action differs from the existing entry, but still overlaps with it since only their masks differ \liron{(different mixtures of ones and wildcards in the $k'$ part)},
 thereby inflicting a failure.
@@
 with an identical one.
+
+\liron{Note that comparing an uninitialized address always succeeds and also writes to it the compared value.
+Initializing required addresses can be safely performed by an a-priori comparison with zero.}
 
@@
     		\State $value \gets \ones{32}\concat\addr$
-    		\State $mask \gets  \ones{32}$
+    		\State $mask \gets  \zeros{32}\concat\ones{32}$
     		\State $match \gets (value,mask)$
@@
  		   \State $action \gets \memmagic$
- 		   \State $flag \gets 0$
-		   \State $\textit{cmd2}\gets \textit{FlowMod}(\textit{match}, op = \add, \textit{flag}, \textit{action}) $
+ 		   %\State $flag \gets 0$
+		   \State $\textit{cmd2}\gets \textit{FlowMod}(\textit{match}, op = \add,  \textit{action}) $
 
@@
 
+\vspace{1mm}
+\noindent\textbf{Note on efficiency.}
+
+\stefan{We conclude by observing that our implementation is switch space efficient,
+in the sense that it requires only a constant number of flow entries per memory address
+and resources scale linearly in the number of controllers. Also, in case of controller failures,
+at most a constant amount of space is lost.}
+
+\stefan{We also note that while we described our scheme in a simplified
+way where numbers continuously increase, it is straight-forward to use a modulo
+scheme. While in our case 32-bit numbers are used, 
+the flow table width (number of bits which can be matched in parallel)
+of modern switches is 40-bits and more.}
+ \liron{I think we should say something like: "... of modern switches can get as high as a few hundreds'. }
 \section{Applications}\label{sec:apps}
@@
 the switch configuration, where the configuration contains both the
-policy and a meta-configuration memory address $\paddr$ that holds the \emph{policy identifier (pid)} .
+policy and a meta-configuration memory address $\paddr$ that holds the \emph{policy identifier (\pid)} .
 In the algorithm, the controller first reads the currently installed
@@
 % and use them in order to make sure that the new policy, with id
-%$pid+1$, replaces the (what considered as) current policy, with id
-%$pid$.
+%$\pid+1$, replaces the (what considered as) current policy, with id
+%$\pid$.
 %Considering the policies identifiers, we send the policy update
 %commands as part of a bundle that also tries to perform
-%$CAS(pid,pid+1)$, thereby insuring that the commands are
+%$CAS(\pid,\pid+1)$, thereby insuring that the commands are
 %performed iff the current policy is still there and preventing other
@@
 a well-known command line tool for monitoring and administering OpenFlow switches. 
-The code is made available together with this paper.
+\stefan{The code is made publicly available at TODO LIRON.}\liron{https://github.com/lironsc/of-sync-lib}
 

### 2015-11-15 ccr-camera/inband-sync.tex
first update (still holding locks)

@@
 %\DeclareGraphicsRule{*}{mps}{*}{<++>}
+
+%to balance columns on the last page
+\usepackage{balance}
 
@@
 In this paper, we propose a synchronization
-framework for control planes  based on the well-known concept of
-\emph{transactions}.
-In particular, we make the case for 
-implementing transactions \emph{in-band}, on the 
-switch itself.
+framework for control planes  based on atomic
+transactions, 
+%In particular, we make the case for 
+implemented \emph{in-band}, on the 
+data-plane switches.
 We argue that this in-band approach is attractive as 
@@
 of controller failures, and we discuss their applications
-for 
- consistent policy composition and fault-tolerant control-planes.
+for consistent policy composition and fault-tolerant control-planes.
 Interestingly, 
-using part of
+by using part of
 the data plane configuration space as a shared memory
-and by leveraging 
- the match-action
-paradigm to realize simple key-value stores, 
+and leveraging  the match-action
+paradigm, 
+%[[PK key-valued stores are not really discussed?
+%to realize simple key-value stores, 
 we can implement our synchronization framework in
@@
 and we report on
-a proof-of-concept accordingly.
+our proof-of-concept implementation.
 \end{abstract}
@@
 
-By consolidating and outsourcing the control over the dataplane switches to a logically
+By consolidating and outsourcing the control over the data-plane switches to a logically
 centralized controller, Software-Defined Networks (SDNs)
@@
 to realize consistent distributed systems.
-In particular, the popular \emph{transactional
-  memory} abstraction~\cite{tm-book} provides a collection of
+In particular, the popular 
+%[[PK it is incorrect to only cite the book for the TM notion]]
+\emph{transactional  memory} abstraction~\cite{HM93,stm-st95,tm-book} 
+%]]
+provides a
+collection of
 concurrent processes with the ability of aggregating sequences of
@@
 We propose to implement our synchronization constructs \emph{in-band},
-on the switch.
+on the data-plane switch.
 An in-band implementation allows us to efficiently alleviate the problems related to
@@
 controllers to solve fundamental agreement tasks in just one message
-exchange \liron{--------why one? ---------}  with the data plane, tolerating asynchrony and failures of
+exchange 
+%[[PK we mean consensus/CAS here, where one message exchange is indeed used
+%\liron{--------why one? ---------}  
+%]]
+with the data plane, tolerating asynchrony and failures of
 any number of controllers.
 
-The key idea of our approach is to use the switch configuration space
+The key idea of our approach is to use the data-plane switch configuration space
 as a \emph{transactional shared memory}.
-The shared memory of the data-plane configuration space 
-can store information about  contention and conflicts
+%[[PK reads a bit strange
+%The shared memory of the data-plane configuration space  
+In addition to the actual data-plane configuration, the memory can store information about  contention and conflicts
 between the controllers.
+%]]
 A transaction can then contain standard control and update operations
@@
 memory.
-We investigate useful synchronization abstractions and propose 
-a simple interface which revolves around policy identifiers (IDs)
-which can be claimed and released atomically and hence consistently. 
-More generally, our synchronization primitives allow the controllers to define
+%[[PK policy IDs come out of the blue, very hard to digest 
+%We 
+%%investigate useful synchronization abstractions and 
+%propose 
+%a simple interface which revolves around network policy identifiers 
+%%[[PK IDs are not used below?
+%%(IDs)
+%%]]
+%which can be claimed and released atomically and hence consistently. 
+%[[PK simplified a bit
+%More generally, 
+The synchronization primitives allow the controllers to define
 general notions of conflicts between configuration updates,
-not restricted to simple conflicts on the overlapping flow spaces.
-In particular, we can define dependencies between mappings of
-independent flows to the physical infrastructure, which is important,
-e.g., for load-balancing applications. \liron{I think this is a bit hard to understand}
+covering simple routing conflicts 
+%conflicts on the overlapping flow spaces 
+as well as more sophisticated multi-flow dependencies appearing in load-balancing applications.   
+%In particular, we can define dependencies between mappings of
+%independent flows to the physical infrastructure, which is important,
+%e.g., for load-balancing applications. 
+%\liron{I think this i s a bit hard to understand}
+%]]
 
@@
 consider the standard \emph{OpenFlow}
- protocol (version 1.4~\cite{openflow}). \stefan{liron can we say that
- for example novikit supports all these features?}\liron{No. Novikit support OF 1.3 and therefore lacks support for bundles.}
- In particular, we show that OpenFlow's match-action paradigm can be instrumented
-to realize simple key-value stores in the configuration space.
-
-Using our abstraction, the controllers can
-solve several important synchronization problems.
+ protocol (version 1.4~\cite{openflow}). 
+%\stefan{liron can we say that
+% for example novikit supports all these features?}\liron{No. Novikit support OF 1.3 and therefore lacks support for bundles.}
+% In particular, 
+%PK ``key-value store'' come out of the blue...
+We show that OpenFlow's match-action paradigm can be instrumented
+to realize 
+%simple key-value stores in the configuration space.
+%
+%\petr{Did not get the flow here: what is ``our abstraction'' here?
+a simple transactional abstraction that allows 
+the controllers to solve several important synchronization problems.
 For example, we describe an OpenFlow implementation of the fundamental
@@
 The atomic CAS operation is known to provide an infinite consensus number,
-and can for instance be used to 
-implement generic replicated control plane services~\cite{Her91} in a consistent and
-fault-tolerant way: a mainstream building block in
-modern distributed system design (Chubby, Google Spanner, Amazon Webservices, etc.).
+and can, for instance, be used to 
+implement a generic consistent and fault-tolerant replicated 
+%[[PK why control-plane here: does not fit the Chubby et al. mentioned later.
+%control-plane 
+service~\cite{Her91} 
+%in a consistent and fault-tolerant way: 
+a mainstream building block in
+modern distributed systems (Chubby, Google Spanner, Amazon Webservices, etc.).
 We discuss a simple CAS-based concurrent policy-update mechanism,
@@
 \includegraphics[width=.56\columnwidth]{loadbal-1.pdf}~\hspace{-.7cm}~\includegraphics[width=.56\columnwidth]{loadbal-2.pdf}\\
-\caption{\emph{Left:} Without synchronization, the two controllers naturally choose to
+\caption{\small \emph{Left:} Without synchronization, the two controllers naturally choose to
 install their flows on the left link, which results in an undesirable
@@
 
-In order to synchronize controllers, we propose
-to use a minimal
-``meta-configuration'' space on the switch,
-to store information
+%[[PK how do we know ``minimal''?
+In order to synchronize controllers, 
+we propose to use a part of the configuration space of the data-plane
+switch for storing  a 
+%minimal
+``meta-data'', containing information
 on contention and conflicts between
-concurrently installed policy updates. Controllers can read and modify
+concurrently installed policy updates. 
+%]]
+Controllers can read and modify
 this meta-data by applying synchronization primitives.
@@
 the ``normal'' configuration which determines the network policy and
-is used to process the data plane traffic (e.g., forwarding rules),
+is used to process the data-plane traffic (e.g., forwarding rules),
 and the \emph{meta-configuration} that contains information used by the
@@
 $\exec\{op_1,\;\ldots\;op_k\}$ and $\execatomic\{op_1,\;\ldots\;op_k\}$
-which execute a sequence of commands $op_1,\ldots,op_k$ in
+which perform a sequence of commands $op_1,\ldots,op_k$ in
 a regular and atomic way, respectively.
-The construct $\execatomic\{op_1,\;\ldots\;op_k\}$  
+In particular, the construct $\execatomic\{op_1,\;\ldots\;op_k\}$  
 returns $\ack$ if and only if all commands in the sequence are performed successfully; otherwise
@@
 
-Therefore, we provide the following transactional operations.
+Therefore, we provide the following (transactional) operations.
 A $\textit{write}(\addr,k)$ operation sets the content at memory location
@@
 Intuitively, by combining compare and write with a set of regular
-switch update commands 
-in a transaction (such as FlowMod in OpenFlow)  we can make sure that the
+switch update commands (e.g., FlowMod in OpenFlow) 
+in a transaction  we can make sure that the
 configuration is changed and the current policy identifier is updated
@@
 However, as is,
-bundling has important limitations, as we will argue in the following.
+bundling has important limitations. %, as we will argue in the following.
 
@@
 overlapping flow spaces or insufficient free space),
-a bundle does not provide any means to modify the switch configuration
-based on application-specific conditions, which depend on the
-current configuration.
-However, as we have seen in our load-balancing
-example, a support for more generic conditions under which configuration
-updates
-should or should not take place, is essential.
-Moreover, it is sometimes desirable to react to conflicts in smarter ways than
-by simply aborting an operation, e.g., by supporting \emph{conditional modifications}.
-
-Nevertheless, we now show that OpenFlow can be used 
-to realize our transactional
-abstractions. Concretely, the OpenFlow match-action paradigm can be used
-to implement a key-value store.
-The OpenFlow
-configuration space can be realized using low-priority rules.
+the bundle construct \emph{per se} 
+does not provide any means to modify the switch configuration
+based on application-specific conditions.
+%[[PK overlapping flow spaces also depend on the switch configuration...
+%, which depend on the
+%current configuration.
+%]]
+But, as we have seen in our load-balancing
+example, supporting such conditions 
+%under which configuration updates should or should not take place, 
+might be essential.
+Second, it is sometimes desirable to react to conflicts in smarter ways than
+by simply aborting an operation, e.g., by providing \emph{conditional modifications}.
+
+%[[PK Why ``nevertheless''? 
+%Nevertheless, 
+We now show that the OpenFlow match-action paradigm 
+can be used  to realize a transactional synchronization abstraction
+with the desired functionality. 
+%Concretely, the OpenFlow match-action paradigm can be used
+%to implement a key-value store.
+%]]
+The 
+%OpenFlow 
+meta-configuration can be realized using low-priority rules.
 Alternatively, we can leverage the possibility to maintain multiple

### 2015-11-15 ccr-camera/inband-sync.tex
one more update (still working on it)

@@
 %[[PK it is incorrect to only cite the book for the TM notion]]
-\emph{transactional  memory} abstraction~\cite{HM93,stm-st95,tm-book} 
+\emph{transactional  memory} abstraction (cf. a survey in~\cite{tm-book}). 
 %]]
@@
 An in-band implementation allows us to efficiently alleviate the problems related to
-coordinating controllers via a control plane
+coordinating controllers via a control-plane
 out-of-band network, in the presence of asynchrony and failures.
@@
 
-\item If $k$ is currently claimed by any controller,
-$\textit{\claimcheck}(k)$ %, $k\in\Nat$,
-aborts the current transaction.
+\item With $\textit{\claimcheck}(k)$, the controller checks if $k$ is
+  currently claimed by any controller, and if so, the current transaction is
+  aborted.
 
@@
 In the simplest setting, a switch maintains one flow table (table 0).
-A data plane packet arriving at a switch is first checked against
+A data-fplane packet arriving at a switch is first checked against
 the rule with the highest priority and in table $0$.
@@
 forming a \emph{pipeline}.
-Data plane packets are checked against flow entries in the pipelined
+Data-plane packets are checked against flow entries in the pipelined
 tables to determine the set of actions the switch should perform on the packet.
-A data plane packet arriving at a switch is first checked against
+A data-plane packet arriving at a switch is first checked against
 the rule with the highest priority and in table $0$.
@@
 %that should be applied to a packet matching the two entries.
+%[[PK shortened
 In order to avoid inconsistencies caused by different rules with
 overlapping match fields, a \emph{FlowMod} command can be equipped with 
-the $\checko$ flag:
+the $\checko$ flag which will be instrumental
+in implementing our synchronization abstractions:
 if the switch maintains a flow entry with an overlapping but not
 identical match part with the same priority, then
-\emph{FlowMod} will fail.
-The $\checko$ flag in a \emph{FlowMod} command helps resolving simple
-conflicts between controllers, and will be instrumental
-in implementing our synchronization abstractions.
+the \emph{FlowMod} command will fail.
+%The $\checko$ flag in a \emph{FlowMod} command helps resolving simple
+%conflicts between controllers, and will be instrumental
+%in implementing our synchronization abstractions.
+%]]
 
@@
 this message, receiving a response, and returning the received configuration.
+\petr{Is it atomic? Should we say that the command returns a ``snapshot'' of the
+  configuration?}
 
@@
 The error message contains $\xid$, the identifier of the first failed command in the
-bundle, and $\ecode$, the \emph{error code} corresponding to the failure.
+bundle, and $\ecode$, the identifier of the failure.
 
@@
 %When properly configured,
-The bundle execution is considered \emph{controller-atomic}
+The bundle execution is \emph{controller-atomic}
 in the sense that all controllers can only observe switch
@@
 making bundles \emph{packet-atomic},
-in the sense that every data plane packet is processed by a configuration
+in the sense that every data-plane packet is processed by a configuration
 before or after the bundle, and not by a configuration resulting from
@@
 sufficient to go over the rules matching every controller identifier.
-In order to distinguish between claim entries and other policy entries, returned by  \textbf{read-state}, we set the action part to be some magic value, $\claimmagic$.
-
-{\small
+In order to distinguish between claim entries and other policy
+entries, returned by  \textbf{read-state},
+%[[PK ``some magic value'' sounds magic... 
+we set the action part to
+be a unique designated value, $\claimmagic$.
+%]]
+\petr{Action part of what? Does it relate to the
+  read-state command? Maybe change order: first talk about claim-magic
+and then about read-state?}
+
+{\footnotesize
 
@@
 
-{\small
+{\footnotesize
 \begin{algorithm}[t]

### 2015-11-16 ccr-camera/inband-sync.tex
a new update

@@
 %
-In Algorithms \ref{alg:claim}-\ref{alg:unclaim},
+In Algorithms~\ref{alg:claim}-\ref{alg:unclaim},
 the match part of the constructed commands represents a concatenation
@@
 This allows multiple controllers to claim the same identifier without
-overriding existing flow entries.
+overriding existing flow entries (Algorithm~\ref{alg:claim}).
 %
@@
 Therefore, they can both be used to claim and unclaim identifiers without affecting each other.
-
-In order to check if an identifier is claimed, a controller tries to add
-a new entry setting the flag \texttt{\checko}, and using a different
-match, $m$, where the controller id, $c$, is shifted $16$ bits to the left, i.e., $m=( c_1\concat *^{16}\concat x_1)$,
-thereby inflicting a failure in case an entry with the same claimed value exists.
-If the check succeeds, we delete the entry.
-In any case since the check entry contains the controller id, it does not conflict with other check entries and there is no negative impact of delaying the deletion.
+In order to distinguish between claim entries and other policy
+entries, %[[PK ``some magic value'' sounds magic... 
+we set the action part of the entries storing claimed ids to
+be a unique designated value, $\claimmagic$.
+
+In order to check if an identifier $x$ is claimed
+(Algorithm~\ref{alg:check}), a controller $c$ tries to add
+a ``tester'' entry, where the match value field is 
+$(c\concat*^{16}\concat x)$  and the flag \texttt{\checko} is set.
+Note that the controller identifier is shifted $16$ bits to the left in
+the match field, which allows us to ensures that the entry is uniquely
+identified, while not conflicting with other entries.  
+This attempt inflicts a failure in case another entry with claimed
+value $x$ exists. 
+If the check succeeds, i.e.,  no entry with claimed value $x$ is
+present, the tester entry is deleted.   
+%[[PK unnecessary?
+%In any case since the check entry contains the controller id, it does not conflict with other check entries and there is no negative impact of delaying the deletion.
+%]]
 %Considering $m_1$ and $m_2$ from the example above,
@@
 
-Note that in order to save space, when checking for claims, also the deletion of the added tester flow entry is included,
-which however has no effect on claims and checks of other controllers; hence, the deletion command does not
-have to be sent as part of a bundle. Therefore, the \texttt{\claimcheck} operation is atomic by nature,
-similar to the \texttt{claim} and \texttt{unclaim} operations, and is
-implemented as essentially one command.
-
+%[[PK too wordy?
+%Note that in order to save space, when checking for claims, also the deletion of the added tester flow entry is included,
+%which however has no effect on claims and checks of other controllers; hence, the deletion command does not
+%have to be sent as part of a bundle. Therefore, the \texttt{\claimcheck} operation is atomic by nature,
+%similar to the \texttt{claim} and \texttt{unclaim} operations, and is
+%implemented as essentially one command.
+Note that the tester rule has no effect on claims and checks of other
+controllers. Hence the add and delete commands do not have to be
+executed atomically (e.g., within a bundle): the \texttt{\claimcheck}
+operation ``takes effect'' at the moment it attempts to add the tester entry.   
+%Hence, the deletion command does not
+%have to be sent as part of a bundle. Therefore, the \texttt{\claimcheck} operation is atomic by nature,
+%similar to the \texttt{claim} and \texttt{unclaim} operations, and is
+%implemented as essentially one command.
+%]]
+
+%[[PK isn't it obvious?
 By employing the \textbf{read-state} command, we may easily implement
 an additional operation that returns all currently claimed ids: it is
-sufficient to go over the rules matching every controller identifier.
-In order to distinguish between claim entries and other policy
-entries, returned by  \textbf{read-state},
-%[[PK ``some magic value'' sounds magic... 
-we set the action part to
-be a unique designated value, $\claimmagic$.
-%]]
-\petr{Action part of what? Does it relate to the
-  read-state command? Maybe change order: first talk about claim-magic
-and then about read-state?}
-
-{\footnotesize
+sufficient to go over the rules matching every controller identifier
+carrying action $\claimmagic$.
+%]]
+%[[PK
+%In order to distinguish between claim entries and other policy
+%entries, returned by  \textbf{read-state},
+%%[[PK ``some magic value'' sounds magic... 
+%we set the action part to
+%be a unique designated value, $\claimmagic$.
+%%]]
+%]]
+
+{\scriptsize
 
@@
 The memory \memwrite and \compare
- operations can be implemented using similar techniques.
- \liron{
- As can be seen in Algorithms \ref{alg:write}-\ref{alg:compare},
-  the value part of the added flow entry match contains the memory address, $addr$, and the mask part contains
-  the written value $k$. More specifically, we use a ternary match string of the form $k'\concat \addr$, where $k'$ is result os replacing every zero bit in $k$ with a wildcard ($*$).}
+ operations are implemented using similar techniques.
+To write a value $k$ in a memory address $\addr$ (Algorithm~\ref{alg:write}), a new flow entry is
+added, in which the value part of the match field contains $\addr$ and
+the mask part contains $k$. 
+Therefore, the effective match string od the entry 
+ is $k'\concat \addr$, where $k'$ is result os replacing every zero bit in $k$ with a wildcard ($*$).
   In order to rewrite a memory location, the old entry needs to be deleted and a new one need to be added atomically.
@@
 %share the same match parts.
-Similarly to claim entries, we set the action to a magic value, $\memmagic$, in order to distinguish them from policy entries returned by \textbf{read-state}.
-
-In order to check if a value is written at a given address,
-we try to add it the same way as in the \memwrite operation,
-but we also set the \texttt{\checko} flag. If the value is not there,
-our action differs from the existing entry, but still overlaps with it since only their masks differ \liron{(different mixtures of ones and wildcards in the $k'$ part)},
-thereby inflicting a failure.
-In case the value is there, our flow entry replaces the existing flow entry
+Similarly to claim entries, we set the action to a magic value,
+$\memmagic$, in order to distinguish them from policy entries.
+%[[PK not needed?
+%returned by \textbf{read-state}.
+%]]
+
+To check whether a value $k$ is written at a given
+address $\addr$ (Algorithm ~\ref{alg:compare}),
+a controller tries to add a new entry the same way as in the \memwrite operation,
+but additionally setting the \texttt{\checko} flag. 
+If a different value $k'$ is currently written, 
+%our action differs from the existing entry, but 
+the entry inflicts a failure, as it overlaps but does not coincide
+with the existing entry.
+Indeed, the value parts of the two entries are identical, but the
+masks differ.
+% \liron{(different mixtures of ones and wildcards in the $k'$ part)},
+%thereby inflicting a failure. 
+In case the currently written value is $k$, the existing flow entry is replaced
 with an identical one.
@@
 Initializing required addresses can be safely performed by an a-priori comparison with zero.}
+\petr{A bit unclear}
 
 %HERE: we can read state by reading flow monitor => read configuraiton abstraction => state of claimer ID objects list of all claimed identifiers
+
+{\scriptsize
 
@@
 at most a constant amount of space is lost.}
+\petr{Unclear which ``resources'' are meant here? Should this
+  discussion be postponed to policy updates, where ids indeed grow
+  linearly with the number of updates?}
 
@@
  \liron{I think we should say something like: "... of modern switches can get as high as a few hundreds'. }
+
 \section{Applications}\label{sec:apps}
 
-Given our mechanisms, we can easily implement
-powerful synchronization primitives such at compare-and-set:
-When executed within a
-transaction, a $\cas(\addr,\textit{old},\textit{new})$ operation 
+Our synchronization primitives can be used to 
+implement powerful synchronization primitives such at
+\emph{compare-and-set} (CAS).
+When executed within a transaction, a $\cas(\addr,\textit{old},\textit{new})$ operation 
 on a memory location $\addr$
@@
 Let us now revisit our load-balancing application
-from the beginning of the paper.
+discssed in Section~\ref{sec:motivation}.
 A simple strategy to balance load in our 2-link example
- fails in the presence of
+ (Figure~\ref{fig:CAS-example}) fails in the presence of
 multiple controllers, even if 
@@
 
-{\footnotesize
+{\scriptsize
 \begin{algorithm}[t]
@@
 
-{\small
+{\scriptsize
 \begin{algorithm}[t]

### 2015-11-16 ccr-camera/inband-sync.tex
release tokens

@@
 
-\vspace{1mm}
-\noindent\textbf{Note on efficiency.}
-
-\stefan{We conclude by observing that our implementation is switch space efficient,
-in the sense that it requires only a constant number of flow entries per memory address
-and resources scale linearly in the number of controllers. Also, in case of controller failures,
-at most a constant amount of space is lost.}
-\petr{Unclear which ``resources'' are meant here? Should this
-  discussion be postponed to policy updates, where ids indeed grow
-  linearly with the number of updates?}
-
-\stefan{We also note that while we described our scheme in a simplified
-way where numbers continuously increase, it is straight-forward to use a modulo
-scheme. While in our case 32-bit numbers are used, 
-the flow table width (number of bits which can be matched in parallel)
-of modern switches is 40-bits and more.}
- \liron{I think we should say something like: "... of modern switches can get as high as a few hundreds'. }
-
 \section{Applications}\label{sec:apps}
@@
 Let us now revisit our load-balancing application
-discssed in Section~\ref{sec:motivation}.
+discussed in Section~\ref{sec:motivation}.
 A simple strategy to balance load in our 2-link example
@@
 
-As we show in the following, our mechanisms can be used 
-to render such algorithms concurrent.
-In particular, we can implement a
-generic template to
-implement concurrent policy compositions of a user-specific
+%[[PK ``render concurrent''? we already discuss a concurrent setting...
+%As we show in the following, our mechanisms can be used 
+%to render such algorithms concurrent.
+Using CAS, we can implement a
+generic template 
+for concurrent policy composition via a user-specific
 $\textsc{\ufunc}()$ function~\cite{stn}.
-This way, for example, by choosing our load-balancing
-function $\textsc{lb\_update}$, 
-the pathological scenario described in 
-Figure~\ref{fig:CAS-example} can be resolved.
+This can be, for example, a specific load-balancing
+function $\textsc{lb\_update}$, %\petr{Did we talk about this function before?}
+which, when applied atomically, excludes the pathological scenario described in 
+Figure~\ref{fig:CAS-example}. 
+We use CAS to exactly achieve this atomicity.
 % if we put
@@
 
-Algorithm~\ref{alg:update} is a more efficient alternative 
-to Algorithm~\ref{alg:simple-update}. It uses our id-claimer abstraction
-to bound the number of used policy identifiers.
-Here, the controller first reads the current policy id and claims to prevent another controller from using it 
-for a
-different policy (Lines~\ref{update:read-state1}-\ref{update:read-state2}).
-Then the controller claims the first \emph{unused} id and computes the update commands to be executed (Line~\ref{update:compute}).
+A less trivial Algorithm~\ref{alg:update} adresses the problem of
+policy identifiers growing without bound in  
+Algorithm~\ref{alg:simple-update} by, additionally, using our id-claimer abstraction.
+%to bound the number of used policy identifiers.
+Here, the controller first reads the current policy id and 
+claims it, which prevents another controller from using the id 
+for a different policy (Lines~\ref{update:read-state1}-\ref{update:read-state2}).Then the controller claims the first \emph{unused} id and computes the update commands to be executed (Line~\ref{update:compute}).
 Finally, within an atomic transaction (\textbf{\execatomic}), it verifies whether the chosen policy id has not
 been claimed and that
-the policy has not been concurrently modified, and then updates the policy.
+the policy has not been concurrently modified, and then updates the
+policy.
+Unlike Algorithm~\ref{alg:simple-update}, this algorithm is only
+\emph{lock-free}, as it allows scenarios in which only one live controller
+succeeds in installing its updates. We can easily make it
+wait-free by
+
+\vspace{1mm}
+\noindent\textbf{A note on efficiency.}
+
+We observe that our implementations of synchronization primitives
+(Algorithms~\ref{alg:claim}-\ref{alg:compare}) 
+%is \emph{switch-space efficient}
+%in the sense that 
+require only a constant number of flow entries per memory address.
+The number of used addresses is constant for  
+Algorithm~\ref{alg:simple-update} and linear in the number of
+controllers for Algorithm~\ref{alg:update}. 
+%[[PK not constant: proportional to the number of failures. But this
+%is kind of obvious.
+%Also, in case of controller failures, at most a constant amount of space is lost.
+%]]
+
+\stefan{We also note that while we described our scheme in a simplified
+way where numbers continuously increase, it is straight-forward to use a modulo
+scheme. While in our case 32-bit numbers are used, 
+the flow table width (number of bits which can be matched in parallel)
+of modern switches is 40-bits and more.}
+
+ \liron{I think we should say something like: "... of modern switches can get as high as a few hundreds'. }
+
+\petr{The discussion is very confusing. Do we need it?}
 
@@
 a well-known command line tool for monitoring and administering OpenFlow switches. 
-\stefan{The code is made publicly available at TODO LIRON.}\liron{https://github.com/lironsc/of-sync-lib}
+The code is publicly available at \url{https://github.com/lironsc/of-sync-lib}.
 
@@
 (controllers) issue thousands of concurrent policy updates
-to the switch, where 
-each update consists in replacing a single rule with another one.
-When verifying the transactions, we could confirm that 
-despite the concurrency, 100 \% of the transactions have been successfully completed.
-Of course, due to the concurrency, multiple transactions also needed to abort,
-due to conflicts.
-
-While our framework can be used as synchronization tool itself, we 
+to the switch, 
+each update consisting in replacing a single rule with another one.
+%[[PK unclear
+%When verifying the transactions, 
+%]]
+We could confirm that 
+despite the concurrency, 100 \% of the updates have been
+eventually completed.
+Of course, due to the concurrency, some transactions also needed to abort
+because of conflicts.
+
+While our framework can be used as a synchronization tool itself, we 
 currently also work on implementing it as an integral part of existing controllers 
 such as Ryu. 
+\petr{Did not get the difference between ``itself'' and ``part of
+  existing controllers''.}
 
@@
 introduce additional points of potential failures.
-Moreover, the switch is the element implementing the
-controller policies anyway.
-Interestingly, the mechanisms presented in this paper are simple and can be implemented using the \emph{standard OpenFlow}
+Moreover, the switch is the element implementing the controller policies anyway.
+The mechanisms presented in this paper are simple and can be implemented using the \emph{standard OpenFlow}
 protocol: they do not require any protocol/hardware extensions as postulated in recent literature~\cite{stn,netpaxos}.
@@
 language abstractions~\cite{pyretic}.
-In general, we understand our paper as a first step, and more
-research is required to investigate and compare useful
-inband synchronization abstractions, also in terms of performance.
+In general, we see our paper as a first step, and more
+work is required to investigate and compare useful
+in-band synchronization abstractions, also in terms of performance.
+
 {

### 2015-11-16 ccr-camera/inband-sync.tex
typos

@@
 policy.
-Unlike Algorithm~\ref{alg:simple-update}, this algorithm is only
-\emph{lock-free}, as it allows scenarios in which only one live controller
-succeeds in installing its updates. We can easily make it
-wait-free by
+%[[PK 
+%We observe that our  algorithm is only
+%\emph{lock-free}, as it allows scenarios in which only one live controller
+%succeeds in installing its updates. We can easily make it
+%wait-free by reading the state of the switch and claiming the current
+%policy id within the same {\execatomic} construct.
+%]]
 
@@
 We could confirm that 
-despite the concurrency, 100 \% of the updates have been
+despite the concurrency, 100 \% of the transactions have been
 eventually completed.
@@
 because of conflicts. 
+\petr{Reads strange: 100\% successfullt complete and some transactions
+  abort...}
 

### 2015-11-17 ccr-camera/inband-sync.tex
some comments and fixes

@@
 We will later discuss later how this can be achieved, for example in 
-today's OpenFlow.
+today's OpenFlow. \liron{two times "later"!}
 
@@
 In the non-atomic mode, no atomicity guarantees are provided.
+\liron{all these jumps from atomic to transactional and back is confusing...}
 
@@
 \petr{Is it atomic? Should we say that the command returns a ``snapshot'' of the
-  configuration?}
+  configuration?}\liron{I am not sure if this is atomic, considering that we execute other things in a bundle then this is practically atomic for us. Perhaps we can leave it unspecified as it is now?}
 
@@
 field in the header space, leaving the remaining fields at default
-values.
+values.\liron{this is not accurate, the match part simply considers only the metadata field.
+no need to discuss other fields, in any case matching doesn't change fields and doesn't assume any values to other fields.}
 In defining the value and the mask,
@@
 the mask part contains $k$. 
-Therefore, the effective match string od the entry 
- is $k'\concat \addr$, where $k'$ is result os replacing every zero bit in $k$ with a wildcard ($*$).
+Therefore, the effective match string of the entry 
+ is $k'\concat \addr$, where $k'$ is result of replacing every zero bit in $k$ with a wildcard ($*$).
   In order to rewrite a memory location, the old entry needs to be deleted and a new one need to be added atomically.
@@
 \petr{A bit unclear}
+
+\liron{SECOND TRY: Note that comparing an address for the first time, without writing it first, always succeeds and also writes to it the compared value.
+In case where such comparison may occur to some address $addr$, it is advised to first perform a safe initialization by calling $\textit{\compare}(\addr,0)$ which fails in case that $addr$ has already been used (and the initialization is not desired).}
 
@@
 claims it, which prevents another controller from using the id 
-for a different policy (Lines~\ref{update:read-state1}-\ref{update:read-state2}).Then the controller claims the first \emph{unused} id and computes the update commands to be executed (Line~\ref{update:compute}).
+for a different policy (Lines~\ref{update:read-state1}-\ref{update:read-state2}).
+Then the controller chooses an \emph{unused} id and computes the update commands to be executed (Line~\ref{update:compute}).
+\liron{I have replaced "claims the first" with "chooses an" following Petr email}
 Finally, within an atomic transaction (\textbf{\execatomic}), it verifies whether the chosen policy id has not
@@
 %in the sense that 
-require only a constant number of flow entries per memory address.
+require only a constant number of flow entries per memory address or claim.
 The number of used addresses is constant for  
@@
 a well-known command line tool for monitoring and administering OpenFlow switches. 
-The code is publicly available at \url{https://github.com/lironsc/of-sync-lib}.
+The code is publicly available at \url{https://github.com/lironsc/of-sync-lib}. \liron{perhaps we can put it as reference? (putting the link in bib)}
 

### 2015-11-18 ccr-camera/inband-sync.tex
cr format
done

--- 2015-11-16 ccr-camera/inband-sync.tex
@@
 %\documentclass[11pt,pdftex,letter]{article}
-\documentclass[conference]{sigcomm-alternate}
+
+%\documentclass[conference]{sigcomm-alternate}
+
+%\documentclass[dvipdfm]{sig-alternate-2006}
+
+\documentclass{sig-alternate-2006}
 
@@
 \centering
-\includegraphics[width=.56\columnwidth]{loadbal-1.pdf}~\hspace{-.7cm}~\includegraphics[width=.56\columnwidth]{loadbal-2.pdf}\\
+\includegraphics[width=.56\columnwidth]{loadbal-1.pdf}~\hspace{-.7cm}
+~\includegraphics[width=.56\columnwidth]{loadbal-2.pdf}\\
 \caption{\small \emph{Left:} Without synchronization, the two controllers naturally choose to
@@
 traffic. 
-We will later discuss later how this can be achieved, for example in 
+We will discuss later how this can be achieved, for example, in 
 today's OpenFlow. 
-
 %One way to achieve this, for example, is to set a low priority to
@@
 sequences of \emph{operations} on the
-switch configuration in an \emph{atomic} (transactional) or \emph{non-atomic} mode.
+switch configuration in an \emph{atomic} or \emph{non-atomic} mode.
 In the atomic mode, either \emph{all} operations in the sequence, henceforth simply
@@
 In the non-atomic mode, no atomicity guarantees are provided.
-
-We introduce two constructs
+%\liron{all these jumps from atomic to transactional and back is confusing...}
+%
+
+%For this purpose, we introduce two constructs:
 %$\textbf{bundle}\{\;\ldots\;\}\rightarrow res$ to refer to represent a
-$\exec\{op_1,\;\ldots\;op_k\}$ and $\execatomic\{op_1,\;\ldots\;op_k\}$
-which perform a sequence of commands $op_1,\ldots,op_k$ in
-a regular and atomic way, respectively.
-In particular, the construct $\execatomic\{op_1,\;\ldots\;op_k\}$  
-returns $\ack$ if and only if all commands in the sequence are performed successfully; otherwise
-$\error(\xid,\ecode)$ is returned, providing
+%$\exec\{\ldots\}$ and
+%$\execatomic\{\ldots\}$.
+In particular, construct $\exec\{op_1,\;\ldots\;op_k\}$ is used to perform a sequence of commands $op_1,\ldots,op_k$ in
+a conventional way.
+Construct $\execatomic\{op_1,\;\ldots\;op_k\}$ ensures that either all
+commands are successfully performed, in which case it returns
+$\ack$, or, if some of the commands cannot be performed, aborts the
+whole sequence and returns
+$\error(\xid,\ecode)$, providing
 the identifier of the first command in the sequence that failed
@@
 
-Therefore, we provide the following (transactional) operations.
+Therefore, we provide the following transactional operations.
 A $\textit{write}(\addr,k)$ operation sets the content at memory location
@@
 this message, receiving a response, and returning the received configuration.
-\petr{Is it atomic? Should we say that the command returns a ``snapshot'' of the
-  configuration?}
+%\petr{Is it atomic? Should we say that the command returns a ``snapshot'' of the
+%  configuration?}
+%\liron{I am not sure if this is atomic, considering that we execute other things in a bundle then this is practically atomic for us. Perhaps we can leave it unspecified as it is now?}
+%[[PK moved this discussion to the algorithms 6 and 7]]
 
@@
 The \textit{match} part in these commands specifies only the $64$bit meta-data
-field in the header space, leaving the remaining fields at default
-values.
+field in the header space.
+%, leaving the remaining fields at default
+%values.
+%\liron{this is not accurate, the match part simply considers only the metadata field.
+no need to discuss other fields, in any case matching doesn't change fields and doesn't assume any values to other fields.}
 In defining the value and the mask,
@@
 with an identical one.
-
-\liron{Note that comparing an uninitialized address always succeeds and also writes to it the compared value.
+%[[PK on initialization
+Before using an address $\addr$, it is advised that a controller safely
+initializes it with a special initial value $0$ by calling
+$\textit{\compare}(\addr,0)$: 
+the first such command succeeds, and any 
+subsequent one will fail. 
+%]]
+
+%\liron{Note that comparing an uninitialized address always succeeds and also writes to it the compared value.
 Initializing required addresses can be safely performed by an a-priori comparison with zero.}
-\petr{A bit unclear}
+%\petr{A bit unclear}
+
+%\liron{SECOND TRY: Note that comparing an address for the first time, without writing it first, always succeeds and also writes to it the compared value.
+In case where such comparison may occur to some address $addr$, it is advised to first perform a safe initialization by calling $\textit{\compare}(\addr,0)$ which fails in case that $addr$ has already been used (and the initialization is not desired).}
 
@@
 with $\pid+1$.
-The latter ensures that if, concurrently, a new policy (with a different identifier) has been installed, the update
+The CAS operation ensures that if, concurrently, a new policy (with a different identifier) has been installed, the update
 fails and takes no effect.
+Note that, since all policy modifications are executed within a
+bundle providing controller-atomicity, the \textbf{read-state} command
+in Line~\ref{line:readstate} appears atomic.  
 Algorithm~\ref{alg:simple-update} can therefore also be used
@@
  		\Repeat
- 			\State $\pid,\textit{policy}\gets$ \textbf{read-state} %switch configuration}
+ 			\State $\pid,\textit{policy}\gets$ \textbf{read-state} \label{line:readstate}%switch configuration}
  			\State $\textit{update\_cmds}\gets \textsc{\ufunc}(policy)$
@@
 
-A less trivial Algorithm~\ref{alg:update} adresses the problem of
-policy identifiers growing without bound in  
-Algorithm~\ref{alg:simple-update} by, additionally, using our id-claimer abstraction.
+Algorithm~\ref{alg:simple-update} is correct assuming that policy
+identifiers may grow without bound. 
+In practice, we can use a module scheme on the 32-bit field of a flow entry
+used for storing the policy identifier, which, in principle,
+may lead to inconsistencies when the relative speeds of controllers is
+very high.   
+Algorithm~\ref{alg:update} adresses this issue 
+by using our id-claimer abstraction.
 %to bound the number of used policy identifiers.
@@
 claims it, which prevents another controller from using the id 
-for a different policy (Lines~\ref{update:read-state1}-\ref{update:read-state2}).Then the controller claims the first \emph{unused} id and computes the update commands to be executed (Line~\ref{update:compute}).
+for a different policy
+(Lines~\ref{update:read-state1}-\ref{update:read-state2}). 
+Then the controller selects any \emph{unused} id and computes the update commands to be executed (Line~\ref{update:compute}).
+%\liron{I have replaced "claims the first" with "chooses an" following Petr email}
 Finally, within an atomic transaction (\textbf{\execatomic}), it verifies whether the chosen policy id has not
 been claimed and that
-the policy has not been concurrently modified, and then updates the
+the policy has not been concurrently modified, and then tries to update the
 policy.
+%
+
 %[[PK 
@@
 \noindent\textbf{A note on efficiency.}
-
+%
 We observe that our implementations of synchronization primitives
@@
 
-\stefan{We also note that while we described our scheme in a simplified
-way where numbers continuously increase, it is straight-forward to use a modulo
-scheme. While in our case 32-bit numbers are used, 
-the flow table width (number of bits which can be matched in parallel)
-of modern switches is 40-bits and more.}
-
- \liron{I think we should say something like: "... of modern switches can get as high as a few hundreds'. }
-
-\petr{The discussion is very confusing. Do we need it?}
+%\stefan{We also note that while we described our scheme in a simplified
+%way where numbers continuously increase, it is straight-forward to use a modulo
+%scheme. While in our case 32-bit numbers are used, 
+%the flow table width (number of bits which can be matched in parallel)
+%of modern switches is 40-bits and more.}
+
+ %\liron{I think we should say something like: "... of modern switches can get as high as a few hundreds'. }
+
+%\petr{The discussion is very confusing. Do we need it?}
 
@@
 a well-known command line tool for monitoring and administering OpenFlow switches. 
-The code is publicly available at \url{https://github.com/lironsc/of-sync-lib}.
+The code is available online at
+\url{https://github.com/lironsc/of-sync-lib}. 
+%\liron{perhaps we can put it as reference? (putting the link in bib)}
 
@@
 For instance, we experimented with a setting where three processes
-(controllers) issue thousands of concurrent policy updates
-to the switch, 
-each update consisting in replacing a single rule with another one.
+(controllers), concurrently issuing $1000$ concurrent policy updates
+each, where every update replaces a single rule with another one.
 %[[PK unclear
@@
 We could confirm that 
-despite the concurrency, 100 \% of the transactions have been
-eventually completed.
-Of course, due to the concurrency, some transactions also needed to abort
-because of conflicts. 
-\petr{Reads strange: 100\% successfullt complete and some transactions
-  abort...}
-
-While our framework can be used as a synchronization tool itself, we 
-currently also work on implementing it as an integral part of existing controllers 
-such as Ryu. 
-\petr{Did not get the difference between ``itself'' and ``part of
-  existing controllers''.}
+despite the concurrency, 100\% of the updates 
+eventually completed in committing their transactions, possibly after
+several aborts caused by conflicts. 
+%Of course, due to the concurrency, some transactions also needed to abort
+%because of conflicts. 
+%\petr{Reads strange: 100\% successfullt complete and some transactions
+ % abort...}
 
@@
 language abstractions~\cite{pyretic}.
-In general, we see our paper as a first step, and more
+We see our paper as a first step, and more
 work is required to investigate and compare useful
 in-band synchronization abstractions, also in terms of performance.
+While our framework can be used as a synchronization library, we 
+also work on integrating it into an existing SDN controller, e.g., Ryu.
 

### 2015-11-18 ccr-camera/inband-sync.tex
minor things from my past notes

@@
 events close to their origin~\cite{devoflow,kandoo}.
-Third, larger SDNs are likely to be operated by multiple administrators~\cite{stn} or may even offer
+Third, larger SDNs are likely to be operated by multiple administrators~\cite{infocom15} or may even offer
 participatory interfaces where different users can install and trigger policy changes
@@
 %[[PK it is incorrect to only cite the book for the TM notion]]
-\emph{transactional  memory} abstraction (cf. a survey in~\cite{tm-book}). 
+\emph{transactional  memory} abstraction (cf.~the survey in~\cite{tm-book})
 %]]
@@
 the missing link for the read-modify-write object
-postulated in CPC and STN~\cite{stn,infocom15}. %\petr{We do not seem to discuss the latter?}
+postulated in CPC and STN~\cite{infocom15}. %\petr{We do not seem to discuss the latter?}
 
@@
 We aim to enforce a consistent use of  \emph{policy identifiers},
-an important notion in consistent policy updates~\cite{network-update,stn}.
+an important notion in consistent policy updates~\cite{network-update,infocom15}.
 Intuitively, a controller should
@@
 
-To make sure that policy identifiers do not grow without bound~\cite{stn}, we
+To make sure that policy identifiers do not grow without bound~\cite{infocom15}, we
 also provide an \emph{id-claimer} object.
@@
 In the simplest setting, a switch maintains one flow table (table 0).
-A data-fplane packet arriving at a switch is first checked against
+A data-plane packet arriving at a switch is first checked against
 the rule with the highest priority and in table $0$.
@@
 for concurrent policy composition via a user-specific
-$\textsc{\ufunc}()$ function~\cite{stn}.
+$\textsc{\ufunc}()$ function~\cite{infocom15}.
 This can be, for example, a specific load-balancing
@@
 The mechanisms presented in this paper are simple and can be implemented using the \emph{standard OpenFlow}
-protocol: they do not require any protocol/hardware extensions as postulated in recent literature~\cite{stn,netpaxos}.
+protocol: they do not require any protocol/hardware extensions as postulated in recent literature~\cite{infocom15,netpaxos}.
 Our work may hence also contribute to the ongoing discussion of what can be implemented
@@
 {
-%\begin{footnotesize}
+\begin{footnotesize}
 \bibliographystyle{abbrv}
 \bibliography{references}  % main.bib is the name of the Bibliography in this case
-%\end{footnotesize}
+\end{footnotesize}
 }

### 2015-11-20 ccr-camera/inband-sync.tex
camera ready proposal

@@
 
-\category{C.2.2}{Computer-Communication Networks}{Network Protocols}
+%\category{C.2.2}{Computer-Communication Networks}{Network Protocols}
 \category{C.2.3}{Network Operations}{Network Management}
@@
 
-\terms{Networks}
-
-\keywords{Software-Defined Networking, Distributed Control Planes, In-band Mechanisms}
+%\terms{Networks}
+
+%\keywords{Software-Defined Networking, Distributed Control Planes, In-band Mechanisms}
 
@@
 all-or-nothing semantics.
+
+\paragraph{Contributions}
 
@@
 %control-plane 
-service~\cite{Her91} 
+service~\cite{Her91}, 
 %in a consistent and fault-tolerant way: 
@@
 the missing link for the read-modify-write object
-postulated in CPC and STN~\cite{infocom15}. %\petr{We do not seem to discuss the latter?}
+postulated in STN~\cite{infocom15}. %\petr{We do not seem to discuss the latter?}
 
@@
 
-\section{Motivation and Example}\label{sec:motivation}
-
+\paragraph{Motivation and Example}
 A transactional distributed system offers all-or-nothing semantics, 
@@
 %
-We distinguish two parts of the configuration space of a switch:
+We distinguish between two parts of the configuration space of a switch:
 the ``normal'' configuration which determines the network policy and
@@
 
-As a case study, in this section, we show how to implement
+Before discussing applications in more detail, 
+as a case study and in order to introduce some notation, 
+we in this section show how to implement
 our transactional approach using today's standard SDN protocol,
- OpenFlow.
-
-\subsection{Background}
-
+ OpenFlow  (version 1.4~\cite{openflow}).\footnote{We note
+however that hardware OpenFlow switches do not always 
+follow the spec as they should~\cite{know}, and will report on a
+prototype implementation later in this paper.}
+
+\paragraph{Background}
 The configuration of an OpenFlow switch is a set of
@@
 
-\subsection{Implementation}
-
+\paragraph{Implementation} 
 At first sight, it may seem that bundling \emph{FlowMod} commands with  the $\checko$
@@
 Note that the controller identifier is shifted $16$ bits to the left in
-the match field, which allows us to ensures that the entry is uniquely
+the match field, which allows us to ensure that the entry is uniquely
 identified, while not conflicting with other entries.  
@@
 the switch configuration, where the configuration contains both the
-policy and a meta-configuration memory address $\paddr$ that holds the \emph{policy identifier (\pid)} .
+policy and a meta-configuration memory address $\paddr$ that holds the \emph{policy identifier (\pid)}.
 In the algorithm, the controller first reads the currently installed
@@
 identifiers may grow without bound. 
-In practice, we can use a module scheme on the 32-bit field of a flow entry
-used for storing the policy identifier, which, in principle,
-may lead to inconsistencies when the relative speeds of controllers is
-very high.   
+In practice, we can use a modulo scheme on the 32-bit field of a flow entry
+used for storing the policy identifiers, which, in principle,
+may lead to inconsistencies if controllers proceed in different speeds.
 Algorithm~\ref{alg:update} adresses this issue 
@@
 %
-We observe that our implementations of synchronization primitives
+Our implementations of synchronization primitives
 (Algorithms~\ref{alg:claim}-\ref{alg:compare}) 
@@
  			\EndIf
- 			\State $\textit{my\_id}\gets$ choose a number from $C\setminus claims$
+ 			\State $\textit{my\_id}\gets$ choose a number from $\{C\setminus claims\}$
  			\State $\textit{update\_cmds}\gets \textsc{\ufunc}(policy)$ \label{update:compute}
@@
 a well-known command line tool for monitoring and administering OpenFlow switches. 
-The code is available online at
-\url{https://github.com/lironsc/of-sync-lib}. 
+The code is made available online\footnote{See \url{https://github.com/lironsc/of-sync-lib}.}. 
 %\liron{perhaps we can put it as reference? (putting the link in bib)}
@@
 For instance, we experimented with a setting where three processes
-(controllers), concurrently issuing $1000$ concurrent policy updates
+(controllers), concurrently issue $1000$ concurrent policy updates
 each, where every update replaces a single rule with another one.
@@
 also work on integrating it into an existing SDN controller, e.g., Ryu. 
+%Moreover, we will consider adding special
+%locking/synchronization primitives to OpenFlow.
 

### 2015-11-20 ccr-camera/inband-sync.tex
final version

@@
 %\title{On the Synchronization Power of the OpenFlow Data Plane}
-\title{In-Band Synchronization for\\Distributed SDN Control Planes\thanks{Supported by the
+\title{In-Band Synchronization for\\Distributed SDN Control Planes\thanks{Supported by the European Research Council (ERC) Starting Grant no.~259085, the
 European Institute of Innovation \& Technology (EIT) project 13153 (Software Defined Networking) 
@@
 to improve load-balancing, and to reduce overheads,
-modules of the control plane %operating on the data plane 
+modules of the control plane 
 should be physically distributed.
-However, %a distributed control plane also introduces
-%new challenges,
+However, 
 in order to guarantee \emph{consistency} of network operation, actions
@@
 transactions, 
-%In particular, we make the case for 
 implemented \emph{in-band}, on the 
@@
 paradigm, 
-%[[PK key-valued stores are not really discussed?
-%to realize simple key-value stores, 
 we can implement our synchronization framework in
@@
 $\textit{\compare}(\addr,0)$: 
-the first such command succeeds, and any 
-subsequent one will fail. 
+the first such command succeeds, 
+and subsequent ones will do no harm.
 %]]
@@
 Let us now revisit our load-balancing application
-discussed in Section~\ref{sec:motivation}.
+discussed in Figure~\ref{fig:CAS-example}.
 A simple strategy to balance load in our 2-link example
@@
 (Lines~\ref{update:read-state1}-\ref{update:read-state2}). 
-Then the controller selects any \emph{unused} id and computes the update commands to be executed (Line~\ref{update:compute}).
+Then the controller chooses any \emph{unused} id and computes the update commands to be executed (Line~\ref{update:compute}).
 %\liron{I have replaced "claims the first" with "chooses an" following Petr email}

### 2015-12-06 ccr-camera/inband-sync.tex
fixed camera submission issues. submitted again.

@@
 
+\pdfpageheight 11in
+\pdfpagewidth 8.5in
 %\documentclass[11pt]{llncs}
@@
 
+\numberofauthors{3}
 \author{
-Liron Schiff$^1$  \quad % \thanks{Supported by the European Research Council (ERC) Starting Grant no. 259085 and by the Israel Science Foundation Grant no.~1386/11.} $^1$,
-Stefan Schmid$^2$ \quad Petr Kuznetsov$^3$ \\
-\small $^1$ Tel Aviv University, Israel \quad $^2$ Aalborg University,
-Denmark \quad $^3$ T\'el\'ecom ParisTech, France
+\alignauthor Liron Schiff \\
+\affaddr{Tel Aviv University, Israel} \\
+%\affaddr{ACM} \\
+%\affaddr{New York, NY} \\
+%\email{griscti@hq.acm.org}
+\alignauthor Stefan Schmid \\
+\affaddr{Aalborg University, Denmark} \\
+%\affaddr{ACM} \\
+%\affaddr{New York, NY} \\
+%\email{murray@hq.acm.org}
+\alignauthor Petr Kuznetsov \\
+\affaddr{T\'el\'ecom ParisTech, France} \\
+%\affaddr{ACM} \\
+%\affaddr{New York, NY} \\
+%\email{rous@hq.acm.org}
 }
+%\author{
+%Liron Schiff$^1$  \quad % \thanks{Supported by the European Research Council (ERC) Starting Grant no. 259085 and by the Israel Science Foundation Grant no.~1386/11.} $^1$,
+%Stefan Schmid$^2$ \quad Petr Kuznetsov$^3$ \\
+%\small $^1$ Tel Aviv University, Israel \quad $^2$ Aalborg University,
+%Denmark \quad $^3$ T\'el\'ecom ParisTech, France
+%}
 
